### 2020-10-06 README.Rmd
Estructura general del proyecto

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Mesoscale Data Assimilation - A PhD Thesis"
```

# mesoda

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/paocorrales/mesoda/master?urlpath=rstudio)

This repository contains the code for my future thesis.

### How to cite

Please cite this compendium as:

> Corrales, (`r format(Sys.Date(), "%Y")`). _Compendium of R code for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`.

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/chapters): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. 
  - [:file\_folder: data](/analysis/data): Links to data used in the analysis. The data is not included here.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your broswer or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text and code is to click on
[binder](https://mybinder.org/v2/gh/paocorrales/mesoda/master?urlpath=rstudio)
to open an instance of RStudio in your browser, which will have the
compendium files ready to work with. Binder uses rocker-project.org
Docker images to ensure a consistent and reproducible computational
environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2020-10-06 R/wrf_project.R
Funciones auxiliares

@@ -1,0 +1,25 @@
+#' Proyección de coordenadas
+#'
+#' Proyecta y reproyecta coordenadas utilizando el paquete proj4
+#'
+#' @param lon,lat longitud y latitud o x e y para calcular la inversa
+#' @param inverse si es FALSE la función devuelve las coordenadas cartográficas (distancia al centro del dominio).
+#' Si es TRUE devuelve lon, lat.
+#' @param round igual a TRUE resuelve problemas de redondeo
+#' @param map_proj recibe un string con la proyección correspondiente en el formato aceptado por proj4
+#' @export
+
+wrf_project <- function(lon, lat, inverse = FALSE, round = TRUE, map_proj = norargentina_lambert) {
+
+  xy <- proj4::project(list(lon, lat), map_proj, inverse = inverse)
+  if (round == TRUE) {
+    list(x = round(xy$x, -3),
+         y = round(xy$y, -4))
+  } else {
+    list(x = xy$x,
+         y = xy$y)
+  }
+}
+
+#' @export
+norargentina_lambert <- "+proj=lcc +lat_1=-30.9659996032715 +lat_2=-30.9659996032715 +lat_0=-30.9660034179688 +lon_0=-63.5670013427734 +a=6370000 +b=6370000"

### 2020-10-06 temp_R/goes_functions.R

@@ -1,0 +1,74 @@
+
+# projection --------------------------------------------------------------
+
+goes_projection <- function(x, y, ncfile) {
+
+    proj_info <- ncdf4::ncatt_get(ncfile, "goes_imager_projection")
+  h <- proj_info$perspective_point_height 
+  
+  map_proj <- paste0("+proj=geos",
+                     " +h=",  proj_info$perspective_point_height,
+                     " +lon_0=", proj_info$longitude_of_projection_origin,
+                     " +sweep=", proj_info$sweep_angle_axis,
+                     " +ellps=GRS80")
+  
+  x_atr <- ncdf4::ncatt_get(ncfile, "x")
+  x <- (x*x_atr$scale_factor + x_atr$add_offset)*h
+  y_atr <- ncdf4::ncatt_get(ncfile, "y")
+  y <- (y*y_atr$scale_factor + y_atr$add_offset)*h
+
+  proj4::project(list(x, y), map_proj, inverse = TRUE)
+}
+
+# ggplot2::ggplot() +
+#   
+#   scattermore::geom_scattermost(data_sub[!is.na(z_norm), .(lon, lat)], interpolate = TRUE, pixels = c(512, 512),
+#                                 viridisLite::viridis(100)[1 + 99*data_sub[!is.na(z_norm), z_norm]]) +
+#   geom_path(data = map, aes(long, lat, group = group)) +
+#   ggplot2::coord_equal(ylim = c(-45, -20), xlim = c(-77, -40))
+
+
+# scale_color_topes -------------------------------------------------------
+
+scale_color_topes <- function(colours = hex, 
+                              limits = c(-90, 50), 
+                              breaks = seq(-90, 50, 20),
+                              guide = guide_colorbar(barheight = 15),
+                              ...) {
+  # https://github.com/garciafido/cima-goes/tree/master/src/cima/goes/img
+  topes <- data.table::fread("paleta_topes")
+  colours <- rgb(topes$V2, topes$V3, topes$V4, maxColorValue = 255)
+  
+  scale_color_gradientn(colours = colours, 
+                        limits = limits, 
+                        breaks = breaks, 
+                        guide = guide, ...) 
+}
[…]

### 2020-10-06 temp_R/help_functions.R

@@ -1,0 +1,528 @@
+# Help functions
+
+
+# read_diag ---------------------------------------------------------------
+# Read diagfiles and tidy uv observations
+
+read_diag <- function(path, variable) {
+  library(foreach)
+  library(doParallel)
+  
+  files <- Sys.glob(paste0(path, ".mem*"))
+  
+  myCluster <- makeCluster(20)
+  registerDoParallel(myCluster)
+  
+  obs <- foreach(f = 1:length(files),
+                 .packages = c("data.table", "metR", "lubridate", "tidyverse"),
+                 .export = c("files", "variable"),
+                 .combine = "rbind") %dopar% {
+                   
+                   # sink("log.txt", append=TRUE)
+                   
+                   diag <- fread(files[f]) %>% 
+                     .[V10 == 1] %>% 
+                     .[, exp := basename(dirname(files[f]))] %>% 
+                     .[, mem := str_extract(files[f], "\\d{3}$")] %>% 
+                     .[, date := ymd_hms(str_extract(files[f], "\\d{14}"))] %>% 
+                     .[, c("V2", "V4") := NULL]
+                   
+                   # cat("Archivo ", basename(files[f]))
+                   
+                   colnames(diag) <- c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag.prep", "obs", "obs.guess", "obs2", "obs.guess2", "rerr", "exp", "mem", "date")
+                   
+                   if ("uv" %in% variable & length(variable) == 1) {
+                     
+                     diag <- diag[var == "uv"] %>% 
+                       melt(measure.vars = c("obs", "obs2", "obs.guess", "obs.guess2")) %>% 
+                       .[, var := if_else(str_detect(variable, "2"), "v", "u")] %>% 
+                       .[, variable := str_remove(variable, "2")] 
+                     
+                     vars <- rlang::syms(setdiff(names(diag), "value")) 
+                     diag <- diag %>% 
+                       distinct(!!!vars, .keep_all = TRUE) %>%  
+                       pivot_wider(names_from = variable, values_from = value) %>% 
+                       setDT %>% 
[…]

### 2020-10-06 temp_R/notas.R

@@ -1,0 +1,90 @@
+obs[, ratio := obs.guess/max(1.3, min(5.6, rerr, na.rm = TRUE), na.rm = TRUE)] %>% 
+  .[usage.flag == 1 & var == "t"] %>% 
+  ggplot(aes(ratio)) +
+  geom_density()
+
+
+diag <- fread("E1/test/test.csv", na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00"))
+
+files <- Sys.glob("*/wrfout_d01_2018-11-20_18:00:00")
+
+ana <- lapply(files, function(f) {
+  ens <- substr(f, 1, 2)
+  message(paste0("Leyendo el archivo ",f))
+  ReadNetCDF(f, vars = c("T", "XLAT", "XLONG"), subset = list(bottom_top = 16)) %>% 
+    .[, c("Time", "south_north", "west_east", "bottom_top") := NULL] %>% 
+    .[, ens := ens] %>% 
+    .[]
+}) %>% 
+  rbindlist()
+
+colnames(ana) <- c("var", "lat", "lon", "ens")
+
+spread <- ana %>% 
+  .[, mean := mean(var), by = .(lon, lat)] %>% 
+  .[, .(sprd = sd(var - mean))]
+
+print(paste0("El spread de T500 es ", spread))
+
+copy(ana) %>% 
+  .[, mean := mean(var), by = .(lon, lat)] %>% 
+  .[, .(sprd = sd(var - mean)), by = .(lon, lat)] %>% 
+  ggplot(aes(lon, lat)) +
+  geom_point(aes(color = sprd)) +
+  scale_color_viridis_c() +
+  labs(title = "Ensemble spread", subtitle = "Temperature, level 16 (~500 hPa)")
+ggsave("T500.png")
+
+ana <- lapply(files, function(f) {
+  ens <- substr(f, 1, 2)
+  message(paste0("Leyendo el archivo ",f))
+  ReadNetCDF(f, vars = c("U", "XLAT_U", "XLONG_U"), subset = list(bottom_top = 16)) %>%
+    .[, c("Time", "south_north", "west_east_stag", "bottom_top") := NULL] %>%
+    .[, ens := ens] %>%
+    .[]
+}) %>%
+  rbindlist()
+
+colnames(ana) <- c("var", "lat", "lon", "ens")
+
+spread <- ana %>% 
+  .[, mean := mean(var), by = .(lon, lat)] %>% 
[…]

### 2020-10-06 temp_R/parse_errtable.R

@@ -1,0 +1,49 @@
+library(tidyverse)
+library(data.table)
+
+string <- read_lines("prepobs_errtable.global")
+
+split_chunks <- function(string, pattern = "OBSERVATION TYPE", negate = FALSE) {
+  start <- stringr::str_which(string, pattern, negate = negate)
+  end <- c(start-1, length(string))[-1]
+  
+  obs_type <- stringr::str_extract(string[start], "\\d+")
+  
+  chunks <- purrr::map(seq_along(start), ~ string[(start[.x]+1):end[.x]])
+  
+  names(chunks) <- obs_type  
+  chunks
+}
+
+string <- split_chunks(string, "OBSERVATION TYPE")
+
+chunk <- string[[1]]
+
+parse_chunk <- function(chunk) {
+  str_split(chunk, " ") %>% 
+    map( ~ str_subset(.x, "^$", negate = TRUE) %>% 
+           setNames(c("P", "T", "Q", "U", "V", "PW")) %>%
+           as.numeric()) %>% 
+    do.call(rbind, .) %>% 
+    as.data.frame()
+}
+
+errtable <- map(string, parse_chunk) %>% 
+  rbindlist(idcol = "type") 
+
+na_fill <- function(x, na.char) {
+  nas <- grepl(na.char, x)
+  x[nas] <- NA
+  return(x)
+}
+
+errtable %>% 
+  mutate(type = as.numeric(type)) %>% 
+  rename(nivel = V1, 
+         t = V2,
+         rh = V3,
+         uv = V4,
+         ps = V5, 
+         pw = V6) %>% 
+  mutate_all(~ifelse(. == 1e+09, NA, .)) %>% 
+  write_csv("errtable.csv")

### 2020-10-06 temp_R/postprocesamiento.R

@@ -1,0 +1,115 @@
+
+# Post procesamiento WRF --------------------------------------------------
+
+## Humedad relativa
+# https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L730
+
+rh <- function(QVAPOR, P, T) {
+  P <- P*0.01      # Debe estar en hPa
+  T <- T - 273.15  # Debe estar en Celsius
+  es <- 6.112 * exp(17.67*(T)/(T + 273.15 - 29.65))
+
+  qvs <- es/(P - (1 - 0.622)*es)
+
+  rh <- 100*pmax(pmin(QVAPOR/qvs, 1), 0)
+
+  return(rh)
+}
+
+## Dew Point
+# https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L970
+
+td <- function(QVAPOR, P) {
+  P <- P*0.01
+
+  QVAPOR <- ifelse(QVAPOR < 0, 0, QVAPOR)
+
+  tdc <- QVAPOR*P / (0.622 + QVAPOR)
+
+  tdc <- ifelse(tdc < 0.001, 0.001, tdc)
+
+  td <- (243.5*log(tdc) - 440.8) / (19.48 - log(tdc))
+
+  return(td)
+}
+
+## T
+# https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L57
+
+tk <- function(T, P, T_BASE = 300){
+  tk <- (T + T_BASE)*(P/100000)^(2/7)
+}
+
+## UV
+# Rota los vientos e interpola a la retícula de masa
+# Solo para Labert
+
+uvmet <- function(ncfile) {
+  library(ncdf4)
+  nc <- nc_open(ncfile)
+
+  # https://github.com/NCAR/wrf-python/blob/b40d1d6e2d4aea3dd2dda03aae18e268b1e9291e/src/wrf/g_uvmet.py#L179
+  true_lat1 <- ncatt_get(nc, 0, "TRUELAT1")[[2]]
+  true_lat2 <- ncatt_get(nc, 0, "TRUELAT2")[[2]]
+  cen_lon <- ncatt_get(nc, 0, "STAND_LON")[[2]]
[…]

### 2020-10-06 analysis/chapters/coeficientes.Rmd
Inicio capitulo de corrección de coeficientes.

---
title: "Coeficientes para la corrección de bias"
output: bookdown::html_document2
bibliography: mesoda_ref.bib
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.path = "../figures/",
	message = FALSE,
	warning = FALSE,
	collapse = TRUE,
	comment = "#>",
	dpi = 300
)
library(mesoda)
library(metR)
library(tidyverse)
library(data.table)
library(here)
library(knitr)
library(kableExtra)

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

sensores <- c("amsua_n15",
              "amsua_n18",
              "amsua_n19",
              "amsua_metop-b",
              "amsua_aqua",
              "airs_aqua",
              "iasi_metop-a",
              "iasi_metop-b", 
              "hirs4_n19",
              "hirs4_metop-a",
              "hirs4_metop-b",
              "mhs_metop-a",
              "mhs_metop-b",
              "mhs_n18",
              "mhs_n19",
              "mhs_metop-a",
              "atms_npp",
              "atms_n20",
              "cris-fsr_npp",
              "cris-fsr_n20"
)
```

La correcta asimilación de observaciones de satélite, en particular radianzas, depende en gran medida de la estimación y corrección del bias asociado a cada observación. Si el bias no es corregido podría afectar el análisis y los pronósticos generados generando impactos negativos.

Uno de los desafíos en la corrección de bias para asimilación regional es la cantidad de observaciones disponibles en cada ciclo de asimilación lo que depende a su vez de la ubicación y orbitas de los distintos satélites. En la región de Sudamérica los distintos satélites polares pasan una o dos veces por día y el dominio escaneado es variable. A esta restricción se suma la disminución de observaciones disponibles debido al límite superior de los modelos regionales.

El sistema GSI-LETKF incorpora un metodo de corrección de bias que estima y actualiza coeficientes $\beta_i$ que se aplican a una regresón lineal previo a la estimación del estado final o análisis [@miyoshi2010]. Esta estimación se calcula aplicando la ecuación:

$$ \delta\beta = (B_{\beta}^{-1} + PR^{-1})^{-1}PR^{-1}[y-H(x)-P^{T}\beta]$$

Esta formulación es equivalente a la utilizada por la versión variacional de GSI y análoga a la ecuación del cálculo del análisis con la inclusión de la corrección de bias. En la ecuación $\delta\beta$ se asume diagonal y sus elementos se inicializan con el valor 0.1 y luego de su estimación que se calcula de manera iterativa, los coeficientes son actualizados:

$$\beta = \beta_b+\delta \beta$$

La corrección del bias se calcula utilizanod una regresión lineal con $N$ predictores independientes $P_i(x)$ y sus correspondientes coeficientes $\beta_i$.

$$BC = \sum_{i=1}^{N} \beta_i P_i(x) $$

La corrección del bias luego es aplicada directamente sobre la innovación $y - H(x)$.

Los predictores utilizados para la corrección asociada al estado de la atmósferea en este sistema corresponden a:

* Global offset
* Zenit angle
* Cloud liquid water (CLW)
* Temperature lapse rate (TLR)
* Temperature lapse rate al cuadrado

Mientras que la corrección asociada al ángulo de observación utiliza un polinomio de grado N.

El cálculo de CLW y TLR depende de cada sensor y se realiza como parte de las rutinas de GSI.

De acuerdo a @zhu2019 los coeficientes utilizados en la corrección del bias deben ser entrenados durante un periodo largo de tiempo y muestra que este tiempo es variable para cada canal de cada sensor. Con esto en mente se propone el siguiente experimento para el entrenamiento de los coeficientes.

### Diseño experimental

Para obtener coeficientes que permitan la correcta asimilación de observaciones de radianzas durante el periodo 18 UTC 20/11/2018 - 12 UTC 23/11/2018 se realizará asimilación continua entre las 18 UTC 11/11/2018 y las 12 UTC del 20/11/2018.

En este periodo de entrenamiento se observaron distintos fenómenos de tiempo severo asociados a sistemas convectivos de mesoescala y periodos de menor actividad por lo que se espera que la variedad de condiciones atmósfericas favorescan el entrenamiento de coeficientes pese al corto periodo de tiempo (9 días). Para garantizar esto y siguiendo a @zhu2019 este entrenamiento se realizará dos veces. Para el primer ciclo de asimilación se utilizaron los coeficientes de corrección de bias utilizados por GFS para la misma hora con la intención de acelerar el entramiento en comparación con una inicialización desde "0". Al finalizar el primer período de asimilación (experimento pre-RAD1) los coeficientes resultantes del último ciclo son utilizados al comenzar el segundo periodo de asimilación en las mismas fechas (pre-RAD2).

El experimento de asimilación se realizará en el dominio que se muestra en la Figura \@ref(fig:dominio) realizando ciclos contínuos a cada hora. El dominio tiene una reticula horizontal de 10 km y el tope del modelo en 50 hPa. En cada ciclo se incluyen las observaciones disponibles en la ventana de asimilación de 1 hora centrada en la hora del análisis incluyendo observaciones convencionales en el prepBUFR, observaciones de estaciones automáticas en el dominio, vientos derivados de satélite y radianzas de sensores en satélites polares listados en la Tabla \@ref(tab:sensores).

```{r dominio, fig.cap="Dominio horizontal del modelo"}
fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>%

  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  # geom_point(aes(color = hgt), size = 5) +
  scale_fill_distiller(name = NULL,
                       type = "seq",
                       palette = "RdYlGn",
                       direction = -1,
                       breaks = seq(0, 6000, 500),
                       guide = guide_colorstrip(barwidth = 20,
                                                barheight = 0.8)) +
  geom_sf(data = map, inherit.aes = FALSE, fill = NA, size = 0.5) +
  coord_sf(ylim = c(-41, -20), xlim = c(-75, -52.5)) +
  labs(
    x = "Longitude", 
    y = "Latitude") +
  theme_minimal() +
  theme(legend.position = "bottom") 
```

```{r sensores}
fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>%
  setnames("!sensor/instr/sat", "sensor_plataforma") %>% 
  .[iuse == 1 & sensor_plataforma %in% sensores, .N, by = .(sensor_plataforma)] %>% 
  separate(col = sensor_plataforma, into = c("sensor", "plataforma"), sep = "_") %>% 
  arrange(sensor) %>% 
  kable(col.names = c("Sensor", "Plataforma", "N° de canales asimilables"),
        caption = "Lista de sensores con la cantidad de canales aceptados por GSI") %>% 
  collapse_rows(columns = 1, valign = "top")  %>% 
  kable_classic_2()
```

#### Radianzas

Ver cantidad de observaciones asimiladas en cada canal y la altura de cada una ¿En todo el periodo?

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2020-10-08 R/read_diag_rad.R
Agrego funcion read_diag_rad y sus dependencias

@@ -1,0 +1,35 @@
+#' Lectura de archivos diag para radianzas
+#'
+#' Internamente define el sensor, plataforma, la fecha y el miembro del ensamble asociado.
+#' El formato de nombre de archivo debería ser "asim_{sensor}_{plat}_{date}.{mem}"
+#'
+#' @param file_list lista de archivos
+#' @param exp nombre del experimento asociado
+#' @export
+read_diag_rad <- function(file_list, exp) {
+
+  diag <- purrr::map(seq_along(file_list), function(f){
+    if (file.size(file_list[f]) == 0) {
+      return(NULL)
+    }
+    meta <- unglue::unglue(basename(file_list[f]), "asim_{sensor}_{plat}_{date}.{mem}")
+    # print(f)
+    out <- data.table::fread(file_list[f])
+    # .[V10 == 1] %>%
+
+    if (file.size(file_list[f]) != 0) {
+      out[, ":="(date = lubridate::ymd_hms(meta[[1]][["date"]]),
+                 mem = meta[[1]][["mem"]],
+                 exp = exp)]
+    }
+    out
+  }) %>%
+    rbindlist()
+
+  colnames(diag) <- c("sensor", "channel", "freq", "lat", "lon", "press", "elev_sup", "dhr", "tb_obs", "tbc", "tbcnob",
+                      "errinv", "qc", "emis", "tlapchn", "rzen", "razi", "rlnd", "rice", "rsnw", "rcld",
+                      "rcldp", paste0("pred", seq(8)), "date", "mem", "exp")
+  return(diag)
+}
+
+.datatable.aware <- TRUE

### 2020-10-08 temp_R/help_functions.R

@@ -7,70 +7,70 @@
 read_diag <- function(path, variable) {
   library(foreach)
   library(doParallel)
-  
+
   files <- Sys.glob(paste0(path, ".mem*"))
-  
+
   myCluster <- makeCluster(20)
   registerDoParallel(myCluster)
-  
+
   obs <- foreach(f = 1:length(files),
                  .packages = c("data.table", "metR", "lubridate", "tidyverse"),
                  .export = c("files", "variable"),
                  .combine = "rbind") %dopar% {
-                   
+
                    # sink("log.txt", append=TRUE)
-                   
-                   diag <- fread(files[f]) %>% 
-                     .[V10 == 1] %>% 
-                     .[, exp := basename(dirname(files[f]))] %>% 
-                     .[, mem := str_extract(files[f], "\\d{3}$")] %>% 
-                     .[, date := ymd_hms(str_extract(files[f], "\\d{14}"))] %>% 
+
+                   diag <- fread(files[f]) %>%
+                     .[V10 == 1] %>%
+                     .[, exp := basename(dirname(files[f]))] %>%
+                     .[, mem := str_extract(files[f], "\\d{3}$")] %>%
+                     .[, date := ymd_hms(str_extract(files[f], "\\d{14}"))] %>%
                      .[, c("V2", "V4") := NULL]
-                   
+
                    # cat("Archivo ", basename(files[f]))
-                   
+
                    colnames(diag) <- c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag.prep", "obs", "obs.guess", "obs2", "obs.guess2", "rerr", "exp", "mem", "date")
-                   
+
                    if ("uv" %in% variable & length(variable) == 1) {
-                     
-                     diag <- diag[var == "uv"] %>% 
-                       melt(measure.vars = c("obs", "obs2", "obs.guess", "obs.guess2")) %>% 
-                       .[, var := if_else(str_detect(variable, "2"), "v", "u")] %>% 
-                       .[, variable := str_remove(variable, "2")] 
-                     
-                     vars <- rlang::syms(setdiff(names(diag), "value")) 
-                     diag <- diag %>% 
[…]

### 2020-10-08 analysis/chapters/coeficientes.Rmd
Algunos graficos para contar cantidad de observaciones

@@
 library(here)
+library(patchwork)
 library(knitr)
@@
 
+La información de radianzas provienen de los sensores AMSU-A, AIRS, MHS, AIRS, IASI, HIRS todos en satélites de orbita polar.
+Estas observaciones son parte del sistema de transferencia de datos que utiliza GFS disponibles cada 6 horas pero asimiladas en ciclos horarios utilizando las observaciones correspondientes a una ventana centrada en la hora del análisis.
+
+Se utilizaron observaciones en aire claro (ver métodos para detectar nubles usado por GSI) y se aplicó un control de calidad específico para cada sensor a fin de rechazar observaciones erroneas.
+A si mismo para tener en cuenta la resolución de cada observación se aplicó un thining de 60 km.
+Este valor fue definido siguiendo la metodología de otros trabajos en el área (@zhu2016, @jones2020, @lin2017a, and @singh2016qms)
+
 Ver cantidad de observaciones asimiladas en cada canal y la altura de cada una ¿En todo el periodo?
+
+```{r n_rad, fig.cap="Cantidad de observaciones potencialmene asimilables en cada ciclo de asimilación para el experimento pre-RAD1"}
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+                    pattern = "asim.", recursive = TRUE, 
+                    full.names = TRUE)
+files <- files[!str_detect(files, pattern = "conv")]
+
+diag_rad <- purrr::map(files, function(f) {
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/{date_ana}/diagfiles/{file}" )
+  read_diag_rad(f, exp = meta[[1]][["exp"]]) %>%
+    .[press > 50 & qc == 0] %>% 
+    .[]
+
+}) %>%
+  rbindlist()
+
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[sensor %in% c("airs", "iasi")] %>% 
+  .[, .N, by = .(date, channel, sensor)] %>% 
+  ggplot(aes(channel, N)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "Canal", y = "Observaciones por ciclo") +
+  theme_minimal() +
+
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[!(sensor %in% c("airs", "iasi"))] %>% 
+  .[, .N, by = .(date, channel, sensor)] %>% 
+  ggplot(aes(channel, N)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "Canal", y = "Observaciones por ciclo") +
+  theme_minimal() +
+  plot_layout(ncol = 1)
+
+```
+
+```{r level_rad}
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[sensor %in% c("airs", "iasi")] %>% 
+  # .[, .N, by = .(date, channel, sensor)] %>% 
+  ggplot(aes(channel, press)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "Canal", y = "Presión (hPa)") +
+  theme_minimal()
+
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[!(sensor %in% c("airs", "iasi"))] %>% 
+  # .[, .N, by = .(date, channel, sensor)] %>% 
+  ggplot(aes(channel, press)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "Canal", y = "Presión (hPa)") +
+  theme_minimal() +
+  plot_layout(ncol = 1)
+```
 

### 2020-10-09 R/read_diag_rad.R
Agrego read_satbias

@@ -26,7 +26,7 @@ read_diag_rad <- function(file_list, exp) {
   }) %>%
     rbindlist()
 
-  colnames(diag) <- c("sensor", "channel", "freq", "lat", "lon", "press", "elev_sup", "dhr", "tb_obs", "tbc", "tbcnob",
+  colnames(diag) <- c("sensor", "channel", "freq", "lat", "lon", "peakwt", "press", "dhr", "tb_obs", "tbc", "tbcnob",
                       "errinv", "qc", "emis", "tlapchn", "rzen", "razi", "rlnd", "rice", "rsnw", "rcld",
                       "rcldp", paste0("pred", seq(8)), "date", "mem", "exp")
   return(diag)

### 2020-10-09 R/read_satbias.R

@@ -1,0 +1,38 @@
+#' Lee archivo "satbias" de coeficientes
+#'
+#' Esta función parsea el archivo satbias y lo ordena en formato tidy.
+#'
+#' @param file archivo a leer
+#' @export
+read_satbias <- function(file) {
+
+  string <- read_lines(file)
+
+  string <- split_chunks(string)
+
+  satbias <- purrr::map(string, parse_chunk) %>%
+    do.call(rbind, .) %>%
+    as.data.frame() %>%
+    setNames(c("id", "sensor", "channel", "tlp1", "tlp2", "nc",
+               paste0("coeff", seq(12)))) %>%
+    dplyr::mutate(across(channel:coeff12, ~as.numeric(as.character(.x))))
+
+  return(satbias)
+}
+
+split_chunks <- function(string) {
+
+  start <- seq(1, length(string), 3)
+  end <- c(start-1, length(string))[-1]
+
+  obs_type <- stringr::str_extract(string[start], "\\d+")
+
+  chunks <- purrr::map(seq_along(start), ~ string[(start[.x]):end[.x]])
+  chunks
+}
+
+parse_chunk <- function(chunk) {
+  stringr::str_split(chunk, " ") %>%
+    purrr::map( ~ stringr::str_subset(.x, "^$", negate = TRUE)) %>%
+    unlist(., recursive = FALSE)
+}

### 2020-10-09 temp_R/help_functions.R

@@ -178,39 +178,39 @@ read_diag_mean <- function(path, variable = c("uv", "p", "t", "q")) {
 
 # read satbias ------------------------------------------------------------
 
-read_satbias <- function(file) {
-
-  library(tidyverse)
-  library(data.table)
-
-  string <- read_lines(file)
-  split_chunks <- function(string) {
-    start <- seq(1, length(string), 3)
-    end <- c(start-1, length(string))[-1]
-
-    obs_type <- stringr::str_extract(string[start], "\\d+")
-
-    chunks <- purrr::map(seq_along(start), ~ string[(start[.x]):end[.x]])
-    chunks
-  }
-
-  string <- split_chunks(string)
-
-  parse_chunk <- function(chunk) {
-    str_split(chunk, " ") %>%
-      map( ~ str_subset(.x, "^$", negate = TRUE)) %>%
-      unlist(., recursive = FALSE)
-  }
-
-  satbias <- map(string, parse_chunk) %>%
-    do.call(rbind, .) %>%
-    as.data.frame() %>%
-    setNames(c("id", "sensor", "channel", "tlp1", "tlp2", "nc",
-               paste0("coeff", seq(12)))) %>%
-    mutate(across(channel:coeff12, ~as.numeric(as.character(.x))))
-
-  return(satbias)
-}
+# read_satbias <- function(file) {
+#
+#   library(tidyverse)
+#   library(data.table)
+#
+#   string <- read_lines(file)
+#   split_chunks <- function(string) {
+#     start <- seq(1, length(string), 3)
+#     end <- c(start-1, length(string))[-1]
+#
+#     obs_type <- stringr::str_extract(string[start], "\\d+")
+#
+#     chunks <- purrr::map(seq_along(start), ~ string[(start[.x]):end[.x]])
+#     chunks
+#   }
+#
+#   string <- split_chunks(string)
+#
+#   parse_chunk <- function(chunk) {
+#     str_split(chunk, " ") %>%
+#       map( ~ str_subset(.x, "^$", negate = TRUE)) %>%
+#       unlist(., recursive = FALSE)
+#   }
+#
+#   satbias <- map(string, parse_chunk) %>%
+#     do.call(rbind, .) %>%
+#     as.data.frame() %>%
+#     setNames(c("id", "sensor", "channel", "tlp1", "tlp2", "nc",
+#                paste0("coeff", seq(12)))) %>%
+#     mutate(across(channel:coeff12, ~as.numeric(as.character(.x))))
[…]

### 2020-10-09 analysis/chapters/coeficientes.Rmd
Evolución de coeficientes amsua

--- 2020-10-06 analysis/chapters/coeficientes.Rmd
@@
 title: "Coeficientes para la corrección de bias"
-output: bookdown::html_document2
+output: 
+  bookdown::html_document2: default
+  bookdown::pdf_document2: default
 bibliography: mesoda_ref.bib
@@
 library(tidyverse)
+library(lubridate)
 library(data.table)
 library(here)
+library(patchwork)
+library(unglue)
 library(knitr)
@@
 
+Distintos trabajos abordan el entrenamiento de los coeficientes de corrección de bias en dominios regionales donde la cobertura de observaciones satélitales no es homogenea de manera similar.
+Por ejemplo @schwartz2012 utiliza un "spin up" de una semana para el entrenamiento de los coeficientes utilizando un método variacional con ciclos de asimilación cada 6 horas en un dominio regional mientras que @liu2012 utiliza un método "offline" durante 3 meses para obtener coeficientes de corrección estables.
+En particular @zhu2019 propone un periodo de entrenamiento de 9 días utilizando un método variacional determina que es necesario repetir este periodo una vez más para obtener coeficientes estables.
+
 Para obtener coeficientes que permitan la correcta asimilación de observaciones de radianzas durante el periodo 18 UTC 20/11/2018 - 12 UTC 23/11/2018 se realizará asimilación continua entre las 18 UTC 11/11/2018 y las 12 UTC del 20/11/2018.
@@
 
-Ver cantidad de observaciones asimiladas en cada canal y la altura de cada una ¿En todo el periodo?
+La información de radianzas provienen de los sensores AMSU-A, AIRS, MHS, AIRS, IASI, HIRS todos en satélites de orbita polar.
+Estas observaciones son parte del sistema de transferencia de datos que utiliza GFS disponibles cada 6 horas pero asimiladas en ciclos horarios utilizando las observaciones correspondientes a una ventana centrada en la hora del análisis.
+
+Se utilizaron observaciones en aire claro (ver métodos para detectar nubles usado por GSI) y se aplicó un control de calidad específico para cada sensor a fin de rechazar observaciones erroneas.
+A si mismo para tener en cuenta la resolución de cada observación se aplicó un thining de 60 km.
+Este valor fue definido siguiendo la metodología de otros trabajos en el área (@zhu2016, @jones2020, @lin2017a, and @singh2016qms).
+
+En la Figura \@ref(fig:n-rad) se muestra la cantidad de observaciones potencialmente asimilables, es decir, que pasan el control de calidad del operador de las observaciones, para cada canal en cada ciclo de asimilación.
+Dejando de lado IASI que no se ve nada porque tiene más de 600 canales, se puede ver que en todo los casos hay outliers.
+Esto tiene que ver con que en cada ciclo solo hay disponibles observaciones de algunos sensores y no siempre la pasada del sensor coincide completamente con el dominio.
+Los momenos donde la cantidad de observaciones aumenta considerablemente coinciden con los momentos donde la pasada del satélite estaba centrada en el dominio.
+En general hay más observaciones en el rango infrarrojo tanto por las observaciones de los sensores multiespectrales (AIRS e IASI) como por el sensor HIRS que en general para todos los canales aporta mayor cantidad de observaciones.
+
+```{r}
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+                    pattern = "asim.", recursive = TRUE, 
+                    full.names = TRUE)
+files <- files[!str_detect(files, pattern = "conv")]
+
+diag_rad <- purrr::map(files, function(f) {
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/{date_ana}/diagfiles/{file}" )
+  read_diag_rad(f, exp = meta[[1]][["exp"]]) %>%
+    .[press > 50 & qc == 0] %>% 
+    .[]
+
+}) %>%
+  rbindlist()
+```
+
+```{r n-rad, fig.cap="Cantidad de observaciones potencialmene asimilables en cada ciclo de asimilación para el experimento pre-RAD1"}
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[sensor %in% c("airs", "iasi")] %>% 
+  .[, .N, by = .(date, channel, sensor)] %>% 
+  ggplot(aes(channel, N)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "", y = "Observaciones por ciclo") +
+  theme_minimal() +
+
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[!(sensor %in% c("airs", "iasi"))] %>% 
+  .[, .N, by = .(date, channel, sensor)] %>% 
+  ggplot(aes(channel, N)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "Canal", y = "Observaciones por ciclo") +
+  theme_minimal() +
+  plot_layout(ncol = 1)
+
+```
+
+En cuanto a la altura de las observaciones (Figura \@ref(fig:leve-rad)), determinada a partir del máximo de la función de peso, muchas observaciones de canales de todos los sensores menos MHS correspoenden a alturas por encima del tope del modelo (50 hPa).
+Respecto de observaciones en el rango de microondas el sensor ATMS aporta la mayor cantidad de observaciones en niveles medios y altos y lo mismo ocurre con AMSU-A.
+En cuanto a las observaciones en el infrarrojo AIRS aporta muchas observaciones entre superficie y 200 hPa mientras que los canales de HIRS4 con observaciones en niveles medios y bajos muestran un comportamiento más variane y aporta en general menos observaciones.
+
+TODO: ¿Qué está pasando con el nivel de presión de IASI? Hay una franja negra en 300 hPa que tiene pinta sospechosa.
+
+```{r level-rad, fig.cap="Nivel de presión asociado a las observaciones potencialmene asimilables en cada ciclo de asimilación para el experimento pre-RAD1"}
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[sensor %in% c("airs", "iasi")] %>% 
+  ggplot(aes(channel, peakwt)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  scale_y_level(name = "Presión (hPa)") +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "", y = "") +
+  theme_minimal() +
+
+diag_rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[!(sensor %in% c("airs", "iasi"))] %>% 
+  ggplot(aes(channel, peakwt)) +
+  geom_boxplot(aes(group = factor(channel))) +
+  scale_x_continuous(breaks = seq) +
+  scale_y_level(name = "Presión (hPa)") +
+  facet_wrap(~sensor, scales = "free") +
+  labs(x = "Canal") +
+  theme_minimal() +
+  plot_layout(ncol = 1)
+```
+
+## Evolución de los coeficientes de corrección.
+
+El entrenamiento de los coeficientes varia sensor a sensor y canal a canal, de acuerdo a lo observado en la bibliografía el tiempo de entrenamiento necesario es variable en cada caso.
+A continuación se muestran algunos ejemplos.
+
+Ver algunos ejemplos para cada sensor (solo alguna plataforma y algún canal), capaz incluyendo amsua n15 para cmparar con zhu 2019
+
+```{r}
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA",
+                    pattern = "satbias_2", 
+                    recursive = TRUE,
+                    full.names = TRUE)
+
+satbias <- map(files, function(f){
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/{date_ana}/satbias_enkf/satbias_{date}")
+  read_satbias(f) %>% 
+    as.data.table() %>% 
+    .[, ":="(date = ymd_hms(meta[[1]][["date"]]),
+             exp = meta[[1]][["exp"]])]
+
+}) %>% 
+  rbindlist()
+
+satbias <- satbias[sensor %in% sensores]
+
+satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
+  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
+
+satbias <- satinfo[satbias, on = c("sensor", "channel")]
+
+this_sensor <- sensores[1]
+
+satbias[sensor == this_sensor & iuse == 1] %>% 
+  .[channel %in% c(5:9)] %>%
+    melt(measure.vars = paste0("coeff", c(1, 3:5, 8))) %>% 
+    ggplot(aes(date, value)) +
+    geom_path(aes(color = variable, linetype = exp), size= 0.5) +
+    scale_color_brewer(type = "qual", palette = "Dark2",
+                       labels = c("coeff1" = "Offset",
+                                  "coeff3" = "CLW",
+                                  "coeff4" = "TLR^2",
+                                  "coeff5" = "TLR",
+                                  "coeff8" = "Emiss")) +
+    scale_linetype_discrete(name = "", labels = c("pre-RAD1", "pre-RAD2")) +
+    scale_x_datetime(date_labels = "%d") +
+    labs(title = paste("Coeficientes de", this_sensor),
+         color = "", x = "Fecha", y = "") +
+    facet_wrap(~channel, scales = "free_y", ncol = 5) +
+    theme_minimal() +
+    theme(legend.position = "bottom")
+
+```
 

### 2020-10-13 analysis/chapters/coeficientes.Rmd
Bias

@@
 	fig.path = "../figures/",
+	fig.retina = TRUE,
+	fig.height = 5,
 	message = FALSE,
@@
 )
+cyan <- rgb(0, 139, 139, maxColorValue = 255)
 ```
@@
 
-En cuanto a la altura de las observaciones (Figura \@ref(fig:leve-rad)), determinada a partir del máximo de la función de peso, muchas observaciones de canales de todos los sensores menos MHS correspoenden a alturas por encima del tope del modelo (50 hPa).
+En cuanto a la altura de las observaciones (Figura \@ref(fig:level-rad)), determinada a partir del máximo de la función de peso, muchas observaciones de canales de todos los sensores menos MHS correspoenden a alturas por encima del tope del modelo (50 hPa).
 Respecto de observaciones en el rango de microondas el sensor ATMS aporta la mayor cantidad de observaciones en niveles medios y altos y lo mismo ocurre con AMSU-A.
@@
 El entrenamiento de los coeficientes varia sensor a sensor y canal a canal, de acuerdo a lo observado en la bibliografía el tiempo de entrenamiento necesario es variable en cada caso.
-A continuación se muestran algunos ejemplos.
-
-Ver algunos ejemplos para cada sensor (solo alguna plataforma y algún canal), capaz incluyendo amsua n15 para cmparar con zhu 2019
+Al mismo tiempo al ser asimilación regional y horaria podría afectar el tiempo que demora la estabilización de los coeficientes de corrección.
 
@@
 
+```
+
+Por ejemplo en la Figura \@ref(fig:coef-amsua-n15) se muestra la evolución de algunos de los coeficientes de canales del AMSU-A en NOAA15 en linea llena para la primera semana de entrenamiento pre-RAD1 y en línea punteada para la segunda semana pre-RAD2.
+Al igual que @zhu2019 se observa que cada coeficiente se estabiliza a una velocidad distinta, por ejemplo el coeficiente de emisividad en el canal 8 si bien hacia el final de la semana 2 muestra un comportamiento similar a la semana 1, los valores no son tan cercanos mientras que el TLR tiene un comportamiento muy similar en casi todo el periodo.
+Lamentablemente @zhu2019 no muestra otros canales (solo canal 5) pero es esperable que cada uno muestre variaciones.
+
+En general algunos coeficientes para algunos canales no requieren casi entrenamiento mientras que a otros capaz les vendría bien una semanita más.
+
+[link a gráficos de todos los sensores y canales](https://www.dropbox.com/sh/a64be0ognp0js0a/AAC6r8AXAGZ45dNFdvF1r8AEa?dl=0)
+
+```{r coef-amsua-n15, fig.cap="Coeficientes de corrección de bias en cada ciclo de asimilación para los canales 5 a 9 del sensor AMSU-A de NOAA-15." }
 this_sensor <- sensores[1]
@@
     theme(legend.position = "bottom")
+```
+Otro ejemplo se observa en la Figura \@ref(fig:coef-mhs-metop-a) correspondiente a los coeficientes el sensor MHS a bordo del satelite METOP-A.
+En este caso por ejemplo para el canal 3 se observa como el Offset se mantuvo aprximadamente constante luego de 6 días de entrenamiento mientras que el TLR en el canal 5 continua variando aún luego de la segunda semana.
 
+```{r coef-mhs-metop-a, fig.cap="Coeficientes de corrección de bias en cada ciclo de asimilación para los canales del sensor MHS en METOP-A." }
+this_sensor <- sensores[12]
+
+satbias[sensor == this_sensor & iuse == 1] %>% 
+  .[channel %in% c(1:5)] %>%
+    melt(measure.vars = paste0("coeff", c(1, 3:5, 8))) %>% 
+    ggplot(aes(date, value)) +
+    geom_path(aes(color = variable, linetype = exp), size= 0.5) +
+    scale_color_brewer(type = "qual", palette = "Dark2",
+                       labels = c("coeff1" = "Offset",
+                                  "coeff3" = "CLW",
+                                  "coeff4" = "TLR^2",
+                                  "coeff5" = "TLR",
+                                  "coeff8" = "Emiss")) +
+    scale_linetype_discrete(name = "", labels = c("pre-RAD1", "pre-RAD2")) +
+    scale_x_datetime(date_labels = "%d") +
+    labs(title = paste("Coeficientes de", this_sensor),
+         color = "", x = "Fecha", y = "") +
+    facet_wrap(~channel, scales = "free_y", ncol = 5) +
+    theme_minimal() +
+    theme(legend.position = "bottom")
+```
+
+Si observamos la corrección de bias para cada canal de la familia de sensores ANSU-A se ven diferencias interesantes entre los dos periodos pre-RAD1 y pre-RAD2.
+Para analizar estas diferencias en la Figura \@ref(fig:bias-amsua) hay que tener en cuenta que no todas las observaciones cumplen las condiciones para ser asimiladas y por lo tanto el BIAS no es corregido o si es corregido los coeficientes no son actualizados en cada ciclo.
+En particular los canales 4 a 7 muestran una mejora en la mediana del bias (más cercana a cero) y/o en la dispersión (menor rango intercuartil y/o menos valores extremos).
+```{r bias-amsua, fig.cap="Bias de las radianzas de AMSU-A en las distintas plataformas y para los dos periodos."}
+
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+                    pattern = "asim.", recursive = TRUE, 
+                    full.names = TRUE)
+files <- files[str_detect(files, pattern = "amsua")]
+
+diag_rad <- purrr::map(files, function(f) {
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/{date_ana}/diagfiles/{file}" )
+  read_diag_rad(f, exp = meta[[1]][["exp"]]) %>%
+    .[]
+
+}) %>%
+  rbindlist()
+
+diag_rad[sensor %in% sensores[1:5] & qc == 0] %>% 
+  ggplot(aes(factor(channel), tbc - tbcnob)) +
+  geom_hline(yintercept = 0, color = "darkgrey") +
+  geom_boxplot(aes(color = exp)) +
+  scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
+  labs(x = "Canal", y = "BIAS (K)",
+       color = "") +
+  facet_wrap(~sensor) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+La corrección del bias para el sensor MHS en cada periodo es menos clara que para AMSU-A (Figura \@ref(fig:bias-mhs)).
+Los cambios en pre-RAD2 son mínimos y en algunos casos negativos.
+Por ejemplo el canal 1 que tiene el máximo cerca de superficie es el que muestra mayor dispersión, esto mejora mucho para el sensor sobre METOP-A y B (aunque en este ultimo caso la mediana se aleja de cero) pero para NOAA19 la dispersión no cambia particularmente (pero la mediana está más cerca de cero).
+
+```{r bias-mhs, fig.cap="Bias de las radianzas de MHS en las distintas plataformas y para los dos periodos."}
+
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+                    pattern = "asim.", recursive = TRUE, 
+                    full.names = TRUE)
+files <- files[str_detect(files, pattern = "mhs")]
+
+diag_rad <- purrr::map(files, function(f) {
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/{date_ana}/diagfiles/{file}" )
+  read_diag_rad(f, exp = meta[[1]][["exp"]]) %>%
+    .[]
+
+}) %>%
+  rbindlist()
+
+diag_rad[qc == 0] %>% 
+  ggplot(aes(factor(channel), tbc - tbcnob)) +
+  geom_hline(yintercept = 0, color = "darkgrey") +
+  geom_boxplot(aes(color = exp)) +
+  scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
+  labs(x = "Canal", y = "BIAS (K)",
+       color = "") +
+  facet_wrap(~sensor) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
 ```
@@
 ```
+
+<script src="https://hypothes.is/embed.js" async></script>

### 2020-10-15 analysis/chapters/coeficientes.Rmd
Mañana intento de nuevo

@@
   theme(legend.position = "bottom")
+
+diag_rad[sensor %in% sensores[1] & qc == 0 & channel %in% c(7:8)] %>% 
+  .[, .(mean_bias = mean(tbc - tbcnob, na.rm = TRUE),
+        mean_mag_bias = mean(abs(tbc - tbcnob), na.rm = TRUE),
+        n = .N), by = .(date, sensor, exp, channel)] %>% 
+  ggplot(aes(date, mean_bias)) +
+  geom_hline(yintercept = 0, color = "darkgrey") +
+  geom_line(aes(color = exp, group = interaction(exp, channel))) +
+  geom_point(aes(color = exp, size = n)) +
+  scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
+  scale_size_area(max_size = 4) +
+  labs(x = "Canal", y = "BIAS (K)",
+       color = "") +
+  facet_wrap(~channel) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
 ```

### 2020-10-16 analysis/chapters/coeficientes.Rmd
Fin del bias

@@
   fig.retina = TRUE,
-	fig.height = 5,
+  # fig.height = 3,
   message = FALSE,
@@
 map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")
+
+geom_mapa <- function() {
+  list(geom_sf(data = map, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE),
+       coord_sf(xlim = c(-76, -52), ylim = c(-41, -20)))
+}
 
@@
 
-```{r dominio, fig.cap="Dominio horizontal del modelo"}
+```{r dominio, fig.cap="Dominio horizontal del modelo", fig.height=4}
 fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
@@
 
-## Evolución de los coeficientes de corrección.
+### Evolución de los coeficientes de corrección.
 
@@
 
-```{r coef-amsua-n15, fig.cap="Coeficientes de corrección de bias en cada ciclo de asimilación para los canales 5 a 9 del sensor AMSU-A de NOAA-15." }
+```{r coef-amsua-n15, fig.cap="Coeficientes de corrección de bias en cada ciclo de asimilación para los canales 5 a 9 del sensor AMSU-A de NOAA-15.", fig.height=3}
 this_sensor <- sensores[1]
@@
 
-```{r coef-mhs-metop-a, fig.cap="Coeficientes de corrección de bias en cada ciclo de asimilación para los canales del sensor MHS en METOP-A." }
+```{r coef-mhs-metop-a, fig.cap="Coeficientes de corrección de bias en cada ciclo de asimilación para los canales del sensor MHS en METOP-A.", fig.height=3}
 this_sensor <- sensores[12]
@@
 
-Si observamos la corrección de bias para cada canal de la familia de sensores ANSU-A se ven diferencias interesantes entre los dos periodos pre-RAD1 y pre-RAD2.
+### Corrección del bias
+
+#### AMSU-A
+
+Si observamos la corrección de bias para cada canal de la familia de sensores AMSU-A se ven diferencias interesantes entre los dos periodos pre-RAD1 y pre-RAD2.
 Para analizar estas diferencias en la Figura \@ref(fig:bias-amsua) hay que tener en cuenta que no todas las observaciones cumplen las condiciones para ser asimiladas y por lo tanto el BIAS no es corregido o si es corregido los coeficientes no son actualizados en cada ciclo.
 En particular los canales 4 a 7 muestran una mejora en la mediana del bias (más cercana a cero) y/o en la dispersión (menor rango intercuartil y/o menos valores extremos).
+
+La evolución de bias a lo largo del tiempo para el sensor sobre NOAA15 (Figura \@ref(fig:bias-temp-amsua)) muestra cambios entre los dos periodos en algunos casos del orden de 0.5 grados.
+En general el comportamiento del bias en el periodo pre-RAD2 mejora respecto del otro periodo y en los casos donde esto no ocurre as diferencias son muy pequeñas.
+
 ```{r bias-amsua, fig.cap="Bias de las radianzas de AMSU-A en las distintas plataformas y para los dos periodos."}
@@
 
+```
+
+```{r bias-temp-amsua, fig.cap="Bias medio de las radianzas de AMSU-A en las distintas plataformas y para los dos periodos a lo largo del tiempo.", fig.height=4}
+
 diag_rad[sensor %in% sensores[1] & qc == 0 & channel %in% c(7:8)] %>% 
@@
   labs(x = "Canal", y = "BIAS (K)",
-       color = "") +
+       color = "", size = "") +
   facet_wrap(~channel) +
@@
 ```
+
+En la Figura \@ref(fig:bias-pred-amsua) se muestra la contribución de cada predictor a la corrección total de bias, o sea lo que resulta de $\beta_i P_i(x)$, sin tener en cuenta la corrección asociada al ángulo (TODO: no lo estoy sacando de los diag files originales, capaz debería hacerlo).
+En primer lugar comparando con el trabajo de @zhu2019 (Fig.
+7) no se ve la variación casi cíclica pero esto muy posiblemente se deba a que en esos experimentos hay observaciones disponibles en cada ciclo de asimilación, lo que no ocurre en los que se presentan acá debido a la asimilación frecuente y al dominio que es más chico.
+En cuanto a valores, hay mayor similitud aún teniendo en cuenta las diferencias entre los distintos canales.
+
+Analizando específicamente los resultados de los experimentos, se ve que la contribución de cada predictor es distinta para los canales mostrados y además es distinta entre sensores.
+Por ejemplo Para el canal 5 del AMSU-A NOAA19 el predictor asociado a emisividad es el que más contribuye mientras que para NOAA18 el TLR^2 tambien contribuye.
+Esto se ve también en canal 7 donde a veces contribuye más el TLR, otras veces el TLR^2.
+
+```{r bias-pred-ansua, fig.cap="Evolución temporal del bias asociado a cada predictor para los sensores AMSU-A en las distintas plataformas", fig.height=3}
+
+diag_rad %>% 
+  .[, bias_total := pred1 + pred2 + pred3 + pred4 + pred5 + pred6 + pred7 + pred8] %>% 
+  melt(measure.vars = c("pred1", "pred2", "pred3", "pred4", 
+                        "pred5", "pred6", "pred7", "pred8", "bias_total")) %>% 
+  .[, .(mean_bias = mean(value, na.rm = TRUE)), by = .(variable, date, exp, channel, sensor)] %>% 
+  .[!(sensor %in% c("amsua_aqua", "amsua_metop-a")) & 
+      channel %in% c(5:8) & variable %in% c("pred1", "pred3", "pred4", "pred5", "pred8", "bias_total")] %>% 
+  ggplot(aes(date, mean_bias)) +
+  geom_line(aes(color = variable, linetype = exp)) +
+  scale_color_brewer(type = "qual", palette = "Dark2",
+                     labels = c("pred1" = "Offset",
+                                "pred3" = "CLW",
+                                "pred4" = "TLR^2",
+                                "pred5" = "TLR",
+                                "pred8" = "Emiss")) +
+  scale_linetype_discrete(name = "", labels = c("pre-RAD1", "pre-RAD2")) +
+  facet_grid(channel~sensor, scales = "free") +
+  labs(color = "", linetype = "",
+       x = "", y = "BIAS predicho") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
+La distribución espacial de la corrección de bias aportada por distintos predictores para un tiempo particular se puede observar en la Figura \@ref(fig:pred-amsua).
+En este caso el TLR^2 muestra valores comparables con la literatura y en particular un máximo que ocupa gran parte del dominio.
+El predictor asociado a la emisividad presenta valores distinto de cero solo en la zona de topografía.
+
+```{r pred-amsua, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 3 de MHS NOAA19 a las 8 UTC del 17/11/2018"}
+
+diag_rad[sensor == "amsua_n19" & date == as_datetime("2018-11-17 08:00:00") & channel == 8] %>% 
+  melt(measure.vars = paste0("pred", c(5))) %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = value)) +
+  scale_color_viridis_c() +
+  geom_mapa() +
+  facet_wrap(~variable) +
+  labs(x = "", y = "", color = "",
+       subtitle = "TLR^2") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  diag_rad[sensor == "amsua_n19" & date == as_datetime("2018-11-17 08:00:00") & channel == 8] %>% 
+  melt(measure.vars = paste0("pred", c(8))) %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = value)) +
+  scale_color_viridis_c() +
+  geom_mapa() +
+  # facet_wrap(~variable) +
+  labs(x = "", y = "", color = "",
+       subtitle = "Emisividad") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+#### MHS
+
 La corrección del bias para el sensor MHS en cada periodo es menos clara que para AMSU-A (Figura \@ref(fig:bias-mhs)).
@@
   theme(legend.position = "bottom")
+
+```
+
+De la misma manera la evolución temporal del bias para este sensor es muy variable y el periodo pre-RAD2 no muestran mejoras marcadas (Figura \@ref(fig:bias-temmp-mhs).
+Los valores de bias aumentan y disminuyen alternativamente [TODO: podría deberse a la cantidad de observaciones?
+eso parece cumplirse en el canal 4 pero se ve lo contrario en el canal 2].
+
+```{r bias-temmp-mhs, fig.cap="Bias medio de las radianzas de MHS para los dos periodos a lo largo del tiempo.", fig.height=4}
+
+diag_rad[qc == 0] %>% 
+  .[, .(mean_bias = mean(tbc - tbcnob, na.rm = TRUE),
+        mean_mag_bias = mean(abs(tbc - tbcnob), na.rm = TRUE),
+        n = .N), by = .(date, sensor, exp, channel)] %>% 
+  ggplot(aes(date, mean_bias)) +
+  geom_hline(yintercept = 0, color = "darkgrey") +
+  geom_line(aes(color = exp, group = interaction(exp, channel))) +
+  geom_point(aes(color = exp, size = n)) +
+  scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
+  scale_size_area(max_size = 4) +
+  labs(x = "Canal", y = "BIAS (K)",
+       color = "", size = "") +
+  facet_wrap(~channel, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+En cuanto a la contribución para la corrección del bias por parte de los distintos predictores en MHS (Figura \@ref(fig:bias-pred-mhs)) se ven mayores diferencias entre los dos periodos.
+Llama poderosamente la atención la magnitud de corrección del bias para el canal 1 para metop a y b donde llega a valores de 30 K.
+El canal 5 muestra un pico que llegan a 5 K mientras que el canal 4 muestra essos valores de manera sostenida para todo el periodo pre-RAD2.
+Para el resto de los canales los valores son similares a lo observado para AMSU-A (Figura \@ref(fig:bias-pred-amsua)).
+Por la escala de las figuras no se observas demasiados detalles respecto de que predictor es el que más contribuye al bias total pero por ejemplo se puede ver que en el canal 3 y para todos los sensores el Offset parece ser el que tiene un mayor aporte.
+
+```{r bias-pred-mhs, fig.cap="Evolución temporal del bias asociado a cada predictor para los sensores MSH en las distintas plataformas", fig.height=5}
+
+diag_rad %>% 
+  .[, bias_total := pred1 + pred2 + pred3 + pred4 + pred5 + pred6 + pred7 + pred8] %>% 
+  melt(measure.vars = c("pred1", "pred2", "pred3", "pred4", 
+                        "pred5", "pred6", "pred7", "pred8", "bias_total")) %>% 
+  .[, .(mean_bias = mean(value, na.rm = TRUE)), by = .(variable, date, exp, channel, sensor)] %>% 
+  .[!(sensor %in% c("amsua_aqua", "amsua_metop-a")) & 
+      channel %in% c(1:5) & variable %in% c("pred1", "pred3", "pred4", "pred5", "pred8", "bias_total")] %>% 
+  ggplot(aes(date, mean_bias)) +
+  geom_line(aes(color = variable, linetype = exp)) +
+  scale_color_brewer(type = "qual", palette = "Dark2",
+                     labels = c("pred1" = "Offset",
+                                "pred3" = "CLW",
+                                "pred4" = "TLR^2",
+                                "pred5" = "TLR",
+                                "pred8" = "Emiss")) +
+  scale_linetype_discrete(name = "", labels = c("pre-RAD1", "pre-RAD2")) +
+  facet_grid(channel~sensor, scales = "free") +
+  labs(color = "", linetype = "",
+       x = "", y = "BIAS predicho") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
+La distribución espacial de la corrección del bias asociado a los predictores de TLR^2 y emisividad para un tiempo en particular se muestran en la Figura \@ref(fig:pred-mhs).
+En este tiempo la pasada del satélite cubrió gran parte del dominio.
+Respecto de la corrección asociada a la emisividad se observan valores distintos de cero en algunas regiones muy localizadas (sucede parecido en otros tiempos) en las zonas de mayor topografía.
+Esto podría deberse a que el nivel de presión asociado a esas observaciones se encuentra más cerca de la superficie?
+El TLR^2 tiene valores pequeños y varian de signo según la región.
+
+```{r pred-mhs, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 3 de MHS NOAA19 a las 8 UTC del 17/11/2018"}
+diag_rad[sensor == "mhs_n19" & date == as_datetime("2018-11-17 08:00:00") & channel == 3] %>% 
+  melt(measure.vars = paste0("pred", c(5))) %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = value)) +
+  scale_color_viridis_c() +
+  geom_mapa() +
+  # facet_wrap(~variable) +
+  labs(x = "", y = "", color = "",
+       subtitle = "TLR^2") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  diag_rad[sensor == "mhs_n19" & date == as_datetime("2018-11-17 08:00:00") & channel == 3] %>% 
+  melt(measure.vars = paste0("pred", c(8))) %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = value)) +
+  scale_color_viridis_c() +
+  geom_mapa() +
+  # facet_wrap(~variable) +
+  labs(x = "", y = "", color = "",
+       subtitle = "Emisividad") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+### Algunos casos de estudio
+
+#### 2018-11-13 01:00Z
+
+```{r}
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+                    pattern = "NPP_2018-11-13_01", recursive = TRUE, 
+                    full.names = TRUE)
+
+pp <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/NPP/NPP_{date}_AN.nc")
+  ReadNetCDF(f, vars = c("precipitacion" = "PP", 
+                                "XLAT", 
+                                "XLONG")) %>% 
+    .[, .(precipitacion = mean(precipitacion, na.rm = TRUE)), by = .(XLAT, XLONG)] %>% 
+    .[, c("x", "y") := wrf_project(XLONG, XLAT, inverse = FALSE)] %>% 
+    .[, exp := meta[[1]][[1]]] %>% 
+    .[]
+}) %>%
+  rbindlist()
+
+pp %>% 
+  dcast(XLAT + XLONG + x + y ~ exp, value.var = "precipitacion") %>% 
+  .[, diff := week2 - week1] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = diff), proj = norargentina_lambert) +
+  scale_fill_divergent() +
+  # scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  theme_minimal()
+
+pp %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = precipitacion), proj = norargentina_lambert,
+                    breaks = seq(0, 25, 2.5)) +
+
+  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  facet_wrap(~exp) +
+theme_minimal()
+
 ```

### 2020-10-16 analysis/chapters/coeficientes.Rmd
casos de estudio

@@
 
-```{r bias-temp-amsua, fig.cap="Bias medio de las radianzas de AMSU-A en las distintas plataformas y para los dos periodos a lo largo del tiempo.", fig.height=4}
+```{r bias-temp-amsua, fig.cap="Bias medio de las radianzas de AMSU-A NOAA 15 en las distintas plataformas y para los dos periodos a lo largo del tiempo.", fig.height=4}
 
@@
 
-```{r pred-amsua, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 3 de MHS NOAA19 a las 8 UTC del 17/11/2018"}
+```{r pred-amsua, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 8 de AMSU-A NOAA19 a las 8 UTC del 17/11/2018"}
 
@@
                          "XLONG")) %>% 
-    .[, .(precipitacion = mean(precipitacion, na.rm = TRUE)), by = .(XLAT, XLONG)] %>% 
     .[, c("x", "y") := wrf_project(XLONG, XLAT, inverse = FALSE)] %>% 
@@
 
-pp %>% 
-  dcast(XLAT + XLONG + x + y ~ exp, value.var = "precipitacion") %>% 
-  .[, diff := week2 - week1] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = diff), proj = norargentina_lambert) +
-  scale_fill_divergent() +
-  # scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5)) +
-  geom_mapa() +
-  labs(fill = "", x = "", y = "") +
-  theme_minimal()
+imerg_path <- "/home/paula.maldonado/datosalertar1/RRA_VERIF/data/raw/imerg_raw"
 
-pp %>% 
+files <- list.files(imerg_path, pattern = "20181113-S00", full.names = TRUE)
+
+pp_imerg <- purrr::map(files, function(f) {
+  meta <- unglue(basename(f), "3B-HHR-L.MS.MRG.3IMERG.{dia}-S{hora_ini}-E{hora_fin}.{algo}.V05B.RT-H5")
+  ReadNetCDF(f, 
+             vars = c(pp = "Grid/precipitationCal"), 
+             subset = list("Grid/lon" = -80:-50, 
+                           "Grid/lat" = -45:-15)) %>% 
+    .[, fecha := ymd_hms(paste0(meta[[1]][[1]], meta[[1]][[2]]))]
+
+}) %>% 
+  rbindlist() %>% 
+  setnames(c("Grid/lon", "Grid/lat"), c("lon", "lat")) %>% 
+  .[, .(pp_acum = sum(pp)), by = .(lon, lat)]
+
+```
+
+```{r caso1, fig.cap="Precipitación acumulada en 60 minutos para cada experimento (arriba), diferencia entre los experimentos (abajo, izquierda) y precipitación acumulada según IMERG (abajo, derecha)", fig.height=7}
+p1 <- pp[, .(precipitacion = mean(precipitacion, na.rm = TRUE)), by = .(x, y, exp)] %>% 
   ggplot(aes(x, y)) +
@@
 
-  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5)) +
+  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5),
+                       limits = c(2.5, NA),
+                       na.value = "white",
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
   geom_mapa() +
   labs(fill = "", x = "", y = "") +
-  facet_wrap(~exp) +
-theme_minimal()
+  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
 
+p2 <- pp[, .(precipitacion = mean(precipitacion, na.rm = TRUE)), by = .(x, y, exp)] %>% 
+  dcast(x + y ~ exp, value.var = "precipitacion") %>% 
+  .[, diff := week2 - week1] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = diff), proj = norargentina_lambert, 
+                    breaks = AnchorBreaks(anchor = 0, binwidth = 2, 
+                                          exclude = 0)) +
+  scale_fill_divergent(breaks = AnchorBreaks(anchor = 0, binwidth = 2, 
+                                             exclude = 0), 
+                       guide = guide_colorstrip(barwidth = 10,
+                                                barheight = 1)) +
+  # scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p3 <- pp_imerg %>% 
+  ggplot(aes(lon, lat)) +
+  geom_contour_fill(aes(z = pp_acum), #proj = norargentina_lambert,
+                    breaks = seq(0, 45, 5)) +
+
+  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 45, 5),
+                       limits = c(5, NA),
+                       na.value = "white",
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+(p1 / (p2 + p3)) + plot_layout(widths = 1) 
+# plot_layout(ncol = 2, widths = c(3, 1), heights = c(1, 1))
+
+```
+
+```{r caso1-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento en el caso 1.", fig.height=7}
+pp[, .(prop = mean(precipitacion > 5)*100), by = .(exp, x, y)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = prop), proj = norargentina_lambert,
+                    breaks = seq(0, 100, 10)) +
+  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
+                       limits = c(10, 100),
+                       na.value = "white",
+                       breaks = seq(0, 100, 10),
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  ggtitle("Probabilidad de precipitación acumulada > 5 mm") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  pp[, .(prop = mean(precipitacion > 25)*100), by = .(exp, x, y)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = prop), proj = norargentina_lambert,
+                    breaks = seq(0, 100, 10)) +
+  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
+                       limits = c(10, 100),
+                       na.value = "white",
+                       breaks = seq(0, 100, 10),
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  ggtitle("Probabilidad de precipitación acumulada > 25 mm") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+  plot_layout(ncol = 1, guides = "collect", widths = 0.5, heights = 2) &
+  theme(legend.position = "bottom") 
+```
+
+#### 2018-11-18 10:00Z
+
+```{r}
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+                    pattern = "NPP_2018-11-18_10", recursive = TRUE, 
+                    full.names = TRUE)
+
+pp <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/NPP/NPP_{date}_AN.nc")
+  ReadNetCDF(f, vars = c("precipitacion" = "PP", 
+                         "XLAT", 
+                         "XLONG")) %>%
+    .[, c("x", "y") := wrf_project(XLONG, XLAT, inverse = FALSE)] %>% 
+    .[, exp := meta[[1]][[1]]] %>% 
+    .[]
+}) %>%
+  rbindlist()
+
+imerg_path <- "/home/paula.maldonado/datosalertar1/RRA_VERIF/data/raw/imerg_raw"
+
+files <- list.files(imerg_path, pattern = "20181118-S09", full.names = TRUE)
+
+pp_imerg <- purrr::map(files, function(f) {
+  meta <- unglue(basename(f), "3B-HHR-L.MS.MRG.3IMERG.{dia}-S{hora_ini}-E{hora_fin}.{algo}.V05B.RT-H5")
+  ReadNetCDF(f, 
+             vars = c(pp = "Grid/precipitationCal"), 
+             subset = list("Grid/lon" = -80:-50, 
+                           "Grid/lat" = -45:-15)) %>% 
+    .[, fecha := ymd_hms(paste0(meta[[1]][[1]], meta[[1]][[2]]))]
+
+}) %>% 
+  rbindlist() %>% 
+  setnames(c("Grid/lon", "Grid/lat"), c("lon", "lat")) %>% 
+  .[, .(pp_acum = sum(pp)), by = .(lon, lat)]
+
+```
+
+```{r caso2, fig.cap="Precipitación acumulada en 10 minutos para cada experimento (arriba), diferencia entre los experimentos (abajo, izquierda) y precipitación acumulada en 30 minutos según IMERG (abajo, derecha)", fig.height=7}
+p1 <- pp[, .(precipitacion = mean(precipitacion, na.rm = TRUE)), by = .(x, y, exp)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = precipitacion), proj = norargentina_lambert,
+                    breaks = seq(0, 25, 2.5)) +
+
+  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5),
+                       limits = c(2.5, NA),
+                       na.value = "white",
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p2 <- pp[, .(precipitacion = mean(precipitacion, na.rm = TRUE)), by = .(x, y, exp)] %>% 
+  dcast(x + y ~ exp, value.var = "precipitacion") %>% 
+  .[, diff := week2 - week1] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = diff), proj = norargentina_lambert, 
+                    breaks = AnchorBreaks(anchor = 0, binwidth = 1, 
+                                          exclude = 0)) +
+  scale_fill_divergent(breaks = AnchorBreaks(anchor = 0, binwidth = 1, 
+                                             exclude = 0), 
+                       guide = guide_colorstrip(barwidth = 10,
+                                                barheight = 1)) +
+  # scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 25, 2.5)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p3 <- pp_imerg %>% 
+  ggplot(aes(lon, lat)) +
+  geom_contour_fill(aes(z = pp_acum), #proj = norargentina_lambert,
+                    breaks = seq(0, 40, 5)) +
+
+  scale_fill_distiller(palette = "YlGnBu", direction = 1, breaks = seq(0, 40, 5),
+                       limits = c(5, NA),
+                       na.value = "white",
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+(p1 / (p2 + p3)) + plot_layout(widths = 1)
+
+```
+
+```{r caso2-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento correspondiente al caso 2", fig.height=7}
+pp[, .(prop = mean(precipitacion > 5)*100), by = .(exp, x, y)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = prop), proj = norargentina_lambert,
+                    breaks = seq(0, 100, 10)) +
+  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
+                       limits = c(10, 100),
+                       na.value = "white",
+                       breaks = seq(0, 100, 10),
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  ggtitle("Probabilidad de precipitación acumulada > 5 mm") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  pp[, .(prop = mean(precipitacion > 25)*100), by = .(exp, x, y)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = prop), proj = norargentina_lambert,
+                    breaks = seq(0, 100, 10)) +
+  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
+                       limits = c(10, 100),
+                       na.value = "white",
+                       breaks = seq(0, 100, 10),
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  geom_mapa() +
+  labs(fill = "", x = "", y = "") +
+  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  ggtitle("Probabilidad de precipitación acumulada > 25 mm") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+  plot_layout(ncol = 1, guides = "collect", widths = 0.5, heights = 2) &
+  theme(legend.position = "bottom") 
 ```

### 2020-10-20 analysis/chapters/coeficientes.Rmd
Bias residual

@@
 ```{r}
-files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7-training/ANA", 
                     pattern = "asim.", recursive = TRUE, 
@@
 diag_rad <- purrr::map(files, function(f) {
-  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/{date_ana}/diagfiles/{file}" )
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7-training/ANA/{exp}/{date_ana}/diagfiles/{file}" )
   read_diag_rad(f, exp = meta[[1]][["exp"]]) %>%
@@
 ```{r}
-files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA",
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7-training/ANA",
                     pattern = "satbias_2", 
@@
 
-  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/{date_ana}/satbias_enkf/satbias_{date}")
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7-training/ANA/{exp}/{date_ana}/satbias_enkf/satbias_{date}")
   read_satbias(f) %>% 
@@
 
-files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7-training/ANA", 
                     pattern = "asim.", recursive = TRUE, 
@@
 
-diag_rad %>% 
+diag_rad[qc == 0] %>% 
   .[, bias_total := pred1 + pred2 + pred3 + pred4 + pred5 + pred6 + pred7 + pred8] %>% 
@@
   theme(legend.position = "bottom")
+
+```
+
+Para estos sensores en particular llama la atención las diferencias entre los dos periodos, por ejemplo en el canal 4 de METOP-B hay un cambio de magnitud del bias que ocurre alrededor de 16 de noviembre y que se mantiene luego en pre-RAD2.
+Para explorar la posibilidad de que este cambio se deba realmente al entrenamiento de los coeficiente y que a partir de esos momentos la corrección del bias esté mejorando, en la Figura \@ref(fig:mhs-bias-res) se muestra la media de la innovación.
+En primer lugar se ven las diferencias bastante marcadas entre la media de la innovación con y sin bias corregido y que en la mayoría de los casos el primer es más cercano a cero.
+En particular para el canal 4 de METOP-B se obsera que la media de la innovación corregida luego de 16 de noviembre se aleja de cero y esto se mantiene también durante el periodo pre-RAD2.
+
+```{r mhs-bias-res, fig.cap="Valor medio de la diferencia entre las observaciones y el pronóstico (O - B) a lo largo del tiempo para cada periodo, sensor y canal"}
+diag_rad %>% 
+  .[qc == 0] %>% 
+  melt(measure.vars = c("tbc", "tbcnob")) %>% 
+  .[, .(ob_mean = mean(value, na.rm = TRUE),
+       ob_sd = sd(value, na.rm = TRUE)), by = .(date, sensor, channel, exp, variable)] %>% 
+  # melt(measure.vars = c("ob_mean", "ob_sd"),  variable.name = "estadistico", 
+  #      value.name = "valor") %>% 
+ggplot(aes(date, ob_mean)) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(color = exp)) +
+  geom_line(aes(color = exp, linetype = variable)) +
+  scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
+  scale_linetype_discrete(labels = c("Bias corregido", "Sin corrección")) +
+  facet_grid(channel~sensor, scales = "free_y") +
+  labs(y = NULL, x = NULL, color = NULL, linetype = NULL) +
+  theme_minimal()
 
@@
 ```{r}
-files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7-training/ANA", 
                     pattern = "NPP_2018-11-13_01", recursive = TRUE, 
@@
 
-  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7/ANA/{exp}/NPP/NPP_{date}_AN.nc")
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7-training/ANA/{exp}/NPP/NPP_{date}_AN.nc")
   ReadNetCDF(f, vars = c("precipitacion" = "PP", 
@@
 ```{r}
-files <- list.files("/home/paola.corrales/datosmunin/EXP/E7/ANA", 
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7-training/ANA", 
                     pattern = "NPP_2018-11-18_10", recursive = TRUE, 

### 2020-10-22 analysis/chapters/coeficientes.Rmd
Fin de analisis, falta revisar grafico de predictores para amsua

@@
   ggplot(aes(channel, N)) +
-  geom_boxplot(aes(group = factor(channel))) +
+  geom_boxplot(aes(group = factor(channel)), color = "darkorange", size = 0.1) +
   facet_wrap(~sensor, scales = "free") +
@@
   ggplot(aes(channel, N)) +
-  geom_boxplot(aes(group = factor(channel))) +
+  geom_boxplot(aes(group = factor(channel)), color = "darkorange") +
   facet_wrap(~sensor, scales = "free") +
@@
   ggplot(aes(channel, peakwt)) +
-  geom_boxplot(aes(group = factor(channel))) +
+  geom_boxplot(aes(group = factor(channel)), color = "darkorange", size = 0.1) +
   scale_y_level(name = "Presión (hPa)") +
@@
   ggplot(aes(channel, peakwt)) +
-  geom_boxplot(aes(group = factor(channel))) +
-  scale_x_continuous(breaks = seq) +
+  geom_boxplot(aes(group = factor(channel)), color = "darkorange") +
+  scale_x_continuous() +
   scale_y_level(name = "Presión (hPa)") +
@@
 Si observamos la corrección de bias para cada canal de la familia de sensores AMSU-A se ven diferencias interesantes entre los dos periodos pre-RAD1 y pre-RAD2.
-Para analizar estas diferencias en la Figura \@ref(fig:bias-amsua) hay que tener en cuenta que no todas las observaciones cumplen las condiciones para ser asimiladas y por lo tanto el BIAS no es corregido o si es corregido los coeficientes no son actualizados en cada ciclo.
-En particular los canales 4 a 7 muestran una mejora en la mediana del bias (más cercana a cero) y/o en la dispersión (menor rango intercuartil y/o menos valores extremos).
+Para analizar estas diferencias en la Figura \@ref(fig:bias-amsua) hay que tener en cuenta que solo incluye las observaciones cumplen las condiciones para ser asimiladas y que este BIAS se obtiene de la diferencia entre la innovación con bias corregido y sin corregir.
+Por lo tanto este es el BIAS corregido por el sistema.
+Llama la atención que en muchos casos el BIAS durante el periodo pre-RAD2 tiene menor dispersión (menor rango intercuartil) y su mediana está más cerca de cero.
+Esto podría deberse a que a las observaciones que finalmente pasan el control de calidad requieren una menor corrección o que durante el periodo pre-RAD1 la corrección es sobreestimada.
 
 La evolución de bias a lo largo del tiempo para el sensor sobre NOAA15 (Figura \@ref(fig:bias-temp-amsua)) muestra cambios entre los dos periodos en algunos casos del orden de 0.5 grados.
-En general el comportamiento del bias en el periodo pre-RAD2 mejora respecto del otro periodo y en los casos donde esto no ocurre as diferencias son muy pequeñas.
+En general para los canales 7 y 8 los valores de bias en el periodo pre-RAD2 son algo menores respecto del otro periodo y en los casos donde esto no ocurre as diferencias son muy pequeñas.
+Esto coincide con lo que se observa en el panel superior derecho de la Figura \@ref(fig:bias-temp-amsua).
 
@@
   scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
-  labs(x = "Canal", y = "BIAS (K)",
+  labs(x = "Canal", y = "BIAS corregido (K)",
        color = "") +
@@
 
-```{r bias-temp-amsua, fig.cap="Bias medio de las radianzas de AMSU-A NOAA 15 en las distintas plataformas y para los dos periodos a lo largo del tiempo.", fig.height=4}
+```{r bias-temp-amsua, fig.cap="Bias medio de las radianzas de AMSU-A NOAA 15 en las distintas plataformas y para los dos periodos a lo largo del tiempo.
+El tamaño de los puntos indica la cantidad de observaciones disponibles en cada momento.", fig.height=3, fig.width=6, fig.align="center"}
 
-diag_rad[sensor %in% sensores[1] & qc == 0 & channel %in% c(7:8)] %>% 
+diag_rad[sensor %in% sensores[1] & qc == 0 & channel %in% c(6:8)] %>% 
   .[, .(mean_bias = mean(tbc - tbcnob, na.rm = TRUE),
@@
   scale_size_area(max_size = 4) +
-  labs(x = "Canal", y = "BIAS (K)",
+  labs(x = "Canal", y = "BIAS corregido (K)",
        color = "", size = "") +
@@
 
-```{r bias-pred-ansua, fig.cap="Evolución temporal del bias asociado a cada predictor para los sensores AMSU-A en las distintas plataformas", fig.height=3}
+```{r bias-pred-amsua, fig.cap="Evolución temporal del bias asociado a cada predictor para los sensores AMSU-A en las distintas plataformas"}
 
@@
 En este caso el TLR^2 muestra valores comparables con la literatura y en particular un máximo que ocupa gran parte del dominio.
-El predictor asociado a la emisividad presenta valores distinto de cero solo en la zona de topografía.
+El predictor asociado a la emisividad presenta valores distinto de cero solo en la zona de topografía donde el nivel de presión de las observaciones está cerca de superficie.
 
-```{r pred-amsua, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 8 de AMSU-A NOAA19 a las 8 UTC del 17/11/2018"}
+```{r pred-amsua, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 8 de AMSU-A NOAA19 a las 8 UTC del 17/11/2018.
+Se incluyen observaciones que no pasaron el control de calidad."}
 
@@
 La corrección del bias para el sensor MHS en cada periodo es menos clara que para AMSU-A (Figura \@ref(fig:bias-mhs)).
-Los cambios en pre-RAD2 son mínimos y en algunos casos negativos.
-Por ejemplo el canal 1 que tiene el máximo cerca de superficie es el que muestra mayor dispersión, esto mejora mucho para el sensor sobre METOP-A y B (aunque en este ultimo caso la mediana se aleja de cero) pero para NOAA19 la dispersión no cambia particularmente (pero la mediana está más cerca de cero).
+Los cambios en pre-RAD2 son muy pequeños y en algunos casos se observan valores muy grandes y mayores a pre-RAD1.
+Por ejemplo el canal 1 que tiene el máximo cerca de superficie es el que muestra mayor dispersión, esto disminuye mucho para el sensor sobre METOP-A y B (aunque en este ultimo caso la mediana se aleja de cero) pero para NOAA19 la dispersión no cambia particularmente (pero la mediana está más cerca de cero).
+Esto podría estar indicando que la corrección de bias con los coeficientes entrenados es mayor.
 
@@
   scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
-  labs(x = "Canal", y = "BIAS (K)",
+  labs(x = "Canal", y = "BIAS Corregido (K)",
        color = "") +
@@
 
-De la misma manera la evolución temporal del bias para este sensor es muy variable y el periodo pre-RAD2 no muestran mejoras marcadas (Figura \@ref(fig:bias-temmp-mhs).
-Los valores de bias aumentan y disminuyen alternativamente [TODO: podría deberse a la cantidad de observaciones?
+De la misma manera la evolución temporal del bias para este sensor es muy variable y el periodo pre-RAD2 no muestran cambios marcados (Figura \@ref(fig:bias-temmp-mhs).
+Los valores de bias corregido aumentan y disminuyen alternativamente [TODO: podría deberse a la cantidad de observaciones?
 eso parece cumplirse en el canal 4 pero se ve lo contrario en el canal 2].
 
-```{r bias-temmp-mhs, fig.cap="Bias medio de las radianzas de MHS para los dos periodos a lo largo del tiempo.", fig.height=4}
+```{r bias-temmp-mhs, fig.cap="Bias medio de las radianzas de MHS para los dos periodos a lo largo del tiempo.El tamaño de los puntos indican la cantidad de observaciones en cada ciclo.", fig.height=4}
 
@@
 En cuanto a la contribución para la corrección del bias por parte de los distintos predictores en MHS (Figura \@ref(fig:bias-pred-mhs)) se ven mayores diferencias entre los dos periodos.
-Llama poderosamente la atención la magnitud de corrección del bias para el canal 1 para metop a y b donde llega a valores de 30 K.
-El canal 5 muestra un pico que llegan a 5 K mientras que el canal 4 muestra essos valores de manera sostenida para todo el periodo pre-RAD2.
+El canal 1 tiene un pico de 5 K al comienzo del periodo y los canales 4 y 5  tienen algunos valores cercanos a 3 K sobre todo en el periodo pre-RAD2.
 Para el resto de los canales los valores son similares a lo observado para AMSU-A (Figura \@ref(fig:bias-pred-amsua)).
-Por la escala de las figuras no se observas demasiados detalles respecto de que predictor es el que más contribuye al bias total pero por ejemplo se puede ver que en el canal 3 y para todos los sensores el Offset parece ser el que tiene un mayor aporte.
+El predictor que más contribuye al bias total es distintos para cada canal, por ejemplo en el canal 3 el más importante es el Offset mientras que para el canal 4 es el TLR^2.
 
@@
 
-Para estos sensores en particular llama la atención las diferencias entre los dos periodos, por ejemplo en el canal 4 de METOP-B hay un cambio de magnitud del bias que ocurre alrededor de 16 de noviembre y que se mantiene luego en pre-RAD2.
-Para explorar la posibilidad de que este cambio se deba realmente al entrenamiento de los coeficiente y que a partir de esos momentos la corrección del bias esté mejorando, en la Figura \@ref(fig:mhs-bias-res) se muestra la media de la innovación.
-En primer lugar se ven las diferencias bastante marcadas entre la media de la innovación con y sin bias corregido y que en la mayoría de los casos el primer es más cercano a cero.
+Para estos sensores en particular llama la atención las diferencias entre los dos periodos, por ejemplo en el canal 4 de METOP-B hay un cambio de magnitud del bias que ocurre alrededor de 16 de noviembre y que se mantiene luego en pre-RAD2 en forma de picos.
+Para explorar la posibilidad de que este cambio se deba realmente al entrenamiento de los coeficiente y que a partir de esos momentos la corrección del bias esté mejorando, en la Figura \@ref(fig:mhs-bias-res) se muestra la media de la innovación con bias corregido (línea llena) y bias sin corregir (línea punteada).
+Se espera que la distribución de la innovación sea aproximadamente gausiana y por lo tanto que la media esté cerca de cero.
+Una desvíación de cero podría indicar que queda un bias *residual* luego de la corrección.
+
+En la figura se ven las diferencias bastante marcadas entre la media de la innovación con y sin bias corregido y que en la mayoría de los casos el primer es más cercano a cero, indicando que la corrección del bias va en la dirección correcta.
 En particular para el canal 4 de METOP-B se obsera que la media de la innovación corregida luego de 16 de noviembre se aleja de cero y esto se mantiene también durante el periodo pre-RAD2.
+Si bien no es posible identificar ninguna causa esto podía indicar que el coeficiente asociado al TLR^2 cambió afectando negativamente a la corrección del bias y esto se mantuvo a lo largo del tiempo.
 
@@
 
-```{r pred-mhs, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 3 de MHS NOAA19 a las 8 UTC del 17/11/2018"}
+```{r pred-mhs, fig.cap="Contribución a la corrección del bias de dos predictores para el canal 3 de MHS NOAA19 a las 8 UTC del 17/11/2018.
+Incluye todas las observaciones de la pasada, aún si no pasan el control de calidad."}
 diag_rad[sensor == "mhs_n19" & date == as_datetime("2018-11-17 08:00:00") & channel == 3] %>% 
@@
 
+A continuación se analizan algunos casos de estudio desde el punto de vista de la precipitación acumulada en una hora comparando el análisis con la precipitación estimada por IMERG.
+
 #### 2018-11-13 01:00Z
+
+El primer caso corresponder a las 01 UTC del 13 de noviembre y corresponde al 8vo ciclo de asimilación.
 
@@
 ```
+
+En la Figura \@ref(fig:caso1) se muestra arriba la precipitación acumulada para la media del ensamble según cada experimento, abajo a la izquierda la diferencia entre los dos campos y a la derecha la precipitación acumulada según IMERG.
+En primer lugar es importante notar que los acumulados son mucho menores en los experimentos pero al mismo tiempo hay que tener en cuenta que en la media de los ensables se pierden las características de los miembros con distintas parametrizaciones.
+En particular algunas parametrizaciones muestran sistemáticamente menor convección afectando la precipitación acumulada.
+
+Las diferencias entre los dos periodos se observan principalmente en toda la franja de precipitación con una disminución en el experimento pre-RAD2 sobre todo en Entre Rios, sur de Uruguay y Rio de la Plata, donde ambos experimentos subestiman la precipitación pero la sobreestiman sobre Santiago del Estero donde no se registró lluvia.
+Esta sobreestimación disminuye en pre-RAD2 y al mismo tiempo corrige la precipitación sobre Córdoba.
+Sobre Jujuy y Salta pre-RAD2 muestra mayor precipitación pero ubicada al oeste de las observaciones.
 
@@
 
+En cuanto a la probabilidad de precipitación acumulada mayor a 5 mm (Figura \@ref(fig:caso1-prob)) las diferencias son importantes en la zona del sur de Santa Fe, Entre Rios y Buenos Aires principalmente.
+En pre-RAD1 la precipitación acumulda parece mucho más extendida en el dominio pero con mayor dispersión entre los miembros del ensable.
+En pre-RAD2 aparecen valores de probabilidad
+de entre 30 y 50% sobre la provincia de Córodba y sobre Salta en la región donde se observa precipitación acumulada.
+Al mismo tiempo sobre Buenos Aires se observa una probabilidad de entre 10 y 20% al sur y mayor al 30% al este que no se observan en la realidad.
+
 ```{r caso1-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento en el caso 1.", fig.height=7}
@@
   theme_minimal() +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom")
 
-  pp[, .(prop = mean(precipitacion > 25)*100), by = .(exp, x, y)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = prop), proj = norargentina_lambert,
-                    breaks = seq(0, 100, 10)) +
-  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
-                       limits = c(10, 100),
-                       na.value = "white",
-                       breaks = seq(0, 100, 10),
-                       guide = guide_colorstrip(barwidth = 15,
-                                                barheight = 1)) +
-  geom_mapa() +
-  labs(fill = "", x = "", y = "") +
-  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
-  ggtitle("Probabilidad de precipitación acumulada > 25 mm") +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-  plot_layout(ncol = 1, guides = "collect", widths = 0.5, heights = 2) &
-  theme(legend.position = "bottom") 
+  # pp[, .(prop = mean(precipitacion > 25)*100), by = .(exp, x, y)] %>% 
+  # ggplot(aes(x, y)) +
+  # geom_contour_fill(aes(z = prop), proj = norargentina_lambert,
+  #                   breaks = seq(0, 100, 10)) +
+  # scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
+  #                      limits = c(10, 100),
+  #                      na.value = "white",
+  #                      breaks = seq(0, 100, 10),
+  #                      guide = guide_colorstrip(barwidth = 15,
+  #                                               barheight = 1)) +
+  # geom_mapa() +
+  # labs(fill = "", x = "", y = "") +
+  # facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  # ggtitle("Probabilidad de precipitación acumulada > 25 mm") +
+  # theme_minimal() +
+  # theme(legend.position = "bottom") +
+  # plot_layout(ncol = 1, guides = "collect", widths = 0.5, heights = 2) &
+  # theme(legend.position = "bottom") 
 ```
@@
 #### 2018-11-18 10:00Z
+
+El segundo caso de estudio corresponde a las 10 UTC del 18 de noviembre casi 7 días luego del comienzo de cada periodo de asimilación continua.
+En este caso la precipitación se da sobre el norte de argentina y sur de Paraguay y Brasil, con algo de precipitación sobre Salta.
 
@@
 ```
+
+En el campo de diferencia en la Figura \@ref(fig:caso2) se observa que en pre-RAD2 disminuye la precipitación acumulada sobre Salta, Jujuy y Tucumán (esta última no se observa en IMERG) mientras que aumenta sobre Formosa y Chaco donde se dan los mayores acumulados de precipitación.
+De nuevo, los paneles superiores muestran la media del ensable por lo que muchas características de la ubicación de la precipitación acumulada se pierden y hay mucha variabilidad entre los miembros.
 
@@
 
+Esto puede verse en la Figura \@ref(fig:caso2-prob) que muestra la probabilidad de precipitación acumulada mayor a 5 mm.
+En pre-RAD2 se ve mayor probabilidad de precipitación por encima de 50% extendida sobre gran parte de la provincia de Formosa mientras que la precipitación observada según IMERG se localiza más hacia el este.
+La precipitación que se observa sobre Misiones en los experimentos no debe ser considerada con mucha confianza ya que se encuentran en el borde del diminio de simulación.
+
 ```{r caso2-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento correspondiente al caso 2", fig.height=7}

### 2020-10-22 analysis/chapters/coeficientes.Rmd
Algunos arreglos en las figuras de coeficientes

@@
   ggplot(aes(channel, N)) +
-  geom_boxplot(aes(group = factor(channel)), color = "darkorange", size = 0.1) +
+  geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
+               outlier.size = 0.5, size = 0.1) +
   facet_wrap(~sensor, scales = "free") +
@@
   ggplot(aes(channel, N)) +
-  geom_boxplot(aes(group = factor(channel)), color = "darkorange") +
+  geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange") +
   facet_wrap(~sensor, scales = "free") +
@@
   ggplot(aes(channel, peakwt)) +
-  geom_boxplot(aes(group = factor(channel)), color = "darkorange", size = 0.1) +
+   geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
+               outlier.size = 0.5, size = 0.1) +
   scale_y_level(name = "Presión (hPa)") +
@@
   ggplot(aes(channel, peakwt)) +
-  geom_boxplot(aes(group = factor(channel)), color = "darkorange") +
+  geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange") +
   scale_x_continuous() +
@@
 
-diag_rad[sensor %in% sensores[1] & qc == 0 & channel %in% c(6:8)] %>% 
+diag_rad[sensor %in% sensores[1] & qc == 0 & channel %in% c(7:8)] %>% 
   .[, .(mean_bias = mean(tbc - tbcnob, na.rm = TRUE),
@@
 
-```{r caso1-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento en el caso 1.", fig.height=7}
+```{r caso1-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento en el caso 1.", fig.height=3}
 pp[, .(prop = mean(precipitacion > 5)*100), by = .(exp, x, y)] %>% 
@@
 
-```{r caso2-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento correspondiente al caso 2", fig.height=7}
+```{r caso2-prob, fig.cap="Probabilidad de precipitación acumulada > 5 mm (arriba) y > 25 mm (abajo) para cada experimento correspondiente al caso 2", fig.height=3}
 pp[, .(prop = mean(precipitacion > 5)*100), by = .(exp, x, y)] %>% 

### 2020-10-22 analysis/chapters/experimento_20181122.Rmd
Inicio capitulo de experimentos

---
title: "Asimilación de distintas observaciones"
output: 
  bookdown::html_document2: default
  bookdown::pdf_document2: default
bibliography: mesoda_ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path = "../figures/",
  fig.retina = TRUE,
  # fig.height = 3,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dpi = 300
)
library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

square <- fread(here("analysis", "data", "derived_data", "dominio_square.csv"))
square2 <- fread(here("analysis", "data", "derived_data", "dominio_square2.csv"))
coord <- fread(here("analysis", "data", "derived_data", "coordenadas.csv"))

geom_mapa <- function() {
  list(geom_sf(data = map, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE),
       coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)))
}

fisica <- data.table(ens = 1:60, 
                     fisica = rep(c("KF-YSU", 
                                    "BMJ-YSU",
                                    "GF-YSU",
                                    "KF-MYJ",
                                    "BMJ-MYJ",
                                    "GF-MYJ",
                                    "KF-MYNN2",
                                    "BMJ-MYNN2",
                                    "GF-MYNN2"), length.out = 60)) %>% setDT()
derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"

```

## Case Study: November 22th, 2020

The experiments focus on a mesoscale convective system developed over central and north-eastern Argentina during November 22th, 2018. This MCS corresponds to IOP08 (Severe, Upscale Growth handoff) of the RELAMPAGO field campaign.

This case stady was selected to focus on the RELAMPAGO period and to take advantage of the observations taken during the intensive observation periods.

```{r caso-gif, fig.cap="GOES-16 channel 13 (clean longwave-IR) during November, 22th.", out.height=450, out.width=600, fig.align="center"}
knitr::include_graphics("https://github.com/paocorrales/Analisis_2018112022/blob/master/fig/caso_20181122.gif?raw=true")
```

### Model and ensemble configuration

To run the simulations the WRF-ARW (V3.9) model with 10 km horizontal resolution and xx vertical levels was used. The GFS deterministic run (0.5° resolution) was used for the initial and boundary conditions.

The LETKF versión of the GSI system was choosen for the assimilation. A 60 member ensemble is initialized using random perturbations and a multi-physics squeme using 9 combinations of 3 Cumulus (Kain–Fritsch, Grell–Freitas, Betts–Miller–Janjic) and 3 PBL (Yonsei University Scheme, Mellor–Yamada–Janjic Scheme, Mellor–Yamada Nakanishi Niino) parameterizations as shown in Table \@ref(tab:fisicas).

```{r fisicas}
fisica[, .(mem = paste(ens, collapse = ", ")), by = .(fisica)] %>% 
  kbl(col.names = c("Parameterizations", "# Member"),
      caption = "Posible Cumulus and PBL combinations and the members in each group.") %>% 
  kable_classic_2(full_width = F)
```

The domain shown in Figure \@ref(fig:dominio) covers the center and north of Argentina and part of Bolivia and Paraguay to capture the development of the MCS during the simulation period (up to 12UTC 2020/11/23).

```{r dominio, fig.cap="Horizontal domain for all simulations and topography.", fig.height=4}
fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>%

  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_distiller(name = NULL,
                       type = "seq",
                       palette = "RdYlGn",
                       direction = -1,
                       breaks = seq(0, 6000, 1000),
                       guide = guide_colorstrip(barwidth = 20,
                                                barheight = 0.8)) +
  geom_mapa() +
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  geom_point(data = square2, aes(lon, lat), size = 0.2, color = "darkgrey") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    x = "Longitude", 
    y = "Latitude") +
  theme_minimal() +
  theme(legend.position = "bottom") 
```

### Experiments configuration

#### Assimilation cycle

Each simulation starts at 12 UTC on Nov 20th with 6 hours of spin up. The assimilation starts at 18 UTC and continue until 12 UTC Nov, 23th every houer. Two ensemble forecasts are initialized from de the analysis at 00 and 06 UTC Nov 22th and continue until 12 UTC Nov 23rd (Figure \@ref(fig:ciclo)).

The analysis is perform in one-hour cetered window using the Observation Operator from GSI to compare the available observations with the first guess at the closest time (FGAT - First Guess at Appropriate Time). Then the LETKF is run to apply the innovations to the analysis for each member.

```{r ciclo, fig.cap="Diagram of analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours. The zoomed figure shows assimilation in one-hour centered window and new boundary conditions from GFSD every 6 hours.", out.height=400, out.width=800, fig.align="center"}
knitr::include_graphics(here("analysis", "figures", "analysis_cycle.svg"))
```
Four experiments were performed using an incremental set of observations each time (Table \@ref(tab:obs)). The CONV experiment uses conventional observations from the prepBUFR file that includes surface observations (land and sea) and profiles from soundings and aircrafts. On the Figure \@ref(fig:mapa-obs), the orange dots are the surface stations and ships locations included in this experiment.

The AUT experiment add to CONV the regional automatic stations networks from Argentina and near countries. The added stations are colored in green in Figure \@ref(fig:mapa-obs). This stations have greater spatial coverage and higher sampling frequency.

SATWND add to AUT the satellite motion vectors from several satellites including GOES-16 and finally the RAD experiments also includes randiances observations from sensors over polar orbit satellites.

```{r obs, eval=FALSE, include=FALSE}
tribble(
  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
  "ADPSFC", "✔", "✔", "✔", "✔",
  "ADPUPA", "✔", "✔", "✔", "✔",
  "AIRCFT", "✔", "✔", "✔", "✔",
  "AIRCAR", "✔", "✔", "✔", "✔",
  "SFCSHP", "✔", "✔", "✔", "✔",
  "ASCATW", "✔", "✔", "✔", "✔", 
  "AUTSFC", " ", "✔", "✔", "✔",
  "SATWND", " ", " ", "✔", "✔",
  "RAD", " ", " ", " ", "✔",
) %>% 
  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic_2(full_width = F)
```

```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots).", fig.height=4}
obsAUT <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
                na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
  .[, c("V2", "V4") := NULL] %>% 
  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))

obsCONV <- fread(here("analysis", "data", "raw_data", "E4_asim_conv_20181121120000.ensmean"), 
                 na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
  .[, c("V2", "V4") := NULL] %>% 
  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))

oficiales <- obsCONV[type != 290, unique(stationID)]

no_oficiales <- unique(obsAUT, by = c("stationID")) %>% 
  .[type != 290] %>% 
  .[!str_detect(stationID, "SMN")] %>%
  .[!stationID %in% oficiales & type != 290, stationID]

unique(obsAUT, by = c("stationID")) %>% 
  .[type != 290] %>% 
  .[!str_detect(stationID, "SMN")] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_mapa() +
  geom_point(alpha = 0.8, size = 2,
             aes(color = stationID %in% oficiales)) +
  scale_color_manual(name =  NULL, values = c("TRUE" = "#FD8002", "FALSE" = "#00695c"),
                     breaks = c("FALSE", "TRUE"),
                     labels = c("Non-official (N = 820)","Official (N = 327)"))+
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```
### Analysis comparison

To observe the impact of assimilating each source of observations on the vertical structure, the average vertical profiles for  the temperature, specific humidity and wind components were calculated over the grey box in Figure \@ref(fig:dominio).

```{r T-diff, fig.cap="Temperature difference between experiments in Kelvin.", fig.height=3}
files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
                    pattern = "perfiles_ana")

perfiles <- purrr::map(files, function(f) {
  exp <- unglue(basename(f), "perfiles_{run}_{exp}.csv")
  fread(f) %>% 
    .[, exp := exp[[1]][["exp"]]]

}) %>% 
  rbindlist() %>% 
   .[, date := ymd_hms(date)] %>% 
  .[date %between% c(as_datetime("2018-11-20 18:00:00"),
                     as_datetime("2018-11-23 12:00:00"))]

perfiles %>% 
  dcast(lev + date ~ exp, value.var = "T") %>% 
  .[, ":="(E5_E4 = E5 - E4,
           E6_E5 = E6 - E5,
           E7_E6 = E7 - E6)] %>% 
  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = value), breaks = seq(-1, 1, 0.2)) +
  scale_fill_divergent(name = NULL,
                       breaks = seq(-1, 1, 0.2),
                       guide = guide_colorstrip(inside = TRUE,
                                                barwidth = 25,
                                                barheight = 0.8)) +
  labs(subtitle = "Temperature difference (K)",
       x = NULL,
       y = "Pressure level") +
  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                         "E6_E5" = "SATWND - AUT",
                                                         "E7_E6" = "RAD - SATWND"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))
```

```{r Q-diff, fig.cap="Specific humidity difference between experiments in g/Kg", fig.height=3}
perfiles %>% 
  dcast(lev + date ~ exp, value.var = "QVAPOR") %>% 
  .[, ":="(E5_E4 = E5 - E4,
           E6_E5 = E6 - E5,
           E7_E6 = E7 - E6)] %>% 
  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
  ggplot(aes(date, lev)) +
    geom_contour_fill(aes(z = value*1000), breaks = seq(-0.0016, 0.0016, 0.0002)*1000) +
  scale_fill_divergent(name = NULL,
                       breaks = seq(-0.0016, 0.0016, 0.0002)*1000,
                       guide = guide_colorstrip(inside = TRUE,
                                                barwidth = 25,
                                                barheight = 0.8)) +

  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
  labs(subtitle = "Specific humidity difference (g/Kg)",
       y = "Pressure level",
       x = NULL) +
  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                         "E6_E5" = "SATWND - AUT",
                                                         "E7_E6" = "RAD - SATWND"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))
```

```{r U-diff, fig.cap="Zonal wind difference between experiments in m/s", fig.height=3}
perfiles %>% 
  dcast(lev + date ~ exp, value.var = "U") %>% 
  .[, ":="(E5_E4 = E5 - E4,
           E6_E5 = E6 - E5,
           E7_E6 = E7 - E6)] %>% 
  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = value), reaks = seq(-1.2, 1.2, 0.2)) +
  scale_fill_divergent(name = NULL,
                       breaks = seq(-1.2, 1.2, 0.2),
                       labels = round(seq(-1.2, 1.2, 0.2), 1),
                       guide = guide_colorstrip(inside = TRUE,
                                                barwidth = 25,
                                                barheight = 0.8)) +
  labs(subtitle = "U difference (m/s)",
       x = NULL,
       y = "Pressure level") +
  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
   facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                         "E6_E5" = "SATWND - AUT",
                                                         "E7_E6" = "RAD - SATWND"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))
```

```{r V-diff, fig.cap="Meridial wind difference between experiments in m/s", fig.height=3}
perfiles %>% 
  dcast(lev + date ~ exp, value.var = "V") %>% 
  .[, ":="(E5_E4 = E5 - E4,
           E6_E5 = E6 - E5,
           E7_E6 = E7 - E6)] %>% 
  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
  ggplot(aes(date, lev)) +
  geom_contour_fill(aes(z = value), reaks = seq(-1.2, 1.2, 0.2)) +
  scale_fill_divergent(name = NULL,
                       breaks = seq(-1.2, 1.2, 0.2),
                       labels = round(seq(-1.2, 1.2, 0.2), 1),
                       guide = guide_colorstrip(inside = TRUE,
                                                barwidth = 25,
                                                barheight = 0.8)) +
  labs(subtitle = "V difference (m/s)",
       x = NULL,
       y = "Pressure level") +
  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
   facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                         "E6_E5" = "SATWND - AUT",
                                                         "E7_E6" = "RAD - SATWND"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

```

```{r q-wind, warning=FALSE, fig.align="center", fig.height=4, fig.width=6, message=FALSE,}
fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
                     exp = c("E4", "E5", "E6", "E7"))

q <- lapply(seq_len(nrow(fcsts)), function(f) {

  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
    .[, exp := fcsts[f, 2]] %>%  
    .[]
}) %>% 
  rbindlist()

q[bottom_top <= 7, .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
  by = .(south_north, west_east, date, exp)] %>% 
  .[coord, on = c("south_north", "west_east")] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = q*1000)) +
  scale_color_distiller(name = NULL,
                        palette = "BrBG", direction = 1,
                        breaks = seq(0, 0.02, 0.002)*1000,
                        guide = guide_colorstrip(inside = FALSE,
                                                 barwidth = 30,
                                                 barheight = 0.8)) +
  geom_vector(aes(dx = u, dy = v), data = function(x) x[is.cross(south_north, west_east, 5)]) +
  scale_mag() +
  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E7 = "RAD"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Low level humidity (g/Kg)",
       x = "Longituted",
       y = "Latitude") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))
```

### 2020-10-23 analysis/chapters/coeficientes.Rmd
Nueva versión post reunion 23/10/2020

@@
 
+```
+
+```{r}
+diag_rad %>% 
+  .[qc == 0 & channel %in% c(5:8)] %>% 
+  melt(measure.vars = c("tbc", "tbcnob")) %>% 
+  .[, .(ob_mean = mean(value, na.rm = TRUE),
+       ob_sd = sd(value, na.rm = TRUE)), by = .(date, sensor, channel, exp, variable)] %>% 
+  # melt(measure.vars = c("ob_mean", "ob_sd"),  variable.name = "estadistico", 
+  #      value.name = "valor") %>% 
+ggplot(aes(date, ob_mean)) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(color = exp)) +
+  geom_line(aes(color = exp, linetype = variable)) +
+  scale_color_manual(values = c("darkorange", cyan), labels = c("pre-RAD1", "pre-RAD2")) +
+  scale_linetype_discrete(labels = c("Bias corregido", "Sin corrección")) +
+  facet_grid(channel~sensor, scales = "free_y") +
+  labs(y = NULL, x = NULL, color = NULL, linetype = NULL) +
+  theme_minimal()
 ```

### 2020-10-23 analysis/chapters/experimento_20181122.Rmd
Arreglo viento

@@
 
-```{r q-wind, warning=FALSE, fig.align="center", fig.height=4, fig.width=6, message=FALSE,}
+```{r q-wind, warning=FALSE, fig.align="center", fig.height=5, fig.width=9, message=FALSE,dev="png"}
 fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
@@
   .[, date := factor(date)] %>% 
-  ggplot(aes(lon, lat)) +
-  geom_point(aes(color = q*1000)) +
-  scale_color_distiller(name = NULL,
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = q*1000), proj = norargentina_lambert,
+                    breaks = seq(0, 0.02, 0.002)*1000) +
+  scale_fill_distiller(name = NULL,
                         palette = "BrBG", direction = 1,
@@
                         guide = guide_colorstrip(inside = FALSE,
-                                                 barwidth = 30,
+                                                 barwidth = 25,
                                                  barheight = 0.8)) +
-  geom_vector(aes(dx = u, dy = v), data = function(x) x[is.cross(south_north, west_east, 5)]) +
-  scale_mag() +
+  # geom_vector(aes(x = lon, y = lat,
+  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
+  #             arrow.angle = 12, size = 0.3) +
+  # scale_mag() +
   geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
@@
   labs(title = "Low level humidity (g/Kg)",
-       x = "Longituted",
-       y = "Latitude") +
-  theme_minimal(base_size = 16) +
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
   theme(legend.position = "bottom",

### 2020-10-27 R/read_diag_rad.R
na string para diag files

@@ -14,7 +14,7 @@ read_diag_rad <- function(file_list, exp) {
     }
     meta <- unglue::unglue(basename(file_list[f]), "asim_{sensor}_{plat}_{date}.{mem}")
     # print(f)
-    out <- data.table::fread(file_list[f])
+    out <- data.table::fread(file_list[f], na.strings = c("0.100000E+12"))
     # .[V10 == 1] %>%
 
     if (file.size(file_list[f]) != 0) {

### 2020-10-27 R/symlog_trans.R
Agrego symlog_trans sacado de https://stackoverflow.com/a/14674703

@@ -1,0 +1,38 @@
+#' Transformación de escalas a symlog
+#'
+#' Permite transformar una escala (colores, ejes, etc.) a symlog.
+#'
+#' @references Un buen samaritano en https://stackoverflow.com/a/14674703
+#'
+#' @export
+symlog_trans <- function(base = 10, thr = 1, scale = 1){
+  trans <- function(x)
+    ifelse(abs(x) < thr, x, sign(x) *
+             (thr + scale * suppressWarnings(log(sign(x) * x / thr, base))))
+
+  inv <- function(x)
+    ifelse(abs(x) < thr, x, sign(x) *
+             base^((sign(x) * x - thr) / scale) * thr)
+
+  breaks <- function(x){
+    sgn <- sign(x[which.max(abs(x))])
+    if(all(abs(x) < thr))
+      scales::pretty_breaks()(x)
+    else if(prod(x) >= 0){
+      if(min(abs(x)) < thr)
+        sgn * unique(c(scales::pretty_breaks()(c(min(abs(x)), thr)),
+                       scales::log_breaks(base)(c(max(abs(x)), thr))))
+      else
+        sgn * log_breaks(base)(sgn * x)
+    } else {
+      if(min(abs(x)) < thr)
+        unique(c(sgn * log_breaks()(c(max(abs(x)), thr)),
+                 pretty_breaks()(c(sgn * thr, x[which.min(abs(x))]))))
+      else
+        unique(c(-log_breaks(base)(c(thr, -x[1])),
+                 pretty_breaks()(c(-thr, thr)),
+                 log_breaks(base)(c(thr, x[2]))))
+    }
+  }
+  trans_new(paste("symlog", thr, base, scale, sep = "-"), trans, inv, breaks)
+}

### 2020-10-27 analysis/chapters/coeficientes.Rmd
Intento de resumen gráfico, no se ve lindo para la semana 2

@@
 )
+
+multiespectrales <- c("airs_aqua",
+              "iasi_metop-a",
+              "iasi_metop-b")
 cyan <- rgb(0, 139, 139, maxColorValue = 255)
@@
 
+#### Resumen para todos los sensores
+
+```{r resume-bias, fig.cap= "Diferencias absoluta entre ambos experimentos para bias residual (color, igual en ambos paneles) y cantidad de observaciones que superan el control de calidad (tamaño de los puntos).
+En circulos negros se indentifican los sesnores y canales donde se asimilaron más observaciones comparando experimentos."}
+files <- list.files("/home/paola.corrales/datosmunin/EXP/E7-training/ANA", 
+                    pattern = "asim.", recursive = TRUE, 
+                    full.names = TRUE)
+files <- files[!str_detect(files, pattern = "conv")]
+
+diag_rad <- purrr::map(files, function(f) {
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/E7-training/ANA/{exp}/{date_ana}/diagfiles/{file}" )
+  read_diag_rad(f, exp = meta[[1]][["exp"]]) %>%
+    .[]
+
+}) %>%
+  rbindlist()
+
+n_obs <- diag_rad[qc == 0, .N, by = .(sensor, channel, exp)] %>% 
+  dcast(sensor + channel ~ exp) %>% 
+  .[, win := fifelse(week2 > week1, "week2", "week1")] %>% 
+  melt(measure.vars = c("week1", "week2"), variable.name = "exp", value.name = "N")
+
+abs_bias <- diag_rad %>% 
+  .[, pasa_qc := fifelse(qc == 0, "pasa qc", "no pasa qc")] %>% 
+  .[, .(bias_tbc = mean(tbc, na.rm = TRUE),
+        bias_tbcnob = mean(tbcnob, na.rm = TRUE)), 
+    by = .(sensor, channel, exp, pasa_qc)] %>% 
+  .[, abs_bias := abs(bias_tbc) - abs(bias_tbcnob)] %>% 
+  dcast(sensor + channel + exp ~ pasa_qc, value.var = "abs_bias") %>% 
+  setnames(c("no pasa qc", "pasa qc"), c("abs_bias_nqc", "abs_bias_qc"))
+
+total <- diag_rad %>% 
+  .[, .(bias_tbc = mean(tbc, na.rm = TRUE),
+        bias_tbcnob = mean(tbcnob, na.rm = TRUE)), 
+    by = .(sensor, channel, exp)] %>% 
+  dcast(sensor + channel ~ exp, value.var = "bias_tbc") %>% 
+  .[, abs_bias := abs(week2) - abs(week1)] %>% 
+  melt(measure.vars = c("week1", "week2"), value.name = "bias_tbc", variable.name = "exp") %>% 
+  # .[abs_bias, on = c("sensor", "channel", "exp")] %>% 
+  .[n_obs, on = c("sensor", "channel", "exp")]
+
+total %>% 
+  .[!(sensor %in% c(multiespectrales, "atms_npp"))] %>% 
+ # dcast(sensor + channel + N ~ exp, value.var = "bias_tbc")
+  ggplot(aes(channel, sensor)) +
+  geom_point(aes(color = abs_bias, size = N)) +
+  geom_point(data = ~.x[win == exp], shape = 21, aes(size = N)) +
+  # geom_text(aes(label = round(abs_bias, 1)), size = 2) +
+  # scale_color_viridis_c(trans = scales::log_trans(), breaks = c(0.1, 1, 10, 100),
+  #                       guide = guide_colourbar(barwidth = 7)) +
+  scale_color_divergent(trans = symlog_trans(), breaks = c(-1, 0, 1, 10),
+                        guide = guide_colourbar(barwidth = 7)) +
+  facet_wrap(~exp, labeller = labeller(exp = c("week1" = "pre-RAD1", "week2" = "pre-RAD2"))) +
+  labs(subtitle = "all obs",
+       color = "Bias\ndifference",
+       size = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
 ### Algunos casos de estudio

### 2020-10-29 R/symlog_trans.R
Fix symlog_trans

@@ -23,16 +23,16 @@ symlog_trans <- function(base = 10, thr = 1, scale = 1){
         sgn * unique(c(scales::pretty_breaks()(c(min(abs(x)), thr)),
                        scales::log_breaks(base)(c(max(abs(x)), thr))))
       else
-        sgn * log_breaks(base)(sgn * x)
+        sgn * scales::log_breaks(base)(sgn * x)
     } else {
       if(min(abs(x)) < thr)
-        unique(c(sgn * log_breaks()(c(max(abs(x)), thr)),
-                 pretty_breaks()(c(sgn * thr, x[which.min(abs(x))]))))
+        unique(c(sgn * scales::log_breaks()(c(max(abs(x)), thr)),
+                 scales::pretty_breaks()(c(sgn * thr, x[which.min(abs(x))]))))
       else
-        unique(c(-log_breaks(base)(c(thr, -x[1])),
-                 pretty_breaks()(c(-thr, thr)),
-                 log_breaks(base)(c(thr, x[2]))))
+        unique(c(-scales::log_breaks(base)(c(thr, -x[1])),
+                 scales::pretty_breaks()(c(-thr, thr)),
+                 scales::log_breaks(base)(c(thr, x[2]))))
     }
   }
-  trans_new(paste("symlog", thr, base, scale, sep = "-"), trans, inv, breaks)
+  scales::trans_new(paste("symlog", thr, base, scale, sep = "-"), trans, inv, breaks)
 }

### 2020-10-29 R/calculate_fss.R
Add FSS

@@ -1,0 +1,33 @@
+#' Wrap para FSS
+#'
+#' Usa la función fss del paquete {verification} pero previamente requiere que las
+#' variables estén en matrices. También puede iterar para distintos q (valor de pp) y
+#' w (tamaño de la caja = w2+1)
+#'
+#' @param fcst matriz con precipitación pronosticada
+#' @param obs matriz con precipitación observada (mismas dimensiones que fcst)
+#' @param q número o vector numérico con umbrales a ser calculados
+#' @param w número o vector numérico con escalas (w2+1)
+#' @param progress lógico, muestra o no el progreso
+#'
+#' @export
+calculate_fss <- function(fcst, obs, q, w, progress = TRUE) {
+  out <- purrr::map_dfr(q, function(q) {
+    fcst_q <- fcst >= q
+    obs_q <- obs >= q
+
+    return <- list(fss = purrr::map_dbl(w, ~ verification::fss(obs_q, fcst_q, .x)),
+                   w = w,
+                   q = rep(q, length(w)))
+
+    if (progress == TRUE) {
+
+      message(paste("Listo q = ", q))
+
+    }
+
+    return(return)
+  })
+
+  return(out)
+}

### 2020-10-29 analysis/chapters/coeficientes.Rmd
Grafico de resumen sin observaciones rechazadas

@@
 
+satinfo <- fread(here("analysis", "data", "raw_data", "satinfo.txt")) %>% 
+  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
+
 n_obs <- diag_rad[qc == 0, .N, by = .(sensor, channel, exp)] %>% 
@@
   # .[abs_bias, on = c("sensor", "channel", "exp")] %>% 
-  .[n_obs, on = c("sensor", "channel", "exp")]
+  .[n_obs, on = c("sensor", "channel", "exp")] %>% 
+  satinfo[., on = c("sensor", "channel")]
 
-total %>% 
+total[iuse == 1] %>% 
   .[!(sensor %in% c(multiespectrales, "atms_npp"))] %>% 

### 2020-11-14 R/calculate_fss.R
Calculate FSS

@@ -11,12 +11,15 @@
 #' @param progress lógico, muestra o no el progreso
 #'
 #' @export
-calculate_fss <- function(fcst, obs, q, w, progress = TRUE) {
-  out <- purrr::map_dfr(q, function(q) {
+calculate_fss <- function(fcst, obs, q, w, progress = TRUE, plan = "cluster", workers = 3) {
+
+  future::plan("cluster", workers = 3)
+
+  out <- furrr::future_map_dfr(q, function(q) {
     fcst_q <- fcst >= q
     obs_q <- obs >= q
 
-    return <- list(fss = purrr::map_dbl(w, ~ verification::fss(obs_q, fcst_q, .x)),
+    return <- list(fss = furrr::future_map_dbl(w, ~ verification::fss(obs_q, fcst_q, .x)),
                    w = w,
                    q = rep(q, length(w)))
 

### 2020-11-14 analysis/chapters/experimento_20181122.Rmd
Nuevo E7

@@
            E6_E5 = E6 - E5,
-           E7_E6 = E7 - E6)] %>% 
+           E7_E6 = E7 - E5)] %>% 
   melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
@@
                        breaks = seq(-1.2, 1.2, 0.2),
-                       labels = round(seq(-1.2, 1.2, 0.2), 1),
+                       # labels = round(seq(-1.2, 1.2, 0.2), 1),
                        guide = guide_colorstrip(inside = TRUE,
@@
 ```
+
+#### FSS for analysis
+
+```{r}
+files <- list.files(here("analysis", "data", "derived_data"), 
+                    full.names = TRUE,
+                    pattern = "fss_3h")
+
+fss <- purrr::map(files, function(f) {
+
+  fread(f) %>% 
+    .[, date := as_datetime(date)]
+}) %>% 
+  rbindlist()
+
+fss %>% 
+  ggplot(aes(date, fss)) +
+  geom_point(aes(color = exp)) +
+  geom_line(aes(color = exp, linetype = factor(w))) +
+  facet_grid(w~q)
+
+```

### 2020-11-14 temp_R/help_functions.R
FSS

@@ -342,26 +342,26 @@ cut_round <- function(x, breaks) {
 }
 
 
-# Wrap FSS ----------------------------------------------------------------
-
-# Usa la función fss del paquete verification pero previamente requiere que las
-# variables estén en matrices. También puede iterar para distintos q (valor de pp) y
-# w (tamaño de la caja = w2+1)
-
-FSS <- function(fcst, obs, q, w) {
-  out <- purrr::map_dfr(q, function(q) {
-    fcst_q <- fcst >= q
-    obs_q <- obs >= q
-
-    return <- list(fss = purrr::map_dbl(w, ~ verification::fss(obs_q, fcst_q, .x)),
-                   w = w,
-                   q = rep(q, length(w)))
-    message(paste("Listo q = ", q))
-    return(return)
-  })
-
-  return(out)
-}
+# # Wrap FSS ----------------------------------------------------------------
+#
+# # Usa la función fss del paquete verification pero previamente requiere que las
+# # variables estén en matrices. También puede iterar para distintos q (valor de pp) y
+# # w (tamaño de la caja = w2+1)
+#
+# FSS <- function(fcst, obs, q, w) {
+#   out <- purrr::map_dfr(q, function(q) {
+#     fcst_q <- fcst >= q
+#     obs_q <- obs >= q
+#
+#     return <- list(fss = purrr::map_dbl(w, ~ verification::fss(obs_q, fcst_q, .x)),
+#                    w = w,
+#                    q = rep(q, length(w)))
+#     message(paste("Listo q = ", q))
+#     return(return)
+#   })
+#
+#   return(out)
+# }
 
 # Parse radiosondes_RELAMPAGO ---------------------------------------------
 

### 2020-11-17 analysis/chapters/experimento_20181122.Rmd
Analisis de diferencia entre experimentos en perfiles verticales.

@@
 ```
+
+Figure with all type of observations (radiosonde, convencional, automatic station, aircraft, ships/buoys, satwind)
+Figure with number of observation throughout the cycles for each type including riadiances.
+
 ### Analysis comparison
@@
 
-```{r T-diff, fig.cap="Temperature difference between experiments in Kelvin.", fig.height=3}
+```{r T-diff, fig.cap="Temperature difference between experiments in Kelvin."}
 files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
@@
 
-```{r Q-diff, fig.cap="Specific humidity difference between experiments in g/Kg", fig.height=3}
+AUT-CONV present a negative difference in low levels with a clear diurnal cycle (Figure \@ref(fig:T-diff)).
+This negative difference indicate that AUT is warmer than CONV.
+It also shows significant differences in the upper levels of the troposphere during the hours of maximun convection.
+The temperature profiles of AUT y SATWND are very similar.
+The difference bewteen the two experiments is smaller than +-0.1 until Nov, 22th 00Z when the convection starts to develope on the domain.
+The differences between RAD y SATWIND (and AUT) are smaller important in low levels at the begining of the experiment and in upper levels at the end indicating that the satellite observations have impact in several levels.
+
+```{r Q-diff, fig.cap="Specific humidity difference between experiments in g/Kg"}
 perfiles %>% 
@@
 
-```{r U-diff, fig.cap="Zonal wind difference between experiments in m/s", fig.height=3}
+Comparing the specific humidity (Figure \@ref(fig:Q-diff)) once again the difference between AUT and CONV is important in low levels showing the impact of assimilating the observations from automatic stations with larger spacial and temporal resolution.
+Between SATWND and AUT the differences are smaller and only important after Nov, 21st 18Z.
+The addition of satellite derived winds produce and increse of humidity around 750 hPa during Nov 21st (0.05 to 0.10 g/kg) and a decrese of humidity between surface and 750 hPa starting at 00Z Nov 22th until the end of the experiment.
+
+The differences between RAD and SATWND and therefore the impact of the radiances have opposite sign to the impact of the observations from automatic stations but with smaller magnituded in low levels.
+A positive difference can also be observed between 850 and 600 hPa.
+
+```{r U-diff, fig.cap="Zonal wind difference between experiments in m/s (filled contours) and zonal wind profile for the experiment with more observations."}
 perfiles %>% 
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value), reaks = seq(-1.2, 1.2, 0.2)) +
+  geom_contour_fill(aes(z = value), breaks = seq(-1.2, 1.4, 0.2)) +
   scale_fill_divergent(name = NULL,
-                       breaks = seq(-1.2, 1.2, 0.2),
-                       # labels = round(seq(-1.2, 1.2, 0.2), 1),
+                       breaks = seq(-1.2, 1.4, 0.2),
+                       labels = function(x) round(x, digits = 1),
                        guide = guide_colorstrip(inside = TRUE,
@@
                                                 barheight = 0.8)) +
+  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7)] %>% 
+                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
+                breaks = seq(0, 30, 3), size = 0.3, linetype = 1, color = "grey30") +
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
+                      melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
+                    breaks = seq(0, 30, 3), color = "grey30", skip = 1, stroke = 0.2) +
   labs(subtitle = "U difference (m/s)",
@@
 
-```{r V-diff, fig.cap="Meridial wind difference between experiments in m/s", fig.height=3}
+In Figure \@ref(fig:U-diff) the difference between the experiments for the zonal wind is shown with the vertical profile for the experiment with more observations.
+For example the left panel shows the difference between AUT and CONV experiments and the profile wind for AUT.
+The wind profiles for the experiments are similar but there are important differences that can be seen on the filled contours.
+
+The differences between AUT and CONV are important in low levels, for the first two days the impact of assimilation observations from automatic stations is a disminution on the wind from the west but an increse of the wind from the west after Nov 22th, 18Z.
+In upper levels the sign of the impact varies throughout the experiment period.
+The assimilation of satellite derived winds produces an impact in midle and upper levels and the incorporation of radiances prodeces impacts on low and upper levels.
+In low levels the impact has the oposite sign comparing with AUT-CONV, this means that RAD experiment shows on average less westerly wind.
+
+```{r V-diff, fig.cap="Meridial wind difference between experiments in m/s"}
 perfiles %>% 
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value), reaks = seq(-1.2, 1.2, 0.2)) +
+  geom_contour_fill(aes(z = value), breaks = seq(-1.8, 1.2, 0.2)) +
   scale_fill_divergent(name = NULL,
-                       breaks = seq(-1.2, 1.2, 0.2),
-                       # labels = round(seq(-1.2, 1.2, 0.2), 1),
+                       breaks = seq(-1.8, 1.2, 0.2),
+                       labels = function(x) round(x, digits = 1),
                        guide = guide_colorstrip(inside = TRUE,
@@
                                                 barheight = 0.8)) +
+  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7)] %>% 
+                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
+                breaks = seq(-7, 0, 1), size = 0.3, linetype = 2, color = "grey30") +
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
+                      melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
+                    breaks = seq(-7, 0, 1), color = "grey30", skip = 1, stroke = 0.2) +
   labs(subtitle = "V difference (m/s)",
@@
 
+ggplot(data = perfiles[exp == "E5" & V < 0], aes(date, lev)) +
+  geom_contour2(aes(z = V))
+
 ```
+
+Figure \@ref(fig:V-diff) is constructed in the same way as Figure \@ref(fig:U-diff) but here only negative meridional wind is ploted.
+Again the wind profiles are similar for each experiment but nevertheless there are differences of +- 1 m/s.
+
+In upper levels and during Nov 22nd and 23rd the impact of assimilating new observations (mainly from automatic stations and radiances) produces an increse of northerly wind.
+In low levels the impact has different sign in AUT-CONV and RAD-SATWND.
 
@@
 
+```{r}
+
+```
+
 #### FSS for analysis

### 2020-11-17 analysis/scripts/wrap_fss.R
wrap para fss

@@ -1,0 +1,106 @@
+library(metR)
+library(tidyverse)
+library(data.table)
+library(lubridate)
+library(unglue)
+library(mesoda)
+
+
+# Constante
+future::plan("cluster", workers = 3)
+imerg_path <- "/home/paula.maldonado/datosalertar1/RRA_VERIF/data/raw/imerg_raw"
+wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
+exp <- "E4"
+
+ini_date <- ymd_hms("20181120180000")
+ciclos <- 68
+
+acumulado <- 3
+q <- c(1, 5, 10) #10mm, para arrancar pensanod en pp acumulada
+w <- c(2, 5, 25, 50) #ancho de cada box w*2+1
+
+# Inicio
+dates <- seq.POSIXt(ini_date, by = "hour", length.out = ciclos)
+
+fss_out <- furrr::future_map_dfr(dates, function(d) {
+
+  print(d)
+  date <- d
+
+  # Observaciones
+  files_imerg <- purrr::map(seq(acumulado), function(l) {
+    list.files(imerg_path, pattern = format(date - hours(l), "%Y%m%d-S%H"), full.names = TRUE)
+  }) %>% unlist()
+
+  pp_imerg <- purrr::map(files_imerg, function(f) {
+    meta <- unglue(basename(f), "3B-HHR-L.MS.MRG.3IMERG.{dia}-S{hora_ini}-E{hora_fin}.{algo}.V05B.RT-H5")
+    ReadNetCDF(f,
+               vars = c(pp = "Grid/precipitationCal"),
+               subset = list("Grid/lon" = -80:-50,
+                             "Grid/lat" = -45:-15)) %>%
+      .[, fecha := ymd_hms(paste0(meta[[1]][[1]], meta[[1]][[3]]))]
+
+  }) %>%
+    rbindlist() %>%
+    setnames(c("Grid/lon", "Grid/lat"), c("lon", "lat")) %>%
+    .[, .(pp_acum = sum(pp)), by = .(lon, lat)]
+
+
+  # Modelo
+
+  files_wrf <-  purrr::map(seq(0, acumulado - 1), function(l) {
+    list.files(path = paste0(wrf_path, exp, "/ANA/", format(date - hours(l), "%Y%m%d%H%M%S")),
+               full.names = TRUE,
+               recursive = TRUE,
+               pattern = "analysis.ensmean")
[…]

### 2020-11-18 R/read_diag_conv.R
Agrego read_diag_conv

@@ -1,0 +1,80 @@
+#' Lectura de archivos diag para observaciones convencionales
+#'
+#' Internamente define la fecha.
+#' El formato de nombre de archivo debería ser "asim_conv_{date}.{mem}"
+#'
+#' @param file_list lista de archivos
+#' @param exp nombre del experimento asociado
+#' @param member miembro del ensamble en caracteres. Si es la media correspode "000"
+#' @param variable vector con las variables que se quieren obtener.
+#' @export
+# read_diag_mean ---------------------------------------------------------------
+# Read mean diagfiles and tidy uv observations
+
+read_diag_conv <- function(file_list, exp, member = "000", variable = c("uv", "p", "t", "q")) {
+
+  obs <- purrr::map(file_list, function(f) {
+
+    if (file.size(f) == 0) {
+      return(NULL)
+    }
+
+    diag <- fread(f) %>%
+      # .[V10 == 1] %>%
+      .[, exp := exp] %>%
+      .[, mem := member] %>%
+      .[, date := ymd_hms(str_extract(basename(f), "\\d{14}"))] %>%
+      .[, c("V2", "V4") := NULL]
+
+    # cat("Archivo ", basename(files[f]))
+
+    colnames(diag) <- c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag.prep", "obs", "obs.guess", "obs2", "obs.guess2", "rerr", "exp", "mem", "date")
+
+    # Parse only wind
+    if ("uv" %in% variable & length(variable) == 1) {
+
+      diag <- diag[var == "uv"] %>%
+        melt(measure.vars = c("obs", "obs2", "obs.guess", "obs.guess2")) %>%
+        .[, var := if_else(str_detect(variable, "2"), "v", "u")] %>%
+        .[, variable := str_remove(variable, "2")]
+
+      vars <- rlang::syms(setdiff(names(diag), "value"))
+      diag <- diag %>%
+        distinct(!!!vars, .keep_all = TRUE) %>%
+        pivot_wider(names_from = variable, values_from = value) %>%
+        setDT %>%
+        .[, id := 1:.N, by = mem]
+
+      # Parse wind and other variables
+    } else if ("uv" %in% variable & length(variable) != 1) {
+
+
+      uv <- diag[var == "uv"] %>%
[…]

### 2020-11-18 temp_R/help_functions.R

@@ -77,77 +77,77 @@ read_diag <- function(path, variable) {
   return(obs)
 }
 
-# read_diag_mean ---------------------------------------------------------------
-# Read mean diagfiles and tidy uv observations
-
-read_diag_mean <- function(path, variable = c("uv", "p", "t", "q")) {
-
-
-  files <- list.files(path, pattern = "asim_conv", full.names = TRUE,  recursive = TRUE)
-
-
-  obs <- purrr::map(files, function(f) {
-
-    if (file.size(f) == 0) {
-      return(NULL)
-    }
-
-    diag <- fread(f) %>%
-      # .[V10 == 1] %>%
-      .[, exp := basename(dirname(f))] %>%
-      .[, mem := "00"] %>%
-      .[, date := ymd_hms(str_extract(f, "\\d{14}"))] %>%
-      .[, c("V2", "V4") := NULL]
-
-    # cat("Archivo ", basename(files[f]))
-
-    colnames(diag) <- c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag.prep", "obs", "obs.guess", "obs2", "obs.guess2", "rerr", "exp", "mem", "date")
-
-    if ("uv" %in% variable & length(variable) == 1) {
-
-      diag <- diag[var == "uv"] %>%
-        melt(measure.vars = c("obs", "obs2", "obs.guess", "obs.guess2")) %>%
-        .[, var := if_else(str_detect(variable, "2"), "v", "u")] %>%
-        .[, variable := str_remove(variable, "2")]
-
-      vars <- rlang::syms(setdiff(names(diag), "value"))
-      diag <- diag %>%
-        distinct(!!!vars, .keep_all = TRUE) %>%
-        pivot_wider(names_from = variable, values_from = value) %>%
-        setDT %>%
-        .[, id := 1:.N, by = mem]
-
-    } else if ("uv" %in% variable & length(variable) != 1) {
-
-
-      uv <- diag[var == "uv"] %>%
-        melt(measure.vars = c("obs", "obs2", "obs.guess", "obs.guess2")) %>%
-        .[, var := if_else(str_detect(variable, "2"), "v", "u")] %>%
-        .[, variable := str_remove(variable, "2")]
-
-      vars <- rlang::syms(setdiff(names(uv), "value"))
-      uv <- uv %>%
-        distinct(!!!vars, .keep_all = TRUE) %>%
[…]

### 2020-11-19 analysis/chapters/experimento_20181122.Rmd
Analisis de humedad en primeros niveles

@@
 
-```{r q-wind, warning=FALSE, fig.align="center", fig.height=5, fig.width=9, message=FALSE,dev="png"}
-fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
+#### Low levels humidity
+
+Specific humidity field averaged over the first 7 levels of the model, from surface to 800 hPa for low topography areas (aproximately).
+
+* AUT add humidity to the model (is correcting a dry model bias or modifing the circulation and transporting humidity from the north?)
+* SATWND is not that different from AUT at list in low levels. The humidity is distributed a little bit different in the domain during 06Z
+* RAD shows less humidity that AUT and SATWND but more than CONV.
+RAD seems to reverse (in part) the effect of incorporating automatic station observations in low levels.
+This can also be seeing in the first levels the difference between RAD and SATWND.
+
+```{r q-wind, warning=FALSE, fig.align="center", message=FALSE,dev="png"}
+fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
                      exp = c("E4", "E5", "E6", "E7"))
@@
 
-q[bottom_top <= 7, .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
+arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10)
+arg_topo[, lon := metR::ConvertLongitude(lon)]
+
+cordillera <- list(
+  ggnewscale::new_scale_fill(),
+  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
+                                         z = h),
+                    breaks = seq(1500, 8000, by = 700)),
+  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
+                      oob = scales::squish))
+
+q[bottom_top %between% c(0, 7), .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
   by = .(south_north, west_east, date, exp)] %>% 
@@
   # scale_mag() +
+  cordillera +
   geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
@@
 
+Is the analysis correcting a dry bias on the model or the humidity is being transported into the domain?
+
+The innovation (OmB) for surface humidity observations is mainly positive indicating that the model has always less humidity than the atmosfere (reallity?) throghout the domain for a specific cycle.
+
 ```{r}
 
+exp <- c("E4", "E5", "E6", "E7")
+
+q_field <- purrr::map(exp, function(f) {  
+  f
+  read_diag_conv(paste0("/home/paola.corrales/datosmunin/EXP/", f,
+                        "/ANA/20181121180000/diagfiles/asim_conv_20181121180000.ensmean"), 
+                 exp = f, member = "000", variable = c("q"))
+}) %>% rbindlist()
+
+q_field[usage.flag == 1 & type %in% c(181, 187) & obs.guess < 0.02] %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = obs.guess)) +
+  scale_color_divergent() +
+  geom_mapa() +
+  facet_wrap(~exp)
+```
+
+The evolution of the mean of the absolute OmB is always positive for all four experiments, AUT and SATWND showing higher values follow by RAD.
+
+```{r}
+files <- list.files("/home/paola.corrales/datosmunin/EXP/", full.names = TRUE, recursive = TRUE,
+                    pattern = "asim_conv_201811*")
+files <- files[!str_detect(files, pattern = "mem")]
+files <- files[!str_detect(files, pattern = "training")]
+files <- files[!str_detect(files, pattern = "prueba")]
+files <- files[!str_detect(files, pattern = "analisis")]
+files <- files[!str_detect(files, pattern = "derived_data")]
+
+conv <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP//{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.ensmean")
+  read_diag_conv(f, exp = meta[[1]][["exp"]], member = "000", variable = c("q"))
+}) %>% rbindlist()
+
+conv[usage.flag == 1 & type %in% c(181, 187), ":="(mean_omb = mean(abs(obs.guess), na.rm = TRUE),
+                                                   n = .N), 
+     by = .(date, exp)] %>% 
+  ggplot(aes(date, mean_omb)) +
+  geom_line(aes(color = exp)) +
+  geom_point(aes(color = exp, size = n)) +
+  scale_size_area(max_size = 3) +
+  scale_color_brewer(palette = "Dark2") +
+  labs(y = "Mean of abs(OmB) (g/g)", x = NULL) +
+  theme_minimal()
+```
+
+```{r}
+a <- ReadNetCDF("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121180000/analysis.ensmean", 
+                vars = c(p = "P", "PB", qv = "QVAPOR", 
+                         lon = "XLONG", lat = "XLAT")) %>% 
+  .[, p := p + PB] %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)]
+
+a[bottom_top == 12] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z =  p/100), proj = norargentina_lambert) +
+  scale_fill_viridis_b()
+
+exp <- c("E4", "E5", "E6", "E7")
+
+pblh <- purrr::map(exp, function(f) {
+
+  ReadNetCDF(paste0("/home/paola.corrales/datosmunin/EXP/", f, "/ANA/20181121180000/analysis.ensmean"), 
+             vars = c("PBLH", lon = "XLONG", lat = "XLAT")) %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+    .[, exp := f]
+
+}) %>% 
+  rbindlist()
+
+pblh[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z =  PBLH), proj = norargentina_lambert,
+                    breaks = seq(0, 5000, 500)) +
+  scale_fill_viridis_b(breaks = seq(0, 5000, 500)) +
+  geom_mapa() +
+  facet_wrap(~ exp) +
+  theme_minimal()
+
+pblh[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
+  .[, .(mean = mean(PBLH)), by = .(exp, south_north)] %>% 
+  ggplot(aes(south_north, mean)) +
+  geom_point(aes(color = exp))
 ```
@@
                     full.names = TRUE,
-                    pattern = "fss_3h")
+                    pattern = "fss_6h")
 

### 2020-11-20 analysis/scripts/wrap_fss.R
Script para calcular FSS ahora también adaptado a FCST (en formato NPP)

@@ -10,13 +10,14 @@ library(mesoda)
 future::plan("cluster", workers = 3)
 imerg_path <- "/home/paula.maldonado/datosalertar1/RRA_VERIF/data/raw/imerg_raw"
 wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E4"
+exp <- "E7"
+run <- "fcst"
 
-ini_date <- ymd_hms("20181120180000")
-ciclos <- 68
+ini_date <- ymd_hms("20181122000000")
+ciclos <- 37
 
-acumulado <- 3
-q <- c(1, 5, 10) #10mm, para arrancar pensanod en pp acumulada
+acumulado <- 6
+q <- c(1, 5, 10) #10mm, para arrancar pensando en pp acumulada
 w <- c(2, 5, 25, 50) #ancho de cada box w*2+1
 
 # Inicio
@@ -48,28 +49,63 @@ fss_out <- furrr::future_map_dfr(dates, function(d) {
 
   # Modelo
 
-  files_wrf <-  purrr::map(seq(0, acumulado - 1), function(l) {
-    list.files(path = paste0(wrf_path, exp, "/ANA/", format(date - hours(l), "%Y%m%d%H%M%S")),
-               full.names = TRUE,
-               recursive = TRUE,
-               pattern = "analysis.ensmean")
-  }) %>% unlist()
+  if (run == "ana") {
+    files_wrf <-  purrr::map(seq(0, acumulado - 1), function(l) {
+      list.files(path = paste0(wrf_path, exp, "/", toupper(run), "/", format(date - hours(l), "%Y%m%d%H%M%S")),
+                 full.names = TRUE,
+                 recursive = TRUE,
+                 pattern = "analysis.ensmean")
+    }) %>% unlist()
 
 
-  pp_wrf <- purrr::map(files_wrf, function(f) {
-    ReadNetCDF(f, vars = c("RAINNC", "RAINC", "RAINSH",
-                           lon = "XLONG", lat = "XLAT")) %>%
-      .[, ":="(pp_acum = RAINNC + RAINC + RAINSH,
-               exp = exp,
-               date = date)] %>%
-      .[, ":="(RAINNC = NULL,
-               RAINC = NULL,
-               RAINSH = NULL,
-               Time = NULL)]
-  }) %>%
-    rbindlist() %>%
-    .[, .(pp_acum = sum(pp_acum, na.rm = TRUE)), by = .(lon, lat, date, exp)]
[…]

### 2020-11-20 analysis/scripts/wrap_fss.R
fix bug

@@ -10,7 +10,7 @@ library(mesoda)
 future::plan("cluster", workers = 3)
 imerg_path <- "/home/paula.maldonado/datosalertar1/RRA_VERIF/data/raw/imerg_raw"
 wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E7"
+exp <- "E6"
 run <- "fcst"
 
 ini_date <- ymd_hms("20181122000000")
@@ -21,7 +21,8 @@ q <- c(1, 5, 10) #10mm, para arrancar pensando en pp acumulada
 w <- c(2, 5, 25, 50) #ancho de cada box w*2+1
 
 # Inicio
-dates <- seq.POSIXt(ini_date, by = "hour", length.out = ciclos)
+dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
+                    length.out = ciclos - acumulado)
 
 fss_out <- furrr::future_map_dfr(dates, function(d) {
 
@@ -76,7 +77,7 @@ fss_out <- furrr::future_map_dfr(dates, function(d) {
 
     lead_time <- as.numeric(difftime(date, ini_date, units = "hours"))
 
-    files_wrf <-  purrr::map(seq(0, acumulado - 1), function(l) {
+    files_wrf <-  purrr::map(c(acumulado, 0), function(l) {
       paste0(wrf_path, exp, "/", toupper(run), "/",
              format(ini_date, "%Y%m%d%H"), "/NPP/NPP_",
              format(ini_date, "%Y"), "-",
@@ -99,11 +100,12 @@ fss_out <- furrr::future_map_dfr(dates, function(d) {
         .[, ":="(XLONG = NULL, XLAT = NULL)] %>%
         .[, .(pp_acum = mean(PP, na.rm = TRUE),
               exp = exp,
-              date = date), by = .(lat, lon)]
+              date = date,
+              t = basename(f)), by = .(lat, lon)]
 
     }) %>%
       rbindlist() %>%
-      .[, .(pp_acum = sum(pp_acum, na.rm = TRUE)), by = .(lon, lat, date, exp)]
+      .[, .(pp_acum = diff(pp_acum)), by = .(lon, lat, date, exp)]
 
   }
 

### 2020-11-20 analysis/chapters/experimento_20181122.Rmd
Arreglo grafico de OmB e incorporo nueva escala para graficos de humedad

@@
 output: 
+  bookdown::pdf_document2: default
   bookdown::html_document2: default
-  bookdown::pdf_document2: default
 bibliography: mesoda_ref.bib
@@
   # fig.height = 3,
+  dev = "png",
   message = FALSE,
@@
 
-geom_mapa <- function() {
-  list(geom_sf(data = map, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE),
+geom_mapa <- function(fill = NA) {
+  list(geom_sf(data = map, fill = fill, color = "black", size = 0.2, inherit.aes = FALSE),
        coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)))
@@
                                     "GF-MYNN2"), length.out = 60)) %>% setDT()
+
+experiments <- c("E4", "E5", "E6", "E7")
+#Color blind friendly but E4 y E7 parecidos en Desaturated
+colores_exp <- c(E4 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
+# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
+
 derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"
@@
 
-ggplot(data = perfiles[exp == "E5" & V < 0], aes(date, lev)) +
-  geom_contour2(aes(z = V))
-
 ```
@@
   ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = q*1000), proj = norargentina_lambert,
+  geom_contour_fill(aes(z = q*1000, fill = stat(level_d)), proj = norargentina_lambert,
                     breaks = seq(0, 0.02, 0.002)*1000) +
-  scale_fill_distiller(name = NULL,
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
                        palette = "BrBG", direction = 1,
-                       breaks = seq(0, 0.02, 0.002)*1000,
-                       guide = guide_colorstrip(inside = FALSE,
+                       limits = c(1, 19),
+                       guide = guide_colorsteps(#inside = FALSE,
                                                 barwidth = 25,
@@
 ```
+```{r}
+
+q[bottom_top %between% c(8, 12), .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
+  by = .(south_north, west_east, date, exp)] %>% 
+  .[coord, on = c("south_north", "west_east")] %>% 
+  .[, date := factor(date)] %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = q*1000, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 0.02, 0.002)*1000) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       limits = c(1, 19),
+                       guide = guide_colorsteps(#inside = FALSE,
+                                                barwidth = 25,
+                                                barheight = 0.8)) +
+  # geom_vector(aes(x = lon, y = lat,
+  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
+  #             arrow.angle = 12, size = 0.3) +
+  # scale_mag() +
+  cordillera +
+  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
+  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
+  facet_grid(date ~ exp, 
+             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Humidity (g/Kg) 800-700 hPa",
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
 
@@
 
-The innovation (OmB) for surface humidity observations is mainly positive indicating that the model has always less humidity than the atmosfere (reallity?) throghout the domain for a specific cycle.
+The innovation (OmB) for surface humidity observations is mainly positive indicating that the model has in general less humidity than the atmosfere throghout the domain for a specific cycle.
 
-```{r}
+```{r omb-q, fig.cap="Observation minus guess for humidity surface observations during Nov, 21st 18Z"}
 
@@
   ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_mapa(fill = "grey90") +
   geom_point(aes(color = obs.guess)) +
-  scale_color_divergent() +
-  geom_mapa() +
-  facet_wrap(~exp)
+  scale_color_divergent(guide = guide_colorbar(barwidth = 25,
+                                                barheight = 0.8)) +
+
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  facet_wrap(~exp, ncol = 4, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  labs(x = NULL, y = NULL, color = "OmB") +
+    theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
 ```
@@
 The evolution of the mean of the absolute OmB is always positive for all four experiments, AUT and SATWND showing higher values follow by RAD.
+The size of the points shows that AUT, SATWND and RAD assimilate more observations.
 
-```{r}
+This result is indicating that the model has an important dry bias and the assimilation of high temporal and spatial resolution observations are correcting this.
+
+TODO: it's the same for all PBL parameterizations?
+
+```{r omb-q-evol, fig.cap="Evolution of the mean OmB for humidity throughout the periodod"}
 files <- list.files("/home/paola.corrales/datosmunin/EXP/", full.names = TRUE, recursive = TRUE,
@@
 
-conv[usage.flag == 1 & type %in% c(181, 187), ":="(mean_omb = mean(abs(obs.guess), na.rm = TRUE),
+conv[usage.flag == 1 & type %in% c(181, 187) & exp %in% experiments, 
+     .(mean_omb = mean(obs.guess, na.rm = TRUE),
                                                    n = .N), 
@@
   scale_size_area(max_size = 3) +
-  scale_color_brewer(palette = "Dark2") +
-  labs(y = "Mean of abs(OmB) (g/g)", x = NULL) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                          E6 = "SATWND", E7 = "RAD")) +
+  labs(y = "Mean of OmB (g/g)", x = NULL,
+       color = NULL, size = "n obs") +
   theme_minimal()
+
+# conv[usage.flag == 1 & type %in% c(181, 187) & exp %in% experiments] %>% 
+#   ggplot(aes(date, obs.guess)) +
+#   geom_point(aes(color = exp, group = date), alpha = 0.1) +
+# scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+#                                           E6 = "SATWND", E7 = "RAD")) +
+#   geom_smooth() +
+#   facet_wrap(~exp) +
+#   labs(y = "Mean of OmB (g/g)", x = NULL,
+#        color = NULL, size = "n obs") +
+#   theme_minimal()
+# ,
+#      .(n = .N), 
+#      by = .(date, exp)] %>% 
+#   ggplot(aes(date, n)) +
+#   geom_line(aes(color = exp)) +
+#   # geom_point(aes(color = exp, size = n)) +
+#   # scale_size_area(max_size = 3) +
+#   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+#                                           E6 = "SATWND", E7 = "RAD")) +
+#   labs(y = "Mean of OmB (g/g)", x = NULL,
+#        color = NULL, size = "n obs") +
+#   theme_minimal()
+
 ```
 
-```{r}
-a <- ReadNetCDF("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121180000/analysis.ensmean", 
-                vars = c(p = "P", "PB", qv = "QVAPOR", 
-                         lon = "XLONG", lat = "XLAT")) %>% 
-  .[, p := p + PB] %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)]
+The negative difference between RAD and SATWND in low levels (\@ref(fig:Q-diff)) and positive values over those levels can be explain by a deeper boundary layer.
+To explore this posibility Figure \@ref(fig:pbl-18Z) shows the zonal mean high of the PBL over the subdomain estimated for the PBL parameterization for Nov 21st, 18Z.
+The mean is calculated over a big area with different hights than can altered the result but but still the hight of the boundary layer for RAD is greater than in the other experiments in many areas.
 
-a[bottom_top == 12] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z =  p/100), proj = norargentina_lambert) +
-  scale_fill_viridis_b()
-
-exp <- c("E4", "E5", "E6", "E7")
-
-pblh <- purrr::map(exp, function(f) {
+```{r pbl-18Z, fig.cap="Zonal mean high of the PBL over the subdomain estimated for the PBL parameterization for Nov 21st, 18Z"}
+pblh18 <- purrr::map(exp, function(f) {
 
   ReadNetCDF(paste0("/home/paola.corrales/datosmunin/EXP/", f, "/ANA/20181121180000/analysis.ensmean"), 
-             vars = c("PBLH", lon = "XLONG", lat = "XLAT")) %>% 
+             vars = c("PBLH", lon = "XLONG", lat = "XLAT"),
+             subset = list(south_north = c(41, 119),
+                           west_east = c(96, 143))) %>% 
     .[, c("x", "y") := wrf_project(lon, lat)] %>% 
@@
 
-pblh[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z =  PBLH), proj = norargentina_lambert,
-                    breaks = seq(0, 5000, 500)) +
-  scale_fill_viridis_b(breaks = seq(0, 5000, 500)) +
-  geom_mapa() +
-  facet_wrap(~ exp) +
-  theme_minimal()
-
-pblh[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
+pblh18[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
   .[, .(mean = mean(PBLH)), by = .(exp, south_north)] %>% 
   ggplot(aes(south_north, mean)) +
-  geom_point(aes(color = exp))
+  geom_point(aes(color = exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                          E6 = "SATWND", E7 = "RAD")) +
+  labs(x = "Grid points (south - north)", y = "mean PBLH (m)", color = NULL) +
+  theme_minimal()
 ```
 
-#### FSS for analysis
+If the average is calculated over a smaller domain to avoid hight terrain and water, it is posible to see the evolution of the PBL hight during the experiment (Figure \@ref(fig:pblh-evol)).
+In CONV and RAD the PBL is higher during the diurnal hours of Nov, 21st with, exceeding AUT and SATWND by about 200 meters.
+These hours coincide with the time when the highest humidity above 800 hPa in RAD is observed.
+After that, the behavior chances and the development of the convection over the domain starts to affect the PBL.
+
+The deeper PBL in RAD can explain the transport of humidity to higher levels. This efect does not appear in CONV because there is no much humidity to transport to upper levels.
+
+TODO: Are there RAD observations on those levels to explain the humidity besides the transport from low levels?
+
+```{r pblh, fig.cap=""}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E[4-7]/ANA/*/analysis.ensmean")
+
+pblh <- purrr::map(files, function(f) {
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{date}/analysis.ensmean")
+
+  ReadNetCDF(paste0(f), 
+             vars = c("PBLH", lon = "XLONG", lat = "XLAT"),
+             subset = list(south_north = c(41, 119),
+                           west_east = c(96, 143))) %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+    .[, exp := meta[[1]][["exp"]]] %>% 
+    .[, date := ymd_hms(meta[[1]][["date"]])]
+
+}) %>% 
+  rbindlist()
+
+# pblh[lat %between% c(-38, -31) & lon %between% c(-64, -59)] %>%
+#   ggplot(aes(x, y)) +
+#   geom_contour_fill(aes(z =  PBLH), proj = norargentina_lambert,
+#                     breaks = seq(0, 5000, 500)) +
+#   scale_fill_viridis_b(breaks = seq(0, 5000, 500)) +
+#   geom_mapa() +
+#   facet_wrap(~ exp) +
+#   theme_minimal()
+
+pblh[, .(mean = mean(PBLH)), by = .(exp, date)] %>% 
+  ggplot(aes(date, mean)) +
+  geom_line(aes(color = exp)) +
+  geom_point(aes(color = exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                          E6 = "SATWND", E7 = "RAD")) +
+  labs(x = NULL, y = "mean PBLH (m)", color = NULL) +
+  theme_minimal()
+```
+
+#### FSS for the analysis
 
@@
   ggplot(aes(date, fss)) +
-  geom_point(aes(color = exp)) +
-  geom_line(aes(color = exp, linetype = factor(w))) +
-  facet_grid(w~q)
+  geom_hline(yintercept = 1, color = "darkgray") +
+  geom_line(aes(color = exp)) +
+  # geom_point(aes(color = exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                          E6 = "SATWND", E7 = "RAD")) +
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", "10" = "10 mm"),
+                                      w = c("2" = "50 km", "5" = "100 km", 
+                                            "25" = "500 km", "50" = "1000 km"))) +
+  labs(color = NULL, x = NULL) +
+  theme_minimal()
 
 ```
+#### FSS for the forecast
+
+```{r}
+files <- list.files(here("analysis", "data", "derived_data"), 
+                    full.names = TRUE,
+                    pattern = "fss_6h_fcst")
+
+fss <- purrr::map(files, function(f) {
+
+  fread(f) %>% 
+    .[, date := as_datetime(date)]
+}) %>% 
+  rbindlist()
+
+fss %>% 
+  ggplot(aes(date, fss)) +
+  geom_hline(yintercept = 1, color = "darkgray") +
+  geom_line(aes(color = exp)) +
+  # geom_point(aes(color = exp)) +
+  scale_color_manual(values = colores_exp) +
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", "10" = "10 mm"),
+                                      w = c("2" = "50 km", "5" = "100 km", 
+                                            "25" = "500 km", "50" = "1000 km"))) +
+  labs(color = NULL, x = NULL) +
+  theme_minimal()
+
+```

### 2020-11-25 analysis/chapters/experimento_20181122.Rmd
Otra opcion de figura con observaciones (espacial)

@@
 
-oficiales <- obsCONV[type != 290, unique(stationID)]
+obsSATW <- fread(here("analysis", "data", "raw_data", "E6_asim_conv_20181121120000.ensmean"), 
+                 na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
+  .[, c("V2", "V4") := NULL] %>% 
+  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))
 
-no_oficiales <- unique(obsAUT, by = c("stationID")) %>% 
-  .[type != 290] %>% 
+oficiales <- obsCONV[type %in% c(181, 281)] %>% unique(by = c("stationID")) %>% 
+  .[, source := "Sfc - Official"]
+
+sondeos <- obsCONV[type == 120] %>% unique(by = c("stationID")) %>% 
+  .[, source := "Radiosonde"]
+
+ascat <- obsCONV[type == 290] %>% unique(by = c("stationID")) %>% 
+  .[, source := "ASCAT"]
+
+aircraf <- obsCONV[type %in% c(130, 131, 133)] %>%# unique(by = c("stationID")) %>% 
+  .[, source := "AIRCRFT"]
+
+satwnd <- obsSATW[type %between% c(240, 260)] %>%# unique(by = c("stationID")) %>% 
+  .[, source := "SATWND"]
+
+no_oficiales <- obsAUT %>% 
+  .[type %in% c(181, 187)] %>% 
   .[!str_detect(stationID, "SMN")] %>%
-  .[!stationID %in% oficiales & type != 290, stationID]
+  .[!stationID %in% oficiales$stationID] %>% 
+  unique(by = c("stationID")) %>% 
+  .[, source := "Sfc - Non-official"]
 
-unique(obsAUT, by = c("stationID")) %>% 
-  .[type != 290] %>% 
-  .[!str_detect(stationID, "SMN")] %>% 
+rbindlist(list(oficiales, no_oficiales, ascat, sondeos, aircraf, satwnd)) %>% 
   ggplot(aes(ConvertLongitude(lon), lat)) +
   geom_mapa() +
-  geom_point(alpha = 0.8, size = 2,
-             aes(color = stationID %in% oficiales)) +
-  scale_color_manual(name =  NULL, values = c("TRUE" = "#FD8002", "FALSE" = "#00695c"),
-                     breaks = c("FALSE", "TRUE"),
+  geom_point(alpha = 1, size = 0.5,
+             aes(color = source, shape = source)) +
+  scale_color_brewer(palette = "Dark2") +
+  # khroma::scale_color_vibrant() +
+  geom_point(data = square, aes(lon, lat), size = 0.2) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(x = NULL, y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+rbindlist(list(no_oficiales, oficiales)) %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_mapa() +
+  geom_point(alpha = 0.8, size = 1,
+             aes(color = source)) +
+  scale_color_manual(name =  NULL, values = c("Sfc - Non-official" = "#00695c", 
+                                              "Sfc - Official" = "#FD8002"),
+                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                      labels = c("Non-official (N = 820)","Official (N = 327)"))+

### 2020-12-04 analysis/chapters/experimento_20181122.Rmd
Muchas figuras nuevas

@@
 }
+
+arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10)
+arg_topo[, lon := metR::ConvertLongitude(lon)]
+
+cordillera <- list(
+  ggnewscale::new_scale_fill(),
+  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
+                                         z = h),
+                    breaks = seq(1500, 8000, by = 700)),
+  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
+                      oob = scales::squish))
 
@@
 
+```{r rad}
+
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/*/diagfiles/asim*ensmean")
+
+files <- files[!str_detect(files, "conv")]
+
+rad <- read_diag_rad(files, "E7")
+
+satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
+  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
+
+rad <- satinfo[rad, on = c("sensor", "channel")]
+
+rad %>% 
+  .[qc == 0 & iuse == 1] %>% 
+  .[, lev := cut_width(peakwt, 50)] %>% 
+  .[, .(omb = mean(tbc),
+              n = .N), by = .(date, lev)] %>% 
+  ggplot(aes(date, fct_rev(lev))) +
+  geom_raster(aes(fill = omb)) +
+  scale_fill_divergent(trans = mesoda::symlog_trans(),
+                       guide = guide_colourbar(barheight = 20))  +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H:%M") +
+  labs(x = NULL, fill = NULL, y = "pressure layer")
+
+rad %>% 
+  .[qc == 0 & iuse == 1] %>% 
+  .[, lev := cut_width(peakwt, 50)] %>% 
+  .[, .(omb = mean(tbc),
+              n = .N), by = .(date, lev)] %>% 
+  ggplot(aes(date, fct_rev(lev))) +
+  geom_raster(aes(fill = n)) +
+  scale_fill_viridis_c(trans = scales::log10_trans(),
+                       guide = guide_colourbar(barheight = 20)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H:%M") +
+  labs(x = NULL, fill = NULL, y = "pressure layer")
+
+```
+
 ### Analysis comparison
@@
   rbindlist() %>% 
-  .[, date := ymd_hms(date)] %>% 
+  .[, date := as_datetime(date)] %>% 
   .[date %between% c(as_datetime("2018-11-20 18:00:00"),
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value), breaks = seq(-1, 1, 0.2)) +
-  scale_fill_divergent(name = NULL,
-                       breaks = seq(-1, 1, 0.2),
-                       guide = guide_colorstrip(inside = TRUE,
-                                                barwidth = 25,
+  geom_contour_fill(aes(z = value, fill = stat(level))) +
+  scale_fill_divergent_discretised(name = NULL,
+                       guide = guide_colorsteps(barwidth = 25,
                                                 barheight = 0.8)) +
@@
 
+```{r T-mse, fig.cap="MSE Temperature between analysis and guess in Kelvin^2."}
+
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/perfiles_mse*")
+
+perfiles_mse <- purrr::map(files, function(f) {
+  exp <- unglue(basename(f), "perfiles_{run}_{exp}.csv")
+  fread(f) %>% 
+    .[, run := exp[[1]][["run"]]] %>% 
+    .[, exp := exp[[1]][["exp"]]]
+
+}) %>% 
+  rbindlist() %>% 
+  .[, date := as_datetime(date)] %>% 
+  .[date %between% c(as_datetime("2018-11-20 20:00:00"),
+                     as_datetime("2018-11-23 12:00:00"))]
+
+perfiles_mse %>% 
+  .[variable == "T"] %>% 
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = mse, fill = stat(level))) +
+  scale_fill_viridis_c(option = "A", direction = -1,
+                       name = NULL, super = ScaleDiscretised,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  labs(subtitle = "Temperature MSE (K^2)",
+       x = NULL,
+       y = "Pressure level") +
+  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+
 ```{r Q-diff, fig.cap="Specific humidity difference between experiments in g/Kg"}
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value*1000), breaks = seq(-0.0016, 0.0016, 0.0002)*1000) +
-  scale_fill_divergent(name = NULL,
-                       breaks = seq(-0.0016, 0.0016, 0.0002)*1000,
-                       guide = guide_colorstrip(inside = TRUE,
-                                                barwidth = 25,
+  geom_contour_fill(aes(z = value*1000, fill = stat(level)), 
+                    breaks = c(seq(-1, 1.6, 0.2), Inf)) +
+  scale_fill_divergent_discretised(name = NULL, limits = c(-1, 1.6),
+                       guide = guide_colorsteps(barwidth = 25,
                                                 barheight = 0.8)) +
@@
 A positive difference can also be observed between 850 and 600 hPa.
+```{r Q-mse, fig.cap="MSE Specific humidity between analysis and guess in g/Kg^2."}
+
+perfiles_mse %>% 
+  .[variable == "QVAPOR"] %>% 
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = mse*1000000, fill = stat(level))) +
+  scale_fill_viridis_c(option = "A", direction = -1,
+                       name = NULL, super = ScaleDiscretised,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  labs(subtitle = "Specific humidity MSE (g/Kg^2)",
+       x = NULL,
+       y = "Pressure level") +
+  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
 
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value), breaks = seq(-1.2, 1.4, 0.2)) +
-  scale_fill_divergent(name = NULL,
-                       breaks = seq(-1.2, 1.4, 0.2),
-                       labels = function(x) round(x, digits = 1),
-                       guide = guide_colorstrip(inside = TRUE,
-                                                barwidth = 25,
+  geom_contour_fill(aes(z = value, fill = stat(level)), breaks = c(seq(-1.6, 2, 0.2), Inf)) +
+  scale_fill_divergent_discretised(name = NULL,
+                       limits = c(-1.6, 2),
+                       guide = guide_colorsteps(barwidth = 25,
                                                 barheight = 0.8)) +
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value), breaks = seq(-1.8, 1.2, 0.2)) +
-  scale_fill_divergent(name = NULL,
-                       breaks = seq(-1.8, 1.2, 0.2),
-                       labels = function(x) round(x, digits = 1),
-                       guide = guide_colorstrip(inside = TRUE,
-                                                barwidth = 25,
+  geom_contour_fill(aes(z = value, fill = stat(level)), 
+                    breaks = c(-Inf, seq(-1.6, 1.2, 0.2), Inf)) +
+  scale_fill_divergent_discretised(name = NULL, limits = c(-1.6, 1.2),
+                       guide = guide_colorsteps(barwidth = 25,
                                                 barheight = 0.8)) +
@@
 
-```{r q-wind, warning=FALSE, fig.align="center", message=FALSE,dev="png"}
+```{r q-low, warning=FALSE, fig.align="center", message=FALSE}
 fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
@@
   rbindlist()
-
-arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10)
-arg_topo[, lon := metR::ConvertLongitude(lon)]
-
-cordillera <- list(
-  ggnewscale::new_scale_fill(),
-  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
-                                         z = h),
-                    breaks = seq(1500, 8000, by = 700)),
-  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
-                      oob = scales::squish))
 
@@
 ```
-```{r}
+
+```{r q-middle}
 
@@
 ```
+```{r v-low}
+q[bottom_top %between% c(0, 7), .(v = mean(v)), 
+  by = .(south_north, west_east, date, exp)] %>% 
+  .[coord, on = c("south_north", "west_east")] %>% 
+  .[, date := factor(date)] %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = v, fill = stat(level_d)), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(name = NULL,
+                       # limits = c(1, 19),
+                       guide = guide_colorsteps(#inside = FALSE,
+                                                barwidth = 25,
+                                                barheight = 0.8)) +
+  # geom_vector(aes(x = lon, y = lat,
+  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
+  #             arrow.angle = 12, size = 0.3) +
+  # scale_mag() +
+  cordillera +
+  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
+  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
+  facet_grid(date ~ exp, 
+             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Low level meridional wind (m/s)",
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+```{r v-middle}
+q[bottom_top %between% c(8, 12), .(v = mean(v)), 
+  by = .(south_north, west_east, date, exp)] %>% 
+  .[coord, on = c("south_north", "west_east")] %>% 
+  .[, date := factor(date)] %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = v, fill = stat(level_d)), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(name = NULL,
+                       # limits = c(1, 19),
+                       guide = guide_colorsteps(#inside = FALSE,
+                                                barwidth = 25,
+                                                barheight = 0.8)) +
+  # geom_vector(aes(x = lon, y = lat,
+  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
+  #             arrow.angle = 12, size = 0.3) +
+  # scale_mag() +
+  cordillera +
+  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
+  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
+  facet_grid(date ~ exp, 
+             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "800-700 hPa meridional wind (m/s)",
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+
+```{r}
+q[date == unique(date)[1:3] & west_east %between% c(70, 180)] %>% 
+  .[, .(q_mean = mean(q),
+        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(south_north, bottom_top)) +
+  geom_contour_fill(aes(z = q_mean*1000, fill = stat(level_d))) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       # limits = c(1, 19),
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  geom_contour(aes(z = v_mean, linetype = factor(-sign(..level..))), 
+               color = "grey30", size = 0.5, binwidth = 2) +
+  geom_text_contour(aes(z = v_mean), color = "grey30", stroke = 0.1, breaks = seq(-10, 6, 2)) +
+  scale_linetype(guide = NULL) +
+  # geom_vector(aes(x = south_north, y = bottom_top,
+  #                 mag = abs(v_mean), angle = fifelse(v_mean > 0, 0, 180)),
+  #             arrow.angle = 6, size = 0.2, skip.x = 10) +
+  # scale_mag() +
+  facet_grid(date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  labs(y =  "Sigma levels", x = "Grid points") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+
+```{r}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/ANA/sondeos/*")
+
+sondeos <- purrr::map(files, function(f){
+
+   fread(f) %>% 
+    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
+    .[, launch_time := as_datetime(launch_time)]
+
+}) %>% 
+  rbindlist()
+
+sondeos %>% 
+  .[variable == "t" & !str_detect(site, "/") & 
+      exp == "E4"] %>% 
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = value), size = 0.5) +
+  scale_color_distiller(palette = "YlOrRd", direction = 1) +
+  # geom_contour(aes(z = value - fcst_value)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  labs(color = "T", y = NULL, x = NULL) +
+  theme_minimal() +
+theme(legend.position = "bottom") +
+
+sondeos %>% 
+  .[variable == "t" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
+  .[, oma := value - fcst_value] %>% 
+  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = oma), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  labs(color = "OmA", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(0.3, 0.7))
+```
+
+```{r}
+sondeos %>% 
+  .[variable == "td" & !str_detect(site, "/") & 
+      exp == "E4"] %>% 
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = value), size = 0.5) +
+  scale_color_distiller(palette = "YlOrRd", direction = 1) +
+  # geom_contour(aes(z = value - fcst_value)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  labs(color = "Td", y = NULL, x = NULL) +
+  theme_minimal() +
+theme(legend.position = "bottom") +
+
+sondeos %>% 
+  .[variable == "td" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
+  .[, oma := value - fcst_value] %>% 
+  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = oma), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  labs(color = "OmA", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(0.3, 0.7))
+```
+
+```{r}
+sondeos %>% 
+  .[variable == "v" & !str_detect(site, "/") & 
+      exp == "E4"] %>% 
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = value), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  labs(color = "V (m/s)", y = NULL, x = NULL) +
+  theme_minimal() +
+theme(legend.position = "bottom") +
+
+sondeos %>% 
+  .[variable == "v" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
+  .[, oma := value - fcst_value] %>% 
+  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = oma), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  labs(color = "OmA", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(0.3, 0.7))
+```
+
+```{r mucape}
+files <- Sys.glob(paste0(derived_data, "mucape/*"))
+
+files <- files[!str_detect(files, c("E3"))]
+
+dates <- c("20181121180000", "20181122000000", "20181122060000")
+
+mucape <- purrr::map(files, function(f) {
+  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]
+
+  if (details$date %in% dates) {
+
+  ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
+    .[, `:=`(exp = details$exp, 
+             date = lubridate::ymd_hms(details$date))] %>%
+    setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
+    .[]
+  }
+}) %>% 
+  rbindlist()
+
+mucape %>% 
+  .[variable == "mcape"] %>% 
+  # .[coord, on = c("south_north", "west_east")] %>% 
+  .[, date := factor(date)] %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = value, fill = stat(level_d)), proj = norargentina_lambert) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  # cordillera +
+  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
+  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
+  facet_grid(date ~ exp, 
+             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Maximum CAPE",
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
 
@@
 ```
+```{r omb-t, fig.cap="Observation minus guess for humidity surface observations during Nov, 21st 18Z"}
+
+exp <- c("E4", "E5", "E6", "E7")
+
+q_field <- purrr::map(exp, function(f) {  
+  f
+  read_diag_conv(paste0("/home/paola.corrales/datosmunin/EXP/", f,
+                        "/ANA/20181121180000/diagfiles/asim_conv_20181121180000.ensmean"), 
+                 exp = f, member = "000", variable = c("t"))
+}) %>% rbindlist()
+
+q_field[usage.flag == 1 & type %in% c(181, 187) & abs(obs.guess) < 15] %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_mapa(fill = "grey90") +
+  geom_point(aes(color = obs.guess)) +
+  scale_color_divergent(guide = guide_colorbar(barwidth = 25,
+                                                barheight = 0.8)) +
+
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  facet_wrap(~exp, ncol = 4, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  labs(x = NULL, y = NULL, color = "OmB") +
+    theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
 
@@
 ```{r omb-q-evol, fig.cap="Evolution of the mean OmB for humidity throughout the periodod"}
-files <- list.files("/home/paola.corrales/datosmunin/EXP/", full.names = TRUE, recursive = TRUE,
-                    pattern = "asim_conv_201811*")
-files <- files[!str_detect(files, pattern = "mem")]
-files <- files[!str_detect(files, pattern = "training")]
-files <- files[!str_detect(files, pattern = "prueba")]
-files <- files[!str_detect(files, pattern = "analisis")]
-files <- files[!str_detect(files, pattern = "derived_data")]
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E[4-7]/ANA/*/diagfiles/asim_conv_201811*.ensmean")
 
@@
 
-  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP//{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.ensmean")
-  read_diag_conv(f, exp = meta[[1]][["exp"]], member = "000", variable = c("q"))
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.ensmean")
+  read_diag_conv(f, exp = meta[[1]][["exp"]], member = "000", variable = c("q")) %>% 
+    .[stationID %in% oficiales$stationID] %>% 
+    .[]
 }) %>% rbindlist()
 
-conv[usage.flag == 1 & type %in% c(181, 187) & exp %in% experiments, 
+conv[type %in% c(181, 187) & exp %in% experiments, 
      .(mean_omb = mean(obs.guess, na.rm = TRUE),
                                                    n = .N), 
-     by = .(date, exp)] %>% 
+     by = .(date, exp, usage.flag)] %>% 
   ggplot(aes(date, mean_omb)) +
@@
   geom_point(aes(color = exp, size = n)) +
-  scale_size_area(max_size = 3) +
+  scale_size_area(max_size = 2) +
   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                           E6 = "SATWND", E7 = "RAD")) +
+  facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
+                                                             "-1" = "rejected"))) +
   labs(y = "Mean of OmB (g/g)", x = NULL,
@@
 ```{r}
-files <- list.files(here("analysis", "data", "derived_data"), 
-                    full.names = TRUE,
-                    pattern = "fss_6h")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_E[4-7]new.csv")
 
@@
                                           E6 = "SATWND", E7 = "RAD")) +
-  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", "10" = "10 mm"),
-                                      w = c("2" = "50 km", "5" = "100 km", 
-                                            "25" = "500 km", "50" = "1000 km"))) +
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
+                                            "10" = "10 mm", "25" = "25 mm"),
+                                      w = c("5" = "50 km", "11" = "100 km", 
+                                            "51" = "500 km", "101" = "1000 km"))) +
   labs(color = NULL, x = NULL) +
@@
 ```
+
 #### FSS for the forecast
@@
 ```{r}
-files <- list.files(here("analysis", "data", "derived_data"), 
-                    full.names = TRUE,
-                    pattern = "fss_6h_fcst")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_*_E[4-7]new.csv")
 
@@
 
+  meta <- unglue(basename(f), "fss_6h_fcst_{ini_date}_{exp}new.csv")
+
   fread(f) %>% 
-    .[, date := as_datetime(date)]
+    .[, date := as_datetime(date)] %>% 
+    .[, ini_date := ymd_h(meta[[1]][["ini_date"]])]
 }) %>% 
@@
   geom_hline(yintercept = 1, color = "darkgray") +
-  geom_line(aes(color = exp)) +
+  geom_line(aes(color = exp, linetype = factor(ini_date))) +
   # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp) +
-  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", "10" = "10 mm"),
-                                      w = c("2" = "50 km", "5" = "100 km", 
-                                            "25" = "500 km", "50" = "1000 km"))) +
-  labs(color = NULL, x = NULL) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                          E6 = "SATWND", E7 = "RAD")) +
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
+                                            "10" = "10 mm", "25" = "25 mm"),
+                                      w = c("5" = "50 km", "11" = "100 km", 
+                                            "51" = "500 km", "101" = "1000 km"))) +
+  labs(color = NULL, x = NULL, linetype = NULL) +
   theme_minimal()

### 2020-12-04 analysis/scripts/wrap_fss.R
Update wrap FSS

@@ -10,11 +10,11 @@ library(mesoda)
 #future::plan("cluster", workers = 3)
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
 wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E6"
+exp <- "E7"
 run <- "fcst"
 
-ini_date <- ymd_hms("20181122000000")
-ciclos <- 37
+ini_date <- ymd_hms("20181122060000")
+ciclos <- 31
 
 acumulado <- 3
 q <- c(1, 5, 10, 25) #10mm, para arrancar pensando en pp acumulada

### 2020-12-04 analysis/scripts/acumulados_imerg.R
Acum IMERG 30hs

@@ -86,3 +86,77 @@ pp_acum_imerg <- furrr::future_map_dfr(dates, function(d) {
 
 saveRDS(pp_acum_imerg, file = paste0("IMERG_", acumulado, "h.rds"))
 
+
+
+# Acumulados 30 horas -----------------------------------------------------
+
+
+imerg_path <- "/home/paula.maldonado/datosalertar1/RRA_VERIF/data/raw/imerg_raw"
+wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
+#exp <- "E4"
+#run <- "ana"
+
+ini_date <- ymd_hms("20181122060000")
+ciclos <- 85
+
+acumulado <- 30
+
+# Inicio
+ini_acum <- ymd_hms("20181122060000")
+dates <- seq.POSIXt(ini_date, by = "hour",
+                    length.out = acumulado)
+
+# Modelo
+pp_wrf <- ReadNetCDF("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181122060000/analysis.ensmean",
+                     vars = c("RAINNC", "RAINC", "RAINSH",
+                         lon = "XLONG", lat = "XLAT")) %>%
+    .[, ":="(pp_acum = RAINNC + RAINC + RAINSH,
+             exp = exp)] %>%
+    .[, ":="(RAINNC = NULL,
+             RAINC = NULL,
+             RAINSH = NULL,
+             Time = NULL)] %>%
+  .[, .(pp_acum = sum(pp_acum, na.rm = TRUE)), by = .(lon, lat, exp)] %>%
+  .[, c("x", "y") := wrf_project(lon, lat)]
+
+
+interpolate <- function(x, y, pp) {
+
+  data <- akima::interp(x, y, pp,
+                        xo = unique(pp_wrf$x),
+                        yo = unique(pp_wrf$y))
+  dimnames(data$z) <- list(x = data$x,
+                           y = data$y)
+  reshape2::melt(data$z)
+}
+
+
+files_imerg <- purrr::map(seq(0, acumulado-1, 1), function(l) {
+  list.files(imerg_path, pattern = format(ini_acum + hours(l), "%Y%m%d-S%H"), full.names = TRUE)
+}) %>% unlist()
+
+pp_imerg <- purrr::map(files_imerg, function(f) {
[…]

### 2020-12-04 analysis/chapters/experimento_20181122.Rmd
Mejoro fig de corte zonal

@@
   geom_contour(aes(z = v_mean, linetype = factor(-sign(..level..))), 
-               color = "grey30", size = 0.5, binwidth = 2) +
-  geom_text_contour(aes(z = v_mean), color = "grey30", stroke = 0.1, breaks = seq(-10, 6, 2)) +
+               color = "grey30", size = 0.3, binwidth = 2) +
+  geom_text_contour(aes(z = v_mean), color = "grey30", size = 4, stroke = 0.1, breaks = seq(-10, 6, 2)) +
   scale_linetype(guide = NULL) +

### 2020-12-04 analysis/scripts/acumulados_wrf.R
script acumulados WRF fcst

@@ -1,0 +1,153 @@
+# Calcula y guarda acumulados del pronóstico para todo el periodo
+
+library(metR)
+library(tidyverse)
+library(data.table)
+library(lubridate)
+library(unglue)
+library(mesoda)
+
+
+# Constante
+imerg_path <- "/home/paula.maldonado/datosalertar1/RRA_VERIF/data/raw/imerg_raw"
+wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
+exp <- "E4"
+run <- "fcst"
+
+ini_date <- ymd_hms("20181122060000")
+ciclos <- 37
+
+acumulado <- 6
+q <- c(5, 10, 25, 50)
+# q <- c(1, 5, 10, 25)
+
+# Inicio
+
+first <- ymd_hms("20181122060000")
+last <- ymd_hms("20181123120000")
+
+lead_time <- c(final = 30, inicial = 0)
+pp_wrf <- purrr::map(seq_along(lead_time), function(f) {
+
+
+  file_wrf <- paste0(wrf_path, exp, "/", toupper(run), "/",
+                     format(ini_date, "%Y%m%d%H"), "/NPP/NPP_",
+                     format(ini_date, "%Y"), "-",
+                     format(ini_date, "%m"), "-",
+                     format(ini_date, "%d"), "_",
+                     format(ini_date, "%H"), "_FC",
+                     formatC(lead_time[f], width = 2, flag = "0"),
+                     ".nc")
+
+
+  if (!file.exists(file_wrf)) {
+    return(NULL)
+  }
+
+  ReadNetCDF(file_wrf, vars = c("PP", "XLAT", "XLONG")) %>%
+    .[, ":="(lon = XLONG, lat = XLAT)] %>%
+    .[, ":="(XLONG = NULL, XLAT = NULL)] %>%
+    .[, ":="(exp = exp,
+             lead_time = names(lead_time)[f])]
+
+}) %>%
+  rbindlist() %>%
+  dcast(ens + lat + lon ~ lead_time, value.var = "PP") %>%
+  .[, pp_acum := final - inicial] %>%
+  .[, ":="(final = NULL,
+           inicial = NULL)] %>%
+  .[, c("x", "y") := wrf_project(lon, lat)]
+
+
+pp_wrf_mean <- pp_wrf[, .(pp_acum = mean(pp_acum)),
[…]

### 2020-12-16 analysis/chapters/experimento_20181122.Rmd
Figuras de RCRV para convencionales

@@
 output: 
+  bookdown::html_document2: default
   bookdown::pdf_document2: default
-  bookdown::html_document2: default
 bibliography: mesoda_ref.bib
@@
 
+scale_date <- function(ini, end, break_bin = "6 hour") {
+  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
+                   labels = function(x) {
+    fifelse(hour(x) == 0, format(x, "%H\n%b %d"), format(x, "%H"))
+  }) 
+}
+
+obs.type <- tribble(
+  ~type, ~code,
+  120, "ADPUPA",
+  130, "AIRCFT-AIREP",
+  131, "AIRCFT-AMDAR",
+  133, "AIRCAR",
+  180, "SFCSHP",
+  181, "ADPSFC",
+  187, "ADPSFC\nno pressure",
+  220, "ADPUPA",
+  221, "ADPUPA-PIBAL",
+  230, "AIRCFT-AIREP",
+  231, "AIRCFT-AMDAR",
+  233, "AIRCAR",
+  280, "SFCSHP",
+  281, "ADPSFC",
+  287, "ADPSFC\nno pressure",
+  290, "ASCATW"
+) %>% 
+  setDT()
+
 fisica <- data.table(ens = 1:60, 
@@
 
-```{r obs, eval=FALSE, include=FALSE}
+```{r obs}
 tribble(
   ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
-  "ADPSFC", "✔", "✔", "✔", "✔",
-  "ADPUPA", "✔", "✔", "✔", "✔",
-  "AIRCFT", "✔", "✔", "✔", "✔",
-  "AIRCAR", "✔", "✔", "✔", "✔",
-  "SFCSHP", "✔", "✔", "✔", "✔",
-  "ASCATW", "✔", "✔", "✔", "✔", 
-  "AUTSFC", " ", "✔", "✔", "✔",
-  "SATWND", " ", " ", "✔", "✔",
-  "RAD", " ", " ", " ", "✔",
+  "ADPSFC", "x", "x", "x", "x",
+  "ADPUPA", "x", "x", "x", "x",
+  "AIRCFT", "x", "x", "x", "x",
+  "AIRCAR", "x", "x", "x", "x",
+  "SFCSHP", "x", "x", "x", "x",
+  "ASCATW", "x", "x", "x", "x", 
+  "AUTSFC", " ", "x", "x", "x",
+  "SATWND", " ", " ", "x", "x",
+  "RAD", " ", " ", " ", "x",
 ) %>% 
@@
 
+```
+
+### Ensemble characteristics
+
+#### Reduced Centered Random Variable
+
+```{r rcrv-sfc, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree", fig.scap=c("Wind", "Temperature")}
+
+fread(here("analysis", "data", "derived_data", "uvt_rcrv_box.csv")) %>%   
+  .[var %in% c("u", "v") & type %in% c(280, 281, 287, 290)] %>% 
+  .[, type.obs := type] %>% 
+  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
+  geom_raster(aes(fill = sd.y)) +
+  geom_mapa() +
+  scale_fill_viridis_c(breaks = seq(0, 4, 0.5),
+                       limits = c(0, 4),
+                       direction = -1,
+                       guide = guide_colorstrip(inside = TRUE,
+                                                barwidth = 20,
+                                                barheight = 0.8)) +
+  facet_grid(var~type.obs, 
+             labeller = labeller(type.obs = c("280" = "SFCSHP",
+                                              "281" = "ADPSFC",
+                                              "287" = "ADPSFC\n no pressure",
+                                              "290" = "ASCATW"),
+                                 var = c("u" = "U wind",
+                                         "v" = "V wind" ))) +
+  labs(x = "",
+       y = "",
+       fill = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+fread(here("analysis", "data", "derived_data", "uvt_rcrv_box.csv")) %>%
+  .[var %in% c("t") & type %in% c(180, 181, 187)] %>% 
+  .[, type.obs := type] %>% 
+  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
+  geom_raster(aes(fill = sd.y)) +
+  geom_mapa() +
+  scale_fill_viridis_c(breaks = seq(0, 5, 0.5),
+                       limits = c(0, 5),
+                       direction = -1,
+                       guide = guide_colorstrip(inside = TRUE,
+                                                barwidth = 20,
+                                                barheight = 0.8)) +
+  facet_grid(var~type.obs, labeller = labeller(type.obs = c("180" = "SFCSHP",
+                                              "181" = "ADPSFC",
+                                              "187" = "ADPSFC\n no pressure"),
+                                              var = c("t" = "Temperature"))) +
+  labs(x = "",
+       y = "",
+       fill = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+
+```{r rcrv-profile, fig.cap="Profiles of mean and standar deviation of rcrv for sounding and aircraft observations."}
+obs.type[fread(here("analysis", "data", "derived_data", "rcrv_uvt_perfil.csv")), on = "type"] %>% 
+  .[type != 130] %>%
+  melt(measure.vars = c("mean.y", "sd.y")) %>% 
+  ggplot(aes(nivel.error, value)) +
+  geom_hline(yintercept = c(0, 1), color = "grey20") +
+  geom_line(aes(color = factor(code), linetype = variable)) +
+  scale_color_brewer(palette = "Set1") +
+  scale_linetype_manual(values = c("dashed", "solid"),
+                        labels = c("Mean RCRV", "SD RCRV")) +
+  scale_x_level() +
+  coord_flip() +
+  facet_wrap(~ var, labeller = labeller(var = c("t" = "Temperature",
+                                                "u" = "U wind",
+                                                "v" = "V wind"))) +
+  labs(y = NULL,
+       linetype = "",
+       color = NULL) +
+ theme_minimal() 
 ```
@@
 
-conv[type %in% c(181, 187) & exp %in% experiments, 
+conv[type %in% c(181, 187) & exp %in% c("E5", "E7") & usage.flag == 1, 
      .(mean_omb = mean(obs.guess, na.rm = TRUE),
@@
                                           E6 = "SATWND", E7 = "RAD")) +
-  facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
-                                                             "-1" = "rejected"))) +
+  # facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
+  #                                                            "-1" = "rejected"))) +
+scale_date(20181120180000, 20181123120000) +
   labs(y = "Mean of OmB (g/g)", x = NULL,
-       color = NULL, size = "n obs") +
+       color = NULL, size = "number\n of obs") +
   theme_minimal()
@@
                                           E6 = "SATWND", E7 = "RAD")) +
+  scale_date(20181121000000, 20181123000000, "12 hours") +
   facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
@@
                                             "51" = "500 km", "101" = "1000 km"))) +
-  labs(color = NULL, x = NULL) +
+  labs(color = NULL, x = NULL, y = "FSS") +
   theme_minimal()

### 2020-12-16 analysis/scripts/acumulados_wrf.R
Acumulados pp para analisis guarda un archivo por tiempo

@@ -62,7 +62,7 @@ pp_wrf_mean <- pp_wrf[, .(pp_acum = mean(pp_acum)),
 
 pp_prop <- purrr::map(q, function(f){
 
-    pp_wrf %>%
+  pp_wrf %>%
     .[, .(prop = mean(pp_acum > f)), by = .(lon, lat, x, y)] %>%
     .[, umbral := f]
 
@@ -111,7 +111,7 @@ pp_wrf <- purrr::map(lead_time, function(f) {
 
 }) %>%
   rbindlist() %>%
-.[, pp_acum := c(0, diff(PP)), by = .(ens, exp, lon, lat)] %>%
+  .[, pp_acum := c(0, diff(PP)), by = .(ens, exp, lon, lat)] %>%
   .[, PP := NULL] %>%
   .[, c("x", "y") := wrf_project(lon, lat)]
 
@@ -152,46 +152,46 @@ q <- c(1, 5, 10, 25, 50)
 dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
                     length.out = ciclos)
 
-fss_out <- purrr::map_dfr(dates, function(d) {
+pp_wrf_out <- purrr::map(dates, function(d) {
 
   date <- d
 
-files_wrf <-  purrr::map(seq(0, acumulado - 1), function(l) {
-  list.files(path = paste0(wrf_path, exp, "/", toupper(run), "/", format(date - hours(l), "%Y%m%d%H%M%S")),
-             full.names = TRUE,
-             recursive = TRUE,
-             pattern = "analysis.mem*")
-}) %>% unlist()
+  files_wrf <-  purrr::map(seq(0, acumulado - 1), function(l) {
+    list.files(path = paste0(wrf_path, exp, "/", toupper(run), "/", format(date - hours(l), "%Y%m%d%H%M%S")),
+               full.names = TRUE,
+               recursive = TRUE,
+               pattern = "analysis.mem*")
+  }) %>% unlist()
 
 
-pp_wrf <- purrr::map(files_wrf, function(f) {
+  pp_wrf <- purrr::map(files_wrf, function(f) {
 
-  metadatos <- unglue(basename(f), "analysis.mem0{mem}")
+    metadatos <- unglue(basename(f), "analysis.mem0{mem}")
 
-  ReadNetCDF(f, vars = c("RAINNC", "RAINC", "RAINSH",
-                         lon = "XLONG", lat = "XLAT")) %>%
[…]

### 2020-12-16 R/cut_round.R
minor changes

@@ -5,6 +5,7 @@
 #'
 #' @export
 cut_round <- function(x, breaks) {
+
   labels <- na.omit((breaks + data.table::shift(breaks, -1))/2)
   cuts <- cut(x, breaks = breaks, labels = labels)
 

### 2020-12-16 R/read_diag_rad.R

@@ -12,7 +12,7 @@ read_diag_rad <- function(file_list, exp) {
     if (file.size(file_list[f]) == 0) {
       return(NULL)
     }
-    meta <- unglue::unglue(basename(file_list[f]), "asim_{sensor}_{plat}_{date}.mem{mem}")
+    meta <- unglue::unglue(basename(file_list[f]), "asim_{sensor}_{plat}_{date}.{mem}")
     # print(f)
     out <- data.table::fread(file_list[f], na.strings = c("0.100000E+12"))
     # .[V10 == 1] %>%

### 2020-12-16 analysis/chapters/experimento_20181122.Rmd
OmB y # obs para radianzas

@@
 
-scale_date <- function(ini, end, break_bin = "6 hour") {
+scale_date <- function(ini = 20181120180000, end = 20181123120000,
+                       break_bin = "6 hour") {
   scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
@@
 
-```{r rad}
+```{r rad, fig.cap="Radiances for each 50 hPa pressure layer over time", fig.scap=c("OmB", "Number of observations")}
 
@@
   .[qc == 0 & iuse == 1] %>% 
-  .[, lev := cut_width(peakwt, 50)] %>% 
+  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
   .[, .(omb = mean(tbc),
               n = .N), by = .(date, lev)] %>% 
+  .[lev != 1025] %>% 
   ggplot(aes(date, fct_rev(lev))) +
-  geom_raster(aes(fill = omb)) +
+  geom_tile(aes(fill = omb)) +
   scale_fill_divergent(trans = mesoda::symlog_trans(),
-                       guide = guide_colourbar(barheight = 20))  +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H:%M") +
-  labs(x = NULL, fill = NULL, y = "pressure layer")
+                       guide = guide_colourbar(barheight = 18))  +
+   # scale_y_level() +
+  scale_date() +
+  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H:%M") +
+  labs(x = NULL, fill = NULL, y = "pressure (hPa)",
+       caption = "OmB") +
+  theme_minimal()
 
@@
   .[qc == 0 & iuse == 1] %>% 
-  .[, lev := cut_width(peakwt, 50)] %>% 
+  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
   .[, .(omb = mean(tbc),
               n = .N), by = .(date, lev)] %>% 
+  .[lev != 1025] %>% 
   ggplot(aes(date, fct_rev(lev))) +
-  geom_raster(aes(fill = n)) +
+  geom_tile(aes(fill = n)) +
   scale_fill_viridis_c(trans = scales::log10_trans(),
-                       guide = guide_colourbar(barheight = 20)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H:%M") +
-  labs(x = NULL, fill = NULL, y = "pressure layer")
+                       guide = guide_colourbar(barheight = 18)) +
+
+  scale_date() +
+  labs(x = NULL, fill = NULL, y = "pressure (hPa)",
+       caption = "Number of observations") +
+  theme_minimal()
 
@@
   geom_line(aes(color = factor(code), linetype = variable)) +
-  scale_color_brewer(palette = "Set1") +
   scale_linetype_manual(values = c("dashed", "solid"),
@@
  theme_minimal() 
+```
+
+```{r rcrv-profile-rad, fig.cap="Profiles of mean and standar deviation of rcrv for radiances observations."}
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:4]
+
+RCRV <- purrr::map(files, function(f) {
+  fread(f)
+}) %>% 
+  rbindlist()
+
+RCRV %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  setDT() %>% 
+  melt(measure.vars = c("mean.y", "sd.y")) %>% 
+  .[, .(value = mean(value, na.rm = TRUE)), by = .(sensor, level, variable)] %>% 
+  ggplot(aes(level, value)) +
+  geom_hline(yintercept = c(0, 1), color = "grey20") +
+  geom_line(aes(color = sensor, linetype = variable)) +
+  scale_linetype_manual(values = c("dashed", "solid"),
+                         labels = c("Mean RCRV", "SD RCRV")) +
+  coord_flip() +
+  scale_x_level() +
+  labs(x = NULL, y = NULL,
+       color = NULL, linetype = NULL) +
+  theme_minimal()
+
 ```

### 2020-12-17 analysis/chapters/experimento_20181122.Rmd
Hovmoller de pp acum

@@
 ```
+```{r rad-wv}
+files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121020000/diagfiles/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121140000/diagfiles/asim_iasi*ensmean"))
+
+rad <- read_diag_rad(files, "E7") %>% 
+  .[, wavenumber := 1/(299792458/(freq*10000000))] %>% 
+  .[wavenumber %between% c(1367, 1422.25)] %>% 
+  satinfo[., on = c("sensor", "channel")]
+
+rad %>% 
+  .[sensor == "iasi_metop-b" & 
+      channel %between% c(330, 335) & 
+      qc == 0 & iuse == 1] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = tbc)) +
+  scale_color_divergent(trans = mesoda::symlog_trans(),
+                        guide = guide_colourbar(barheight = 18)) +
+  geom_mapa() +
+  facet_wrap(~date) +
+  labs(x = NULL, y = NULL, color = NULL,
+       caption = "1410.75 cm^-1") +
+  theme_minimal()
+```
 
@@
 
+### Verification
+
+```{r pp-hov, fig.cap="Hövmoller of hourly acumulated precipitation (for ensemble mean) over the inner domain."}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*mean*")
+
+ppacum_1h <- purrr::map(files, function(f) {
+
+  readRDS(f)
+
+}) %>% rbindlist() %>% 
+  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)]
+
+readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
+  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
+  .[end_date %between% c(as_datetime("20181120190000"), as_datetime("20181123120000"))] %>% 
+    .[, .(pp_acum = mean(pp_acum)), by = .(end_date, y)] %>% 
+  ggplot(aes(end_date, y/100000)) +
+  geom_contour(aes(z = pp_acum, color = ..level..), 
+               breaks = seq(0, 12, 0.5)) +
+  scale_color_viridis_c(guide = NULL, breaks = seq(0, 12, 0.5)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  labs(x = NULL, y = NULL, color = "PP acum", 
+       subtitle = "IMERG") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  ppacum_1h %>% 
+  .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y)] %>% 
+  ggplot(aes(date, y/100000)) +
+  geom_contour(aes(z = pp_acum, color = ..level..), 
+               breaks = seq(0, 12, 0.5)) +
+  scale_color_viridis_c(breaks = seq(0, 12, 1), limits = c(0, 12),
+                        guide = guide_colorstrip(barwidth = 25,
+                                                 barheight = 1)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_wrap(~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                E6 = "SATWND", E7 = "RAD"))) +
+  labs(x = NULL, y = NULL, color = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+ plot_layout(ncol = 2, widths = c(0.4, 0.6), guides = "collect") & theme(legend.position = 'bottom')
+```
+
+```{r}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*prop*")
+
+ppacum_1h <- purrr::map(files, function(f) {
+
+  readRDS(f)
+
+}) %>% rbindlist() %>% 
+  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)]
+
+ppacum_1h %>% 
+.[, .(prop = mean(prop)), by = .(exp, date, y, umbral)] %>% 
+  .[umbral %in% c(1, 5)] %>% 
+ggplot(aes(date, y)) +
+  geom_contour_fill(aes(z = prop*100, fill = stat(level)),
+                    breaks = seq(0, 100, 10)) +
+  scale_color_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
+                       limits = c(10, 100),
+                       na.value = "white",
+                       super = ScaleDiscretised,
+                       breaks = seq(0, 100, 10),
+                       guide = guide_colorstrip(barwidth = 15,
+                                                barheight = 1)) +
+  # scale_date() +
+  labs(fill = "", x = "", y = "") +
+  facet_grid(umbral~exp) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
 #### FSS for the analysis

### 2020-12-17 R/scale_color_dbz.R
scala de colores para reflectividad

@@ -1,0 +1,37 @@
+#' Reflectivity color pallete
+#'
+#' @param colour
+#' @param limits
+#' @param guide
+#' @param ...
+#'
+#' @export
+scale_color_dbz <- function(colours = hex,
+                              limits = c(0, 60),
+
+                              guide = guide_colorbar(barheight = 15),
+                              ...) {
+
+  colours <- rgb(ref$R, ref$G, ref$B, maxColorValue = 1)
+
+  scale_color_gradientn(colours = colours,
+                        limits = limits,
+                        guide = guide, ...)
+}
+
+ref <- dplyr::tribble(
+  ~R, ~G, ~B,
+   1.0/255.0, 159.0/255.0, 244.0/255.0,
+   3.0/255.0,   0.0/255.0, 244.0/255.0,
+   2.0/255.0, 253.0/255.0,   2.0/255.0,
+   1.0/255.0, 197.0/255.0,   1.0/255.0,
+   0.0/255.0, 142.0/255.0,   0.0/255.0,
+ 253.0/255.0, 248.0/255.0,   2.0/255.0,
+ 229.0/255.0, 188.0/255.0,   0.0/255.0,
+ 253.0/255.0, 149.0/255.0,   0.0/255.0,
+ 253.0/255.0,   0.0/255.0,   0.0/255.0,
+# 212.0/255.0,   0.0/255.0,   0.0/255.0,
+ 188.0/255.0,   0.0/255.0,   0.0/255.0,
+ 248.0/255.0,   0.0/255.0, 253.0/255.0,
+ 152.0/255.0,  84.0/255.0, 198.0/255.0
+)

### 2020-12-17 analysis/chapters/experimento_20181122.Rmd
Figura reflectividad

@@
   ggplot(aes(date, y)) +
-  geom_contour_fill(aes(z = prop*100, fill = stat(level)),
+  geom_contour_fill(aes(z = prop*100, fill = stat(level_d)),
                     breaks = seq(0, 100, 10)) +
-  scale_color_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")), 
+  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")),
                        limits = c(10, 100),
@@
                        super = ScaleDiscretised,
-                       breaks = seq(0, 100, 10),
-                       guide = guide_colorstrip(barwidth = 15,
+                       # breaks = seq(0, 100, 10),
+                       guide = guide_colorsteps(barwidth = 15,
                                                 barheight = 1)) +
-  # scale_date() +
+  # scale_color_viridis_c() +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_grid(umbral ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                       E6 = "SATWND", E7 = "RAD"),
+                                               umbral = c("1" = "1 mm/h",
+                                                          "5" = "5 mm/h"))) +
   labs(fill = "", x = "", y = "") +
-  facet_grid(umbral~exp) +
   theme_minimal() +
   theme(legend.position = "bottom")
+```
+
+#### Reflectivity
+
+```{r}
+
+dates <- rep(c("20181122100000", "20181122220000"), 1, each = 4)
+files <- paste0(derived_data, "dbz/maxdbz_ana_", experiments, "_", dates, ".nc")
+
+dbz <- purrr::map(files, function(f){
+
+  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
+  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             date = ymd_hms(meta[[1]][["date"]]))] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] 
+}) %>% 
+  rbindlist()
+
+dbz %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
+                    proj = norargentina_lambert,
+                    breaks = seq(0, 60, 5),) +
+  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
+                  guide = guide_colorsteps(barheight = 18)) +
+  geom_mapa() +
+  facet_grid(date~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                   E6 = "SATWND", E7 = "RAD"))) +
+  labs(x = NULL, y = NULL, fill = NULL) +
+  theme_minimal()
 ```

### 2020-12-18 analysis/chapters/experimento_20181122.Rmd
Post reunión 18/12

@@
 # colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
-
+satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
+  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
 derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"
@@
 
-satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
-  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
-
 rad <- satinfo[rad, on = c("sensor", "channel")]
@@
 ```{r rad-wv}
-files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121020000/diagfiles/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121140000/diagfiles/asim_iasi*ensmean"))
-
-rad <- read_diag_rad(files, "E7") %>% 
+files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121020000/diagfiles*/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121140000/diagfiles*/asim_iasi_metop-b*ensmean"))
+
+rad <- purrr::map(files, function(f) {
+
+  rad <- read_diag_rad(f, "E7") %>% 
+    .[, run := basename(dirname(f))] %>% 
     .[, wavenumber := 1/(299792458/(freq*10000000))] %>% 
@@
     satinfo[., on = c("sensor", "channel")]
+}) %>% rbindlist()
 
@@
   .[sensor == "iasi_metop-b" & 
-      channel %between% c(330, 335) & 
+      channel == 335 & 
       qc == 0 & iuse == 1] %>% 
@@
   scale_color_divergent(trans = mesoda::symlog_trans(),
-                        guide = guide_colourbar(barheight = 18)) +
+                        guide = guide_colourbar(barheight = 15, barwidth = 1)) +
   geom_mapa() +
-  facet_wrap(~date) +
-  labs(x = NULL, y = NULL, color = NULL,
-       caption = "1410.75 cm^-1") +
-  theme_minimal()
+  facet_grid(run~date, labeller = labeller(run = c("diagfiles" = "OmB",
+                                                   "diagfiles_ana" = "OmA"))) +
+  labs(x = NULL, y = NULL, color = NULL) +
+  theme_minimal() +
+
+  rad %>% 
+  .[sensor == "iasi_metop-b" & 
+      channel == 335 & 
+      # qc == 0 & 
+      iuse == 1] %>% 
+  dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 
+  .[, ":="(tbc = diagfiles - diagfiles_ana,
+           run = "AmB")] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = tbc)) +
+  scale_color_divergent(guide = guide_colourbar(barheight = 15, barwidth = 1)) +
+  geom_mapa() +
+  labs(x = NULL, y = NULL, color = NULL) +
+  facet_grid(run~date) +
+  theme_minimal() +
+  plot_layout(ncol = 2, widths = c(0.4, 0.6))
 ```
@@
 
-```{r}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/ANA/sondeos/*")
-
-sondeos <- purrr::map(files, function(f){
-
-  fread(f) %>% 
-    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
-    .[, launch_time := as_datetime(launch_time)]
-
+```{r temp}
+fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
+                     exp = c("E4", "E5", "E6", "E7"))
+
+t <- lapply(seq_len(nrow(fcsts)), function(f) {
+
+  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
+    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
+    .[, t := ReadNetCDF(paste0(derived_data, "/t_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(t = "temp"), out = "vector")] %>% 
+    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
+    .[, exp := fcsts[f, 2]] %>%  
+    .[]
 }) %>% 
@@
 
-sondeos %>% 
-  .[variable == "t" & !str_detect(site, "/") & 
-      exp == "E4"] %>% 
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = value), size = 0.5) +
-  scale_color_distiller(palette = "YlOrRd", direction = 1) +
-  # geom_contour(aes(z = value - fcst_value)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  labs(color = "T", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  sondeos %>% 
-  .[variable == "t" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
-  .[, oma := value - fcst_value] %>% 
-  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = oma), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
-  labs(color = "OmA", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(0.3, 0.7))
-```
-
-```{r}
-sondeos %>% 
-  .[variable == "td" & !str_detect(site, "/") & 
-      exp == "E4"] %>% 
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = value), size = 0.5) +
-  scale_color_distiller(palette = "YlOrRd", direction = 1) +
-  # geom_contour(aes(z = value - fcst_value)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  labs(color = "Td", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  sondeos %>% 
-  .[variable == "td" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
-  .[, oma := value - fcst_value] %>% 
-  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = oma), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
-  labs(color = "OmA", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(0.3, 0.7))
-```
-
-```{r}
-sondeos %>% 
-  .[variable == "v" & !str_detect(site, "/") & 
-      exp == "E4"] %>% 
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = value), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  labs(color = "V (m/s)", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  sondeos %>% 
-  .[variable == "v" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
-  .[, oma := value - fcst_value] %>% 
-  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = oma), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
-  labs(color = "OmA", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(0.3, 0.7))
+t[date == unique(date)[1:3] & west_east %between% c(70, 180)] %>% 
+  .[, .(t_mean = mean(t),
+        u_mean = mean(u),
+        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
+  .[, ":="(dir = atan2(v_mean, u_mean)*180/pi,
+           mag = Mag(u_mean, v_mean))] %>%
+  .[, date := factor(date)] %>% 
+  ggplot(aes(south_north, bottom_top)) +
+  geom_contour_fill(aes(z = t_mean, fill = stat(level_d))) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "OrRd", direction = 1,
+                       # limits = c(1, 19),
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  # geom_contour(aes(z = v_mean, linetype = factor(-sign(..level..))), 
+  #              color = "grey30", size = 0.3, binwidth = 2) +
+  # geom_text_contour(aes(z = v_mean), color = "grey30", size = 4, stroke = 0.1, breaks = seq(-10, 6, 2)) +
+  # scale_linetype(guide = NULL) +
+  geom_arrow(aes(mag = mag, angle = dir),
+             arrow.angle = 8, size = 0.2, skip.x = 10, skip.y = 3) +
+  scale_mag() +
+  facet_grid(date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                     E6 = "SATWND", E7 = "RAD"))) +
+  labs(y =  "Sigma levels", x = "Grid points") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
 ```
@@
 ```
-
-#### FSS for the forecast
+#### Soundings
+
+```{r message=FALSE, warning=FALSE}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/ANA/sondeos/*")
+
+sondeos <- purrr::map(files, function(f){
+
+  fread(f) %>% 
+    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
+    .[, launch_time := as_datetime(launch_time)]
+
+}) %>% 
+  rbindlist()
+
+IOP <- tribble(
+  ~iop, ~ini, ~end,
+  "IOP07", ymd_hms("20181121150000"), ymd_hms("20181121210000"),
+  "IOP08", ymd_hms("20181122140000"), ymd_hms("20181122200000")
+) %>% setDT()
+
+sondeos %>% 
+  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
+                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
+  .[variable == "t" & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
+  .[, oma := value - fcst_value] %>% 
+  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = oma), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  facet_grid(exp ~ iop, scales = "free_x", labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                                       E6 = "SATWND", E7 = "RAD"))) +
+  labs(color = "OmA", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  sondeos %>% 
+  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
+                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
+  .[variable == "v" & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
+  .[, oma := value - fcst_value] %>% 
+  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = oma), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  facet_grid(exp ~ iop, scales = "free_x", labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                                       E6 = "SATWND", E7 = "RAD"))) +
+  labs(color = "OmA", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(0.5, 0.5))
+```
+
+```{r}
+
+sondeos %>% 
+  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
+                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
+  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
+  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
+  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2)), 
+           BIAS = mean(value - fcst_value)), by = .(iop, exp, variable, lev)] %>% 
+  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
+  ggplot(aes(lev/1000, value.est)) +
+  geom_line(aes(color = exp, linetype = estadistico)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(iop~variable, scales = "free_x") +
+  labs(x = NULL, y = NULL, color = NULL, linetype = NULL) +
+  theme_minimal()
+```
+
+```{r}
+sondeos %>% 
+  .[variable == "td" & !str_detect(site, "/") & 
+      exp == "E4"] %>%, 
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = value), size = 0.5) +
+  scale_color_distiller(palette = "YlOrRd", direction = 1) +
+  # geom_contour(aes(z = value - fcst_value)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  labs(color = "Td", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  sondeos %>% 
+  .[variable == "td" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
+  .[, oma := value - fcst_value] %>% 
+  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = oma), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E7 = "RAD"))) +
+  labs(color = "OmA", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(0.3, 0.7))
+```
+
+```{r}
+sondeos %>% 
+  .[variable == "v" & !str_detect(site, "/") & 
+      exp == "E4"] %>% 
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = value), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  labs(color = "V (m/s)", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  sondeos %>% 
+  .[variable == "v" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
+  .[, oma := value - fcst_value] %>% 
+  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
+  ggplot(aes(launch_time, alt/1000)) +
+  geom_point(aes(color = oma), size = 0.5) +
+  scale_color_divergent() +
+  # geom_contour(aes(z = value - fcst_value)) +
+  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
+  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E7 = "RAD"))) +
+  labs(color = "OmA", y = NULL, x = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(0.3, 0.7))
+```
+
+### Forecast
 

### 2021-01-07 analysis/chapters/experimento_20181122.Rmd
Impacto de las obs?

@@
       channel == 335 & 
-      qc == 0 & iuse == 1] %>% 
+      errinv %between% c(0.0000316227, 1000) & 
+      qc == 0 & 
+      iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
   .[, date := factor(date)] %>% 
@@
       channel == 335 & 
-      # qc == 0 & 
-      iuse == 1] %>% 
+      errinv %between% c(0.0000316227, 1000) & 
+      qc == 0 & 
+      iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
+  dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 
+  .[, ":="(tbc = diagfiles - diagfiles_ana,
+           run = "AmB")] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = tbc)) +
+  scale_color_divergent(guide = guide_colourbar(barheight = 15, barwidth = 1)) +
+  geom_mapa() +
+  labs(x = NULL, y = NULL, color = NULL) +
+  facet_grid(run~date) +
+  theme_minimal() +
+  plot_layout(ncol = 2, widths = c(0.4, 0.6))
+```
+```{r}
+files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121010000/diagfiles*/asim_*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121130000/diagfiles*/asim_*ensmean"))
+
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121130000/diagfiles*/asim_*ensmean")
+
+files <- files[str_detect(files, "conv", negate = TRUE)]
+
+rad <- purrr::map(files, function(f) {
+
+  rad <- read_diag_rad(f, "E7") %>% 
+    .[, run := basename(dirname(f))] %>% 
+    .[, wavenumber := 1/(299792458/(freq*10000000))] %>% 
+    # .[wavenumber %between% c(1367, 1422.25)] %>% 
+    satinfo[., on = c("sensor", "channel")]
+}) %>% rbindlist()
+
+rad %>% 
+  .[run == "diagfiles"] %>% 
+  .[errinv %between% c(0.0000316227, 1000) & 
+      qc == 0 & 
+      iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = sensor, shape = sensor), size = 0.8) +
+  # scale_color_divergent(trans = mesoda::symlog_trans(),
+  #                       guide = guide_colourbar(barheight = 15, barwidth = 1)) +
+  scale_color_brewer(palette = "Dark2") +
+  geom_mapa() +
+  # facet_grid(run~date, labeller = labeller(run = c("diagfiles" = "OmB",
+  #                                                  "diagfiles_ana" = "OmA"))) +
+  labs(x = NULL, y = NULL, color = NULL, shape = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+  rad %>% 
+  .[sensor == "iasi_metop-b" & 
+      channel == 335 & 
+      errinv %between% c(0.0000316227, 1000) & 
+      qc == 0 & 
+      iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
   dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 

### 2021-01-07 analysis/scripts/acum_update.R
Acumulados

@@ -1,0 +1,86 @@
+exp <- "E7"
+ini_date <- ymd_hms("20181120180000")
+ciclos <- 50
+
+dates <- seq.POSIXt(ini_date, by = "hour",
+                    length.out = ciclos)
+
+
+
+
+for (d in seq_along(dates)) {
+
+  print(d)
+
+  path_ana <- paste0("~/datosmunin/EXP/", exp, "/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.ensmean")
+  path_gue <- paste0("~/datosmunin/EXP/", exp, "/GUESS/", format(dates[d], "%Y%m%d%H%M%S"), "/wrfarw.ensmean")
+
+  update <- rbind(ReadNetCDF(path_ana,
+                             vars = c(p = "P", "PB", qv = "QVAPOR",
+                                      lon = "XLONG", lat = "XLAT")) %>%
+                    .[, p := p + PB] %>%
+                    .[, ":="(Time = NULL,
+                             west_east = NULL,
+                             south_north = NULL,
+                             PB = NULL,
+                             run = "ana")],
+
+                  ReadNetCDF(path_gue,
+                             vars = c(p = "P", "PB", qv = "QVAPOR",
+                                      lon = "XLONG", lat = "XLAT")) %>%
+                    .[, p := p + PB] %>%
+                    .[, ":="(Time = NULL,
+                             west_east = NULL,
+                             south_north = NULL,
+                             PB = NULL,
+                             run = "guess")]) %>%
+    # .[, ":="(p_lev = cut_round(p/100, breaks = seq(0, 1050, 50)),
+    .[, ":="(exp = exp,
+             date = dates[d])] %>%
+    .[, .(qv = mean(qv)), by = .(lon, lat, bottom_top, run, exp, date)] %>%
+    dcast(bottom_top + lon + lat + exp + date ~ run, value.var = "qv") %>%
+    .[, ":="(qv = ana - guess)] %>%
+    .[, c("x", "y") := wrf_project(lon, lat)]
+
+  mean_qv <- update[, .(qv = mean(qv)), by = .(exp, date, bottom_top)]
+
+  if (d == 1) {
+    global_qv <- mean_qv
+    # espacial_qv <- update
+  } else {
+    global_qv <- rbind(global_qv, mean_qv)
[…]

### 2021-01-11 analysis/chapters/experimento_20181122.Rmd
Update graficos

--- 2020-12-18 analysis/chapters/experimento_20181122.Rmd
@@
 
-fisica <- data.table(ens = 1:60, 
+fisica <- data.table(mem = as.character(formatC(1:60, flag = "0", width = 3)),
                      fisica = rep(c("KF-YSU", 
@@
 
-```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots).", fig.height=4}
+```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots)."}
 obsAUT <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
@@
   .[, source := "Sfc - Non-official"]
-
-rbindlist(list(oficiales, no_oficiales, ascat, sondeos, aircraf, satwnd)) %>% 
-  ggplot(aes(ConvertLongitude(lon), lat)) +
-  geom_mapa() +
-  geom_point(alpha = 1, size = 0.5,
-             aes(color = source, shape = source)) +
-  scale_color_brewer(palette = "Dark2") +
-  # khroma::scale_color_vibrant() +
-  geom_point(data = square, aes(lon, lat), size = 0.2) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(x = NULL, y = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom")
 
@@
                      breaks = c("Sfc - Non-official", "Sfc - Official"),
-                     labels = c("Non-official (N = 820)","Official (N = 327)"))+
+                     labels = c("Non-official (N = 820)","Official (N = 327)"),
+                     guide = guide_legend(override.aes = list(size = 2)))+
   geom_point(data = square, aes(lon, lat), size = 0.2) +
@@
   theme_minimal() + 
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") +
+
+  plot_spacer() +
+
+  rbindlist(list(ascat, sondeos, aircraf, satwnd)) %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_mapa() +
+  geom_point(alpha = 1,
+             aes(color = source, shape = source, size = source)) +
+  scale_color_manual(values = c("darkorange", "purple", "cyan4", "grey70"),
+                     guide = guide_legend(override.aes = list(size = 2))) +
+  scale_size_manual(values = c(0.7, 0.5, 1.5, 0.1)) +
+  # scale_color_brewer(palette = "Dark2") +
+  # khroma::scale_color_vibrant() +
+  geom_point(data = square, aes(lon, lat), size = 0.2) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(x = NULL, y = NULL, color = NULL, shape = NULL, size = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 3, widths = c(0.47, 0.06, 0.47))
 ```
@@
       channel == 335 & 
-      # qc == 0 & 
-      iuse == 1] %>% 
+      errinv %between% c(0.0000316227, 1000) & 
+      qc == 0 & 
+      iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
   dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 
@@
   plot_layout(ncol = 2, widths = c(0.4, 0.6))
+```
+```{r omb-qv-acum}
+
+global_qv_E6 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E6.rds")
+global_qv_E7 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E7.rds")
+
+rbind(global_qv_E6, global_qv_E7) %>%
+  # .[exp == "E7"] %>% 
+  .[, nivel := fcase(bottom_top %in% c(1:8), "1 to 8",
+                     bottom_top %in% c(9:13), "9 to 13",
+                     bottom_top %in% c(14:20), "14 to 20",
+                     bottom_top %in% c(21:38), "21 to 38")] %>%
+  .[, acum := cumsum(qv), by = .(bottom_top, exp)] %>%
+  # .[bottom_top %in% c(1:18)] %>%
+  ggplot(aes(date, acum*1000)) +
+  geom_line(aes(color = exp, linetype = exp, group = interaction(bottom_top, exp))) +
+  # geom_label(data = ~.x[date %in% sample(unique(date), 38)] %>% group_by(exp, bottom_top) %>% sample_n(size = 1),
+   geom_label(data = ~.x[date %in% unique(date)[67]],
+             aes(label = bottom_top, color = exp, 
+                 group = interaction(bottom_top, exp)),
+             size = 3, fill = "white", show.legend = FALSE,
+             label.size = NA, label.padding = unit(0.1, "lines")) +
+  scale_date() +
+  scale_linetype_manual(values = c(2, 1), labels = c(E4 = "CONV", E5 = "AUT",
+                                     E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  facet_wrap(~nivel, scales = "free_y") +
+  labs(x = NULL, y = "Accumulated specific humidity (g/Kg)",
+       color = NULL, linetype = NULL) +
+  theme_minimal()
+
+```
+
+```{r}
+files <- Sys.glob("~/datosmunin/EXP/E7/ANA/2018112*/diagfiles/asim*.ensmean")
+
+files <- files[str_detect(files, "conv", negate = TRUE)]
+
+rad <- read_diag_rad(files, "E7") %>% 
+  satinfo[., on = c("sensor", "channel")] %>% 
+  .[,  type :=  fifelse(sensor %in% multiespectrales, "multiespectral", "otro")] %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_")
+
+rad[errinv %between% c(0.0000316227, 1000) & 
+      qc == 0 & 
+      iuse == 1 &
+      peakwt %between% c(0.001, 1200), 
+    .(tbc_mean = mean(tbc),
+      tbcnob_mean = mean(tbcnob)), by = .(sensor, channel, date)] %>% 
+  .[, diff := abs(tbcnob_mean) - abs(tbc_mean)] %>% 
+  # .[, range(diff, na.rm = TRUE)]
+  melt(id.vars = c("sensor", "channel", "date")) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
+                 alpha= 0.3, position = "identity") +
+  geom_density(aes(color = variable, y = stat(density))) +
+  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
+  scale_fill_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
+  scale_color_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
+  geom_vline(xintercept = 0, color = "grey40") +
+  coord_cartesian(xlim = c(-5, 5)) +
+  facet_wrap(~sensor) +
+  labs(fill = NULL, x = NULL, color = NULL) +
+  theme_minimal()
+
+rad[errinv %between% c(0.0000316227, 1000) & 
+      qc == 0 & 
+      iuse == 1 &
+      peakwt %between% c(0.001, 1200), 
+    .(tbc_mean = mean(tbc),
+      tbcnob_mean = mean(tbcnob)), by = .(sensor, channel, date)] %>% 
+  .[, diff := abs(tbcnob_mean) - abs(tbc_mean)] %>% 
+  # .[, range(diff, na.rm = TRUE)]
+  melt(id.vars = c("sensor", "channel", "date")) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
+                 alpha= 0.3, position = "identity") +
+  geom_density(aes(color = variable)) +
+  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
+  scale_color_viridis_d(labels = c("OmB BC", "OmB NoBC", "Difference")) +
+  scale_fill_viridis_d(labels = c("OmB BC", "OmB NoBC", "Difference")) +
+  geom_vline(xintercept = 0, color = "grey40") +
+  coord_cartesian(xlim = c(-5, 5)) +
+  # facet_wrap(~sensor) +
+  labs(fill = NULL, x = NULL, color = NULL) +
+  theme_minimal()
+
 ```
@@
 
+```{r omb-q-evol, fig.cap="Evolution of the mean OmB for humidity throughout the periodod"}
+files <- c(Sys.glob("~/datosmunin/EXP/E5/ANA/*/diagfiles/asim_conv_201811*.mem*"),
+           Sys.glob("~/datosmunin/EXP/E7/ANA/*/diagfiles/asim_conv_201811*.mem*"))
+
+conv <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.mem{mem}")
+  read_diag_conv(f, exp = meta[[1]][["exp"]], member = meta[[1]][["mem"]], variable = c("q")) %>% 
+    .[stationID %in% oficiales$stationID] %>% 
+    .[]
+}) %>% rbindlist() %>% 
+  fisica[., on = "mem"] %>% 
+  separate(fisica,  into = c("conv", "pbl"))
+
+conv[type %in% c(181, 187) & exp %in% c("E5", "E7") & usage.flag == 1, 
+     .(mean_omb = mean(obs.guess, na.rm = TRUE),
+       n = .N), 
+     by = .(date, exp, pbl, usage.flag)] %>% 
+  ggplot(aes(date, mean_omb)) +
+  geom_line(aes(color = exp, linetype = pbl)) +
+  geom_point(aes(color = exp, size = n/60)) +
+  scale_size_area(max_size = 2) +
+  scale_linetype_manual(values = c(2, 1, 4)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  # facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
+  #                                                            "-1" = "rejected"))) +
+  scale_date(20181120180000, 20181123120000) +
+  labs(y = "Mean of OmB (g/g)", x = NULL, linetype = NULL,
+       color = NULL, size = "number\n of obs") +
+  theme_minimal()
+```
+
 The negative difference between RAD and SATWND in low levels (\@ref(fig:Q-diff)) and positive values over those levels can be explain by a deeper boundary layer.
@@
 
-```{r}
+```{r dbz-mean}
 
@@
 
+```{r dbz-ens}
+
+files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[4-7]_0*"))
+files <- files[str_detect(files, "20181122000000", negate = TRUE)]
+
+dbz <- purrr::map(files, function(f){
+
+  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{mem}_{date}.nc")
+  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             mem = meta[[1]][["mem"]],
+             date = ymd_hms(meta[[1]][["date"]]))] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] 
+}) %>% 
+  rbindlist() %>% 
+  fisica[., on = "mem"] %>% 
+  separate(fisica, into = c("cumulus", "pbl"))
+
+dbz %>% 
+  .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
+  .[date == unique(date)[1]] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
+                    proj = norargentina_lambert,
+                    breaks = seq(0, 60, 5),) +
+  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
+                  guide = guide_colorsteps(barheight = 18)) +
+  geom_mapa() +
+  facet_grid(cumulus~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                   E6 = "SATWND", E7 = "RAD"))) +
+  labs(x = NULL, y = NULL, fill = NULL) +
+  theme_minimal() +
+
+  dbz %>% 
+  .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
+  .[date == unique(date)[2]] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
+                    proj = norargentina_lambert,
+                    breaks = seq(0, 60, 5),) +
+  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
+                  guide = guide_colorsteps(barheight = 18)) +
+  geom_mapa() +
+  facet_grid(cumulus~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                   E6 = "SATWND", E7 = "RAD"))) +
+  labs(x = NULL, y = NULL, fill = NULL) +
+  theme_minimal() +
+  plot_layout(ncol = 1)
+```
 #### FSS for the analysis
@@
 ```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_E[4-7]new.csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-7]new.csv")
 

### 2021-01-11 analysis/scripts/acumulados_wrf.R
acumulados y fss listos!

@@ -139,7 +139,7 @@ saveRDS(pp_wrf_mean, paste0(exp, "_fcst_", format(ini_date, "%Y%m%d%H"), "_mean_
 
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
 wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E6"
+exp <- "E7"
 run <- "ana"
 
 ini_date <- ymd_hms("20181120180000")

### 2021-01-11 analysis/scripts/wrap_fss.R

@@ -10,7 +10,7 @@ library(mesoda)
 #future::plan("cluster", workers = 3)
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
 wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E5"
+exp <- "E7"
 run <- "ana_ens"
 
 ini_date <- ymd_hms("20181120180000")

### 2021-01-14 analysis/chapters/experimento_20181122.Rmd
Figuras listas, ponele.

@@
 geom_mapa <- function(fill = NA) {
-  list(geom_sf(data = map, fill = fill, color = "black", size = 0.2, inherit.aes = FALSE),
+  list(geom_sf(data = map, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
        coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)))
@@
 
+multiespectrales <- c("airs_aqua",
+              "iasi_metop-a",
+              "iasi_metop-b")
+
 experiments <- c("E4", "E5", "E6", "E7")
@@
 ```{r fisicas}
-fisica[, .(mem = paste(ens, collapse = ", ")), by = .(fisica)] %>% 
+fisica[, .(mem = paste(mem, collapse = ", ")), by = .(fisica)] %>% 
   kbl(col.names = c("Parameterizations", "# Member"),
@@
 
-```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots)."}
+```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots).", fig.height=4, fig.width=8}
 obsAUT <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
@@
   theme_minimal() + 
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom", legend.justification = "left",
+         legend.key.size = unit(0.8, "lines")) +
 
@@
   theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 3, widths = c(0.47, 0.06, 0.47))
+  theme(legend.position = "bottom", legend.justification = "left",
+        legend.key.size = unit(0.8, "lines")) +
+
+  plot_layout(ncol = 3, widths = c(0.49, 0.02, 0.49))
 ```
@@
 
-rad <- read_diag_rad(files, "E7")
-
-rad <- satinfo[rad, on = c("sensor", "channel")]
+rad <- read_diag_rad(files, "E7") %>% 
+  satinfo[., on = c("sensor", "channel")]
 
@@
 ```
+
+```{r level-rad, fig.cap="Nivel de presión asociado a las observaciones en cada ciclo de asimilación"}
+rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[sensor %in% c("airs", "iasi") &
+      # errinv %between% c(0.0000316227, 1000) & 
+      # qc == 0 & 
+      # iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
+  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
+  ggplot(aes(channel, Q50)) +
+   # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
+   #             outlier.size = 0.5, size = 0.1) +
+  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
+  geom_line(color = "darkorange") +
+  scale_y_level(name = "Pressure (hPa)") +
+  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
+  labs(x = "", y = "") +
+  theme_minimal() +
+
+rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[!sensor %in% c("airs", "iasi") &
+      # errinv %between% c(0.0000316227, 1000) & 
+      # qc == 0 & 
+      # iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
+  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
+  ggplot(aes(channel, Q50)) +
+   # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
+   #             outlier.size = 0.5, size = 0.1) +
+  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
+  geom_line(color = "darkorange") +
+  scale_y_level(name = "Pressure (hPa)") +
+  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
+  labs(y = "", x = "Channel index") +
+  theme_minimal() +
+  plot_layout(ncol = 1)
+```
+
 ```{r rad-wv}
@@
 
+```{r}
+rm(rad)
+```
+
 ### Ensemble characteristics
@@
 
-```{r T-mse, fig.cap="MSE Temperature between analysis and guess in Kelvin^2."}
+```{r T-mse, eval=FALSE, fig.cap="MSE", include=FALSE}
 
@@
 A positive difference can also be observed between 850 and 600 hPa.
-```{r Q-mse, fig.cap="MSE Specific humidity between analysis and guess in g/Kg^2."}
+```{r Q-mse, eval=FALSE, fig.cap="MSE", include=FALSE}
 
@@
 
-```{r q-low, warning=FALSE, fig.align="center", message=FALSE}
+```{r q-low, eval=FALSE, fig.align="center", message=FALSE, warning=FALSE, include=FALSE}
 fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
@@
 
-```{r q-middle}
+```{r q-middle, eval=FALSE, include=FALSE}
 
@@
 ```
-```{r v-low}
+```{r v-low, eval=FALSE, include=FALSE}
 q[bottom_top %between% c(0, 7), .(v = mean(v)), 
@@
 ```
-```{r v-middle}
+
+```{r v-middle, eval=FALSE, include=FALSE}
 q[bottom_top %between% c(8, 12), .(v = mean(v)), 
@@
 
-```{r}
-q[date == unique(date)[1:3] & west_east %between% c(70, 180)] %>% 
-  .[, .(q_mean = mean(q),
-        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
-  .[, date := factor(date)] %>% 
-  ggplot(aes(south_north, bottom_top)) +
-  geom_contour_fill(aes(z = q_mean*1000, fill = stat(level_d))) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "BrBG", direction = 1,
-                       # limits = c(1, 19),
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  geom_contour(aes(z = v_mean, linetype = factor(-sign(..level..))), 
-               color = "grey30", size = 0.3, binwidth = 2) +
-  geom_text_contour(aes(z = v_mean), color = "grey30", size = 4, stroke = 0.1, breaks = seq(-10, 6, 2)) +
-  scale_linetype(guide = NULL) +
-  # geom_vector(aes(x = south_north, y = bottom_top,
-  #                 mag = abs(v_mean), angle = fifelse(v_mean > 0, 0, 180)),
-  #             arrow.angle = 6, size = 0.2, skip.x = 10) +
-  # scale_mag() +
-  facet_grid(date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                     E6 = "SATWND", E7 = "RAD"))) +
-  labs(y =  "Sigma levels", x = "Grid points") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r temp}
+```{r temp, eval=FALSE, include=FALSE}
 fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
@@
 
-```{r mucape}
+```{r mucape, eval=FALSE, include=FALSE}
 files <- Sys.glob(paste0(derived_data, "mucape/*"))
@@
 
+```{r summary, fig.height=12, fig.width=10}
+files <- Sys.glob(paste0(derived_data, "pw/*")) 
+files <- files[!str_detect(files, c("E3"))]
+
+dates <- c("20181122000000")
+
+pw <- purrr::map(files, function(f) {
+  details <- unglue::unglue(basename(f), "pw_ana_{exp}_{date}.nc")[[1]]
+
+  if (details$date %in% dates) {
+
+    ReadNetCDF(f, vars = c(value = "pw", lon = "XLONG", lat = "XLAT")) %>% 
+      .[, `:=`(exp = details$exp, 
+               date = lubridate::ymd_hms(details$date))] %>%
+      .[]
+  }
+}) %>% 
+  rbindlist() %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)]
+
+files <- Sys.glob(paste0(derived_data, "mucape/*")) 
+files <- files[!str_detect(files, c("E3"))]
+
+mucape <- purrr::map(files, function(f) {
+  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]
+
+  if (details$date %in% dates) {
+
+    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
+      .[, `:=`(exp = details$exp, 
+               date = lubridate::ymd_hms(details$date))] %>%
+      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
+      .[]
+  }
+}) %>% 
+  rbindlist()
+
+fcsts <- expand_grid(fcsts = c("20181122000000"),
+                     exp = c("E4", "E5", "E6", "E7"))
+
+ana <- lapply(seq_len(nrow(fcsts)), function(f) {
+
+  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
+    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
+    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
+    .[, theta := ReadNetCDF(paste0(derived_data, "/theta_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(th = "theta"), out = "vector")] %>%
+    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
+    .[, exp := fcsts[f, 2]] %>%  
+    .[]
+}) %>% 
+  rbindlist()
+
+cor <- ana[bottom_top %in% c(2, 10)] %>% 
+  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
+  .[, cortante := Mag(u_10 - u_2, v_10 - v_2)] %>% 
+  .[, ":="(u_2 = NULL, 
+           u_10 = NULL,
+           v_2 = NULL,
+           v_10 = NULL)]
+
+ana[bottom_top %between% c(1, 10), .(v_mean = mean(v)), 
+  by = .(south_north, west_east, date, exp)] %>% 
+  .[pw, on = .NATURAL] %>% 
+ggplot(aes(x, y)) +
+ # geom_point(aes(color = v_mean)) +
+  geom_contour_fill(aes(z = value, fill = stat(level_d)),
+                    proj = norargentina_lambert, alpha = 0.7) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "YlGn", direction = 1,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  geom_contour2(aes(z = v_mean, linetype = factor(-sign(v_mean))),
+                proj = norargentina_lambert, color = "darkorange", 
+                breaks = c(-10, -5, -2, 0, 2, 5, 10)) +
+   geom_text_contour(aes(z = v_mean), size = 4, stroke = 0.05,
+                     proj = norargentina_lambert, color = "darkorange", 
+                breaks = c(-10, -5, -2, 0, 2, 5, 10),
+                label.placement = label_placement_n(1)) +
+  scale_linetype(guide = NULL) +
+  cordillera +
+  geom_mapa() +
+  facet_grid(. ~ exp, 
+             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+ana[bottom_top %between% c(2, 10), .(theta_mean = mean(theta)), 
+  by = .(south_north, west_east, date, exp)] %>% 
+   .[pw, on = .NATURAL] %>% 
+ggplot(aes(x, y)) +
+ # geom_point(aes(color = v_mean)) +
+  geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
+                    proj = norargentina_lambert, alpha = 0.7,
+                    breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "RdYlBu", direction = -1,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  cordillera +
+  geom_mapa() +
+  facet_grid(. ~ exp, 
+             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3),
+        strip.text = element_blank(), 
+        strip.background = element_blank()) +
+
+mucape %>% 
+  .[variable == "mcape"] %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  .[cor, on = .NATURAL] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = value, fill = stat(level_d)), 
+                    proj = norargentina_lambert, limits = c(500, 6000)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  # geom_contour2(aes(z = cortante),
+  #               proj = norargentina_lambert, color = "black", size = 0.4,
+  #               breaks = c(10, 25)) +
+  geom_point(data = ~.x[cortante >= 10 & is.cross(x, y, skip = 1)], 
+             aes(x = lon, y = lat, 
+                 alpha = factor(cut_round(cortante, breaks = c(5, 15, 40)))),
+             size = 0.00002, shape = 20) +
+  scale_alpha_manual(values = c(0.3, 0.7), guide = NULL) +
+  cordillera +
+  geom_mapa() +
+  facet_grid(. ~ exp, 
+             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E7 = "RAD"))) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3),
+        strip.text = element_blank(), 
+        strip.background = element_blank()) +
+
+    plot_layout(ncol = 1, heights = c(1, 1, 1)) 
+```
+
+```{r}
+coord <- mucape[variable == "mcape", c("south_north", "lat")] %>% 
+  .[, .(lat = mean(lat)), by = south_north]
+
+ana[west_east %between% c(70, 180)] %>% 
+  .[, .(q_mean = mean(q),
+        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
+  coord[., on = "south_north"] %>% 
+  ggplot(aes(lat, bottom_top)) +
+  geom_contour_fill(aes(z = q_mean*1000, fill = stat(level_d))) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       # limits = c(1, 19),
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  geom_contour2(aes(z = v_mean, linetype = factor(-sign(..level..))), 
+               color = "grey30", size = 0.3, binwidth = 2) +
+  geom_text_contour(aes(z = v_mean), color = "grey30", 
+                    size = 3, stroke = 0.05, breaks = seq(-10, 6, 2)) +
+  scale_linetype(guide = NULL) +
+  scale_x_latitude(breaks = seq(-40, -20, 5)) +
+  facet_wrap( ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                     E6 = "SATWND", E7 = "RAD"))) +
+  labs(y =  "Sigma levels", x = "Grid points") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) 
+```
+
 Is the analysis correcting a dry bias on the model or the humidity is being transported into the domain?
@@
 
-```{r omb-q-evol, fig.cap="Evolution of the mean OmB for humidity throughout the periodod"}
+```{r omb-q-evol, eval=FALSE, fig.cap="Evolution", include=FALSE}
 files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E[4-7]/ANA/*/diagfiles/asim_conv_201811*.ensmean")
@@
   theme_minimal()
+
+rm(conv)
 
@@
 
-```{r omb-q-evol, fig.cap="Evolution of the mean OmB for humidity throughout the periodod"}
+```{r omb-q-evol-param, eval=FALSE, fig.cap="Evolution", include=FALSE}
 files <- c(Sys.glob("~/datosmunin/EXP/E5/ANA/*/diagfiles/asim_conv_201811*.mem*"),
@@
   theme_minimal()
+
+rm(conv)
 ```
@@
 
-```{r pbl-18Z, fig.cap="Zonal mean high of the PBL over the subdomain estimated for the PBL parameterization for Nov 21st, 18Z"}
+```{r pbl-18Z, eval=FALSE, fig.cap="Zonal mean high of the PBL over the subdomain estimated for the PBL parameterization for Nov 21st", include=FALSE}
 pblh18 <- purrr::map(exp, function(f) {
@@
 
-```{r pblh, fig.cap=""}
+```{r pblh, eval=FALSE, fig.cap="", include=FALSE}
 files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E[4-7]/ANA/*/analysis.ensmean")
@@
 
-  readRDS(f)
+  readRDS(f) %>% 
+    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)]
 
@@
 readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
+  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
   .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
   .[end_date %between% c(as_datetime("20181120190000"), as_datetime("20181123120000"))] %>% 
-  .[, .(pp_acum = mean(pp_acum)), by = .(end_date, y)] %>% 
-  ggplot(aes(end_date, y/100000)) +
-  geom_contour(aes(z = pp_acum, color = ..level..), 
-               breaks = seq(0, 12, 0.5)) +
-  scale_color_viridis_c(guide = NULL, breaks = seq(0, 12, 0.5)) +
+  .[, .(pp_acum = mean(pp_acum),
+        lat = mean(lat)), by = .(end_date, y)] %>% 
+  ggplot(aes(end_date, lat)) +
+  geom_contour2(aes(z = pp_acum, color = ..level..), 
+               breaks = seq(1, 15, 1)) +
+  scale_color_viridis_c(guide = NULL, breaks = seq(1, 14, 1), limits = c(1, 15)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  scale_y_latitude(breaks = seq(-40, -20, 5)) +
   labs(x = NULL, y = NULL, color = "PP acum", 
@@
   ppacum_1h %>% 
-  .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y)] %>% 
-  ggplot(aes(date, y/100000)) +
-  geom_contour(aes(z = pp_acum, color = ..level..), 
-               breaks = seq(0, 12, 0.5)) +
-  scale_color_viridis_c(breaks = seq(0, 12, 1), limits = c(0, 12),
-                        guide = guide_colorstrip(barwidth = 25,
+  .[, lat := lat[1], by = .(y)] %>% 
+  .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y, lat)] %>% 
+  ggplot(aes(date, lat)) +
+  # geom_raster(aes(fill = pp_acum)) +
+  geom_contour2(aes(z = pp_acum, color = ..level..),
+               breaks = seq(1, 15, 1)) +
+  scale_color_viridis_c(breaks = seq(1, 15, 1), limits = c(1, 15),
+                        guide = guide_colorstrip(inside = TRUE,
+                                                 barwidth = 25,
                                                  barheight = 1)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  scale_y_latitude(breaks = seq(-40, -20, 5)) +
   facet_wrap(~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*prop*")
@@
   scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barheight = 18)) +
+                  guide = guide_colorsteps(barheight = 18, 
+                                           barwidth = 1)) +
   geom_mapa() +
@@
   labs(x = NULL, y = NULL, fill = NULL) +
-  theme_minimal() +
+  theme_minimal()
 
@@
   scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barheight = 18)) +
+                  guide = guide_colorsteps(barheight = 18, barwidth = 1)) +
   geom_mapa() +
@@
   labs(x = NULL, y = NULL, fill = NULL) +
-  theme_minimal() +
-  plot_layout(ncol = 1)
+  theme_minimal()
+
+  rm(dbz)
 ```
@@
   .[variable == "td" & !str_detect(site, "/") & 
-      exp == "E4"] %>%, 
+      exp == "E4"] %>%
 ggplot(aes(launch_time, alt/1000)) +

### 2021-01-16 analysis/chapters/experimento_20181122.Rmd
Cambios menores en figuras que van en metodología. Mapa ahora no incluye provincias de paises limitrofes. Nor Sorry.

@@
 
-map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")
+map_arg <- rnaturalearth::ne_states(country = c("argentina"), 
+                                      returnclass = "sf")
+map_limitrofes <- rnaturalearth::ne_countries(country = c("Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")
 
@@
 geom_mapa <- function(fill = NA) {
-  list(geom_sf(data = map, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+  list(geom_sf(data = map_arg, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+       geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
        coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)))
@@
 sondeos <- obsCONV[type == 120] %>% unique(by = c("stationID")) %>% 
-  .[, source := "Radiosonde"]
+  .[, source := "ADPUPA"]
 
 ascat <- obsCONV[type == 290] %>% unique(by = c("stationID")) %>% 
-  .[, source := "ASCAT"]
+  .[, source := "ASCATW"]
 
 aircraf <- obsCONV[type %in% c(130, 131, 133)] %>%# unique(by = c("stationID")) %>% 
-  .[, source := "AIRCRFT"]
+  .[, source := "AIRCFT /\nAIRCAR"]
 
@@
   geom_point(alpha = 0.8, size = 1,
-             aes(color = source)) +
-  scale_color_manual(name =  NULL, values = c("Sfc - Non-official" = "#00695c", 
+             aes(color = source, shape = source)) +
+  scale_shape_manual(name = "Surface\nstations",  breaks = c("Sfc - Non-official", "Sfc - Official"),
+                     labels = c("ASWS","CSWS"), values = c(15, 17),
+                     guide = guide_legend(override.aes = list(size = 2))) +
+  scale_color_manual(name =  "Surface\nstations", values = c("Sfc - Non-official" = "#00695c", 
                                               "Sfc - Official" = "#FD8002"),
                      breaks = c("Sfc - Non-official", "Sfc - Official"),
-                     labels = c("Non-official (N = 820)","Official (N = 327)"),
+                     labels = c("ASWS","CSWS"),
                      guide = guide_legend(override.aes = list(size = 2)))+
@@
                      guide = guide_legend(override.aes = list(size = 2))) +
-  scale_size_manual(values = c(0.7, 0.5, 1.5, 0.1)) +
+  scale_size_manual(values = c(1.5, 0.7, 0.5, 0.1)) +
   # scale_color_brewer(palette = "Dark2") +
@@
 
-  plot_layout(ncol = 3, widths = c(0.49, 0.02, 0.49))
+  plot_layout(ncol = 3, widths = c(0.49, 0.02, 0.49)) +
+  plot_annotation(tag_levels = "a", tag_suffix = ")")
 ```
@@
   geom_tile(aes(fill = n)) +
-  scale_fill_viridis_c(trans = scales::log10_trans(),
+  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
                        guide = guide_colourbar(barheight = 18)) +
@@
   scale_date() +
-  labs(x = NULL, fill = NULL, y = "pressure (hPa)",
-       caption = "Number of observations") +
+  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
   theme_minimal()

### 2021-01-19 analysis/chapters/experimento_20181122.Rmd
Agrego scale_y scale_x a geom_mapa

@@
        geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
-       coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)))
+       coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)),
+       scale_x_longitude(ticks = 5),
+       scale_y_latitude(ticks = 5))
 }
@@
   geom_point(data = square2, aes(lon, lat), size = 0.2, color = "darkgrey") +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(
-    x = "Longitude", 
-    y = "Latitude") +
-  theme_minimal() +
-  theme(legend.position = "bottom") 
+  theme_minimal() +
+  theme(legend.position = "bottom", legend.box.spacing = margin(0, 0, 0, 0)) 
 ```
@@
   coord_cartesian(xlim = c(-5, 5)) +
-  facet_wrap(~sensor) +
+  facet_wrap(~sensor, labeller = labeller(sensor = toupper)) +
   labs(fill = NULL, x = NULL, color = NULL) +

### 2021-01-19 analysis/chapters/experimento_20181122.Rmd
Cambio varios graficos para unificar estetica

@@
   geom_point(data = square, aes(lon, lat), size = 0.2) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(x = NULL, y = NULL) +
   theme_minimal() + 
@@
 
-  plot_spacer() +
+  # plot_spacer() +
 
@@
   geom_point(data = square, aes(lon, lat), size = 0.2) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(x = NULL, y = NULL, color = NULL, shape = NULL, size = NULL) +
+  labs(color = NULL, shape = NULL, size = NULL) +
   theme_minimal() +
@@
 
-  plot_layout(ncol = 3, widths = c(0.49, 0.02, 0.49)) +
+  plot_layout(ncol = 3, widths = c(1, 1)) +
   plot_annotation(tag_levels = "a", tag_suffix = ")")
@@
   scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
-                       guide = guide_colourbar(barheight = 18)) +
+                       guide = guide_colourbar(barheight = 18,
+                                               barwidth = 0.8)) +
 
@@
                                          "v" = "V wind" ))) +
-  labs(x = "",
-       y = "",
-       fill = "") +
+  labs(fill = "") +
   theme_minimal() +
@@
                                                var = c("t" = "Temperature"))) +
-  labs(x = "",
-       y = "",
-       fill = "") +
+  labs(fill = "") +
   theme_minimal() +
@@
                         labels = c("Mean RCRV", "SD RCRV")) +
-  scale_x_level() +
+  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
   coord_flip() +
@@
   geom_line(aes(color = sensor, linetype = variable)) +
+  scale_color_discrete(labels = toupper) +
   scale_linetype_manual(values = c("dashed", "solid"),
@@
   coord_flip() +
-  scale_x_level() +
+  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
   labs(x = NULL, y = NULL,
@@
        x = NULL,
-       y = "Pressure level") +
-  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+       y = NULL) +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  # scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
   facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
@@
 
-  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
   labs(subtitle = "Specific humidity difference (g/Kg)",
-       y = "Pressure level",
+       y = NULL,
        x = NULL) +
@@
                   melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                breaks = seq(0, 30, 3), size = 0.3, linetype = 1, color = "grey30") +
+                breaks = seq(0, 30, 3), size = 0.2, linetype = 1, color = "grey30") +
   geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                    breaks = seq(0, 30, 3), color = "grey30", skip = 1, stroke = 0.2) +
+                    breaks = seq(0, 30, 3), color = "grey30", skip = 1, 
+                    size = 4, stroke = 0.1) +
   labs(subtitle = "U difference (m/s)",
        x = NULL,
-       y = "Pressure level") +
-  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+       y = NULL) +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
   facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
@@
                   melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                breaks = seq(-7, 0, 1), size = 0.3, linetype = 2, color = "grey30") +
+                breaks = seq(-7, 0, 1), size = 0.2, linetype = 2, color = "grey30") +
   geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                    breaks = seq(-7, 0, 1), color = "grey30", skip = 1, stroke = 0.2) +
+                    breaks = seq(-7, 0, 1), color = "grey30", skip = 1, 
+                    size = 4, stroke = 0.1) +
   labs(subtitle = "V difference (m/s)",
        x = NULL,
-       y = "Pressure level") +
-  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+       y = NULL) +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
   facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
@@
 ggplot(aes(x, y)) +
- # geom_point(aes(color = v_mean)) +
   geom_contour_fill(aes(z = value, fill = stat(level_d)),
@@
                                          E6 = "SATWND", E7 = "RAD"))) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(x = NULL,
-       y = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        legend.box.spacing = margin(0, 0, 0, 0),
         panel.ontop = TRUE,

### 2021-01-22 analysis/chapters/experimento_20181122.Rmd
Fix cotours V-diff

@@
                   melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                breaks = seq(-7, 0, 1), size = 0.2, linetype = 2, color = "grey30") +
+                breaks = seq(-9, 0, 1), size = 0.2, linetype = 2, color = "grey30") +
   geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                    breaks = seq(-7, 0, 1), color = "grey30", skip = 1, 
+                    breaks = seq(-9, 0, 1), color = "grey30", skip = 1, 
                     size = 4, stroke = 0.1) +
@@
 
-```{r summary, fig.height=12, fig.width=10}
+```{r summary, fig.height=10, fig.width=10}
 files <- Sys.glob(paste0(derived_data, "pw/*")) 
@@
   theme(legend.position = "bottom",
-        legend.box.spacing = margin(0, 0, 0, 0),
+        # legend.box.spacing = margin(0, 0, 0, 0),
         panel.ontop = TRUE,
@@
   theme(legend.position = "bottom",
-        legend.box.spacing = margin(0, 0, 0, 0),
+        # legend.box.spacing = margin(1, 1, 1, 1),
         panel.ontop = TRUE,
@@
                                                      E6 = "SATWND", E7 = "RAD"))) +
-  labs(y =  "Sigma levels", x = "Grid points") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
+  labs(y =  "Sigma levels") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        legend.box.spacing = margin(0, 0, 0, 0),
         panel.ontop = TRUE,
@@
 ```
+
 ```{r omb-t, fig.cap="Observation minus guess for humidity surface observations during Nov, 21st 18Z"}
@@
   .[, .(pp_acum = mean(pp_acum),
-        lat = mean(lat)), by = .(end_date, y)] %>% 
+        lat = mean(lat),
+        exp = "IMERG"), by = .(end_date, y)] %>% 
   ggplot(aes(end_date, lat)) +
@@
   scale_y_latitude(breaks = seq(-40, -20, 5)) +
-  labs(x = NULL, y = NULL, color = "PP acum", 
-       subtitle = "IMERG") +
+  facet_wrap(~ exp) +
+  labs(x = NULL, y = NULL, color = NULL) +
   theme_minimal() +
@@
                                                  barwidth = 25,
-                                                 barheight = 1)) +
+                                                 barheight = 0.8)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
@@
   scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barheight = 18)) +
+                  guide = guide_colorsteps(barwidth = 0.8, 
+                                           barheight = 20)) +
   geom_mapa() +
@@
   labs(x = NULL, y = NULL, fill = NULL) +
-  theme_minimal()
-```
-
-```{r dbz-ens}
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+
+```{r dbz-ens, fig.height=10, fig.width=7}
 
@@
   .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
-  .[date == unique(date)[1]] %>% 
+    .[date %in% unique(date)[1:2]] %>% 
+    .[, date := factor(date)] %>% 
   ggplot(aes(x, y)) +
@@
   scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barheight = 18, 
-                                           barwidth = 1)) +
+                  guide = guide_colorsteps(barheight = 0.8, 
+                                           barwidth = 20)) +
   geom_mapa() +
-  facet_grid(cumulus~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                   E6 = "SATWND", E7 = "RAD"))) +
-  labs(x = NULL, y = NULL, fill = NULL) +
-  theme_minimal()
-
-  dbz %>% 
-  .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
-  .[date == unique(date)[2]] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
-                    proj = norargentina_lambert,
-                    breaks = seq(0, 60, 5),) +
-  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barheight = 18, barwidth = 1)) +
-  geom_mapa() +
-  facet_grid(cumulus~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                   E6 = "SATWND", E7 = "RAD"))) +
-  labs(x = NULL, y = NULL, fill = NULL) +
-  theme_minimal()
+  ggh4x::facet_nested(date + cumulus ~ exp, 
+                      labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                   E6 = "SATWND", E7 = "RAD"))) +
+  labs(fill = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
 
@@
   .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
-  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2)), 
-           BIAS = mean(value - fcst_value)), by = .(iop, exp, variable, lev)] %>% 
+  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
+           BIAS = mean(value - fcst_value, na.rm = TRUE)), by = .(iop, exp, variable, lev)] %>% 
   melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
   ggplot(aes(lev/1000, value.est)) +
+  geom_hline(yintercept = 0, color = "grey30") +
   geom_line(aes(color = exp, linetype = estadistico)) +
@@
   coord_flip() +
-  facet_grid(iop~variable, scales = "free_x") +
-  labs(x = NULL, y = NULL, color = NULL, linetype = NULL) +
+  facet_grid(iop~variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
   theme_minimal()

### 2021-01-22 analysis/chapters/experimento_20181122.Rmd
pw + viento arreglado?

@@
 
-ana[bottom_top %between% c(1, 10), .(v_mean = mean(v)), 
+ana[bottom_top %between% c(2, 7), .(v_mean = mean(v)), 
   by = .(south_north, west_east, date, exp)] %>% 
@@
   geom_contour_fill(aes(z = value, fill = stat(level_d)),
-                    proj = norargentina_lambert, alpha = 0.7) +
+                    proj = norargentina_lambert) +
   scale_fill_distiller(name = NULL, super = ScaleDiscretised,
@@
                 proj = norargentina_lambert, color = "darkorange", 
-                breaks = c(-10, -5, -2, 0, 2, 5, 10)) +
-   geom_text_contour(aes(z = v_mean), size = 4, stroke = 0.05,
+                breaks = c(-10, -5, 0), size = 0.4) +
+   geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
                      proj = norargentina_lambert, color = "darkorange", 
-                breaks = c(-10, -5, -2, 0, 2, 5, 10),
+                breaks = c(-10, -5, 0), skip = 0,
                 label.placement = label_placement_n(1)) +
-  scale_linetype(guide = NULL) +
-  cordillera +
+  scale_linetype_manual(guide = NULL, values = c(2, 1)) +
+  # cordillera +
   geom_mapa() +
@@
   geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
-                    proj = norargentina_lambert, alpha = 0.7,
+                    proj = norargentina_lambert,
                     breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
@@
                                                 barheight = 0.8)) +
-  cordillera +
+  # cordillera +
   geom_mapa() +
@@
   scale_alpha_manual(values = c(0.3, 0.7), guide = NULL) +
-  cordillera +
+  # cordillera +
   geom_mapa() +

### 2021-02-02 R/read_diag_rad.R
Arreglo bug en read_diag

@@ -5,6 +5,9 @@
 #'
 #' @param file_list lista de archivos
 #' @param exp nombre del experimento asociado
+#'
+#' @importFrom magrittr %>%
+#'
 #' @export
 read_diag_rad <- function(file_list, exp) {
 
@@ -25,7 +28,7 @@ read_diag_rad <- function(file_list, exp) {
     }
     out
   }) %>%
-    rbindlist()
+    data.table::rbindlist()
 
   colnames(diag) <- c("sensor", "channel", "freq", "lat", "lon", "peakwt", "press", "dhr", "tb_obs", "tbc", "tbcnob",
                       "varch", "errinv", "qc", "emis", "tlapchn", "rzen", "razi", "rlnd", "rice", "rsnw", "rcld",
@@ -34,3 +37,5 @@ read_diag_rad <- function(file_list, exp) {
 }
 
 .datatable.aware <- TRUE
+
+

### 2021-02-02 analysis/scripts/acum_update.R
acum update

@@ -1,6 +1,6 @@
-exp <- "E7"
+exp <- "E6"
 ini_date <- ymd_hms("20181120180000")
-ciclos <- 50
+ciclos <- 67
 
 dates <- seq.POSIXt(ini_date, by = "hour",
                     length.out = ciclos)
@@ -52,8 +52,9 @@ for (d in seq_along(dates)) {
     # espacial_qv$qv <- espacial_qv$qv + update$qv
   }
 
-
 }
+
+write_rds(global_qv, paste0("~/mesoda/analysis/data/derived_data/omb_acum_qv_", exp, ".rds"))
 
 
 rbind(global_qv_E6, global_qv_E7) %>%

### 2021-02-02 analysis/chapters/experimento_20181122.Rmd
Cambios en algunas figuras. Mejora la figura de resumen

@@
   cordillera +
-  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
-  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
+  geom_mapa() +
   facet_grid(date ~ exp, 
@@
                                          E6 = "SATWND", E7 = "RAD"))) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
   labs(title = "Low level meridional wind (m/s)",
@@
 
-```{r summary, fig.height=10, fig.width=10}
+```{r summary, fig.height=10, fig.width=9}
 files <- Sys.glob(paste0(derived_data, "pw/*")) 
@@
                                                 barheight = 0.8)) +
-  geom_contour2(aes(z = v_mean, linetype = factor(-sign(v_mean))),
+  geom_contour2(aes(z = v_mean, size = stat(level)),
                 proj = norargentina_lambert, color = "darkorange", 
-                breaks = c(-10, -5, 0), size = 0.4) +
-   geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
-                     proj = norargentina_lambert, color = "darkorange", 
-                breaks = c(-10, -5, 0), skip = 0,
-                label.placement = label_placement_n(1)) +
-  scale_linetype_manual(guide = NULL, values = c(2, 1)) +
-  # cordillera +
+                breaks = c(-10, -5)) +
+  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
+                        labels = c("-10 m/s", "-5 m/s"),
+                        name = NULL) +
+   # geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
+   #                   proj = norargentina_lambert, color = "darkorange", 
+   #              breaks = c(-10, -5), skip = 0,
+   #              label.placement = label_placement_n(1)) +
+  # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
+  cordillera +
   geom_mapa() +
@@
                                                 barheight = 0.8)) +
-  # cordillera +
+  cordillera +
   geom_mapa() +
@@
              size = 0.00002, shape = 20) +
-  scale_alpha_manual(values = c(0.3, 0.7), guide = NULL) +
-  # cordillera +
+  scale_alpha_manual(values = c(0.3, 0.7), name = NULL, labels = c(">5 m/s", ">15 m/s"),
+                     guide = guide_legend(override.aes = list(size = 2))) +
+  cordillera +
   geom_mapa() +
@@
 
-    plot_layout(ncol = 1, heights = c(1, 1, 1))
+    plot_layout(ncol = 1, heights = c(1, 1, 1)) +
+  plot_annotation(tag_levels = "a", tag_suffix = ")")
 ```
@@
 
-ana[west_east %between% c(70, 180)] %>% 
+ana[west_east %between% c(70, 180) & bottom_top %between% c(1, 25)] %>% 
   .[, .(q_mean = mean(q),
@@
   geom_text_contour(aes(z = v_mean), color = "grey30", 
-                    size = 3, stroke = 0.05, breaks = seq(-10, 6, 2)) +
+                    size = 3, stroke = 0.05, breaks = seq(-10, 6, 2),
+                    label.placement = label_placement_n(1)) +
   scale_linetype(guide = NULL) +

### 2021-02-02 analysis/chapters/experimento_20181122.Rmd
Versión simplificada de figura de la innovación de qv acumulada

@@
   # .[exp == "E7"] %>% 
-  .[, nivel := fcase(bottom_top %in% c(1:8), "1 to 8",
-                     bottom_top %in% c(9:13), "9 to 13",
-                     bottom_top %in% c(14:20), "14 to 20",
+  .[, nivel := fcase(bottom_top %in% c(1:8), "lower",
+                     bottom_top %in% c(9:13), "middle",
+                     bottom_top %in% c(14:20), "upper",
                      bottom_top %in% c(21:38), "21 to 38")] %>%
+  .[ bottom_top %between% c(1, 20)] %>% 
   .[, acum := cumsum(qv), by = .(bottom_top, exp)] %>%
+  .[, acum := mean(acum), by = .(nivel, exp, date)] %>%
   # .[bottom_top %in% c(1:18)] %>%
   ggplot(aes(date, acum*1000)) +
-  geom_line(aes(color = exp, linetype = exp, group = interaction(bottom_top, exp))) +
+  geom_line(aes(color = exp, linetype = exp, group = interaction(nivel, exp))) +
   # geom_label(data = ~.x[date %in% sample(unique(date), 38)] %>% group_by(exp, bottom_top) %>% sample_n(size = 1),
    geom_label(data = ~.x[date %in% unique(date)[67]],
-             aes(label = bottom_top, color = exp, 
-                 group = interaction(bottom_top, exp)),
+             aes(label = nivel, color = exp, 
+                 group = interaction(nivel, exp)),
              size = 3, fill = "white", show.legend = FALSE,
@@
                                                       E6 = "SATWND", E7 = "RAD")) +
-  facet_wrap(~nivel, scales = "free_y") +
+  # facet_wrap(~nivel, scales = "free_y") +
   labs(x = NULL, y = "Accumulated specific humidity (g/Kg)",
        color = NULL, linetype = NULL) +
-  theme_minimal()
+  theme_minimal() +
+  theme(legend.position = "bottom")
 

### 2021-02-05 R/scale_color_dbz.R
Nueva escala de colores para dbz, igual no se parece a la del SMN

@@ -6,20 +6,47 @@
 #' @param ...
 #'
 #' @export
-scale_color_dbz <- function(colours = hex,
-                              limits = c(0, 60),
+scale_color_dbz <- function(colours = ref,
+                              limits = c(-30, 75),
 
                               guide = guide_colorbar(barheight = 15),
                               ...) {
 
-  colours <- rgb(ref$R, ref$G, ref$B, maxColorValue = 1)
+  # colours <- rgb(ref$R, ref$G, ref$B, maxColorValue = 1)
 
-  scale_color_gradientn(colours = colours,
+  scale_color_gradientn(colours = ref,
                         limits = limits,
                         guide = guide, ...)
 }
 
-ref <- dplyr::tribble(
+ref <- c(
+#   "#646464", # ND
+   "#ccffff", # -30
+   "#cc99cc", # -25
+   "#996699", # -20
+   "#663366", # -15
+   "#cccc99", # -10
+   "#999966", # -5
+   "#646464", # 0
+   "#04e9e7", # 5
+   "#019ff4", # 10
+   "#0300f4", # 15
+   "#02fd02", # 20
+   "#01c501", # 25
+   "#008e00", # 30
+   "#fdf802", # 35
+   "#e5bc00", # 40
+   "#fd9500", # 45
+   "#fd0000", # 50
+   "#d40000", # 55
+   "#bc0000", # 60
+   "#f800fd", # 65
+   "#9854c6", # 70
+   "#fdfdfd" # 75
+)
+
+
+ref2 <- dplyr::tribble(
   ~R, ~G, ~B,
    1.0/255.0, 159.0/255.0, 244.0/255.0,
    3.0/255.0,   0.0/255.0, 244.0/255.0,

### 2021-02-05 analysis/chapters/experimento_20181122.Rmd
Arreglo figura con acumulados de pp, ahora tiene sentido.

@@
 ```{r pp-hov, fig.cap="Hövmoller of hourly acumulated precipitation (for ensemble mean) over the inner domain."}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*mean*")
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*mean_acum_1h.rds")
 
@@
   geom_contour2(aes(z = pp_acum, color = ..level..), 
-               breaks = seq(1, 15, 1)) +
-  scale_color_viridis_c(guide = NULL, breaks = seq(1, 14, 1), limits = c(1, 15)) +
+               breaks = seq(1, 12, 1)) +
+  scale_color_viridis_c(guide = NULL, breaks = seq(1, 12, 1), limits = c(1, 12)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
@@
   geom_contour2(aes(z = pp_acum, color = ..level..),
-               breaks = seq(1, 15, 1)) +
-  scale_color_viridis_c(breaks = seq(1, 15, 1), limits = c(1, 15),
+               breaks = seq(1, 12, 1)) +
+  scale_color_viridis_c(breaks = seq(1, 12, 1), limits = c(1, 12),
                         guide = guide_colorstrip(inside = TRUE,
@@
                     proj = norargentina_lambert,
-                    breaks = seq(0, 60, 5),) +
+                    breaks = seq(-30, 75, 5)) +
   scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barwidth = 0.8, 
-                                           barheight = 20)) +
+                  guide = guide_colorsteps(barwidth = 20, 
+                                           barheight = 0.8)) +
+  ggnewscale::new_scale_fill() +
+    geom_contour_fill(aes(z = max_dbz), fill = "white", 
+                    proj = norargentina_lambert,
+                    breaks = c(-Inf, 15)) +
   geom_mapa() +
@@
                     proj = norargentina_lambert,
-                    breaks = seq(0, 60, 5),) +
+                    breaks = seq(-30, 75, 5),) +
   scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
@@
                                            barwidth = 20)) +
+  ggnewscale::new_scale_fill() +
+  geom_contour_fill(aes(z = max_dbz), fill = "white", 
+                    proj = norargentina_lambert,
+                    breaks = c(-Inf, 10)) +
   geom_mapa() +
@@
 
-  rm(dbz)
+  # rm(dbz)
 ```
@@
 fss %>% 
+  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
   ggplot(aes(date, fss)) +
@@
                                                       E6 = "SATWND", E7 = "RAD")) +
-  scale_date(20181121000000, 20181123000000, "12 hours") +
+  scale_date(20181122000000, 20181123000000, "12 hours") +
+  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
   facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 

### 2021-02-05 analysis/chapters/experimento_20181122.Rmd
FSS solo para 1 y 25 mm y 10 y 100 km. escala de refectividad correcta, creo.

@@
 
-dates <- rep(c("20181122100000", "20181122220000"), 1, each = 4)
-files <- paste0(derived_data, "dbz/maxdbz_ana_", experiments, "_", dates, ".nc")
+dates <- c("20181122100000", "20181122130000", 
+           "20181122160000", "20181122190000")
+files <- paste0(derived_data, "dbz/maxdbz_ana_E7_", dates, ".nc")
 
@@
 
+values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
+22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
+42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
+61.5, 63, 64.5, 66, 69, 76.5)
+
+colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
+"#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
+"#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
+"#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
+"#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
+"#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
+"#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
+"#95E3C5", "#85E0BD")
+
+dbz %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, guide = guide_colorsteps(barwidth = 30, 
+                                                                         barheight = 0.8)) +
+
+  # scale_color_dbz(name = NULL, aesthetics = "fill", 
+  #                 guide = guide_colorstrip(barwidth = 20, 
+  #                                          barheight = 0.8,
+  #                                          inside = TRUE)) +
+  # ggnewscale::new_scale_fill() +
+  # geom_contour_fill(aes(z = max_dbz), fill = "white", 
+  #                   proj = norargentina_lambert,
+  #                   breaks = c(-Inf, 10)) +
+  geom_mapa() +
+  facet_wrap(~ date, ncol = 4) +
+  labs(x = NULL, y = NULL, fill = NULL, color = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
+
+```{r dbz-mean}
+
+dates <- rep(c("20181122100000"), 1, each = 4)
+files <- paste0(derived_data, "dbz/maxdbz_ana_", experiments, "_", dates, ".nc")
+
+dbz <- purrr::map(files, function(f){
+
+  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
+  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             date = ymd_hms(meta[[1]][["date"]]))] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] 
+}) %>% 
+  rbindlist()
+
 dbz %>% 
@@
                     proj = norargentina_lambert,
-                    breaks = seq(-30, 75, 5)) +
-  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barwidth = 20, 
-                                           barheight = 0.8)) +
-  ggnewscale::new_scale_fill() +
-    geom_contour_fill(aes(z = max_dbz), fill = "white", 
-                    proj = norargentina_lambert,
-                    breaks = c(-Inf, 15)) +
+                    breaks = seq(15, 75, 5)) +
+  scale_fill_manual(values = ref, drop = FALSE, guide = guide_colorsteps(barwidth = 20, 
+                                                                         barheight = 0.8)) +
+
   geom_mapa() +
-  facet_grid(date~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+  facet_wrap(~exp, ncol = 2, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                                    E6 = "SATWND", E7 = "RAD"))) +
@@
 
-```{r dbz-ens, fig.height=10, fig.width=7}
-
-files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[4-7]_0*"))
-files <- files[str_detect(files, "20181122000000", negate = TRUE)]
+```{r dbz-ens}
+files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[4-7]_0*_20181122100000.nc"))
 
@@
   .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
-    .[date %in% unique(date)[1:2]] %>% 
+    # .[date %in% unique(date)[1:2]] %>% 
     .[, date := factor(date)] %>% 
   ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
                     proj = norargentina_lambert,
-                    breaks = seq(-30, 75, 5),) +
-  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barheight = 0.8, 
-                                           barwidth = 20)) +
-  ggnewscale::new_scale_fill() +
-  geom_contour_fill(aes(z = max_dbz), fill = "white", 
-                    proj = norargentina_lambert,
-                    breaks = c(-Inf, 10)) +
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    drop = FALSE, guide = guide_colorsteps(barwidth = 0.8, 
+                                                                         barheight = 30)) +
   geom_mapa() +
@@
   theme_minimal() +
-  theme(legend.position = "bottom",
+  theme(#legend.position = "bottom",
         legend.box.spacing = margin(0, 0, 0, 0),
@@
 ```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-7]new.csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-7].csv")
 
@@
                                             "10" = "10 mm", "25" = "25 mm"),
-                                      w = c("5" = "50 km", "11" = "100 km", 
+                                      w = c("1" = "10 km", "5" = "50 km", "11" = "100 km", 
                                             "51" = "500 km", "101" = "1000 km"))) +
@@
 
+fss %>% 
+  .[w %in% c(1, 11) & q %in% c(1, 25)] %>% 
+  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
+  ggplot(aes(date, fss)) +
+  geom_hline(yintercept = 1, color = "darkgray") +
+  geom_line(aes(color = exp)) +
+  # geom_point(aes(color = exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date(20181122000000, 20181123000000, "12 hours") +
+  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
+                                            "10" = "10 mm", "25" = "25 mm"),
+                                      w = c("1" = "10 km", "5" = "50 km", "11" = "100 km", 
+                                            "51" = "500 km", "101" = "1000 km"))) +
+  labs(color = NULL, x = NULL, y = "FSS") +
+  theme_minimal()
 ```

### 2021-02-19 analysis/chapters/experimento_20181122.Rmd
FSS de los pronósticos. Podría dar mejor

@@
 
-cor <- ana[bottom_top %in% c(2, 10)] %>% 
+cor <- ana[bottom_top %in% c(1, 16)] %>% 
   dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
-  .[, cortante := Mag(u_10 - u_2, v_10 - v_2)] %>% 
+  .[, cortante := Mag(u_16 - u_1, v_16 - v_1)] %>% 
   .[, ":="(u_2 = NULL, 
@@
   geom_contour2(aes(z = v_mean, size = stat(level)),
-                proj = norargentina_lambert, color = "darkorange", 
+                proj = norargentina_lambert, color = "blueviolet", 
                 breaks = c(-10, -5)) +
@@
   #               breaks = c(10, 25)) +
-  geom_point(data = ~.x[cortante >= 10 & is.cross(x, y, skip = 1)], 
-             aes(x = lon, y = lat, 
-                 alpha = factor(cut_round(cortante, breaks = c(5, 15, 40)))),
-             size = 0.00002, shape = 20) +
-  scale_alpha_manual(values = c(0.3, 0.7), name = NULL, labels = c(">5 m/s", ">15 m/s"),
-                     guide = guide_legend(override.aes = list(size = 2))) +
+  geom_contour2(aes(z = cortante, size = stat(level)),
+                proj = norargentina_lambert, color = "black", 
+                breaks = c(15, 30)) +
+  scale_size_continuous(breaks = c(15, 30), range = c(0.4, 0.8),
+                        labels = c("15 m/s", "30 m/s"),
+                        name = NULL) +
+  # geom_point(data = ~.x[cortante >= 15 & is.cross(x, y, skip = 1)], 
+  #            aes(x = lon, y = lat, 
+  #                alpha = factor(cut_round(cortante, breaks = c(15, 25, Inf)))),
+  #            size = 0.00002, shape = 20) +
+  # scale_alpha_manual(values = c(0.3, 0.7), name = NULL, #labels = c(">5 m/s", ">15 m/s"),
+  #                    guide = guide_legend(override.aes = list(size = 2))) +
   cordillera +
@@
 ```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_*_E[4-7]new.csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[4-7]_ens.csv")
 
@@
 
-  meta <- unglue(basename(f), "fss_6h_fcst_{ini_date}_{exp}new.csv")
+  meta <- unglue(basename(f), "fss_6h_fcst_ens_{ini_date}_{exp}_ens.csv")
 
@@
                                                       E6 = "SATWND", E7 = "RAD")) +
+  scale_date(20181122000000, 20181123000000, "12 hours") +
   facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                             "10" = "10 mm", "25" = "25 mm"),
-                                      w = c("5" = "50 km", "11" = "100 km", 
+                                      w = c("1" = "1 km", "5" = "50 km", 
+                                            "11" = "100 km", 
                                             "51" = "500 km", "101" = "1000 km"))) +

### 2021-02-19 analysis/scripts/wrap_fss.R
Cambios para calcular el FSS del ensamble de pronosticos

@@ -10,15 +10,15 @@ library(mesoda)
 #future::plan("cluster", workers = 3)
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
 wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E7"
-run <- "ana_ens"
+exp <- "E4"
+run <- "fcst_ens"
 
-ini_date <- ymd_hms("20181120180000")
-ciclos <- 67
+ini_date <- ymd_hms("20181122060000")
+ciclos <- 31
 
-acumulado <- 3
+acumulado <- 6
 q <- c(1, 5, 10, 25) #10mm, para arrancar pensando en pp acumulada
-w <- c(5, 11, 51, 101) #ancho de cada box en puntos de grilla
+w <- c(1, 5, 11, 51, 101) #ancho de cada box en puntos de grilla
 
 # Inicio
 dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
@@ -70,6 +70,13 @@ fss_out <- purrr::map_dfr(dates, function(d) {
 
     pp_wrf <- readRDS(files_wrf)
 
+  } else if (run == "fcst_ens") {
+
+    files_wrf <- paste0(wrf_path, "derived_data/ppacum/", exp, "_fcst_", format(ini_date, "%Y%m%d%H"), "_prop_acum_",
+                        acumulado, "h.rds")
+
+    pp_wrf <- readRDS(files_wrf) %>%
+      .[date == d]
 
   } else {
 
@@ -127,6 +134,18 @@ fss_out <- purrr::map_dfr(dates, function(d) {
       fss <- fcst[, calculate_fss(pp[[1]], obs[, pp[[1]]],
                                   q = qi, w = w, binary = FALSE),
                   by = .(exp, date)]
+    } else if (run == "fcst_ens") {
+
+      fcst <- pp_wrf[umbral == qi] %>%
+        .[, .(pp = list(dcast(.SD, x ~ y, value.var = "prop") %>%
+                          .[, -1] %>%
+                          as.matrix())),
+          by = .(exp)]
+
+      fss <- fcst[, calculate_fss(pp[[1]], obs[, pp[[1]]],
+                                  q = qi, w = w, binary = FALSE),
+                  by = .(exp)] %>%
+        .[, date := d]
     } else {
       fcst <- pp_wrf %>%
[…]

### 2021-02-19 analysis/scripts/acumulados_wrf.R
Acumulados de fcst a 6 horas.

@@ -83,7 +83,7 @@ q <- c(1, 5, 10, 25, 50)
 
 # Inicio
 
-lead_time <- seq(0, 30, acumulado)
+lead_time <- seq(0, 36, acumulado)
 
 pp_wrf <- purrr::map(lead_time, function(f) {
 
@@ -134,18 +134,98 @@ pp_prop <- purrr::map(q, function(f){
 saveRDS(pp_prop, paste0(exp, "_fcst_", format(ini_date, "%Y%m%d%H"), "_prop_acum_1h.rds"))
 saveRDS(pp_wrf_mean, paste0(exp, "_fcst_", format(ini_date, "%Y%m%d%H"), "_mean_acum_1h.rds"))
 
+# Acumulados cada 6 horas -----------------------------------------------------
+
+imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
+wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
+exp <- "E7"
+run <- "fcst"
+
+ini_date <- ymd_hms("20181122060000")
+ciclos <- 31
+acumulado <- 6
+
+dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
+                    length.out = ciclos - acumulado)
+
+
+q <- c(1, 5, 10, 25, 50)
+
+# Inicio
+
+pp_wrf <- purrr::map(dates, function(d) {
+
+  print(d)
+  date <- d
+
+  lead_time <- as.numeric(difftime(date, ini_date, units = "hours"))
+
+  files_wrf <-  purrr::map(c(acumulado, 0), function(l) {
+    paste0(wrf_path, exp, "/", toupper(run), "/",
+           format(ini_date, "%Y%m%d%H"), "/NPP/NPP_",
+           format(ini_date, "%Y"), "-",
+           format(ini_date, "%m"), "-",
+           format(ini_date, "%d"), "_",
+           format(ini_date, "%H"), "_FC",
+           formatC(lead_time - l, width = 2, flag = "0"),
+           ".nc")
+  }) %>% unlist()
+
+
+  pp_wrf <- purrr::map(files_wrf, function(f) {
+
+    if (!file.exists(f)) {
+      return(NULL)
+    }
+
+    ReadNetCDF(f, vars = c("PP", "XLAT", "XLONG")) %>%
[…]

### 2021-02-19 analysis/chapters/forecast_20181122.Rmd
Init analisis de los pronosticos

--- 2021-02-05 analysis/chapters/experimento_20181122.Rmd
@@
 ---
-title: "Asimilación de distintas observaciones"
+title: "FCST"
 output: 
@@
 
-## Case Study: November 22th, 2020
+### FSS
 
-The experiments focus on a mesoscale convective system developed over central and north-eastern Argentina during November 22th, 2018.
-This MCS corresponds to IOP08 (Severe, Upscale Growth handoff) of the RELAMPAGO field campaign.
-
-This case stady was selected to focus on the RELAMPAGO period and to take advantage of the observations taken during the intensive observation periods.
-
-```{r caso-gif, fig.cap="GOES-16 channel 13 (clean longwave-IR) during November, 22th.", out.height=450, out.width=600, fig.align="center"}
-knitr::include_graphics("https://github.com/paocorrales/Analisis_2018112022/blob/master/fig/caso_20181122.gif?raw=true")
-```
-
-### Model and ensemble configuration
-
-To run the simulations the WRF-ARW (V3.9) model with 10 km horizontal resolution and xx vertical levels was used.
-The GFS deterministic run (0.5° resolution) was used for the initial and boundary conditions.
-
-The LETKF versión of the GSI system was choosen for the assimilation.
-A 60 member ensemble is initialized using random perturbations and a multi-physics squeme using 9 combinations of 3 Cumulus (Kain–Fritsch, Grell–Freitas, Betts–Miller–Janjic) and 3 PBL (Yonsei University Scheme, Mellor–Yamada–Janjic Scheme, Mellor–Yamada Nakanishi Niino) parameterizations as shown in Table \@ref(tab:fisicas).
-
-```{r fisicas}
-fisica[, .(mem = paste(mem, collapse = ", ")), by = .(fisica)] %>% 
-  kbl(col.names = c("Parameterizations", "# Member"),
-      caption = "Posible Cumulus and PBL combinations and the members in each group.") %>% 
-  kable_classic_2(full_width = F)
-```
-
-The domain shown in Figure \@ref(fig:dominio) covers the center and north of Argentina and part of Bolivia and Paraguay to capture the development of the MCS during the simulation period (up to 12UTC 2020/11/23).
-
-```{r dominio, fig.cap="Horizontal domain for all simulations and topography.", fig.height=4}
-fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>%
-
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
-                    breaks = seq(0, 6000, 500)) +
-  scale_fill_distiller(name = NULL,
-                       type = "seq",
-                       palette = "RdYlGn",
-                       direction = -1,
-                       breaks = seq(0, 6000, 1000),
-                       guide = guide_colorstrip(barwidth = 20,
-                                                barheight = 0.8)) +
-  geom_mapa() +
-  geom_point(data = square, aes(lon, lat), size = 0.2) +
-  geom_point(data = square2, aes(lon, lat), size = 0.2, color = "darkgrey") +
-  theme_minimal() +
-  theme(legend.position = "bottom", legend.box.spacing = margin(0, 0, 0, 0)) 
-```
-
-### Experiments configuration
-
-#### Assimilation cycle
-
-Each simulation starts at 12 UTC on Nov 20th with 6 hours of spin up.
-The assimilation starts at 18 UTC and continue until 12 UTC Nov, 23th every houer.
-Two ensemble forecasts are initialized from de the analysis at 00 and 06 UTC Nov 22th and continue until 12 UTC Nov 23rd (Figure \@ref(fig:ciclo)).
-
-The analysis is perform in one-hour cetered window using the Observation Operator from GSI to compare the available observations with the first guess at the closest time (FGAT - First Guess at Appropriate Time).
-Then the LETKF is run to apply the innovations to the analysis for each member.
-
-```{r ciclo, fig.cap="Diagram of analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
-The zoomed figure shows assimilation in one-hour centered window and new boundary conditions from GFSD every 6 hours.", out.height=400, out.width=800, fig.align="center"}
-knitr::include_graphics(here("analysis", "figures", "analysis_cycle.svg"))
-```
-Four experiments were performed using an incremental set of observations each time (Table \@ref(tab:obs)).
-The CONV experiment uses conventional observations from the prepBUFR file that includes surface observations (land and sea) and profiles from soundings and aircrafts.
-On the Figure \@ref(fig:mapa-obs), the orange dots are the surface stations and ships locations included in this experiment.
-
-The AUT experiment add to CONV the regional automatic stations networks from Argentina and near countries.
-The added stations are colored in green in Figure \@ref(fig:mapa-obs).
-This stations have greater spatial coverage and higher sampling frequency.
-
-SATWND add to AUT the satellite motion vectors from several satellites including GOES-16 and finally the RAD experiments also includes randiances observations from sensors over polar orbit satellites.
-
-```{r obs}
-tribble(
-  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
-  "ADPSFC", "x", "x", "x", "x",
-  "ADPUPA", "x", "x", "x", "x",
-  "AIRCFT", "x", "x", "x", "x",
-  "AIRCAR", "x", "x", "x", "x",
-  "SFCSHP", "x", "x", "x", "x",
-  "ASCATW", "x", "x", "x", "x", 
-  "AUTSFC", " ", "x", "x", "x",
-  "SATWND", " ", " ", "x", "x",
-  "RAD", " ", " ", " ", "x",
-) %>% 
-  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
-      caption = "Observation types assimilated in each experiment.") %>% 
-  kable_classic_2(full_width = F)
-```
-
-```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots).", fig.height=4, fig.width=8}
-obsAUT <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
-                na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
-  .[, c("V2", "V4") := NULL] %>% 
-  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))
-
-obsCONV <- fread(here("analysis", "data", "raw_data", "E4_asim_conv_20181121120000.ensmean"), 
-                 na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
-  .[, c("V2", "V4") := NULL] %>% 
-  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))
-
-obsSATW <- fread(here("analysis", "data", "raw_data", "E6_asim_conv_20181121120000.ensmean"), 
-                 na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
-  .[, c("V2", "V4") := NULL] %>% 
-  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))
-
-oficiales <- obsCONV[type %in% c(181, 281)] %>% unique(by = c("stationID")) %>% 
-  .[, source := "Sfc - Official"]
-
-sondeos <- obsCONV[type == 120] %>% unique(by = c("stationID")) %>% 
-  .[, source := "ADPUPA"]
-
-ascat <- obsCONV[type == 290] %>% unique(by = c("stationID")) %>% 
-  .[, source := "ASCATW"]
-
-aircraf <- obsCONV[type %in% c(130, 131, 133)] %>%# unique(by = c("stationID")) %>% 
-  .[, source := "AIRCFT /\nAIRCAR"]
-
-satwnd <- obsSATW[type %between% c(240, 260)] %>%# unique(by = c("stationID")) %>% 
-  .[, source := "SATWND"]
-
-no_oficiales <- obsAUT %>% 
-  .[type %in% c(181, 187)] %>% 
-  .[!str_detect(stationID, "SMN")] %>%
-  .[!stationID %in% oficiales$stationID] %>% 
-  unique(by = c("stationID")) %>% 
-  .[, source := "Sfc - Non-official"]
-
-rbindlist(list(no_oficiales, oficiales)) %>% 
-  ggplot(aes(ConvertLongitude(lon), lat)) +
-  geom_mapa() +
-  geom_point(alpha = 0.8, size = 1,
-             aes(color = source, shape = source)) +
-  scale_shape_manual(name = "Surface\nstations",  breaks = c("Sfc - Non-official", "Sfc - Official"),
-                     labels = c("ASWS","CSWS"), values = c(15, 17),
-                     guide = guide_legend(override.aes = list(size = 2))) +
-  scale_color_manual(name =  "Surface\nstations", values = c("Sfc - Non-official" = "#00695c", 
-                                              "Sfc - Official" = "#FD8002"),
-                     breaks = c("Sfc - Non-official", "Sfc - Official"),
-                     labels = c("ASWS","CSWS"),
-                     guide = guide_legend(override.aes = list(size = 2)))+
-  geom_point(data = square, aes(lon, lat), size = 0.2) +
-  theme_minimal() + 
-  theme(legend.position = "bottom", legend.justification = "left",
-         legend.key.size = unit(0.8, "lines")) +
-
-  # plot_spacer() +
-
-  rbindlist(list(ascat, sondeos, aircraf, satwnd)) %>% 
-  ggplot(aes(ConvertLongitude(lon), lat)) +
-  geom_mapa() +
-  geom_point(alpha = 1,
-             aes(color = source, shape = source, size = source)) +
-  scale_color_manual(values = c("darkorange", "purple", "cyan4", "grey70"),
-                     guide = guide_legend(override.aes = list(size = 2))) +
-  scale_size_manual(values = c(1.5, 0.7, 0.5, 0.1)) +
-  # scale_color_brewer(palette = "Dark2") +
-  # khroma::scale_color_vibrant() +
-  geom_point(data = square, aes(lon, lat), size = 0.2) +
-  labs(color = NULL, shape = NULL, size = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom", legend.justification = "left",
-        legend.key.size = unit(0.8, "lines")) +
-
-  plot_layout(ncol = 3, widths = c(1, 1)) +
-  plot_annotation(tag_levels = "a", tag_suffix = ")")
-```
-
-Figure with all type of observations (radiosonde, convencional, automatic station, aircraft, ships/buoys, satwind)
-Figure with number of observation throughout the cycles for each type including riadiances.
-
-```{r rad, fig.cap="Radiances for each 50 hPa pressure layer over time", fig.scap=c("OmB", "Number of observations")}
-
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/*/diagfiles/asim*ensmean")
-
-files <- files[!str_detect(files, "conv")]
-
-rad <- read_diag_rad(files, "E7") %>% 
-  satinfo[., on = c("sensor", "channel")]
-
-rad %>% 
-  .[qc == 0 & iuse == 1] %>% 
-  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
-  .[, .(omb = mean(tbc),
-        n = .N), by = .(date, lev)] %>% 
-  .[lev != 1025] %>% 
-  ggplot(aes(date, fct_rev(lev))) +
-  geom_tile(aes(fill = omb)) +
-  scale_fill_divergent(trans = mesoda::symlog_trans(),
-                       guide = guide_colourbar(barheight = 18))  +
-  # scale_y_level() +
-  scale_date() +
-  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H:%M") +
-  labs(x = NULL, fill = NULL, y = "pressure (hPa)",
-       caption = "OmB") +
-  theme_minimal()
-
-rad %>% 
-  .[qc == 0 & iuse == 1] %>% 
-  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
-  .[, .(omb = mean(tbc),
-        n = .N), by = .(date, lev)] %>% 
-  .[lev != 1025] %>% 
-  ggplot(aes(date, fct_rev(lev))) +
-  geom_tile(aes(fill = n)) +
-  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
-                       guide = guide_colourbar(barheight = 18,
-                                               barwidth = 0.8)) +
-
-  scale_date() +
-  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
-  theme_minimal()
-
-```
-
-```{r level-rad, fig.cap="Nivel de presión asociado a las observaciones en cada ciclo de asimilación"}
-rad %>% 
-  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
-  .[sensor %in% c("airs", "iasi") &
-      # errinv %between% c(0.0000316227, 1000) & 
-      # qc == 0 & 
-      # iuse == 1 &
-      peakwt %between% c(0.001, 1200)] %>% 
-  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
-  ggplot(aes(channel, Q50)) +
-   # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
-   #             outlier.size = 0.5, size = 0.1) +
-  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
-  geom_line(color = "darkorange") +
-  scale_y_level(name = "Pressure (hPa)") +
-  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
-  labs(x = "", y = "") +
-  theme_minimal() +
-
-rad %>% 
-  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
-  .[!sensor %in% c("airs", "iasi") &
-      # errinv %between% c(0.0000316227, 1000) & 
-      # qc == 0 & 
-      # iuse == 1 &
-      peakwt %between% c(0.001, 1200)] %>% 
-  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
-  ggplot(aes(channel, Q50)) +
-   # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
-   #             outlier.size = 0.5, size = 0.1) +
-  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
-  geom_line(color = "darkorange") +
-  scale_y_level(name = "Pressure (hPa)") +
-  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
-  labs(y = "", x = "Channel index") +
-  theme_minimal() +
-  plot_layout(ncol = 1)
-```
-
-```{r rad-wv}
-files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121020000/diagfiles*/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121140000/diagfiles*/asim_iasi_metop-b*ensmean"))
-
-rad <- purrr::map(files, function(f) {
-
-  rad <- read_diag_rad(f, "E7") %>% 
-    .[, run := basename(dirname(f))] %>% 
-    .[, wavenumber := 1/(299792458/(freq*10000000))] %>% 
-    .[wavenumber %between% c(1367, 1422.25)] %>% 
-    satinfo[., on = c("sensor", "channel")]
-}) %>% rbindlist()
-
-rad %>% 
-  .[sensor == "iasi_metop-b" & 
-      channel == 335 & 
-      errinv %between% c(0.0000316227, 1000) & 
-      qc == 0 & 
-      iuse == 1 &
-      peakwt %between% c(0.001, 1200)] %>% 
-  .[, date := factor(date)] %>% 
-  ggplot(aes(ConvertLongitude(lon), lat)) +
-  geom_point(aes(color = tbc)) +
-  scale_color_divergent(trans = mesoda::symlog_trans(),
-                        guide = guide_colourbar(barheight = 15, barwidth = 1)) +
-  geom_mapa() +
-  facet_grid(run~date, labeller = labeller(run = c("diagfiles" = "OmB",
-                                                   "diagfiles_ana" = "OmA"))) +
-  labs(x = NULL, y = NULL, color = NULL) +
-  theme_minimal() +
-
-  rad %>% 
-  .[sensor == "iasi_metop-b" & 
-      channel == 335 & 
-      errinv %between% c(0.0000316227, 1000) & 
-      qc == 0 & 
-      iuse == 1 &
-      peakwt %between% c(0.001, 1200)] %>% 
-  dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 
-  .[, ":="(tbc = diagfiles - diagfiles_ana,
-           run = "AmB")] %>% 
-  .[, date := factor(date)] %>% 
-  ggplot(aes(ConvertLongitude(lon), lat)) +
-  geom_point(aes(color = tbc)) +
-  scale_color_divergent(guide = guide_colourbar(barheight = 15, barwidth = 1)) +
-  geom_mapa() +
-  labs(x = NULL, y = NULL, color = NULL) +
-  facet_grid(run~date) +
-  theme_minimal() +
-  plot_layout(ncol = 2, widths = c(0.4, 0.6))
-```
-```{r omb-qv-acum}
-
-global_qv_E6 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E6.rds")
-global_qv_E7 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E7.rds")
-
-rbind(global_qv_E6, global_qv_E7) %>%
-  # .[exp == "E7"] %>% 
-  .[, nivel := fcase(bottom_top %in% c(1:8), "lower",
-                     bottom_top %in% c(9:13), "middle",
-                     bottom_top %in% c(14:20), "upper",
-                     bottom_top %in% c(21:38), "21 to 38")] %>%
-  .[ bottom_top %between% c(1, 20)] %>% 
-  .[, acum := cumsum(qv), by = .(bottom_top, exp)] %>%
-  .[, acum := mean(acum), by = .(nivel, exp, date)] %>%
-  # .[bottom_top %in% c(1:18)] %>%
-  ggplot(aes(date, acum*1000)) +
-  geom_line(aes(color = exp, linetype = exp, group = interaction(nivel, exp))) +
-  # geom_label(data = ~.x[date %in% sample(unique(date), 38)] %>% group_by(exp, bottom_top) %>% sample_n(size = 1),
-   geom_label(data = ~.x[date %in% unique(date)[67]],
-             aes(label = nivel, color = exp, 
-                 group = interaction(nivel, exp)),
-             size = 3, fill = "white", show.legend = FALSE,
-             label.size = NA, label.padding = unit(0.1, "lines")) +
-  scale_date() +
-  scale_linetype_manual(values = c(2, 1), labels = c(E4 = "CONV", E5 = "AUT",
-                                     E6 = "SATWND", E7 = "RAD")) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  # facet_wrap(~nivel, scales = "free_y") +
-  labs(x = NULL, y = "Accumulated specific humidity (g/Kg)",
-       color = NULL, linetype = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-
-```
+Acumulados a 6 hs
 
 ```{r}
-files <- Sys.glob("~/datosmunin/EXP/E7/ANA/2018112*/diagfiles/asim*.ensmean")
-
-files <- files[str_detect(files, "conv", negate = TRUE)]
-
-rad <- read_diag_rad(files, "E7") %>% 
-  satinfo[., on = c("sensor", "channel")] %>% 
-  .[,  type :=  fifelse(sensor %in% multiespectrales, "multiespectral", "otro")] %>% 
-  separate(sensor, into = c("sensor", "plat"), sep = "_")
-
-rad[errinv %between% c(0.0000316227, 1000) & 
-      qc == 0 & 
-      iuse == 1 &
-      peakwt %between% c(0.001, 1200), 
-    .(tbc_mean = mean(tbc),
-      tbcnob_mean = mean(tbcnob)), by = .(sensor, channel, date)] %>% 
-  .[, diff := abs(tbcnob_mean) - abs(tbc_mean)] %>% 
-  # .[, range(diff, na.rm = TRUE)]
-  melt(id.vars = c("sensor", "channel", "date")) %>% 
-  ggplot(aes(value)) +
-  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
-                 alpha= 0.3, position = "identity") +
-  geom_density(aes(color = variable, y = stat(density))) +
-  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
-  scale_fill_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
-  scale_color_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
-  geom_vline(xintercept = 0, color = "grey40") +
-  coord_cartesian(xlim = c(-5, 5)) +
-  facet_wrap(~sensor, labeller = labeller(sensor = toupper)) +
-  labs(fill = NULL, x = NULL, color = NULL) +
-  theme_minimal()
-
-rad[errinv %between% c(0.0000316227, 1000) & 
-      qc == 0 & 
-      iuse == 1 &
-      peakwt %between% c(0.001, 1200), 
-    .(tbc_mean = mean(tbc),
-      tbcnob_mean = mean(tbcnob)), by = .(sensor, channel, date)] %>% 
-  .[, diff := abs(tbcnob_mean) - abs(tbc_mean)] %>% 
-  # .[, range(diff, na.rm = TRUE)]
-  melt(id.vars = c("sensor", "channel", "date")) %>% 
-  ggplot(aes(value)) +
-  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
-                 alpha= 0.3, position = "identity") +
-  geom_density(aes(color = variable)) +
-  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
-  scale_color_viridis_d(labels = c("OmB BC", "OmB NoBC", "Difference")) +
-  scale_fill_viridis_d(labels = c("OmB BC", "OmB NoBC", "Difference")) +
-  geom_vline(xintercept = 0, color = "grey40") +
-  coord_cartesian(xlim = c(-5, 5)) +
-  # facet_wrap(~sensor) +
-  labs(fill = NULL, x = NULL, color = NULL) +
-  theme_minimal()
-
-```
-
-```{r}
-rm(rad)
-```
-
-### Ensemble characteristics
-
-#### Reduced Centered Random Variable
-
-```{r rcrv-sfc, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree", fig.scap=c("Wind", "Temperature")}
-
-fread(here("analysis", "data", "derived_data", "uvt_rcrv_box.csv")) %>%   
-  .[var %in% c("u", "v") & type %in% c(280, 281, 287, 290)] %>% 
-  .[, type.obs := type] %>% 
-  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
-  geom_raster(aes(fill = sd.y)) +
-  geom_mapa() +
-  scale_fill_viridis_c(breaks = seq(0, 4, 0.5),
-                       limits = c(0, 4),
-                       direction = -1,
-                       guide = guide_colorstrip(inside = TRUE,
-                                                barwidth = 20,
-                                                barheight = 0.8)) +
-  facet_grid(var~type.obs, 
-             labeller = labeller(type.obs = c("280" = "SFCSHP",
-                                              "281" = "ADPSFC",
-                                              "287" = "ADPSFC\n no pressure",
-                                              "290" = "ASCATW"),
-                                 var = c("u" = "U wind",
-                                         "v" = "V wind" ))) +
-  labs(fill = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-fread(here("analysis", "data", "derived_data", "uvt_rcrv_box.csv")) %>%
-  .[var %in% c("t") & type %in% c(180, 181, 187)] %>% 
-  .[, type.obs := type] %>% 
-  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
-  geom_raster(aes(fill = sd.y)) +
-  geom_mapa() +
-  scale_fill_viridis_c(breaks = seq(0, 5, 0.5),
-                       limits = c(0, 5),
-                       direction = -1,
-                       guide = guide_colorstrip(inside = TRUE,
-                                                barwidth = 20,
-                                                barheight = 0.8)) +
-  facet_grid(var~type.obs, labeller = labeller(type.obs = c("180" = "SFCSHP",
-                                                            "181" = "ADPSFC",
-                                                            "187" = "ADPSFC\n no pressure"),
-                                               var = c("t" = "Temperature"))) +
-  labs(fill = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r rcrv-profile, fig.cap="Profiles of mean and standar deviation of rcrv for sounding and aircraft observations."}
-obs.type[fread(here("analysis", "data", "derived_data", "rcrv_uvt_perfil.csv")), on = "type"] %>% 
-  .[type != 130] %>%
-  melt(measure.vars = c("mean.y", "sd.y")) %>% 
-  ggplot(aes(nivel.error, value)) +
-  geom_hline(yintercept = c(0, 1), color = "grey20") +
-  geom_line(aes(color = factor(code), linetype = variable)) +
-  scale_linetype_manual(values = c("dashed", "solid"),
-                        labels = c("Mean RCRV", "SD RCRV")) +
-  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
-  coord_flip() +
-  facet_wrap(~ var, labeller = labeller(var = c("t" = "Temperature",
-                                                "u" = "U wind",
-                                                "v" = "V wind"))) +
-  labs(y = NULL,
-       linetype = "",
-       color = NULL) +
-  theme_minimal() 
-```
-
-```{r rcrv-profile-rad, fig.cap="Profiles of mean and standar deviation of rcrv for radiances observations."}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:4]
-
-RCRV <- purrr::map(files, function(f) {
-  fread(f)
-}) %>% 
-  rbindlist()
-
-RCRV %>% 
-  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
-  setDT() %>% 
-  melt(measure.vars = c("mean.y", "sd.y")) %>% 
-  .[, .(value = mean(value, na.rm = TRUE)), by = .(sensor, level, variable)] %>% 
-  ggplot(aes(level, value)) +
-  geom_hline(yintercept = c(0, 1), color = "grey20") +
-  geom_line(aes(color = sensor, linetype = variable)) +
-  scale_color_discrete(labels = toupper) +
-  scale_linetype_manual(values = c("dashed", "solid"),
-                        labels = c("Mean RCRV", "SD RCRV")) +
-  coord_flip() +
-  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
-  labs(x = NULL, y = NULL,
-       color = NULL, linetype = NULL) +
-  theme_minimal()
-
-```
-
-### Analysis comparison
-
-To observe the impact of assimilating each source of observations on the vertical structure, the average vertical profiles for  the temperature, specific humidity and wind components were calculated over the grey box in Figure \@ref(fig:dominio).
-
-```{r T-diff, fig.cap="Temperature difference between experiments in Kelvin."}
-files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
-                    pattern = "perfiles_ana")
-
-perfiles <- purrr::map(files, function(f) {
-  exp <- unglue(basename(f), "perfiles_{run}_{exp}.csv")
-  fread(f) %>% 
-    .[, exp := exp[[1]][["exp"]]]
-
-}) %>% 
-  rbindlist() %>% 
-  .[, date := as_datetime(date)] %>% 
-  .[date %between% c(as_datetime("2018-11-20 18:00:00"),
-                     as_datetime("2018-11-23 12:00:00"))]
-
-perfiles %>% 
-  dcast(lev + date ~ exp, value.var = "T") %>% 
-  .[, ":="(E5_E4 = E5 - E4,
-           E6_E5 = E6 - E5,
-           E7_E6 = E7 - E6)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
-  ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value, fill = stat(level))) +
-  scale_fill_divergent_discretised(name = NULL,
-                                   guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
-  labs(subtitle = "Temperature difference (K)",
-       x = NULL,
-       y = NULL) +
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  # scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
-                                                         "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-AUT-CONV present a negative difference in low levels with a clear diurnal cycle (Figure \@ref(fig:T-diff)).
-This negative difference indicate that AUT is warmer than CONV.
-It also shows significant differences in the upper levels of the troposphere during the hours of maximun convection.
-The temperature profiles of AUT y SATWND are very similar.
-The difference bewteen the two experiments is smaller than +-0.1 until Nov, 22th 00Z when the convection starts to develope on the domain.
-The differences between RAD y SATWIND (and AUT) are smaller important in low levels at the begining of the experiment and in upper levels at the end indicating that the satellite observations have impact in several levels.
-
-```{r T-mse, eval=FALSE, fig.cap="MSE", include=FALSE}
-
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/perfiles_mse*")
-
-perfiles_mse <- purrr::map(files, function(f) {
-  exp <- unglue(basename(f), "perfiles_{run}_{exp}.csv")
-  fread(f) %>% 
-    .[, run := exp[[1]][["run"]]] %>% 
-    .[, exp := exp[[1]][["exp"]]]
-
-}) %>% 
-  rbindlist() %>% 
-  .[, date := as_datetime(date)] %>% 
-  .[date %between% c(as_datetime("2018-11-20 20:00:00"),
-                     as_datetime("2018-11-23 12:00:00"))]
-
-perfiles_mse %>% 
-  .[variable == "T"] %>% 
-  ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = mse, fill = stat(level))) +
-  scale_fill_viridis_c(option = "A", direction = -1,
-                       name = NULL, super = ScaleDiscretised,
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  labs(subtitle = "Temperature MSE (K^2)",
-       x = NULL,
-       y = "Pressure level") +
-  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r Q-diff, fig.cap="Specific humidity difference between experiments in g/Kg"}
-perfiles %>% 
-  dcast(lev + date ~ exp, value.var = "QVAPOR") %>% 
-  .[, ":="(E5_E4 = E5 - E4,
-           E6_E5 = E6 - E5,
-           E7_E6 = E7 - E5)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
-  ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value*1000, fill = stat(level)), 
-                    breaks = c(seq(-1, 1.6, 0.2), Inf)) +
-  scale_fill_divergent_discretised(name = NULL, limits = c(-1, 1.6),
-                                   guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
-
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  labs(subtitle = "Specific humidity difference (g/Kg)",
-       y = NULL,
-       x = NULL) +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
-                                                         "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-Comparing the specific humidity (Figure \@ref(fig:Q-diff)) once again the difference between AUT and CONV is important in low levels showing the impact of assimilating the observations from automatic stations with larger spacial and temporal resolution.
-Between SATWND and AUT the differences are smaller and only important after Nov, 21st 18Z.
-The addition of satellite derived winds produce and increse of humidity around 750 hPa during Nov 21st (0.05 to 0.10 g/kg) and a decrese of humidity between surface and 750 hPa starting at 00Z Nov 22th until the end of the experiment.
-
-The differences between RAD and SATWND and therefore the impact of the radiances have opposite sign to the impact of the observations from automatic stations but with smaller magnituded in low levels.
-A positive difference can also be observed between 850 and 600 hPa.
-```{r Q-mse, eval=FALSE, fig.cap="MSE", include=FALSE}
-
-perfiles_mse %>% 
-  .[variable == "QVAPOR"] %>% 
-  ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = mse*1000000, fill = stat(level))) +
-  scale_fill_viridis_c(option = "A", direction = -1,
-                       name = NULL, super = ScaleDiscretised,
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  labs(subtitle = "Specific humidity MSE (g/Kg^2)",
-       x = NULL,
-       y = "Pressure level") +
-  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r U-diff, fig.cap="Zonal wind difference between experiments in m/s (filled contours) and zonal wind profile for the experiment with more observations."}
-perfiles %>% 
-  dcast(lev + date ~ exp, value.var = "U") %>% 
-  .[, ":="(E5_E4 = E5 - E4,
-           E6_E5 = E6 - E5,
-           E7_E6 = E7 - E6)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
-  ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value, fill = stat(level)), breaks = c(seq(-1.6, 2, 0.2), Inf)) +
-  scale_fill_divergent_discretised(name = NULL,
-                                   limits = c(-1.6, 2),
-                                   guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
-  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7)] %>% 
-                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                breaks = seq(0, 30, 3), size = 0.2, linetype = 1, color = "grey30") +
-  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
-                      melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                    breaks = seq(0, 30, 3), color = "grey30", skip = 1, 
-                    size = 4, stroke = 0.1) +
-  labs(subtitle = "U difference (m/s)",
-       x = NULL,
-       y = NULL) +
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
-                                                         "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-In Figure \@ref(fig:U-diff) the difference between the experiments for the zonal wind is shown with the vertical profile for the experiment with more observations.
-For example the left panel shows the difference between AUT and CONV experiments and the profile wind for AUT.
-The wind profiles for the experiments are similar but there are important differences that can be seen on the filled contours.
-
-The differences between AUT and CONV are important in low levels, for the first two days the impact of assimilation observations from automatic stations is a disminution on the wind from the west but an increse of the wind from the west after Nov 22th, 18Z.
-In upper levels the sign of the impact varies throughout the experiment period.
-The assimilation of satellite derived winds produces an impact in midle and upper levels and the incorporation of radiances prodeces impacts on low and upper levels.
-In low levels the impact has the oposite sign comparing with AUT-CONV, this means that RAD experiment shows on average less westerly wind.
-
-```{r V-diff, fig.cap="Meridial wind difference between experiments in m/s"}
-perfiles %>% 
-  dcast(lev + date ~ exp, value.var = "V") %>% 
-  .[, ":="(E5_E4 = E5 - E4,
-           E6_E5 = E6 - E5,
-           E7_E6 = E7 - E6)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
-  ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value, fill = stat(level)), 
-                    breaks = c(-Inf, seq(-1.6, 1.2, 0.2), Inf)) +
-  scale_fill_divergent_discretised(name = NULL, limits = c(-1.6, 1.2),
-                                   guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
-  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7)] %>% 
-                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                breaks = seq(-9, 0, 1), size = 0.2, linetype = 2, color = "grey30") +
-  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
-                      melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                    breaks = seq(-9, 0, 1), color = "grey30", skip = 1, 
-                    size = 4, stroke = 0.1) +
-  labs(subtitle = "V difference (m/s)",
-       x = NULL,
-       y = NULL) +
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
-  scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
-                                                         "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-```
-
-Figure \@ref(fig:V-diff) is constructed in the same way as Figure \@ref(fig:U-diff) but here only negative meridional wind is ploted.
-Again the wind profiles are similar for each experiment but nevertheless there are differences of +- 1 m/s.
-
-In upper levels and during Nov 22nd and 23rd the impact of assimilating new observations (mainly from automatic stations and radiances) produces an increse of northerly wind.
-In low levels the impact has different sign in AUT-CONV and RAD-SATWND.
-
-#### Low levels humidity
-
-Specific humidity field averaged over the first 7 levels of the model, from surface to 800 hPa for low topography areas (aproximately).
-
-* AUT add humidity to the model (is correcting a dry model bias or modifing the circulation and transporting humidity from the north?)
-* SATWND is not that different from AUT at list in low levels. The humidity is distributed a little bit different in the domain during 06Z
-* RAD shows less humidity that AUT and SATWND but more than CONV.
-RAD seems to reverse (in part) the effect of incorporating automatic station observations in low levels.
-This can also be seeing in the first levels the difference between RAD and SATWND.
-
-```{r q-low, eval=FALSE, fig.align="center", message=FALSE, warning=FALSE, include=FALSE}
-fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
-                     exp = c("E4", "E5", "E6", "E7"))
-
-q <- lapply(seq_len(nrow(fcsts)), function(f) {
-
-  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
-    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
-    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
-    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
-    .[, exp := fcsts[f, 2]] %>%  
-    .[]
-}) %>% 
-  rbindlist()
-
-q[bottom_top %between% c(0, 7), .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
-  by = .(south_north, west_east, date, exp)] %>% 
-  .[coord, on = c("south_north", "west_east")] %>% 
-  .[, date := factor(date)] %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = q*1000, fill = stat(level_d)), proj = norargentina_lambert,
-                    breaks = seq(0, 0.02, 0.002)*1000) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "BrBG", direction = 1,
-                       limits = c(1, 19),
-                       guide = guide_colorsteps(#inside = FALSE,
-                         barwidth = 25,
-                         barheight = 0.8)) +
-  # geom_vector(aes(x = lon, y = lat,
-  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
-  #             arrow.angle = 12, size = 0.3) +
-  # scale_mag() +
-  cordillera +
-  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
-  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
-  facet_grid(date ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(title = "Low level humidity (g/Kg)",
-       x = NULL,
-       y = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r q-middle, eval=FALSE, include=FALSE}
-
-q[bottom_top %between% c(8, 12), .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
-  by = .(south_north, west_east, date, exp)] %>% 
-  .[coord, on = c("south_north", "west_east")] %>% 
-  .[, date := factor(date)] %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = q*1000, fill = stat(level_d)), proj = norargentina_lambert,
-                    breaks = seq(0, 0.02, 0.002)*1000) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "BrBG", direction = 1,
-                       limits = c(1, 19),
-                       guide = guide_colorsteps(#inside = FALSE,
-                         barwidth = 25,
-                         barheight = 0.8)) +
-  # geom_vector(aes(x = lon, y = lat,
-  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
-  #             arrow.angle = 12, size = 0.3) +
-  # scale_mag() +
-  cordillera +
-  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
-  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
-  facet_grid(date ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(title = "Humidity (g/Kg) 800-700 hPa",
-       x = NULL,
-       y = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-```{r v-low, eval=FALSE, include=FALSE}
-q[bottom_top %between% c(0, 7), .(v = mean(v)), 
-  by = .(south_north, west_east, date, exp)] %>% 
-  .[coord, on = c("south_north", "west_east")] %>% 
-  .[, date := factor(date)] %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = v, fill = stat(level_d)), proj = norargentina_lambert) +
-  scale_fill_divergent_discretised(name = NULL,
-                                   # limits = c(1, 19),
-                                   guide = guide_colorsteps(#inside = FALSE,
-                                     barwidth = 25,
-                                     barheight = 0.8)) +
-  # geom_vector(aes(x = lon, y = lat,
-  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
-  #             arrow.angle = 12, size = 0.3) +
-  # scale_mag() +
-  cordillera +
-  geom_mapa() +
-  facet_grid(date ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  labs(title = "Low level meridional wind (m/s)",
-       x = NULL,
-       y = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r v-middle, eval=FALSE, include=FALSE}
-q[bottom_top %between% c(8, 12), .(v = mean(v)), 
-  by = .(south_north, west_east, date, exp)] %>% 
-  .[coord, on = c("south_north", "west_east")] %>% 
-  .[, date := factor(date)] %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = v, fill = stat(level_d)), proj = norargentina_lambert) +
-  scale_fill_divergent_discretised(name = NULL,
-                                   # limits = c(1, 19),
-                                   guide = guide_colorsteps(#inside = FALSE,
-                                     barwidth = 25,
-                                     barheight = 0.8)) +
-  # geom_vector(aes(x = lon, y = lat,
-  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
-  #             arrow.angle = 12, size = 0.3) +
-  # scale_mag() +
-  cordillera +
-  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
-  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
-  facet_grid(date ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(title = "800-700 hPa meridional wind (m/s)",
-       x = NULL,
-       y = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r temp, eval=FALSE, include=FALSE}
-fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
-                     exp = c("E4", "E5", "E6", "E7"))
-
-t <- lapply(seq_len(nrow(fcsts)), function(f) {
-
-  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
-    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
-    .[, t := ReadNetCDF(paste0(derived_data, "/t_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(t = "temp"), out = "vector")] %>% 
-    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
-    .[, exp := fcsts[f, 2]] %>%  
-    .[]
-}) %>% 
-  rbindlist()
-
-t[date == unique(date)[1:3] & west_east %between% c(70, 180)] %>% 
-  .[, .(t_mean = mean(t),
-        u_mean = mean(u),
-        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
-  .[, ":="(dir = atan2(v_mean, u_mean)*180/pi,
-           mag = Mag(u_mean, v_mean))] %>%
-  .[, date := factor(date)] %>% 
-  ggplot(aes(south_north, bottom_top)) +
-  geom_contour_fill(aes(z = t_mean, fill = stat(level_d))) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "OrRd", direction = 1,
-                       # limits = c(1, 19),
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  # geom_contour(aes(z = v_mean, linetype = factor(-sign(..level..))), 
-  #              color = "grey30", size = 0.3, binwidth = 2) +
-  # geom_text_contour(aes(z = v_mean), color = "grey30", size = 4, stroke = 0.1, breaks = seq(-10, 6, 2)) +
-  # scale_linetype(guide = NULL) +
-  geom_arrow(aes(mag = mag, angle = dir),
-             arrow.angle = 8, size = 0.2, skip.x = 10, skip.y = 3) +
-  scale_mag() +
-  facet_grid(date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                     E6 = "SATWND", E7 = "RAD"))) +
-  labs(y =  "Sigma levels", x = "Grid points") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-```
-
-```{r mucape, eval=FALSE, include=FALSE}
-files <- Sys.glob(paste0(derived_data, "mucape/*"))
-
-files <- files[!str_detect(files, c("E3"))]
-
-dates <- c("20181121180000", "20181122000000", "20181122060000")
-
-mucape <- purrr::map(files, function(f) {
-  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]
-
-  if (details$date %in% dates) {
-
-    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
-      .[, `:=`(exp = details$exp, 
-               date = lubridate::ymd_hms(details$date))] %>%
-      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
-      .[]
-  }
-}) %>% 
-  rbindlist()
-
-mucape %>% 
-  .[variable == "mcape"] %>% 
-  # .[coord, on = c("south_north", "west_east")] %>% 
-  .[, date := factor(date)] %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = value, fill = stat(level_d)), proj = norargentina_lambert) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "YlOrRd", direction = 1,
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  # cordillera +
-  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
-  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
-  facet_grid(date ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(title = "Maximum CAPE",
-       x = NULL,
-       y = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-```
-
-```{r summary, fig.height=10, fig.width=9}
-files <- Sys.glob(paste0(derived_data, "pw/*")) 
-files <- files[!str_detect(files, c("E3"))]
-
-dates <- c("20181122000000")
-
-pw <- purrr::map(files, function(f) {
-  details <- unglue::unglue(basename(f), "pw_ana_{exp}_{date}.nc")[[1]]
-
-  if (details$date %in% dates) {
-
-    ReadNetCDF(f, vars = c(value = "pw", lon = "XLONG", lat = "XLAT")) %>% 
-      .[, `:=`(exp = details$exp, 
-               date = lubridate::ymd_hms(details$date))] %>%
-      .[]
-  }
-}) %>% 
-  rbindlist() %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)]
-
-files <- Sys.glob(paste0(derived_data, "mucape/*")) 
-files <- files[!str_detect(files, c("E3"))]
-
-mucape <- purrr::map(files, function(f) {
-  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]
-
-  if (details$date %in% dates) {
-
-    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
-      .[, `:=`(exp = details$exp, 
-               date = lubridate::ymd_hms(details$date))] %>%
-      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
-      .[]
-  }
-}) %>% 
-  rbindlist()
-
-fcsts <- expand_grid(fcsts = c("20181122000000"),
-                     exp = c("E4", "E5", "E6", "E7"))
-
-ana <- lapply(seq_len(nrow(fcsts)), function(f) {
-
-  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
-    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
-    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
-    .[, theta := ReadNetCDF(paste0(derived_data, "/theta_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(th = "theta"), out = "vector")] %>%
-    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
-    .[, exp := fcsts[f, 2]] %>%  
-    .[]
-}) %>% 
-  rbindlist()
-
-cor <- ana[bottom_top %in% c(2, 10)] %>% 
-  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
-  .[, cortante := Mag(u_10 - u_2, v_10 - v_2)] %>% 
-  .[, ":="(u_2 = NULL, 
-           u_10 = NULL,
-           v_2 = NULL,
-           v_10 = NULL)]
-
-ana[bottom_top %between% c(2, 7), .(v_mean = mean(v)), 
-  by = .(south_north, west_east, date, exp)] %>% 
-  .[pw, on = .NATURAL] %>% 
-ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = value, fill = stat(level_d)),
-                    proj = norargentina_lambert) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "YlGn", direction = 1,
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  geom_contour2(aes(z = v_mean, size = stat(level)),
-                proj = norargentina_lambert, color = "darkorange", 
-                breaks = c(-10, -5)) +
-  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
-                        labels = c("-10 m/s", "-5 m/s"),
-                        name = NULL) +
-   # geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
-   #                   proj = norargentina_lambert, color = "darkorange", 
-   #              breaks = c(-10, -5), skip = 0,
-   #              label.placement = label_placement_n(1)) +
-  # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
-  cordillera +
-  geom_mapa() +
-  facet_grid(. ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        # legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3)) +
-
-ana[bottom_top %between% c(2, 10), .(theta_mean = mean(theta)), 
-  by = .(south_north, west_east, date, exp)] %>% 
-   .[pw, on = .NATURAL] %>% 
-ggplot(aes(x, y)) +
- # geom_point(aes(color = v_mean)) +
-  geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
-                    proj = norargentina_lambert,
-                    breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "RdYlBu", direction = -1,
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  cordillera +
-  geom_mapa() +
-  facet_grid(. ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        # legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3),
-        strip.text = element_blank(), 
-        strip.background = element_blank()) +
-
-mucape %>% 
-  .[variable == "mcape"] %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-  .[cor, on = .NATURAL] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = value, fill = stat(level_d)), 
-                    proj = norargentina_lambert, limits = c(500, 6000)) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "YlOrRd", direction = 1,
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  # geom_contour2(aes(z = cortante),
-  #               proj = norargentina_lambert, color = "black", size = 0.4,
-  #               breaks = c(10, 25)) +
-  geom_point(data = ~.x[cortante >= 10 & is.cross(x, y, skip = 1)], 
-             aes(x = lon, y = lat, 
-                 alpha = factor(cut_round(cortante, breaks = c(5, 15, 40)))),
-             size = 0.00002, shape = 20) +
-  scale_alpha_manual(values = c(0.3, 0.7), name = NULL, labels = c(">5 m/s", ">15 m/s"),
-                     guide = guide_legend(override.aes = list(size = 2))) +
-  cordillera +
-  geom_mapa() +
-  facet_grid(. ~ exp, 
-             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                         E6 = "SATWND", E7 = "RAD"))) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        # legend.box.spacing = margin(1, 1, 1, 1),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3),
-        strip.text = element_blank(), 
-        strip.background = element_blank()) +
-
-    plot_layout(ncol = 1, heights = c(1, 1, 1)) +
-  plot_annotation(tag_levels = "a", tag_suffix = ")")
-```
-
-```{r}
-coord <- mucape[variable == "mcape", c("south_north", "lat")] %>% 
-  .[, .(lat = mean(lat)), by = south_north]
-
-ana[west_east %between% c(70, 180) & bottom_top %between% c(1, 25)] %>% 
-  .[, .(q_mean = mean(q),
-        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
-  coord[., on = "south_north"] %>% 
-  ggplot(aes(lat, bottom_top)) +
-  geom_contour_fill(aes(z = q_mean*1000, fill = stat(level_d))) +
-  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
-                       palette = "BrBG", direction = 1,
-                       # limits = c(1, 19),
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  geom_contour2(aes(z = v_mean, linetype = factor(-sign(..level..))), 
-               color = "grey30", size = 0.3, binwidth = 2) +
-  geom_text_contour(aes(z = v_mean), color = "grey30", 
-                    size = 3, stroke = 0.05, breaks = seq(-10, 6, 2),
-                    label.placement = label_placement_n(1)) +
-  scale_linetype(guide = NULL) +
-  scale_x_latitude(breaks = seq(-40, -20, 5)) +
-  facet_wrap( ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                     E6 = "SATWND", E7 = "RAD"))) +
-  labs(y =  "Sigma levels") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3)) 
-```
-
-Is the analysis correcting a dry bias on the model or the humidity is being transported into the domain?
-
-The innovation (OmB) for surface humidity observations is mainly positive indicating that the model has in general less humidity than the atmosfere throghout the domain for a specific cycle.
-
-```{r omb-q, fig.cap="Observation minus guess for humidity surface observations during Nov, 21st 18Z"}
-
-exp <- c("E4", "E5", "E6", "E7")
-
-q_field <- purrr::map(exp, function(f) {  
-  f
-  read_diag_conv(paste0("/home/paola.corrales/datosmunin/EXP/", f,
-                        "/ANA/20181121180000/diagfiles/asim_conv_20181121180000.ensmean"), 
-                 exp = f, member = "000", variable = c("q"))
-}) %>% rbindlist()
-
-q_field[usage.flag == 1 & type %in% c(181, 187) & obs.guess < 0.02] %>% 
-  ggplot(aes(ConvertLongitude(lon), lat)) +
-  geom_mapa(fill = "grey90") +
-  geom_point(aes(color = obs.guess)) +
-  scale_color_divergent(guide = guide_colorbar(barwidth = 25,
-                                               barheight = 0.8)) +
-
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  facet_wrap(~exp, ncol = 4, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                         E6 = "SATWND", E7 = "RAD"))) +
-  labs(x = NULL, y = NULL, color = "OmB") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-```
-
-```{r omb-t, fig.cap="Observation minus guess for humidity surface observations during Nov, 21st 18Z"}
-
-exp <- c("E4", "E5", "E6", "E7")
-
-q_field <- purrr::map(exp, function(f) {  
-  f
-  read_diag_conv(paste0("/home/paola.corrales/datosmunin/EXP/", f,
-                        "/ANA/20181121180000/diagfiles/asim_conv_20181121180000.ensmean"), 
-                 exp = f, member = "000", variable = c("t"))
-}) %>% rbindlist()
-
-q_field[usage.flag == 1 & type %in% c(181, 187) & abs(obs.guess) < 15] %>% 
-  ggplot(aes(ConvertLongitude(lon), lat)) +
-  geom_mapa(fill = "grey90") +
-  geom_point(aes(color = obs.guess)) +
-  scale_color_divergent(guide = guide_colorbar(barwidth = 25,
-                                               barheight = 0.8)) +
-
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  facet_wrap(~exp, ncol = 4, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                         E6 = "SATWND", E7 = "RAD"))) +
-  labs(x = NULL, y = NULL, color = "OmB") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-```
-
-The evolution of the mean of the absolute OmB is always positive for all four experiments, AUT and SATWND showing higher values follow by RAD.
-The size of the points shows that AUT, SATWND and RAD assimilate more observations.
-
-This result is indicating that the model has an important dry bias and the assimilation of high temporal and spatial resolution observations are correcting this.
-
-TODO: it's the same for all PBL parameterizations?
-
-```{r omb-q-evol, eval=FALSE, fig.cap="Evolution", include=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E[4-7]/ANA/*/diagfiles/asim_conv_201811*.ensmean")
-
-conv <- purrr::map(files, function(f) {
-
-  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.ensmean")
-  read_diag_conv(f, exp = meta[[1]][["exp"]], member = "000", variable = c("q")) %>% 
-    .[stationID %in% oficiales$stationID] %>% 
-    .[]
-}) %>% rbindlist()
-
-conv[type %in% c(181, 187) & exp %in% c("E5", "E7") & usage.flag == 1, 
-     .(mean_omb = mean(obs.guess, na.rm = TRUE),
-       n = .N), 
-     by = .(date, exp, usage.flag)] %>% 
-  ggplot(aes(date, mean_omb)) +
-  geom_line(aes(color = exp)) +
-  geom_point(aes(color = exp, size = n)) +
-  scale_size_area(max_size = 2) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  # facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
-  #                                                            "-1" = "rejected"))) +
-  scale_date(20181120180000, 20181123120000) +
-  labs(y = "Mean of OmB (g/g)", x = NULL,
-       color = NULL, size = "number\n of obs") +
-  theme_minimal()
-
-rm(conv)
-
-# conv[usage.flag == 1 & type %in% c(181, 187) & exp %in% experiments] %>% 
-#   ggplot(aes(date, obs.guess)) +
-#   geom_point(aes(color = exp, group = date), alpha = 0.1) +
-# scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-#                                           E6 = "SATWND", E7 = "RAD")) +
-#   geom_smooth() +
-#   facet_wrap(~exp) +
-#   labs(y = "Mean of OmB (g/g)", x = NULL,
-#        color = NULL, size = "n obs") +
-#   theme_minimal()
-# ,
-#      .(n = .N), 
-#      by = .(date, exp)] %>% 
-#   ggplot(aes(date, n)) +
-#   geom_line(aes(color = exp)) +
-#   # geom_point(aes(color = exp, size = n)) +
-#   # scale_size_area(max_size = 3) +
-#   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-#                                           E6 = "SATWND", E7 = "RAD")) +
-#   labs(y = "Mean of OmB (g/g)", x = NULL,
-#        color = NULL, size = "n obs") +
-#   theme_minimal()
-
-```
-
-```{r omb-q-evol-param, eval=FALSE, fig.cap="Evolution", include=FALSE}
-files <- c(Sys.glob("~/datosmunin/EXP/E5/ANA/*/diagfiles/asim_conv_201811*.mem*"),
-           Sys.glob("~/datosmunin/EXP/E7/ANA/*/diagfiles/asim_conv_201811*.mem*"))
-
-conv <- purrr::map(files, function(f) {
-
-  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.mem{mem}")
-  read_diag_conv(f, exp = meta[[1]][["exp"]], member = meta[[1]][["mem"]], variable = c("q")) %>% 
-    .[stationID %in% oficiales$stationID] %>% 
-    .[]
-}) %>% rbindlist() %>% 
-  fisica[., on = "mem"] %>% 
-  separate(fisica,  into = c("conv", "pbl"))
-
-conv[type %in% c(181, 187) & exp %in% c("E5", "E7") & usage.flag == 1, 
-     .(mean_omb = mean(obs.guess, na.rm = TRUE),
-       n = .N), 
-     by = .(date, exp, pbl, usage.flag)] %>% 
-  ggplot(aes(date, mean_omb)) +
-  geom_line(aes(color = exp, linetype = pbl)) +
-  geom_point(aes(color = exp, size = n/60)) +
-  scale_size_area(max_size = 2) +
-  scale_linetype_manual(values = c(2, 1, 4)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  # facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
-  #                                                            "-1" = "rejected"))) +
-  scale_date(20181120180000, 20181123120000) +
-  labs(y = "Mean of OmB (g/g)", x = NULL, linetype = NULL,
-       color = NULL, size = "number\n of obs") +
-  theme_minimal()
-
-rm(conv)
-```
-
-The negative difference between RAD and SATWND in low levels (\@ref(fig:Q-diff)) and positive values over those levels can be explain by a deeper boundary layer.
-To explore this posibility Figure \@ref(fig:pbl-18Z) shows the zonal mean high of the PBL over the subdomain estimated for the PBL parameterization for Nov 21st, 18Z.
-The mean is calculated over a big area with different hights than can altered the result but but still the hight of the boundary layer for RAD is greater than in the other experiments in many areas.
-
-```{r pbl-18Z, eval=FALSE, fig.cap="Zonal mean high of the PBL over the subdomain estimated for the PBL parameterization for Nov 21st", include=FALSE}
-pblh18 <- purrr::map(exp, function(f) {
-
-  ReadNetCDF(paste0("/home/paola.corrales/datosmunin/EXP/", f, "/ANA/20181121180000/analysis.ensmean"), 
-             vars = c("PBLH", lon = "XLONG", lat = "XLAT"),
-             subset = list(south_north = c(41, 119),
-                           west_east = c(96, 143))) %>% 
-    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-    .[, exp := f]
-
-}) %>% 
-  rbindlist()
-
-pblh18[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
-  .[, .(mean = mean(PBLH)), by = .(exp, south_north)] %>% 
-  ggplot(aes(south_north, mean)) +
-  geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  labs(x = "Grid points (south - north)", y = "mean PBLH (m)", color = NULL) +
-  theme_minimal()
-```
-
-If the average is calculated over a smaller domain to avoid hight terrain and water, it is posible to see the evolution of the PBL hight during the experiment (Figure \@ref(fig:pblh-evol)).
-In CONV and RAD the PBL is higher during the diurnal hours of Nov, 21st with, exceeding AUT and SATWND by about 200 meters.
-These hours coincide with the time when the highest humidity above 800 hPa in RAD is observed.
-After that, the behavior chances and the development of the convection over the domain starts to affect the PBL.
-
-The deeper PBL in RAD can explain the transport of humidity to higher levels. This efect does not appear in CONV because there is no much humidity to transport to upper levels.
-
-TODO: Are there RAD observations on those levels to explain the humidity besides the transport from low levels?
-
-```{r pblh, eval=FALSE, fig.cap="", include=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E[4-7]/ANA/*/analysis.ensmean")
-
-pblh <- purrr::map(files, function(f) {
-  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{date}/analysis.ensmean")
-
-  ReadNetCDF(paste0(f), 
-             vars = c("PBLH", lon = "XLONG", lat = "XLAT"),
-             subset = list(south_north = c(41, 119),
-                           west_east = c(96, 143))) %>% 
-    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-    .[, exp := meta[[1]][["exp"]]] %>% 
-    .[, date := ymd_hms(meta[[1]][["date"]])]
-
-}) %>% 
-  rbindlist()
-
-# pblh[lat %between% c(-38, -31) & lon %between% c(-64, -59)] %>%
-#   ggplot(aes(x, y)) +
-#   geom_contour_fill(aes(z =  PBLH), proj = norargentina_lambert,
-#                     breaks = seq(0, 5000, 500)) +
-#   scale_fill_viridis_b(breaks = seq(0, 5000, 500)) +
-#   geom_mapa() +
-#   facet_wrap(~ exp) +
-#   theme_minimal()
-
-pblh[, .(mean = mean(PBLH)), by = .(exp, date)] %>% 
-  ggplot(aes(date, mean)) +
-  geom_line(aes(color = exp)) +
-  geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  labs(x = NULL, y = "mean PBLH (m)", color = NULL) +
-  theme_minimal()
-```
-
-### Verification
-
-```{r pp-hov, fig.cap="Hövmoller of hourly acumulated precipitation (for ensemble mean) over the inner domain."}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*mean_acum_1h.rds")
-
-ppacum_1h <- purrr::map(files, function(f) {
-
-  readRDS(f) %>% 
-    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)]
-
-}) %>% rbindlist() %>% 
-  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)]
-
-readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
-  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
-  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
-  .[end_date %between% c(as_datetime("20181120190000"), as_datetime("20181123120000"))] %>% 
-  .[, .(pp_acum = mean(pp_acum),
-        lat = mean(lat),
-        exp = "IMERG"), by = .(end_date, y)] %>% 
-  ggplot(aes(end_date, lat)) +
-  geom_contour2(aes(z = pp_acum, color = ..level..), 
-               breaks = seq(1, 12, 1)) +
-  scale_color_viridis_c(guide = NULL, breaks = seq(1, 12, 1), limits = c(1, 12)) +
-  scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  scale_y_latitude(breaks = seq(-40, -20, 5)) +
-  facet_wrap(~ exp) +
-  labs(x = NULL, y = NULL, color = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  ppacum_1h %>% 
-  .[, lat := lat[1], by = .(y)] %>% 
-  .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y, lat)] %>% 
-  ggplot(aes(date, lat)) +
-  # geom_raster(aes(fill = pp_acum)) +
-  geom_contour2(aes(z = pp_acum, color = ..level..),
-               breaks = seq(1, 12, 1)) +
-  scale_color_viridis_c(breaks = seq(1, 12, 1), limits = c(1, 12),
-                        guide = guide_colorstrip(inside = TRUE,
-                                                 barwidth = 25,
-                                                 barheight = 0.8)) +
-  scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  scale_y_latitude(breaks = seq(-40, -20, 5)) +
-  facet_wrap(~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                E6 = "SATWND", E7 = "RAD"))) +
-  labs(x = NULL, y = NULL, color = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(0.4, 0.6), guides = "collect") & theme(legend.position = 'bottom')
-```
-
-```{r eval=FALSE, include=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*prop*")
-
-ppacum_1h <- purrr::map(files, function(f) {
-
-  readRDS(f)
-
-}) %>% rbindlist() %>% 
-  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)]
-
-ppacum_1h %>% 
-  .[, .(prop = mean(prop)), by = .(exp, date, y, umbral)] %>% 
-  .[umbral %in% c(1, 5)] %>% 
-  ggplot(aes(date, y)) +
-  geom_contour_fill(aes(z = prop*100, fill = stat(level_d)),
-                    breaks = seq(0, 100, 10)) +
-  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")),
-                       limits = c(10, 100),
-                       na.value = "white",
-                       super = ScaleDiscretised,
-                       # breaks = seq(0, 100, 10),
-                       guide = guide_colorsteps(barwidth = 15,
-                                                barheight = 1)) +
-  # scale_color_viridis_c() +
-  scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  facet_grid(umbral ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                       E6 = "SATWND", E7 = "RAD"),
-                                               umbral = c("1" = "1 mm/h",
-                                                          "5" = "5 mm/h"))) +
-  labs(fill = "", x = "", y = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-```
-
-#### Reflectivity
-
-```{r dbz-mean}
-
-dates <- rep(c("20181122100000", "20181122220000"), 1, each = 4)
-files <- paste0(derived_data, "dbz/maxdbz_ana_", experiments, "_", dates, ".nc")
-
-dbz <- purrr::map(files, function(f){
-
-  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
-  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
-    .[, ":="(exp = meta[[1]][["exp"]],
-             date = ymd_hms(meta[[1]][["date"]]))] %>% 
-    .[, c("x", "y") := wrf_project(lon, lat)] 
-}) %>% 
-  rbindlist()
-
-dbz %>% 
-  .[, date := factor(date)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
-                    proj = norargentina_lambert,
-                    breaks = seq(-30, 75, 5)) +
-  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barwidth = 20, 
-                                           barheight = 0.8)) +
-  ggnewscale::new_scale_fill() +
-    geom_contour_fill(aes(z = max_dbz), fill = "white", 
-                    proj = norargentina_lambert,
-                    breaks = c(-Inf, 15)) +
-  geom_mapa() +
-  facet_grid(date~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                   E6 = "SATWND", E7 = "RAD"))) +
-  labs(x = NULL, y = NULL, fill = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-```
-
-```{r dbz-ens, fig.height=10, fig.width=7}
-
-files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[4-7]_0*"))
-files <- files[str_detect(files, "20181122000000", negate = TRUE)]
-
-dbz <- purrr::map(files, function(f){
-
-  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{mem}_{date}.nc")
-  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
-    .[, ":="(exp = meta[[1]][["exp"]],
-             mem = meta[[1]][["mem"]],
-             date = ymd_hms(meta[[1]][["date"]]))] %>% 
-    .[, c("x", "y") := wrf_project(lon, lat)] 
-}) %>% 
-  rbindlist() %>% 
-  fisica[., on = "mem"] %>% 
-  separate(fisica, into = c("cumulus", "pbl"))
-
-  dbz %>% 
-  .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
-    .[date %in% unique(date)[1:2]] %>% 
-    .[, date := factor(date)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
-                    proj = norargentina_lambert,
-                    breaks = seq(-30, 75, 5),) +
-  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
-                  guide = guide_colorsteps(barheight = 0.8, 
-                                           barwidth = 20)) +
-  ggnewscale::new_scale_fill() +
-  geom_contour_fill(aes(z = max_dbz), fill = "white", 
-                    proj = norargentina_lambert,
-                    breaks = c(-Inf, 10)) +
-  geom_mapa() +
-  ggh4x::facet_nested(date + cumulus ~ exp, 
-                      labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                   E6 = "SATWND", E7 = "RAD"))) +
-  labs(fill = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-  # rm(dbz)
-```
-#### FSS for the analysis
-
-```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-7]new.csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[4-7]_ens.csv")
 
@@
 
-  fread(f) %>% 
-    .[, date := as_datetime(date)]
-}) %>% 
-  rbindlist()
-
-fss %>% 
-  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
-  ggplot(aes(date, fss)) +
-  geom_hline(yintercept = 1, color = "darkgray") +
-  geom_line(aes(color = exp)) +
-  # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  scale_date(20181122000000, 20181123000000, "12 hours") +
-  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
-  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
-                                            "10" = "10 mm", "25" = "25 mm"),
-                                      w = c("5" = "50 km", "11" = "100 km", 
-                                            "51" = "500 km", "101" = "1000 km"))) +
-  labs(color = NULL, x = NULL, y = "FSS") +
-  theme_minimal()
-
-```
-#### Soundings
-
-```{r message=FALSE, warning=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/ANA/sondeos/*")
-
-sondeos <- purrr::map(files, function(f){
-
-  fread(f) %>% 
-    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
-    .[, launch_time := as_datetime(launch_time)]
-
-}) %>% 
-  rbindlist()
-
-IOP <- tribble(
-  ~iop, ~ini, ~end,
-  "IOP07", ymd_hms("20181121150000"), ymd_hms("20181121210000"),
-  "IOP08", ymd_hms("20181122140000"), ymd_hms("20181122200000")
-) %>% setDT()
-
-sondeos %>% 
-  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
-                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
-  .[variable == "t" & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
-  .[, oma := value - fcst_value] %>% 
-  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = oma), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  facet_grid(exp ~ iop, scales = "free_x", labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                                       E6 = "SATWND", E7 = "RAD"))) +
-  labs(color = "OmA", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  sondeos %>% 
-  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
-                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
-  .[variable == "v" & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
-  .[, oma := value - fcst_value] %>% 
-  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = oma), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  # scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  facet_grid(exp ~ iop, scales = "free_x", labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                                       E6 = "SATWND", E7 = "RAD"))) +
-  labs(color = "OmA", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(0.5, 0.5))
-```
-
-```{r}
-
-sondeos %>% 
-  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
-                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
-  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
-  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
-  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
-           BIAS = mean(value - fcst_value, na.rm = TRUE)), by = .(iop, exp, variable, lev)] %>% 
-  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
-  ggplot(aes(lev/1000, value.est)) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-```
-
-```{r}
-sondeos %>% 
-  .[variable == "td" & !str_detect(site, "/") & 
-      exp == "E4"] %>%
-ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = value), size = 0.5) +
-  scale_color_distiller(palette = "YlOrRd", direction = 1) +
-  # geom_contour(aes(z = value - fcst_value)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  labs(color = "Td", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  sondeos %>% 
-  .[variable == "td" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
-  .[, oma := value - fcst_value] %>% 
-  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = oma), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
-  labs(color = "OmA", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(0.3, 0.7))
-```
-
-```{r}
-sondeos %>% 
-  .[variable == "v" & !str_detect(site, "/") & 
-      exp == "E4"] %>% 
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = value), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  labs(color = "V (m/s)", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  sondeos %>% 
-  .[variable == "v" & !str_detect(site, "/") & !is.na(fcst_value)] %>% 
-  .[, oma := value - fcst_value] %>% 
-  # melt(measure.vars = c("value", "oma"), variable.name = "var") %>%  
-  ggplot(aes(launch_time, alt/1000)) +
-  geom_point(aes(color = oma), size = 0.5) +
-  scale_color_divergent() +
-  # geom_contour(aes(z = value - fcst_value)) +
-  coord_cartesian(ylim = c(0, 20), expand = c(0, 0)) +
-  scale_x_datetime(date_breaks = "12 hours", date_labels = "%b %d \n %H%M") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
-  labs(color = "OmA", y = NULL, x = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(0.3, 0.7))
-```
-
-### Forecast
-
-```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_*_E[4-7]new.csv")
-
-fss <- purrr::map(files, function(f) {
-
-  meta <- unglue(basename(f), "fss_6h_fcst_{ini_date}_{exp}new.csv")
+  meta <- unglue(basename(f), "fss_6h_fcst_ens_{ini_date}_{exp}_ens.csv")
 
@@
 ```
+
+### Soundings
+
+```{r message=FALSE, warning=FALSE}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
+files <- files[str_detect(files, "_01_")]
+
+sondeos <- purrr::map(files, function(f){
+
+  fread(f) %>% 
+    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
+    .[, launch_time := as_datetime(launch_time)] %>% 
+    .[, fcst_exp := basename(dirname(dirname(f)))]
+
+}) %>% 
+  rbindlist()
+
+IOP <- tribble(
+  ~iop, ~ini, ~end,
+  "IOP07", ymd_hms("20181121150000"), ymd_hms("20181121210000"),
+  "IOP08", ymd_hms("20181122140000"), ymd_hms("20181122200000")
+) %>% setDT()
+
+```
+
+```{r}
+
+sondeos %>% 
+  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
+                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
+  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
+  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
+  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
+           BIAS = mean(value - fcst_value, na.rm = TRUE)), 
+    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
+  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
+  ggplot(aes(lev/1000, value.est)) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  geom_line(aes(color = exp, linetype = estadistico)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+```
+
+### Observaciones de superficie
+
+```{r}
+# files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112206/interp*"))
+# 
+# obs <- purrr::map(files, function(f) {
+#   print(f)
+#   read_rds(f) %>% 
+#     .[, id := 1:.N, by = ens] %>% 
+#     .[, fcst_exp := "2018112206"] %>% 
+#     .[, exp := NULL] %>% 
+#     melt(measure.vars = c("E4", "E5", "E6", "E7"), 
+#          variable.name = "exp",
+#          value.name = "fcst") %>% 
+#     .[, ":="(obs = if_else(usage.flag == 1, obs, NA_real_),
+#              fcst = if_else(usage.flag == 1, fcst, NA_real_))] %>% 
+#     .[, c(list(obs = mean(obs, na.rm = TRUE),
+#                fcst = mean(fcst, na.rm = TRUE)),
+#           .SD[ens == 1]), by = .(id, exp), .SDcols = -c("obs", "fcst")] %>% 
+#     .[usage.flag == 1]
+#   
+# }) %>% 
+#   rbindlist()
+
+obs_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean.rds"))
+obs_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean.rds"))
+
+rbind(obs_00, obs_06) %>% 
+  .[usage.flag == 1] %>% 
+  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
+        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
+  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
+  .[estadistico == "rmse"] %>% 
+  ggplot(aes(date, value)) +
+  geom_line(aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+
+rbind(obs_00, obs_06) %>% 
+  .[usage.flag == 1] %>% 
+  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
+        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
+  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
+  .[estadistico == "bias"] %>% 
+  ggplot(aes(date, value)) +
+  geom_line(aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+```
+
+### Temperatura de brillo
+
+```{r eval=FALSE, include=FALSE}
+
+scale_color_topes <- function(colours = hex, 
+                              limits = c(-90, 50), 
+                              breaks = seq(-90, 50, 20),
+                              guide = guide_colorbar(barheight = 15),
+                              ...) {
+  # https://github.com/garciafido/cima-goes/tree/master/src/cima/goes/img
+  topes <- data.table::fread("../../temp_R/paleta_topes")
+  colours <- rgb(topes$V2, topes$V3, topes$V4, maxColorValue = 255)
+
+  scale_color_gradientn(colours = colours, 
+                        limits = limits, 
+                        breaks = breaks, 
+                        guide = guide, ...) 
+}
+
+purrr::walk(c(0:30), function(h) {
+
+  file_path <- paste0("~/datosmunin/EXP/E[4-7]/FCST/2018112200/CRTM/01/G16_20181122000000_", formatC(h, flag = "0", width = 3), "_00.nc")
+
+  files <- Sys.glob(file_path)
+
+  crtm <- purrr::map(files, function(f) {
+
+    meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/FCST/{fcst_exp}/CRTM/{mem}/G16_{fcst}_{hour}_{minutes}.nc")
+
+    ReadNetCDF(f, vars = c(lat = "XLAT", lon = "XLONG", "CH10")) %>% 
+      .[, CH10 := CH10 - 273.15] %>% 
+      .[, ":="(exp = meta[[1]][["exp"]],
+               mem = meta[[1]][["mem"]],
+               fcst_exp = meta[[1]][["fcst_exp"]],
+               date = ymd_hms(meta[[1]][["fcst"]]) + 
+                 hours(as.numeric(meta[[1]][["hour"]])) +
+                 minutes(as.numeric(meta[[1]][["minutes"]])))]
+
+  }) %>% rbindlist()
+
+  crtm %>% 
+    ggplot(aes(lon, lat)) +
+    geom_point(aes(color = CH10), size = 0.7) +
+    # scale_color_viridis_c() +
+    scale_color_topes(guide = guide_colorbar(barheight = 20)) +
+    geom_mapa() +
+    facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                 E6 = "SATWND", E7 = "RAD"))) +
+    labs(caption = crtm[,date[1]]) +
+    theme_minimal() +
+    theme(panel.ontop = TRUE,
+          panel.grid = element_line(linetype = 3))
+
+  ggsave(paste0("CH10_2018112200_", format(crtm[,date[1]], "%Y%m%d%H%M"), ".png"))
+
+})
+```

### 2021-02-19 analysis/scripts/biasvar_coef.R
Arreglo update del bias

@@ -29,8 +29,8 @@ pred <- diag$predictors[, as.vector(QCmask), this_channel]
 Bdiag <- rep(.1, npred)
 
 
-  # coef_prior <- coef_gfs[sensor == paste(this_sensor, plat, sep = "_") & channel == this_channel, 7:18]
-  # coef_prior <- matrix(as.vector(t(coef_prior)))
+  coef_prior <- coef_gfs[sensor == paste(this_sensor, plat, sep = "_") & channel == this_channel, 7:18]
+  coef_prior <- matrix(as.vector(t(coef_prior)))
   coef_prior <- rep(0, npred)
 
 biasvar <- c(0.000001, 0.00001, 0.0001, 0.001, 0.01, 0.1, 1, 10, 100, 1000)
@@ -42,7 +42,7 @@ for (i in seq_along(biasvar)) {
 
   coef_est <- estimate_coef(Xobs, Xmod, pred, coef_prior, Rdiag, Bdiag)
 
-  bias_est <- get_bias_correction(pred, coef_est)
+  bias_est <- get_bias_correction(pred, coef_est + coef_prior)
 
   if (i == 1) {
     dt <- data.table("xobs" = c(Xobs), "xmod" = c(Xmod), "bias" = bias_est, biasv = biasvar[i])

### 2021-02-19 analysis/scripts/main_coef.R
Arreglo update coeficientes

@@ -1,4 +1,5 @@
 library(reticulate)
+library(tidyverse)
 library(data.table)
 library(ggplot2)
 
@@ -30,8 +31,8 @@ out_coef <- purrr::map(sensor_list, function(f) {
     #Sensor
     nchanl <- diag[[1]]$nuchan
 
-    out_sensor <- purrr::map(nchanl, function(n) {
-      this_channel <- n
+    out_sensor <- purrr::map(seq_along(nchanl), function(n) {
+      channel_index <- n
 
       this_sensor <- str_trim(diag[[1]]$sensor)
       plat <- str_trim(diag[[1]]$sat)
@@ -42,12 +43,12 @@ out_coef <- purrr::map(sensor_list, function(f) {
       coef_prior <- rep(0, npred)
 
       out_channel <- purrr::map(diag, function(x) {
-        QCmask <- matrix(x$qc[, this_channel] == 0)
+        QCmask <- matrix(x$qc[, channel_index] == 0)
 
-        Rdiag <- 1/x$errinv[, this_channel][QCmask]
-        Xobs <- matrix(x$tb_obs[, this_channel][QCmask])
-        Xmod <- Xobs - matrix(x$tbcnob[, this_channel][QCmask])
-        pred <- x$predictors[, as.vector(QCmask), this_channel]
+        Rdiag <- 1/x$errinv[, channel_index][QCmask]
+        Xobs <- matrix(x$tb_obs[, channel_index][QCmask])
+        Xmod <- Xobs - matrix(x$tbcnob[, channel_index][QCmask])
+        pred <- x$predictors[, as.vector(QCmask), channel_index]
 
         list(Rdiag = Rdiag, Xobs = Xobs, Xmod = Xmod, pred = pred)
       } ) %>%
@@ -58,7 +59,7 @@ out_coef <- purrr::map(sensor_list, function(f) {
 
       if (length(out_channel$Xobs) > 1) {
         coef_est <- estimate_coef(out_channel$Xobs, out_channel$Xmod,
-                                  out_channel$pred, coef_prior, out_channel$Rdiag, Bdiag)
+                                  out_channel$pred, coef_prior, out_channel$Rdiag, Bdiag) + coef_prior
 
         bias_est <- get_bias_correction(out_channel$pred, coef_est)
 
@@ -69,7 +70,7 @@ out_coef <- purrr::map(sensor_list, function(f) {
       }
 
       coef_out <- data.table(sensor = paste(this_sensor, plat, sep = "_"),
-                             channel = this_channel,
[…]

### 2021-02-24 analysis/chapters/forecast_20181122.Rmd
Arreglo figura de observaciones de superficie

@@
 ```{r}
-# files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112206/interp*"))
+# files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112200/interp*"))
 # 
@@
 #     .[, id := 1:.N, by = ens] %>%
-#     .[, fcst_exp := "2018112206"] %>% 
+#     .[, fcst_exp := "2018112200"] %>%
 #     .[, exp := NULL] %>%
@@
   ggplot(aes(date, value)) +
+  geom_hline(yintercept = 0, color = "grey60") +
   geom_line(aes(color = exp, linetype = fcst_exp)) +

### 2021-03-03 analysis/scripts/sondeos_est.R
0rdeno

@@ -4,7 +4,7 @@ library(tidyverse)
 library(lubridate)
 library(data.table)
 
-miembros <- 3
+miembros <- 60
 
 IOP <- tribble(
   ~iop, ~ini, ~end,

### 2021-03-03 analysis/chapters/forecast_20181122.Rmd
Sondeos, todas las versiones

@@
 
+miembros <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_agregados.rds") %>% 
+  .[, mem := formatC(mem, width = 3, flag = "0")] %>% 
+  fisica[., on = "mem"] %>% 
+  separate(fisica, into = c("cumulus", "pbl"))
+
+media <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_media.rds") %>% 
+  .[, .(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
+        BIAS = mean(value - fcst_value, na.rm = TRUE)), 
+    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
+  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est")
+
 IOP <- tribble(
@@
 
+```{r rmse-sondeos-media}
+extremos <- miembros %>% 
+  .[, .(max_rmse = max(RMSE),
+        min_rmse = min(RMSE)), by = .(exp, fcst_exp, lev, variable)]
+
+media %>% 
+  .[estadistico == "RMSE"] %>% 
+  extremos[., on = .NATURAL] %>% 
+  ggplot(aes(lev/1000, value.est)) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  geom_ribbon(aes(ymin = min_rmse, 
+                  ymax = max_rmse, fill = exp), alpha = 0.2) +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E7 = "RAD")) +
+  geom_line(aes(color = exp, linetype = estadistico)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+
+media %>% 
+  .[estadistico == "RMSE"] %>% 
+  miembros[., on = .NATURAL] %>% 
+  ggplot(aes(lev/1000, value.est)) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  geom_line(alpha = 0.1, 
+            aes(y = RMSE, color = exp, group = interaction(exp, mem))) +
+  geom_line(aes(color = exp, linetype = estadistico)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+```
+
+```{r rmse-sondeos-param}
+pbl <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, lev, variable)]
+
+media %>% 
+  .[estadistico == "RMSE"] %>% 
+  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
+  .[, ":="(pbl = "Mean",
+           RMSE = value.est)] %>% 
+  .[, value.est := NULL] %>% 
+  rbind(., pbl) %>% 
+  ggplot(aes(lev/1000, RMSE)) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  geom_line(aes(y = RMSE, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
+  scale_linetype_manual(values = c(1, 2, 3, 4)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp ~ variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+
+cum <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, lev, variable)]
+
+media %>% 
+  .[estadistico == "RMSE"] %>% 
+  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
+  .[, ":="(cumulus = "Mean",
+           RMSE = value.est)] %>% 
+  .[, value.est := NULL] %>% 
+  rbind(., cum) %>% 
+  ggplot(aes(lev/1000, RMSE)) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  geom_line(aes(y = RMSE, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
+  scale_linetype_manual(values = c(4, 2, 3, 1)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp ~ variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+
+```
+
+```{r bias-sondeos-media}
+extremos <- miembros %>% 
+  .[, .(max_bias = max(BIAS),
+        min_bias = min(BIAS)), by = .(exp, fcst_exp, lev, variable)]
+
+media %>% 
+  .[estadistico == "BIAS"] %>% 
+  extremos[., on = .NATURAL] %>% 
+  ggplot(aes(lev/1000, value.est)) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  geom_ribbon(aes(ymin = min_bias, 
+                  ymax = max_bias, fill = exp), alpha = 0.2) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E7 = "RAD")) +
+  geom_line(aes(color = exp, linetype = estadistico)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+
+media %>% 
+  .[estadistico == "BIAS"] %>% 
+  miembros[., on = .NATURAL] %>% 
+  ggplot(aes(lev/1000, value.est)) +
+
+  geom_line(alpha = 0.1, 
+            aes(y = BIAS, color = exp, group = interaction(exp, mem))) +
+  geom_line(aes(color = exp, linetype = estadistico)) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal() geom_hline(yintercept = 0, color = "grey30") +
+```
+
+```{r bias-sondeos-param}
+pbl <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, lev, variable)]
+
+media %>% 
+  .[estadistico == "BIAS"] %>% 
+  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
+  .[, ":="(pbl = "Mean",
+           BIAS = value.est)] %>% 
+  .[, value.est := NULL] %>% 
+  rbind(., pbl) %>% 
+  ggplot(aes(lev/1000, BIAS)) +
+  geom_line(aes(y = BIAS, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(1, 2, 3, 4)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp ~ variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+
+cum <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, lev, variable)]
+
+media %>% 
+  .[estadistico == "BIAS"] %>% 
+  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
+  .[, ":="(cumulus = "Mean",
+           BIAS = value.est)] %>% 
+  .[, value.est := NULL] %>% 
+  rbind(., cum) %>% 
+  ggplot(aes(lev/1000, BIAS)) +
+  geom_line(aes(y = BIAS, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(4, 2, 3, 1)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  coord_flip() +
+  facet_grid(fcst_exp ~ variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+
+```
+
 ### Observaciones de superficie
@@
 obs_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean.rds"))
+
+obs_ens_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds"))
+obs_ens_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds"))
 
@@
   theme_minimal()
+```
+
+```{r}
+
+extremos <- rbind(obs_ens_00, obs_ens_06) %>% 
+  .[, .(max_rmse = max(RMSE),
+        min_rmse = min(RMSE)), by = .(var, exp, fcst_exp, date)]
+
+miembros <- rbind(obs_ens_00, obs_ens_06)
+
+rbind(obs_00, obs_06) %>% 
+  .[usage.flag == 1] %>% 
+  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
+        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
+  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
+  .[estadistico == "rmse"] %>% 
+  extremos[., on = .NATURAL] %>% 
+  ggplot(aes(date, value)) +
+  geom_ribbon(aes(ymin = min_rmse, 
+                  ymax = max_rmse, fill = exp), alpha = 0.2) +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E7 = "RAD")) +
+  geom_line(aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+
+rbind(obs_00, obs_06) %>% 
+  .[usage.flag == 1] %>% 
+  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
+        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
+  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
+  .[estadistico == "rmse"] %>% 
+  miembros[., on = .NATURAL] %>% 
+  ggplot(aes(date, value)) +
+  geom_line(alpha = 0.1, 
+            aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
+  geom_line(aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+
+```
+
+```{r}
+miembros %>% 
+  .[, mean_RMSE := mean(RMSE), by =  .(exp, fcst_exp, date, var)] %>% 
+  ggplot(aes(date, RMSE)) +
+  geom_line(aes(color = exp, linetype = fcst_exp, 
+                grup = interaction(exp, fcst_exp, mem)), alpha = 0.1) +
+  geom_line(aes(y = mean_RMSE, color = exp, linetype = fcst_exp)) +
+   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+```
+
+### Precipitación
+
+```{r}
+files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_mean_acum_30h.rds"))
+
+pp_acum <- purrr::map(files, function(f) {
+
+  meta <- unglue(basename(f), "{exp}_fcst_{fcst_init}_mean_acum_30h.rds")
+
+  read_rds(f) %>% 
+    .[, pp_acum := if_else(pp_acum < 0, 0, pp_acum)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_init = meta[[1]][["fcst_init"]])]
+
+}) %>% rbindlist()
+
+pp_acum %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
+                    proj = norargentina_lambert,
+                    breaks = seq(20, 300, 20)) +
+  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
+                                                barheight = 0.8), 
+                       super = ScaleDiscretised,
+                       limits = c(20, 300),
+                       na.value = "white") +
+  geom_point(data = square, aes(lon, lat), size = 0.1) +
+  geom_mapa() +
+  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  labs(fill = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  read_rds(paste0(derived_data, "ppacum/IMERG_30h.rds")) %>% 
+  .[, run := "IMERG"] %>%
+  .[,  c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
+                    proj = norargentina_lambert,
+                    breaks = seq(20, 300, 20)) +
+  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
+                                                barheight = 0.8), 
+                       super = ScaleDiscretised,
+                       limits = c(20, 300),
+                       na.value = "white") +
+  geom_point(data = square, aes(lon, lat), size = 0.1) +
+  geom_mapa() +
+  facet_wrap(~run) +
+  labs(fill = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(3, 1), guides = "collect") & theme(legend.position = 'bottom')
+
+```
+
+```{r}
+files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_prop_acum_30h.rds"))
+
+pp_prop <- purrr::map(files, function(f) {
+
+  meta <- unglue(basename(f), "{exp}_fcst_{fcst_init}_prop_acum_30h.rds")
+
+  read_rds(f) %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_init = meta[[1]][["fcst_init"]])]
+
+}) %>% rbindlist()
+
+pp_prop %>% 
+  .[umbral == 100] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = prop*100, fill = stat(level_d)), 
+                    proj = norargentina_lambert,
+                    breaks = seq(10, 100, 10)) +
+  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")),
+                       limits = c(10, 100),
+                       guide = guide_colorsteps(barwidth = 20,
+                                                barheight = 0.8),
+                       super = ScaleDiscretised,
+                       na.value = "white") +
+  geom_point(data = square, aes(lon, lat), size = 0.1) +
+  geom_mapa() +
+  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  labs(fill = "100 mm") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  read_rds(paste0(derived_data, "ppacum/IMERG_30h.rds")) %>% 
+  .[, run := "IMERG"] %>%
+  .[,  c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
+                    proj = norargentina_lambert,
+                    breaks = seq(20, 300, 20)) +
+  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
+                                                barheight = 0.8), 
+                       super = ScaleDiscretised,
+                       limits = c(20, 300),
+                       na.value = "white") +
+  geom_point(data = square, aes(lon, lat), size = 0.1) +
+  geom_mapa() +
+  facet_wrap(~run) +
+  labs(fill = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(3, 1)) & theme(legend.position = 'bottom')
+
 ```

### 2021-03-03 analysis/scripts/superficie_est.R
Calculo de RMSE y BIAS para obs de superficie

@@ -1,0 +1,74 @@
+
+fisica <- data.table(mem = as.character(formatC(1:60, flag = "0", width = 3)),
+                     fisica = rep(c("KF-YSU",
+                                    "BMJ-YSU",
+                                    "GF-YSU",
+                                    "KF-MYJ",
+                                    "BMJ-MYJ",
+                                    "GF-MYJ",
+                                    "KF-MYNN2",
+                                    "BMJ-MYNN2",
+                                    "GF-MYNN2"), length.out = 60)) %>% setDT()
+
+files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112200/interp*"))
+
+obs <- purrr::map(files, function(f) {
+  print(f)
+  read_rds(f) %>%
+    .[, id := 1:.N, by = ens] %>%
+    .[, fcst_exp := "2018112200"] %>%
+    .[, mem := formatC(ens, flag = "0", width = 3)] %>%
+    .[, exp := NULL] %>%
+    melt(measure.vars = c("E4", "E5", "E6", "E7"),
+         variable.name = "exp",
+         value.name = "fcst") %>%
+    .[, ":="(obs = if_else(usage.flag == 1, obs, NA_real_),
+             fcst = if_else(usage.flag == 1, fcst, NA_real_))] %>%
+    .[usage.flag == 1] %>%
+    .[, .(RMSE = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
+          BIAS = mean(obs - fcst, na.rm = TRUE)), by = .(var, exp, mem, fcst_exp, date)]
+  # .[, c(list(RMSE = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
+  #            BIAS = mean(obs - fcst, na.rm = TRUE)),
+  #       .SD[mem == 1]), by = .(id, exp, mem), .SDcols = -c("obs", "fcst")] %>%
+
+
+}) %>%
+  rbindlist()
+
+write_rds(obs, paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds"))
+
+
+
+# mean --------------------------------------------------------------------
+
+files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112206/interp*"))
+
+obs <- purrr::map(files, function(f) {
+  print(f)
+  a <- read_rds(f) %>%
[…]

### 2021-03-05 analysis/chapters/forecast_20181122.Rmd
Campos de variables

@@
   labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal() geom_hline(yintercept = 0, color = "grey30") +
+  theme_minimal() 
 ```
@@
 
-obs_ens_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds"))
-obs_ens_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds"))
+media <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean2.rds")),
+               read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean2.rds")))
 
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
+miembros <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds")),
+                  read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds")))
+
+media %>% 
+  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico") %>% 
+  .[estadistico == "RMSE"] %>% 
   ggplot(aes(date, value)) +
@@
 
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "bias"] %>% 
+media %>% 
+  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico") %>% 
+  .[estadistico == "BIAS"] %>% 
   ggplot(aes(date, value)) +
@@
 
-extremos <- rbind(obs_ens_00, obs_ens_06) %>% 
+extremos <- miembros %>% 
   .[, .(max_rmse = max(RMSE),
@@
 
-miembros <- rbind(obs_ens_00, obs_ens_06)
-
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
+media %>% 
   extremos[., on = .NATURAL] %>% 
-  ggplot(aes(date, value)) +
+  .[exp == "E7"] %>% 
+  ggplot(aes(date, RMSE)) +
   geom_ribbon(aes(ymin = min_rmse, 
-                  ymax = max_rmse, fill = exp), alpha = 0.2) +
+                  ymax = max_rmse, fill = exp, group = interaction(exp, fcst_exp)), alpha = 0.2) +
   scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
@@
 
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
-  miembros[., on = .NATURAL] %>% 
-  ggplot(aes(date, value)) +
-  geom_line(alpha = 0.1, 
+media %>% 
+  .[, mem := "000"] %>% 
+  rbind(., miembros) %>% 
+  .[exp == "E7"] %>% 
+  ggplot(aes(date, RMSE)) +
+  geom_line(data = ~.x[mem != "000"], alpha = 0.1, 
             aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
-  geom_line(aes(color = exp, linetype = fcst_exp)) +
+  geom_line(data = ~.x[mem == "000"], aes(color = exp, linetype = fcst_exp)) +
   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
@@
 ```{r}
-miembros %>% 
+copy(miembros) %>% 
   .[, mean_RMSE := mean(RMSE), by =  .(exp, fcst_exp, date, var)] %>% 
@@
   facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+```
+```{r}
+
+extremos <- miembros %>% 
+  .[, .(max_bias = max(BIAS),
+        min_bias = min(BIAS)), by = .(var, exp, fcst_exp, date)]
+
+media %>% 
+  extremos[., on = .NATURAL] %>% 
+  .[exp == "E7"] %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_ribbon(aes(ymin = min_bias, 
+                  ymax = max_bias, fill = exp, group = interaction(exp, fcst_exp)), alpha = 0.2) +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E7 = "RAD")) +
+  geom_line(aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+
+media %>% 
+  .[, mem := "000"] %>% 
+  rbind(., miembros) %>% 
+  .[exp == "E7"] %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_line(data = ~.x[mem != "000"], alpha = 0.1, 
+            aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
+  geom_line(data = ~.x[mem == "000"], aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+
+```
+
+```{r}
+cum <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, var, date)]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
+  .[, ":="(cumulus = "Mean")] %>% 
+  rbind(., cum) %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_line(aes(color = exp, linetype = cumulus)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(4, 2, 3, 1)) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+
+cum <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, var, date)]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
+  .[, ":="(cumulus = "Mean")] %>% 
+  rbind(., cum) %>% 
+  ggplot(aes(date, RMSE)) +
+  geom_line(aes(color = exp, linetype = cumulus)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(4, 2, 3, 1)) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+```
+```{r}
+pbl <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, var, date)]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
+  .[, ":="(pbl = "Mean")] %>% 
+  rbind(., pbl) %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_line(aes(color = exp, linetype = pbl)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(1, 2, 3, 4)) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+
+pbl <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, var, date)]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
+  .[, ":="(pbl = "Mean")] %>% 
+  rbind(., pbl) %>% 
+  ggplot(aes(date, RMSE)) +
+  geom_line(aes(color = exp, linetype = pbl)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(1, 2, 3, 4)) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
   theme_minimal()
@@
 ```
+
+### Campos de variables
+
+```{r}
+df <- data.frame(
+  lead_time = rep(c(6, 12, 18), 4),
+  exp = rep(c("E4", "E5", "E6", "E7"), each = 3)
+)
+
+files <- purrr::map2_chr(df$lead_time, df$exp, function(l, e) {
+  paste0("/home/paola.corrales/datosmunin/EXP/", e, "/FCST/2018112200/NPP/NPP_2018-11-22_00_FC", 
+         formatC(l, width = 2, flag = "0"), ".nc")
+})
+
+coord <- ReadNetCDF(files[1], vars = c("XLAT", "XLONG"))
+
+var_00 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("Q", "T", "V"), subset = list(lev = c(800, 1000))) %>% 
+    .[, .(Q = mean(Q),
+          T = mean(T),
+          V = mean(V)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+
+mcape_00 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("MCAPE", "MDBZ")) %>% 
+    .[, .(MCAPE = mean(MCAPE, na.rm = TRUE),
+          MDBZ = mean(MDBZ, na.rm = TRUE)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+```
+
+```{r}
+df <- data.frame(
+  lead_time = rep(c(0, 6, 12), 4),
+  exp = rep(c("E4", "E5", "E6", "E7"), each = 3)
+)
+
+files <- purrr::map2_chr(df$lead_time, df$exp, function(l, e) {
+  paste0("/home/paola.corrales/datosmunin/EXP/", e, "/FCST/2018112206/NPP/NPP_2018-11-22_06_FC", 
+         formatC(l, width = 2, flag = "0"), ".nc")
+})
+
+coord <- ReadNetCDF(files[1], vars = c("XLAT", "XLONG"))
+
+var_06 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("Q", "T", "V"), subset = list(lev = c(800, 1000))) %>% 
+    .[, .(Q = mean(Q),
+          T = mean(T),
+          V = mean(V)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+
+mcape_06 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("MCAPE", "MDBZ")) %>% 
+    .[, .(MCAPE = mean(MCAPE, na.rm = TRUE),
+          MDBZ = mean(MDBZ, na.rm = TRUE)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+```
+
+```{r}
+var_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = Q, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 20, 2)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       limits = c(1, 20),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+var_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = Q, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 20, 2)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       limits = c(1, 20),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+```{r}
+var_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = V, fill = stat(level_d)), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(name = NULL,
+                                   guide = guide_colorsteps(#inside = FALSE,
+                                     barwidth = 25,
+                                     barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+var_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = V, fill = stat(level_d)), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(name = NULL,
+                                   guide = guide_colorsteps(#inside = FALSE,
+                                     barwidth = 25,
+                                     barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
+
+```{r}
+var_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = T - 273.15, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 40, 5)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "OrRd", direction = 1,
+                       limits = c(1, 40),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+var_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = T - 273.15, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 40, 5)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "OrRd", direction = 1,
+                       limits = c(1, 40),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+```{r}
+mcape_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MCAPE, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 4500, 500)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+mcape_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MCAPE, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 4500, 500)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+```{r}
+values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
+22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
+42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
+61.5, 63, 64.5, 66, 69, 76.5)
+
+colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
+"#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
+"#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
+"#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
+"#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
+"#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
+"#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
+"#95E3C5", "#85E0BD")
+
+mcape_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MDBZ, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    drop = FALSE, guide = guide_colorsteps(barwidth = 0.8, 
+                                                                         barheight = 30)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+mcape_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MDBZ, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    drop = FALSE, guide = guide_colorsteps(barwidth = 0.8, 
+                                                                         barheight = 30)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E7 = "RAD"))) +
+  theme_minimal() +
+  theme(panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```

### 2021-03-09 analysis/chapters/experimento_20181122.Rmd
Actualizo figura dbz con los datos del SMN

@@
 
+proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"
 ```
@@
 
-```{r dbz-mean}
+```{r dbz-read}
+
+files <- Sys.glob("/home/jruiz/share/DATA/DATOS_RADAR/MOSAICO/*.nc")[seq(1, 12, 3)]
+# Solo un par de horarios 10, 13, 16 y 19 UTC
+
+dbz_obs <- purrr::map(files, function(f) {
+
+  ReadNetCDF(f, vars = "DBZH_cor") %>%
+    .[!is.na(DBZH_cor)] %>% 
+    .[, .(DBZH_cor = max(DBZH_cor, na.rm = TRUE)), by = .(x, y)] %>% 
+    .[, max_dbz := 10 * log10(DBZH_cor)] %>% 
+    .[, ":="(date = ymd_hms(str_remove(basename(f), ".nc")),
+             DBZH_cor = NULL)] %>% 
+    .[, c("lon", "lat") := proj4::project(list(x, y), proj_radar, inverse = TRUE)]
+
+}) %>% rbindlist() %>% 
+  .[, run := "OBSERVATIONS"]
 
@@
 
-dbz <- purrr::map(files, function(f){
+dbz_ana <- purrr::map(files, function(f){
 
@@
 }) %>% 
-  rbindlist()
+  rbindlist()  %>% 
+  .[, run := "RAD"]
 
@@
              "#95E3C5", "#85E0BD")
-
-dbz %>% 
+```
+
+```{r dbz-mean}
+dbz_obs %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+                    proj = proj_radar,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 30, 
+                                             barheight = 0.8)) +
+  geom_mapa() +
+  facet_grid(run ~ date) +
+  labs(x = NULL, y = NULL, fill = NULL, color = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        # legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  dbz_ana %>% 
   .[, date := factor(date)] %>% 
@@
                     breaks = values) +
-  scale_fill_manual(values = colours, labels = function(x) JumpBy(x, 3, fill = ""),
-                    drop = FALSE, guide = guide_colorsteps(barwidth = 30, 
-                                                                         barheight = 0.8)) +
-
-  # scale_color_dbz(name = NULL, aesthetics = "fill", 
-  #                 guide = guide_colorstrip(barwidth = 20, 
-  #                                          barheight = 0.8,
-  #                                          inside = TRUE)) +
-  # ggnewscale::new_scale_fill() +
-  # geom_contour_fill(aes(z = max_dbz), fill = "white", 
-  #                   proj = norargentina_lambert,
-  #                   breaks = c(-Inf, 10)) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 30, 
+                                             barheight = 0.8)) +
+
   geom_mapa() +
-  facet_wrap(~ date, ncol = 4) +
+  facet_grid(run ~ date) +
   labs(x = NULL, y = NULL, fill = NULL, color = NULL) +
@@
   theme(legend.position = "bottom",
-        legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  plot_layout(ncol = 1, guides = "collect") & 
+  theme(legend.position = 'bottom')
 
@@
   .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
-  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
+  .[, lev := cut_round(alt, c(seq(0, 3000, 500), seq(4000, 21000, 1000)))] %>% 
   .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 

### 2021-03-09 analysis/scripts/interp_obs_parallel.R
q/1000

@@ -19,7 +19,7 @@ fcst_path <- "/home/paola.corrales/datosmunin/EXP"
 
 ini_date <- ymd_hms("20181122000000")
 
-lead_time <- 2
+lead_time <- 36
 
 # out <- foreach(l = 0:lead_time),
 # .packages = c("data.table", "metR", "lubridate", "interp", "dplyr"),
@@ -60,7 +60,8 @@ purrr::map_dbl(c(0:lead_time), function(l) {
     fcst_exp <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{fcst_exp}/FCST/{info}")
     fcst <- ReadNetCDF(f, vars = c("XLONG", "XLAT", p = "PSFC",
                                    t = "T2",q = "Q2", u = "U10", v = "V10")) %>%
-      .[, fcst_exp := fcst_exp[[1]][["fcst_exp"]]]
+      .[, fcst_exp := fcst_exp[[1]][["fcst_exp"]]] %>%
+      .[, q := q/1000]
 
   }) %>% rbindlist()
 
@@ -105,8 +106,7 @@ purrr::map_dbl(c(0:lead_time), function(l) {
     write_rds(temp, path_out)
 
   l
-}) %>% rbindlist()
-
+})
 
 
 

### 2021-03-09 analysis/scripts/sondeos_est.R
Saco Sao Borja por sondeo malo y lejos del resto

@@ -39,6 +39,7 @@ for (m in seq_len(miembros)) {
     .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
                      launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>%
     .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>%
+    .[site != "Sao Borja, Brazil"] %>%
     .[, lev := cut_round(alt, c(seq(0, 3000, 500), seq(4000, 21000, 1000)))]
 
 

### 2021-03-10 analysis/chapters/experimento_20181122.Rmd
Saco Sao Borja de los sondeos.

@@
   .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
+  .[site != "Sao Borja, Brazil"] %>% 
   .[, lev := cut_round(alt, c(seq(0, 3000, 500), seq(4000, 21000, 1000)))] %>% 

### 2021-03-11 analysis/chapters/forecast_20181122.Rmd
Graficos de resumen de RMSE y BIAS

@@
 ```{r message=FALSE, warning=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
-files <- files[str_detect(files, "_01_")]
-
-sondeos <- purrr::map(files, function(f){
-
-  fread(f) %>% 
-    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
-    .[, launch_time := as_datetime(launch_time)] %>% 
-    .[, fcst_exp := basename(dirname(dirname(f)))]
-
-}) %>% 
-  rbindlist()
+# files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
+# files <- files[str_detect(files, "_01_")]
+# 
+# sondeos <- purrr::map(files, function(f){
+#   
+#   fread(f) %>% 
+#     .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
+#     .[, launch_time := as_datetime(launch_time)] %>% 
+#     .[, fcst_exp := basename(dirname(dirname(f)))]
+#   
+# }) %>% 
+#   rbindlist()
 
@@
 
-sondeos %>% 
-  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
-                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
-  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
-  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
-  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
-           BIAS = mean(value - fcst_value, na.rm = TRUE)), 
-    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
-  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
+media %>% 
   ggplot(aes(lev/1000, value.est)) +
@@
 
-```{r rmse-sondeos-media}
+```{r rmse-sondeos-ensamble}
 extremos <- miembros %>% 
@@
 ```{r rmse-sondeos-param}
-pbl <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, lev, variable)]
+pbl <- miembros[, .(RMSE = mean(RMSE),
+                    tipo = "pbl"), 
+                by = .(exp, fcst_exp, pbl, lev, variable)] %>% 
+  setnames("pbl", "param")
 
@@
   .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
-  .[, ":="(pbl = "Mean",
-           RMSE = value.est)] %>% 
+  .[, ":="(param = "Mean",
+           RMSE = value.est, 
+           tipo = "Mean")] %>% 
   .[, value.est := NULL] %>% 
@@
   # geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(y = RMSE, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
+  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
   scale_linetype_manual(values = c(1, 2, 3, 4)) +
@@
 
-cum <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, lev, variable)]
+cum <- miembros[, .(RMSE = mean(RMSE),
+                    tipo = "cumulus"), 
+                by = .(exp, fcst_exp, cumulus, lev, variable)] %>% 
+  setnames("cumulus", "param")
 
@@
   .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
-  .[, ":="(cumulus = "Mean",
-           RMSE = value.est)] %>% 
+  .[, ":="(param = "Mean",
+           RMSE = value.est,
+           tipo = "Mean")] %>% 
   .[, value.est := NULL] %>% 
@@
   # geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(y = RMSE, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
+  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
   scale_linetype_manual(values = c(4, 2, 3, 1)) +
@@
 
-```
-
-```{r bias-sondeos-media}
+rbind(pbl, cum) %>% 
+  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
+    by = .(variable, exp, param, fcst_exp)] %>% 
+  ggplot(aes(RMSE_mean)) +
+  geom_histogram(aes(fill = exp), 
+                 position = "identity", 
+                 alpha = 0.5) +
+   scale_fill_manual(values = colores_exp, 
+                     labels = c(E4 = "CONV", E5 = "AUT",
+                                E6 = "SATWND", E7 = "RAD")) +
+  facet_grid(fcst_exp ~ variable, scales = "free_x")
+
+rbind(pbl, cum) %>%
+  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
+    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(RMSE_mean, exp)) +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  facet_grid(fcst_exp ~ variable, scales = "free_x") +
+  labs(y = NULL, x = "mean RMSE", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
+
+```
+
+```{r bias-sondeos-ensamble}
 extremos <- miembros %>% 
@@
 ```{r bias-sondeos-param}
-pbl <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, lev, variable)]
+pbl <- miembros[, .(BIAS = mean(BIAS),
+                    tipo = "pbl"), 
+                by = .(exp, fcst_exp, pbl, lev, variable)] %>% 
+  setnames("pbl", "param")
 
@@
   .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
-  .[, ":="(pbl = "Mean",
-           BIAS = value.est)] %>% 
+  .[, ":="(param = "Mean",
+           BIAS = value.est,
+           tipo = "Mean")] %>% 
   .[, value.est := NULL] %>% 
@@
   ggplot(aes(lev/1000, BIAS)) +
-  geom_line(aes(y = BIAS, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
+  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
   geom_hline(yintercept = 0, color = "grey30") +
@@
 
-cum <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, lev, variable)]
+cum <- miembros[, .(BIAS = mean(BIAS),
+                    tipo = "cumulus"), 
+                by = .(exp, fcst_exp, cumulus, lev, variable)] %>% 
+  setnames("cumulus", "param")
 
@@
   .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
-  .[, ":="(cumulus = "Mean",
-           BIAS = value.est)] %>% 
+  .[, ":="(param = "Mean",
+           BIAS = value.est,
+           tipo = "Mean")] %>% 
   .[, value.est := NULL] %>% 
@@
   ggplot(aes(lev/1000, BIAS)) +
-  geom_line(aes(y = BIAS, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
+  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
   geom_hline(yintercept = 0, color = "grey30") +
@@
   theme_minimal()
+
+rbind(pbl, cum) %>%
+  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
+    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(BIAS_mean, exp)) +
+  geom_vline(xintercept = 0, color = "grey70") +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  facet_grid(fcst_exp ~ variable, scales = "free_x") +
+  labs(y = NULL, x = "mean BIAS", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
 
@@
 media <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean2.rds")),
-               read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean2.rds")))
+               read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean2.rds"))) %>% 
+  .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
+           BIAS = fifelse(var == "q", BIAS*1000, BIAS))]
 
 miembros <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds")),
-                  read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds")))
+                  read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds"))) %>% 
+    .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
+           BIAS = fifelse(var == "q", BIAS*1000, BIAS))]
 
@@
   separate(fisica, into = c("cumulus", "pbl")) %>% 
-  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, var, date)]
+  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
+   setnames("cumulus", "param") %>% 
+  .[, tipo := "cumulus"]
 
@@
   .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
-  .[, ":="(cumulus = "Mean")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
   rbind(., cum) %>% 
   ggplot(aes(date, BIAS)) +
-  geom_line(aes(color = exp, linetype = cumulus)) +
+  geom_line(aes(color = exp, linetype = param)) +
   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
@@
 
+pbl <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
+   setnames("pbl", "param") %>% 
+  .[, tipo := "pbl"]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
+  rbind(., pbl) %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_line(aes(color = exp, linetype = param)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(1, 2, 3, 4)) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+
+rbind(pbl, cum) %>%
+  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
+    by = .(var, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(BIAS_mean, exp)) +
+  geom_vline(xintercept = 0, color = "grey70") +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  facet_grid(fcst_exp ~ var, scales = "free_x") +
+  labs(y = NULL, x = "mean BIAS", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
+```
+
+```{r}
 cum <- fisica[miembros, on = .NATURAL] %>% 
   separate(fisica, into = c("cumulus", "pbl")) %>% 
-  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, var, date)]
+  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
+   setnames("cumulus", "param") %>% 
+  .[, tipo := "cumulus"]
 
@@
   .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
-  .[, ":="(cumulus = "Mean")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
   rbind(., cum) %>% 
   ggplot(aes(date, RMSE)) +
-  geom_line(aes(color = exp, linetype = cumulus)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  # geom_hline(yintercept = 0, color = "grey30") +
+  geom_line(aes(color = exp, linetype = param)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD")) +
+  geom_hline(yintercept = 0, color = "grey30") +
   scale_linetype_manual(values = c(4, 2, 3, 1)) +
@@
   theme_minimal()
-```
-```{r}
+
 pbl <- fisica[miembros, on = .NATURAL] %>% 
   separate(fisica, into = c("cumulus", "pbl")) %>% 
-  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, var, date)]
-
-media %>% 
-  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
-  .[, ":="(pbl = "Mean")] %>% 
-  rbind(., pbl) %>% 
-  ggplot(aes(date, BIAS)) +
-  geom_line(aes(color = exp, linetype = pbl)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  scale_linetype_manual(values = c(1, 2, 3, 4)) +
-  scale_date() +
-  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
-  facet_grid(var~fcst_exp, scales = "free_y") +
-  theme_minimal()
-
-pbl <- fisica[miembros, on = .NATURAL] %>% 
-  separate(fisica, into = c("cumulus", "pbl")) %>% 
-  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, var, date)]
+  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
+   setnames("pbl", "param") %>% 
+  .[, tipo := "pbl"]
 
@@
   .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
-  .[, ":="(pbl = "Mean")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
   rbind(., pbl) %>% 
   ggplot(aes(date, RMSE)) +
-  geom_line(aes(color = exp, linetype = pbl)) +
+  geom_line(aes(color = exp, linetype = param)) +
   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
@@
   theme_minimal()
+
+rbind(pbl, cum) %>%
+  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
+    by = .(var, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(RMSE_mean, exp)) +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E7 = "RAD")) +
+  facet_grid(fcst_exp ~ var, scales = "free_x") +
+  labs(y = NULL, x = "mean RMSE", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
 ```

### 2021-03-14 analysis/chapters/paper_ana_20181122.Rmd
Init paper | Dominio

--- 2021-03-03 analysis/chapters/forecast_20181122.Rmd
@@
 ---
-title: "FCST"
-output: 
-  bookdown::html_document2: default
-  bookdown::pdf_document2: default
-bibliography: mesoda_ref.bib
+title: "Untitled"
+output: html_document
 ---
@@
 
-### FSS
+## Introduction
 
-Acumulados a 6 hs
+## Data and Methods
 
-```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[4-7]_ens.csv")
+```{r dominio, fig.cap="The horizontal domain used on the simulations (black box) and the inner domain used for the validation / evaluation (grey box).
+The topography in meters is also shown (filled contours).", fig.height=4}
 
-fss <- purrr::map(files, function(f) {
-
-  meta <- unglue(basename(f), "fss_6h_fcst_ens_{ini_date}_{exp}_ens.csv")
-
-  fread(f) %>% 
-    .[, date := as_datetime(date)] %>% 
-    .[, ini_date := ymd_h(meta[[1]][["ini_date"]])]
-}) %>% 
-  rbindlist()
-
-fss %>% 
-  ggplot(aes(date, fss)) +
-  geom_hline(yintercept = 1, color = "darkgray") +
-  geom_line(aes(color = exp, linetype = factor(ini_date))) +
-  # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  scale_date(20181122000000, 20181123000000, "12 hours") +
-  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
-                                            "10" = "10 mm", "25" = "25 mm"),
-                                      w = c("1" = "1 km", "5" = "50 km", 
-                                            "11" = "100 km", 
-                                            "51" = "500 km", "101" = "1000 km"))) +
-  labs(color = NULL, x = NULL, linetype = NULL) +
-  theme_minimal()
-
-```
-
-### Soundings
-
-```{r message=FALSE, warning=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
-files <- files[str_detect(files, "_01_")]
-
-sondeos <- purrr::map(files, function(f){
-
-  fread(f) %>% 
-    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
-    .[, launch_time := as_datetime(launch_time)] %>% 
-    .[, fcst_exp := basename(dirname(dirname(f)))]
-
-}) %>% 
-  rbindlist()
-
-miembros <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_agregados.rds") %>% 
-  .[, mem := formatC(mem, width = 3, flag = "0")] %>% 
-  fisica[., on = "mem"] %>% 
-  separate(fisica, into = c("cumulus", "pbl"))
-
-media <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_media.rds") %>% 
-  .[, .(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
-        BIAS = mean(value - fcst_value, na.rm = TRUE)), 
-    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
-  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est")
-
-IOP <- tribble(
-  ~iop, ~ini, ~end,
-  "IOP07", ymd_hms("20181121150000"), ymd_hms("20181121210000"),
-  "IOP08", ymd_hms("20181122140000"), ymd_hms("20181122200000")
-) %>% setDT()
-
-```
-
-```{r}
-
-sondeos %>% 
-  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
-                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
-  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
-  .[, lev := cut_round(alt, seq(0, 20000, 1000))] %>% 
-  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
-           BIAS = mean(value - fcst_value, na.rm = TRUE)), 
-    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
-  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
-  ggplot(aes(lev/1000, value.est)) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-```
-
-```{r rmse-sondeos-media}
-extremos <- miembros %>% 
-  .[, .(max_rmse = max(RMSE),
-        min_rmse = min(RMSE)), by = .(exp, fcst_exp, lev, variable)]
-
-media %>% 
-  .[estadistico == "RMSE"] %>% 
-  extremos[., on = .NATURAL] %>% 
-  ggplot(aes(lev/1000, value.est)) +
-  # geom_hline(yintercept = 0, color = "grey30") +
-  geom_ribbon(aes(ymin = min_rmse, 
-                  ymax = max_rmse, fill = exp), alpha = 0.2) +
-  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                                   E6 = "SATWND", E7 = "RAD")) +
-  geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-
-media %>% 
-  .[estadistico == "RMSE"] %>% 
-  miembros[., on = .NATURAL] %>% 
-  ggplot(aes(lev/1000, value.est)) +
-  # geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(alpha = 0.1, 
-            aes(y = RMSE, color = exp, group = interaction(exp, mem))) +
-  geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-```
-
-```{r rmse-sondeos-param}
-pbl <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, lev, variable)]
-
-media %>% 
-  .[estadistico == "RMSE"] %>% 
-  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
-  .[, ":="(pbl = "Mean",
-           RMSE = value.est)] %>% 
-  .[, value.est := NULL] %>% 
-  rbind(., pbl) %>% 
-  ggplot(aes(lev/1000, RMSE)) +
-  # geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(y = RMSE, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
-  scale_linetype_manual(values = c(1, 2, 3, 4)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp ~ variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-
-cum <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, lev, variable)]
-
-media %>% 
-  .[estadistico == "RMSE"] %>% 
-  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
-  .[, ":="(cumulus = "Mean",
-           RMSE = value.est)] %>% 
-  .[, value.est := NULL] %>% 
-  rbind(., cum) %>% 
-  ggplot(aes(lev/1000, RMSE)) +
-  # geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(y = RMSE, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
-  scale_linetype_manual(values = c(4, 2, 3, 1)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp ~ variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-
-```
-
-```{r bias-sondeos-media}
-extremos <- miembros %>% 
-  .[, .(max_bias = max(BIAS),
-        min_bias = min(BIAS)), by = .(exp, fcst_exp, lev, variable)]
-
-media %>% 
-  .[estadistico == "BIAS"] %>% 
-  extremos[., on = .NATURAL] %>% 
-  ggplot(aes(lev/1000, value.est)) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  geom_ribbon(aes(ymin = min_bias, 
-                  ymax = max_bias, fill = exp), alpha = 0.2) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                                   E6 = "SATWND", E7 = "RAD")) +
-  geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-
-media %>% 
-  .[estadistico == "BIAS"] %>% 
-  miembros[., on = .NATURAL] %>% 
-  ggplot(aes(lev/1000, value.est)) +
-
-  geom_line(alpha = 0.1, 
-            aes(y = BIAS, color = exp, group = interaction(exp, mem))) +
-  geom_line(aes(color = exp, linetype = estadistico)) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal() geom_hline(yintercept = 0, color = "grey30") +
-```
-
-```{r bias-sondeos-param}
-pbl <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, lev, variable)]
-
-media %>% 
-  .[estadistico == "BIAS"] %>% 
-  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
-  .[, ":="(pbl = "Mean",
-           BIAS = value.est)] %>% 
-  .[, value.est := NULL] %>% 
-  rbind(., pbl) %>% 
-  ggplot(aes(lev/1000, BIAS)) +
-  geom_line(aes(y = BIAS, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  scale_linetype_manual(values = c(1, 2, 3, 4)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp ~ variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-
-cum <- miembros[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, lev, variable)]
-
-media %>% 
-  .[estadistico == "BIAS"] %>% 
-  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
-  .[, ":="(cumulus = "Mean",
-           BIAS = value.est)] %>% 
-  .[, value.est := NULL] %>% 
-  rbind(., cum) %>% 
-  ggplot(aes(lev/1000, BIAS)) +
-  geom_line(aes(y = BIAS, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
-  geom_hline(yintercept = 0, color = "grey30") +
-  scale_linetype_manual(values = c(4, 2, 3, 1)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
-  facet_grid(fcst_exp ~ variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
-
-```
-
-### Observaciones de superficie
-
-```{r}
-# files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112200/interp*"))
-# 
-# obs <- purrr::map(files, function(f) {
-#   print(f)
-#   read_rds(f) %>%
-#     .[, id := 1:.N, by = ens] %>%
-#     .[, fcst_exp := "2018112200"] %>%
-#     .[, exp := NULL] %>%
-#     melt(measure.vars = c("E4", "E5", "E6", "E7"),
-#          variable.name = "exp",
-#          value.name = "fcst") %>%
-#     .[, ":="(obs = if_else(usage.flag == 1, obs, NA_real_),
-#              fcst = if_else(usage.flag == 1, fcst, NA_real_))] %>%
-#     .[, c(list(obs = mean(obs, na.rm = TRUE),
-#                fcst = mean(fcst, na.rm = TRUE)),
-#           .SD[ens == 1]), by = .(id, exp), .SDcols = -c("obs", "fcst")] %>%
-#     .[usage.flag == 1]
-# 
-# }) %>%
-#   rbindlist()
-
-obs_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean.rds"))
-obs_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean.rds"))
-
-obs_ens_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds"))
-obs_ens_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds"))
-
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
-  ggplot(aes(date, value)) +
-  geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  scale_date() +
-  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
-  facet_wrap(~ var, scales = "free_y") +
-  theme_minimal()
-
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "bias"] %>% 
-  ggplot(aes(date, value)) +
-  geom_hline(yintercept = 0, color = "grey60") +
-  geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  scale_date() +
-  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
-  facet_wrap(~ var, scales = "free_y") +
-  theme_minimal()
-```
-
-```{r}
-
-extremos <- rbind(obs_ens_00, obs_ens_06) %>% 
-  .[, .(max_rmse = max(RMSE),
-        min_rmse = min(RMSE)), by = .(var, exp, fcst_exp, date)]
-
-miembros <- rbind(obs_ens_00, obs_ens_06)
-
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
-  extremos[., on = .NATURAL] %>% 
-  ggplot(aes(date, value)) +
-  geom_ribbon(aes(ymin = min_rmse, 
-                  ymax = max_rmse, fill = exp), alpha = 0.2) +
-  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                                   E6 = "SATWND", E7 = "RAD")) +
-  geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  scale_date() +
-  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
-  facet_wrap(~ var, scales = "free_y") +
-  theme_minimal()
-
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
-  miembros[., on = .NATURAL] %>% 
-  ggplot(aes(date, value)) +
-  geom_line(alpha = 0.1, 
-            aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
-  geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  scale_date() +
-  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
-  facet_wrap(~ var, scales = "free_y") +
-  theme_minimal()
-
-```
-
-```{r}
-miembros %>% 
-  .[, mean_RMSE := mean(RMSE), by =  .(exp, fcst_exp, date, var)] %>% 
-  ggplot(aes(date, RMSE)) +
-  geom_line(aes(color = exp, linetype = fcst_exp, 
-                grup = interaction(exp, fcst_exp, mem)), alpha = 0.1) +
-  geom_line(aes(y = mean_RMSE, color = exp, linetype = fcst_exp)) +
-   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  scale_date() +
-  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
-  facet_wrap(~ var, scales = "free_y") +
-  theme_minimal()
-```
-
-### Precipitación
-
-```{r}
-files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_mean_acum_30h.rds"))
-
-pp_acum <- purrr::map(files, function(f) {
-
-  meta <- unglue(basename(f), "{exp}_fcst_{fcst_init}_mean_acum_30h.rds")
-
-  read_rds(f) %>% 
-    .[, pp_acum := if_else(pp_acum < 0, 0, pp_acum)] %>% 
-    .[, ":="(exp = meta[[1]][["exp"]],
-             fcst_init = meta[[1]][["fcst_init"]])]
-
-}) %>% rbindlist()
-
-pp_acum %>%
+fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>%
 
   ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
-                    proj = norargentina_lambert,
-                    breaks = seq(20, 300, 20)) +
-  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
-                                                barheight = 0.8), 
-                       super = ScaleDiscretised,
-                       limits = c(20, 300),
-                       na.value = "white") +
-  geom_point(data = square, aes(lon, lat), size = 0.1) +
+  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
+                    breaks = seq(0, 6000, 500)) +
+  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
+                      name = NULL,
+                       breaks = seq(0, 6000, 1000),
+                       guide = guide_colorstrip(barwidth = 20,
+                                                barheight = 0.8)) +
   geom_mapa() +
-  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                          E6 = "SATWND", E7 = "RAD"))) +
-  labs(fill = NULL) +
+  geom_point(data = square, aes(lon, lat), size = 0.2) +
+  geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
   theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  read_rds(paste0(derived_data, "ppacum/IMERG_30h.rds")) %>% 
-  .[, run := "IMERG"] %>%
-  .[,  c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
-                    proj = norargentina_lambert,
-                    breaks = seq(20, 300, 20)) +
-  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
-                                                barheight = 0.8), 
-                       super = ScaleDiscretised,
-                       limits = c(20, 300),
-                       na.value = "white") +
-  geom_point(data = square, aes(lon, lat), size = 0.1) +
-  geom_mapa() +
-  facet_wrap(~run) +
-  labs(fill = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(3, 1), guides = "collect") & theme(legend.position = 'bottom')
-
+  theme(legend.position = "bottom", legend.box.spacing = margin(0, 0, 0, 0)) 
 ```
-
-```{r}
-files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_prop_acum_30h.rds"))
-
-pp_prop <- purrr::map(files, function(f) {
-
-  meta <- unglue(basename(f), "{exp}_fcst_{fcst_init}_prop_acum_30h.rds")
-
-  read_rds(f) %>% 
-    .[, ":="(exp = meta[[1]][["exp"]],
-             fcst_init = meta[[1]][["fcst_init"]])]
-
-}) %>% rbindlist()
-
-pp_prop %>% 
-  .[umbral == 100] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = prop*100, fill = stat(level_d)), 
-                    proj = norargentina_lambert,
-                    breaks = seq(10, 100, 10)) +
-  scale_fill_gradientn(colours = rev(Redmonder::redmonder.pal(10, name = "qMSOStd")),
-                       limits = c(10, 100),
-                       guide = guide_colorsteps(barwidth = 20,
-                                                barheight = 0.8),
-                       super = ScaleDiscretised,
-                       na.value = "white") +
-  geom_point(data = square, aes(lon, lat), size = 0.1) +
-  geom_mapa() +
-  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                          E6 = "SATWND", E7 = "RAD"))) +
-  labs(fill = "100 mm") +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  read_rds(paste0(derived_data, "ppacum/IMERG_30h.rds")) %>% 
-  .[, run := "IMERG"] %>%
-  .[,  c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
-                    proj = norargentina_lambert,
-                    breaks = seq(20, 300, 20)) +
-  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
-                                                barheight = 0.8), 
-                       super = ScaleDiscretised,
-                       limits = c(20, 300),
-                       na.value = "white") +
-  geom_point(data = square, aes(lon, lat), size = 0.1) +
-  geom_mapa() +
-  facet_wrap(~run) +
-  labs(fill = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-
-  plot_layout(ncol = 2, widths = c(3, 1)) & theme(legend.position = 'bottom')
-
-```
-
-### Temperatura de brillo
-
-```{r eval=FALSE, include=FALSE}
-
-scale_color_topes <- function(colours = hex, 
-                              limits = c(-90, 50), 
-                              breaks = seq(-90, 50, 20),
-                              guide = guide_colorbar(barheight = 15),
-                              ...) {
-  # https://github.com/garciafido/cima-goes/tree/master/src/cima/goes/img
-  topes <- data.table::fread("../../temp_R/paleta_topes")
-  colours <- rgb(topes$V2, topes$V3, topes$V4, maxColorValue = 255)
-
-  scale_color_gradientn(colours = colours, 
-                        limits = limits, 
-                        breaks = breaks, 
-                        guide = guide, ...) 
-}
-
-purrr::walk(c(0:30), function(h) {
-
-  file_path <- paste0("~/datosmunin/EXP/E[4-7]/FCST/2018112200/CRTM/01/G16_20181122000000_", formatC(h, flag = "0", width = 3), "_00.nc")
-
-  files <- Sys.glob(file_path)
-
-  crtm <- purrr::map(files, function(f) {
-
-    meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/FCST/{fcst_exp}/CRTM/{mem}/G16_{fcst}_{hour}_{minutes}.nc")
-
-    ReadNetCDF(f, vars = c(lat = "XLAT", lon = "XLONG", "CH10")) %>% 
-      .[, CH10 := CH10 - 273.15] %>% 
-      .[, ":="(exp = meta[[1]][["exp"]],
-               mem = meta[[1]][["mem"]],
-               fcst_exp = meta[[1]][["fcst_exp"]],
-               date = ymd_hms(meta[[1]][["fcst"]]) + 
-                 hours(as.numeric(meta[[1]][["hour"]])) +
-                 minutes(as.numeric(meta[[1]][["minutes"]])))]
-
-  }) %>% rbindlist()
-
-  crtm %>% 
-    ggplot(aes(lon, lat)) +
-    geom_point(aes(color = CH10), size = 0.7) +
-    # scale_color_viridis_c() +
-    scale_color_topes(guide = guide_colorbar(barheight = 20)) +
-    geom_mapa() +
-    facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                 E6 = "SATWND", E7 = "RAD"))) +
-    labs(caption = crtm[,date[1]]) +
-    theme_minimal() +
-    theme(panel.ontop = TRUE,
-          panel.grid = element_line(linetype = 3))
-
-  ggsave(paste0("CH10_2018112200_", format(crtm[,date[1]], "%Y%m%d%H%M"), ".png"))
-
-})
-```

### 2021-03-14 analysis/chapters/paper_ana_20181122.Rmd
Tabla 1 actualizada

@@
 ```{r dominio, fig.cap="The horizontal domain used on the simulations (black box) and the inner domain used for the validation / evaluation (grey box).
-The topography in meters is also shown (filled contours).", fig.height=4}
+The topography in meters is also shown (filled contours).", fig.height=5}
+
+oficiales <- fread(here("analysis", "data", "raw_data", "E4_asim_conv_20181121120000.ensmean"), 
+                   na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
+  .[, c("V2", "V4") := NULL] %>% 
+  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>% 
+  .[type %in% c(181, 281)] %>% unique(by = c("stationID")) %>%
+  .[!str_detect(stationID, pattern = "[A-Z]")] %>% 
+  .[, source := "Sfc - Official"]
+
+no_oficiales <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
+                      na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
+  .[, c("V2", "V4") := NULL] %>% 
+  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>% 
+  .[type %in% c(181, 187)] %>% 
+  .[!str_detect(stationID, "SMN")] %>%
+  .[!(stationID %in% oficiales$stationID)] %>% 
+  unique(by = c("stationID")) %>% 
+  .[, source := "Sfc - Non-official"]
+
+obs <- rbind(no_oficiales, oficiales)
 
@@
   .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-
   ggplot(aes(x, y)) +
@@
   geom_mapa() +
+  geom_point(data = obs, aes(ConvertLongitude(lon), lat, 
+                             color = source, shape = source), 
+             size = 0.9, alpha = 0.8) + 
+  scale_shape_manual(name = "Surface stations",  
+                     breaks = c("Sfc - Non-official", "Sfc - Official"),
+                     labels = c("ASWS","CSWS"), values = c(15, 17),
+                     guide = guide_legend(override.aes = list(size = 2))) +
+  scale_color_manual(name =  "Surface stations", 
+                     values = c("Sfc - Non-official" = "#00695c", 
+                                "Sfc - Official" = "#FD8002"),
+                     breaks = c("Sfc - Non-official", "Sfc - Official"),
+                     labels = c("ASWS","CSWS"),
+                     guide = guide_legend(override.aes = list(size = 2))) +
   geom_point(data = square, aes(lon, lat), size = 0.2) +
@@
   theme_minimal() +
-  theme(legend.position = "bottom", legend.box.spacing = margin(0, 0, 0, 0)) 
+  theme(legend.position = "bottom", 
+        legend.box = "vertical",
+        legend.margin = margin(),
+        legend.box.spacing = margin(0, 0, 0, 0)) 
 ```
+
+```{r obs}
+tribble(
+  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
+  "PREPBUFR", "x", "x", "x", "x",
+  "AUTSFC", " ", "x", "x", "x",
+  "SATWND", " ", " ", "x", "x",
+  "RAD", " ", " ", " ", "x",
+) %>% 
+  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
+      caption = "Observation types assimilated in each experiment.") %>% 
+  kable_classic_2(full_width = F)
+```

### 2021-03-14 analysis/chapters/paper_ana_20181122.Rmd
Tabla radianzas

@@
 ```
+
+```{r}
+fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
+  .[, ":="(sensor = toupper(sensor),
+           plataforma = toupper(plataforma),
+           prop = scales::percent(prop/100, accuracy = 0.01, trim = TRUE))] %>% 
+  .[order(.[,'sensor']), ] %>% 
+  .[] %>% 
+  kbl(col.names = c("Sensor", "Plataform", "Number of assimilated\nchannels", "Percentage over total"),
+      caption = "List of the available sensors in several platforms, the number of accepted channels and the percentage of assimilated observations calculated over all radiance observations.") %>% 
+  collapse_rows(1) %>% 
+  kable_classic_2(full_width = F)
+```

### 2021-03-15 analysis/scripts/figuras_sondeos.R
Codigo que genera fguras de sondeos vs modelo

@@ -1,0 +1,89 @@
+
+# leo sondeos
+# files <- list.files(path = "/home/paola.corrales/datosmunin/DA/DA_DATA/OBS_RELAMPAGO/sondeos_raw",
+#                     pattern = "cls", full.names = TRUE)
+#
+# sondeos <- purrr::map(files, ~ read_radiosonde_relampago(.x)) %>%
+#   rbindlist()
+# sondeos %>%
+#   unique(by = c("site", "nominal_launch_time")) %>%
+# .[nominal_launch_time %between% c(as_datetime("2018-11-20 18:00:00"),
+#                                   as_datetime("2018-11-23 12:00:00"))] %>%
+#   .[, c("site", "lon", "lat", "nominal_launch_time")] %>%
+#   fwrite("analysis/data/derived_data/lista_sondeos.csv")
+
+run <- "ANA"
+
+colores_exp <- c(E4 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311")
+
+files <- Sys.glob(paste0("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos/", run, "/*"))
+
+sondeos <- purrr::map(files, ~ fread(.x)) %>%
+  rbindlist() %>%
+  .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>%
+  .[, var_type := fcase(variable %in% c("t", "td"), "temperatura",
+                        variable %in% c("u", "v"), "viento")]
+
+lista_sondeos <- fread("analysis/data/derived_data/lista_sondeos.csv")
+
+for (s in seq_len(nrow(lista_sondeos))) {
+
+# s <- 70
+print(s)
+this_site <- lista_sondeos$site[s]
+this_lauch_time <- lista_sondeos$nominal_launch_time[s]
+
+
+sondeos[site == this_site & nominal_launch_time == as_datetime(this_lauch_time) &
+          !is.na(var_type) & alt %between% c(0, 5000)]  %>%
+  ggplot(aes(alt, value)) +
+  geom_line(data = ~.x[exp == "E7"], aes(linetype = variable)) +
+  geom_line(aes(y = fcst_value, color = exp, linetype = variable)) +
+  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT",
+                                                      E6 = "SATWND", E7 = "RAD")) +
[…]

### 2021-03-15 analysis/scripts/count_obs.R
Count obs

@@ -1,0 +1,61 @@
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim*ensmean")
+
+diag <- read_diag_conv(files, "E6", variable = "uv")
+
+diag[type %between% c(240, 260) & var == "u" & usage.flag == 1 &
+       date %between% c(as_datetime("2018-11-20 18:00:00"), as_datetime("2018-11-23 12:00:00"))] %>%
+  .[, sat_type := case_when(
+    type == 240 ~ "GOES SW winds",
+    type == 241 ~ "India",
+    type == 242 ~ "JMA Visible",
+    type == 243 ~ "EUMETSAT visible",
+    type == 244 ~ "AVHRR winds",
+    type == 245 ~ "GOES IR",
+    type == 246 ~ "GOES WV cloud top",
+    type == 247 ~ "GOES WV deep layer",
+    type == 248 ~ "GOES cloud top (sounder)",
+    type == 249 ~ "GOES deep layer (sounder)",
+    type == 250 ~ "JMA WV deep layer",
+    type == 251 ~ "GOES visible",
+    type == 252 ~ "JMA IR winds",
+    type == 253 ~ "EUMETSAT IR winds",
+    type == 254 ~ "EUMETSAT WV deep layer winds",
+    type == 257 ~ "MODIS IR",
+    type == 258 ~ "MODIS WV cloud top",
+    type == 259 ~ "MODIS WV deep layer winds",
+    type == 260 ~ "VIIR IR winds")] %>%
+  separate(sat_type, into = c("sat", "wn"), remove = FALSE) %>%
+  .[, .N, by = sat]
+
+
+diag <- read_diag_conv(files, "E6", variable = "t")
+
+diag[type %between% c(130, 134) & usage.flag == 1] %>%
+  .[, .N, by = type]
+
+
+multiespectrales <- c("airs_aqua",
+                      "iasi_metop-a",
+                      "iasi_metop-b")
+
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/*/diagfiles/asim*ensmean")
+
+files <- files[str_detect(files, "conv", negate = TRUE)]
+
+diag <- read_diag_rad(files, "E7")
+
+diag[, tipo := fifelse(sensor %in% multiespectrales, "multiespectral", "no-multiespectral")] %>%
[…]

### 2021-03-16 analysis/scripts/count_obs.R
Porcentaje de convencionales y satwnd por experimento

@@ -59,3 +59,62 @@ diag[, tipo := fifelse(sensor %in% multiespectrales, "multiespectral", "no-multi
 asimilable <- satinfo[iuse == 1, .N, by = sensor]
 
 diag[tipo == "no-multiespectral", unique(sensor)]
+
+
+
+# Porcentaje de obs por experimento ---------------------------------------
+
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E4/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+conv <- read_diag_conv(files, exp = "E4", member = "000")
+
+conv[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                          type %in% c(120, 220, 221), "ADPUPA",
+                          type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                          type %in% c(290), "ASCATW",
+                          type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                          type %in% c(240:260), "SATWND")]
+
+conv[usage.flag == 1 & rerr != 1.0e+10] %>%
+  .[, N := .N] %>%
+  .[, .(exp = "E4",
+        N = .N,
+        proporcion = .N/unique(N)), by = bufr_code] %>%
+  fwrite(here("analysis/data/derived_data/count_obs_E4.csv"))
+
+
+files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+aut <- read_diag_conv(files, exp = "E5", member = "000")
+
+aut[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                         type %in% c(120, 220, 221), "ADPUPA",
+                         type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                         type %in% c(290), "ASCATW",
+                         type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                         type %in% c(240:260), "SATWND")]
+
+aut[usage.flag == 1 & rerr != 1.0e+10] %>%
+  .[, N := .N] %>%
+  .[, .(exp = "E5",
+        N = .N,
+        proporcion = .N/unique(N)), by = bufr_code] %>%
+  fwrite(here("analysis/data/derived_data/count_obs_E5.csv"))
+
[…]

### 2021-03-16 analysis/chapters/paper_ana_20181122.Rmd
Intento de grafico resumen de observaciones convencionales y satwnd

@@
                              color = source, shape = source), 
-             size = 0.9, alpha = 0.8) + 
+             size = 1, alpha = 0.8) + 
   scale_shape_manual(name = "Surface stations",  
@@
 ```{r}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E4/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+conv <- read_diag_conv(files, exp = "E4", member = "000")
+
+conv[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                          type %in% c(120, 220, 221), "ADPUPA",
+                          type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                          type %in% c(290), "ASCATW",
+                          type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                          type %in% c(240:260), "SATWND")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+aut <- read_diag_conv(files, exp = "E5", member = "000")
+
+aut[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                         type %in% c(120, 220, 221), "ADPUPA",
+                         type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                         type %in% c(290), "ASCATW",
+                         type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                         type %in% c(240:260), "SATWND")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+satwnd <- read_diag_conv(files, exp = "E6", member = "000")
+
+satwnd[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                            type %in% c(120, 220, 221), "ADPUPA",
+                            type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                            type %in% c(290), "ASCATW",
+                            type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                            type %in% c(240:260), "SATWND")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+p1 <- rbind(conv, aut, satwnd) %>% 
+  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
+  .[, .(count_obs = .N), by = .(exp, lon.box, lat.box, date)] %>% 
+  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lon.box, lat.box)] %>% 
+  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
+  geom_raster(aes(fill = obs_cycle), alpha = 0.8) +
+  scale_fill_viridis_c(option = "E", trans = scales::log10_trans(),
+                       guide = guide_colorbar(barwidth = 20)) +
+  geom_mapa() +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E7 = "RAD"))) +
+  labs(fill = "Mean number of\nobs per cycle") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p2 <- rbind(conv, aut, satwnd) %>% 
+  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
+  .[, ":="(lev.box = cut_round(pressure, breaks = seq(50, 1050, 50)))] %>% 
+  .[, .(count_obs = .N), by = .(exp, lev.box, date)] %>% 
+  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lev.box)] %>% 
+  ggplot(aes(exp, lev.box)) +
+  geom_tile(aes(fill = obs_cycle)) +
+  scale_fill_viridis_c(option = "E", trans = scales::log10_trans(),
+                       guide = guide_colorbar(barheight = NULL)) +
+  scale_x_discrete(labels = c("CONV", "AUT", "SATWND")) +
+  scale_y_reverse() +
+  labs(x = NULL, y = "Pressure level (hPa)", fill = "Mean number of\nobs per cycle") +
+  theme_minimal() +
+  theme(aspect.ratio = 2.5/1, plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p3 <- rbind(conv, aut, satwnd) %>% 
+  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
+  .[, .(count_obs = .N), by = .(exp, date)] %>% 
+  ggplot(aes(date, count_obs)) +
+  geom_line(aes(color = exp)) +
+  scale_color_manual(values = colores_exp, 
+                     labels = c(E4 = "CONV", E5 = "AUT", 
+                                E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(y = "Mean number of obs per cycle", color = NULL, x = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        legend.position = "bottom", 
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+cowplot::plot_grid(p1, NULL, p2, p3, ncol = 3,  rel_widths = c(1.3, -0.3, 1))
+
+```
+
+```{r}
+
+```
+
+```{r}
 fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 

### 2021-03-16 analysis/chapters/paper_ana_20181122.Rmd
Figura extra con sondeos de relampago

@@
         legend.box.spacing = margin(0, 0, 0, 0))
+
+lista_sondeos <- fread(here("analysis/data/derived_data/lista_sondeos.csv")) %>% 
+  .[, periodo := fcase(nominal_launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")),  "IOP07", 
+                       nominal_launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")),  "IOP08", 
+                       default = "Others")
+  ]
+
+fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
+                    breaks = seq(0, 6000, 500)) +
+  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
+                      name = NULL,
+                      breaks = seq(0, 6000, 1000),
+                      guide = guide_colorstrip(barwidth = 20,
+                                               barheight = 0.8)) +
+  geom_mapa() +
+  coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
+  geom_jitter(data = lista_sondeos, aes(lon, lat, 
+                                        color = periodo,
+                                        shape = periodo), 
+              alpha = 0.5, size = 1.5, width = 0.03, height = 0.03) +
+  scale_color_brewer(palette = "Dark2",
+                     guide = guide_legend(override.aes = list(size = 2,
+                                                              alpha = 1))) +
+  # geom_point(data = obs, aes(ConvertLongitude(lon), lat, 
+  #                            color = source, shape = source), 
+  #            size = 1, alpha = 0.8) + 
+  # scale_shape_manual(name = "Surface stations",  
+  #                    breaks = c("Sfc - Non-official", "Sfc - Official"),
+  #                    labels = c("ASWS","CSWS"), values = c(15, 17),
+  #                    guide = guide_legend(override.aes = list(size = 2))) +
+  # scale_color_manual(name =  "Surface stations", 
+  #                    values = c("Sfc - Non-official" = "#00695c", 
+  #                               "Sfc - Official" = "#FD8002"),
+  #                    breaks = c("Sfc - Non-official", "Sfc - Official"),
+#                    labels = c("ASWS","CSWS"),
+#                    guide = guide_legend(override.aes = list(size = 2))) +
+# geom_point(data = square, aes(lon, lat), size = 0.2) +
+  # geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
+labs(color = NULL, shape = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom", 
+        legend.box = "vertical",
+        legend.margin = margin(),
+        legend.box.spacing = margin(0, 0, 0, 0))
 ```
@@
   geom_raster(aes(fill = obs_cycle), alpha = 0.8) +
-  scale_fill_viridis_c(option = "E", trans = scales::log10_trans(),
+  scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
                        guide = guide_colorbar(barwidth = 20)) +
@@
   geom_tile(aes(fill = obs_cycle)) +
-  scale_fill_viridis_c(option = "E", trans = scales::log10_trans(),
+  scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
                        guide = guide_colorbar(barheight = NULL)) +
@@
 ```{r}
+files <- Sys.glob(here("analysis/data/derived_data/count_obs_E*"))
 
+purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
+  .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
+  dcast(bufr_code ~ exp, value.var = "porcentaje") %>% 
+  .[order(.[,'bufr_code']), ] %>% 
+  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
+  kable_classic_2(full_width = F)
+
+purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
+  # .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
+  dcast(bufr_code ~ exp, value.var = "N") %>% 
+  .[order(.[,'bufr_code']), ] %>% 
+  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
+  kable_classic_2(full_width = F)
 ```

### 2021-03-17 analysis/scripts/campos_ana.R
Figuras de t y td en bach

@@ -1,0 +1,191 @@
+library(tidyverse)
+library(lubridate)
+library(data.table)
+library(unglue)
+library(metR)
+library(mesoda)
+source(here::here("temp_R/postprocesamiento.R"))
+
+geom_mapa <- function(fill = NA) {
+  list(geom_sf(data = map_arg, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+       geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+       coord_sf(ylim = c(-34, -29), xlim = c(-66, -63)),
+       scale_x_longitude(ticks = ),
+       scale_y_latitude(ticks = 1))
+}
+
+
+wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
+ini_date <- ymd_hms("20181120180000")
+ciclos <- 67
+
+dates <- seq(ini_date, by = "hour",
+                    length.out = ciclos)
+
+# Td2m
+
+for (d in seq_along(dates)) {
+
+  print(dates[d])
+
+
+  files_ana <- Sys.glob(paste0(wrf_path, "E[4-7]/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.ensmean"))
+
+  files_gue <- Sys.glob(paste0(wrf_path, "E[4-7]/GUESS/", format(dates[d], "%Y%m%d%H%M%S"), "/wrfarw.ensmean"))
+
+  ana <- purrr::map(files_ana, function(f) {
+
+    metadatos <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{fecha}/analysis.ensmean")
+
+
+    ReadNetCDF(f, vars = c("QVAPOR", "P", "PB",
+                           lon = "XLONG", lat = "XLAT"),
+               subset = list(bottom_top = c(1:3))) %>%
+      .[, ":="(td = td(QVAPOR, P+PB),
+               exp = metadatos[[1]][["exp"]],
+               date = metadatos[[1]][["fecha"]])] %>%
+      .[, .(td2 = mean(td)), by = .(lon, lat, exp, date)]
+  }) %>%
+    rbindlist() %>%
+    .[, c("x", "y") := wrf_project(lon, lat)]
+
+  guess <- purrr::map(files_gue, function(f) {
+
[…]

### 2021-04-11 analysis/chapters/experimento_20181122.Rmd
Comparación E7 y E8

@@
 oficiales <- obsCONV[type %in% c(181, 281)] %>% unique(by = c("stationID")) %>%
+  .[!str_detect(stationID, pattern = "[A-Z]")] %>% 
   .[, source := "Sfc - Official"]
@@
   .[!str_detect(stationID, "SMN")] %>%
-  .[!stationID %in% oficiales$stationID] %>% 
+  .[!(stationID %in% oficiales$stationID)] %>% 
   unique(by = c("stationID")) %>% 
@@
 ```{r rad-wv}
-files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121020000/diagfiles*/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121140000/diagfiles*/asim_iasi_metop-b*ensmean"))
+files <- c(Sys.glob("/home/paola.corrales/datosmunin/EXP/E8/ANA/20181121020000/diagfiles*/asim_iasi*ensmean"), Sys.glob("/home/paola.corrales/datosmunin/EXP/E8/ANA/20181121140000/diagfiles*/asim_iasi_metop-b*ensmean"))
 
@@
   .[sensor == "iasi_metop-b" & 
-      channel == 335 & 
+      channel == 3049 & 
       errinv %between% c(0.0000316227, 1000) & 
@@
 ```{r}
-files <- Sys.glob("~/datosmunin/EXP/E7/ANA/2018112*/diagfiles/asim*.ensmean")
+files <- Sys.glob("~/datosmunin/EXP/E8/ANA/2018112*/diagfiles/asim*.ensmean")
 
@@
 
-rad <- read_diag_rad(files, "E7") %>% 
+rad <- read_diag_rad(files, "E8") %>% 
   satinfo[., on = c("sensor", "channel")] %>% 
@@
   geom_density(aes(color = variable, y = stat(density))) +
+  geom_rug(data = ~.x[variable == "diff"]) +
   # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
@@
   geom_density(aes(color = variable)) +
+  geom_rug(data = ~.x[variable == "diff"]) +
   # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
@@
            E6_E5 = E6 - E5,
-           E7_E6 = E7 - E6)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
+           E7_E6 = E7 - E6,
+           E8_E6 = E8 - E6)] %>% 
+  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6", "E8_E6")) %>% 
   ggplot(aes(date, lev)) +
@@
   # scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
+  facet_wrap(~variable, ncol = 2,
+             labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                          "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
+                                                         "E7_E6" = "RAD - SATWND",
+                                                         "E8_E6" = "NEW RAD - SATWND"))) +
   theme_minimal() +
@@
            E6_E5 = E6 - E5,
-           E7_E6 = E7 - E5)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6")) %>% 
+           E7_E6 = E7 - E6,
+           E8_E6 = E8 - E6)] %>% 
+  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6", "E8_E6")) %>% 
   ggplot(aes(date, lev)) +
@@
        x = NULL) +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
+ facet_wrap(~variable, ncol = 2,
+             labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                          "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
+                                                         "E7_E6" = "RAD - SATWND",
+                                                         "E8_E6" = "NEW RAD - SATWND"))) +
   theme_minimal() +
@@
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
+  facet_wrap(~variable, ncol = 2,
+             labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                          "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
-  theme_minimal() +
+                                                         "E7_E6" = "RAD - SATWND",
+                                                         "E8_E6" = "NEW RAD - SATWND"))) +
   theme(legend.position = "bottom",
@@
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  facet_wrap(~variable, labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
+  facet_wrap(~variable, ncol = 2,
+             labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
                                                          "E6_E5" = "SATWND - AUT",
-                                                         "E7_E6" = "RAD - SATWND"))) +
+                                                         "E7_E6" = "RAD - SATWND",
+                                                         "E8_E6" = "NEW RAD - SATWND"))) +
   theme_minimal() +

### 2021-04-23 analysis/chapters/paper_ana_20181122.Rmd
Codigo de figuras de metodos en primera buena versión

--- 2021-03-14 analysis/chapters/paper_ana_20181122.Rmd
@@
 title: "Untitled"
-output: html_document
+output:
+  html_document: default
+  pdf_document: default
 ---
@@
 #Color blind friendly but E4 y E7 parecidos en Desaturated
-colores_exp <- c(E4 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
+colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
 # colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
@@
 
-```{r dominio, fig.cap="The horizontal domain used on the simulations (black box) and the inner domain used for the validation / evaluation (grey box).
+```{r dominio, fig.cap="Figure 1.
+a) The horizontal domain used on the simulations (black box) and the inner domain used for the experiment comparison (red box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles) .
+b) Locations of radiosonde launches during the experiment.
 The topography in meters is also shown (filled contours).", fig.height=5}
 
-oficiales <- fread(here("analysis", "data", "raw_data", "E4_asim_conv_20181121120000.ensmean"), 
+oficiales <- fread(here("analysis", "data", "raw_data", "E2_asim_conv_20181121120000.ensmean"), 
                    na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
@@
 
-fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+lista_sondeos <- fread(here("analysis/data/derived_data/lista_sondeos.csv")) %>% 
+  .[, periodo := fcase(nominal_launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")),  "IOP07", 
+                       nominal_launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")),  "IOP08", 
+                       default = "Others")
+  ]
+
+dominio <- fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)]
+
+dominio %>% 
   ggplot(aes(x, y)) +
@@
                              color = source, shape = source), 
-             size = 0.9, alpha = 0.8) + 
-  scale_shape_manual(name = "Surface stations",  
+             size = 1, alpha = 0.8) + 
+  scale_shape_manual(name = "a) Surface stations",  
                      breaks = c("Sfc - Non-official", "Sfc - Official"),
@@
                      guide = guide_legend(override.aes = list(size = 2))) +
-  scale_color_manual(name =  "Surface stations", 
+  scale_color_manual(name =  "a) Surface stations", 
                      values = c("Sfc - Non-official" = "#00695c", 
@@
   theme(legend.position = "bottom", 
-        legend.box = "vertical",
-        legend.margin = margin(),
-        legend.box.spacing = margin(0, 0, 0, 0)) 
+        legend.box = "vertical") +
+        # ,
+        # legend.margin = margin(),
+        # legend.box.spacing = margin(0, 0, 0, 0)) +
+        # 
+dominio %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
+                    breaks = seq(0, 6000, 500)) +
+  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
+                      name = NULL,
+                      breaks = seq(0, 6000, 1000),
+                      guide = guide_colorstrip(barwidth = 20,
+                                               barheight = 0.8)) +
+  geom_mapa() +
+  coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
+  geom_jitter(data = lista_sondeos, aes(lon, lat, 
+                                        color = periodo,
+                                        shape = periodo), 
+              alpha = 0.5, size = 1.5, width = 0.03, height = 0.03) +
+  scale_color_brewer(palette = "Dark2",
+                     guide = guide_legend(override.aes = list(size = 2,
+                                                              alpha = 1))) +
+  labs(color = "b) Soundings", shape = "b) Soundings") +
+  theme_minimal() +
+  theme(legend.position = "bottom", 
+        legend.box = "vertical") +
+# ,
+#         legend.margin = margin(),
+#         legend.box.spacing = margin(0, 0, 0, 0)) +
+  plot_layout(ncol = 2, guides = "collect") +
+  plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
+  theme(legend.position = 'bottom', legend.box = "vertical")
+```
+
+```{r ciclo, fig.cap="Diagram of analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
+The zoomed figure shows assimilation in one-hour centered window and new boundary conditions from GFSD every 6 hours.", out.height=400, out.width=800, fig.align="center"}
+knitr::include_graphics(here("analysis", "figures", "analysis_cycle.svg"))
 ```
@@
 
+```{r count_obs, fig.height=6, fig.width=10}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+conv <- read_diag_conv(files, exp = "E2", member = "000")
+
+conv[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                          type %in% c(120, 220, 221), "ADPUPA",
+                          type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                          type %in% c(290), "ASCATW",
+                          type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                          type %in% c(240:260), "SATWND")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+aut <- read_diag_conv(files, exp = "E5", member = "000")
+
+aut[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                         type %in% c(120, 220, 221), "ADPUPA",
+                         type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                         type %in% c(290), "ASCATW",
+                         type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                         type %in% c(240:260), "SATWND")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]
+
+satwnd <- read_diag_conv(files, exp = "E6", member = "000")
+
+satwnd[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                            type %in% c(120, 220, 221), "ADPUPA",
+                            type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                            type %in% c(290), "ASCATW",
+                            type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                            type %in% c(240:260), "SATWND")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+p1 <- rbind(conv, aut, satwnd) %>% 
+  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
+  .[, .(count_obs = .N), by = .(exp, lon.box, lat.box, date)] %>% 
+  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lon.box, lat.box)] %>% 
+  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
+  geom_raster(aes(fill = obs_cycle), alpha = 0.8) +
+  scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
+                       guide = guide_colorbar(barwidth = 20)) +
+  geom_mapa() +
+  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E7 = "RAD"))) +
+  labs(fill = "Mean number of\nobs per cycle") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p2 <- rbind(conv, aut, satwnd) %>% 
+  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
+  .[, ":="(lev.box = cut_round(pressure, breaks = seq(50, 1050, 50)))] %>% 
+  .[, .(count_obs = .N), by = .(exp, lev.box, date)] %>% 
+  .[, .(obs_cycle = mean(count_obs, na.rm = TRUE)), by = .(exp, lev.box)] %>% 
+  ggplot(aes(lev.box, obs_cycle)) +
+  geom_line(aes(color = exp)) +
+  geom_point(aes(color = exp, shape = exp), size = 2) +
+    scale_color_manual(values = colores_exp, 
+                     labels = c(E2 = "CONV", E5 = "AUT", 
+                                E6 = "SATWND", E7 = "RAD")) +
+  scale_shape_manual(values = c(15, 16, 17), labels = c(E2 = "CONV", E5 = "AUT", 
+                                E6 = "SATWND", E7 = "RAD")) +
+  # geom_tile(aes(fill = obs_cycle)) +
+  # scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
+  #                      guide = guide_colorbar(barheight = NULL)) +
+  scale_y_log10() +
+  scale_x_reverse() +
+  coord_flip() +
+  labs(color = NULL, shape = NULL,
+       x = "Pressure level (hPa)", y = "Mean number of\nobs per cycle") +
+  theme_minimal() +
+  theme(aspect.ratio = 2.5/1, 
+        legend.position = "bottom",
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+p3 <- rbind(conv, aut, satwnd) %>% 
+  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
+  .[, .(count_obs = .N), by = .(exp, date)] %>% 
+  ggplot(aes(date, count_obs)) +
+  geom_line(aes(color = exp)) +
+  geom_point(aes(color = exp, shape = exp), size = 2) +
+  scale_color_manual(values = colores_exp, 
+                     labels = c(E2 = "CONV", E5 = "AUT", 
+                                E6 = "SATWND", E7 = "RAD")) +
+  scale_shape_manual(values = c(15, 16, 17), labels = c(E2 = "CONV", E5 = "AUT", 
+                                E6 = "SATWND", E7 = "RAD")) +
+  scale_date() +
+  labs(y = "Number of obs per cycle", x = NULL,
+       color = NULL, shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        legend.position = "bottom", 
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+cowplot::plot_grid(p1, NULL, p2, p3, ncol = 3,  rel_widths = c(1.3, -0.3, 1))
+
+```
+
+```{r}
+files <- Sys.glob(here("analysis/data/derived_data/count_obs_E*"))[c(1, 3, 4)]
+
+purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
+  .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
+  dcast(bufr_code ~ exp, value.var = "porcentaje") %>% 
+  .[order(.[,'bufr_code']), ] %>% 
+  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
+  kable_classic_2(full_width = F)
+
+purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
+  # .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
+  dcast(bufr_code ~ exp, value.var = "N") %>% 
+  .[order(.[,'bufr_code']), ] %>% 
+  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
+  kable_classic_2(full_width = F)
+```
+
 ```{r}
@@
       caption = "List of the available sensors in several platforms, the number of accepted channels and the percentage of assimilated observations calculated over all radiance observations.") %>% 
-  collapse_rows(1) %>% 
+  collapse_rows(1,valign = "top") %>% 
+  # column_spec(1, width = "3cm", latex_valign = "p") %>% 
   kable_classic_2(full_width = F)
 ```
+
+```{r rad, fig.cap="Number of radiance observations for each 50 hPa pressure layer over time"}
+
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E8/ANA/*/diagfiles/asim*ensmean")
+
+files <- files[!str_detect(files, "conv")]
+
+rad <- read_diag_rad(files, "E8") %>% 
+  satinfo[., on = c("sensor", "channel")]
+
+rad %>% 
+  .[qc == 0 & iuse == 1] %>% 
+  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
+  .[, .(omb = mean(tbc),
+        n = .N), by = .(date, lev)] %>% 
+  .[lev != 1025] %>% 
+  ggplot(aes(date, fct_rev(lev))) +
+  geom_tile(aes(fill = n)) +
+  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
+                       guide = guide_colourbar(barheight = 18,
+                                               barwidth = 0.8)) +
+
+  scale_date() +
+  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
+  theme_minimal()
+
+```
+```{r level-rad, fig.cap="Nivel de presión asociado a las observaciones en cada ciclo de asimilación"}
+
+rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[sensor %in% c("airs", "iasi") &
+      # errinv %between% c(0.0000316227, 1000) & 
+      # qc == 0 & 
+      # iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
+  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
+  ggplot(aes(channel, Q50)) +
+  # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
+  #             outlier.size = 0.5, size = 0.1) +
+  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
+  geom_line(color = "darkorange") +
+  scale_y_level(name = "Pressure (hPa)") +
+  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
+  labs(x = "", y = "") +
+  theme_minimal() +
+
+  rad %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  .[!sensor %in% c("airs", "iasi") &
+      # errinv %between% c(0.0000316227, 1000) & 
+      # qc == 0 & 
+      # iuse == 1 &
+      peakwt %between% c(0.001, 1200)] %>% 
+  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
+  ggplot(aes(channel, Q50)) +
+  # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
+  #             outlier.size = 0.5, size = 0.1) +
+  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
+  geom_line(color = "darkorange") +
+  scale_y_level(name = "Pressure (hPa)") +
+  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
+  labs(y = "", x = "Channel index") +
+  theme_minimal() +
+  plot_layout(ncol = 1)
+```
+
+## Results
+
+### Ensemble reliability
+
+```{r rcrv-sfc, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree", fig.scap=c("Wind", "Temperature")}
+
+rbind(fread(here("analysis", "data", "derived_data", "rcrv_V_box.csv")),
+      fread(here("analysis", "data", "derived_data", "rcrv_t_box.csv"))) %>%   
+  # .[var %in% c("u", "v") & type %in% c(281, 287)] %>% 
+  .[, type.obs := type] %>% 
+  .[, type.obs := fcase(type.obs == 181, 281,
+                        type.obs == 187, 287,
+                        type.obs == 281, 281,
+                        type.obs == 287, 287)] %>% 
+  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
+  geom_raster(aes(fill = sd.y)) +
+  geom_mapa() +
+  scale_fill_viridis_c(breaks = seq(0, 5, 0.5),
+                       limits = c(0, 5),
+                       direction = -1,
+                       guide = guide_colorstrip(inside = TRUE,
+                                                barwidth = 20,
+                                                barheight = 0.8)) +
+  facet_grid(var~type.obs, 
+             labeller = labeller(type.obs = c("281" = "ADPSFC",
+                                              "287" = "ADPSFC\n no pressure",
+                                              "181" = "ADPSFC",
+                                              "187" = "ADPSFC\n no pressure"),
+                                 var = c("v" = "Wind",
+                                         "t" = "Temperature"))) +
+  labs(fill = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
+
+```{r rcrv-profile, fig.cap="Profiles of mean and standar deviation of rcrv for sounding and aircraft observations."}
+
+rbind(fread(here("analysis", "data", "derived_data", "rcrv_V_perfil.csv")), 
+      fread(here("analysis", "data", "derived_data", "rcrv_t_perfil.csv"))) %>% 
+  obs.type[., on = "type"] %>%   
+  .[type != 130] %>%
+  melt(measure.vars = c("mean.y", "sd.y")) %>% 
+  ggplot(aes(error.level, value)) +
+  geom_hline(yintercept = c(0, 1), color = "grey20") +
+  geom_line(aes(color = factor(code), linetype = variable)) +
+  scale_linetype_manual(values = c("dashed", "solid"),
+                        labels = c("Mean RCRV", "SD RCRV")) +
+  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
+  coord_flip() +
+  facet_wrap(~ var, labeller = labeller(var = c("t" = "Temperature",
+                                                "u" = "U wind",
+                                                "v" = "Wind"))) +
+  labs(y = NULL,
+       linetype = "",
+       color = NULL) +
+  theme_minimal() 
+```

### 2021-04-23 analysis/scripts/RCRV_satwnd.R
Calculo de RCRV para satwind

@@ -1,0 +1,42 @@
+source("help_functions.R")
+path <- "/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim_conv_2018112*"
+
+perfiles <- read_diag(path, variable = c("uv")) %>%
+  .[, c("error", "nivel.error") := input_obs_error(var, type, pressure)]
+
+RCRV <- get_RCRV(perfiles[type %between% c(240, 260)], tipo = c("perfil")) %>%
+  .[, sat_type := case_when(
+    type == 240 ~ "GOES SW winds",
+    type == 241 ~ "India",
+    type == 242 ~ "JMA Visible",
+    type == 243 ~ "EUMETSAT visible",
+    type == 244 ~ "AVHRR winds",
+    type == 245 ~ "GOES IR",
+    type == 246 ~ "GOES WV cloud top",
+    type == 247 ~ "GOES WV deep layer",
+    type == 248 ~ "GOES cloud top (sounder)",
+    type == 249 ~ "GOES deep layer (sounder)",
+    type == 250 ~ "JMA WV deep layer",
+    type == 251 ~ "GOES visible",
+    type == 252 ~ "JMA IR winds",
+    type == 253 ~ "EUMETSAT IR winds",
+    type == 254 ~ "EUMETSAT WV deep layer winds",
+    type == 257 ~ "MODIS IR",
+    type == 258 ~ "MODIS WV cloud top",
+    type == 259 ~ "MODIS WV deep layer winds",
+    type == 260 ~ "VIIR IR winds")]
+
+RCRV %>%
+  melt(measure.vars = c("mean.y", "sd.y")) %>%
+  ggplot(aes(nivel.error, value)) +
+  geom_hline(yintercept = c(0, 1), color = "darkgrey") +
+  geom_line(aes(color = factor(sat_type), linetype = variable)) +
+  scale_color_brewer(palette = "Set1") +
+  scale_x_level() +
+  coord_flip() +
+  facet_wrap(~ var, scales = "free_x") +
+  labs(title = "Reduced Centered Random Variable",
+       subtitle = "Satellite wind",
+       linetype = "",
+       color = "Obs type") +
+  theme_minimal()

### 2021-04-23 analysis/scripts/2018112300_CONV.R
Analisis de caso CONV 2018112300

@@ -18,8 +18,9 @@ geom_mapa <- function(fill = NA) {
 diag <- read_diag_conv("/home/paola.corrales/datosmunin/EXP/E4/ANA/20181123000000/diagfiles/asim_conv_20181123000000.ensmean", exp = "CONV") %>%
   .[, ":="(lon = ConvertLongitude(lon),
            obs = fcase(var == "t", obs - 273.15,
-                       var == "q", obs*1000,
-                       var %in% c("v", "u"), obs))]
+                       var %in% c("q", "v", "u"), obs),
+           run = "guess")]
+
 
 diag[usage.flag == 1 & lat %between% c(-34.5, -28.5) & lon %between% c(-66.5, -62.5), .N, by = .(var)]
 
@@ -30,6 +31,188 @@ diag[usage.flag == 1 & lat %between% c(-32, -30) & lon %between% c(-66, -64)] %>
   scale_y_level() +
   facet_wrap(~var, scales = "free_x")
 
+
+diag_ana <- read_diag_conv("/home/paola.corrales/datosmunin/asim_conv_20181123000000.ensmean", exp = "CONV") %>%
+  .[, ":="(lon = ConvertLongitude(lon),
+           obs = fcase(var == "t", obs - 273.15,
+                       var %in% c("q", "v", "u"), obs),
+           run = "analisis")]
+
+
+rbind(diag, diag_ana) %>%
+  .[usage.flag == 1 & lat %between% c(-32, -30) & lon %between% c(-66, -64)] %>%
+  .[var %in% c("q") & type == 120] %>%
+  .[, guess := obs - obs.guess] %>%
+  # .[]
+  # .[, td := td(obs, pressure)] %>%
+  melt(measure.vars = c("obs", "guess")) %>%
+  ggplot(aes(pressure, value*1000)) +
+  geom_line(data = ~.x[variable == "obs"]) +
+  geom_point(data = ~.x[variable == "obs"], aes(shape = "obs")) +
+  geom_line(data = ~.x[variable != "obs"], aes(color = run)) +
+  geom_point(data = ~.x[variable != "obs"], aes(color = run)) +
+  coord_flip() +
+  scale_x_level() +
+  labs(shape = NULL, color = NULL, y = "q (g/Kg)") +
+  theme_minimal()
+facet_wrap(~variable, scales = "free_x")
+
+
+rbind(diag, diag_ana) %>%
+  .[usage.flag == 1 & lat %between% c(-32, -30) & lon %between% c(-66, -64)] %>%
[…]

### 2021-04-26 analysis/chapters/experimento_20181122.Rmd
Comparación entre experimentos

@@
 
-experiments <- c("E4", "E5", "E6", "E7")
+experiments <- c("E4", "E5", "E6", "E7", "E8")
 #Color blind friendly but E4 y E7 parecidos en Desaturated
@@
 global_qv_E7 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E7.rds")
-
-rbind(global_qv_E6, global_qv_E7) %>%
+global_qv_E8 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E8.rds")
+
+rbind(global_qv_E6, global_qv_E7, global_qv_E8) %>%
   # .[exp == "E7"] %>% 
@@
   scale_date() +
-  scale_linetype_manual(values = c(2, 1), labels = c(E4 = "CONV", E5 = "AUT",
-                                                     E6 = "SATWND", E7 = "RAD")) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  # facet_wrap(~nivel, scales = "free_y") +
+  # scale_linetype_manual(values = c(2, 1, 3), labels = c(E4 = "CONV", E5 = "AUT",
+  #                                                    E6 = "SATWND", E7 = "RAD", E8 = "NEW RAD")) +
+  # scale_color_manual(values = c(colores_exp, "black"), labels = c(E4 = "CONV", E5 = "AUT",
+  #                                                     E6 = "SATWND", E7 = "RAD", E8 = "NEW RAD")) +
+  # # facet_wrap(~nivel, scales = "free_y") +
   labs(x = NULL, y = "Accumulated specific humidity (g/Kg)",
@@
 ```{r rcrv-profile-rad, fig.cap="Profiles of mean and standar deviation of rcrv for radiances observations."}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:4]
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:6]
 
 RCRV <- purrr::map(files, function(f) {
-  fread(f)
+  meta <- unglue::unglue(basename(f), "rcrv_{sensor}_perfil.csv")
+
+  fread(f) %>%
+    .[, sensor := meta[[1]][["sensor"]]]
 }) %>%
@@
   dcast(lev + date ~ exp, value.var = "T") %>% 
-  .[, ":="(E5_E4 = E5 - E4,
+  .[, ":="(E5_E2 = E5 - E2,
            E6_E5 = E6 - E5,
@@
            E8_E6 = E8 - E6)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6", "E8_E6")) %>% 
+  melt(measure.vars = c("E5_E2", "E6_E5", "E7_E6", "E8_E6")) %>% 
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value, fill = stat(level))) +
+  geom_contour_fill(aes(z = value, fill = stat(level)),
+                         breaks = c(seq(-2, 2, 0.2), Inf)) +
   scale_fill_divergent_discretised(name = NULL,
@@
   facet_wrap(~variable, ncol = 2,
-             labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
+             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
                                                          "E6_E5" = "SATWND - AUT",
@@
   dcast(lev + date ~ exp, value.var = "QVAPOR") %>% 
-  .[, ":="(E5_E4 = E5 - E4,
+  .[, ":="(E5_E2 = E5 - E2,
            E6_E5 = E6 - E5,
@@
            E8_E6 = E8 - E6)] %>% 
-  melt(measure.vars = c("E5_E4", "E6_E5", "E7_E6", "E8_E6")) %>% 
+  melt(measure.vars = c("E5_E2", "E6_E5", "E7_E6", "E8_E6")) %>% 
   ggplot(aes(date, lev)) +
   geom_contour_fill(aes(z = value*1000, fill = stat(level)), 
-                    breaks = c(seq(-1, 1.6, 0.2), Inf)) +
-  scale_fill_divergent_discretised(name = NULL, limits = c(-1, 1.6),
+                    breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
+  scale_fill_divergent_discretised(name = NULL, limits = c(-1.6, 1.6),
                                    guide = guide_colorsteps(barwidth = 25,
@@
  facet_wrap(~variable, ncol = 2,
-             labeller = labeller(variable = c("E5_E4" = "AUT - CONV",
+             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
                                                          "E6_E5" = "SATWND - AUT",
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = value, fill = stat(level)), breaks = c(seq(-1.6, 2, 0.2), Inf)) +
+  # geom_contour(aes(z = value)) +
+  geom_contour_fill(aes(z = value, fill = stat(level)), breaks = c(seq(-2, 2, 0.2), Inf)) +
   scale_fill_divergent_discretised(name = NULL,
-                                   limits = c(-1.6, 2),
+                                   # limits = c(-2, 2),
                                    guide = guide_colorsteps(barwidth = 25,
                                                             barheight = 0.8)) +
-  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7)] %>% 
+  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7, E8_E6 = E8)] %>%
                   melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
                 breaks = seq(0, 30, 3), size = 0.2, linetype = 1, color = "grey30") +
-  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7, E8_E6 = E8)] %>%
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
@@
                                                          "E8_E6" = "NEW RAD - SATWND"))) +
+  theme_minimal() +
   theme(legend.position = "bottom",
@@
   geom_contour_fill(aes(z = value, fill = stat(level)), 
-                    breaks = c(-Inf, seq(-1.6, 1.2, 0.2), Inf)) +
-  scale_fill_divergent_discretised(name = NULL, limits = c(-1.6, 1.2),
+                    breaks = c(-Inf, seq(-1.6, 1.6, 0.2), Inf)) +
+  scale_fill_divergent_discretised(name = NULL, limits = c(-1.6, 1.6),
                                    guide = guide_colorsteps(barwidth = 25,
                                                             barheight = 0.8)) +
-  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7)] %>% 
+  geom_contour2(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7, E8_E6 = E8)] %>% 
                   melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
                 breaks = seq(-9, 0, 1), size = 0.2, linetype = 2, color = "grey30") +
-  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 =E7)] %>% 
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E4 = E5, E6_E5 = E6, E7_E6 = E7, E8_E6 = E8)] %>% 
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
@@
 ```{r q-low, eval=FALSE, fig.align="center", message=FALSE, warning=FALSE, include=FALSE}
-fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
-                     exp = c("E4", "E5", "E6", "E7"))
+fcsts <- expand_grid(fcsts = c("20181122000000"),
+                     exp = c("E4", "E5", "E6", "E7", "E8"))
 
@@
   # cordillera +
-  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
-  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
+  geom_mapa() +
   facet_grid(date ~ exp, 
@@
 
-files <- Sys.glob(paste0(derived_data, "mucape/*")) 
+files <- Sys.glob(paste0(derived_data, "mucape/mcape_ana_E*_20181122000000.nc")) 
 files <- files[!str_detect(files, c("E3"))]
@@
 fcsts <- expand_grid(fcsts = c("20181122000000"),
-                     exp = c("E4", "E5", "E6", "E7"))
+                     exp = c("E4", "E5", "E6", "E7", "E8"))
 
@@
            "20181122160000", "20181122190000")
-files <- paste0(derived_data, "dbz/maxdbz_ana_E7_", dates, ".nc")
+files <- paste0(derived_data, "dbz/maxdbz_ana_E8_", dates, ".nc")
 
@@
 
-dates <- rep(c("20181122100000"), 1, each = 4)
+dates <- rep(c("20181122110000"), 1, each = 5)
 files <- paste0(derived_data, "dbz/maxdbz_ana_", experiments, "_", dates, ".nc")
@@
 dbz %>% 
+  .[exp %in% c("E5", "E6", "E8")] %>% 
   .[, date := factor(date)] %>% 
@@
                     proj = norargentina_lambert,
-                    breaks = seq(15, 75, 5)) +
-  scale_fill_manual(values = ref, drop = FALSE, guide = guide_colorsteps(barwidth = 20, 
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 30, 
                                              barheight = 0.8)) +
@@
   geom_mapa() +
-  facet_wrap(~exp, ncol = 2, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                         E6 = "SATWND", E7 = "RAD"))) +
+  facet_wrap(~exp, ncol = 3, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                         E6 = "SATWND", E7 = "RAD",
+                                                         E8 = "NEW RAD"))) +
   labs(x = NULL, y = NULL, fill = NULL) +
@@
 ```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-7].csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-8].csv")
 
@@
   # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = c(colores_exp, "black"), labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD", E8 = "NEW RAD")) +
   scale_date(20181122000000, 20181123000000, "12 hours") +
@@
 ```{r message=FALSE, warning=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/ANA/sondeos/*")
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos/ANA/*")
 
@@
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = c(colores_exp, "black"), labels = c(E4 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E7 = "RAD", E8 = "NEW RAD")) +
   coord_flip() +

### 2021-04-26 analysis/chapters/paper_ana_20181122.Rmd
Agrego figuras de resultados

@@
 library(kableExtra)
+library(tagger)
 
@@
 
-```{r rcrv-sfc, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree", fig.scap=c("Wind", "Temperature")}
+```{r rcrv-sfc, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree"}
 
@@
                         type.obs == 287, 287)] %>% 
+  .[, var := factor(var, levels = c("v", "t"))] %>% 
   ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
@@
                                          "t" = "Temperature"))) +
+  tag_facets() +
   labs(fill = "") +
@@
 
-```{r rcrv-profile, fig.cap="Profiles of mean and standar deviation of rcrv for sounding and aircraft observations."}
+```{r rcrv-profile, fig.cap="Profiles of the mean and standard deviation of RCRV for upper-air observations.", fig.subcap=c("Sounding and aircraft observations.", "Satellite-derived wind.", "Radiances observations.")}
 
@@
   facet_wrap(~ var, labeller = labeller(var = c("t" = "Temperature",
-                                                "u" = "U wind",
                                                 "v" = "Wind"))) +
+  tag_facets() +
   labs(y = NULL,
@@
        color = NULL) +
-  theme_minimal() 
+  theme_minimal() +
+  theme(legend.position = "bottom", 
+        legend.box = "vertical",
+        aspect.ratio = 2/1)
+
+read_csv(here::here("analysis", "data", "derived_data", "rcrv_satwind.csv")) %>% 
+  melt(measure.vars = c("mean.y", "sd.y")) %>%
+  ggplot(aes(nivel.error, value)) +
+  geom_hline(yintercept = c(0, 1), color = "darkgrey") +
+  geom_line(aes(color = factor(sat_type), linetype = variable)) +
+  scale_color_brewer(palette = "Set1") +
+  scale_linetype_manual(values = c("dashed", "solid"),
+                        labels = c("Mean RCRV", "SD RCRV")) +
+  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
+  coord_flip() +
+  facet_wrap(~ var, scales = "free_x", labeller = labeller(var = c("u" = "U wind",
+                                                                   "v" = "V wind"))) +
+  labs(linetype = NULL,
+       color = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom", 
+        legend.box = "vertical",
+        aspect.ratio = 2/1)
+
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:6]
+
+RCRV <- purrr::map(files, function(f) {
+  meta <- unglue::unglue(basename(f), "rcrv_{sensor}_perfil.csv")
+
+  fread(f) %>%
+    .[, sensor := meta[[1]][["sensor"]]]
+}) %>%
+  rbindlist()
+
+RCRV %>% 
+  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
+  setDT() %>% 
+  melt(measure.vars = c("mean.y", "sd.y")) %>% 
+  .[, .(value = mean(value, na.rm = TRUE)), by = .(sensor, level, variable)] %>% 
+  ggplot(aes(level, value)) +
+  geom_hline(yintercept = c(0, 1), color = "grey20") +
+  geom_line(aes(color = sensor, linetype = variable)) +
+  scale_color_discrete(labels = toupper) +
+  scale_linetype_manual(values = c("dashed", "solid"),
+                        labels = c("Mean RCRV", "SD RCRV")) +
+  coord_flip() +
+  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
+  labs(x = NULL, y = NULL,
+       color = NULL, linetype = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom", 
+        legend.box = "vertical",
+        aspect.ratio = 2/1)
+
 ```
+
+### Impacts of assimilated observations
+
+```{r TQ-diff, fig.cap="Difference between experiments for the average vertical profile of variables over time.", fig.scap=c("Temperature (°C).", "Specific Humidity (g/kg)."), fig.height=3.5}
+files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
+                    pattern = "perfiles_ana")
+
+perfiles <- purrr::map(files, function(f) {
+  exp <- unglue(basename(f), "perfiles_{run}_{exp}.csv")
+  fread(f) %>% 
+    .[, exp := exp[[1]][["exp"]]]
+
+}) %>% 
+  rbindlist() %>% 
+  .[, date := as_datetime(date)] %>% 
+  .[date %between% c(as_datetime("2018-11-20 18:00:00"),
+                     as_datetime("2018-11-23 12:00:00"))]
+
+perfiles %>% 
+  dcast(lev + date ~ exp, value.var = "T") %>% 
+  .[, ":="(E5_E2 = E5 - E2,
+           E6_E5 = E6 - E5,
+           E8_E6 = E8 - E6)] %>% 
+  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>% 
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = value, fill = stat(level)),
+                         breaks = c(seq(-2, 2, 0.2), Inf)) +
+  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
+                                                            barheight = 0.8)) +
+  labs(fill = "Temperature (K)",
+       x = NULL,
+       y = NULL) +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_wrap(~variable, ncol = 3,
+             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
+                                                         "E6_E5" = "SATWND - AUT",
+                                                         "E8_E6" = "RAD - SATWND"))) +
+  tag_facets(position = list(x = 0.1, y = 0.96)) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+perfiles %>% 
+  dcast(lev + date ~ exp, value.var = "QVAPOR") %>% 
+  .[, ":="(E5_E2 = E5 - E2,
+           E6_E5 = E6 - E5,
+           E8_E6 = E8 - E6)] %>% 
+  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>% 
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = value*1000, fill = stat(level)), 
+                    breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
+  scale_fill_divergent_discretised(limits = c(-1.6, 1.6),
+                                   guide = guide_colorsteps(barwidth = 25,
+                                                            barheight = 0.8)) +
+
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  labs(fill = "Specific humidity (g/Kg)",
+       y = NULL,
+       x = NULL) +
+ facet_wrap(~variable, ncol = 3,
+             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
+                                                         "E6_E5" = "SATWND - AUT",
+                                                         "E8_E6" = "RAD - SATWND"))) +
+  tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
+
+```{r UV-diff, fig.cap="Difference between experiments for the average vertical profile of wind over time (shaded) and U and V (only northerly wind) profiles for the experiment with more observations (contours).", fig.height=3.5}
+perfiles %>% 
+  dcast(lev + date ~ exp, value.var = "U") %>% 
+  .[, ":="(E5_E2 = E5 - E2,
+           E6_E5 = E6 - E5,
+           E8_E6 = E8 - E6)] %>% 
+  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>% 
+  ggplot(aes(date, lev)) +
+  # geom_contour(aes(z = value)) +
+  geom_contour_fill(aes(z = value, fill = stat(level)), breaks = c(seq(-2, 2, 0.2), Inf)) +
+  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
+                                                            barheight = 0.8)) +
+  geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>%
+                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
+                breaks = seq(0, 30, 3), size = 0.2, linetype = 1, color = "grey30") +
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>%
+                      melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
+                    breaks = seq(0, 30, 3), color = "grey30", skip = 1,
+                    size = 4, stroke = 0.1) +
+  labs(fill = "U (m/s)",
+       x = NULL,
+       y = NULL) +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_wrap(~variable, ncol = 3,
+             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
+                                                         "E6_E5" = "SATWND - AUT",
+                                                         "E8_E6" = "RAD - SATWND"))) +
+   tag_facets(position = list(x = 0.1, y = 0.96)) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+perfiles %>% 
+  dcast(lev + date ~ exp, value.var = "V") %>% 
+  .[, ":="(E5_E2 = E5 - E2,
+           E6_E5 = E6 - E5,
+           E8_E6 = E8 - E6)] %>% 
+  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>% 
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = value, fill = stat(level)), 
+                    breaks = c(-Inf, seq(-1.6, 1.6, 0.2), Inf)) +
+  scale_fill_divergent_discretised(limits = c(-1.6, 1.6),
+                                   guide = guide_colorsteps(barwidth = 25,
+                                                            barheight = 0.8)) +
+  geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
+                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
+                breaks = seq(-9, 0, 1), size = 0.2, linetype = 2, color = "grey30") +
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
+                      melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
+                    breaks = seq(-9, 0, 1), color = "grey30", skip = 1, 
+                    size = 4, stroke = 0.1) +
+  labs(fill = "V (m/s)",
+       x = NULL,
+       y = NULL) +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_wrap(~variable, ncol = 3,
+             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
+                                                         "E6_E5" = "SATWND - AUT",
+                                                         "E8_E6" = "RAD - SATWND"))) +
+  tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
+
+```{r summary-fields, fig.cap="a) Precipitable water (shaded) and northerly wind (contours, m/s), b) Average potential temperature for the PBL and c) Maximum CAPE and ~0-6 km wind shear over 15 m/s and 30 m/s for each experiment.
+All fields correspond to 00 UTC Nov, 22.", fig.height=10, fig.width=9}
+
+files <- Sys.glob(paste0(derived_data, "pw/*"))[c(1, 4, 5, 7)]
+
+dates <- c("20181122000000")
+
+pw <- purrr::map(files, function(f) {
+  details <- unglue::unglue(basename(f), "pw_ana_{exp}_{date}.nc")[[1]]
+
+  if (details$date %in% dates) {
+
+    ReadNetCDF(f, vars = c(value = "pw", lon = "XLONG", lat = "XLAT")) %>% 
+      .[, `:=`(exp = details$exp, 
+               date = lubridate::ymd_hms(details$date))] %>%
+      .[]
+  }
+}) %>% 
+  rbindlist() %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)]
+
+files <- Sys.glob(paste0(derived_data, "mucape/mcape_ana_E*_20181122000000.nc"))[c(1, 4, 5, 7)]
+
+mucape <- purrr::map(files, function(f) {
+  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]
+
+  if (details$date %in% dates) {
+
+    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
+      .[, `:=`(exp = details$exp, 
+               date = lubridate::ymd_hms(details$date))] %>%
+      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
+      .[]
+  }
+}) %>% 
+  rbindlist()
+
+fcsts <- expand_grid(fcsts = c("20181122000000"),
+                     exp = c("E2", "E5", "E6", "E8"))
+
+ana <- lapply(seq_len(nrow(fcsts)), function(f) {
+
+  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
+    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
+    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
+    .[, theta := ReadNetCDF(paste0(derived_data, "/theta_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(th = "theta"), out = "vector")] %>%
+    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
+    .[, exp := fcsts[f, 2]] %>%  
+    .[]
+}) %>% 
+  rbindlist()
+
+cor <- ana[bottom_top %in% c(1, 16)] %>% 
+  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
+  .[, cortante := Mag(u_16 - u_1, v_16 - v_1)] %>% 
+  .[, ":="(u_1 = NULL, 
+           u_16 = NULL,
+           v_1 = NULL,
+           v_16 = NULL)]
+
+ana[bottom_top %between% c(2, 7), .(v_mean = mean(v)), 
+    by = .(south_north, west_east, date, exp)] %>% 
+  .[pw, on = .NATURAL] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = value, fill = stat(level_d)),
+                    proj = norargentina_lambert) +
+  scale_fill_continuous_divergingx(super = ScaleDiscretised,
+                       palette = "PRGn", mid = 25,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  geom_contour2(aes(z = v_mean, size = stat(level)),
+                proj = norargentina_lambert, color = "black", 
+                breaks = c(-10, -5)) +
+  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
+                        labels = c("-10", "-5")) +
+  # geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
+  #                   proj = norargentina_lambert, color = "darkorange", 
+  #              breaks = c(-10, -5), skip = 0,
+  #              label.placement = label_placement_n(1)) +
+  # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
+  labs(fill = "Precipitable\nwater (Kg/m²)", size = "V (m/s)") +
+  cordillera +
+  geom_mapa() +
+  facet_grid(. ~ exp, 
+             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        # legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  ana[bottom_top %between% c(2, 10), .(theta_mean = mean(theta)), 
+      by = .(south_north, west_east, date, exp)] %>% 
+  .[pw, on = .NATURAL] %>% 
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = v_mean)) +
+  geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
+                    proj = norargentina_lambert,
+                    breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
+  scale_fill_distiller(super = ScaleDiscretised,
+                       palette = "RdYlBu", direction = -1,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  labs(fill = "Potential\ntemperature (K)") +
+  cordillera +
+  geom_mapa() +
+  facet_grid(. ~ exp, 
+             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        # legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3),
+        strip.text = element_blank(), 
+        strip.background = element_blank()) +
+
+  mucape %>% 
+  .[variable == "mcape"] %>% 
+  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+  .[cor, on = .NATURAL] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = value, fill = stat(level_d)), 
+                    proj = norargentina_lambert, limits = c(500, 6000)) +
+  scale_fill_distiller(super = ScaleDiscretised,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  # geom_contour2(aes(z = cortante),
+  #               proj = norargentina_lambert, color = "black", size = 0.4,
+  #               breaks = c(10, 25)) +
+  geom_contour2(aes(z = cortante, size = stat(level)),
+                proj = norargentina_lambert, color = "black", 
+                breaks = c(15, 30)) +
+  scale_size_continuous(breaks = c(15, 30), range = c(0.8, 0.4),
+                        labels = c("15", "30")) +
+  # geom_point(data = ~.x[cortante >= 15 & is.cross(x, y, skip = 1)], 
+  #            aes(x = lon, y = lat, 
+  #                alpha = factor(cut_round(cortante, breaks = c(15, 25, Inf)))),
+  #            size = 0.00002, shape = 20) +
+  # scale_alpha_manual(values = c(0.3, 0.7), name = NULL, #labels = c(">5 m/s", ">15 m/s"),
+  #                    guide = guide_legend(override.aes = list(size = 2))) +
+  labs(fill = "MCAPE (J/Kg)", size = "Wind\nshear (m/s)") +
+  cordillera +
+  geom_mapa() +
+  facet_grid(. ~ exp, 
+             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                         E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        # legend.box.spacing = margin(1, 1, 1, 1),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3),
+        strip.text = element_blank(), 
+        strip.background = element_blank()) +
+
+  plot_layout(ncol = 1, heights = c(1, 1, 1)) +
+  plot_annotation(tag_levels = "a", tag_suffix = ")")
+```
+
+```{r q_vert, fig.cap="Specific humidity (g/kg, shaded) and meridional wind (m/s, contours) average over the inner domain for each latitude for 00 UTC Nov, 22."}
+
+coord <- mucape[variable == "mcape", c("south_north", "lat")] %>% 
+  .[, .(lat = mean(lat)), by = south_north]
+
+ana[west_east %between% c(70, 180) & bottom_top %between% c(1, 25)] %>% 
+  .[, .(q_mean = mean(q),
+        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
+  coord[., on = "south_north"] %>% 
+  ggplot(aes(lat, bottom_top)) +
+  geom_contour_fill(aes(z = q_mean*1000, fill = stat(level_d))) +
+  scale_fill_distiller(super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       # limits = c(1, 19),
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  geom_contour2(aes(z = v_mean, linetype = factor(-sign(..level..))), 
+                color = "grey30", size = 0.3, binwidth = 2) +
+  geom_text_contour(aes(z = v_mean), color = "grey30", 
+                    size = 3, stroke = 0.05, breaks = seq(-10, 6, 2),
+                    label.placement = label_placement_n(1)) +
+  scale_linetype(guide = NULL) +
+  scale_x_latitude(breaks = seq(-40, -20, 5)) +
+  facet_wrap( ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                 E6 = "SATWND", E8 = "RAD"))) +
+  labs(y =  "Sigma levels", fill = "Specific\nhumidity (g/Kg)") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) 
+```

### 2021-04-27 analysis/chapters/paper_ana_20181122.Rmd
Agrego figuras de la sección verificación

@@
 #Color blind friendly but E4 y E7 parecidos en Desaturated
-colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
+colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311") 
 # colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
@@
   facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
+                                               E6 = "SATWND", E8 = "RAD"))) +
   labs(fill = "Mean number of\nobs per cycle") +
@@
                      labels = c(E2 = "CONV", E5 = "AUT", 
-                                E6 = "SATWND", E7 = "RAD")) +
+                                E6 = "SATWND", E8 = "RAD")) +
   scale_shape_manual(values = c(15, 16, 17), labels = c(E2 = "CONV", E5 = "AUT", 
@@
 ```
+
+### Verification
+
+```{r pp-hov, fig.cap="Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right, for the ensemble mean) over the inner domain."}
+
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*mean_acum_1h.rds")
+
+ppacum_1h <- purrr::map(files, function(f) {
+
+  readRDS(f) %>% 
+    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)]
+
+}) %>% rbindlist() %>% 
+  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] 
+}
+
+readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
+  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
+  .[end_date %between% c(as_datetime("20181120190000"), as_datetime("20181123120000"))] %>% 
+  .[, .(pp_acum = mean(pp_acum),
+        lat = mean(lat),
+        exp = "IMERG"), by = .(end_date, y)] %>% 
+  ggplot(aes(end_date, lat)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(1, 13, 1), Inf)) +
+  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 13, 1)) +
+  # scale_fill_viridis_c(option = "A",
+  scale_fill_distiller(palette = "YlGnBu",
+                       direction = 1,
+                       super = ScaleDiscretised,
+                       guide = guide_colorsteps(inside = TRUE,
+                                                barwidth = 25,
+                                                barheight = 0.8)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  scale_y_latitude(breaks = seq(-40, -20, 5)) +
+  facet_wrap(~ exp) +
+  labs(x = NULL, y = NULL, fill = "Acumulated\nprecipitation (mm/h)") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  ppacum_1h %>% 
+  .[exp %in% c("E2", "E5", "E6", "E8")] %>%
+  .[, lat := lat[1], by = .(y)] %>% 
+  .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y, lat)] %>% 
+  ggplot(aes(date, lat)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(1, 13, 1), Inf)) +
+  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 13, 1)) +
+   # scale_fill_viridis_c(option = "A",
+  scale_fill_distiller(palette = "YlGnBu",
+                       direction = 1,
+                       super = ScaleDiscretised,
+                       guide = guide_colorsteps(inside = TRUE,
+                                                barwidth = 25,
+                                                barheight = 0.8)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  scale_y_latitude(breaks = seq(-40, -20, 5), labels = NULL) +
+  facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                E6 = "SATWND", E8 = "RAD"))) +
+  labs(x = NULL, y = NULL, fill = "Acumulated\nprecipitation (mm/h)") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+
+  plot_layout(ncol = 2, widths = c(0.2, 0.8), guides = "collect") & theme(legend.position = 'bottom')
+```
+
+```{r fss, fig.cap="FSS calculated over a 6-hours moving window for different thresholds and scales."}
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2-8].csv")
+
+fss <- purrr::map(files, function(f) {
+
+  fread(f) %>% 
+    .[, date := as_datetime(date)]
+}) %>% 
+  rbindlist()
+
+fss %>% 
+  .[w %in% c(1, 11) & q %in% c(1, 25) & exp %in% c("E2", "E5", "E6", "E8")] %>% 
+  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
+  ggplot(aes(date, fss)) +
+  geom_hline(yintercept = 1, color = "darkgray") +
+  geom_line(aes(color = exp)) +
+  # geom_point(aes(color = exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  scale_date(20181122000000, 20181123000000, "12 hours") +
+  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
+                                            "10" = "10 mm", "25" = "25 mm"),
+                                      w = c("1" = "10 km", "5" = "50 km", "11" = "100 km", 
+                                            "51" = "500 km", "101" = "1000 km"))) +
+  labs(color = NULL, x = NULL, y = "FSS") +
+  theme_minimal()
+```
+
+```{r soundings, fig.cap="RMSE and Bias of temperature, dew point temperature and wind components calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8."}
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos/ANA/sondeo_E[2568]*")
+
+sondeos <- purrr::map(files, function(f){
+
+  fread(f) %>% 
+    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
+    .[, launch_time := as_datetime(launch_time)]
+
+}) %>% 
+  rbindlist()
+
+IOP <- tribble(
+  ~iop, ~ini, ~end,
+  "IOP07", ymd_hms("20181121150000"), ymd_hms("20181121210000"),
+  "IOP08", ymd_hms("20181122140000"), ymd_hms("20181122200000")
+) %>% setDT()
+
+sondeos %>% 
+  .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
+                   launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>% 
+  .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>% 
+  .[site != "Sao Borja, Brazil"] %>% 
+  .[, lev := cut_round(alt, c(seq(0, 3000, 500), seq(4000, 21000, 1000)))] %>% 
+  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
+           BIAS = mean(value - fcst_value, na.rm = TRUE)), by = .(iop, exp, variable, lev)] %>% 
+  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
+  ggplot(aes(lev/1000, value.est)) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  geom_line(aes(color = exp, linetype = estadistico)) +
+  scale_color_manual(values = c(colores_exp, "black"), labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  coord_flip() +
+  facet_grid(iop~variable, scales = "free_x", 
+             labeller = labeller(variable = c("t" = "Temperature",
+                                              "td" = "Dew point\ntemperature",
+                                              "u" = "U wind",
+                                              "v" = "V wind"))) +
+  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+  theme_minimal()
+```

### 2021-04-27 analysis/chapters/paper_ana_20181122.Rmd
Figura de reflectividad

@@
 ```
+
+```{r dbz-mean}
+
+rad_loc <- read_csv(here("analysis", "data", "derived_data", "radar_loc.csv"))
+
+files <- Sys.glob("/home/jruiz/share/DATA/DATOS_RADAR/MOSAICO/*.nc")[seq(1, 12, 3)]
+# Solo un par de horarios 10, 13, 16 y 19 UTC
+
+dbz_obs <- purrr::map(files, function(f) {
+
+  ReadNetCDF(f, vars = "DBZH_cor") %>%
+    .[!is.na(DBZH_cor)] %>% 
+    .[, .(DBZH_cor = max(DBZH_cor, na.rm = TRUE)), by = .(x, y)] %>% 
+    .[, max_dbz := 10 * log10(DBZH_cor)] %>% 
+    .[, ":="(date = ymd_hms(str_remove(basename(f), ".nc")),
+             DBZH_cor = NULL)] %>% 
+    .[, c("lon", "lat") := proj4::project(list(x, y), proj_radar, inverse = TRUE)]
+
+}) %>% rbindlist() %>% 
+  .[, run := "OBSERVATIONS"]
+
+dates <- c("20181122100000", "20181122130000", 
+           "20181122160000", "20181122190000")
+files <- paste0(derived_data, "dbz/maxdbz_ana_E8_", dates, ".nc")
+
+dbz_ana <- purrr::map(files, function(f){
+
+  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
+  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             date = ymd_hms(meta[[1]][["date"]]))] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] 
+}) %>% 
+  rbindlist()  %>% 
+  .[, run := "RAD"]
+
+values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
+            22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
+            42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
+            61.5, 63, 64.5, 66, 69, 76.5)
+
+colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
+             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
+             "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
+             "#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
+             "#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
+             "#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
+             "#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
+             "#95E3C5", "#85E0BD")
+
+dbz_obs %>% 
+  # .[date == unique(date)[1]] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+                    proj = proj_radar,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 30, 
+                                             barheight = 0.8)) +
+  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
+  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
+  geom_mapa() +
+  facet_grid(run ~ date) +
+  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        # legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  dbz_ana %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 30, 
+                                             barheight = 0.8)) +
+    # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
+ geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
+  geom_mapa() +
+  facet_grid(run ~ date) +
+  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  plot_layout(ncol = 1, guides = "collect") & 
+  theme(legend.position = 'bottom')
+
+```

### 2021-04-27 analysis/scripts/radar_loc.R
Localización y rango de los radares.

@@ -1,0 +1,36 @@
+library(tidyverse)
+library(data.table)
+library(mesoda)
+
+
+files <- Sys.glob("/home/jruiz/share/DATA/DATOS_RADAR/MOSAICO/*.nc")[1]
+
+rad_loc <- ReadNetCDF(files, vars = c(lat = "radar_latitude", lon = "radar_longitude")) %>%
+  .[, c("x", "y") := proj4::project(list(lon, lat), proj_radar, inverse = FALSE)]
+
+grilla <- CJ(x = seq(-680000, 990000, length.out = 100), y = seq(-1200000, 1100000, length.out = 100))
+
+
+ranges <- map(rad_loc$nradar, function(n) {
+
+  this_radar <- rad_loc[nradar == n]
+
+  grilla[, ":="(x_radar = this_radar$x, y_radar = this_radar$y)] %>%
+    .[, z := Mag(x - x_radar, y - y_radar)] %>%
+    .[]
+
+  range <- StatContour$compute_group(grilla, breaks = 240000) %>%
+    setDT() %>%
+    .[, c("lon", "lat") := proj4::project(list(x, y), proj_radar, inverse = TRUE)]
+
+
+}) %>% rbindlist(idcol = "ID")
+
+write_csv(ranges, here::here("analysis", "data", "derived_data", "radar_loc.csv"))
+
+
+
+ranges %>%
+  ggplot(aes(lon, lat)) +
+  geom_path(aes(group = ID)) +
+  geom_mapa()

### 2021-04-29 analysis/scripts/RCRV_satwnd.R
Actualizo RCRV para satwnd, ahora u y v van juntos

@@ -1,10 +1,19 @@
-source("help_functions.R")
+library(tidyverse)
+library(data.table)
+library(mesoda)
+library(metR)
+
+source(here::here("temp_R", "help_functions.R"))
 path <- "/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim_conv_2018112*"
 
 perfiles <- read_diag(path, variable = c("uv")) %>%
-  .[, c("error", "nivel.error") := input_obs_error(var, type, pressure)]
+  .[, c("error", "nivel.error") := input_obs_error(var, type, pressure, path_to_errtable = "analysis/data/raw_data/errtable.csv")]
 
-RCRV <- get_RCRV(perfiles[type %between% c(240, 260)], tipo = c("perfil")) %>%
+perfiles <- perfiles[type %between% c(240, 260)] %>%
+  .[, var := "v"] %>%
+  setnames("nivel.error", "error.level")
+
+RCRV <- get_RCRV(perfiles[type %between% c(240, 260)], tipo.var = c("perfil"), calculo = "total") %>%
   .[, sat_type := case_when(
     type == 240 ~ "GOES SW winds",
     type == 241 ~ "India",
@@ -26,15 +35,17 @@ RCRV <- get_RCRV(perfiles[type %between% c(240, 260)], tipo = c("perfil")) %>%
     type == 259 ~ "MODIS WV deep layer winds",
     type == 260 ~ "VIIR IR winds")]
 
+write_csv(na.omit(RCRV), here::here("analysis", "data", "derived_data", "rcrv_satwind.csv"))
+
 RCRV %>%
   melt(measure.vars = c("mean.y", "sd.y")) %>%
-  ggplot(aes(nivel.error, value)) +
+  ggplot(aes(error.level, value)) +
   geom_hline(yintercept = c(0, 1), color = "darkgrey") +
   geom_line(aes(color = factor(sat_type), linetype = variable)) +
   scale_color_brewer(palette = "Set1") +
   scale_x_level() +
   coord_flip() +
-  facet_wrap(~ var, scales = "free_x") +
+  # facet_wrap(~ var, scales = "free_x") +
   labs(title = "Reduced Centered Random Variable",
        subtitle = "Satellite wind",
        linetype = "",

### 2021-04-29 analysis/chapters/paper_ana_20181122.Rmd
Arreglo figura de RCRV profiles

@@
   #             outlier.size = 0.5, size = 0.1) +
-  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
-  geom_line(color = "darkorange") +
+  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4) +
+  geom_line() +
   scale_y_level(name = "Pressure (hPa)") +
@@
   geom_hline(yintercept = c(0, 1), color = "grey20") +
-  geom_line(aes(color = factor(code), linetype = variable)) +
+  geom_line(aes(linetype = variable, color = code)) +
   scale_linetype_manual(values = c("dashed", "solid"),
                         labels = c("Mean RCRV", "SD RCRV")) +
+  scale_color_brewer(palette = "Dark2") +
+  guides(linetype = guide_legend(order = 1),
+        color = guide_legend(order = 10)) +
+  # scale_color_discrete() +
   scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
@@
   tag_facets() +
-  labs(y = NULL,
-       linetype = "",
-       color = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom", 
-        legend.box = "vertical",
-        aspect.ratio = 2/1)
+  labs(x = NULL, y = NULL,
+      linetype = NULL, color = "Upper air\nobservations") +
+  theme_minimal() +
+  theme(legend.box = "vertical",
+    aspect.ratio = 4/1) +
 
@@
   melt(measure.vars = c("mean.y", "sd.y")) %>%
-  ggplot(aes(nivel.error, value)) +
-  geom_hline(yintercept = c(0, 1), color = "darkgrey") +
-  geom_line(aes(color = factor(sat_type), linetype = variable)) +
-  scale_color_brewer(palette = "Set1") +
+  ggplot(aes(error.level, value)) +
+  geom_hline(yintercept = c(0, 1), color = "grey20") +
+  geom_line(aes(color = sat_type, linetype = variable)) +
   scale_linetype_manual(values = c("dashed", "solid"),
                         labels = c("Mean RCRV", "SD RCRV")) +
+  scale_color_brewer(palette = "Set1") +
   scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
@@
   facet_wrap(~ var, scales = "free_x", labeller = labeller(var = c("u" = "U wind",
-                                                                   "v" = "V wind"))) +
-  labs(linetype = NULL,
-       color = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom", 
-        legend.box = "vertical",
-        aspect.ratio = 2/1)
-
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:6]
-
-RCRV <- purrr::map(files, function(f) {
+                                                                   "v" = "Wind"))) +
+  tag_facets(tag_pool = c("c")) +
+  labs(x = NULL, y = NULL,
+      linetype = NULL, color = "Satellite-derived\nwind") +
+  theme_minimal() +
+  theme(legend.box = "vertical",
+    aspect.ratio = 2/1) +
+
+  purrr::map(Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:6], function(f) {
     meta <- unglue::unglue(basename(f), "rcrv_{sensor}_perfil.csv")
-
         fread(f) %>%
@@
   }) %>%
-  rbindlist()
-
-RCRV %>% 
+  rbindlist() %>% 
   separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
@@
   melt(measure.vars = c("mean.y", "sd.y")) %>% 
-  .[, .(value = mean(value, na.rm = TRUE)), by = .(sensor, level, variable)] %>% 
+  .[, .(value = mean(value, na.rm = TRUE),
+        var = "Brightness\ntemperature"), by = .(sensor, level, variable)] %>% 
   ggplot(aes(level, value)) +
@@
   geom_line(aes(color = sensor, linetype = variable)) +
-  scale_color_discrete(labels = toupper) +
   scale_linetype_manual(values = c("dashed", "solid"),
                         labels = c("Mean RCRV", "SD RCRV")) +
+  scale_color_discrete(labels = toupper) +
   coord_flip() +
   scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
+  facet_wrap(~ var, scales = "free_x") +
+  tag_facets(tag_pool = c("d")) +
   labs(x = NULL, y = NULL,
-       color = NULL, linetype = NULL) +
-  theme_minimal() +
-  theme(legend.position = "bottom", 
+      linetype = NULL, color = "Radiances\nobservations") +
+  theme_minimal() +
+  theme(#legend.position = "bottom", 
     legend.box = "vertical",
-        aspect.ratio = 2/1)
+    aspect.ratio = 2/1) +
+
+  plot_layout(ncol = 3, widths = c(1, 1, 1), guides = "collect") & theme(legend.box = "vertical")
 
@@
                     proj = norargentina_lambert) +
-  scale_fill_continuous_divergingx(super = ScaleDiscretised,
+  colorspace::scale_fill_continuous_divergingx(super = ScaleDiscretised,
                                                palette = "PRGn", mid = 25,

### 2021-04-30 analysis/chapters/paper_ana_20181122.Rmd
Agrego dbz de CONV

@@
 ---
-title: "Untitled"
+title: Assimilation of conventional and satellite observations in a deep convection
+  case during RELAMPAGO using the WRF-GSI-LETFK system.
+author:
+- name: Paola Belén Corrales
+  email: paola.corrales@cima.fcen.uba.ar
+  affiliation: Some Institute of Technology
+  footnote: 1
+- name: Bob Security
+  email: bob@example.com
+  affiliation: Another University
+- name: Cat Memes
+  email: cat@example.com
+  affiliation: Another University
+  footnote: 2
+- name: Derek Zoolander
+  email: derek@example.com
+  affiliation: Some Institute of Technology
+  footnote: 2
+date: "`r Sys.Date()`"
 output:
-  html_document: default
-  pdf_document: default
+  rticles::elsevier_article: 
+    keep_tex: true
+abstract: |
+  This is the abstract.
+  It consists of two paragraphs.
+journal: Atmospheric Research
+layout: preprint
+address:
+- code: Some Institute of Technology
+  address: Department, Street, City, State, Zip
+- code: Another University
+  address: Department, Street, City, State, Zip
+bibliography: mybibfile.bib
+csl: elsevier-harvard.csl
+footnote:
+- code: 1
+  text: Corresponding Author
+- code: 2
+  text: Equal contribution
 ---
@@
 ```{r ciclo, fig.cap="Diagram of analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
-The zoomed figure shows assimilation in one-hour centered window and new boundary conditions from GFSD every 6 hours.", out.height=400, out.width=800, fig.align="center"}
-knitr::include_graphics(here("analysis", "figures", "analysis_cycle.svg"))
+The zoomed figure shows assimilation in one-hour centered window and new boundary conditions from GFSD every 6 hours."}
+knitr::include_graphics(here("analysis", "figures", "analysis_cycle.pdf"))
 ```
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/*ana*mean_acum_1h.rds")
+files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/E[2568]_ana*mean_acum_1h.rds")
 
@@
   .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)]
-}
 
@@
   theme_minimal()
+```
+
+```{r dbz-mean, fig.cap="Figure 14. Observed maximum reflectivity in the column (upper row) and ensemble mean maximum reflectivity for CONV and RAD at different times (bottom two rows)."}
+proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"
+
+rad_loc <- read_csv(here("analysis", "data", "derived_data", "radar_loc.csv"))
+
+files <- Sys.glob("/home/jruiz/share/DATA/DATOS_RADAR/MOSAICO/*.nc")[seq(1, 12, 3)]
+# Solo un par de horarios 10, 13, 16 y 19 UTC
+
+dbz_obs <- purrr::map(files, function(f) {
+
+  ReadNetCDF(f, vars = "DBZH_cor") %>%
+    .[!is.na(DBZH_cor)] %>% 
+    .[, .(DBZH_cor = max(DBZH_cor, na.rm = TRUE)), by = .(x, y)] %>% 
+    .[, max_dbz := 10 * log10(DBZH_cor)] %>% 
+    .[, ":="(date = ymd_hms(str_remove(basename(f), ".nc")),
+             DBZH_cor = NULL)] %>% 
+    .[, c("lon", "lat") := proj4::project(list(x, y), proj_radar, inverse = TRUE)]
+
+}) %>% rbindlist() %>% 
+  .[, run := "OBSERVATIONS"]
+
+dates <- c("20181122100000", "20181122130000", 
+           "20181122160000", "20181122190000")
+files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[28]_", dates, ".nc"))
+
+dbz_ana <- purrr::map(files, function(f){
+
+  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
+  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             date = ymd_hms(meta[[1]][["date"]]))] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] 
+}) %>% 
+  rbindlist()  %>% 
+  .[, run := fifelse(exp == "E8", "RAD", "CONV")]
+
+values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
+            22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
+            42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
+            61.5, 63, 64.5, 66, 69, 76.5)
+
+colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
+             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
+             "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
+             "#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
+             "#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
+             "#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
+             "#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
+             "#95E3C5", "#85E0BD")
+
+dbz_obs %>% 
+  # .[date == unique(date)[1]] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+                    proj = proj_radar,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 30, 
+                                             barheight = 0.8)) +
+  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
+  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
+  geom_mapa() +
+  facet_grid(run ~ date) +
+  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        # legend.box.spacing = margin(0, 0, 0, 0),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  dbz_ana %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 30, 
+                                             barheight = 0.8)) +
+  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
+  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
+  geom_mapa() +
+  facet_grid(run ~ date) +
+  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  plot_layout(ncol = 1, heights = c(1, 2), guides = "collect") & 
+  theme(legend.position = 'bottom')
+
 ```
@@
 ```
-
-```{r dbz-mean}
-
-rad_loc <- read_csv(here("analysis", "data", "derived_data", "radar_loc.csv"))
-
-files <- Sys.glob("/home/jruiz/share/DATA/DATOS_RADAR/MOSAICO/*.nc")[seq(1, 12, 3)]
-# Solo un par de horarios 10, 13, 16 y 19 UTC
-
-dbz_obs <- purrr::map(files, function(f) {
-
-  ReadNetCDF(f, vars = "DBZH_cor") %>%
-    .[!is.na(DBZH_cor)] %>% 
-    .[, .(DBZH_cor = max(DBZH_cor, na.rm = TRUE)), by = .(x, y)] %>% 
-    .[, max_dbz := 10 * log10(DBZH_cor)] %>% 
-    .[, ":="(date = ymd_hms(str_remove(basename(f), ".nc")),
-             DBZH_cor = NULL)] %>% 
-    .[, c("lon", "lat") := proj4::project(list(x, y), proj_radar, inverse = TRUE)]
-
-}) %>% rbindlist() %>% 
-  .[, run := "OBSERVATIONS"]
-
-dates <- c("20181122100000", "20181122130000", 
-           "20181122160000", "20181122190000")
-files <- paste0(derived_data, "dbz/maxdbz_ana_E8_", dates, ".nc")
-
-dbz_ana <- purrr::map(files, function(f){
-
-  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
-  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
-    .[, ":="(exp = meta[[1]][["exp"]],
-             date = ymd_hms(meta[[1]][["date"]]))] %>% 
-    .[, c("x", "y") := wrf_project(lon, lat)] 
-}) %>% 
-  rbindlist()  %>% 
-  .[, run := "RAD"]
-
-values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
-            22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
-            42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
-            61.5, 63, 64.5, 66, 69, 76.5)
-
-colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
-             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
-             "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
-             "#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
-             "#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
-             "#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
-             "#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
-             "#95E3C5", "#85E0BD")
-
-dbz_obs %>% 
-  # .[date == unique(date)[1]] %>% 
-  .[, date := factor(date)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
-                    proj = proj_radar,
-                    breaks = values) +
-  scale_fill_manual(values = colours, 
-                    labels = function(x) JumpBy(x, 3, fill = ""),
-                    drop = FALSE, 
-                    guide = guide_colorsteps(barwidth = 30, 
-                                             barheight = 0.8)) +
-  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
-  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
-  geom_mapa() +
-  facet_grid(run ~ date) +
-  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        # legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3)) +
-
-  dbz_ana %>% 
-  .[, date := factor(date)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
-                    proj = norargentina_lambert,
-                    breaks = values) +
-  scale_fill_manual(values = colours, 
-                    labels = function(x) JumpBy(x, 3, fill = ""),
-                    drop = FALSE, 
-                    guide = guide_colorsteps(barwidth = 30, 
-                                             barheight = 0.8)) +
-  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
-  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
-  geom_mapa() +
-  facet_grid(run ~ date) +
-  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3)) +
-
-  plot_layout(ncol = 1, guides = "collect") & 
-  theme(legend.position = 'bottom')
-
-```

### 2021-04-30 analysis/chapters/experimento_20181122.Rmd
Sacar el bordecito del dominio no afecta el FSS

@@
            "20181122160000", "20181122190000")
-files <- paste0(derived_data, "dbz/maxdbz_ana_E8_", dates, ".nc")
+files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[2568]_", dates, ".nc"))
 
@@
              date = ymd_hms(meta[[1]][["date"]]))] %>% 
-    .[, c("x", "y") := wrf_project(lon, lat)] 
+    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+    .[, run := meta[[1]][["exp"]]]
 }) %>% 
-  rbindlist()  %>% 
-  .[, run := "RAD"]
+  rbindlist()
 
@@
                                              barheight = 0.8)) +
+  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
   geom_mapa() +
@@
                                              barheight = 0.8)) +
+  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
 
@@
 ```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-8].csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2-8].csv")
 
@@
   # geom_point(aes(color = exp)) +
-  scale_color_manual(values = c(colores_exp, "black"), labels = c(E4 = "CONV", E5 = "AUT", 
+  scale_color_manual(values = c("purple", colores_exp, "black"), labels = c(E4 = "CONV", E5 = "AUT", 
                                                       E6 = "SATWND", E7 = "RAD", E8 = "NEW RAD")) +

### 2021-05-28 R/rh.R
Agrego funciones de postprocesamiento

@@ -1,0 +1,20 @@
+#' Humedad relativa
+#'
+#' @param QVAPOR vector con la humedad especifica en kg/kg
+#' @param P vector con la presión en Pa
+#' @param T vector con la temperatura en kelvin
+#'
+#' @export
+# https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L730
+
+rh <- function(QVAPOR, P, T) {
+  P <- P*0.01      # Debe estar en hPa
+  T <- T - 273.15  # Debe estar en Celsius
+  es <- 6.112 * exp(17.67*(T)/(T + 273.15 - 29.65))
+
+  qvs <- es/(P - (1 - 0.622)*es)
+
+  rh <- 100*pmax(pmin(QVAPOR/qvs, 1), 0)
+
+  return(rh)
+}

### 2021-05-28 R/scale_color_dbz.R

@@ -1,64 +1,0 @@
-#' Reflectivity color pallete
-#'
-#' @param colour
-#' @param limits
-#' @param guide
-#' @param ...
-#'
-#' @export
-scale_color_dbz <- function(colours = ref,
-                              limits = c(15, 75),
-
-                              guide = guide_colorbar(barheight = 15),
-                              ...) {
-
-  # colours <- rgb(ref$R, ref$G, ref$B, maxColorValue = 1)
-
-  scale_color_gradientn(colours = ref,
-                        limits = limits,
-                        guide = guide, ...)
-}
-
-ref <- c(
-#   "#646464", # ND
-   # "#ccffff", # -30
-   # "#cc99cc", # -25
-   # "#996699", # -20
-   # "#663366", # -15
-   # "#cccc99", # -10
-   # "#999966", # -5
-   # "#646464", # 0
-   # "#04e9e7", # 5
-   # "#019ff4", # 10
-   "#0300f4", # 15
-   "#02fd02", # 20
-   "#01c501", # 25
-   "#008e00", # 30
-   "#fdf802", # 35
-   "#e5bc00", # 40
-   "#fd9500", # 45
-   "#fd0000", # 50
-   "#d40000", # 55
-   "#bc0000", # 60
-   "#f800fd", # 65
-   "#9854c6", # 70
-   "#fdfdfd" # 75
-)
-
-
-ref2 <- dplyr::tribble(
-  ~R, ~G, ~B,
-   1.0/255.0, 159.0/255.0, 244.0/255.0,
-   3.0/255.0,   0.0/255.0, 244.0/255.0,
-   2.0/255.0, 253.0/255.0,   2.0/255.0,
-   1.0/255.0, 197.0/255.0,   1.0/255.0,
-   0.0/255.0, 142.0/255.0,   0.0/255.0,
- 253.0/255.0, 248.0/255.0,   2.0/255.0,
- 229.0/255.0, 188.0/255.0,   0.0/255.0,
- 253.0/255.0, 149.0/255.0,   0.0/255.0,
- 253.0/255.0,   0.0/255.0,   0.0/255.0,
-# 212.0/255.0,   0.0/255.0,   0.0/255.0,
- 188.0/255.0,   0.0/255.0,   0.0/255.0,
- 248.0/255.0,   0.0/255.0, 253.0/255.0,
- 152.0/255.0,  84.0/255.0, 198.0/255.0
-)

### 2021-05-28 R/td.R

@@ -1,0 +1,24 @@
+#' Dew Point
+#'
+#' @param QVAPOR vector con la humedad especifica en kg/kg
+#' @param P vector con la presión en Pa
+#'
+#'  @export
+td <- function(QVAPOR, P) {
+
+  # qv en kg/kg
+  # p en pascales
+  P <- P*0.01
+
+  QVAPOR <- ifelse(QVAPOR < 0, 0, QVAPOR)
+
+  tdc <- QVAPOR*P / (0.622 + QVAPOR)
+
+  tdc <- ifelse(tdc < 0.001, 0.001, tdc)
+
+  td <- (243.5*log(tdc) - 440.8) / (19.48 - log(tdc))
+
+  return(td)
+}
+
+# https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L970

### 2021-05-28 R/tk.R

@@ -1,0 +1,12 @@
+#' Temperatura
+#'
+#' @param T vector con la temperatura potencial en kelvin
+#' @param P vector con la presión en Pa
+#' @param T_BASE valor numérico con la temperatura de base del WRF, usualmente 300 K
+#'
+#' @export
+# https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L57
+
+tk <- function(T, P, T_BASE = 300){
+  tk <- (T + T_BASE)*((P/100000)^(2/7))
+}

### 2021-05-28 R/uvmet.R

@@ -1,0 +1,61 @@
+#' Rota los vientos e interpola a la retícula de masa
+#'
+#' La función calcula la rotación correcta e interpola a la reticula de masa.
+#' Solo funciona para Labert
+#'
+#' @param ncfile path al archivo netCDF donde están los datos
+#'
+#' @export
+uvmet <- function(ncfile) {
+
+  nc <- ncdf4::nc_open(ncfile)
+
+  # https://github.com/NCAR/wrf-python/blob/b40d1d6e2d4aea3dd2dda03aae18e268b1e9291e/src/wrf/g_uvmet.py#L179
+  true_lat1 <- ncdf4::ncatt_get(nc, 0, "TRUELAT1")[[2]]
+  true_lat2 <- ncdf4::ncatt_get(nc, 0, "TRUELAT2")[[2]]
+  cen_lon <- ncdf4::ncatt_get(nc, 0, "STAND_LON")[[2]]
+
+  lon <- ncdf4::ncvar_get(nc, "XLONG")
+  lat <- ncdf4::ncvar_get(nc, "XLAT")
+
+  u <- ncdf4::ncvar_get(nc, "U")
+  v <- ncdf4::ncvar_get(nc, "V")
+
+  rpd <- pi/180.0
+
+  if ((abs(true_lat1 - true_lat2) > 0.1) & (abs(true_lat2 - 90.) > 0.1)) {
+    cone = (log(cos(true_lat1*rpd)) -
+              log(cos(true_lat2*rpd)))
+    cone = (cone /
+              (log(tan((45.- abs(true_lat1/2.))*rpd))
+               - log(tan((45.- abs(true_lat2/2.)) *
+                           rpd))))
+  } else {
+    cone = sin(abs(true_lat1)*rpd)
+  }
+
+  # https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L801
+
+  longca <- lon - cen_lon
+
+  longca[longca > 180] <- longca[longca > 180] - 360
+  longca[longca < -180] <- longca[longca < -180] + 360
+
+  longcb <- longca*cone*rpd            # Hemisferio norte
+  longcb[lat < 0] <- -longcb[lat < 0]  # Corrección si el dominio está en el HS
+
+  longca <- cos(longcb)
+  longcb <- sin(longcb)
+
+  # destagger
[…]

### 2021-05-28 R/v_circulation.R

@@ -1,0 +1,95 @@
+library(mesoda)
+library(metR)
+library(tidyverse)
+library(lubridate)
+library(data.table)
+library(here)
+library(patchwork)
+library(unglue)
+library(knitr)
+library(kableExtra)
+library(tagger)
+
+levs <- c(200, 300, 500, 600)
+
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E[2,5,6,8]/ANA/20181122180000/analysis.ensmean")
+
+coord <- ReadNetCDF(files[1], vars = c(lon = "XLONG", lat = "XLAT"), subset = list(bottom_top = 1))
+
+ana <- purrr::map(files, function(f) {
+  descriptores <- unglue::unglue(f, c("/home/paola.corrales/datosmunin3/EXP/{exp}/ANA/{date}/analysis.ensmean"))
+
+  # Leo pronóstico con algo de post procesamiento
+  fcst <- ReadNetCDF(f, vars = c(p = "P", "PB", t = "T", qv = "QVAPOR",
+                                        lon = "XLONG", lat = "XLAT")) %>%
+    .[, p := p + PB] %>%
+    .[, t := (t + 300)*(p/100000)^(2/7)] %>%
+    # .[, td := td(qv, p) + 273.15] %>%
+    .[, ":="(Time = NULL,
+             PB = NULL)] %>%
+    .[, c("u", "v") := uvmet(f)] %>%
+    .[, ":="(date = ymd_hms(descriptores[[1]][["date"]]),
+             exp = descriptores[[1]][["exp"]])] %>%
+    .[, .(lev = levs,
+          t = approx(p, t, xout = levs*100)$y,
+          qv = approx(t, qv, xout = levs*100)$y,
+          u = approx(p, u, xout = levs*100)$y,
+          v = approx(p, v, xout = levs*100)$y),
+      by = .(date, south_north, west_east, exp)] %>%
+    coord[., on = .NATURAL]
+
+}) %>%
+  rbindlist()
+
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/dbz/maxdbz_ana_E[2,5,6,8]_20181122180000.nc")
+
+dbz <- purrr::map(files,  function(f) {
+
+  exp <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/derived_data/dbz/maxdbz_ana_{exp}_20181122180000.nc")
[…]

### 2021-05-28 temp_R/help_functions.R

@@ -10,7 +10,7 @@ read_diag <- function(path, variable) {
 
   files <- Sys.glob(paste0(path, ".mem*"))
 
-  myCluster <- makeCluster(20)
+  myCluster <- makeCluster(10)
   registerDoParallel(myCluster)
 
   obs <- foreach(f = 1:length(files),
@@ -368,7 +368,7 @@ cut_round <- function(x, breaks) {
 library(data.table)
 library(dplyr)
 
-read_radiosonde_relampago <- function(file){
+read_radiosondeo_relampago <- function(file){
   # Leo línea por línea
   lines <- readLines(file)
 

### 2021-05-28 temp_R/postprocesamiento.R

@@ -1,102 +1,105 @@
 
 # Post procesamiento WRF --------------------------------------------------
 
-## Humedad relativa
-# https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L730
+# ## Humedad relativa
+# # https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L730
+#
+# rh <- function(QVAPOR, P, T) {
+#   P <- P*0.01      # Debe estar en hPa
+#   T <- T - 273.15  # Debe estar en Celsius
+#   es <- 6.112 * exp(17.67*(T)/(T + 273.15 - 29.65))
+#
+#   qvs <- es/(P - (1 - 0.622)*es)
+#
+#   rh <- 100*pmax(pmin(QVAPOR/qvs, 1), 0)
+#
+#   return(rh)
+# }
 
-rh <- function(QVAPOR, P, T) {
-  P <- P*0.01      # Debe estar en hPa
-  T <- T - 273.15  # Debe estar en Celsius
-  es <- 6.112 * exp(17.67*(T)/(T + 273.15 - 29.65))
+# ## Dew Point
+# # https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L970
+#
+# td <- function(QVAPOR, P) {
+#
+# # qv en kg/kg
+# # p en pascales
+#   P <- P*0.01
+#
+#   QVAPOR <- ifelse(QVAPOR < 0, 0, QVAPOR)
+#
+#   tdc <- QVAPOR*P / (0.622 + QVAPOR)
+#
+#   tdc <- ifelse(tdc < 0.001, 0.001, tdc)
+#
+#   td <- (243.5*log(tdc) - 440.8) / (19.48 - log(tdc))
+#
+#   return(td)
+# }
 
-  qvs <- es/(P - (1 - 0.622)*es)
+# ## T
+# # https://github.com/NCAR/wrf-python/blob/d9585354c0e2a75a0f7c1d6b200d353f5e4eb084/fortran/wrf_user.f90#L57
+#
+# tk <- function(T, P, T_BASE){
+#   tk <- (T + T_BASE)*((P/100000)^(2/7))
+# }
 
[…]

### 2021-05-28 analysis/scripts/altura_estaciones.R
Analisis de la altura de las estaciones

@@ -1,0 +1,59 @@
+library(mesoda)
+library(metR)
+library(tidyverse)
+library(lubridate)
+library(data.table)
+library(here)
+
+dominio <- fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>%
+  .[, c("x", "y") := wrf_project(lon, lat)]
+
+
+oficiales <- fread("/home/paola.corrales/E2_asim_conv_20181121120000.ensmean",
+                   na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00", "0.100E+12")) %>%
+  .[, c("V2", "V4") := NULL] %>%
+  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "alt", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>%
+  .[type %in% c(181, 281)] %>% unique(by = c("stationID")) %>%
+  .[!str_detect(stationID, pattern = "[A-Z]")] %>%
+  .[, source := "Sfc - Official"]
+
+no_oficiales <- fread("/home/paola.corrales/E5_asim_conv_20181121120000.ensmean",
+                      na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00", "0.100E+12")) %>%
+  .[, c("V2", "V4") := NULL] %>%
+  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "alt", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>%
+  .[type %in% c(181, 187)] %>%
+  unique(by = c("stationID")) %>%
+  .[!str_detect(stationID, "SMN")] %>%
+  .[!(stationID %in% oficiales$stationID)] %>%
+  .[, source := "Sfc - Non-official"]
+
+obs <- rbind(no_oficiales, oficiales) %>%
+  .[, c("x", "y") := wrf_project(ConvertLongitude(lon), lat)]
+
+
+alturas <- dominio %>%
+  .[, Interpolate(hgt ~ x + y, x.out = obs$x, y.out = obs$y, grid = FALSE)] %>%
+  .[obs, on = .NATURAL] %>%
+  .[, diff := alt - hgt] %>%
+  .[, diff_b := fcase(abs(diff) %between% c(0, 100), "[0, 100]",
+                      abs(diff) %between% c(100, 300), "[100, 300]",
[…]

### 2021-05-28 analysis/chapters/paper_ana_20181122.Rmd
Tiene forma?

@@
 journal: Atmospheric Research
-layout: preprint
+layout: 5p
 address:
@@
   address: Department, Street, City, State, Zip
-bibliography: mybibfile.bib
+bibliography: paper_ANA.bib
 csl: elsevier-harvard.csl
@@
   setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
-derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"
-
-```
-
-## Introduction
-
-## Data and Methods
-
-```{r dominio, fig.cap="Figure 1.
-a) The horizontal domain used on the simulations (black box) and the inner domain used for the experiment comparison (red box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles) .
-b) Locations of radiosonde launches during the experiment.
-The topography in meters is also shown (filled contours).", fig.height=5}
+derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
+
+```
+
+# Introduction
+
+Severe weather events cause significant human and material losses around the world.
+A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
+The north and central region of Argentina has been shown to have the highest frequency of favorable conditions for high-impact meteorological events [@brooks2003] and the highest occurrence of large hail events [@cecil2012].
+This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014].
+
+Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to several factors.
+On the one hand, mesoscale phenomena have limited predictability, while on the other hand, there is great difficulty in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and in the order of minutes).
+To solve this problem, data assimilation (DA) at the mesoscale is a topic that has gained increasing interest since it can provide appropriate initial conditions for the initialization of high-resolution forecasts [@sun2014].
+
+DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations.
+However surface observations, which directly measure the atmospheric state, have limited availability in South America.
+Other types of observations with relatively high spatial and temporal resolution, such as satellite radiance observations, measure only indirect quantities and are more difficult to incorporate to DA systems.
+The assimilation of conventional observations from surface stations have an important role at improving the initial conditions that lead to better forecasts.
+For example, @ha2014 showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
+[Agregar algún ejemplo local que no sea el RRA]
+
+Given the sparsity of the local conventional observing network it is important to explore the potential benefits of including additional sources of observations at a regional scale.
+Recently Zhu et al studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
+Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the AIRS sensor in a regional application (cita: RRA, otros de Maru?).
+Moreover Bao et al studied the impact on temperature and humidity forecast over western USA while assimilating directly microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels due the microwave observations but an opposite impact for infrared data.
+Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al.
+2016).
+
+Having analyses with higher temporal resolution than those available from global forecasting systems (e.g.
+at hourly or sub-hourly) and using an increased number of regional observations that are not included in global models are some of the advantages of assimilation at these scales.
+Previous work has shown promising results on mesoscale DA in the region (goncalvesdegoncalves2015, Dillon 2019, Dillon 2020 submitted) using different approaches such as LETKF; (Hunt et al, 2007) and 3D-Var ().
+In particular, the use of ensemble-based methods have shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances.
+Moreover, the use of ensemble models also provides valuable information about the uncertainties in the forecast.
+Since 2016, the Argentinian National Meteorological Service (ANMS) has been running a convective-scale downscaling forecast (skabar2018).
+Recently a rapid-refresh ensemble-based DA system has been experimentally implemented during the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) Field Campaign [central Argentina, November-December 2018, Nesbitt 2021] with promising results [Dillon 2020 submitted].
+This system assimilated data from high-resolution surface networks, AMDAR data from local aircraft flights, soundings, AIRS retrievals, high-resolution GOES-16 wind estimates, and local radar data every hour with grid spacing of 10 km.
+The forecast initialized from the analyses shows an overall similar performance to cold-start initializations from the Global Forecast System analysis, and even a positive impact in some cases.
+
+One important challenge of the DA method is to assess the performance of a regional data assimilation system at representing the true atmospheric state and the skill of the forecasts initialized using the analysis as initial condition.
+In this sense, the observations and measurements taken during the intense observing period of RELAMPAGO are an invaluable resource to validate a regional data assimilation system.
+
+The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
+The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during 22-23 November 2018 during the IOP period of the RELAMPAGO field campaign.
+On 22 November a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
+The MCS lasts for 42 hours traveling approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 18 UTC 23 November.
+To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged the MSC.
+The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system.
+A multi-model ensemble is used in the data assimilation experiments allowing an evaluation of the performance of different model configurations in this particular case.
+
+The rest of the paper is organized as follows: a description of the assimilation systems and assimilated observations is provided in Section XX, results are discussed in …. …..
+
+# Data and Methods
+
+The model simulations for the case of study were performed using the version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al., n.d.) model.
+The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
+The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25° resolution) [] every 6 hours.
+The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulation period.
+
+The Local Ensemble Transform Kalman Filter (LETKF; ) implementation (V1.3) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Hu et al.
+2018) is chosen to generate the analysis.
+A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within +- 30 minutes of the analysis time are assimilated.
+Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
+For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al.
+2006) was used as an observation operator to calculate model-simulated brightness temperatures.
+
+To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localizations radius of  0.4 (in log pressure coordinates) as in [RRA] for all types of observations.
+A relaxation-to-prior inflation [Whitaker and Hamill] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
+
+We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis and random perturbations are used to generate the initial ensemble perturbations.
+The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) with a smooth time evolution as in \citep{Necker2020,Maldonado2021}.
+In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
+These random perturbations are also applied at the boundaries in order to keep the ensemble spread within the domain.
+
+In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system.
+We use 9 different model configuration consisting on the combinations of 3 moist convection schemes  (Kain–Fritsch [], Grell–Freitas [], and Betts–Miller–Janjic []) and 3 planetary boundary layer schemes (Yonsei University Scheme [], Mellor–Yamada–Janjic Scheme [], and Mellor–Yamada Nakanishi Niino []).
+All ensemble members used the same land-surface model (Noah-MP, ) and microphysics (WRF single-moment 6–class squeme []), and radiation processes (RRTMG shortwave and longwave scheme []) parameterizations.
+
+```{r dominio, fig.cap="a) The horizontal domain used on the simulations (black box) and the inner domain used for the experiment comparison (red box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles) .
+b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP07, orange triangles are radiosondes launched during IOP08 and purple square are radiosondes launched outside the IOPs.
+The topography in meters is also shown (filled contours).", out.width="110%"}
 
@@
                       guide = guide_colorstrip(barwidth = 20,
-                                               barheight = 0.8)) +
+                                               barheight = 0.5)) +
   geom_mapa() +
@@
   geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
-  theme_minimal() +
+  theme_minimal(base_size = 9) +
   theme(legend.position = "bottom", 
@@
   labs(color = "b) Soundings", shape = "b) Soundings") +
-  theme_minimal() +
+  theme_minimal(base_size = 9) +
   theme(legend.position = "bottom", 
@@
   #         legend.box.spacing = margin(0, 0, 0, 0)) +
-  plot_layout(ncol = 2, guides = "collect") +
+  plot_layout(ncol = 2, widths = c(4, 1), guides = "collect") +
   plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
@@
 
-```{r ciclo, fig.cap="Diagram of analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
-The zoomed figure shows assimilation in one-hour centered window and new boundary conditions from GFSD every 6 hours."}
-knitr::include_graphics(here("analysis", "figures", "analysis_cycle.pdf"))
+### Observations
+
+#### Conventional and Satellite-derived wind
+
+Conventional and satellite-derived wind observations are part of the Global Data Assimilation System (GDAS) [] except for Automatic Surface Weather Stations (ASWS).
+These observations come from public and private automatic station networks over Southern South America and are part of RELAMPAGO Data Set repository (García, F., et al.
+2019).
+
+We used conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP (Keyser 2013) that consist of surface observations (land and sea) and upper-air observations from radiosondes and aircrafts (Figure \@ref(fig:dominio)).
+The frequency of these observations varied between 1 hour for surface stations and 24 hours for most radiosondes in the region.
+
+Satellite-derived wind observations also come from GDAS and includes information from GOES-16 (estimated from the visible, infrared and water vapour channels) and METEOSAT 8 and 11 (visible and water vapour channels).
+Due to the domain covered by each of these satellites, GOES-16 is the primary source of this type of observation (99 % of this type of observation).
+
+Table 2 lists all the observation types (i.e., pressure, temperature, specific humidity and wind) available for each source, together with their associated errors.
+The observation errors were specified following the GSI default configuration.
+In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind).
+In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and model-simulated observation based on the first-guess)  with a predefined threshold (also included in Table 2).
+
+#### Satellite radiances
+
+Satellite radiances available through the GDAS datastream which consist of infrared and microwave observations are used in this paper.
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 3).
+Since the regional domain is located in mid-latitudes, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
+For this reason, radiance observations were not available in all analysis cycles and the number of observations varied significantly among cycles.
+This is evident in Figure 2 b) that shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
+In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in the period.
+The vertical location of each radiance was estimated as the model level at which its weighting function was maximized as calculated by the CRTM.
+The multispectral sensors have a good vertical distribution and are able to sense throughout the entire atmosphere.
+The channels whose maximum weighting function is in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.
+
+The channels accepted for assimilation and their associated error were defined using the default GSI configuration.
+The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
+Firstly a spatial data thinning is applied using a 60 km grid following [Singh et al 2016, Jones et al 2013, Lin et al 2017], where the observations to be assimilated are chosen based on their distance to the model grid points, to the observation quality (based on available data quality predictors) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
+Also, observations over the sea or sea-ice surface are preferred from those over land or snow [cita a la guía avanzada].
+
+The thinned observations were then bias corrected.
+The bias correction (BC) has an air-mass dependent and an angle-dependent component (Zhu 2014) and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$.
+Then, the bias corrected brightness temperature can be obtained as:
+
+$$\mathit{BT_bc}  =\mathit{ BT} +  \sum{i = 0}{N} \beta_i p_i (x)$$
+
+In GSI $p_0 = 1$ a constant offset bias correction term.
+The following predictors are:  the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight , and the emissivity sensitivity.
+Scan angle dependent bias is modeled as a 4th-order polynomial [Zhu 2014].
+
+The $\beta_i$ coefficients are usually trained to achieve the best fit between the simulation and the observations.
+In the GSI system these coefficients are trained using a variational estimation method (Zhu 2014) which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
+The coefficients initialized from the 18 UTC Nov, 11 2018 GFS coefficients were trained during one week of continuous assimilation and using the same setup as in the experiments described in the next section.
+Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments to the predictor coefficients during each cycle.
+
+During the quality control step, GSI rejects all observations if the innovation is greater than a predefined threshold for each channel and sensor.
+For microwave sensors, the quality control routines also use the scattering and Liquid Water Path (LWP) indexes calculated from brightness temperature channel differences (Weston et al., 2019; Zhu et al., 2016).
+The quality control routine for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
+It also checks for cloud-affected observations using the transmittance profile calculated within the CRTM algorithms and performs an emissivity check for observations over land.
+In the experiment, only clear-sky observations are used as defined by the internal subroutines of GSI that use the LWP for microwave and the transmittance profile for infrared observations that could be potentially affected by clouds.
+
+```{r}
+files <- Sys.glob(here("analysis/data/derived_data/count_obs_E*"))[c(1, 3, 4)]
+
+purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
+  .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
+  dcast(bufr_code ~ exp, value.var = "porcentaje") %>% 
+  .[order(.[,'bufr_code']), ] %>% 
+  kbl(booktabs = T, col.names = c("Obs type", "CONV", "AUT", "SATWND"),
+      caption = "Observation types assimilated in each experiment.") %>% 
+  kable_classic(full_width = F)
+
+# purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
+#   # .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
+#   dcast(bufr_code ~ exp, value.var = "N") %>% 
+#   .[order(.[,'bufr_code']), ] %>% 
+#   kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
+#   kable_classic_2(full_width = F)
+```
+
+```{r}
+fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
+  .[, ":="(sensor = toupper(sensor),
+           plataforma = toupper(plataforma),
+           prop = scales::percent(prop/100, accuracy = 0.01, trim = TRUE))] %>% 
+  .[order(.[,'sensor']), ] %>% 
+  .[] %>% 
+  kbl(booktabs = T, col.names = c("Sensor", "Plataform", "Assimilated\nchannels", "Percentage over total"),
+      caption = "List of the available sensors in several platforms, the number of accepted channels and the percentage of assimilated observations calculated over all radiance observations.") %>% 
+  collapse_rows(1,valign = "top") %>% 
+  # column_spec(1, width = "3cm", latex_valign = "p") %>% 
+  kable_classic_2(full_width = F)
 ```
@@
 ) %>% 
-  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
+  kbl(booktabs = T, col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
       caption = "Observation types assimilated in each experiment.") %>% 
@@
 
-```{r count_obs, fig.height=6, fig.width=10}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]
+```{r dist_obs, fig.cap="a) Horizontal spatial distribution of the mean available observations per analysis cycle for CONV, AUT and SATWND.
+b) Number of radiance observations assimilated at each analysis cycle and pressure level."}
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]
 
@@
 
-files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
+files <- files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
 
@@
 
-files <- files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]
+files <- files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]
 
@@
 
-p1 <- rbind(conv, aut, satwnd) %>% 
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")
+
+files <- files[!str_detect(files, "conv")]
+
+rad <- read_diag_rad(files, "E8") %>% 
+  satinfo[., on = c("sensor", "channel")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+rbind(conv, aut, satwnd) %>% 
   .[usage.flag == 1 & rerr != 1.0e+10] %>% 
@@
   scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
-                       guide = guide_colorbar(barwidth = 20)) +
+                       guide = guide_colorbar(barwidth = 10,
+                                              barheight = 0.5)) +
   geom_mapa() +
@@
   labs(fill = "Mean number of\nobs per cycle") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
-p2 <- rbind(conv, aut, satwnd) %>% 
+  theme_minimal( base_size = 6) +
+  theme(legend.position = "bottom",
+        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
+
+rad %>% 
+  .[qc == 0 & iuse == 1] %>% 
+  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
+  .[, .(omb = mean(tbc),
+        n = .N), by = .(date, lev, exp)] %>% 
+  .[lev != 1025] %>% 
+  ggplot(aes(date, fct_rev(lev))) +
+  geom_tile(aes(fill = n)) +
+  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
+                       guide = guide_colourbar(barheight = 8,
+                                               barwidth = 0.5)) +
+
+  scale_date() +
+   facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E8 = "RAD"))) +
+  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
+  theme_minimal(base_size = 6) +
+  theme(aspect.ratio = 1/1.5) +
+
+  plot_layout(ncol = 2, widths = c(1.6, 1)) + plot_annotation(tag_levels = 'a', tag_suffix = ")") &
+  theme(plot.tag = element_text(size = 6))
+```
+
+```{r count_obs, fig.cap="a) Number of available observations per cycle during the experiment. a) Vertical spatial distribution of the mean available observations per analysis cycle."}
+rad2 <- rad[, .(press, iuse, error, exp, date)] %>% 
+  setnames(c("press", "iuse", "error"), c("pressure", "usage.flag", "rerr"))
+
+rbind(conv, aut, satwnd) %>%
+  .[, .(pressure, usage.flag, rerr, exp, date)] %>% 
+  rbind(., rad2) %>% 
+  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
+  .[, .(count_obs = .N), by = .(exp, date)] %>% 
+  ggplot(aes(date, count_obs)) +
+  geom_line(aes(color = exp), size = 0.2) +
+  geom_point(aes(color = exp, shape = exp), size = 1) +
+  scale_color_manual(values = colores_exp, 
+                     labels = c(E2 = "CONV", E5 = "AUT", 
+                                E6 = "SATWND", E8 = "RAD")) +
+  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "AUT", 
+                                                        E6 = "SATWND", E8 = "RAD")) +
+  scale_y_log10() +
+  scale_date() +
+  labs(y = "Number of obs per cycle", x = NULL,
+       color = NULL, shape = NULL) +
+  theme_minimal(base_size = 6) +
+  theme(legend.position = "bottom", 
+        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
+
+  rbind(conv, aut, satwnd) %>%
+  .[, .(pressure, usage.flag, rerr, exp, date)] %>% 
+  rbind(., rad2) %>% 
   .[usage.flag == 1 & rerr != 1.0e+10] %>% 
@@
   ggplot(aes(lev.box, obs_cycle)) +
-  geom_line(aes(color = exp)) +
-  geom_point(aes(color = exp, shape = exp), size = 2) +
+  geom_line(aes(color = exp), size = 0.2) +
+  geom_point(aes(color = exp, shape = exp), size = 1) +
   scale_color_manual(values = colores_exp, 
@@
                                 E6 = "SATWND", E8 = "RAD")) +
-  scale_shape_manual(values = c(15, 16, 17), labels = c(E2 = "CONV", E5 = "AUT", 
-                                                        E6 = "SATWND", E7 = "RAD")) +
+  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "AUT", 
+                                                        E6 = "SATWND", E8 = "RAD")) +
   # geom_tile(aes(fill = obs_cycle)) +
@@
        x = "Pressure level (hPa)", y = "Mean number of\nobs per cycle") +
-  theme_minimal() +
-  theme(aspect.ratio = 2.5/1, 
+  theme_minimal(base_size = 6) +
+  theme(legend.position = "bottom",
+        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
+
+  plot_layout(ncol = 2, widths = c(4, 1), guides = "collect") + 
+  plot_annotation(tag_levels = 'a', tag_suffix = ")") &
+  theme(plot.tag = element_text(size = 6), 
         legend.position = "bottom",
+        legend.margin = unit(c(0, 0, 0, 0), "cm"),
         plot.margin = unit(c(0, 0, 0, 0), "cm"))
 
-p3 <- rbind(conv, aut, satwnd) %>% 
-  .[usage.flag == 1 & rerr != 1.0e+10] %>% 
-  .[, .(count_obs = .N), by = .(exp, date)] %>% 
-  ggplot(aes(date, count_obs)) +
-  geom_line(aes(color = exp)) +
-  geom_point(aes(color = exp, shape = exp), size = 2) +
-  scale_color_manual(values = colores_exp, 
-                     labels = c(E2 = "CONV", E5 = "AUT", 
-                                E6 = "SATWND", E7 = "RAD")) +
-  scale_shape_manual(values = c(15, 16, 17), labels = c(E2 = "CONV", E5 = "AUT", 
-                                                        E6 = "SATWND", E7 = "RAD")) +
-  scale_date() +
-  labs(y = "Number of obs per cycle", x = NULL,
-       color = NULL, shape = NULL) +
-  theme_minimal() +
-  theme(aspect.ratio = 1/2,
-        legend.position = "bottom", 
-        plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
-cowplot::plot_grid(p1, NULL, p2, p3, ncol = 3,  rel_widths = c(1.3, -0.3, 1))
-
-```
-
-```{r}
-files <- Sys.glob(here("analysis/data/derived_data/count_obs_E*"))[c(1, 3, 4)]
-
-purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
-  .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
-  dcast(bufr_code ~ exp, value.var = "porcentaje") %>% 
-  .[order(.[,'bufr_code']), ] %>% 
-  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
-  kable_classic_2(full_width = F)
-
-purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
-  # .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
-  dcast(bufr_code ~ exp, value.var = "N") %>% 
-  .[order(.[,'bufr_code']), ] %>% 
-  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
-  kable_classic_2(full_width = F)
-```
-
-```{r}
-fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
-  .[, ":="(sensor = toupper(sensor),
-           plataforma = toupper(plataforma),
-           prop = scales::percent(prop/100, accuracy = 0.01, trim = TRUE))] %>% 
-  .[order(.[,'sensor']), ] %>% 
-  .[] %>% 
-  kbl(col.names = c("Sensor", "Plataform", "Number of assimilated\nchannels", "Percentage over total"),
-      caption = "List of the available sensors in several platforms, the number of accepted channels and the percentage of assimilated observations calculated over all radiance observations.") %>% 
-  collapse_rows(1,valign = "top") %>% 
-  # column_spec(1, width = "3cm", latex_valign = "p") %>% 
-  kable_classic_2(full_width = F)
-```
-
-```{r rad, fig.cap="Number of radiance observations for each 50 hPa pressure layer over time"}
-
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E8/ANA/*/diagfiles/asim*ensmean")
-
-files <- files[!str_detect(files, "conv")]
-
-rad <- read_diag_rad(files, "E8") %>% 
-  satinfo[., on = c("sensor", "channel")]
-
-rad %>% 
-  .[qc == 0 & iuse == 1] %>% 
-  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
-  .[, .(omb = mean(tbc),
-        n = .N), by = .(date, lev)] %>% 
-  .[lev != 1025] %>% 
-  ggplot(aes(date, fct_rev(lev))) +
-  geom_tile(aes(fill = n)) +
-  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
-                       guide = guide_colourbar(barheight = 18,
-                                               barwidth = 0.8)) +
-
-  scale_date() +
-  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
-  theme_minimal()
-
-```
-```{r level-rad, fig.cap="Nivel de presión asociado a las observaciones en cada ciclo de asimilación"}
-
-rad %>% 
-  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
-  .[sensor %in% c("airs", "iasi") &
-      # errinv %between% c(0.0000316227, 1000) & 
-      # qc == 0 & 
-      # iuse == 1 &
-      peakwt %between% c(0.001, 1200)] %>% 
-  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
-  ggplot(aes(channel, Q50)) +
-  # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
-  #             outlier.size = 0.5, size = 0.1) +
-  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4) +
-  geom_line() +
-  scale_y_level(name = "Pressure (hPa)") +
-  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
-  labs(x = "", y = "") +
-  theme_minimal() +
-
-  rad %>% 
-  separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
-  .[!sensor %in% c("airs", "iasi") &
-      # errinv %between% c(0.0000316227, 1000) & 
-      # qc == 0 & 
-      # iuse == 1 &
-      peakwt %between% c(0.001, 1200)] %>% 
-  .[, as.list(quantile(peakwt, c(0.05, 0.5, 0.95))) %>% setNames(c("Q5", "Q50", "Q95")), by = .(sensor, channel)] %>% 
-  ggplot(aes(channel, Q50)) +
-  # geom_boxplot(aes(group = factor(channel)), outlier.color = "darkorange", 
-  #             outlier.size = 0.5, size = 0.1) +
-  geom_ribbon(aes(ymin = Q5, ymax = Q95), alpha = 0.4, fill = "darkorange") +
-  geom_line(color = "darkorange") +
-  scale_y_level(name = "Pressure (hPa)") +
-  facet_wrap(~sensor, scales = "free", labeller = labeller(sensor = toupper)) +
-  labs(y = "", x = "Channel index") +
-  theme_minimal() +
-  plot_layout(ncol = 1)
-```
+```
+
+#### Validation dataset
+
+In order to evaluate the performance of the ensemble based data assimilation system presented in this article, we use part of the following observations datasets:
+
+* The Multi-Network Composite Highest Resolution Radiosonde Data [] from the RELAMPAGO field campaign database consists of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
+Only the radiosondes that did not enter the assimilation system were used to perform the validation.
+The experiment period covers the Intensive Observation Periods (IOP) 7 and 8 during which XXXX radiosondes were launched in a small area Figure 1 b).
+
+* The Satellite precipitation estimation IMERG half-hourly Final  Run of  0.01◦ resolution (Huffman et al.
+2018) was used as a reference state to validate the skill of the short-range forecasts to represent the precipitation over the domain.
+
+* Radar observations are used to perform a qualitative and visual assessment of the convection features.
+The data comes from 9 radars located on the domain and is provided by the recently developed Argentine C-band Doppler dual-polarization weather radar network \citep{deelia2017} with a temporal frequency of 10 minutes.
+For this work only the maximum reflectivity in the column (COLMAX) field close to the analysis time was used.
+
+### Experimental design
+
+To investigate the impact of different observations upon the analysis, four data assimilation experiments were performed using different observation sets (Table \@ref(tab:obs)).
+The CONV experiment uses only convectional observations from PREPBUFR.
+The orange triangles in Figure \@ref(fig:dominio) a), are the surface stations and ship locations included in this experiment.
+
+In a second experiment, referred to as AUT, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from ASWS.
+The automatic stations are indicated as green squares in Figure \@ref(fig:dominio) a).
+These stations have greater spatial coverage and a higher sampling frequency (10 minutes in most cases).
+
+In the third experiment, referred to as SATWND, we assimilate all the observations of the AUT experiment and the satellite-derived motion winds.
+Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbit satellites as described in Section XX.
+
+The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure 2 a).
+The larger values of assimilated observations over the center of the domain and the first pressure levels correspond to the ASWS observations.
+In Figure 3 a) the number of assimilated observations over time is shown.
+In this figure the peak/maximum at 12 UTC each day for CONV, AUT and SATWND is produced by the operational observations like radiosondes that are launched at this time.
+The strong variability in the number of radiance observations per cycle is also noticeable.
+The vertical distribution of mean number of observations per cycle shows a maximum in low levels due to the ASWS observations and a local maximum around 400-300 hPa in SATWND is due to the satellite-derived winds.
+Above 800 hpa, radiance observations are in the majority.
+
+All the experiments start at 12 UTC Nov 20, 2018 with 6 hours of simulation spin up (without assimilation).
+The assimilation cycles start at 18 UTC 20 Nov, 20 and continue until 12 UTC Nov, 23 (67 hours/assimilation cycles).
+Two independent ensemble forecasts whose validation is discussed in a companion paper or on part II of this paper were initialized from the analysis at 00 and 06 UTC Nov, 22 and continue until 12 UTC Nov, 23 (Figure 3).
+
+```{r ciclo, fig.cap="Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
+The zoomed section shows the assimilation for each hour that is performed within a one-hour centered window and new boundary conditions from GFSD every 6 hours.
+Also the two Intensive Observation Periods (IOP) from the RELAMPAGO Field Campaign are shown."}
+knitr::include_graphics(here("analysis", "figures", "analysis_cycle.pdf"))
+```
+
+### Verification methods
+
+A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper.
+These aspects include a validation of how uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations not used in the assimilation.
+To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use  the Reduced Centered Random Variable (RCRV, Candille et al., 2007) which is define as  defined as:
+
+$$RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}$$,
+
+Where $x_o$ is the error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble
+The average of  $RCRV$ computed over all the realizations represents the weighted bias between the ensemble and observation:
+
+$$\mathit{mean RCRV} = E[RCRV]$$
+
+The standard deviation of this variable measures the systematic over- or under- dispersion of the ensemble:
+
+$$\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}$$
+
+Where $M$ is the sample size.
+The second moment of the $RCRV$ measures the agreement of the ensemble spread and the specific observational error with the observed amplitude of the forecast error.
+A reliable system will have no bias, $mean RCRV = 0$, and a standard deviation equal to 1, $sd RCRV = 1$.
+If the ensemble has a positive (negative) bias, $mean RCRV$ will be positive (negative).
+A $sd RCRV > 1$ ($sd RCRV < 1$), indicates that the ensemble is underdispersive (overdispersive).
+
+The fit of the first-guess and analysis to a set of independent observations is computed based on the Root Mean Square Error (RMSE) and the BIAS:
+
+$$\mathit{RMSE} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)^{2}}$$
+
+$$\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i  - O_i)$$
+
+Where $O$ and $X$ stand for independent observations and the analysis respectively.
+
+For the comparison of the first-guess’ precipitation with the precipitation estimates we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al.
+2021 and adapted from roberts2008.
+
+$$\mathit{FSS} = 1-\frac{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i)^{2}+\frac{1}{N}\sum_{i=1}^{N} ({P_o}_i)^{2}} $$,
+
+where ${P_x}$ and ${P_o}$ are the probability of occurrence of specified rainfall accumulation of the simulated ensemble and the reference atmospheric state, respectively.
+To evaluate the grid points fraction in which the precipitation exceeds certain thresholds, the probability of occurrences for the whole ensemble was used, avoiding the systematic underestimation of high precipitation values by the ensemble mean.
+The FSS was applied to the 6 hr accumulated precipitation ensemble first-guess (i.e.
+adding the precipitation summed in the 6 previous analysis cycles).
 
@@
 ### Ensemble reliability
+
+To investigate the ability of the ensemble to represent the possible states of the atmosphere taking into account the uncertainties of the model and the observations we calculated the mean and sd of the first-guess RCRV for the RAD experiment between 18 UTC Nov, 20 and 12 UTC Nov, 23.
+As this experiment assimilates all types of observations used in this work, it is possible to analyse the reliability of the ensemble by comparing with each type of observation.
+Figure 5 shows the sd RCRV for surface observations computed and box-averaged to a 2.5° grid.
+The wind (Figure 5 a) has a $sd RCRV$ close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error.
+For the temperature (Figure 6 b and d), the results are similar except for some areas to the west of the domain where sd RCRV can be as high as 4.5.
+These areas were studied in detail and it was found that the mean innovation was overly large.
+The main reason for this is the complexity of the topography, usually associated with differences between the model topography and the station height leading to relatively large errors.
+
+Figure 6 shows the mean and standard deviation of the RCRV for the upper-air observations.
+Figure 7 a and b show the RCRV statistics in soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
+Both ADPUPA and AIRCFR show a general good agreement between the ensemble spread and the observation error while AIRCAR presents an irregular profile with values that suggests that the error for this type of observation is overestimated.
+ADPUPA and AIRCAR also present a mean RCRV profile near zero on middle and upper levels but ADPUPA shows a cold bias at low levels, a characteristic already studied in Ruiz 2010 and Dillon 2021.
+
+For satellite-derived wind observations (Figure 7 c)) in low levels, the profiles of b and d that are not close to the expected values with a negative bias and an overestimation of the observation error.
+This characteristic persists throughout all levels.
+Wind estimation derived from water vapor channels shows a bias close to zero in upper levels.
+
+The mean RCRV profiles calculated from the radiance observations (Figure 7 d)) shows almost no bias and the same happens if the mean RCRV is calculated over each channel of each sensor indicating that the bias correction algorithm worked as expected.
+The sd RCRV values are less than 1 for all sensors possible due to an overestimation of the observation errors.
+
+Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work   Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate.
 
@@
   scale_linetype_manual(values = c("dashed", "solid"),
-                        labels = c("Mean RCRV", "SD RCRV")) +
+                        labels = c("mean RCRV", "sd RCRV")) +
   scale_color_brewer(palette = "Dark2") +
@@
   read_csv(here::here("analysis", "data", "derived_data", "rcrv_satwind.csv")) %>% 
+  setDT() %>% 
+  .[!(sat_type %in% c("GOES visible", "EUMETSAT visible") & error.level == 700)] %>% 
   melt(measure.vars = c("mean.y", "sd.y")) %>%
@@
   scale_linetype_manual(values = c("dashed", "solid"),
-                        labels = c("Mean RCRV", "SD RCRV")) +
+                        labels = c("mean RCRV", "sd RCRV")) +
   scale_color_brewer(palette = "Set1") +
@@
   scale_linetype_manual(values = c("dashed", "solid"),
-                        labels = c("Mean RCRV", "SD RCRV")) +
+                        labels = c("mean RCRV", "sd RCRV")) +
   scale_color_discrete(labels = toupper) +
@@
 
-```{r TQ-diff, fig.cap="Difference between experiments for the average vertical profile of variables over time.", fig.scap=c("Temperature (°C).", "Specific Humidity (g/kg)."), fig.height=3.5}
+In this section we  investigate the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
+The analysis is performed over a smaller domain (red box in Figure 1) to avoid high topography effects and to focus on the region most directly affected by the MCS.
+Figure 8 a to c shows the difference between AUT and CONV in the spatially averaged vertical profile of temperature.
+The assimilation of  ASWS observations results in a colder planetary boundary layer (PBL).
+This cooling effect has a clear diurnal cycle that is stronger during night time.
+(Figure 8 a).
+The impact of ASWS observations is not limited to the PBL, as significant impact  in mid to upper levels is also found during the mature stage of the MCS.
+The warming in AUT present between 500 and 300 hPa is produced by the development of stronger convection in this experiment compared to CONV.
+SATWND observations have an almost negligible impact on the mean temperature (Figure 8 b)).
+The assimilation of radiances produces significant impacts in low level temperature.
+These observations produce a warming effect in the analyzed PBL which partially compensate for the cooling effect of ASWS observations, particularly during the first 24 hours.
+After that period the impact of assimilating radiances becomes smaller in the first levels.
+The impact of the radiances is also important during the mature stage of the MCS and extended to the end of the simulation period.
+
+Comparing the specific humidity in the experiments (Figure 8 d-f) the impact of assimilating  ASWS with larger spatial and temporal resolution is most significant at low levels (Figure 8 d)).
+The PBL in the AUT experiment is consistently moister than in the CONV experiment.
+The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [Ruiz 2010, algún otro del SMN?].
+The impact of satellite-derived winds are smaller and only important after 18 UTC Nov, 21 at low levels.
+As for the temperature, the impact of RAD observations reduces the impact of ASWS at low levels until 12 UTC Nov, 22.
+Radiances also reduce low-middle level moisture above the boundary layer where the impact of surface observations is not so important.
+This effect is significant during the development of the MCS between 00 and 12 UTC Nov, 22.
+
+The impacts on the wind components are shown in Figure 9.
+The shaded values correspond to the difference between the experiments while the contours represent the spatially averaged vertical wind profile for the experiment highlighted in bold on each figure.
+For example, Figure 9 a) shows the difference between AUT and CONV (shaded) and the averaged zonal wind for AUT as is the experiment with more assimilated observations on that panel.
+
+The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
+Therefore, these observations generate a reduction of the northerly flow associated with the South American low level jet during night time.
+After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain.
+At upper levels, the sign of the impact varies throughout the experiment period becoming larger  during the development of the MCS.
+During Nov 22 and 23 the impact of assimilating ASWS in AUT, produces an increase of northerly wind in upper levels.
+This could indicate a stronger development of convection and therefore an increase in the outflow of convective cells.
+The assimilation of satellite-derived winds produces impacts only in middle and upper levels.
+The impact in the northerly wind is an increase of 0.8 m/s on average.
+
+The assimilation of radiance observations produced a reduction in the westerly wind compared with SATWIND in low and upper levels.
+For the meridional wind, these observations produce an enhancement of the northerly low-level flow of $1 ms^{-1}$ on average, opposite to what is generated by the assimilation of ASWS observations.
+On upper levels and during Nov 22 and 23 the impact of assimilating radiances have the opposite effect on the meridional wind when compared with the impact of the ASWS observations.
+
+To summarize these results, ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly at low levels.
+These changes in the planetary boundary layer can substantially affect the development of the convection.
+On the other hand, radiances also produce important impacts but these are, in general, opposite to those of ASWS observatios.
+Spatial fields of different variables at 00 UTC Nov, 22 will be analyzed next to obtain an overview  of the environment prior to the development of the MSC.
+
+To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22.
+At that moment the first convection cells were developing on the southern region of the domain along the cold front.
+Figure 10 a) shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
+We found that the moist tongue extending over Northern and central Argentina is enhanced by the assimilation of denser surface observations.
+The moisture increase is particularly stronger at the southern tip of this tongue at the boundary of the relatively colder air mass where the convection leading to the MCS was triggered.
+As a result AUT and SATWND experiments are very similar, with values of precipitable water over 55 kg/m2 north of 30°S and a similar vertical distribution of specific humidity (not shown).
+RAD has lower precipitable water content than previous experiments but not as low as CONV.
+The distribution of moisture at low and medium levels observed in RAD can be explained by a combination of the drying that generates the assimilation of radiances and a higher moisture transport due to the stronger northerly wind with values above 10 m/s over the centre of the domain.
+
+The potential temperature on the PBL (Figure 11 b) resembles the characteristics observed on the temperature profiles in Figure 8 a and c.
+On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL due to the assimilation of radiance observations.
+Figure 11 c) shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 km wind shear.
+The values of MCAPE in CONV do not exceed 2000 JKg-1 while the rest of the experiments show maximum MCAPE over 4000 JKg-1.
+The decrease of the specific humidity in lower levels present in RAD could explain the slightly lower values of MCAPE that is partially compensated by an increase in the temperature in the PBL, compared with AUT or SATWND.
+The wind shear is more intense in RAD  reaching  values over 15 ms-1 over the area of maximum MCAPE at the southern tip of the northerly low-level flow.
+Stronger wind shear between 0 and 6 km has been associated with the development of more intense and organized MCS [Chen 2015, ] and also with conditions favorable for supercells (e.g.
+Markowsky and Richardson 2010).
+Particularly in this case we can see that on the Southern tip of the strong CAPE region, the 0-6 km wind shear is over 15 ms-1 which is considered favorable for supercell occurrence, while in the CONV experiment the intersection of the unstable warm air and the high shear is slightly smaller and occurs at lower MCAPE values.
+
+```{r TQ-diff, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of temperature (a, b, and c, in K) and specific humidity (d, e, and f in gkg-1).", fig.scap=c("Temperature (°C).", "Specific Humidity (g/kg)."), fig.height=3.5}
 files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
@@
   facet_wrap(~variable, ncol = 3,
-             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
-                                              "E6_E5" = "SATWND - AUT",
-                                              "E8_E6" = "RAD - SATWND"))) +
+             labeller = labeller(variable = c("E5_E2" = "**AUT** - CONV",
+                                              "E6_E5" = "**SATWND** - AUT",
+                                              "E8_E6" = "**RAD** - SATWND"))) +
   tag_facets(position = list(x = 0.1, y = 0.96)) +
@@
         panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
+        panel.grid = element_line(linetype = 3), 
+        strip.text = ggtext::element_markdown())
 
@@
   geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
-                  melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                breaks = seq(-9, 0, 1), size = 0.2, linetype = 2, color = "grey30") +
+                  melt(id.vars = c("lev", "date")) %>% unique(), 
+                aes(z = value, linetype = factor(-sign(..level..))), 
+                breaks = seq(-10, 4, 2), size = 0.2, color = "grey30") +
   geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
-                    breaks = seq(-9, 0, 1), color = "grey30", skip = 1, 
+                    breaks = seq(-10, 4, 2), color = "grey30", skip = 1, 
                     size = 4, stroke = 0.1) +
@@
        y = NULL) +
+  scale_linetype(guide = NULL) +
   scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
@@
   facet_wrap(~variable, ncol = 3,
-             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
-                                              "E6_E5" = "SATWND - AUT",
-                                              "E8_E6" = "RAD - SATWND"))) +
+             labeller = labeller(variable = c("E5_E2" = "**AUT** - CONV",
+                                              "E6_E5" = "**SATWND** - AUT",
+                                              "E8_E6" = "**RAD** - SATWND"))) +
   tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
@@
         panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
+        panel.grid = element_line(linetype = 3),
+        strip.text = ggtext::element_markdown())
 
@@
 
+Analysis verifications are performed to assess the skill of each experiment to represent the atmospheric circulation, particularly associated with the MSC development and precipitation.
+Figure 12 shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67°W and 54.5°W as a function of time and latitude in the different experiments, and estimated precipitation by IMERG.
+The heaviest precipitation (over 12 mm/h) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the experiments.
+In all the experiments, the accumulated precipitation in the short-range forecasts  underestimate the precipitation.
+This is particularly evident on CONV where the convection initiation is significantly delayed with respect to the observations and occurs further north with respect to the observed initiation.
+AUT, SATWND and RAD better capture the timing and location of convective initiation.
+On RAD the initiation is delayed respecto to AUT and SATWND possibly related to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
+After 18 UTC Nov, 22 RAD shows improvements in precipitation rate and its distribution compared to the other experiments due to the enhanced development of the conveccion.
+
+To quantify the spatial match between the estimated precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure 13).
+All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, afterwards FSS values improve during the mature and dissipation stages of the MCS.
+The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the maturity and dissipation stages of the MCS.
+AUT and  SATWND show similar FSSs indicating that satellite-derived winds assimilation has little impact on the precipitation for the present case study.
+The assimilation of radiance observations improves the skill of the 1-hour forecasted precipitation as the FSS in RAD is higher, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
+The enhancement is particularly important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).
+
+To complement  the analysis of the representation of the MCS evolution by the different experiments, Figure 14 shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments  ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed reflectivity.
+These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations.
+In addition, these experiments are the worst and best performing experiments with respect to precipitation.
+Overall none of the short range forecasts is able to capture the mesoscale details in the precipitation distribution.
+This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately capture the strength of the convective band associated with the MCS.
+Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field.
+RAD better represents the observed features of the MSC showing an organized pattern over the centre domain at 10 and 13 UTC.
+The convection cells that initiate after 16 UTC in the northeast region are also present in the 1-hr forecast.
+CONV also captures the location of the MCS, but the convection seems to be less organised and much weaker than in RAD.
+Before and after the times shown in Figure 14, the agreement between experiment and observations is quite high in the regions where radar data are available.
+
+Individual radiosondes were also compared with the experiments to analyse the specific characteristics of the atmosphere before and after the development of the MCS.
+
+Figure 15 shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO IOP07 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP08 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+During the first period, no significant meteorological events were observed on the domain of interest, thus the synoptic situation corresponds to a warm day with an important humidity transport from the centre of South America.
+For the temperature, CONV presents a warm bias on the PBL of 1°C and an RMSE between 1 and 2.5°C that are slightly reduced  by the assimilation of ASWS and radiances.
+The RMSE for temperature remains between 0.5 and 2°C throughout the profile and the assimilation has almost no effect in middle and upper levels.
+The dew point temperature shows a small dry bias in low levels that is corrected on AUT and SATWND.
+On middle levels (below 10 km) the bias is large and positive (8 - 10°C), indicating a more moist atmosphere in the analysis.
+The RMSE also grows with height where small variations of humidity produces major chances in dew point temperatures.
+The zonal wind has a significant positive bias around 10 km of 10 m/s, this means that all experiments present a stronger westerly wind that is in part corrected by the assimilation of radiance observations.
+Finally, the meridional wind shows small values of bias and RMSE through the vertical profile but the assimilation of radiance observations increases the northerly wind near the surface.
+
+For the second period, corresponding to IOP08, none of the radiosondes was directly affected by the convective processes since the MSC crossed RELAMPAGO's region of study before the beginning of the observation period.
+Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
+Between 5 and 12 km, the bias is cold and only the ASWS observations improves the analysis.
+The RMSE for AUT, SATWND and RAD remains below 2°C for almost all levels with small improvement in upper levels where radiance observations have the most impact.
+In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
+The moist bias in conjunction with the underestimation of the precipitation might be indicating that the convection processes on the model are not converting the available moisture into precipitation or that the model fails to represent boundary layer processes correctly.
+SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7°C.
+Once again, the zonal wind is overestimated on the analyses with a positive bias on middle and upper layers.
+For this period only RAD shows an improvement on the upper layers of the troposphere.
+In low levels the meridional wind presents a positive bias, indicating an underestimation of the northerly wind principally in AUT, SATWND and RAD.
+In middle levels both AUT and SATWND have almost no bias while RAD shows a positive bias of 2.5 m/s while its RMSE does not exceed 5 m/s.
+
+To summarize, the results suggest that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study.
+While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels.
+The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the CSM leading to an increase in the accumulated precipitation.
+
 ```{r pp-hov, fig.cap="Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right, for the ensemble mean) over the inner domain."}
 
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/ppacum/E[2568]_ana*mean_acum_1h.rds")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2568]_ana*mean_acum_1h.rds")
 
@@
 
-```{r dbz-mean, fig.cap="Figure 14. Observed maximum reflectivity in the column (upper row) and ensemble mean maximum reflectivity for CONV and RAD at different times (bottom two rows)."}
+```{r dbz-mean, eval=FALSE, fig.cap="Observed maximum reflectivity in the column (upper row) and ensemble mean maximum reflectivity for CONV and RAD at different times (bottom two rows).", include=FALSE}
 proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"
@@
 
-files <- Sys.glob("/home/jruiz/share/DATA/DATOS_RADAR/MOSAICO/*.nc")[seq(1, 12, 3)]
+files <- Sys.glob("/home/jruiz/datosmunin3/datos/DATOS_RADAR/MOSAICO/*.nc")[seq(1, 12, 3)]
 # Solo un par de horarios 10, 13, 16 y 19 UTC
@@
 ```{r soundings, fig.cap="RMSE and Bias of temperature, dew point temperature and wind components calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8."}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos/ANA/sondeo_E[2568]*")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos/ANA/sondeo_E[2568]_ana*")
 
@@
   .[, lev := cut_round(alt, c(seq(0, 3000, 500), seq(4000, 21000, 1000)))] %>% 
-  .[, ":="(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
-           BIAS = mean(value - fcst_value, na.rm = TRUE)), by = .(iop, exp, variable, lev)] %>% 
+  .[, ":="(RMSE = mean(sqrt((fcst_value - value)^2), na.rm = TRUE), 
+           BIAS = mean(fcst_value - value, na.rm = TRUE)), by = .(iop, exp, variable, lev)] %>% 
   melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
@@
 ```
+
+## Conclusions

### 2021-05-28 analysis/scripts/cor_ana.R
Graficos de correlacion entre t y td

@@ -1,0 +1,179 @@
+library(tidyverse)
+library(lubridate)
+library(data.table)
+library(unglue)
+library(metR)
+library(mesoda)
+source(here::here("temp_R/postprocesamiento.R"))
+
+fisica <- data.table(mem = as.character(formatC(1:60, flag = "0", width = 3)),
+                     fisica = rep(c("KF-YSU",
+                                    "BMJ-YSU",
+                                    "GF-YSU",
+                                    "KF-MYJ",
+                                    "BMJ-MYJ",
+                                    "GF-MYJ",
+                                    "KF-MYNN2",
+                                    "BMJ-MYNN2",
+                                    "GF-MYNN2"), length.out = 60)) %>% setDT()
+# geom_mapa <- function(fill = NA) {
+#   list(geom_sf(data = map_arg, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+#        geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+#        coord_sf(ylim = c(-34, -29), xlim = c(-66, -63)),
+#        scale_x_longitude(ticks = 1.5),
+#        scale_y_latitude(ticks = 1.5))
+# }
+
+coord <- ReadNetCDF("/home/paola.corrales/datosmunin3/EXP/E2/ANA/20181120180000/analysis.ensmean",
+                    vars = c(lon = "XLONG", lat = "XLAT")) %>%
+  .[, Time := NULL]
+
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
+ini_date <- ymd_hms("20181120180000")
+ciclos <- 67
+
+dates <- seq(ini_date, by = "hour",
+             length.out = ciclos)
+
+cor <- list()
+cor_fisicas <- list()
+
+for (d in seq_along(dates)) {
+
+  print(dates[d])
+
+
+  files_ana <- Sys.glob(paste0(wrf_path, "E8/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.mem*"))
+
+  # files_gue <- Sys.glob(paste0(wrf_path, "E[2,5,6,8]/GUESS/", format(dates[d], "%Y%m%d%H%M%S"), "/wrfarw.mem*"))
+
+  ana <- purrr::map(files_ana, function(f) {
+
+    metadatos <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/ANA/{fecha}/analysis.mem{mem}")
+
+
[…]

### 2021-06-24 analysis/chapters/paper_ana_20181122.Rmd
Intro con referencias. Y otras yerbas asociadas al yaml

@@
 ---
-title: Assimilation of conventional and satellite observations in a deep convection
-  case during RELAMPAGO using the WRF-GSI-LETFK system.
+title: "Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis"
 author:
@@
   email: paola.corrales@cima.fcen.uba.ar
-  affiliation: Some Institute of Technology
+  affiliation: UBA,CIMA,CNRS
   footnote: 1
-- name: Bob Security
-  email: bob@example.com
-  affiliation: Another University
-- name: Cat Memes
-  email: cat@example.com
-  affiliation: Another University
-  footnote: 2
-- name: Derek Zoolander
-  email: derek@example.com
-  affiliation: Some Institute of Technology
-  footnote: 2
+- name: Victoria Galligani
+  affiliation: UBA,CIMA,CNRS
+- name: Juan Ruiz
+  affiliation: UBA,CIMA,CNRS
+- name: Luiz Sapucci
+  affiliation: INPE
+- name: María Eugenia Dillon
+  affiliation: SMN,CONICET
+- name: Yanina García Skabar
+  affiliation: SMN,CONICET,CNRS
+- name: Maximiliano Sacco
+  affiliation: SMN
 date: "`r Sys.Date()`"
@@
 abstract: |
-  This is the abstract.
-  It consists of two paragraphs.
+  The impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF is evaluated.
+We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November, 22th, 2018.
+The accumulated precipitation observed during the transit of the MCS was quite high with values over 200 mm over the northern part of the domain.
+This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during Nov.
+2018.
+
+  We used the GSI-4DLETKF to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
+A multiphysics approach is also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
+We conducted four assimilation experiments using different sets of observations: CONV with conventional observations from prepBUFR, AUT that uses CONV observations plus automatic station networks, SATWND add satellite-derived winds and RAD includes satellite radiances from AMSU, HIRS, MHS, ATMS, AIRS and IASI.
+We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study.
+The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
+
 journal: Atmospheric Research
-layout: 5p
+layout: authoryear,preprint,review,12pt
+#layout: final,5p,times,twocolumn,authoryear
 address:
-- code: Some Institute of Technology
-  address: Department, Street, City, State, Zip
-- code: Another University
-  address: Department, Street, City, State, Zip
+- code: UBA
+  address: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
+- code: CIMA
+  address: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
+- code: CNRS
+  address: CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.
+- code: SMN
+  address: Servicio Meteorológico Nacional de Argentina
+- code: CONICET
+  address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas)
+- code: INPE
+  address: National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies
 bibliography: paper_ANA.bib
@@
   text: Corresponding Author
-- code: 2
-  text: Equal contribution
+keywords: Regional Data Assimilation, Surface Obsevations, Satellite Observations 
+numbersections: true
 ---
@@
 knitr::opts_chunk$set(
-  echo = FALSE,
+	eval = FALSE,
 	fig.path = "../figures/",
 	fig.retina = TRUE,
-  # fig.height = 3,
-  dev = "png",
 	message = FALSE,
@@
 	comment = "#>",
-  dpi = 300
+	dev = "png",
+	dpi = 300,
+	include = FALSE
 )
@@
 
-arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10)
-arg_topo[, lon := metR::ConvertLongitude(lon)]
-
-cordillera <- list(
-  ggnewscale::new_scale_fill(),
-  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
-                                         z = h),
-                    breaks = seq(1500, 8000, by = 700)),
-  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
-                      oob = scales::squish))
+# arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10)
+# arg_topo[, lon := metR::ConvertLongitude(lon)]
+# 
+# cordillera <- list(
+#   ggnewscale::new_scale_fill(),
+#   geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
+#                                          z = h),
+#                     breaks = seq(1500, 8000, by = 700)),
+#   scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
+#                       oob = scales::squish))
 
@@
 A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
-The north and central region of Argentina has been shown to have the highest frequency of favorable conditions for high-impact meteorological events [@brooks2003] and the highest occurrence of large hail events [@cecil2012].
+Southern South America has been shown to have the highest frequency of favorable conditions for high-impact meteorological events [@brooks2003] and the highest occurrence of large hail events [@cecil2012].
 This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014].
@@
 However surface observations, which directly measure the atmospheric state, have limited availability in South America. 
-Other types of observations with relatively high spatial and temporal resolution, such as satellite radiance observations, measure only indirect quantities and are more difficult to incorporate to DA systems.
+Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products. 
+Radiance observations particularly, are indirect measurements and are more difficult to incorporate to DA systems as they require complex forward models.
+
 The assimilation of conventional observations from surface stations have an important role at improving the initial conditions that lead to better forecasts. 
 For example, @ha2014 showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
-[Agregar algún ejemplo local que no sea el RRA]
 
 Given the sparsity of the local conventional observing network it is important to explore the potential benefits of including additional sources of observations at a regional scale. 
-Recently Zhu et al studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
-Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the AIRS sensor in a regional application (cita: RRA, otros de Maru?).
-Moreover Bao et al studied the impact on temperature and humidity forecast over western USA while assimilating directly microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels due the microwave observations but an opposite impact for infrared data.
-Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al.
-2016).
+Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
+Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
+@miyoshi2012a assimilated observations, atmospheric temperature and humidity profile data from the AIRS and found improvement on the tracking of two typhoons. 
+@dillon2019 also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation. 
+Moreover @bao2015 studied the impact on temperature and humidity forecast over western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts [@gao2015].
 
 Having analyses with higher temporal resolution than those available from global forecasting systems (e.g.
-at hourly or sub-hourly) and using an increased number of regional observations that are not included in global models are some of the advantages of assimilation at these scales.
-Previous work has shown promising results on mesoscale DA in the region (goncalvesdegoncalves2015, Dillon 2019, Dillon 2020 submitted) using different approaches such as LETKF; (Hunt et al, 2007) and 3D-Var ().
+at hourly or sub-hourly) and using an increased number of regional observations that are not assimilated in global DA systems are some of the advantages of regional DA.
+Previous work has shown promising results on mesoscale DA in the region using different approaches such as LETKF; [@dillon2016; @dillonInreview] and 3D-Var [@goncalvesdegoncalves2015]. 
 In particular, the use of ensemble-based methods have shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances.
-Moreover, the use of ensemble models also provides valuable information about the uncertainties in the forecast.
-Since 2016, the Argentinian National Meteorological Service (ANMS) has been running a convective-scale downscaling forecast (skabar2018).
-Recently a rapid-refresh ensemble-based DA system has been experimentally implemented during the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) Field Campaign [central Argentina, November-December 2018, Nesbitt 2021] with promising results [Dillon 2020 submitted].
-This system assimilated data from high-resolution surface networks, AMDAR data from local aircraft flights, soundings, AIRS retrievals, high-resolution GOES-16 wind estimates, and local radar data every hour with grid spacing of 10 km.
+Moreover, the use of ensembles also provides valuable information about the uncertainties in the forecast.
+Recently a rapid-refresh ensemble-based DA system has been experimentally implemented during the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) Field Campaign (central Argentina, November-December 2018, @nesbitt2021) with promising results [@dillonInreview].
+This system assimilated data from high-resolution surface networks, AMDAR data from local aircraft flights, soundings, AIRS retrievals, high-resolution GOES-16 wind estimates, and local radar data every hour with a grid spacing of 10 km.
 The forecast initialized from the analyses shows an overall similar performance to cold-start initializations from the Global Forecast System analysis, and even a positive impact in some cases.
 
-One important challenge of the DA method is to assess the performance of a regional data assimilation system at representing the true atmospheric state and the skill of the forecasts initialized using the analysis as initial condition.
+One important challenge of regional data assimilation systems is the assessment of their representation of the true atmospheric state and the skill of the forecasts initialized using their analysis as initial conditions.
 In this sense, the observations and measurements taken during the intense observing period of RELAMPAGO are an invaluable resource to validate a regional data assimilation system.
@@
 The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
-The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during 22-23 November 2018 during the IOP period of the RELAMPAGO field campaign.
+The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system. 
+The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during 22-23 November 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
 On 22 November a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
-The MCS lasts for 42 hours traveling approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 18 UTC 23 November.
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged the MSC.
-The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system.
-A multi-model ensemble is used in the data assimilation experiments allowing an evaluation of the performance of different model configurations in this particular case.
-
-The rest of the paper is organized as follows: a description of the assimilation systems and assimilated observations is provided in Section XX, results are discussed in …. …..
+The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours. 
+To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MSC. 
+The paper is organized as follows.
+The data assimilation system, the experimental design and observations used is presented in section 2.
+Results are discussed in section 3 and finally, conclusions are summarized in section 4.
 
@@
 
-The model simulations for the case of study were performed using the version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al., n.d.) model.
+The model simulations for the case study were performed using the version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2018) model.
 The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. 
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25° resolution) [] every 6 hours.
-The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulation period.
-
-The Local Ensemble Transform Kalman Filter (LETKF; ) implementation (V1.3) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Hu et al.
-2018) is chosen to generate the analysis.
-A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within +- 30 minutes of the analysis time are assimilated.
+The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25° horizontal resolution and 6 hour temporal resolution, @cisl_rda_ds084.1). 
+The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.
+
+The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @hu2018).
+A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
 Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. 
-For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al.
-2006) was used as an observation operator to calculate model-simulated brightness temperatures.
-
-To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localizations radius of  0.4 (in log pressure coordinates) as in [RRA] for all types of observations.
+For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) was used as an observation operator to calculate model-simulated brightness temperatures.
+
+To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in [RRA] for all types of observations.
 A relaxation-to-prior inflation [Whitaker and Hamill] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
 
-We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis and random perturbations are used to generate the initial ensemble perturbations.
-The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) with a smooth time evolution as in \citep{Necker2020,Maldonado2021}.
+We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis.
+Since the publicly available GFS ensemble members change at each cycle (alguna fuente) we used random perturbations to generate the initial ensemble perturbations.
+This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
+The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) with a smooth time evolution as in [@necker2020, @maldonado2021].
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
@@
 In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system.
-We use 9 different model configuration consisting on the combinations of 3 moist convection schemes  (Kain–Fritsch [], Grell–Freitas [], and Betts–Miller–Janjic []) and 3 planetary boundary layer schemes (Yonsei University Scheme [], Mellor–Yamada–Janjic Scheme [], and Mellor–Yamada Nakanishi Niino []).
-All ensemble members used the same land-surface model (Noah-MP, ) and microphysics (WRF single-moment 6–class squeme []), and radiation processes (RRTMG shortwave and longwave scheme []) parameterizations.
+We use 9 different model configurations consisting on the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
+All ensemble members used the same land-surface model (Noah-MP, @chen2001) and microphysics (WRF single-moment 6–class squeme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
 
@@
 
-## Results
+# Results
 
@@
                         type.obs == 287, 287)] %>% 
+  .[, sd.y := mean(sd.y), by = .(lon.box, lat.box, var)] %>% 
   .[, var := factor(var, levels = c("v", "t"))] %>% 
@@
                                                 barwidth = 20,
-                                                barheight = 0.8)) +
-  facet_grid(var~type.obs, 
-             labeller = labeller(type.obs = c("281" = "ADPSFC",
-                                              "287" = "ADPSFC\n no pressure",
-                                              "181" = "ADPSFC",
-                                              "187" = "ADPSFC\n no pressure"),
-                                 var = c("v" = "Wind",
+                                                barheight = 0.5)) +
+  facet_wrap(~var, labeller = labeller(var = c("v" = "Wind",
                                                "t" = "Temperature"))) +
+  # facet_grid(var~type.obs, 
+  #            labeller = labeller(type.obs = c("281" = "ADPSFC",
+  #                                             "287" = "ADPSFC\n no pressure",
+  #                                             "181" = "ADPSFC",
+  #                                             "187" = "ADPSFC\n no pressure"),
+  #                                var = c("v" = "Wind",
+  #                                        "t" = "Temperature"))) +
   tag_facets() +
   labs(fill = "") +
-  theme_minimal() +
+  theme_minimal(base_size = 6) +
   theme(legend.position = "bottom",
@@
 
-```{r TQ-diff, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of temperature (a, b, and c, in K) and specific humidity (d, e, and f in gkg-1).", fig.scap=c("Temperature (°C).", "Specific Humidity (g/kg)."), fig.height=3.5}
+```{r TQ-diff, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of temperature (a, b, and c, in K) and specific humidity (d, e, and f in gkg-1).", fig.env="figure*", fig.height=3.5}
 files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
@@
                     breaks = c(seq(-2, 2, 0.2), Inf)) +
+  geom_contour2(aes(z = value), color = "white", size = 0.1,
+                    breaks = c(seq(-2, 2, 0.2), Inf)) +
   scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
+                                                            barheight = 0.5)) +
   labs(fill = "Temperature (K)",
@@
                     breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
+  geom_contour2(aes(z = value), color = "white", size = 0.1,
+                    breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
   scale_fill_divergent_discretised(limits = c(-1.6, 1.6),
                                    guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
+                                                            barheight = 0.5)) +
 
@@
 
-```{r UV-diff, fig.cap="Difference between experiments for the average vertical profile of wind over time (shaded) and U and V (only northerly wind) profiles for the experiment with more observations (contours).", fig.height=3.5}
+```{r UV-diff, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of u wind (a, b, and c, in ms-1) and v wind (d, e, and f in ms-1).
+In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f).", fig.height=3.5, fig.env="figure*"}
 perfiles %>% 
@@
   scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
+                                                            barheight = 0.5)) +
+  geom_contour2(aes(z = value), color = "white", size = 0.1,
+                    breaks = c(seq(-2, 2, 0.2), Inf)) +
   geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>%
@@
                     breaks = c(-Inf, seq(-1.6, 1.6, 0.2), Inf)) +
+  geom_contour2(aes(z = value), color = "white", size = 0.1,
+                    breaks = c(-Inf, seq(-1.6, 1.6, 0.2), Inf)) +
   scale_fill_divergent_discretised(limits = c(-1.6, 1.6),
                                    guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.8)) +
+                                                            barheight = 0.5)) +
   geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
@@
 ```{r summary-fields, fig.cap="a) Precipitable water (shaded) and northerly wind (contours, m/s), b) Average potential temperature for the PBL and c) Maximum CAPE and ~0-6 km wind shear over 15 m/s and 30 m/s for each experiment.
-All fields correspond to 00 UTC Nov, 22.", fig.height=10, fig.width=9}
+All fields correspond to 00 UTC Nov, 22.", fig.height=10, fig.width=9, fig.env="figure*"}
 
@@
                                                guide = guide_colorsteps(barwidth = 25,
-                                                                        barheight = 0.8)) +
+                                                                        barheight = 0.5)) +
   geom_contour2(aes(z = v_mean, size = stat(level)),
@@
   labs(fill = "Precipitable\nwater (Kg/m²)", size = "V (m/s)") +
-  cordillera +
+  #cordillera +
   geom_mapa() +
@@
                        guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
+                                                barheight = 0.5)) +
   labs(fill = "Potential\ntemperature (K)") +
-  cordillera +
+  #cordillera +
   geom_mapa() +
@@
                        guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
+                                                barheight = 0.5)) +
   # geom_contour2(aes(z = cortante),
@@
   labs(fill = "MCAPE (J/Kg)", size = "Wind\nshear (m/s)") +
-  cordillera +
+  #cordillera +
   geom_mapa() +
@@
 
-```{r q_vert, fig.cap="Specific humidity (g/kg, shaded) and meridional wind (m/s, contours) average over the inner domain for each latitude for 00 UTC Nov, 22."}
-
-coord <- mucape[variable == "mcape", c("south_north", "lat")] %>% 
-  .[, .(lat = mean(lat)), by = south_north]
-
-ana[west_east %between% c(70, 180) & bottom_top %between% c(1, 25)] %>% 
-  .[, .(q_mean = mean(q),
-        v_mean = mean(v)), by = .(bottom_top, south_north, date, exp)] %>% 
-  coord[., on = "south_north"] %>% 
-  ggplot(aes(lat, bottom_top)) +
-  geom_contour_fill(aes(z = q_mean*1000, fill = stat(level_d))) +
-  scale_fill_distiller(super = ScaleDiscretised,
-                       palette = "BrBG", direction = 1,
-                       # limits = c(1, 19),
-                       guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.8)) +
-  geom_contour2(aes(z = v_mean, linetype = factor(-sign(..level..))), 
-                color = "grey30", size = 0.3, binwidth = 2) +
-  geom_text_contour(aes(z = v_mean), color = "grey30", 
-                    size = 3, stroke = 0.05, breaks = seq(-10, 6, 2),
-                    label.placement = label_placement_n(1)) +
-  scale_linetype(guide = NULL) +
-  scale_x_latitude(breaks = seq(-40, -20, 5)) +
-  facet_wrap( ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
-                                                 E6 = "SATWND", E8 = "RAD"))) +
-  labs(y =  "Sigma levels", fill = "Specific\nhumidity (g/Kg)") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        legend.box.spacing = margin(0, 0, 0, 0),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3)) 
-```
-
 ### Verification
@@
   ggplot(aes(end_date, lat)) +
-  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(1, 13, 1), Inf)) +
-  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 13, 1)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(0.5, 13, 0.5), Inf)) +
+  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(0.5, 13, 0.5)) +
   # scale_fill_viridis_c(option = "A",
@@
                        direction = 1,
+                        labels = function(x) JumpBy(x, 2, fill = ""),
                        super = ScaleDiscretised,
@@
                                                 barwidth = 25,
-                                                barheight = 0.8)) +
+                                                barheight = 0.5)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
@@
   ggplot(aes(date, lat)) +
-  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(1, 13, 1), Inf)) +
-  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 13, 1)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(0.5, 13, 0.5), Inf)) +
+  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(0.5, 13, 0.5)) +
   # scale_fill_viridis_c(option = "A",
@@
                        direction = 1,
+                       labels = function(x) JumpBy(x, 2, fill = ""),
                        super = ScaleDiscretised,
@@
 
-## Conclusions
+# Conclusions
+
+# References

### 2021-06-25 analysis/chapters/paper_ana_20181122.Rmd
Figura del dominio aproximadamente lista.

@@
 output: 
-  rticles::elsevier_article: 
-    keep_tex: true
+  bookdown::pdf_book: 
+      base_format: rticles::elsevier_article
 abstract: |
@@
 journal: Atmospheric Research
-layout: authoryear,preprint,review,12pt
-#layout: final,5p,times,twocolumn,authoryear
+#layout: authoryear,preprint,review,12pt
+layout: final,5p,times,twocolumn,authoryear
 address:
@@
 numbersections: true
+# header-includes: 
+#  - \usepackage{gensymb}
+#  - \usepackage{subfig}
+#  - \usepackage[inline]{showlabels}
+#  - \usepackage{chngcntr}
+#  - \usepackage{natbib}
+#  - \usepackage{lineno}
+#  - \linenumbers 
 ---
@@
 knitr::opts_chunk$set(
-	eval = FALSE,
+	echo = FALSE,
 	fig.path = "../figures/",
@@
 	dev = "png",
-	dpi = 300,
-	include = FALSE
+	dpi = 300
 )
@@
 library(tagger)
+library(cowplot)
 
@@
 
-To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in [RRA] for all types of observations.
-A relaxation-to-prior inflation [Whitaker and Hamill] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
+To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillonInreview for all types of observations.
+A relaxation-to-prior inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
 
 We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis.
-Since the publicly available GFS ensemble members change at each cycle (alguna fuente) we used random perturbations to generate the initial ensemble perturbations.
+Since the publicly available GFS ensemble members change at each cycle **(alguna fuente)** we used random perturbations to generate the initial ensemble perturbations.
 This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
-The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) with a smooth time evolution as in [@necker2020, @maldonado2021].
+The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5° horizontal grid spacing with a smooth time evolution as in [@necker2020; @maldonado2021].
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
@@
 
-```{r dominio, fig.cap="a) The horizontal domain used on the simulations (black box) and the inner domain used for the experiment comparison (red box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles) .
-b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP07, orange triangles are radiosondes launched during IOP08 and purple square are radiosondes launched outside the IOPs.
-The topography in meters is also shown (filled contours).", out.width="110%"}
+(ref:dominio) a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (green box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOPs.
+The topography in meters is also shown (filled contours).
+
+```{r dominio, fig.cap="(ref:dominio)", out.width="100%"}
 
@@
 lista_sondeos <- fread(here("analysis/data/derived_data/lista_sondeos.csv")) %>% 
-  .[, periodo := fcase(nominal_launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")),  "IOP07", 
-                       nominal_launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")),  "IOP08", 
+  .[, periodo := fcase(nominal_launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")),  "IOP 7", 
+                       nominal_launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")),  "IOP 8", 
                        default = "Others")
@@
 
-dominio %>% 
+g1 <- dominio %>% 
   ggplot(aes(x, y)) +
@@
                       breaks = seq(0, 6000, 1000),
-                      guide = guide_colorstrip(barwidth = 20,
-                                               barheight = 0.5)) +
+                      guide = NULL) +
   geom_mapa() +
@@
              size = 1, alpha = 0.8) + 
-  scale_shape_manual(name = "a) Surface stations",  
+  geom_rect(aes(xmin = -66.5, xmax = -61.5, ymin = -35.5, ymax = -29), 
+            color = "#7CAE00", alpha = 0) +
+  scale_shape_manual(name = NULL,  
                      breaks = c("Sfc - Non-official", "Sfc - Official"),
@@
                      guide = guide_legend(override.aes = list(size = 2))) +
-  scale_color_manual(name =  "a) Surface stations", 
+  scale_color_manual(name =  NULL, 
                      values = c("Sfc - Non-official" = "#00695c", 
@@
   geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
-  theme_minimal(base_size = 9) +
-  theme(legend.position = "bottom", 
-        legend.box = "vertical") +
-  # ,
-  # legend.margin = margin(),
-  # legend.box.spacing = margin(0, 0, 0, 0)) +
-  # 
-  dominio %>% 
+  theme_minimal(base_size = 10) +
+  theme(legend.box = "horizontal",
+        legend.position = c(0.1, 0.95),
+        legend.background = element_rect(fill = "white", color = "white"))
+
+  temp <- dominio %>% 
   ggplot(aes(x, y)) +
@@
   coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
+  # geom_jitter(data = lista_sondeos, aes(lon, lat, 
+  #                                       color = periodo,
+  #                                       shape = periodo), 
+  #             alpha = 0.5, size = 1.5, width = 0.03, height = 0.03) +
+  # scale_color_brewer(palette = "Dark2",
+  #                    guide = guide_legend(override.aes = list(size = 2,
+  #                                                             alpha = 1))) +
+  # labs(color = "b) Soundings", shape = "b) Soundings") +
+  theme_minimal(base_size = 10) +
+  theme(legend.position = "bottom")
+
+  legend <- get_legend(temp)
+
+  g2 <- dominio %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
+                    breaks = seq(0, 6000, 500)) +
+  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
+                      name = NULL,
+                      breaks = seq(0, 6000, 1000),
+                      guide = NULL) +
+  geom_mapa() +
+  coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
   geom_jitter(data = lista_sondeos, aes(lon, lat, 
@@
                                                               alpha = 1))) +
-  labs(color = "b) Soundings", shape = "b) Soundings") +
-  theme_minimal(base_size = 9) +
-  theme(legend.position = "bottom", 
-        legend.box = "vertical") +
-  # ,
-  #         legend.margin = margin(),
-  #         legend.box.spacing = margin(0, 0, 0, 0)) +
-  plot_layout(ncol = 2, widths = c(4, 1), guides = "collect") +
-  plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
-  theme(legend.position = 'bottom', legend.box = "vertical")
+  labs(color = NULL, shape = NULL) +
+  theme_minimal(base_size = 10) +
+  theme(legend.box = "horizontal",
+        legend.position = c(0.16, 0.92),
+        legend.background = element_rect(fill = "white", color = "white"))
+
+  ggdraw(plot_grid(plot_grid(g1, g2, ncol = 2, rel_widths = c(1, 0.5), 
+                             labels = c("a)", "b)"), label_size = 10), legend, 
+                   ncol = 1, rel_heights = c(1, 0.05)))
+
 ```
@@
 
-```{r dist_obs, fig.cap="a) Horizontal spatial distribution of the mean available observations per analysis cycle for CONV, AUT and SATWND.
+```{r dist_obs, eval=FALSE, fig.cap="a) Horizontal spatial distribution of the mean available observations per analysis cycle for CONV, AUT and SATWND.
 b) Number of radiance observations assimilated at each analysis cycle and pressure level."}
@@
 
-```{r count_obs, fig.cap="a) Number of available observations per cycle during the experiment. a) Vertical spatial distribution of the mean available observations per analysis cycle."}
+```{r count_obs, eval=FALSE, fig.cap="a) Number of available observations per cycle during the experiment. a) Vertical spatial distribution of the mean available observations per analysis cycle."}
 rad2 <- rad[, .(press, iuse, error, exp, date)] %>% 
@@
 
-```{r rcrv-sfc, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree"}
+```{r rcrv-sfc, eval=FALSE, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree"}
 
@@
 
-```{r rcrv-profile, fig.cap="Profiles of the mean and standard deviation of RCRV for upper-air observations.", fig.subcap=c("Sounding and aircraft observations.", "Satellite-derived wind.", "Radiances observations.")}
+```{r rcrv-profile, eval=FALSE, fig.cap="Profiles of the mean and standard deviation of RCRV for upper-air observations.", fig.subcap=c("Sounding and aircraft observations.", "Satellite-derived wind.", "Radiances observations.")}
 
@@
 
-```{r TQ-diff, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of temperature (a, b, and c, in K) and specific humidity (d, e, and f in gkg-1).", fig.env="figure*", fig.height=3.5}
+```{r TQ-diff, eval=FALSE, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of temperature (a, b, and c, in K) and specific humidity (d, e, and f in gkg-1).", fig.env="figure*", fig.height=3.5}
 files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
@@
 
-```{r UV-diff, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of u wind (a, b, and c, in ms-1) and v wind (d, e, and f in ms-1).
+```{r UV-diff, eval=FALSE, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of u wind (a, b, and c, in ms-1) and v wind (d, e, and f in ms-1).
 In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f).", fig.height=3.5, fig.env="figure*"}
@@
 
-```{r summary-fields, fig.cap="a) Precipitable water (shaded) and northerly wind (contours, m/s), b) Average potential temperature for the PBL and c) Maximum CAPE and ~0-6 km wind shear over 15 m/s and 30 m/s for each experiment.
+```{r summary-fields, eval=FALSE, fig.cap="a) Precipitable water (shaded) and northerly wind (contours, m/s), b) Average potential temperature for the PBL and c) Maximum CAPE and ~0-6 km wind shear over 15 m/s and 30 m/s for each experiment.
 All fields correspond to 00 UTC Nov, 22.", fig.height=10, fig.width=9, fig.env="figure*"}
@@
 
-```{r pp-hov, fig.cap="Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right, for the ensemble mean) over the inner domain."}
+```{r pp-hov, eval=FALSE, fig.cap="Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right, for the ensemble mean) over the inner domain."}
 
@@
 
-```{r fss, fig.cap="FSS calculated over a 6-hours moving window for different thresholds and scales."}
+```{r fss, eval=FALSE, fig.cap="FSS calculated over a 6-hours moving window for different thresholds and scales."}
 files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2-8].csv")
@@
 
-```{r dbz-mean, eval=FALSE, fig.cap="Observed maximum reflectivity in the column (upper row) and ensemble mean maximum reflectivity for CONV and RAD at different times (bottom two rows).", include=FALSE}
+```{r dbz-mean, eval=FALSE, eval=FALSE, fig.cap="Observed maximum reflectivity in the column (upper row) and ensemble mean maximum reflectivity for CONV and RAD at different times (bottom two rows).", include=FALSE}
 proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"
@@
 
-```{r soundings, fig.cap="RMSE and Bias of temperature, dew point temperature and wind components calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8."}
+```{r soundings, eval=FALSE, fig.cap="RMSE and Bias of temperature, dew point temperature and wind components calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8."}
 files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos/ANA/sondeo_E[2568]_ana*")

### 2021-06-27 analysis/chapters/paper_ana_20181122.Rmd
Avance, las figuras siguen dando vueltas por donde quieren

@@
 	echo = FALSE,
+	cache = TRUE,
 	fig.path = "../figures/",
@@
   .[!str_detect(stationID, "SMN")] %>%
+  .[!str_detect(stationID, "^SC")] %>%
   .[!(stationID %in% oficiales$stationID)] %>% 
@@
 
-### Observations
-
-#### Conventional and Satellite-derived wind
-
-Conventional and satellite-derived wind observations are part of the Global Data Assimilation System (GDAS) [] except for Automatic Surface Weather Stations (ASWS).
-These observations come from public and private automatic station networks over Southern South America and are part of RELAMPAGO Data Set repository (García, F., et al.
-2019).
-
-We used conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP (Keyser 2013) that consist of surface observations (land and sea) and upper-air observations from radiosondes and aircrafts (Figure \@ref(fig:dominio)).
-The frequency of these observations varied between 1 hour for surface stations and 24 hours for most radiosondes in the region.
+## Observations
+
+### Conventional and Satellite-derived wind
+
+The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS) **[ver cita]**.
+We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircrafts.
+The orange triangles in Figure \@ref(fig:dominio) a), are the surface stations and ship locations included in this experiment.
+The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.
 
 Satellite-derived wind observations also come from GDAS and includes information from GOES-16 (estimated from the visible, infrared and water vapour channels) and METEOSAT 8 and 11 (visible and water vapour channels).
-Due to the domain covered by each of these satellites, GOES-16 is the primary source of this type of observation (99 % of this type of observation).
-
-Table 2 lists all the observation types (i.e., pressure, temperature, specific humidity and wind) available for each source, together with their associated errors.
+Due to the domain covered by each of these satellites, GOES-16 is the primary source of this type of observation (99 % of observation).
+
+We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of  various public and private automatic station networks over Southern South America and are part of RELAMPAGO Data Set repository [@garcia2019].
+These stations are indicated as green squares in Figure \@ref(fig:dominio) a).
+They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases.
+The variables measured by the stations varied between networks; all stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.
+
+Table \@ref(tab:table-obs) lists all the observation types (i.e., pressure, temperature, specific humidity and wind) available for each source, together with their associated errors.
 The observation errors were specified following the GSI default configuration.
 In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind).
-In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and model-simulated observation based on the first-guess)  with a predefined threshold (also included in Table 2).
-
-#### Satellite radiances
-
-Satellite radiances available through the GDAS datastream which consist of infrared and microwave observations are used in this paper.
-This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 3).
-Since the regional domain is located in mid-latitudes, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
-For this reason, radiance observations were not available in all analysis cycles and the number of observations varied significantly among cycles.
-This is evident in Figure 2 b) that shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
-In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in the period.
-The vertical location of each radiance was estimated as the model level at which its weighting function was maximized as calculated by the CRTM.
-The multispectral sensors have a good vertical distribution and are able to sense throughout the entire atmosphere.
-The channels whose maximum weighting function is in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.
+In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \@ref(tab:table-obs)).
+
+```{r table-obs}
+fread(here("analysis/data/derived_data/table1.csv")) %>% 
+  # str()
+  kbl(caption = "Observation errors and gross check thresholds for different observation platforms and types.",
+      col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
+      booktabs = TRUE,
+      escape = FALSE) %>% 
+  kable_classic_2(full_width = FALSE) %>% 
+  kable_styling(font_size = 6,
+                position = "center") %>% 
+  column_spec(1, width = "3.5em") %>% 
+  column_spec(2, width = "4.5em") %>% 
+  column_spec(3, width = "5em") %>% 
+  column_spec(4:5, width = "7em") %>% 
+  collapse_rows(columns = 1:2, valign = "middle", latex_hline = "major") %>% 
+  footnote(symbol = c("Observation error varied with height",
+                      "Observations over 600 hPa are rejected",
+                      "Observation error depends on the report type"),
+           symbol_manual = c("*", "**", "+"))
+```
+
+### Satellite radiances {#sat}
+
+Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study.
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \@ref(tab:table-rad)).
+Since the regional domain is located in the mid-latitudes and all satellite platforms are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
+For this reason, the number of satellite observations varied significantly among cycles.
+This is evident in Figure \@ref(fig:obs-cycle) a) which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
+In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiments.
+The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM.
+The multispectral sensors have a good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
+The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.
 
@@
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-Firstly a spatial data thinning is applied using a 60 km grid following [Singh et al 2016, Jones et al 2013, Lin et al 2017], where the observations to be assimilated are chosen based on their distance to the model grid points, to the observation quality (based on available data quality predictors) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
-Also, observations over the sea or sea-ice surface are preferred from those over land or snow [cita a la guía avanzada].
+Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, no the observation quality (based on available data quality predictors) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
+Also, observations over the sea are preferred from those over land or snow [@hu2018].
 
 The thinned observations were then bias corrected.
-The bias correction (BC) has an air-mass dependent and an angle-dependent component (Zhu 2014) and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$.
+The bias correction (BC) has an air-mass dependent and an angle-dependent component [@zhu2014] and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$.
 Then, the bias corrected brightness temperature can be obtained as:
@@
 
-In GSI $p_0 = 1$ a constant offset bias correction term.
-The following predictors are:  the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight , and the emissivity sensitivity.
-Scan angle dependent bias is modeled as a 4th-order polynomial [Zhu 2014].
+GSI has a constant offset bias correction term ($p_0 = 1$).
+The following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
+Scan angle-dependent bias is modelled as a 4th-order polynomial [@zhu2014].
 
 The $\beta_i$ coefficients are usually trained to achieve the best fit between the simulation and the observations.
-In the GSI system these coefficients are trained using a variational estimation method (Zhu 2014) which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
-The coefficients initialized from the 18 UTC Nov, 11 2018 GFS coefficients were trained during one week of continuous assimilation and using the same setup as in the experiments described in the next section.
-Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments to the predictor coefficients during each cycle.
-
-During the quality control step, GSI rejects all observations if the innovation is greater than a predefined threshold for each channel and sensor.
-For microwave sensors, the quality control routines also use the scattering and Liquid Water Path (LWP) indexes calculated from brightness temperature channel differences (Weston et al., 2019; Zhu et al., 2016).
-The quality control routine for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
-It also checks for cloud-affected observations using the transmittance profile calculated within the CRTM algorithms and performs an emissivity check for observations over land.
-In the experiment, only clear-sky observations are used as defined by the internal subroutines of GSI that use the LWP for microwave and the transmittance profile for infrared observations that could be potentially affected by clouds.
-
-```{r}
+In the GSI system these coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
+The coefficients initialized from the 18 UTC Nov, 11 2018 GFS coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
+Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments  in the estimated coefficients at each cycle.
+
+In our experiments, only clear-sky observations are used.
+For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes[@weston2019; @zhu2016].
+For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
+Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
+It also performs an emissivity check for observations over land for both infrared and microwave radiances.
+
+```{r eval=FALSE, include=FALSE}
 files <- Sys.glob(here("analysis/data/derived_data/count_obs_E*"))[c(1, 3, 4)]
@@
 
-```{r}
+```{r table-rad}
 fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
@@
            plataforma = toupper(plataforma),
-           prop = scales::percent(prop/100, accuracy = 0.01, trim = TRUE))] %>% 
+           prop = paste0(prop, " \\", "%"))] %>% 
   .[order(.[,'sensor']), ] %>% 
   .[] %>% 
-  kbl(booktabs = T, col.names = c("Sensor", "Plataform", "Assimilated\nchannels", "Percentage over total"),
+  kbl(booktabs = TRUE,
+      escape = FALSE,
+      col.names = linebreak(c("Sensor", "Plataform", "Assimilated\nchannels", "Percentage\nover total")),
       caption = "List of the available sensors in several platforms, the number of accepted channels and the percentage of assimilated observations calculated over all radiance observations.") %>% 
+  kable_styling(font_size = 7) %>% 
   collapse_rows(1,valign = "top") %>% 
   # column_spec(1, width = "3cm", latex_valign = "p") %>% 
-  kable_classic_2(full_width = F)
-```
-
-```{r obs}
+  kable_classic_2(full_width = TRUE)
+```
+
+```{r table-exp}
 tribble(
@@
 
-```{r dist_obs, eval=FALSE, fig.cap="a) Horizontal spatial distribution of the mean available observations per analysis cycle for CONV, AUT and SATWND.
-b) Number of radiance observations assimilated at each analysis cycle and pressure level."}
+(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AUT c) SATWND and d) RAD experiments calculate over 2.5º boxes.
+
+```{r obs-horizontal, fig.cap="(ref:obs-horizontal)", out.width="100%"}
 files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]
@@
 
-rbind(conv, aut, satwnd) %>% 
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")
+
+files <- files[str_detect(files, "conv")]
+
+rad_conv <- read_diag_conv(files, exp = "E8", member = "000")
+
+rad_conv[, bufr_code := fcase(type %in% c(181, 187, 281, 287), "ADPSFC",
+                            type %in% c(120, 220, 221), "ADPUPA",
+                            type %in% c(130, 131, 133, 230, 231, 233), "AIRCFT",
+                            type %in% c(290), "ASCATW",
+                            type %in% c(180, 280, 183, 283, 184, 284), "SFCSHP",
+                            type %in% c(240:260), "SATWND")] %>% 
+  .[, ":="(lon.box = cut_round(lon, breaks = seq(284, 309, 2.5)),
+           lat.box = cut_round(lat, breaks = seq(-42, -17, 2.5)))]
+
+rad <- rad[, .(press, iuse, error, exp, date, lon.box, lat.box)] %>% 
+  setnames(c("press", "iuse", "error"), c("pressure", "usage.flag", "rerr"))
+
+rbind(conv, aut, satwnd, rad_conv) %>%
+  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
+  rbind(., rad) %>% 
   .[usage.flag == 1 & rerr != 1.0e+10] %>% 
@@
                                                E6 = "SATWND", E8 = "RAD"))) +
+  tagger::tag_facets() +
   labs(fill = "Mean number of\nobs per cycle") +
-  theme_minimal( base_size = 6) +
+  theme_minimal(base_size = 8) +
   theme(legend.position = "bottom",
-        plot.margin = unit(c(0, 0, 0, 0), "cm")) +
-
-rad %>% 
-  .[qc == 0 & iuse == 1] %>% 
-  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
-  .[, .(omb = mean(tbc),
-        n = .N), by = .(date, lev, exp)] %>% 
-  .[lev != 1025] %>% 
-  ggplot(aes(date, fct_rev(lev))) +
-  geom_tile(aes(fill = n)) +
-  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
-                       guide = guide_colourbar(barheight = 8,
-                                               barwidth = 0.5)) +
-
-  scale_date() +
-   facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E8 = "RAD"))) +
-  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
-  theme_minimal(base_size = 6) +
-  theme(aspect.ratio = 1/1.5) +
-
-  plot_layout(ncol = 2, widths = c(1.6, 1)) + plot_annotation(tag_levels = 'a', tag_suffix = ")") &
-  theme(plot.tag = element_text(size = 6))
-```
-
-```{r count_obs, eval=FALSE, fig.cap="a) Number of available observations per cycle during the experiment. a) Vertical spatial distribution of the mean available observations per analysis cycle."}
-rad2 <- rad[, .(press, iuse, error, exp, date)] %>% 
-  setnames(c("press", "iuse", "error"), c("pressure", "usage.flag", "rerr"))
-
-rbind(conv, aut, satwnd) %>%
-  .[, .(pressure, usage.flag, rerr, exp, date)] %>% 
-  rbind(., rad2) %>% 
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+```
+
+(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle and divided into 50 hPa-depth vertical layers for the CONV (blue line), AUT (light blue line), SATWND (orange line) and RAD (red line) experiments.
+
+```{r obs-cycle, fig.cap="(ref:obs-cycle)", fig.env="figure*"}
+
+rbind(conv, aut, satwnd, rad_conv) %>%
+  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
+  rbind(., rad) %>% 
   .[usage.flag == 1 & rerr != 1.0e+10] %>% 
@@
 
-  rbind(conv, aut, satwnd) %>%
-  .[, .(pressure, usage.flag, rerr, exp, date)] %>% 
-  rbind(., rad2) %>% 
+rbind(conv, aut, satwnd, rad_conv) %>%
+  .[, .(pressure, usage.flag, rerr, exp, date, lon.box, lat.box)] %>% 
+  rbind(., rad) %>%  
   .[usage.flag == 1 & rerr != 1.0e+10] %>% 
@@
 
-#### Validation dataset
-
-In order to evaluate the performance of the ensemble based data assimilation system presented in this article, we use part of the following observations datasets:
-
-* The Multi-Network Composite Highest Resolution Radiosonde Data [] from the RELAMPAGO field campaign database consists of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
-Only the radiosondes that did not enter the assimilation system were used to perform the validation.
-The experiment period covers the Intensive Observation Periods (IOP) 7 and 8 during which XXXX radiosondes were launched in a small area Figure 1 b).
-
-* The Satellite precipitation estimation IMERG half-hourly Final  Run of  0.01◦ resolution (Huffman et al.
-2018) was used as a reference state to validate the skill of the short-range forecasts to represent the precipitation over the domain.
+### Validation dataset
+
+To evaluate the performance of the ensemble based data assimilation system presented in this article, we use the following observations datasets:
+
+* The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
+Only the measurements/soundings that did not enter the assimilation system were used for validation.
+The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio) b).
+
+* The Satellite precipitation estimation IMERG Final Run with 0.01◦ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill 1-hour forecasts to represent the precipitation over the domain.
 
 * Radar observations are used to perform a qualitative and visual assessment of the convection features.
-The data comes from 9 radars located on the domain and is provided by the recently developed Argentine C-band Doppler dual-polarization weather radar network \citep{deelia2017} with a temporal frequency of 10 minutes.
-For this work only the maximum reflectivity in the column (COLMAX) field close to the analysis time was used.
-
-### Experimental design
+The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes.
+For this work only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
+
+## Experimental design
 
@@
 
-```{r ciclo, fig.cap="Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
+```{r cycle, fig.cap="Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
 The zoomed section shows the assimilation for each hour that is performed within a one-hour centered window and new boundary conditions from GFSD every 6 hours.
-Also the two Intensive Observation Periods (IOP) from the RELAMPAGO Field Campaign are shown."}
+Also the two Intensive Observation Periods (IOP) from the RELAMPAGO Field Campaign are shown.", out.width="100%"}
 knitr::include_graphics(here("analysis", "figures", "analysis_cycle.pdf"))
@@
 
-### Verification methods
+## Verification methods
 
@@
 
-### Ensemble reliability
+## Ensemble reliability
 
@@
 
-### Impacts of assimilated observations
+## Impacts of assimilated observations
 
@@
 
-### Verification
+## Verification
 

### 2021-06-28 analysis/chapters/paper_ana_20181122.Rmd
datos y metodos revisados

@@
 The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. 
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25° horizontal resolution and 6 hour temporal resolution, @cisl_rda_ds084.1). 
+The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal resolution and 6 hour temporal resolution, @cisl_rda_ds084.1). 
 The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.
@@
 This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
-The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5° horizontal grid spacing with a smooth time evolution as in [@necker2020; @maldonado2021].
+The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in [@necker2020; @maldonado2021].
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
@@
 We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircrafts.
-The orange triangles in Figure \@ref(fig:dominio) a), are the surface stations and ship locations included in this experiment.
+The orange triangles in Figure \@ref(fig:dominio)a, are the surface stations and ship locations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.
@@
 We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of  various public and private automatic station networks over Southern South America and are part of RELAMPAGO Data Set repository [@garcia2019].
-These stations are indicated as green squares in Figure \@ref(fig:dominio) a).
+These stations are indicated as green squares in Figure \@ref(fig:dominio)a.
 They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases.
@@
 For this reason, the number of satellite observations varied significantly among cycles.
-This is evident in Figure \@ref(fig:obs-cycle) a) which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
+This is evident in Figure \@ref(fig:obs-cycle)a which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
 In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiments.
@@
 For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
-Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
+Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60$^{\circ}$) to reject channels near the visible range that can be contaminated with reflection.
 It also performs an emissivity check for observations over land for both infrared and microwave radiances.
@@
 
-```{r table-exp}
-tribble(
-  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
-  "PREPBUFR", "x", "x", "x", "x",
-  "AUTSFC", " ", "x", "x", "x",
-  "SATWND", " ", " ", "x", "x",
-  "RAD", " ", " ", " ", "x",
-) %>% 
-  kbl(booktabs = T, col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
-      caption = "Observation types assimilated in each experiment.") %>% 
-  kable_classic_2(full_width = F)
-```
-
-(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AUT c) SATWND and d) RAD experiments calculate over 2.5º boxes.
-
-```{r obs-horizontal, fig.cap="(ref:obs-horizontal)", out.width="100%"}
+(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AUT c) SATWND and d) RAD experiments calculate over 2.5$^{\circ}$ boxes.
+
+```{r obs-horizontal, fig.cap="(ref:obs-horizontal)", out.width="100%", fig.height=7}
 files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]
@@
   scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
-                       guide = guide_colorbar(barwidth = 10,
-                                              barheight = 0.5)) +
+                       guide = guide_colorbar(barwidth = 20,
+                                              barheight = 0.5, 
+                                              title.position = "left", 
+                                              title.vjust = 1,
+                                              frame.colour = "black")) +
   geom_mapa() +
@@
   labs(fill = "Mean number of\nobs per cycle") +
-  theme_minimal(base_size = 8) +
+  theme_minimal(base_size = 13) +
   theme(legend.position = "bottom",
@@
 
-(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle and divided into 50 hPa-depth vertical layers for the CONV (blue line), AUT (light blue line), SATWND (orange line) and RAD (red line) experiments.
-
-```{r obs-cycle, fig.cap="(ref:obs-cycle)", fig.env="figure*"}
+(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle and divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
+
+```{r obs-cycle, fig.cap="(ref:obs-cycle)", fig.env="figure*", fig.height=3.5, fig.width=7}
 
@@
        color = NULL, shape = NULL) +
-  theme_minimal(base_size = 6) +
+  theme_minimal(base_size = 10) +
   theme(legend.position = "bottom", 
@@
        x = "Pressure level (hPa)", y = "Mean number of\nobs per cycle") +
-  theme_minimal(base_size = 6) +
+  theme_minimal(base_size = 10) +
   theme(legend.position = "bottom",
@@
 Only the measurements/soundings that did not enter the assimilation system were used for validation.
-The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio) b).
-
-* The Satellite precipitation estimation IMERG Final Run with 0.01◦ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill 1-hour forecasts to represent the precipitation over the domain.
+The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio)b.
+
+* The Satellite precipitation estimation IMERG Final Run with 0.01$^{\circ}$ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill 1-hour forecasts to represent the precipitation over the domain.
 
@@
 
-To investigate the impact of different observations upon the analysis, four data assimilation experiments were performed using different observation sets (Table \@ref(tab:obs)).
+To investigate the impact of different observations upon the analysis, four data assimilation experiments were performed using different observation sets (Table \@ref(tab:table-exp)).
 The CONV experiment uses only convectional observations from PREPBUFR.
-The orange triangles in Figure \@ref(fig:dominio) a), are the surface stations and ship locations included in this experiment.
-
 In a second experiment, referred to as AUT, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from ASWS.
-The automatic stations are indicated as green squares in Figure \@ref(fig:dominio) a).
-These stations have greater spatial coverage and a higher sampling frequency (10 minutes in most cases).
-
 In the third experiment, referred to as SATWND, we assimilate all the observations of the AUT experiment and the satellite-derived motion winds.
-Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbit satellites as described in Section XX.
-
-The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure 2 a).
-The larger values of assimilated observations over the center of the domain and the first pressure levels correspond to the ASWS observations.
-In Figure 3 a) the number of assimilated observations over time is shown.
-In this figure the peak/maximum at 12 UTC each day for CONV, AUT and SATWND is produced by the operational observations like radiosondes that are launched at this time.
-The strong variability in the number of radiance observations per cycle is also noticeable.
-The vertical distribution of mean number of observations per cycle shows a maximum in low levels due to the ASWS observations and a local maximum around 400-300 hPa in SATWND is due to the satellite-derived winds.
-Above 800 hpa, radiance observations are in the majority.
-
-All the experiments start at 12 UTC Nov 20, 2018 with 6 hours of simulation spin up (without assimilation).
-The assimilation cycles start at 18 UTC 20 Nov, 20 and continue until 12 UTC Nov, 23 (67 hours/assimilation cycles).
-Two independent ensemble forecasts whose validation is discussed in a companion paper or on part II of this paper were initialized from the analysis at 00 and 06 UTC Nov, 22 and continue until 12 UTC Nov, 23 (Figure 3).
-
-```{r cycle, fig.cap="Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
-The zoomed section shows the assimilation for each hour that is performed within a one-hour centered window and new boundary conditions from GFSD every 6 hours.
-Also the two Intensive Observation Periods (IOP) from the RELAMPAGO Field Campaign are shown.", out.width="100%"}
+Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbit satellites as described in Section \@ref(sat).
+
+The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \@ref(fig:obs-horizontal).
+The larger number of assimilated observations over the center and east of the domain correspond to the ASWS observations.
+In this region we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet.
+In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown.
+Local maxima at 12 and 00 UTC  found mainly in CONV is produced by the operational soundings.
+The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
+The maxima after 12 and 00 UTC correspond to the contribution of the multispectral sensors.
+The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b shows a maximum in low levels due to the ASWS observations.
+Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa).
+Above 850 hPa, most of the observations correspond to radiance observations.
+
+All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
+The initial ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov 20 (Figure \@ref(fig:cycle)).
+
+```{r table-exp}
+tribble(
+  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
+  "PREPBUFR", "x", "x", "x", "x",
+  "AUTSFC", " ", "x", "x", "x",
+  "SATWND", " ", " ", "x", "x",
+  "RAD", " ", " ", " ", "x",
+) %>% 
+  kbl(booktabs = TRUE, col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
+      align = "lcccc",
+      caption = "Observation types assimilated in each experiment.") %>% 
+  kable_classic_2(full_width = TRUE) %>% 
+  column_spec(1, width = "5em") %>% 
+  column_spec(2:5, width = "3.5em")
+```
+
+(ref:cycle) Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
+The zoomed section shows the assimilation for each hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
+Also the two Intensive Observation Periods (IOP) missions from the RELAMPAGO Field Campaign are shown.
+
+```{r cycle, fig.cap="(ref:cycle)", out.width="100%"}
+
 knitr::include_graphics(here("analysis", "figures", "analysis_cycle.pdf"))
+
 ```
@@
 A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper.
-These aspects include a validation of how uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations not used in the assimilation.
-To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use  the Reduced Centered Random Variable (RCRV, Candille et al., 2007) which is define as  defined as:
-
-$$RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}$$,
-
-Where $x_o$ is the error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble
-The average of  $RCRV$ computed over all the realizations represents the weighted bias between the ensemble and observation:
+These aspects include a validation of how uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations which are not assimilated .
+To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:
+
+$$RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}$$
+
+where $x_o$ is the error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
+The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
 
@@
 
-The standard deviation of this variable measures the systematic over- or under- dispersion of the ensemble:
+The standard deviation of the $RCRV$ or $sd RCRV$ measures the agreement of the ensemble spread and the observational error with respect to the distance between the ensemble mean and the observations, and then the systematic over- or under- dispersion of the ensemble:
 
@@
 
-Where $M$ is the sample size.
-The second moment of the $RCRV$ measures the agreement of the ensemble spread and the specific observational error with the observed amplitude of the forecast error.
-A reliable system will have no bias, $mean RCRV = 0$, and a standard deviation equal to 1, $sd RCRV = 1$.
+where $M$ is the sample size.
+A reliable system will have no bias ($mean RCRV = 0$), and a standard deviation equal to 1, $sd RCRV = 1$.
 If the ensemble has a positive (negative) bias, $mean RCRV$ will be positive (negative).
@@
 
-Where $O$ and $X$ stand for independent observations and the analysis respectively.
-
-For the comparison of the first-guess’ precipitation with the precipitation estimates we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al.
-2021 and adapted from roberts2008.
-
-$$\mathit{FSS} = 1-\frac{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i)^{2}+\frac{1}{N}\sum_{i=1}^{N} ({P_o}_i)^{2}} $$,
-
-where ${P_x}$ and ${P_o}$ are the probability of occurrence of specified rainfall accumulation of the simulated ensemble and the reference atmospheric state, respectively.
-To evaluate the grid points fraction in which the precipitation exceeds certain thresholds, the probability of occurrences for the whole ensemble was used, avoiding the systematic underestimation of high precipitation values by the ensemble mean.
-The FSS was applied to the 6 hr accumulated precipitation ensemble first-guess (i.e.
-adding the precipitation summed in the 6 previous analysis cycles).
+where $O$ and $X$ stand for independent observations and the analysis respectively.
+
+For the comparison of the first-guess’ precipitation with the precipitation estimates we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following @maldonado2021 and adaptations from @roberts2008.
+
+$$\mathit{FSS} = 1-\frac{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i)^{2}+\frac{1}{N}\sum_{i=1}^{N} ({P_o}_i)^{2}} $$
+
+where ${P_x}$ and is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and ${P_o}$ is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise.
+This probabilistic formulation uses the information from the full ensemble thus avoiding the smoothing effect of the ensemble mean which can particularly affect localized extreme precipitation values.
+The FSS was applied to the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
 
@@
 
-```{r rcrv-sfc, eval=FALSE, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree"}
+(ref:rcrv-sfc) sd RCRV calculated for surface observations of a) wind, and b) temperature, averaged over 2.5º boxes.
+
+```{r rcrv-sfc, fig.cap="(ref:rcrv-sfc)", out.width="100%", fig.height=5}
 
@@
       fread(here("analysis", "data", "derived_data", "rcrv_t_box.csv"))) %>%   
-  # .[var %in% c("u", "v") & type %in% c(281, 287)] %>% 
   .[, type.obs := type] %>% 
@@
                        direction = -1,
-                       guide = guide_colorstrip(inside = TRUE,
+                       guide = guide_colorsteps(inside = TRUE,
                                                 barwidth = 20,
-                                                barheight = 0.5)) +
+                                                barheight = 0.5, 
+                                                title.position = "left", 
+                                                title.vjust = 1,
+                                              frame.colour = "black")) +
   facet_wrap(~var, labeller = labeller(var = c("v" = "Wind",
                                                "t" = "Temperature"))) +
-  # facet_grid(var~type.obs, 
-  #            labeller = labeller(type.obs = c("281" = "ADPSFC",
-  #                                             "287" = "ADPSFC\n no pressure",
-  #                                             "181" = "ADPSFC",
-  #                                             "187" = "ADPSFC\n no pressure"),
-  #                                var = c("v" = "Wind",
-  #                                        "t" = "Temperature"))) +
   tag_facets() +
-  labs(fill = "") +
-  theme_minimal(base_size = 6) +
+  labs(fill = "sd RCRV") +
+  theme_minimal(base_size = 13) +
   theme(legend.position = "bottom",
@@
 
-```{r rcrv-profile, eval=FALSE, fig.cap="Profiles of the mean and standard deviation of RCRV for upper-air observations.", fig.subcap=c("Sounding and aircraft observations.", "Satellite-derived wind.", "Radiances observations.")}
+(ref:rcrv-profile) Vertical profiles of mean RCRV and sd RCRV for  a) and b) sounding and aircraft observations, c) satellite-derived winds, and d) brightness temperature observations.
+
+```{r rcrv-profile, fig.cap="(ref:rcrv-profile)", fig.env="figure*", fig.height=3.5, fig.width=7}
 
@@
       linetype = NULL, color = "Upper air\nobservations") +
-  theme_minimal() +
+    guides(color = guide_legend(
+                 keyheight = 0.4,
+                 default.unit = "cm")) +
+  theme_minimal(base_size = 6) +
+  # theme(legend.box = "vertical",
+  #   aspect.ratio = 4/1) +
     theme(legend.box = "vertical",
-    aspect.ratio = 4/1) +
+        legend.spacing.y = unit(0.1, 'mm')) +
 
@@
   .[!(sat_type %in% c("GOES visible", "EUMETSAT visible") & error.level == 700)] %>% 
+  .[, sat_type := fifelse(sat_type == "EUMETSAT WV deep layer winds", "EUMETSAT WV", sat_type)] %>% 
   melt(measure.vars = c("mean.y", "sd.y")) %>%
@@
       linetype = NULL, color = "Satellite-derived\nwind") +
-  theme_minimal() +
+  guides(color = guide_legend(
+                 keyheight = 0.4,
+                 default.unit = "cm")) +
+  theme_minimal(base_size = 6) +
+  # theme(legend.box = "vertical",
+  #   aspect.ratio = 2/1) +
   theme(legend.box = "vertical",
-    aspect.ratio = 2/1) +
+        legend.spacing.y = unit(0.1, 'mm')) +
 
@@
       linetype = NULL, color = "Radiances\nobservations") +
-  theme_minimal() +
-  theme(#legend.position = "bottom", 
-    legend.box = "vertical",
-    aspect.ratio = 2/1) +
-
-  plot_layout(ncol = 3, widths = c(1, 1, 1), guides = "collect") & theme(legend.box = "vertical")
+    guides(color = guide_legend(
+                 keyheight = 0.4,
+                 default.unit = "cm")) +
+  theme_minimal(base_size = 6) +
+  # theme(legend.box = "vertical",
+  #   aspect.ratio = 2/1) +
+  theme(legend.box = "vertical",
+        legend.spacing.y = unit(0.1, 'mm')) +
+
+  plot_layout(ncol = 3, widths = c(1, 1, 1), guides = "collect") & 
+  theme(legend.box = "vertical",
+        legend.spacing.y = unit(0.1, 'mm'))
 

### 2021-07-18 analysis/chapters/paper_ana_20181122.Rmd
Impacts of assimilated observations

@@
 
-(ref:cycle) Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23rd 12 UTC plus and spin up period of 6 hours.
+(ref:cycle) Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
 The zoomed section shows the assimilation for each hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
@@
 
-To investigate the ability of the ensemble to represent the possible states of the atmosphere taking into account the uncertainties of the model and the observations we calculated the mean and sd of the first-guess RCRV for the RAD experiment between 18 UTC Nov, 20 and 12 UTC Nov, 23.
-As this experiment assimilates all types of observations used in this work, it is possible to analyse the reliability of the ensemble by comparing with each type of observation.
-Figure 5 shows the sd RCRV for surface observations computed and box-averaged to a 2.5° grid.
-The wind (Figure 5 a) has a $sd RCRV$ close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error.
-For the temperature (Figure 6 b and d), the results are similar except for some areas to the west of the domain where sd RCRV can be as high as 4.5.
-These areas were studied in detail and it was found that the mean innovation was overly large.
-The main reason for this is the complexity of the topography, usually associated with differences between the model topography and the station height leading to relatively large errors.
-
-Figure 6 shows the mean and standard deviation of the RCRV for the upper-air observations.
-Figure 7 a and b show the RCRV statistics in soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
-Both ADPUPA and AIRCFR show a general good agreement between the ensemble spread and the observation error while AIRCAR presents an irregular profile with values that suggests that the error for this type of observation is overestimated.
-ADPUPA and AIRCAR also present a mean RCRV profile near zero on middle and upper levels but ADPUPA shows a cold bias at low levels, a characteristic already studied in Ruiz 2010 and Dillon 2021.
-
-For satellite-derived wind observations (Figure 7 c)) in low levels, the profiles of b and d that are not close to the expected values with a negative bias and an overestimation of the observation error.
-This characteristic persists throughout all levels.
-Wind estimation derived from water vapor channels shows a bias close to zero in upper levels.
-
-The mean RCRV profiles calculated from the radiance observations (Figure 7 d)) shows almost no bias and the same happens if the mean RCRV is calculated over each channel of each sensor indicating that the bias correction algorithm worked as expected.
-The sd RCRV values are less than 1 for all sensors possible due to an overestimation of the observation errors.
-
-Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work   Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate.
+To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid for the RAD experiment.
+As this experiment assimilates all types of observations used in this work, it is possible to analyse the reliability of the ensemble by comparing it with each type of observations.
+$sd RCRV$ for  wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error.
+For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas to the west of the domain where the $sd RCRV$ can be as high as 4.5.
+In this region the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.
+
+Figure \@ref(fig:rcrv-profile) shows the mean and standard deviation of the RCRV for the upper-air observations.
+Figure \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
+Both ADPUPA and AIRCFR show a general good agreement between the ensemble spread and the observation error.
+As sounding observations and the associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread.
+AIRCAR presents an irregular profile with values that suggests that the error for this type of observation is overestimated.
+ADPUPA and AIRCAR also present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dillonInreview.
+
+For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behaviour with a negative bias and an overestimation of the observation error.
+Wind estimations derived from water vapor channels are abundant over 500 hPa  where their bias is close to zero.
+The only exception are the EUMETSAT observations which contribute very little in the region.
+
+The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) shows almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor indicating that the bias correction algorithm works as expected.
+The $sd RCRV$ values are less than 1 for all sensors possibly due to an overestimation of the observation errors to reduce the influence of potentially erroneous observations.
+
+Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
+Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate.
 
@@
 
-In this section we  investigate the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
-The analysis is performed over a smaller domain (red box in Figure 1) to avoid high topography effects and to focus on the region most directly affected by the MCS.
-Figure 8 a to c shows the difference between AUT and CONV in the spatially averaged vertical profile of temperature.
-The assimilation of  ASWS observations results in a colder planetary boundary layer (PBL).
-This cooling effect has a clear diurnal cycle that is stronger during night time.
-(Figure 8 a).
-The impact of ASWS observations is not limited to the PBL, as significant impact  in mid to upper levels is also found during the mature stage of the MCS.
-The warming in AUT present between 500 and 300 hPa is produced by the development of stronger convection in this experiment compared to CONV.
-SATWND observations have an almost negligible impact on the mean temperature (Figure 8 b)).
-The assimilation of radiances produces significant impacts in low level temperature.
-These observations produce a warming effect in the analyzed PBL which partially compensate for the cooling effect of ASWS observations, particularly during the first 24 hours.
-After that period the impact of assimilating radiances becomes smaller in the first levels.
-The impact of the radiances is also important during the mature stage of the MCS and extended to the end of the simulation period.
-
-Comparing the specific humidity in the experiments (Figure 8 d-f) the impact of assimilating  ASWS with larger spatial and temporal resolution is most significant at low levels (Figure 8 d)).
-The PBL in the AUT experiment is consistently moister than in the CONV experiment.
-The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [Ruiz 2010, algún otro del SMN?].
-The impact of satellite-derived winds are smaller and only important after 18 UTC Nov, 21 at low levels.
-As for the temperature, the impact of RAD observations reduces the impact of ASWS at low levels until 12 UTC Nov, 22.
-Radiances also reduce low-middle level moisture above the boundary layer where the impact of surface observations is not so important.
-This effect is significant during the development of the MCS between 00 and 12 UTC Nov, 22.
-
-The impacts on the wind components are shown in Figure 9.
-The shaded values correspond to the difference between the experiments while the contours represent the spatially averaged vertical wind profile for the experiment highlighted in bold on each figure.
-For example, Figure 9 a) shows the difference between AUT and CONV (shaded) and the averaged zonal wind for AUT as is the experiment with more assimilated observations on that panel.
+In this section we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
+The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS.
+Figure \@ref(fig:TQ-diff)a-c shows the difference between AUT and CONV in the spatially averaged vertical profile of temperature.
+By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
+During the first day, the assimilation of ASWS observations results in a colder PBL.
+This cooling effect has a clear diurnal cycle being stronger during night time.
+(Figure \@ref(fig:TQ-diff)a).
+During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
+The warm difference  shown in AUT with respect to CONV  in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV.
+This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection.
+Figure \@ref(fig:TQ-diff)b, indicates that SATWND observations have an almost negligible impact on the mean temperature through the experiment.
+This result is unexpected since mid-to-upper level winds can have a profound impact in the organization and evolution of MCSs over the region.
+During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensate for the cooling effect of ASWS observations.
+No clear systematic impact is found above the PBL during this period.
+During the second day, the impact of radiance observations is found through the troposphere with a distribution which is similar to the impact found in the AUT experiment but with opposite sign.
+
+Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \@ref(fig:TQ-diff)d).
+The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the night time.
+The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@ruiz2010].
+We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL.
+In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
+As for the temperature, the systematic impact of satellite-derived winds is smaller and only important after 18 UTC Nov, 21 at low levels.
+Figure \@ref(fig:TQ-diff)f shows that, during the first day of the experiment, radiances reduce low-middle level moisture.
+The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov, 22.
+
+The impacts on the wind components are shown in Figure Figure \@ref(fig:UV-diff) along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
 
 The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
-Therefore, these observations generate a reduction of the northerly flow associated with the South American low level jet during night time.
+There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with stronger reduction of the northerly wind during night hours.
+This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment.
 After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain.
-At upper levels, the sign of the impact varies throughout the experiment period becoming larger  during the development of the MCS.
-During Nov 22 and 23 the impact of assimilating ASWS in AUT, produces an increase of northerly wind in upper levels.
-This could indicate a stronger development of convection and therefore an increase in the outflow of convective cells.
-The assimilation of satellite-derived winds produces impacts only in middle and upper levels.
-The impact in the northerly wind is an increase of 0.8 m/s on average.
-
-The assimilation of radiance observations produced a reduction in the westerly wind compared with SATWIND in low and upper levels.
-For the meridional wind, these observations produce an enhancement of the northerly low-level flow of $1 ms^{-1}$ on average, opposite to what is generated by the assimilation of ASWS observations.
-On upper levels and during Nov 22 and 23 the impact of assimilating radiances have the opposite effect on the meridional wind when compared with the impact of the ASWS observations.
-
-To summarize these results, ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly at low levels.
-These changes in the planetary boundary layer can substantially affect the development of the convection.
-On the other hand, radiances also produce important impacts but these are, in general, opposite to those of ASWS observatios.
-Spatial fields of different variables at 00 UTC Nov, 22 will be analyzed next to obtain an overview  of the environment prior to the development of the MSC.
-
-To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22.
-At that moment the first convection cells were developing on the southern region of the domain along the cold front.
-Figure 10 a) shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
-We found that the moist tongue extending over Northern and central Argentina is enhanced by the assimilation of denser surface observations.
-The moisture increase is particularly stronger at the southern tip of this tongue at the boundary of the relatively colder air mass where the convection leading to the MCS was triggered.
-As a result AUT and SATWND experiments are very similar, with values of precipitable water over 55 kg/m2 north of 30°S and a similar vertical distribution of specific humidity (not shown).
-RAD has lower precipitable water content than previous experiments but not as low as CONV.
-The distribution of moisture at low and medium levels observed in RAD can be explained by a combination of the drying that generates the assimilation of radiances and a higher moisture transport due to the stronger northerly wind with values above 10 m/s over the centre of the domain.
-
-The potential temperature on the PBL (Figure 11 b) resembles the characteristics observed on the temperature profiles in Figure 8 a and c.
-On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL due to the assimilation of radiance observations.
-Figure 11 c) shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 km wind shear.
-The values of MCAPE in CONV do not exceed 2000 JKg-1 while the rest of the experiments show maximum MCAPE over 4000 JKg-1.
-The decrease of the specific humidity in lower levels present in RAD could explain the slightly lower values of MCAPE that is partially compensated by an increase in the temperature in the PBL, compared with AUT or SATWND.
-The wind shear is more intense in RAD  reaching  values over 15 ms-1 over the area of maximum MCAPE at the southern tip of the northerly low-level flow.
-Stronger wind shear between 0 and 6 km has been associated with the development of more intense and organized MCS [Chen 2015, ] and also with conditions favorable for supercells (e.g.
+During the second day of the experiment the systematic impact on the wind field is larger at mid and upper levels.
+During Nov 22 and 23 the impact of assimilating ASWS, produces an increase of northerly wind in upper levels.
+This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
+Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
+
+The assimilation of radiances produced a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure  \@ref(fig:UV-diff)c and f).
+For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of ASWS observations.
+On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind.
+When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind.
+
+To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles).
+At that moment the first convective cells were developing on the southern region of the domain along the cold front.
+Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
+We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
+The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
+AUT and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
+RAD has lower precipitable water content than AUT and SATWND but not higher than CONV.
+The distribution of moisture at low levels in RAD seems to be the results of the combination of the moistening effect of assimilating ASWS, partially compensated by the assimilation of radiance observations, and a reduced meridional moisture transport due to the weaker northerly flow (with respect to CONV) over the centre of the domain.
+
+The potential temperature on the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed on the temperature profiles in Figure \@ref(fig:TQ-diff)a-c.
+On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL than AUT due to the assimilation of radiance observations.
+Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear.
+The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$.
+MCAPE in the RAD experiment is lower compared to CONV or SATWND.
+This is consistent with less humidity in the PBL with respect to these experiments and maybe partially compensated by a slightly warmer PBL in the RAD experiment.
+The 0-6 km wind shear is more intense in AUT, SATWND and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
+Moreover in this same region, these experiments show larger MCAPE values.
+Wind shear over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells (e.g.
 Markowsky and Richardson 2010).
-Particularly in this case we can see that on the Southern tip of the strong CAPE region, the 0-6 km wind shear is over 15 ms-1 which is considered favorable for supercell occurrence, while in the CONV experiment the intersection of the unstable warm air and the high shear is slightly smaller and occurs at lower MCAPE values.
-
-```{r TQ-diff, eval=FALSE, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of temperature (a, b, and c, in K) and specific humidity (d, e, and f in gkg-1).", fig.env="figure*", fig.height=3.5}
+
+(ref:TQ-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg-1$).
+
+```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=6, fig.align="center",fig.pos="t"}
 files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
@@
   scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.5)) +
+                                                            barheight = 0.3,
+                                                            title.position = "left", 
+                                                            title.vjust = 1,
+                                                            frame.colour = "black")) +
   labs(fill = "Temperature (K)",
@@
   tag_facets(position = list(x = 0.1, y = 0.96)) +
-  theme_minimal() +
+  theme_minimal(base_size = 8) +
   theme(legend.position = "bottom",
         panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
+        panel.grid = element_line(linetype = 3)) +
 
@@
                                    guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.5)) +
+                                                            barheight = 0.3, 
+                                                            title.position = "left", 
+                                                            title.vjust = 1,
+                                                            frame.colour = "black")) +
 
@@
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  labs(fill = "Specific humidity (g/Kg)",
+  labs(fill = latex2exp::TeX("Specific \nhumidity ($gKg^{-1}$)"),
        y = NULL,
@@
   tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
-  theme_minimal() +
+  theme_minimal(base_size = 8) +
   theme(legend.position = "bottom",
         panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-```
-
-```{r UV-diff, eval=FALSE, fig.cap="Difference between experiments AUT-CONV (a and b), SATWND-AUT (b and e) and RAD-SATWND (c-f) for the spatially averaged vertical profile of u wind (a, b, and c, in ms-1) and v wind (d, e, and f in ms-1).
-In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f).", fig.height=3.5, fig.env="figure*"}
+        panel.grid = element_line(linetype = 3)) +
+
+  plot_layout(ncol = 1, heights = c(1, 1))
+
+```
+
+(ref:UV-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in ms-1) and v wind (d, e, and f in $ms^{-1}$).
+In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f) are shown.
+
+```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=6, fig.align="center", fig.pos="t"}
 perfiles %>% 
@@
   scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.5)) +
+                                                            barheight = 0.3, 
+                                                            title.position = "left", 
+                                                            title.vjust = 1,
+                                                            frame.colour = "black")) +
   geom_contour2(aes(z = value), color = "white", size = 0.1,
@@
                     size = 4, stroke = 0.1) +
-  labs(fill = "U (m/s)",
+  labs(fill = latex2exp::TeX("U ($ms^{-1}$)"),
        x = NULL,
@@
   tag_facets(position = list(x = 0.1, y = 0.96)) +
-  theme_minimal() +
+  theme_minimal(base_size = 8) +
   theme(legend.position = "bottom",
@@
         panel.grid = element_line(linetype = 3), 
-        strip.text = ggtext::element_markdown())
+        strip.text = ggtext::element_markdown()) +
 
@@
                                    guide = guide_colorsteps(barwidth = 25,
-                                                            barheight = 0.5)) +
+                                                            barheight = 0.3, 
+                                                            title.position = "left", 
+                                                            title.vjust = 1,
+                                                            frame.colour = "black")) +
   geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
@@
                     size = 4, stroke = 0.1) +
-  labs(fill = "V (m/s)",
+  labs(fill = latex2exp::TeX("V ($ms^{-1}$)"),
        x = NULL,
@@
   tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
-  theme_minimal() +
+  theme_minimal(base_size = 8) +
   theme(legend.position = "bottom",
@@
         panel.grid = element_line(linetype = 3),
-        strip.text = ggtext::element_markdown())
-
-```
-
-```{r summary-fields, eval=FALSE, fig.cap="a) Precipitable water (shaded) and northerly wind (contours, m/s), b) Average potential temperature for the PBL and c) Maximum CAPE and ~0-6 km wind shear over 15 m/s and 30 m/s for each experiment.
-All fields correspond to 00 UTC Nov, 22.", fig.height=10, fig.width=9, fig.env="figure*"}
+        strip.text = ggtext::element_markdown()) +
+    plot_layout(ncol = 1, heights = c(1, 1))
+
+```
+
+(ref:summary-fields) Precipitable water (shaded, $kgm^{-2}$) and average northerly wind over the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, $ms^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $ms^{-1}$ for each experiment.
+All fields correspond to 00 UTC Nov, 22.
+
+```{r summary-fields, fig.cap="(ref:summary-fields)", out.width="100%", fig.height=12, fig.width=10.5, fig.env="figure*"}
 
@@
 
-ana[bottom_top %between% c(2, 7), .(v_mean = mean(v)), 
+ana[bottom_top %between% c(1, 7), .(v_mean = mean(v)), 
     by = .(south_north, west_east, date, exp)] %>% 
@@
   # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
-  labs(fill = "Precipitable\nwater (Kg/m²)", size = "V (m/s)") +
+  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1}")) +
   #cordillera +
@@
 
-  ana[bottom_top %between% c(2, 10), .(theta_mean = mean(theta)), 
+  ana[bottom_top %between% c(1, 10), .(theta_mean = mean(theta)), 
       by = .(south_north, west_east, date, exp)] %>% 
@@
   #                    guide = guide_legend(override.aes = list(size = 2))) +
-  labs(fill = "MCAPE (J/Kg)", size = "Wind\nshear (m/s)") +
+  labs(fill = latex2exp::TeX("MCAPE ($JKg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($ms^{-1}$)")) +
   #cordillera +

### 2021-07-18 analysis/chapters/paper_ana_20181122.Rmd
Validation

@@
                                                guide = guide_colorsteps(barwidth = 25,
-                                                                        barheight = 0.5)) +
+                                                                        barheight = 0.5, 
+                                                                        title.position = "left", 
+                                                                        title.vjust = 1,
+                                                                        frame.colour = "black")) +
   geom_contour2(aes(z = v_mean, size = stat(level)),
@@
                        guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.5)) +
+                                                barheight = 0.5, 
+                                                title.position = "left", 
+                                                title.vjust = 1,
+                                                frame.colour = "black")) +
   labs(fill = "Potential\ntemperature (K)") +
@@
                        guide = guide_colorsteps(barwidth = 25,
-                                                barheight = 0.5)) +
+                                                barheight = 0.5, 
+                                                title.position = "left", 
+                                                title.vjust = 1,
+                                                frame.colour = "black")) +
   # geom_contour2(aes(z = cortante),
@@
 
-Analysis verifications are performed to assess the skill of each experiment to represent the atmospheric circulation, particularly associated with the MSC development and precipitation.
-Figure 12 shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67°W and 54.5°W as a function of time and latitude in the different experiments, and estimated precipitation by IMERG.
-The heaviest precipitation (over 12 mm/h) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the experiments.
+Firstly we analyse the impact of different observation types in terms of the representation of the MCS and its associated precipitation.
+Figure \@ref(fig_pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and estimated precipitation by IMERG.
+The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the experiments.
 In all the experiments, the accumulated precipitation in the short-range forecasts underestimate the precipitation.
@@
 AUT, SATWND and RAD better capture the timing and location of convective initiation.
-On RAD the initiation is delayed respecto to AUT and SATWND possibly related to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
-After 18 UTC Nov, 22 RAD shows improvements in precipitation rate and its distribution compared to the other experiments due to the enhanced development of the conveccion.
-
-To quantify the spatial match between the estimated precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure 13).
-All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, afterwards FSS values improve during the mature and dissipation stages of the MCS.
-The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the maturity and dissipation stages of the MCS.
-AUT and  SATWND show similar FSSs indicating that satellite-derived winds assimilation has little impact on the precipitation for the present case study.
-The assimilation of radiance observations improves the skill of the 1-hour forecasted precipitation as the FSS in RAD is higher, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
-The enhancement is particularly important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).
-
-To complement  the analysis of the representation of the MCS evolution by the different experiments, Figure 14 shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments  ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed reflectivity.
-These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations.
-In addition, these experiments are the worst and best performing experiments with respect to precipitation.
-Overall none of the short range forecasts is able to capture the mesoscale details in the precipitation distribution.
-This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately capture the strength of the convective band associated with the MCS.
-Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field.
-RAD better represents the observed features of the MSC showing an organized pattern over the centre domain at 10 and 13 UTC.
-The convection cells that initiate after 16 UTC in the northeast region are also present in the 1-hr forecast.
-CONV also captures the location of the MCS, but the convection seems to be less organised and much weaker than in RAD.
-Before and after the times shown in Figure 14, the agreement between experiment and observations is quite high in the regions where radar data are available.
-
-Individual radiosondes were also compared with the experiments to analyse the specific characteristics of the atmosphere before and after the development of the MCS.
-
-Figure 15 shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO IOP07 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP08 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
-During the first period, no significant meteorological events were observed on the domain of interest, thus the synoptic situation corresponds to a warm day with an important humidity transport from the centre of South America.
-For the temperature, CONV presents a warm bias on the PBL of 1°C and an RMSE between 1 and 2.5°C that are slightly reduced  by the assimilation of ASWS and radiances.
-The RMSE for temperature remains between 0.5 and 2°C throughout the profile and the assimilation has almost no effect in middle and upper levels.
-The dew point temperature shows a small dry bias in low levels that is corrected on AUT and SATWND.
-On middle levels (below 10 km) the bias is large and positive (8 - 10°C), indicating a more moist atmosphere in the analysis.
-The RMSE also grows with height where small variations of humidity produces major chances in dew point temperatures.
-The zonal wind has a significant positive bias around 10 km of 10 m/s, this means that all experiments present a stronger westerly wind that is in part corrected by the assimilation of radiance observations.
-Finally, the meridional wind shows small values of bias and RMSE through the vertical profile but the assimilation of radiance observations increases the northerly wind near the surface.
-
-For the second period, corresponding to IOP08, none of the radiosondes was directly affected by the convective processes since the MSC crossed RELAMPAGO's region of study before the beginning of the observation period.
-Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias is cold and only the ASWS observations improves the analysis.
-The RMSE for AUT, SATWND and RAD remains below 2°C for almost all levels with small improvement in upper levels where radiance observations have the most impact.
-In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
-The moist bias in conjunction with the underestimation of the precipitation might be indicating that the convection processes on the model are not converting the available moisture into precipitation or that the model fails to represent boundary layer processes correctly.
-SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7°C.
-Once again, the zonal wind is overestimated on the analyses with a positive bias on middle and upper layers.
-For this period only RAD shows an improvement on the upper layers of the troposphere.
-In low levels the meridional wind presents a positive bias, indicating an underestimation of the northerly wind principally in AUT, SATWND and RAD.
-In middle levels both AUT and SATWND have almost no bias while RAD shows a positive bias of 2.5 m/s while its RMSE does not exceed 5 m/s.
-
-To summarize, the results suggest that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study.
-While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels.
-The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the CSM leading to an increase in the accumulated precipitation.
-
-```{r pp-hov, eval=FALSE, fig.cap="Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right, for the ensemble mean) over the inner domain."}
+On RAD the initiation is delayed with respect to AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
+After 18 UTC Nov, 22 RAD shows improvements in precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the conveccion.
+
+(ref:pp-hov) Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right, for the ensemble mean) calculated over the small domain (red box in Figure \@ref(fig:dominio)a).
+
+```{r pp-hov, fig.cap="(ref:pp-hov)", fig.height=4, fig.env="figure*", fig.pos="h"}
 
@@
                        super = ScaleDiscretised,
-                       guide = guide_colorsteps(inside = TRUE,
-                                                barwidth = 25,
-                                                barheight = 0.5)) +
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.3, 
+                                                title.position = "left", 
+                                                title.vjust = 1,
+                                                frame.colour = "black")) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
@@
   labs(x = NULL, y = NULL, fill = "Acumulated\nprecipitation (mm/h)") +
-  theme_minimal() +
+  theme_minimal(base_size = 9) +
   theme(legend.position = "bottom") +
@@
 
-```{r fss, eval=FALSE, fig.cap="FSS calculated over a 6-hours moving window for different thresholds and scales."}
+To quantify the spatial match between the observed precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss).
+All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature and dissipation stages of the MCS.
+The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the maturity and dissipation stages of the MCS.
+AUT and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
+The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
+The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).
+
+(ref:fss) FSS calculated over a 6-hours moving window for a) 1 mm and b) 25 mm with a scale of 10 km and c) 1 mm and a d) 25 mm with a scale of 100 km, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
+
+```{r fss, fig.cap="(ref:fss)", fig.width=3.5, fig.height=3.5, fig.pos="t"}
 files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2-8].csv")
@@
                                             "51" = "500 km", "101" = "1000 km"))) +
+  tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
   labs(color = NULL, x = NULL, y = "FSS") +
-  theme_minimal()
-```
-
-```{r dbz-mean, eval=FALSE, eval=FALSE, fig.cap="Observed maximum reflectivity in the column (upper row) and ensemble mean maximum reflectivity for CONV and RAD at different times (bottom two rows).", include=FALSE}
+  theme_minimal(base_size = 8) +
+  theme(legend.position = "bottom")
+```
+
+To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed maximum reflectivity.
+These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations.
+In addition, these experiments are the worst and best performing experiments in terms of the short range precipitation forecast skill.
+Overall none of the short range forecasts is able to capture the mesoscale details in the precipitation distribution.
+This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately capture the strength of the convective band associated with the MCS.
+Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field.
+RAD better represents the observed features of the MSC showing a more strong and organized MCS over the domain center at 10 and 13 UTC (first and second columns in Figure \@ref(fig:dbz-mean)).
+The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain  are well captured by  both experiments, but better represented in terms of strength in the RAD experiment.
+CONV captures the location of the MCS, but the convection seems to be less organised and much weaker than in RAD.
+Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between experiment and observations is quite high in the regions where radar data are available.
+
+Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments.
+The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection.
+For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that are slightly reduced by the assimilation of ASWS and radiances.
+The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND and RAD is small above the PBL.
+The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND and RAD.
+On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C), indicating a more moist atmosphere in the analysis.
+Overall, the RMSE grows with height, this is probably because at upper levels, moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
+The zonal wind has a significant positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced by assimilating satellite radiances.
+It is somewhat surprising that no significant effect is observed by assimilating satellite derived winds.
+The assimilation of ASWS and radiance observations increases the northerly wind bias and RMSE near the surface.
+
+(ref:dbz-mean) Observed maximum reflectivity in the column (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third rows) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22.
+
+```{r dbz-mean, fig.cap="(ref:dbz-mean)", out.width="100%", fig.height=8, fig.width=8, fig.env="figure*", fig.pos="h"}
 proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"
@@
                     drop = FALSE, 
-                    guide = guide_colorsteps(barwidth = 30, 
-                                             barheight = 0.8)) +
+                    guide = guide_colorsteps(barwidth = 25,
+                                             barheight = 0.4, 
+                                             title.position = "left", 
+                                             title.vjust = 1,
+                                             frame.colour = "black")) +
   # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
@@
   labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
-  theme_minimal() +
+  theme_minimal(base_size = 9) +
   theme(legend.position = "bottom",
@@
 
-```{r soundings, eval=FALSE, fig.cap="RMSE and Bias of temperature, dew point temperature and wind components calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8."}
+For the IOP 8 (bottom row in Figure \@ref(fig:soundings)), the densely observed area was behind the MCS, but far enough from it to be directly affected by its mesoscale circulations.
+This area was also behind the cold front thus affected by low-level cold advection.
+Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
+Between 5 and 12 km, the bias in each experiment is cold.
+AUT, SATWND and RAD show a lower bias due to the assimilations of ASWS.
+The RMSE for AUT, SATWND and RAD remains below 2$^{\circ}$C for almost all levels with significant  improvement in upper levels where radiance observations have the largest impact.
+In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
+This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region.
+The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
+SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C.
+The zonal wind is overestimated on the analyses with a positive bias at middle levels.
+For this period only RAD shows an improvement in the upper troposphere.
+At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND and RAD.
+In fact low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
+
+(ref:soundings) RMSE (solid line) and Bias (dash line) of a) temperature, b) dew point temperature, c) u wind and d) v wind calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+The blue line correspond to CONV, the ligh blue line to AUT, SATWND is represented with an orange line and RAD with a red line.
+
+```{r soundings, fig.cap="(ref:soundings)", fig.env="figure*", fig.pos="t", fig.align="center"}
 files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos/ANA/sondeo_E[2568]_ana*")
@@
   facet_grid(iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature",
-                                              "td" = "Dew point\ntemperature",
-                                              "u" = "U wind",
-                                              "v" = "V wind"))) +
+             labeller = labeller(variable = c("t" = "a) Temperature",
+                                              "td" = "b) Dew point temperature",
+                                              "u" = "c) U wind",
+                                              "v" = "d) V wind"),
+                                 iop = c("IOP07" = "IOP 7",
+                                         "IOP08" = "IOP 8"))) +
   labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
-  theme_minimal()
+  theme_minimal(base_size = 9) +
+  theme(legend.position = "bottom")
 ```
@@
 
+To summarize these results, firstly, ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly at low levels.
+These changes in the planetary boundary layer can substantially affect the development of the convection.
+Secondly, radiances also produce important impacts but these are, in general, opposite to those of ASWS observatios.
+Spatial fields of different variables at 00 UTC Nov, 22 will be analyzed next to obtain an overview of the environment prior to the development of the MSC.
+
+To summarize, the results suggest that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study.
+While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels.
+The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
+
 # References

### 2021-07-19 analysis/chapters/paper_ana_20181122.Rmd
Algunas correcciones menores.

@@
 We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November, 22th, 2018.
-The accumulated precipitation observed during the transit of the MCS was quite high with values over 200 mm over the northern part of the domain.
+The accumulated precipitation observed during the transit of the MCS was quite high, exceeding 200 mm over the northern part of the domain.
 This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during Nov.
@@
 Given the sparsity of the local conventional observing network it is important to explore the potential benefits of including additional sources of observations at a regional scale. 
-Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
 Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
@@
 Moreover @bao2015 studied the impact on temperature and humidity forecast over western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
 Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts [@gao2015].
@@
 This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
-The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in [@necker2020; @maldonado2021].
+The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021.
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
@@
 We use 9 different model configurations consisting on the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
-All ensemble members used the same land-surface model (Noah-MP, @chen2001) and microphysics (WRF single-moment 6–class squeme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
+All ensemble members used the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class squeme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
 
@@
 The orange triangles in Figure \@ref(fig:dominio)a, are the surface stations and ship locations included in this experiment.
+The orange triangles in Figure \@ref(fig:dominio) a), are the surface station locations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.
@@
 Satellite-derived wind observations also come from GDAS and includes information from GOES-16 (estimated from the visible, infrared and water vapour channels) and METEOSAT 8 and 11 (visible and water vapour channels).
-Due to the domain covered by each of these satellites, GOES-16 is the primary source of this type of observation (99 % of observation).
-
-We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of  various public and private automatic station networks over Southern South America and are part of RELAMPAGO Data Set repository [@garcia2019].
+Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations).
+
+We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of  various public and private automatic station networks over Southern South America and are included in RELAMPAGO Data Set repository [@garcia2019].
 These stations are indicated as green squares in Figure \@ref(fig:dominio)a.
@@
 In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind).
-In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \@ref(tab:table-obs)).
+In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and the model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \@ref(tab:table-obs)).
 
@@
 This is evident in Figure \@ref(fig:obs-cycle)a which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
-In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiments.
+In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiment.
 The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM.
@@
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, no the observation quality (based on available data quality predictors) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
+Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, on the observation quality (based on available data quality information) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
 Also, observations over the sea are preferred from those over land or snow [@hu2018].
@@
 
-$$\mathit{BT_bc} =\mathit{ BT} + \sum{i = 0}{N} \beta_i p_i (x)$$
-
-GSI has a constant offset bias correction term ($p_0 = 1$).
-The following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
+$$\mathit{BT_{bc}} =\mathit{ BT} + \sum_{i = 0}^{N} \beta_i p_i (x)$$
+
+GSI has a constant offset bias correction term ($p_0 = 1$) and the following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
 Scan angle-dependent bias is modelled as a 4th-order polynomial [@zhu2014].
 
-The $\beta_i$ coefficients are usually trained to achieve the best fit between the simulation and the observations.
-In the GSI system these coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
-The coefficients initialized from the 18 UTC Nov, 11 2018 GFS coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
+In the GSI system the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
+The coefficients initialized from the 18 UTC Nov, 11 2018 GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
 Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.
@@
 In our experiments, only clear-sky observations are used.
-For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes[@weston2019; @zhu2016].
+For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes [@weston2019; @zhu2016].
 For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
@@
 
-(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle and divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
+(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
 
@@
 * The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
-Only the measurements/soundings that did not enter the assimilation system were used for validation.
-The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio)b.
-
-* The Satellite precipitation estimation IMERG Final Run with 0.01$^{\circ}$ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill 1-hour forecasts to represent the precipitation over the domain.
+Only the soundings that did not enter the assimilation system were used for validation.
+The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio)b).
+
+* The Satellite precipitation estimation IMERG Final Run with 0.01$^{\circ}$ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
 
@@
 The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
-The maxima after 12 and 00 UTC correspond to the contribution of the multispectral sensors.
-The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b shows a maximum in low levels due to the ASWS observations.
+The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors.
+The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b) shows a maximum in low levels due to the ASWS observations.
 Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa).
@@
   "PREPBUFR", "x", "x", "x", "x",
-  "AUTSFC", " ", "x", "x", "x",
+  "ASWS", " ", "x", "x", "x",
   "SATWND", " ", " ", "x", "x",
@@
 A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper.
-These aspects include a validation of how uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations which are not assimilated .
+These aspects include a validation of how the uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations which are not assimilated.
+
 To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:
@@
 
-where $x_o$ is the error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
+where $x_o$ is the observation and it error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
 The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
@@
 where $M$ is the sample size.
-A reliable system will have no bias ($mean RCRV = 0$), and a standard deviation equal to 1, $sd RCRV = 1$.
+A reliable system will have no bias ($mean RCRV = 0$), and a standard deviation equal to 1 ($sd RCRV = 1$).
 If the ensemble has a positive (negative) bias, $mean RCRV$ will be positive (negative).
@@
 
-where ${P_x}$ and is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and ${P_o}$ is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise.
+where ${P_x}$ is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and ${P_o}$ is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise.
 This probabilistic formulation uses the information from the full ensemble thus avoiding the smoothing effect of the ensemble mean which can particularly affect localized extreme precipitation values.
@@
   # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
-  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1}")) +
+  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
   #cordillera +

### 2021-07-26 analysis/chapters/paper_ana_20181122.Rmd
Revisión del texto

@@
 
-To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid for the RAD experiment.
+To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment.
 As this experiment assimilates all types of observations used in this work, it is possible to analyse the reliability of the ensemble by comparing it with each type of observations.
-$sd RCRV$ for  wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error.
+Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid.
+The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error.
 For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas to the west of the domain where the $sd RCRV$ can be as high as 4.5.
@@
 Both ADPUPA and AIRCFR show a general good agreement between the ensemble spread and the observation error.
-As sounding observations and the associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread.
-AIRCAR presents an irregular profile with values that suggests that the error for this type of observation is overestimated.
-ADPUPA and AIRCAR also present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dillonInreview.
-
-For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behaviour with a negative bias and an overestimation of the observation error.
+As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread.
+AIRCAR presents an irregular profile with $sd RCRV$ values that suggests that the error for this type of observation is overestimated.
+ADPUPA and AIRCAR present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dilloninreview.
+
+For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, where the are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behaviour with a negative bias and an overestimation of the observation error.
 Wind estimations derived from water vapor channels are abundant over 500 hPa  where their bias is close to zero.
@@
 
-The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) shows almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor indicating that the bias correction algorithm works as expected.
+The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) shows almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor.
+This indicates that the bias correction algorithm works as expected.
 The $sd RCRV$ values are less than 1 for all sensors possibly due to an overestimation of the observation errors to reduce the influence of potentially erroneous observations.
@@
 Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
-Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate.
+Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate for the system.
 
@@
 The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS.
-Figure \@ref(fig:TQ-diff)a-c shows the difference between AUT and CONV in the spatially averaged vertical profile of temperature.
+Figure \@ref(fig:TQ-diff)a-c shows the difference between experiments in the spatially averaged vertical profile of temperature.
 By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
 During the first day, the assimilation of ASWS observations results in a colder PBL.
-This cooling effect has a clear diurnal cycle being stronger during night time.
-(Figure \@ref(fig:TQ-diff)a).
+This cooling effect has a clear diurnal cycle being stronger during night time (Figure \@ref(fig:TQ-diff)a).
 During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
@@
 No clear systematic impact is found above the PBL during this period.
-During the second day, the impact of radiance observations is found through the troposphere with a distribution which is similar to the impact found in the AUT experiment but with opposite sign.
+During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AUT experiment but with opposite sign.
 
@@
 In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
-As for the temperature, the systematic impact of satellite-derived winds is smaller and only important after 18 UTC Nov, 21 at low levels.
-Figure \@ref(fig:TQ-diff)f shows that, during the first day of the experiment, radiances reduce low-middle level moisture.
+As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels.
+Figure \@ref(fig:TQ-diff)f shows that radiances reduce low-middle level moisture during the first day of the experiment.
 The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov, 22.
@@
 After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain.
-During the second day of the experiment the systematic impact on the wind field is larger at mid and upper levels.
+After the initiation of the convective cells the systematic impact on the wind field is larger at mid and upper levels.
 During Nov 22 and 23 the impact of assimilating ASWS, produces an increase of northerly wind in upper levels.
@@
 On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind.
-When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind.
-
-To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles).
+When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in  Figure /@ref(fig:UV-diff)f.
+
+To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles).
 At that moment the first convective cells were developing on the southern region of the domain along the cold front.
@@
 AUT and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
-RAD has lower precipitable water content than AUT and SATWND but not higher than CONV.
-The distribution of moisture at low levels in RAD seems to be the results of the combination of the moistening effect of assimilating ASWS, partially compensated by the assimilation of radiance observations, and a reduced meridional moisture transport due to the weaker northerly flow (with respect to CONV) over the centre of the domain.
+RAD has lower precipitable water content than AUT and SATWND but higher than CONV.
+The distribution of moisture at low levels in RAD seems to be the results of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the centre of the domain compare to CONV.
 
@@
 The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$.
-MCAPE in the RAD experiment is lower compared to CONV or SATWND.
+MCAPE in the RAD experiment is lower compared to AUT or SATWND.
 This is consistent with less humidity in the PBL with respect to these experiments and maybe partially compensated by a slightly warmer PBL in the RAD experiment.
@@
 Moreover in this same region, these experiments show larger MCAPE values.
-Wind shear over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells (e.g.
-Markowsky and Richardson 2010).
+Wind shear over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowsky2010].
 
@@
 Firstly we analyse the impact of different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \@ref(fig_pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and estimated precipitation by IMERG.
-The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the experiments.
+Figure \@ref(fig:pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and the estimated precipitation by IMERG.
+The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period.
 In all the experiments, the accumulated precipitation in the short-range forecasts underestimate the precipitation.
-This is particularly evident on CONV where the convection initiation is significantly delayed with respect to the observations and occurs further north with respect to the observed initiation.
+This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation.
 AUT, SATWND and RAD better capture the timing and location of convective initiation.
 On RAD the initiation is delayed with respect to AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
-After 18 UTC Nov, 22 RAD shows improvements in precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the conveccion.
+After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the conveccion.
 
@@
 
-To quantify the spatial match between the observed precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss).
-All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature and dissipation stages of the MCS.
-The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the maturity and dissipation stages of the MCS.
+To quantify the spatial match between the observed precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
+All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature stage of the MCS.
+The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
 AUT and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
@@
 
-Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+Finally, Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
 IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments.
@@
 Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold.
-AUT, SATWND and RAD show a lower bias due to the assimilations of ASWS.
+Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND and RAD to the assimilations of ASWS.
 The RMSE for AUT, SATWND and RAD remains below 2$^{\circ}$C for almost all levels with significant  improvement in upper levels where radiance observations have the largest impact.

### 2021-07-27 analysis/chapters/paper_ana_20181122.Rmd
Arreglo figuras, la 12 se sigue resistiendo.

@@
   # labs(color = "b) Soundings", shape = "b) Soundings") +
-  theme_minimal(base_size = 10) +
+  theme_minimal(base_size = 11) +
   theme(legend.position = "bottom")
@@
   labs(color = NULL, shape = NULL) +
-  theme_minimal(base_size = 10) +
+  theme_minimal(base_size = 11) +
   theme(legend.box = "horizontal",
@@
 ggdraw(plot_grid(plot_grid(g1, g2, ncol = 2, rel_widths = c(1, 0.5), 
-                           labels = c("a)", "b)"), label_size = 10), legend, 
+                           labels = c("a)", "b)"), label_size = 11), legend, 
                  ncol = 1, rel_heights = c(1, 0.05)))
@@
   theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 14),
         plot.margin = unit(c(0, 0, 0, 0), "cm"))
@@
   plot_annotation(tag_levels = 'a', tag_suffix = ")") &
-  theme(plot.tag = element_text(size = 6), 
+  theme(plot.tag = element_text(size = 8), 
         legend.position = "bottom",
@@
 All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
-The initial ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov 20 (Figure \@ref(fig:cycle)).
+The initial ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \@ref(fig:cycle)).
 
@@
   theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 14),
         panel.ontop = TRUE,
@@
        linetype = NULL, color = "Upper air\nobservations") +
-  guides(color = guide_legend(
-    keyheight = 0.4,
-    default.unit = "cm")) +
-  theme_minimal(base_size = 6) +
-  # theme(legend.box = "vertical",
-  #   aspect.ratio = 4/1) +
+  guides(color = guide_legend(keyheight = 0.4, default.unit = "cm"),
+         linetype = guide_legend(keyheight = 0.4, default.unit = "cm")) +
+  theme_minimal(base_size = 8) +
   theme(legend.box = "vertical",
-        legend.spacing.y = unit(0.1, 'mm')) +
+        legend.spacing.y = unit(0.1, 'mm'),
+        tagger.panel.tag.text = element_text(size = 8)) +
 
@@
   scale_color_brewer(palette = "Set1") +
-  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
+  scale_x_level(breaks = c(1000, 500, 250, 125, 50), limits = c(1000, 50)) +
   coord_flip() +
@@
        linetype = NULL, color = "Satellite-derived\nwind") +
-  guides(color = guide_legend(
-    keyheight = 0.4,
-    default.unit = "cm")) +
-  theme_minimal(base_size = 6) +
-  # theme(legend.box = "vertical",
-  #   aspect.ratio = 2/1) +
+    guides(color = guide_legend(keyheight = 0.4, default.unit = "cm"),
+         linetype = guide_legend(keyheight = 0.4, default.unit = "cm")) +
+  theme_minimal(base_size = 8) +
   theme(legend.box = "vertical",
-        legend.spacing.y = unit(0.1, 'mm')) +
+        legend.spacing.y = unit(0.1, 'mm'),
+        tagger.panel.tag.text = element_text(size = 8)) +
 
@@
        linetype = NULL, color = "Radiances\nobservations") +
-  guides(color = guide_legend(
-    keyheight = 0.4,
-    default.unit = "cm")) +
-  theme_minimal(base_size = 6) +
-  # theme(legend.box = "vertical",
-  #   aspect.ratio = 2/1) +
+  guides(color = guide_legend(keyheight = 0.4, default.unit = "cm"),
+         linetype = guide_legend(keyheight = 0.4, default.unit = "cm")) +
+  theme_minimal(base_size = 8) +
   theme(legend.box = "vertical",
-        legend.spacing.y = unit(0.1, 'mm')) +
+        legend.spacing.y = unit(0.1, 'mm'),
+        tagger.panel.tag.text = element_text(size = 8)) +
 
@@
   theme(legend.box = "vertical",
-        legend.spacing.y = unit(0.1, 'mm'))
+        legend.spacing.y = unit(0.1, 'mm'),
+        tagger.panel.tag.text = element_text(size = 8))
 
@@
 On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind.
-When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in  Figure /@ref(fig:UV-diff)f.
+When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in  Figure \@ref(fig:UV-diff)f.
 
@@
   theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 8),
         panel.ontop = TRUE,
@@
   facet_wrap(~ exp) +
-  labs(x = NULL, y = NULL, fill = "Acumulated\nprecipitation (mm/h)") +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
   theme_minimal(base_size = 9) +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 9)) +
 
@@
                                                           E6 = "SATWND", E8 = "RAD"))) +
-  labs(x = NULL, y = NULL, fill = "Acumulated\nprecipitation (mm/h)") +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
   theme_minimal(base_size = 9) +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 9)) +
 
@@
 To quantify the spatial match between the observed precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
-All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature stage of the MCS.
+All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22 with an improvement during the mature stage of the MCS.
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
@@
 
-To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed maximum reflectivity.
+To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity.
 These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations.
@@
 
-dbz_obs %>% 
+p1 <- dbz_obs %>% 
   # .[date == unique(date)[1]] %>% 
@@
   geom_mapa() +
-  facet_grid(run ~ date) +
+  coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)) +
+  scale_x_longitude(ticks = 5, labels = NULL) +
+  facet_grid(run ~ date, labeller = labeller(date = c("2018-11-22 10:00:00" = "10 UTC",
+                                                      "2018-11-22 13:00:00" = "13 UTC",
+                                                      "2018-11-22 16:00:00" = "16 UTC",
+                                                      "2018-11-22 19:00:00" = "19 UTC"))) +
   labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
   theme_minimal(base_size = 9) +
-  theme(legend.position = "bottom",
-        # legend.box.spacing = margin(0, 0, 0, 0),
+  theme(legend.position = "none",
+        axis.text.x = element_blank(),
+        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
         panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3)) +
-
-  dbz_ana %>% 
+        panel.grid = element_line(linetype = 3))
+
+p2 <- dbz_ana %>% 
   .[, date := factor(date)] %>% 
@@
   geom_mapa() +
-  facet_grid(run ~ date) +
+  facet_grid(run ~ date, labeller = labeller(date = c("2018-11-22 10:00:00" = "10 UTC",
+                                                      "2018-11-22 13:00:00" = "13 UTC",
+                                                      "2018-11-22 16:00:00" = "16 UTC",
+                                                      "2018-11-22 19:00:00" = "19 UTC"))) +
   labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
@@
   theme(legend.position = "bottom",
+        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
         strip.background = element_blank(),
@@
         panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3)) +
-
-  plot_layout(ncol = 1, heights = c(1, 2), guides = "collect") & 
-  theme(legend.position = 'bottom')
+        panel.grid = element_line(linetype = 3))
+
+legend <- get_legend(
+  # create some space to the left of the legend
+  p2 + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"), 
+             legend.box.margin = margin(0, 0, 0, 0))
+)
+
+p2 <- p2 + theme(legend.position = "none")
+
+plot_grid(p1 + theme(legend.position = "none",
+                     plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")),
+          p2 + theme(legend.position = "none",
+                     plot.margin = margin(t = -5, r = 0, b = -2, l = 0, unit = "pt")), 
+          legend, ncol = 1, rel_heights = c(1.1, 2, 0.1))
 

### 2021-07-27 analysis/chapters/coeficientes.Rmd
Revisión de epigrafes

@@
 ```{r}
-files <- list.files("/home/paola.corrales/datosmunin/EXP/E7-training/ANA", 
+files <- list.files("/home/paola.corrales/datosmunin3/EXP/E7-training/ANA", 
                     pattern = "asim.", recursive = TRUE, 

### 2021-07-27 analysis/chapters/experimento_20181122.Rmd
Old stuff

@@
   setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
-derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"
+derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
 
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/derived_data/perfiles_mse*")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/perfiles_rmsd*")
 
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = mse, fill = stat(level))) +
+  geom_contour_fill(aes(z = rmsd, fill = stat(level))) +
   scale_fill_viridis_c(option = "A", direction = -1,
@@
                                                 barheight = 0.8)) +
-  labs(subtitle = "Temperature MSE (K^2)",
+  labs(subtitle = "Temperature RMSD (K)",
        x = NULL,
@@
   scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
-  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                               E6 = "SATWND", E7 = "RAD"))) +
+  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E8 = "RAD"))) +
   theme_minimal() +
@@
 A positive difference can also be observed between 850 and 600 hPa.
+
 ```{r Q-mse, eval=FALSE, fig.cap="MSE", include=FALSE}
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = mse*1000000, fill = stat(level))) +
+  geom_contour_fill(aes(z = rmsd, fill = stat(level))) +
   scale_fill_viridis_c(option = "A", direction = -1,
@@
                                                 barheight = 0.8)) +
-  labs(subtitle = "Specific humidity MSE (g/Kg^2)",
+  labs(subtitle = "Specific humidity RMSD (g/Kg)",
        x = NULL,
@@
 
+```{r U-mse, eval=FALSE, fig.cap="MSE", include=FALSE}
+
+perfiles_mse %>% 
+  .[variable == "U"] %>% 
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = rmsd, fill = stat(level))) +
+  scale_fill_viridis_c(option = "A", direction = -1,
+                       name = NULL, super = ScaleDiscretised,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  labs(subtitle = "U RMSD (m/s)",
+       x = NULL,
+       y = "Pressure level") +
+  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+
 In Figure \@ref(fig:U-diff) the difference between the experiments for the zonal wind is shown with the vertical profile for the experiment with more observations.
@@
 
+```
+
+```{r V-mse, eval=FALSE, fig.cap="MSE", include=FALSE}
+
+perfiles_mse %>% 
+  .[variable == "V"] %>% 
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = rmsd, fill = stat(level))) +
+  scale_fill_viridis_c(option = "A", direction = -1,
+                       name = NULL, super = ScaleDiscretised,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.8)) +
+  labs(subtitle = "V RMSD (m/s)",
+       x = NULL,
+       y = "Pressure level") +
+  scale_y_level(name = "Pressure level", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_x_datetime(expand = c(0,0), date_labels = "%HZ \n %b-%d", date_breaks = "24 hours") +
+  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                               E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
 ```
@@
   plot_annotation(tag_levels = "a", tag_suffix = ")")
+```
+
+```{r}
+files <- Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122*.nc"))
+
+dates <- c("20181122000000", "20181122060000", 
+           "20181122120000", "20181122180000",
+           "20181123000000")
+
+mucape <- purrr::map(files, function(f) {
+  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]
+
+  if (details$date %in% dates) {
+
+    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
+      .[, `:=`(exp = details$exp, 
+               date = lubridate::ymd_hms(details$date))] %>%
+      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
+      .[]
+}
+}) %>% 
+  rbindlist()
+
+fcsts <- expand_grid(fcsts = dates,
+                     exp = c("E2", "E5", "E6", "E8"))
+
+cor <- lapply(seq_len(nrow(fcsts)), function(f) {
+
+  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
+    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
+    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
+    .[, theta := ReadNetCDF(paste0(derived_data, "/theta_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(th = "theta"), out = "vector")] %>%
+    .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
+    .[, exp := fcsts[f, 2]] %>%  
+    .[]
+}) %>% 
+  rbindlist() %>% 
+.[bottom_top %in% c(1, 16)] %>% 
+  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
+  .[, cortante := Mag(u_16 - u_1, v_16 - v_1)] %>% 
+  .[, ":="(u_2 = NULL, 
+           u_10 = NULL,
+           v_2 = NULL,
+           v_10 = NULL)]
 ```

### 2021-07-27 temp.R
Scripts

@@ -1,0 +1,114 @@
+library(tidyverse)
+library(lubridate)
+library(data.table)
+library(unglue)
+library(metR)
+library(mesoda)
+library(patchwork)
+source(here::here("temp_R/postprocesamiento.R"))
+
+map_arg <- rnaturalearth::ne_states(country = c("argentina"),
+                                    returnclass = "sf")
+map_limitrofes <- rnaturalearth::ne_countries(country = c("Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")
+
+geom_mapa <- function(fill = NA) {
+  list(geom_sf(data = map_arg, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+       geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
+       coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)),
+       scale_x_longitude(ticks = 5),
+       scale_y_latitude(ticks = 5))
+}
+
+wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
+ini_date <- ymd_hms("20181120180000")
+ciclos <- 67
+
+dates <- seq(ini_date, by = "hour",
+             length.out = ciclos)
+
+
+
+for (d in seq_along(dates)) {
+
+  print(dates[d])
+
+
+  files_ana <- Sys.glob(paste0(wrf_path, "E[2,5,6,8]/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.ensmean"))
+
+  files_gue <- Sys.glob(paste0(wrf_path, "E[2,5,6,8]/GUESS/", format(dates[d], "%Y%m%d%H%M%S"), "/wrfarw.ensmean"))
+
+  ana <- purrr::map(files_ana, function(f) {
+
+    metadatos <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{fecha}/analysis.ensmean")
+
+
+    ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", u = "P")) %>%
+      .[, c("u", "v") := uvmet(f)] %>%
+      .[, ":="(exp = metadatos[[1]][["exp"]],
+               date = metadatos[[1]][["fecha"]])] %>%
+      .[bottom_top == 7]
+  }) %>%
+    rbindlist() %>%
[…]

### 2021-07-27 analysis/chapters/paper_ana_20181122.Rmd
Paper completo por primera vez!

@@
   affiliation: SMN
+- name: Craig S. Schwartz
+  affiliation: NCAR
 date: "`r Sys.Date()`"
@@
   The impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF is evaluated.
-We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November, 22th, 2018.
+We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November 22, 2018.
 The accumulated precipitation observed during the transit of the MCS was quite high, exceeding 200 mm over the northern part of the domain.
@@
 - code: SMN
-  address: Servicio Meteorológico Nacional de Argentina
+  address: Servicio Meteorológico Nacional de Argentina.
 - code: CONICET
-  address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas)
+  address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).
 - code: INPE
-  address: National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies
-bibliography: paper_ANA.bib
+  address: National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.
+- code: NCAR
+  address: National Center for Atmospheric Research, Boulder, Colorado.
+bibliography: 
+  - paper_ANA.bib 
+  - packages.bib
 csl: elsevier-harvard.csl
@@
 
-# arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10)
-# arg_topo[, lon := metR::ConvertLongitude(lon)]
-# 
-# cordillera <- list(
-#   ggnewscale::new_scale_fill(),
-#   geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
-#                                          z = h),
-#                     breaks = seq(1500, 8000, by = 700)),
-#   scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
-#                       oob = scales::squish))
+arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10, 
+                                file.dir = "~/mesoda/analysis/data/derived_data/")
+arg_topo[, lon := metR::ConvertLongitude(lon)]
+
+cordillera <- list(
+  ggnewscale::new_scale_fill(),
+  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat,
+                                         z = h),
+                    breaks = seq(1500, 8000, by = 700)),
+  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
+                      oob = scales::squish))
 
@@
 DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations. 
-However surface observations, which directly measure the atmospheric state, have limited availability in South America. 
+However, surface observations, which directly measure the atmospheric state, have limited availability in South America. 
 Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products. 
-Radiance observations particularly, are indirect measurements and are more difficult to incorporate to DA systems as they require complex forward models.
-
-The assimilation of conventional observations from surface stations have an important role at improving the initial conditions that lead to better forecasts. 
+Radiance observations particularly, are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
+
+The assimilation of conventional observations from surface stations has an important role at improving the initial conditions that lead to better forecasts. 
 For example, @ha2014 showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
 
-Given the sparsity of the local conventional observing network it is important to explore the potential benefits of including additional sources of observations at a regional scale. 
+Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale. 
 Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
@@
 Previous work has shown promising results on mesoscale DA in the region using different approaches such as LETKF; [@dillon2016; @dillonInreview] and 3D-Var [@goncalvesdegoncalves2015]. 
-In particular, the use of ensemble-based methods have shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances.
+In particular, the use of ensemble-based methods has shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances.
 Moreover, the use of ensembles also provides valuable information about the uncertainties in the forecast.
@@
 The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system. 
-The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during 22-23 November 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
-On 22 November a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
+The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
+In Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours. 
@@
 The paper is organized as follows.
-The data assimilation system, the experimental design and observations used is presented in section 2.
+The data assimilation system, the experimental design and observations used are presented in section 2.
 Results are discussed in section 3 and finally, conclusions are summarized in section 4.
@@
 
-The model simulations for the case study were performed using the version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2018) model.
+The model simulations for the case study were performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2018) model.
 The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. 
@@
 In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system.
-We use 9 different model configurations consisting on the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
-All ensemble members used the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class squeme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
-
-(ref:dominio) a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (green box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
-b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOPs.
+We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
+All ensemble members used the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
+
+(ref:dominio) a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions.
 The topography in meters is also shown (filled contours).
@@
   geom_rect(aes(xmin = -66.5, xmax = -61.5, ymin = -35.5, ymax = -29), 
-            color = "#7CAE00", alpha = 0) +
+            color = "#40BDEC", alpha = 0) +
   scale_shape_manual(name = NULL,  
@@
   coord_sf(ylim = c(-35.5, -29), xlim = c(-66.5, -61.5)) +
-  # geom_jitter(data = lista_sondeos, aes(lon, lat, 
-  #                                       color = periodo,
-  #                                       shape = periodo), 
-  #             alpha = 0.5, size = 1.5, width = 0.03, height = 0.03) +
-  # scale_color_brewer(palette = "Dark2",
-  #                    guide = guide_legend(override.aes = list(size = 2,
-  #                                                             alpha = 1))) +
-  # labs(color = "b) Soundings", shape = "b) Soundings") +
   theme_minimal(base_size = 11) +
@@
 
-Satellite-derived wind observations also come from GDAS and includes information from GOES-16 (estimated from the visible, infrared and water vapour channels) and METEOSAT 8 and 11 (visible and water vapour channels).
+Satellite-derived wind observations also come from GDAS and include information from GOES-16 (estimated from the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (visible and water vapor channels).
 Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations).
 
-We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of  various public and private automatic station networks over Southern South America and are included in RELAMPAGO Data Set repository [@garcia2019].
+We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of various public and private automatic station networks over Southern South America and are included in the RELAMPAGO Data Set repository [@garcia2019].
 These stations are indicated as green squares in Figure \@ref(fig:dominio)a.
@@
 
-Table \@ref(tab:table-obs) lists all the observation types (i.e., pressure, temperature, specific humidity and wind) available for each source, together with their associated errors.
+Table \@ref(tab:table-obs) lists all the observation types (i.e., pressure, temperature, specific humidity, and wind) available for each source, together with their associated errors.
 The observation errors were specified following the GSI default configuration.
@@
   # str()
-  kbl(caption = "Observation errors and gross check thresholds for different observation platforms and types.",
+  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and it source, the available variables, the observation error and the gross check thresholds used.",
       col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
@@
   collapse_rows(columns = 1:2, valign = "middle", latex_hline = "major") %>% 
-  footnote(symbol = c("Observation error varied with height",
-                      "Observations over 600 hPa are rejected",
-                      "Observation error depends on the report type"),
+  footnote(symbol = c("Observation error varied with height.",
+                      "Observations over 600 hPa are rejected.",
+                      "Observation error depends on the report type."),
            symbol_manual = c("*", "**", "+"))
@@
 The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM.
-The multispectral sensors have a good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
+The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
 The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.
@@
 GSI has a constant offset bias correction term ($p_0 = 1$) and the following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
-Scan angle-dependent bias is modelled as a 4th-order polynomial [@zhu2014].
-
-In the GSI system the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
-The coefficients initialized from the 18 UTC Nov, 11 2018 GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
+Scan angle-dependent bias is modeled as a 4th-order polynomial [@zhu2014].
+
+In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
+The coefficients initialized from the 18 UTC Nov 11, 2018, GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
 Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.
@@
 It also performs an emissivity check for observations over land for both infrared and microwave radiances.
-
-```{r eval=FALSE, include=FALSE}
-files <- Sys.glob(here("analysis/data/derived_data/count_obs_E*"))[c(1, 3, 4)]
-
-purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
-  .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
-  dcast(bufr_code ~ exp, value.var = "porcentaje") %>% 
-  .[order(.[,'bufr_code']), ] %>% 
-  kbl(booktabs = T, col.names = c("Obs type", "CONV", "AUT", "SATWND"),
-      caption = "Observation types assimilated in each experiment.") %>% 
-  kable_classic(full_width = F)
-
-# purrr::map(files, ~fread(.x)) %>% rbindlist() %>% 
-#   # .[, porcentaje := scales::percent(proporcion, accuracy = 0.01, trim = TRUE)] %>% 
-#   dcast(bufr_code ~ exp, value.var = "N") %>% 
-#   .[order(.[,'bufr_code']), ] %>% 
-#   kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND")) %>% 
-#   kable_classic_2(full_width = F)
-```
 
@@
       col.names = linebreak(c("Sensor", "Plataform", "Assimilated\nchannels", "Percentage\nover total")),
-      caption = "List of the available sensors in several platforms, the number of accepted channels and the percentage of assimilated observations calculated over all radiance observations.") %>% 
+      caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>%
   kable_styling(font_size = 7) %>% 
-  collapse_rows(1,valign = "top") %>% 
-  # column_spec(1, width = "3cm", latex_valign = "p") %>% 
+  collapse_rows(1, valign = "top") %>% 
   kable_classic_2(full_width = TRUE)
@@
 
-(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
+(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
 
@@
 
-To evaluate the performance of the ensemble based data assimilation system presented in this article, we use the following observations datasets:
+To evaluate the performance of the ensemble-based data assimilation system presented in this article, we use the following observations datasets:
 
@@
 The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \@ref(fig:obs-horizontal).
-The larger number of assimilated observations over the center and east of the domain correspond to the ASWS observations.
-In this region we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet.
+The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations.
+In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet.
 In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown.
@@
   ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
-  "PREPBUFR", "x", "x", "x", "x",
-  "ASWS", " ", "x", "x", "x",
-  "SATWND", " ", " ", "x", "x",
-  "RAD", " ", " ", " ", "x",
+  "Conventional (PREPBUFR)", "x", "x", "x", "x",
+  "Conventional (ASWS)", " ", "x", "x", "x",
+  "Satellite-derived winds", " ", " ", "x", "x",
+  "Radiances", " ", " ", " ", "x",
 ) %>% 
@@
       align = "lcccc",
-      caption = "Observation types assimilated in each experiment.") %>% 
+      caption = "Observation types assimilated in each experiment.",
+      escape = FALSE) %>% 
   kable_classic_2(full_width = TRUE) %>% 
-  column_spec(1, width = "5em") %>% 
-  column_spec(2:5, width = "3.5em")
+  column_spec(1, width = "8em") %>% 
+  column_spec(2:3, width = "2.5em", latex_valign = "m") %>% 
+  column_spec(4:5, width = "3em", latex_valign = "m")
 ```
@@
 (ref:cycle) Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
-The zoomed section shows the assimilation for each hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
-Also the two Intensive Observation Periods (IOP) missions from the RELAMPAGO Field Campaign are shown.
+The zoomed section shows the hourly assimilation hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
+The two Intensive Observation Periods (IOP) missions from the RELAMPAGO Field Campaign are shown.
 
@@
 
-where $x_o$ is the observation and it error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
+where $x_o$ is the observation and its error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
 The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
@@
 
-For the comparison of the first-guess’ precipitation with the precipitation estimates we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following @maldonado2021 and adaptations from @roberts2008.
+For the comparison of the first-guess’ precipitation with the precipitation estimates, we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following @maldonado2021 and adaptations from @roberts2008.
 
@@
 
+## Computation procedures
+
+We performed all analysis in this paper using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
+All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2019].
+All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019].
+
 # Results
@@
 To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment.
-As this experiment assimilates all types of observations used in this work, it is possible to analyse the reliability of the ensemble by comparing it with each type of observations.
+As this experiment assimilates all types of observations used in this work, it is possible to analyze the reliability of the ensemble by comparing it with each type of observation.
 Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid.
-The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error.
+The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error.
 For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas to the west of the domain where the $sd RCRV$ can be as high as 4.5.
-In this region the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.
+In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.
 
@@
 Figure \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
-Both ADPUPA and AIRCFR show a general good agreement between the ensemble spread and the observation error.
+Both ADPUPA and AIRCFR show a generally good agreement between the ensemble spread and the observation error.
 As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread.
-AIRCAR presents an irregular profile with $sd RCRV$ values that suggests that the error for this type of observation is overestimated.
-ADPUPA and AIRCAR present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dilloninreview.
-
-For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, where the are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behaviour with a negative bias and an overestimation of the observation error.
+AIRCAR presents an irregular profile with $sd RCRV$ values that suggest that the error for this type of observation is overestimated.
+ADPUPA and AIRCAR present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dillonInreview.
+
+For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behavior with a negative bias and an overestimation of the observation error.
 Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero.
@@
 
-The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) shows almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor.
+The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) show almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor.
 This indicates that the bias correction algorithm works as expected.
@@
 
-(ref:rcrv-sfc) sd RCRV calculated for surface observations of a) wind, and b) temperature, averaged over 2.5º boxes.
+(ref:rcrv-sfc) $sd RCRV$ calculated for surface observations (from PREPBUFR and ASWS) of a) wind, and b) temperature, averaged over 2.5º boxes.
 
@@
   tag_facets() +
-  labs(fill = "sd RCRV") +
+  labs(fill = latex2exp::TeX("sd RCRV")) +
   theme_minimal(base_size = 13) +
@@
 
-(ref:rcrv-profile) Vertical profiles of mean RCRV and sd RCRV for  a) and b) sounding and aircraft observations, c) satellite-derived winds, and d) brightness temperature observations.
+(ref:rcrv-profile) Vertical profiles of $mean RCRV$ (dash line) and $sd RCRV$ (solid line) for  a) temperature and b) wind of sounding and aircraft observations, c) satellite-derived wind observations, and d) brightness temperature observations.
 
@@
   facet_wrap(~ var, scales = "free_x", labeller = labeller(var = c("u" = "U wind",
-                                                                   "v" = "Wind"))) +
+                                                                   "v" = "Satellite-derived\nwind"))) +
   tag_facets(tag_pool = c("c")) +
@@
 
-In this section we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
+In this section, we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
 The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS.
@@
 During the first day, the assimilation of ASWS observations results in a colder PBL.
-This cooling effect has a clear diurnal cycle being stronger during night time (Figure \@ref(fig:TQ-diff)a).
+This cooling effect has a clear diurnal cycle being stronger during nighttime (Figure \@ref(fig:TQ-diff)a).
 During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
@@
 Figure \@ref(fig:TQ-diff)b, indicates that SATWND observations have an almost negligible impact on the mean temperature through the experiment.
-This result is unexpected since mid-to-upper level winds can have a profound impact in the organization and evolution of MCSs over the region.
-During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensate for the cooling effect of ASWS observations.
+This result is unexpected since mid-to-upper level winds can have a profound impact on the organization and evolution of MCSs over the region.
+During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of ASWS observations.
 No clear systematic impact is found above the PBL during this period.
-During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AUT experiment but with opposite sign.
+During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AUT experiment but with the opposite sign.
 
 Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \@ref(fig:TQ-diff)d).
-The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the night time.
+The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the nighttime.
 The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@ruiz2010].
@@
 The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
-There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with stronger reduction of the northerly wind during night hours.
+There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with a stronger reduction of the northerly wind during night hours.
 This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment.
@@
 After the initiation of the convective cells the systematic impact on the wind field is larger at mid and upper levels.
-During Nov 22 and 23 the impact of assimilating ASWS, produces an increase of northerly wind in upper levels.
+During Nov 22 and 23 the impact of assimilating ASWS produces an increase of northerly wind in upper levels.
 This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
-Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
+Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
 
@@
 
-To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles).
+To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles).
 At that moment the first convective cells were developing on the southern region of the domain along the cold front.
@@
 RAD has lower precipitable water content than AUT and SATWND but higher than CONV.
-The distribution of moisture at low levels in RAD seems to be the results of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the centre of the domain compare to CONV.
+The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compare to CONV.
 
@@
 MCAPE in the RAD experiment is lower compared to AUT or SATWND.
-This is consistent with less humidity in the PBL with respect to these experiments and maybe partially compensated by a slightly warmer PBL in the RAD experiment.
-The 0-6 km wind shear is more intense in AUT, SATWND and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
-Moreover in this same region, these experiments show larger MCAPE values.
+This is consistent with less humidity in the PBL with respect to these experiments and may be partially compensated by a slightly warmer PBL in the RAD experiment.
+The 0-6 km wind shear is more intense in AUT, SATWND, and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
+Moreover, in this same region, these experiments show larger MCAPE values.
 Wind shear over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowsky2010].
 
-(ref:TQ-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg-1$).
+(ref:TQ-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a).
 
@@
 
-(ref:UV-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in ms-1) and v wind (d, e, and f in $ms^{-1}$).
+(ref:UV-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a).
 In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f) are shown.
 
-```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=6, fig.align="center", fig.pos="t"}
+```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=6, fig.align="center", fig.pos="th"}
 perfiles %>% 
@@
                     breaks = seq(0, 30, 3), color = "grey30", skip = 1,
-                    size = 4, stroke = 0.1) +
+                    size = 2, stroke = 0.1) +
   labs(fill = latex2exp::TeX("U ($ms^{-1}$)"),
@@
                     breaks = seq(-10, 4, 2), color = "grey30", skip = 1, 
-                    size = 4, stroke = 0.1) +
+                    size = 2, stroke = 0.1) +
   labs(fill = latex2exp::TeX("V ($ms^{-1}$)"),
@@
 
-(ref:summary-fields) Precipitable water (shaded, $kgm^{-2}$) and average northerly wind over the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, $ms^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $ms^{-1}$ for each experiment.
+(ref:summary-fields) a) Precipitable water (shaded, $kgm^{-2}$) and average northerly wind on the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, $ms^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $ms^{-1}$ for each experiment.
 All fields correspond to 00 UTC Nov, 22.
@@
                         labels = c("-10", "-5")) +
-  # geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
-  #                   proj = norargentina_lambert, color = "darkorange", 
-  #              breaks = c(-10, -5), skip = 0,
-  #              label.placement = label_placement_n(1)) +
-  # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
   labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
-  #cordillera +
+  cordillera +
   geom_mapa() +
@@
   theme(legend.position = "bottom",
-        # legend.box.spacing = margin(0, 0, 0, 0),
         panel.ontop = TRUE,
@@
   ggplot(aes(x, y)) +
-  # geom_point(aes(color = v_mean)) +
   geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
@@
   labs(fill = "Potential\ntemperature (K)") +
-  #cordillera +
+  cordillera +
   geom_mapa() +
@@
                                                 frame.colour = "black")) +
-  # geom_contour2(aes(z = cortante),
-  #               proj = norargentina_lambert, color = "black", size = 0.4,
-  #               breaks = c(10, 25)) +
   geom_contour2(aes(z = cortante, size = stat(level)),
@@
                         labels = c("15", "30")) +
-  # geom_point(data = ~.x[cortante >= 15 & is.cross(x, y, skip = 1)], 
-  #            aes(x = lon, y = lat, 
-  #                alpha = factor(cut_round(cortante, breaks = c(15, 25, Inf)))),
-  #            size = 0.00002, shape = 20) +
-  # scale_alpha_manual(values = c(0.3, 0.7), name = NULL, #labels = c(">5 m/s", ">15 m/s"),
-  #                    guide = guide_legend(override.aes = list(size = 2))) +
   labs(fill = latex2exp::TeX("MCAPE ($JKg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($ms^{-1}$)")) +
-  #cordillera +
+  cordillera +
   geom_mapa() +
@@
   theme(legend.position = "bottom",
-        # legend.box.spacing = margin(1, 1, 1, 1),
         panel.ontop = TRUE,
@@
 
-Firstly we analyse the impact of different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \@ref(fig:pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and the estimated precipitation by IMERG.
+Firstly we analyze the impact of different observation types in terms of the representation of the MCS and its associated precipitation.
+Figure \@ref(fig:pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitude range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and the estimated precipitation by IMERG.
 The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period.
-In all the experiments, the accumulated precipitation in the short-range forecasts underestimate the precipitation.
+In all the experiments, the accumulated precipitation in the short-range forecasts underestimates the precipitation.
 This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation.
-AUT, SATWND and RAD better capture the timing and location of convective initiation.
+AUT, SATWND, and RAD better capture the timing and location of convective initiation.
 On RAD the initiation is delayed with respect to AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
-After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the conveccion.
-
-(ref:pp-hov) Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right, for the ensemble mean) calculated over the small domain (red box in Figure \@ref(fig:dominio)a).
+After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
+
+(ref:pp-hov) Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right), for the ensemble mean of each experiment, averaged over the inner domain (red box in Figure \@ref(fig:dominio)a).
+Contours drawn every 0.5 $mmh^{-1}$, starting at 0.5 $mmh^{-1}$.
 
@@
 
-To quantify the spatial match between the observed precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
+To quantify the spatial match between the observed precipitation and the short-range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
 All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22 with an improvement during the mature stage of the MCS.
@@
 
-(ref:fss) FSS calculated over a 6-hours moving window for a) 1 mm and b) 25 mm with a scale of 10 km and c) 1 mm and a d) 25 mm with a scale of 100 km, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
+(ref:fss) FSS calculated over a 6-hours moving window for a) 1 mm and b) 25 mm on a 10 km and c) 1 mm and a d) 25 mm on a 100 km scale, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
 
@@
 These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations.
-In addition, these experiments are the worst and best performing experiments in terms of the short range precipitation forecast skill.
-Overall none of the short range forecasts is able to capture the mesoscale details in the precipitation distribution.
+In addition, these experiments are the worst and best performing experiments in terms of the short-range precipitation forecast skill.
+Overall none of the short-range forecasts is able to capture the mesoscale details in the precipitation distribution.
 This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately capture the strength of the convective band associated with the MCS.
@@
 The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in the RAD experiment.
-CONV captures the location of the MCS, but the convection seems to be less organised and much weaker than in RAD.
+CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
 Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between experiment and observations is quite high in the regions where radar data are available.
@@
 The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection.
-For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that are slightly reduced by the assimilation of ASWS and radiances.
-The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND and RAD is small above the PBL.
-The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND and RAD.
+For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that is slightly reduced by the assimilation of ASWS and radiances.
+The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND, and RAD is small above the PBL.
+The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND, and RAD.
 On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C), indicating a more moist atmosphere in the analysis.
@@
 The zonal wind has a significant positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced by assimilating satellite radiances.
-It is somewhat surprising that no significant effect is observed by assimilating satellite derived winds.
+It is somewhat surprising that no significant effect is observed by assimilating satellite-derived winds.
 The assimilation of ASWS and radiance observations increases the northerly wind bias and RMSE near the surface.
 
-(ref:dbz-mean) Observed maximum reflectivity in the column (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third rows) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22.
-
-```{r dbz-mean, fig.cap="(ref:dbz-mean)", out.width="100%", fig.height=8, fig.width=8, fig.env="figure*", fig.pos="h"}
+(ref:dbz-mean) Observed maximum reflectivity in the column (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third rows) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.
+
+```{r dbz-mean, fig.cap="(ref:dbz-mean)", out.width="100%", fig.height=6.5, fig.width=8, fig.env="figure*", fig.pos="h"}
 proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"
@@
   rbindlist()  %>% 
-  .[, run := fifelse(exp == "E8", "RAD", "CONV")]
+  .[, run := fifelse(exp == "E8", "RAD", "CONV")] %>% 
+  .[, ":="(south_north = NULL,
+           west_east = NULL,
+           exp = NULL)]
 
@@
 
-p1 <- dbz_obs %>% 
-  # .[date == unique(date)[1]] %>% 
-  .[, date := factor(date)] %>% 
+rbind(dbz_obs, dbz_ana) %>% 
+  .[, ":="(date = factor(date),
+           run = fct_relevel(run, "OBSERVATIONS", "CONV", "RAD"))] %>% 
   ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+  geom_contour_fill(data = ~.x[run == "OBSERVATIONS"], aes(z = max_dbz, fill = stat(level)), 
+                    proj = proj_radar,
+                    breaks = values) +
+  geom_contour_fill(data = ~.x[run != "OBSERVATIONS"], aes(z = max_dbz, fill = stat(level)), 
                     proj = proj_radar,
@@
   theme_minimal(base_size = 9) +
-  theme(legend.position = "none",
+  theme(legend.position = "bottom",
         axis.text.x = element_blank(),
@@
 
-p2 <- dbz_ana %>% 
-  .[, date := factor(date)] %>% 
-  ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
-                    proj = norargentina_lambert,
-                    breaks = values) +
-  scale_fill_manual(values = colours, 
-                    labels = function(x) JumpBy(x, 3, fill = ""),
-                    drop = FALSE, 
-                    guide = guide_colorsteps(barwidth = 25,
-                                             barheight = 0.4, 
-                                             title.position = "left", 
-                                             title.vjust = 1,
-                                             frame.colour = "black")) +
-  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
-  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
-  geom_mapa() +
-  facet_grid(run ~ date, labeller = labeller(date = c("2018-11-22 10:00:00" = "10 UTC",
-                                                      "2018-11-22 13:00:00" = "13 UTC",
-                                                      "2018-11-22 16:00:00" = "16 UTC",
-                                                      "2018-11-22 19:00:00" = "19 UTC"))) +
-  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
-  theme_minimal(base_size = 9) +
-  theme(legend.position = "bottom",
-        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        panel.ontop = TRUE,
-        panel.grid = element_line(linetype = 3))
-
-legend <- get_legend(
-  # create some space to the left of the legend
-  p2 + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"), 
-             legend.box.margin = margin(0, 0, 0, 0))
-)
-
-p2 <- p2 + theme(legend.position = "none")
-
-plot_grid(p1 + theme(legend.position = "none",
-                     plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")),
-          p2 + theme(legend.position = "none",
-                     plot.margin = margin(t = -5, r = 0, b = -2, l = 0, unit = "pt")), 
-          legend, ncol = 1, rel_heights = c(1.1, 2, 0.1))
-
 ```
@@
 Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND and RAD to the assimilations of ASWS.
-The RMSE for AUT, SATWND and RAD remains below 2$^{\circ}$C for almost all levels with significant  improvement in upper levels where radiance observations have the largest impact.
+Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilations of ASWS.
+The RMSE for AUT, SATWND, and RAD remains below 2$^{\circ}$C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact.
 In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
@@
 For this period only RAD shows an improvement in the upper troposphere.
-At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND and RAD.
-In fact low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
+At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND, and RAD.
+In fact, low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
 
@@
 
-To summarize these results, firstly, ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly at low levels.
-These changes in the planetary boundary layer can substantially affect the development of the convection.
-Secondly, radiances also produce important impacts but these are, in general, opposite to those of ASWS observatios.
-Spatial fields of different variables at 00 UTC Nov, 22 will be analyzed next to obtain an overview of the environment prior to the development of the MSC.
-
-To summarize, the results suggest that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study.
-While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels.
-The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
+We performed four experiments to assess the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances, into a regional frequent-update ensemble-based data assimilation system.
+We used a case study approach of a massive MCS developed over Southern South America on Nov 22, 2018.
+
+We evaluated the reliability of the ensemble using the RCRV calculated for each type of observation used in RAD.
+Conventional observations including surface observation and upper air observations from soundings and aircraft show a $sd RCRV$ near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
+For satellite-derived winds and radiance observations, the $sd RCRV$ is lower than 1.
+This could be the result of an underdispersive ensemble or an overestimation of the observation errors.
+Taking into account that the errors of this type of observation are usually overestimated to avoid strong impacts on the analysis for poor quality observations and the results obtained for more reliable observations such as soundings, we conclude that the assembly has a reasonable spread.
+
+In terms of the analysis, we found that ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly on the PBL.
+In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed case study.
+We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature state of the MCS leading to an increase in the accumulated precipitation compared with the other experiments.
+However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates.
+While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels available for this case study.
+A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels with a revision of the quality control of these observations.
+
+Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during the IOPs 7 and 8.
+For the pre-convective environment (IOP 7), we observed an improvement on low level when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind.
+In particular, ASWS observations correct a dry bias already observed by @ruiz2010.
+For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels for the temperature but the opposite effect for the dew point temperature.
+In this case, the analyses have a moist bias that increases with the assimilation of ASWS and radiances observations.
+We observed a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
+It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period and they require further analysis.
+
+To summarize this study, we conclude that the assimilation of surface observations with higher spatial and temporal resolution and clear-sky radiances from polar orbit satellites had an overly positive impact on the development of the studied MCS and its associated precipitation.
+In the future, we will analyze whether these impacts carry over in time to the independent forecasts initialized from the analyses.
+
+# Acknowledgements
+
+We are very thankful to the Atmospheric and Sea Research Center (CIMA), the University of Buenos Aires (UBA) and the National Scientific and Technical Research Council (CONICET) who support this study.
+We acknowledge the Argentinian National Meteorological Service (ANMS) for kindly providing the radar observations used for validation.
+We also acknowledge the Cheyenne HPC resources (doi:10.5065/D6RX99HX) from NCAR's Computational and Information Systems Laboratory, National Science Foundation (project code UIUC0012).
+Also, PICT 2017-2033 and PICT 2016-0710 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
 

### 2021-07-27 analysis/chapters/paper_ana_20181122.tex
Paper completo por primera vez!

\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Atmospheric Research} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing

% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis}
    \author[UBA,CIMA,CNRS]{Paola Belén Corrales\corref{1}}
   \ead{paola.corrales@cima.fcen.uba.ar} 
    \author[UBA,CIMA,CNRS]{Victoria Galligani}

    \author[UBA,CIMA,CNRS]{Juan Ruiz}

    \author[INPE]{Luiz Sapucci}

    \author[SMN,CONICET]{María Eugenia Dillon}

    \author[SMN,CONICET,CNRS]{Yanina García Skabar}

    \author[SMN]{Maximiliano Sacco}

    \author[NCAR]{Craig S. Schwartz}

      \address[UBA]{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.}
    \address[CIMA]{CONICET -- Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.}
    \address[CNRS]{CNRS -- IRD -- CONICET -- UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.}
    \address[SMN]{Servicio Meteorológico Nacional de Argentina.}
    \address[CONICET]{CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).}
    \address[INPE]{National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.}
    \address[NCAR]{National Center for Atmospheric Research, Boulder, Colorado.}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  The impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF is evaluated. We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November 22, 2018. The accumulated precipitation observed during the transit of the MCS was quite high, exceeding 200 mm over the northern part of the domain. This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during Nov.~2018.

  We used the GSI-4DLETKF to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance. A multiphysics approach is also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations). We conducted four assimilation experiments using different sets of observations: CONV with conventional observations from prepBUFR, AUT that uses CONV observations plus automatic station networks, SATWND add satellite-derived winds and RAD includes satellite radiances from AMSU, HIRS, MHS, ATMS, AIRS and IASI. We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
  \end{abstract}
   \begin{keyword} Regional Data Assimilation, Surface Obsevations, Satellite Observations\end{keyword}
 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Severe weather events cause significant human and material losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
Southern South America has been shown to have the highest frequency of favorable conditions for high-impact meteorological events (Brooks et al., 2003) and the highest occurrence of large hail events (Cecil and Blankenship, 2012).
This is also confirmed by observational evidence and high impact weather reports (Matsudo et al., 2015; Rasmussen et al., 2014).

Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to several factors.
On the one hand, mesoscale phenomena have limited predictability, while on the other hand, there is great difficulty in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and in the order of minutes).
To solve this problem, data assimilation (DA) at the mesoscale is a topic that has gained increasing interest since it can provide appropriate initial conditions for the initialization of high-resolution forecasts (Sun et al., 2014).

DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations.
However, surface observations, which directly measure the atmospheric state, have limited availability in South America.
Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products.
Radiance observations particularly, are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.

The assimilation of conventional observations from surface stations has an important role at improving the initial conditions that lead to better forecasts.
For example, Ha and Snyder (2014) showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.

Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale.
Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e.~AIRS and IASI) in a regional domain application.
Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature and humidity profile data from the AIRS and found improvement on the tracking of two typhoons.
Dillon et al. (2019) also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation.
Moreover Bao et al. (2015) studied the impact on temperature and humidity forecast over western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. Recently Zhu et al. (2019) studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al., 2015).

Having analyses with higher temporal resolution than those available from global forecasting systems (e.g.~at hourly or sub-hourly) and using an increased number of regional observations that are not assimilated in global DA systems are some of the advantages of regional DA.
Previous work has shown promising results on mesoscale DA in the region using different approaches such as LETKF; (Dillon et al., n.d., 2016) and 3D-Var (gustavo Goncalves de Goncalves et al., 2015).
In particular, the use of ensemble-based methods has shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances.
Moreover, the use of ensembles also provides valuable information about the uncertainties in the forecast.
Recently a rapid-refresh ensemble-based DA system has been experimentally implemented during the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) Field Campaign (central Argentina, November-December 2018, Nesbitt et al. (2021)) with promising results (Dillon et al., n.d.).
This system assimilated data from high-resolution surface networks, AMDAR data from local aircraft flights, soundings, AIRS retrievals, high-resolution GOES-16 wind estimates, and local radar data every hour with a grid spacing of 10 km.
The forecast initialized from the analyses shows an overall similar performance to cold-start initializations from the Global Forecast System analysis, and even a positive impact in some cases.

One important challenge of regional data assimilation systems is the assessment of their representation of the true atmospheric state and the skill of the forecasts initialized using their analysis as initial conditions.
In this sense, the observations and measurements taken during the intense observing period of RELAMPAGO are an invaluable resource to validate a regional data assimilation system.

The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system.
The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
In Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MSC.
The paper is organized as follows. The data assimilation system, the experimental design and observations used are presented in section 2. Results are discussed in section 3 and finally, conclusions are summarized in section 4.

\hypertarget{data-and-methods}{%
\section{Data and Methods}\label{data-and-methods}}

The model simulations for the case study were performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al. (2018)) model.
The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal resolution and 6 hour temporal resolution, National Centers for Environmental Prediction, National Weather Service, NOAA, U.S. Department of Commerce (2015)).
The domain covers the area indicated in Figure \ref{fig:dominio} to capture the development of the MCS during the simulated period.

The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, Hunt et al. (2007)) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Hu et al. (2018)).
A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within \(\pm\) 30 minutes of the analysis time are assimilated.
Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) was used as an observation operator to calculate model-simulated brightness temperatures.

To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in Dillon et al. (n.d.) for all types of observations.
A relaxation-to-prior inflation (Whitaker and Hamill, 2012) is applied with an inflation parameter \(\alpha=0.9\) to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.

We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis. Since the publicly available GFS ensemble members change at each cycle \textbf{(alguna fuente)} we used random perturbations to generate the initial ensemble perturbations. This method also helps to prevent an underestimation of the ensemble spread (Ouaraini et al., 2015). The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in Necker et al. (2020) and Maldonado et al. (2021). In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. These random perturbations are also applied at the boundaries in order to keep the ensemble spread within the domain.

In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system. We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain--Fritsch (Kain, 2004), Grell--Freitas (Grell and Freitas, 2013), and Betts--Miller--Janjic (Janjić, 1994)) and 3 planetary boundary layer schemes (Yonsei University Scheme (Hong, Noh, et al., 2006), Mellor--Yamada--Janjic Scheme (Janjić, 1994), and Mellor--Yamada Nakanishi Niino (Nakanishi and Niino, 2009)). All ensemble members used the same land-surface model (Noah-MP, Chen and Dudhia (2001)), microphysics (WRF single-moment 6--class scheme (Hong, Kim, et al., 2006)), and radiation processes (RRTMG shortwave and longwave scheme (Iacono et al., 2008)) parameterizations.

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions. The topography in meters is also shown (filled contours).}\label{fig:dominio}
\end{figure}

\hypertarget{observations}{%
\subsection{Observations}\label{observations}}

\hypertarget{conventional-and-satellite-derived-wind}{%
\subsubsection{Conventional and Satellite-derived wind}\label{conventional-and-satellite-derived-wind}}

The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS) \textbf{{[}ver cita{]}}. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircrafts. The orange triangles in Figure \ref{fig:dominio}a, are the surface stations and ship locations included in this experiment. The orange triangles in Figure \ref{fig:dominio} a), are the surface station locations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.

Satellite-derived wind observations also come from GDAS and include information from GOES-16 (estimated from the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (visible and water vapor channels). Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 \% of the observations).

We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of various public and private automatic station networks over Southern South America and are included in the RELAMPAGO Data Set repository (Garcia et al., 2019). These stations are indicated as green squares in Figure \ref{fig:dominio}a. They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases. The variables measured by the stations varied between networks; all stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.

Table \ref{tab:table-obs} lists all the observation types (i.e., pressure, temperature, specific humidity, and wind) available for each source, together with their associated errors. The observation errors were specified following the GSI default configuration. In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind). In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and the model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \ref{tab:table-obs}).

\begin{table}

\caption{\label{tab:table-obs}Characteristics of the assimilated observations: The code for each observation type and it source, the available variables, the observation error and the gross check thresholds used.}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5em}>{\raggedright\arraybackslash}p{4.5em}>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{7em}>{\raggedright\arraybackslash}p{7em}}
\toprule
Code & Platform & Variable & Error & Gross check\\
\midrule
 &  & Pressure & 1-1.6 $hPa^*$ & 3.6 $hPa$\\

 &  & Temperature & 1.5 $K$ & 7 $K$\\

 &  & Specific humidity & 20 \% & 8 $gKg^{-1}$\\

\multirow{-4}{3.5em}{\raggedright\arraybackslash CSWS   ASWS} & \multirow{-4}{4.5em}{\raggedright\arraybackslash Surface weather stations} & Wind & 2.2 $ms^{-1}$ & 6 $ms^{-1}$\\
\cmidrule{1-5}
 &  & Pressure & 1.1-1.2 $hPa^{**}$ & 4 $hPa$\\

 &  & Temperature & 0.8-1.5 $K^*$ & 8 $K$\\

 &  & Specific humidity & 20 \% & 8 $gKg^{-1}$\\

\multirow{-4}{3.5em}{\raggedright\arraybackslash ADPUPA} & \multirow{-4}{4.5em}{\raggedright\arraybackslash Radiosondes} & Wind & 1.4-3 $ms^{-1}$* & 8 $ms^{-1}$\\
\cmidrule{1-5}
 &  & Temperature & 1.47-2.5 $K^+$ & 7 $K$\\

\multirow{-2}{3.5em}{\raggedright\arraybackslash AIRCFT} & \multirow{-2}{4.5em}{\raggedright\arraybackslash Aircrafts} & Wind & 2.4-3.6 $ms^{-1+}$ & 6.5-7.5 $ms^{-1+}$\\
\cmidrule{1-5}
ASCATW & Advance Scatterometers & Wind & 1.5 $ms^{-1}$ & 5 $ms^{-1}$\\
\cmidrule{1-5}
 &  & Pressure & 1.3 $hPa$ & 4 $hPa$\\

 &  & Temperature & 2.5 $K$ & 7 $K$\\

 &  & Specific humidity & 20 \% & 8 $gKg^{-1}$\\

\multirow{-4}{3.5em}{\raggedright\arraybackslash SFCSHP} & \multirow{-4}{4.5em}{\raggedright\arraybackslash Ships and Buoys} & Wind & 2.5 $ms^{-1}$ & 5 $ms^{-1}$\\
\cmidrule{1-5}
SATWND & Satellite-derived winds & Wind & 3.8-8 $ms^{-1*+}$ & 1.3-2.5 $ms^{-1+}$\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{*} Observation error varied with height.}\\
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{**} Observations over 600 hPa are rejected.}\\
\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{+} Observation error depends on the report type.}\\
\end{tabular}
\end{table}

\hypertarget{sat}{%
\subsubsection{Satellite radiances}\label{sat}}

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \ref{tab:table-rad}). Since the regional domain is located in the mid-latitudes and all satellite platforms are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. This is evident in Figure \ref{fig:obs-cycle}a which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure. In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 \% of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere. The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.

The channels accepted for assimilation and their associated error were defined using the default GSI configuration. The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. Firstly a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, on the observation quality (based on available data quality information) and on the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred from those over land or snow (Hu et al., 2018).

The thinned observations were then bias corrected. The bias correction (BC) has an air-mass dependent and an angle-dependent component (Zhu et al., 2014) and it is calculated as a multi-linear function of N predictors \(p_i(x)\), with associated coefficients \(\beta_i\). Then, the bias corrected brightness temperature can be obtained as:

\[\mathit{BT_{bc}} =\mathit{ BT} + \sum_{i = 0}^{N} \beta_i p_i (x)\]

GSI has a constant offset bias correction term (\(p_0 = 1\)) and the following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity. Scan angle-dependent bias is modeled as a 4th-order polynomial (Zhu et al., 2014).

In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves for the \(\beta_i\) that provides the best fit between the simulation and the observations. The coefficients initialized from the 18 UTC Nov 11, 2018, GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section. Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.

In our experiments, only clear-sky observations are used. For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes (Weston et al., 2019; Zhu et al., 2016). For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms. Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60\(^{\circ}\)) to reject channels near the visible range that can be contaminated with reflection. It also performs an emissivity check for observations over land for both infrared and microwave radiances.

\begin{table}

\caption{\label{tab:table-rad}List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedright}X}
\toprule
Sensor & Plataform & \makecell[r]{Assimilated\\channels} & \makecell[l]{Percentage\\over total}\\
\midrule
AIRS & AQUA & 118 & 39.37 \%\\
\cmidrule{1-4}
 & AQUA & 7 & 1.13 \%\\
\cmidrule{2-4}
 & N19 & 13 & 1.55 \%\\
\cmidrule{2-4}
 & N15 & 13 & 3.93 \%\\
\cmidrule{2-4}
 & N18 & 13 & 3.17 \%\\
\cmidrule{2-4}
\multirow[t]{-5}{*}{\raggedright\arraybackslash AMSUA} & METOP-A & 13 & 1.3 \%\\
\cmidrule{1-4}
 & N20 & 21 & 1.16 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash ATMS} & NPP & 21 & 0.86 \%\\
\cmidrule{1-4}
 & N19 & 14 & 1.79 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash HIRS4} & METOP-A & 14 & 1.37 \%\\
\cmidrule{1-4}
 & METOP-A & 174 & 40.07 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash IASI} & METOP-B & 174 & 2.94 \%\\
\cmidrule{1-4}
 & N19 & 5 & 0.44 \%\\
\cmidrule{2-4}
 & METOP-A & 5 & 0.46 \%\\
\cmidrule{2-4}
\multirow[t]{-3}{*}{\raggedright\arraybackslash MHS} & METOP-B & 5 & 0.44 \%\\
\bottomrule
\end{tabu}
\end{table}

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/obs-horizontal-1} \caption{Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AUT c) SATWND and d) RAD experiments calculate over 2.5\(^{\circ}\) boxes.}\label{fig:obs-horizontal}
\end{figure}

\begin{figure*}
\includegraphics{../figures/obs-cycle-1} \caption{a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.}\label{fig:obs-cycle}
\end{figure*}

\hypertarget{validation-dataset}{%
\subsubsection{Validation dataset}\label{validation-dataset}}

To evaluate the performance of the ensemble-based data assimilation system presented in this article, we use the following observations datasets:

\begin{itemize}
\item
  The Multi-Network Composite Highest Resolution Radiosonde Data (Earth Observing Laboratory, 2020) from the RELAMPAGO field campaign database consisting of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements. Only the soundings that did not enter the assimilation system were used for validation. The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \ref{fig:dominio}b).
\item
  The Satellite precipitation estimation IMERG Final Run with 0.01\(^{\circ}\) spatial resolution and 30 minutes temporal resolution (Huffman et al., 2018) was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
\item
  Radar observations are used to perform a qualitative and visual assessment of the convection features. The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network (de Elía et al., 2017) with a temporal frequency of 10 minutes. For this work only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
\end{itemize}

\hypertarget{experimental-design}{%
\subsection{Experimental design}\label{experimental-design}}

To investigate the impact of different observations upon the analysis, four data assimilation experiments were performed using different observation sets (Table \ref{tab:table-exp}). The CONV experiment uses only convectional observations from PREPBUFR. In a second experiment, referred to as AUT, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from ASWS. In the third experiment, referred to as SATWND, we assimilate all the observations of the AUT experiment and the satellite-derived motion winds. Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbit satellites as described in Section \ref{sat}.

The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \ref{fig:obs-horizontal}. The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations. In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet. In Figure \ref{fig:obs-cycle}a the number of assimilated observations over time is shown. Local maxima at 12 and 00 UTC found mainly in CONV is produced by the operational soundings. The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage. The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors. The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the ASWS observations. Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa). Above 850 hPa, most of the observations correspond to radiance observations.

All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles). The initial ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \ref{fig:cycle}).

\begin{table}

\caption{\label{tab:table-exp}Observation types assimilated in each experiment.}
\centering
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{8em}>{\centering\arraybackslash}m{2.5em}>{\centering\arraybackslash}m{2.5em}>{\centering\arraybackslash}m{3em}>{\centering\arraybackslash}m{3em}}
\toprule
Obs type & CONV & AUT & SATWND & RAD\\
\midrule
Conventional (PREPBUFR) & x & x & x & x\\
Conventional (ASWS) &  & x & x & x\\
Satellite-derived winds &  &  & x & x\\
Radiances &  &  &  & x\\
\bottomrule
\end{tabu}
\end{table}

\begin{figure}
\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours. The zoomed section shows the hourly assimilation hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours. The two Intensive Observation Periods (IOP) missions from the RELAMPAGO Field Campaign are shown.}\label{fig:cycle}
\end{figure}

\hypertarget{verification-methods}{%
\subsection{Verification methods}\label{verification-methods}}

A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper. These aspects include a validation of how the uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations which are not assimilated.

To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use the Reduced Centered Random Variable (RCRV, Candille et al. (2007)) which is defined as:

\[RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}\]

where \(x_o\) is the observation and its error \(\sigma_o\), the mean \(m\), and the standard deviation \(\sigma\) of the ensemble.
The average of \(RCRV\) computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:

\[\mathit{mean RCRV} = E[RCRV]\]

The standard deviation of the \(RCRV\) or \(sd RCRV\) measures the agreement of the ensemble spread and the observational error with respect to the distance between the ensemble mean and the observations, and then the systematic over- or under- dispersion of the ensemble:

\[\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}\]

where \(M\) is the sample size. A reliable system will have no bias (\(mean RCRV = 0\)), and a standard deviation equal to 1 (\(sd RCRV = 1\)). If the ensemble has a positive (negative) bias, \(mean RCRV\) will be positive (negative). A \(sd RCRV > 1\) (\(sd RCRV < 1\)), indicates that the ensemble is underdispersive (overdispersive).

The fit of the first-guess and analysis to a set of independent observations is computed based on the Root Mean Square Error (RMSE) and the BIAS:

\[\mathit{RMSE} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)^{2}}\]

\[\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)\]

where \(O\) and \(X\) stand for independent observations and the analysis respectively.

For the comparison of the first-guess' precipitation with the precipitation estimates, we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al. (2021) and adaptations from Roberts (2008).

\[\mathit{FSS} = 1-\frac{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i)^{2}+\frac{1}{N}\sum_{i=1}^{N} ({P_o}_i)^{2}} \]

where \({P_x}\) is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and \({P_o}\) is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise. This probabilistic formulation uses the information from the full ensemble thus avoiding the smoothing effect of the ensemble mean which can particularly affect localized extreme precipitation values. The FSS was applied to the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

\hypertarget{computation-procedures}{%
\subsection{Computation procedures}\label{computation-procedures}}

We performed all analysis in this paper using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2019; Xie, 2015). All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{ensemble-reliability}{%
\subsection{Ensemble reliability}\label{ensemble-reliability}}

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the \(mean RCRV\) and the \(sd RCRV\) for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyze the reliability of the ensemble by comparing it with each type of observation. Figure 5 shows the \(sd RCRV\) for surface observations box-averaged to a 2.5° grid. The \(sd RCRV\) for wind observations (Figure \ref{fig:rcrv-sfc}a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error. For the temperature (Figure \ref{fig:rcrv-sfc}b), the results are similar except that for some areas to the west of the domain where the \(sd RCRV\) can be as high as 4.5. In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.

Figure \ref{fig:rcrv-profile} shows the mean and standard deviation of the RCRV for the upper-air observations. Figure \ref{fig:rcrv-profile}a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT). Both ADPUPA and AIRCFR show a generally good agreement between the ensemble spread and the observation error. As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread. AIRCAR presents an irregular profile with \(sd RCRV\) values that suggest that the error for this type of observation is overestimated. ADPUPA and AIRCAR present a \(mean RCRV\) profile near zero on middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in Ruiz et al. (2010) and Dillon et al. (n.d.).

For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) in low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) shows a larger departure from the expected behavior with a negative bias and an overestimation of the observation error. Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero. The only exception are the EUMETSAT observations which contribute very little in the region.

The mean RCRV profiles calculated from the radiance observations (Figure \ref{fig:rcrv-profile}d) show almost no bias and the same happens if the \(mean RCRV\) is calculated over each channel of each sensor. This indicates that the bias correction algorithm works as expected. The \(sd RCRV\) values are less than 1 for all sensors possibly due to an overestimation of the observation errors to reduce the influence of potentially erroneous observations.

Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work. Moreover, these results suggest the inflation parameter \(\beta = 0.9\) is adequate for the system.

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/rcrv-sfc-1} \caption{\(sd RCRV\) calculated for surface observations (from PREPBUFR and ASWS) of a) wind, and b) temperature, averaged over 2.5º boxes.}\label{fig:rcrv-sfc}
\end{figure}

\begin{figure*}
\includegraphics{../figures/rcrv-profile-1} \caption{Vertical profiles of \(mean RCRV\) (dash line) and \(sd RCRV\) (solid line) for a) temperature and b) wind of sounding and aircraft observations, c) satellite-derived wind observations, and d) brightness temperature observations.}\label{fig:rcrv-profile}
\end{figure*}

\hypertarget{impacts-of-assimilated-observations}{%
\subsection{Impacts of assimilated observations}\label{impacts-of-assimilated-observations}}

In this section, we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection. The analysis is performed over a smaller domain (red box in Figure \ref{fig:dominio}a) to focus on the region most directly affected by the MCS. Figure \ref{fig:TQ-diff}a-c shows the difference between experiments in the spatially averaged vertical profile of temperature. By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state. During the first day, the assimilation of ASWS observations results in a colder PBL. This cooling effect has a clear diurnal cycle being stronger during nighttime (Figure \ref{fig:TQ-diff}a). During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS. The warm difference shown in AUT with respect to CONV in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV. This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection. Figure \ref{fig:TQ-diff}b, indicates that SATWND observations have an almost negligible impact on the mean temperature through the experiment. This result is unexpected since mid-to-upper level winds can have a profound impact on the organization and evolution of MCSs over the region.
During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of ASWS observations. No clear systematic impact is found above the PBL during this period. During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AUT experiment but with the opposite sign.

Comparing the specific humidity in the experiments (Figure \ref{fig:TQ-diff}d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \ref{fig:TQ-diff}d). The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the nighttime. The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region (Ruiz et al., 2010). We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL. In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections. As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels. Figure \ref{fig:TQ-diff}f shows that radiances reduce low-middle level moisture during the first day of the experiment. The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov, 22.

The impacts on the wind components are shown in Figure Figure \ref{fig:UV-diff} along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.

The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis. There is a diurnal cycle in the impact of surface weather stations on the meridional velocity (Figure \ref{fig:UV-diff}d) with a stronger reduction of the northerly wind during night hours. This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment. After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain. After the initiation of the convective cells the systematic impact on the wind field is larger at mid and upper levels. During Nov 22 and 23 the impact of assimilating ASWS produces an increase of northerly wind in upper levels. This could be a consequence of a stronger MCS with an increased polar side upper level outflow. Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.

The assimilation of radiances produced a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c and f). For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of ASWS observations. On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind. When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in Figure \ref{fig:UV-diff}f.~

To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles). At that moment the first convective cells were developing on the southern region of the domain along the cold front. Figure \ref{fig:summary-fields}a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours). We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations. The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place. AUT and SATWND experiments are very similar, with values of precipitable water over 55 \(kgm^{-2}\) north of 30\(^{\circ}\)S and a similar vertical distribution of specific humidity (not shown). RAD has lower precipitable water content than AUT and SATWND but higher than CONV. The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compare to CONV.

The potential temperature on the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed on the temperature profiles in Figure \ref{fig:TQ-diff}a-c.~On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL than AUT due to the assimilation of radiance observations. Figure \ref{fig:summary-fields}c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 \(km\) wind shear. The values of MCAPE in CONV do not exceed 2000 \(JKg^{-1}\) while the rest of the experiments show maximum MCAPE over 4000 \(JKg^{-1}\). MCAPE in the RAD experiment is lower compared to AUT or SATWND. This is consistent with less humidity in the PBL with respect to these experiments and may be partially compensated by a slightly warmer PBL in the RAD experiment. The 0-6 km wind shear is more intense in AUT, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values. Moreover, in this same region, these experiments show larger MCAPE values. Wind shear over 15 \(ms^{-1}\) have been associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowsky and Richardson, 2010).

\begin{figure*}[t]

{\centering \includegraphics{../figures/TQ-diff-1}

}

\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a).}\label{fig:TQ-diff}
\end{figure*}

\begin{figure*}[th]

{\centering \includegraphics{../figures/UV-diff-1}

}

\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a). In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f) are shown.}\label{fig:UV-diff}
\end{figure*}

\begin{figure*}
\includegraphics[width=1\linewidth]{../figures/summary-fields-1} \caption{a) Precipitable water (shaded, \(kgm^{-2}\)) and average northerly wind on the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, \(ms^{-1}\)), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and \textasciitilde0-6 km wind shear over 15 and 30 \(ms^{-1}\) for each experiment. All fields correspond to 00 UTC Nov, 22.}\label{fig:summary-fields}
\end{figure*}

\hypertarget{verification}{%
\subsection{Verification}\label{verification}}

Firstly we analyze the impact of different observation types in terms of the representation of the MCS and its associated precipitation. Figure \ref{fig:pp-hov} shows the first-guess hourly accumulated precipitation averaged over a longitude range between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments, and the estimated precipitation by IMERG. The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period. In all the experiments, the accumulated precipitation in the short-range forecasts underestimates the precipitation. This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation. AUT, SATWND, and RAD better capture the timing and location of convective initiation. On RAD the initiation is delayed with respect to AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue. After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.

\begin{figure*}[h]
\includegraphics{../figures/pp-hov-1} \caption{Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right), for the ensemble mean of each experiment, averaged over the inner domain (red box in Figure \ref{fig:dominio}a). Contours drawn every 0.5 \(mmh^{-1}\), starting at 0.5 \(mmh^{-1}\).}\label{fig:pp-hov}
\end{figure*}

To quantify the spatial match between the observed precipitation and the short-range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}). All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22 with an improvement during the mature stage of the MCS. The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS. AUT and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study. The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22. The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).

\begin{figure}
\centering
\includegraphics{../figures/fss-1.png}
\caption{\label{fig:fss}FSS calculated over a 6-hours moving window for a) 1 mm and b) 25 mm on a 10 km and c) 1 mm and a d) 25 mm on a 100 km scale, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.}
\end{figure}

To complement the analysis, Figure \ref{fig:dbz-mean} shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity. These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations. In addition, these experiments are the worst and best performing experiments in terms of the short-range precipitation forecast skill. Overall none of the short-range forecasts is able to capture the mesoscale details in the precipitation distribution. This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately capture the strength of the convective band associated with the MCS. Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field. RAD better represents the observed features of the MSC showing a more strong and organized MCS over the domain center at 10 and 13 UTC (first and second columns in Figure \ref{fig:dbz-mean}). The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in the RAD experiment. CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD. Before and after the times shown in Figure \ref{fig:dbz-mean}, the agreement between experiment and observations is quite high in the regions where radar data are available.

Finally, Figure \ref{fig:soundings} shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes). IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments. The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection. For the temperature, CONV presents a warm bias on the PBL of 1\(^{\circ}\)C and an RMSE between 1 and 2.5\(^{\circ}\)C that is slightly reduced by the assimilation of ASWS and radiances. The RMSE for temperature remains between 0.5 and 2\(^{\circ}\)C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND, and RAD is small above the PBL. The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND, and RAD. On middle levels (below 10 km) the bias is large and positive (8 - 10\(^{\circ}\)C), indicating a more moist atmosphere in the analysis. Overall, the RMSE grows with height, this is probably because at upper levels, moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures. The zonal wind has a significant positive bias of approximately 10 \(ms^{-1}\) in the upper troposphere which is mainly reduced by assimilating satellite radiances. It is somewhat surprising that no significant effect is observed by assimilating satellite-derived winds. The assimilation of ASWS and radiance observations increases the northerly wind bias and RMSE near the surface.

\begin{figure*}[h]
\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Observed maximum reflectivity in the column (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third rows) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.}\label{fig:dbz-mean}
\end{figure*}

For the IOP 8 (bottom row in Figure \ref{fig:soundings}), the densely observed area was behind the MCS, but far enough from it to be directly affected by its mesoscale circulations. This area was also behind the cold front thus affected by low-level cold advection. Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances. Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilations of ASWS. The RMSE for AUT, SATWND, and RAD remains below 2\(^{\circ}\)C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact. In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations. This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region. The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL. SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7\(^{\circ}\)C. The zonal wind is overestimated on the analyses with a positive bias at middle levels. For this period only RAD shows an improvement in the upper troposphere. At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND, and RAD. In fact, low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).

\begin{figure*}[t]

{\centering \includegraphics{../figures/soundings-1}

}

\caption{RMSE (solid line) and Bias (dash line) of a) temperature, b) dew point temperature, c) u wind and d) v wind calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8. The blue line correspond to CONV, the ligh blue line to AUT, SATWND is represented with an orange line and RAD with a red line.}\label{fig:soundings}
\end{figure*}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

We performed four experiments to assess the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances, into a regional frequent-update ensemble-based data assimilation system. We used a case study approach of a massive MCS developed over Southern South America on Nov 22, 2018.

We evaluated the reliability of the ensemble using the RCRV calculated for each type of observation used in RAD. Conventional observations including surface observation and upper air observations from soundings and aircraft show a \(sd RCRV\) near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations. For satellite-derived winds and radiance observations, the \(sd RCRV\) is lower than 1. This could be the result of an underdispersive ensemble or an overestimation of the observation errors. Taking into account that the errors of this type of observation are usually overestimated to avoid strong impacts on the analysis for poor quality observations and the results obtained for more reliable observations such as soundings, we conclude that the assembly has a reasonable spread.

In terms of the analysis, we found that ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly on the PBL. In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed case study. We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature state of the MCS leading to an increase in the accumulated precipitation compared with the other experiments. However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates. While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels available for this case study. A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels with a revision of the quality control of these observations.

Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during the IOPs 7 and 8. For the pre-convective environment (IOP 7), we observed an improvement on low level when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind. In particular, ASWS observations correct a dry bias already observed by Ruiz et al. (2010). For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels for the temperature but the opposite effect for the dew point temperature. In this case, the analyses have a moist bias that increases with the assimilation of ASWS and radiances observations. We observed a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL. It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period and they require further analysis.

To summarize this study, we conclude that the assimilation of surface observations with higher spatial and temporal resolution and clear-sky radiances from polar orbit satellites had an overly positive impact on the development of the studied MCS and its associated precipitation. In the future, we will analyze whether these impacts carry over in time to the independent forecasts initialized from the analyses.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We are very thankful to the Atmospheric and Sea Research Center (CIMA), the University of Buenos Aires (UBA) and the National Scientific and Technical Research Council (CONICET) who support this study. We acknowledge the Argentinian National Meteorological Service (ANMS) for kindly providing the radar observations used for validation. We also acknowledge the Cheyenne HPC resources (doi:10.5065/D6RX99HX) from NCAR's Computational and Information Systems Laboratory, National Science Foundation (project code UIUC0012). Also, PICT 2017-2033 and PICT 2016-0710 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-allaire2019}{}%
Allaire, J., Horner, J., Xie, Y., Marti, V., Porte, N., 2019. Markdown: Render markdown with the c library 'sundown'.

\leavevmode\hypertarget{ref-bao2015}{}%
Bao, Y., Xu, J., Powell Jr., A.M., Shao, M., Min, J., Pan, Y., 2015. Impacts of AMSU-A, MHS and IASI data assimilation on temperature and humidity forecasts with GSI--WRF over the western United States. Atmos. Meas. Tech. 8, 4231--4242. doi:\href{https://doi.org/10.5194/amt-8-4231-2015}{10.5194/amt-8-4231-2015}

\leavevmode\hypertarget{ref-brooks2003}{}%
Brooks, H.E., Lee, J.W., Craven, J.P., 2003. The spatial distribution of severe thunderstorm and tornado environments from global reanalysis data. Atmospheric Research 67-68, 73--94. doi:\href{https://doi.org/10.1016/S0169-8095(03)00045-0}{10.1016/S0169-8095(03)00045-0}

\leavevmode\hypertarget{ref-campitelli2020}{}%
Campitelli, E., 2020. metR: Tools for Easier Analysis of Meteorological Fields.

\leavevmode\hypertarget{ref-candille2007}{}%
Candille, G., Côté, C., Houtekamer, P.L., Pellerin, G., 2007. Verification of an Ensemble Prediction System against Observations. Mon. Wea. Rev. 135, 2688--2699. doi:\href{https://doi.org/10.1175/MWR3414.1}{10.1175/MWR3414.1}

\leavevmode\hypertarget{ref-cecil2012}{}%
Cecil, D.J., Blankenship, C.B., 2012. Toward a Global Climatology of Severe Hailstorms as Estimated by Satellite Passive Microwave Imagers. Journal of Climate 25, 687--703. doi:\href{https://doi.org/10.1175/JCLI-D-11-00130.1}{10.1175/JCLI-D-11-00130.1}

\leavevmode\hypertarget{ref-chen2001}{}%
Chen, F., Dudhia, J., 2001. Coupling an Advanced Land Surface--Hydrology Model with the Penn State--NCAR MM5 Modeling System. Part I: Model Implementation and Sensitivity. Monthly Weather Review 129, 569--585. doi:\href{https://doi.org/10.1175/1520-0493(2001)129\%3C0569:CAALSH\%3E2.0.CO;2}{10.1175/1520-0493(2001)129\textless0569:CAALSH\textgreater2.0.CO;2}

\leavevmode\hypertarget{ref-chen2015}{}%
Chen, Q., Fan, J., Hagos, S., Gustafson, W.I., Berg, L.K., 2015. Roles of wind shear at different vertical levels: Cloud system organization and properties. Journal of Geophysical Research: Atmospheres 120, 6551--6574. doi:\href{https://doi.org/10.1002/2015JD023253}{10.1002/2015JD023253}

\leavevmode\hypertarget{ref-Cheyenne2019}{}%
Computational and Information Systems Laboratory, 2019. Cheyenne: HPE/SGI ICE XA System (University Community Computing). National Center for Atmospheric Research Boulder, CO. doi:\href{https://doi.org/doi:10.5065/D6RX99HX}{doi:10.5065/D6RX99HX}

\leavevmode\hypertarget{ref-deelia2017}{}%
de Elía, R., Vidal, L., Lohigorry, P., 2017. El SMN y la red argentina de radares meteorológicos {[}WWW Document{]}. URL \url{http://hdl.handle.net/20.500.12160/625}

\leavevmode\hypertarget{ref-dillon2019}{}%
Dillon, M.E., García Skabar, Y., Kalnay, E., Ruiz, J.J., Collini, E.A., 2019. Sensibilidad de un sistema de asimilación de datos por ensambles a diferentes configuraciones, implementado en el sur de Sudamérica. Meteorológica 44, 15--34.

\leavevmode\hypertarget{ref-dillonInreview}{}%
Dillon, M.E., Maldonado, P., Corrales, P., García Skabar, Y., Ruiz, J.J., Sacco, M., Cutraro, F., Mingari, L., Matsudo, C., Vidal, L., Rugna, M., Hobouchian, M.P., Salio, P., Nesbitt, S., Saulo, C., Kalnay, E., Miyoshi, T., n.d. A Rapid Refresh ensemble based Data Assimilation and Forecast system for the RELAMPAGO field campaign. Atmospheric Research.

\leavevmode\hypertarget{ref-dillon2016}{}%
Dillon, M.E., Skabar, Y.G., Ruiz, J., Kalnay, E., Collini, E.A., Echevarría, P., Saucedo, M., Miyoshi, T., Kunii, M., 2016. Application of the WRF-LETKF Data Assimilation System over Southern South America: Sensitivity to Model Physics. Wea. Forecasting 31, 217--236. doi:\href{https://doi.org/10.1175/WAF-D-14-00157.1}{10.1175/WAF-D-14-00157.1}

\leavevmode\hypertarget{ref-dowle2020}{}%
Dowle, M., Srinivasan, A., 2020. Data.Table: Extension of 'data.frame'.

\leavevmode\hypertarget{ref-sondeos}{}%
Earth Observing Laboratory, U. -, 2020. Multi-network composite highest resolution radiosonde data. Version 1.3. UCAR/NCAR - earth observing laboratory.

\leavevmode\hypertarget{ref-gao2015}{}%
Gao, F., Huang, X.-Y., Jacobs, N.A., Wang, H., 2015. Assimilation of wind speed and direction observations: Results from real observation experiments. Tellus A: Dynamic Meteorology and Oceanography 67, 27132. doi:\href{https://doi.org/10.3402/tellusa.v67.27132}{10.3402/tellusa.v67.27132}

\leavevmode\hypertarget{ref-garcia2019}{}%
Garcia, F., Ruiz, J., Salio, P., Bechis, H., Nesbitt, S., 2019. Argentina mesonet data. Version 1.1. UCAR/NCAR - earth observing laboratory.

\leavevmode\hypertarget{ref-grell2013}{}%
Grell, G.A., Freitas, S.R., 2013. A scale and aerosol aware stochastic convective parameterization for weather and air quality modeling. Atmos. Chem. Phys. Discuss. 13, 23845--23893. doi:\href{https://doi.org/10.5194/acpd-13-23845-2013}{10.5194/acpd-13-23845-2013}

\leavevmode\hypertarget{ref-goncalvesdegoncalves2015}{}%
gustavo Goncalves de Goncalves, L., Sapucci, L., Vendrasco, E., de Mattos, J.G., Ferreira, C., Khamis, E., Cruz, N., 2015. A rapid update data assimilation cycle over South America using 3DVar and EnKF, in: The 20th International TOVS Study Conference (ITSC-20). The 20th International TOVS Study Conference (ITSC-20), Lake Geneva, Wisconsin, USA.

\leavevmode\hypertarget{ref-ha2014}{}%
Ha, S.-Y., Snyder, C., 2014. Influence of Surface Observations in Mesoscale Data Assimilation Using an Ensemble Kalman Filter. Monthly Weather Review 142, 1489--1508. doi:\href{https://doi.org/10.1175/MWR-D-13-00108.1}{10.1175/MWR-D-13-00108.1}

\leavevmode\hypertarget{ref-han2006}{}%
Han, Y., Van Delst, P., Liu, Q., Weng, F., Yan, B., Treadon, R., Derber, J., 2006. JCSDA Community Radiative Transfer Model (CRTM)---version 1, NOAA Technical Report NESDIS 122. Washington, D.C.

\leavevmode\hypertarget{ref-hong2006a}{}%
Hong, S.-Y., Kim, J.-H., Lim, J.-o., Dudhia, J., 2006. The WRF Single Moment 6-Class Microphysics Scheme (WSM6). Journal of the Korean Meteorological Society 42, 129--151.

\leavevmode\hypertarget{ref-hong2006}{}%
Hong, S.-Y., Noh, Y., Dudhia, J., 2006. A New Vertical Diffusion Package with an Explicit Treatment of Entrainment Processes. Mon. Wea. Rev. 134, 2318--2341. doi:\href{https://doi.org/10.1175/MWR3199.1}{10.1175/MWR3199.1}

\leavevmode\hypertarget{ref-hu2018}{}%
Hu, M., Ge, G., Zhou, C., Stark, D., Shao, H., Newman, K., Beck, J., Zhang, X., 2018. Grid-point Statistical Interpolation (GSI) User's Guide Version 3.7. Developmental Testbed Center.

\leavevmode\hypertarget{ref-huffman2018}{}%
Huffman, G., Bolvin, D., Braithwaite, D., Hsu, K., Joyce, R., Kidd, C., Nelkin, E., Sorooshian, S., Tan, J., Xie, P., 2018. NASA Global Precipitation Measurement (GPM) Integrated Multi-satellitE Retrievals for GPM (IMERG). National Aeronautics and Space Administration (NASA).

\leavevmode\hypertarget{ref-hunt2007}{}%
Hunt, B.R., Kostelich, E.J., Szunyogh, I., 2007. Efficient data assimilation for spatiotemporal chaos: A local ensemble transform Kalman filter. Physica D: Nonlinear Phenomena 230, 112--126. doi:\href{https://doi.org/10.1016/j.physd.2006.11.008}{10.1016/j.physd.2006.11.008}

\leavevmode\hypertarget{ref-iacono2008}{}%
Iacono, M.J., Delamere, J.S., Mlawer, E.J., Shephard, M.W., Clough, S.A., Collins, W.D., 2008. Radiative forcing by long-lived greenhouse gases: Calculations with the AER radiative transfer models. J. Geophys. Res. 113, D13103. doi:\href{https://doi.org/10.1029/2008JD009944}{10.1029/2008JD009944}

\leavevmode\hypertarget{ref-janjic1994}{}%
Janjić, Z.I., 1994. The Step-Mountain Eta Coordinate Model: Further Developments of the Convection, Viscous Sublayer, and Turbulence Closure Schemes. Mon. Wea. Rev. 122, 927--945. doi:\href{https://doi.org/10.1175/1520-0493(1994)122\%3C0927:TSMECM\%3E2.0.CO;2}{10.1175/1520-0493(1994)122\textless0927:TSMECM\textgreater2.0.CO;2}

\leavevmode\hypertarget{ref-jones2013}{}%
Jones, T.A., Otkin, J.A., Stensrud, D.J., Knopfmeier, K., 2013. Assimilation of Satellite Infrared Radiances and Doppler Radar Observations during a Cool Season Observing System Simulation Experiment. Mon. Wea. Rev. 141, 3273--3299. doi:\href{https://doi.org/10.1175/MWR-D-12-00267.1}{10.1175/MWR-D-12-00267.1}

\leavevmode\hypertarget{ref-kain2004}{}%
Kain, J.S., 2004. The Kain--Fritsch Convective Parameterization: An Update. JOURNAL OF APPLIED METEOROLOGY 43, 12.

\leavevmode\hypertarget{ref-lin2017a}{}%
Lin, H., Weygandt, S.S., Benjamin, S.G., Hu, M., 2017. Satellite Radiance Data Assimilation within the Hourly Updated Rapid Refresh. Wea. Forecasting 32, 1273--1287. doi:\href{https://doi.org/10.1175/WAF-D-16-0215.1}{10.1175/WAF-D-16-0215.1}

\leavevmode\hypertarget{ref-maldonado2021}{}%
Maldonado, P., Ruiz, J., Saulo, C., 2021. Sensitivity to Initial and Boundary Perturbations in Convective-Scale Ensemble-Based Data Assimilation: Imperfect-Model OSSEs. SOLA 17, 96--102. doi:\href{https://doi.org/10.2151/sola.2021-015}{10.2151/sola.2021-015}

\leavevmode\hypertarget{ref-markowsky2010}{}%
Markowsky, P., Richardson, Y., 2010. Organization of Isolated Convection, in: Mesoscale Meteorology in Midlatitudes. John Wiley \& Sons, Ltd, pp. 201--244. doi:\href{https://doi.org/10.1002/9780470682104.ch8}{10.1002/9780470682104.ch8}

\leavevmode\hypertarget{ref-matsudo2015}{}%
Matsudo, C., García Skabar, Y., Ruiz, J., Vidal, L., Salio, P., 2015. Verification of WRF-ARW convective-resolving forecasts over Southeastern South America. Mausam 66, 445--456.

\leavevmode\hypertarget{ref-miyoshi2012a}{}%
Miyoshi, T., Kunii, M., 2012. Using AIRS retrievals in the WRF-LETKF system to improve regional numerical weather prediction. Tellus A: Dynamic Meteorology and Oceanography 64, 18408. doi:\href{https://doi.org/10.3402/tellusa.v64i0.18408}{10.3402/tellusa.v64i0.18408}

\leavevmode\hypertarget{ref-nakanishi2009}{}%
Nakanishi, M., Niino, H., 2009. Development of an Improved Turbulence Closure Model for the Atmospheric Boundary Layer. JMSJ 87, 895--912. doi:\href{https://doi.org/10.2151/jmsj.87.895}{10.2151/jmsj.87.895}

\leavevmode\hypertarget{ref-cisl_rda_ds084.1}{}%
National Centers for Environmental Prediction, National Weather Service, NOAA, U.S. Department of Commerce, 2015. NCEP GFS 0.25 degree global forecast grids historical archive.

\leavevmode\hypertarget{ref-necker2020}{}%
Necker, T., Geiss, S., Weissmann, M., Ruiz, J., Miyoshi, T., Lien, G., 2020. A convective‐scale 1,000‐member ensemble simulation and potential applications. Q.J.R. Meteorol. Soc 146, 1423--1442. doi:\href{https://doi.org/10.1002/qj.3744}{10.1002/qj.3744}

\leavevmode\hypertarget{ref-nesbitt2021}{}%
Nesbitt, S.W., Salio, P.V., Ávila, E., Bitzer, P., Carey, L., Chandrasekar, V., Deierling, W., Dominguez, F., Dillon, M.E., Garcia, C.M., Gochis, D., Goodman, S., Hence, D.A., Kosiba, K.A., Kumjian, M.R., Lang, T., Luna, L.M., Marquis, J., Marshall, R., McMurdie, L.A., Nascimento, E.L., Rasmussen, K.L., Roberts, R., Rowe, A.K., Ruiz, J.J., Sabbas, E.F.M.T.S., Saulo, A.C., Schumacher, R.S., Skabar, Y.G., Machado, L.A.T., Trapp, R.J., Varble, A., Wilson, J., Wurman, J., Zipser, E.J., Arias, I., Bechis, H., Grover, M.A., 2021. A storm safari in Subtropical South America: Proyecto RELAMPAGO. Bulletin of the American Meteorological Society -1, 1--64. doi:\href{https://doi.org/10.1175/BAMS-D-20-0029.1}{10.1175/BAMS-D-20-0029.1}

\leavevmode\hypertarget{ref-ouaraini2015}{}%
Ouaraini, R.E., Berre, L., Fischer, C., Sayouty, E.H., 2015. Sensitivity of regional ensemble data assimilation spread to perturbations of lateral boundary conditions. Tellus A: Dynamic Meteorology and Oceanography 67, 28502. doi:\href{https://doi.org/10.3402/tellusa.v67.28502}{10.3402/tellusa.v67.28502}

\leavevmode\hypertarget{ref-rasmussen2014}{}%
Rasmussen, K.L., Zuluaga, M.D., Houze, R.A., 2014. Severe convection and lightning in subtropical South America. Geophysical Research Letters 41, 7359--7366. doi:\href{https://doi.org/10.1002/2014GL061767}{10.1002/2014GL061767}

\leavevmode\hypertarget{ref-rcoreteam2020}{}%
R Core Team, 2020. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria.

\leavevmode\hypertarget{ref-roberts2008}{}%
Roberts, N., 2008. Assessing the spatial and temporal variation in the skill of precipitation forecasts from an NWP model. Meteorological Applications 15, 163--169. doi:\href{https://doi.org/10.1002/met.57}{10.1002/met.57}

\leavevmode\hypertarget{ref-ruiz2010}{}%
Ruiz, J.J., Saulo, C., Nogués-Paegle, J., 2010. WRF Model Sensitivity to Choice of Parameterization over South America: Validation against Surface Variables. Mon. Wea. Rev. 138, 3342--3355. doi:\href{https://doi.org/10.1175/2010MWR3358.1}{10.1175/2010MWR3358.1}

\leavevmode\hypertarget{ref-singh2016}{}%
Singh, R., Ojha, S.P., Kishtawal, C.M., Pal, P.K., Kiran Kumar, A.S., 2016. Impact of the assimilation of INSAT-3D radiances on short-range weather forecasts: Assimilation of INSAT-3D Radiances. Q.J.R. Meteorol. Soc. 142, 120--131. doi:\href{https://doi.org/10.1002/qj.2636}{10.1002/qj.2636}

\leavevmode\hypertarget{ref-skamarock2018}{}%
Skamarock, W.C., Klemp, J.B., Dudhia, J., Gill, D.O., Barker, D.M., Duda, M.G., Huang, X.-Y., Wang, W., Powers, J.G., 2018. A Description of the Advanced Research WRF Version 3.

\leavevmode\hypertarget{ref-sun2014}{}%
Sun, J., Xue, M., Wilson, J.W., Zawadzki, I., Ballard, S.P., Onvlee-Hooimeyer, J., Joe, P., Barker, D.M., Li, P.-W., Golding, B., Xu, M., Pinto, J., 2014. Use of NWP for Nowcasting Convective Precipitation: Recent Progress and Challenges. Bulletin of the American Meteorological Society 95, 409--426. doi:\href{https://doi.org/10.1175/BAMS-D-11-00263.1}{10.1175/BAMS-D-11-00263.1}

\leavevmode\hypertarget{ref-weston2019}{}%
Weston, P., Geer, A., Bormann, N., Bormann, N., 2019. Investigations into the assimilation of AMSU-A in the presence of cloud and precipitation {[}WWW Document{]}. EUMETSAT/ECMWF Fellowship Programme Research Report. doi:\href{https://doi.org/10.21957/ewahn9ce}{10.21957/ewahn9ce}

\leavevmode\hypertarget{ref-whitaker2012}{}%
Whitaker, J.S., Hamill, T.M., 2012. Evaluating Methods to Account for System Errors in Ensemble Data Assimilation. Mon. Wea. Rev. 140, 3078--3089. doi:\href{https://doi.org/10.1175/MWR-D-11-00276.1}{10.1175/MWR-D-11-00276.1}

\leavevmode\hypertarget{ref-wickham2009}{}%
Wickham, H., 2009. Ggplot2: Elegant Graphics for Data Analysis, Use R! Springer-Verlag, New York. doi:\href{https://doi.org/10.1007/978-0-387-98141-3}{10.1007/978-0-387-98141-3}

\leavevmode\hypertarget{ref-xie2015}{}%
Xie, Y., 2015. Dynamic documents with R and knitr, Second. ed. Chapman and Hall/CRC, Boca Raton, Florida.

\leavevmode\hypertarget{ref-zhu2019}{}%
Zhu, K., Xue, M., Pan, Y., Hu, M., Benjamin, S.G., Weygandt, S.S., Lin, H., 2019. The Impact of Satellite Radiance Data Assimilation within a Frequently Updated Regional Forecast System Using a GSI-based Ensemble Kalman Filter. Adv. Atmos. Sci. 36, 1308--1326. doi:\href{https://doi.org/10.1007/s00376-019-9011-3}{10.1007/s00376-019-9011-3}

\leavevmode\hypertarget{ref-zhu2014}{}%
Zhu, Y., Derber, J., Collard, A., Dee, D., Treadon, R., Gayno, G., Jung, J.A., 2014. Enhanced radiance bias correction in the National Centers for Environmental Prediction's Gridpoint Statistical Interpolation data assimilation system. Quarterly Journal of the Royal Meteorological Society 140, 1479--1492. doi:\href{https://doi.org/10.1002/qj.2233}{10.1002/qj.2233}

\leavevmode\hypertarget{ref-zhu2016}{}%
Zhu, Y., Liu, E., Mahajan, R., Thomas, C., Groff, D., Van Delst, P., Collard, A., Kleist, D., Treadon, R., Derber, J.C., 2016. All-Sky Microwave Radiance Assimilation in NCEP's GSI Analysis System. Mon. Wea. Rev. 144, 4709--4735. doi:\href{https://doi.org/10.1175/MWR-D-15-0445.1}{10.1175/MWR-D-15-0445.1}

\end{document}

### 2021-08-17 analysis/chapters/paper_ana_20181122.Rmd
Enviado al resto de los autores

@@
 ---
-title: "Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis"
+title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis"
 author:
@@
 abstract: |
-  The impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF is evaluated.
+  In this paper we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF.
 We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November 22, 2018.
-The accumulated precipitation observed during the transit of the MCS was quite high, exceeding 200 mm over the northern part of the domain.
-This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during Nov.
-2018.
-
-  We used the GSI-4DLETKF to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
+The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
+This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
+
+  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
 A multiphysics approach is also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
-We conducted four assimilation experiments using different sets of observations: CONV with conventional observations from prepBUFR, AUT that uses CONV observations plus automatic station networks, SATWND add satellite-derived winds and RAD includes satellite radiances from AMSU, HIRS, MHS, ATMS, AIRS and IASI.
-We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
+We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-
 journal: Atmospheric Research
@@
   cache = TRUE,
+  cache.extra = 42,
   fig.path = "../figures/",
@@
 Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products. 
-Radiance observations particularly, are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
-
-The assimilation of conventional observations from surface stations has an important role at improving the initial conditions that lead to better forecasts. 
+Radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
+
+The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts. 
 For example, @ha2014 showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
@@
 Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
-@miyoshi2012a assimilated observations, atmospheric temperature and humidity profile data from the AIRS and found improvement on the tracking of two typhoons. 
+@miyoshi2012a assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons. 
 @dillon2019 also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation. 
-Moreover @bao2015 studied the impact on temperature and humidity forecast over western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
-Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
+Moreover, @bao2015 studied the impact on temperature and humidity forecast over the western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
 Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts [@gao2015].
@@
 
-The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
+The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
 The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system. 
 The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
-In Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
+On Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours. 
@@
 The paper is organized as follows.
-The data assimilation system, the experimental design and observations used are presented in section 2.
+The data assimilation system, the experimental design, and the observations used are presented in section 2.
 Results are discussed in section 3 and finally, conclusions are summarized in section 4.
@@
 The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. 
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal resolution and 6 hour temporal resolution, @cisl_rda_ds084.1). 
+The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal resolution and 6-hour temporal resolution, @cisl_rda_ds084.1). 
 The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.
@@
 We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis.
-Since the publicly available GFS ensemble members change at each cycle **(alguna fuente)** we used random perturbations to generate the initial ensemble perturbations.
+Since the publicly available GFS ensemble members change at each cycle we used random perturbations to generate the initial ensemble perturbations.
 This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
@@
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
-These random perturbations are also applied at the boundaries in order to keep the ensemble spread within the domain.
+These random perturbations are also applied at the boundaries to keep the ensemble spread within the domain.
 
@@
 
-The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS) **[ver cita]**.
-We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircrafts.
+The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS).
+We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
 The orange triangles in Figure \@ref(fig:dominio)a, are the surface stations and ship locations included in this experiment.
@@
 Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study.
-This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \@ref(tab:table-rad)).
-Since the regional domain is located in the mid-latitudes and all satellite platforms are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
+Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
 For this reason, the number of satellite observations varied significantly among cycles.
-This is evident in Figure \@ref(fig:obs-cycle)a which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
+This is evident in Figure 3 a) which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
 In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiment.
@@
 The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
-The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.
+The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top (50 hPa) used in the experiments.
 
@@
 Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, on the observation quality (based on available data quality information) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
-Also, observations over the sea are preferred from those over land or snow [@hu2018].
+Also, observations over the sea are preferred over those over land or snow [@hu2018].
 
@@
 For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
-Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60$^{\circ}$) to reject channels near the visible range that can be contaminated with reflection.
+Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
 It also performs an emissivity check for observations over land for both infrared and microwave radiances.
@@
 
-* The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
+* The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
 Only the soundings that did not enter the assimilation system were used for validation.
@@
 The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes.
-For this work only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
+For this work, only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
 
@@
 A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper.
-These aspects include a validation of how the uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations which are not assimilated.
+These aspects include a validation of how the uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations that are not assimilated.
 
@@
 We performed all analysis in this paper using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
-All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2019].
+All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2020].
 All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019].
@@
 The impacts on the wind components are shown in Figure Figure \@ref(fig:UV-diff) along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
-
 The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
@@
 After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain.
-After the initiation of the convective cells the systematic impact on the wind field is larger at mid and upper levels.
+After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels.
 During Nov 22 and 23 the impact of assimilating ASWS produces an increase of northerly wind in upper levels.
@@
 
-## Verification
+## Validation against independent observations
 
@@
 AUT, SATWND, and RAD better capture the timing and location of convective initiation.
-On RAD the initiation is delayed with respect to AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
+On RAD the initiation is delayed when we compare with AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
 After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
@@
 To quantify the spatial match between the observed precipitation and the short-range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
-All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22 with an improvement during the mature stage of the MCS.
+All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature stage of the MCS.
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
@@
 
-To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity.
-These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations.
+To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed maximum reflectivity.
+These experiments were chosen because they represent the analysis with the minimum and the maximum number of assimilated observations.
 In addition, these experiments are the worst and best performing experiments in terms of the short-range precipitation forecast skill.
@@
 In terms of the analysis, we found that ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly on the PBL.
-In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed case study.
+In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
 We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature state of the MCS leading to an increase in the accumulated precipitation compared with the other experiments.
@@
 While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels available for this case study.
-A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels with a revision of the quality control of these observations.
+A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels.
 

### 2021-08-17 analysis/chapters/paper_ana_20181122.tex
Enviado al resto de los autores

@@
             pdfauthor={},
-            pdftitle={Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis},
+            pdftitle={Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis},
             colorlinks=false,
@@
 
-  \title{Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis}
+  \title{Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis}
     \author[UBA,CIMA,CNRS]{Paola Belén Corrales\corref{1}}
@@
   \begin{abstract}
-  The impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF is evaluated.
+  In this paper we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF.
 We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November 22, 2018.
-The accumulated precipitation observed during the transit of the MCS was quite high, exceeding 200 mm over the northern part of the domain.
-This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during Nov.~2018.
-
-  We used the GSI-4DLETKF to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
+The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
+This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
+
+  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
 A multiphysics approach is also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
-We conducted four assimilation experiments using different sets of observations: CONV with conventional observations from prepBUFR, AUT that uses CONV observations plus automatic station networks, SATWND add satellite-derived winds and RAD includes satellite radiances from AMSU, HIRS, MHS, ATMS, AIRS and IASI.
-We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed case study.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
+We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
@@
 Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products.
-Radiance observations particularly, are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
-
-The assimilation of conventional observations from surface stations has an important role at improving the initial conditions that lead to better forecasts.
+Radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
+
+The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts.
 For example, Ha and Snyder (2014) showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
@@
 Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e.~AIRS and IASI) in a regional domain application.
-Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature and humidity profile data from the AIRS and found improvement on the tracking of two typhoons.
+Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons.
 Dillon et al. (2019) also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation.
-Moreover Bao et al.
-(2015) studied the impact on temperature and humidity forecast over western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+Moreover, Bao et al.
+(2015) studied the impact on temperature and humidity forecast over the western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
 Recently Zhu et al.
-(2019) studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
+(2019) studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
 Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al., 2015).
@@
 
-The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
+The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
 The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system.
 The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
-In Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
+On Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
@@
 The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal resolution and 6 hour temporal resolution, National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
+The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal resolution and 6-hour temporal resolution, National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
 Department of Commerce (2015)).
@@
 We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis.
-Since the publicly available GFS ensemble members change at each cycle \textbf{(alguna fuente)} we used random perturbations to generate the initial ensemble perturbations.
+Since the publicly available GFS ensemble members change at each cycle we used random perturbations to generate the initial ensemble perturbations.
 This method also helps to prevent an underestimation of the ensemble spread (Ouaraini et al., 2015).
@@
 
-The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS) \textbf{{[}ver cita{]}}.
-We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircrafts.
+The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS).
+We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
 The orange triangles in Figure \ref{fig:dominio}a, are the surface stations and ship locations included in this experiment.
@@
 Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study.
-This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \ref{tab:table-rad}).
-Since the regional domain is located in the mid-latitudes and all satellite platforms are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
+Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
 For this reason, the number of satellite observations varied significantly among cycles.
-This is evident in Figure \ref{fig:obs-cycle}a which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
+This is evident in Figure 3 a) which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
 In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 \% of the total amount of assimilated radiances in our experiment.
@@
 Firstly a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, on the observation quality (based on available data quality information) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
-Also, observations over the sea are preferred from those over land or snow (Hu et al., 2018).
+Also, observations over the sea are preferred over those over land or snow (Hu et al., 2018).
 
@@
 For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
-Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60\(^{\circ}\)) to reject channels near the visible range that can be contaminated with reflection.
+Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
 It also performs an emissivity check for observations over land for both infrared and microwave radiances.
@@
 \item
-  The Multi-Network Composite Highest Resolution Radiosonde Data (Earth Observing Laboratory, 2020) from the RELAMPAGO field campaign database consisting of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
+  The Multi-Network Composite Highest Resolution Radiosonde Data (Earth Observing Laboratory, 2020) from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
 Only the soundings that did not enter the assimilation system were used for validation.
@@
 The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network (de Elía et al., 2017) with a temporal frequency of 10 minutes.
-For this work only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
+For this work, only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
 \end{itemize}
@@
 We performed all analysis in this paper using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
-All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2019; Xie, 2015).
+All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2020; Xie, 2015).
 All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019).
@@
 The impacts on the wind components are shown in Figure Figure \ref{fig:UV-diff} along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
-
 The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
@@
 
-\hypertarget{verification}{%
-\subsection{Verification}\label{verification}}
+\hypertarget{validation-against-independent-observations}{%
+\subsection{Validation against independent observations}\label{validation-against-independent-observations}}
 
@@
 To quantify the spatial match between the observed precipitation and the short-range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}).
-All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22 with an improvement during the mature stage of the MCS.
+All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature stage of the MCS.
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
@@
 
-To complement the analysis, Figure \ref{fig:dbz-mean} shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity.
-These experiments were chosen because they represent the analysis with minimum and maximum number of assimilated observations.
+To complement the analysis, Figure \ref{fig:dbz-mean} shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed maximum reflectivity.
+These experiments were chosen because they represent the analysis with the minimum and the maximum number of assimilated observations.
 In addition, these experiments are the worst and best performing experiments in terms of the short-range precipitation forecast skill.
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-allaire2019}{}%
-Allaire, J., Horner, J., Xie, Y., Marti, V., Porte, N., 2019. Markdown: Render markdown with the c library 'sundown'.
+\leavevmode\hypertarget{ref-allaire2020}{}%
+Allaire, J.J., Xie {[}aut, Y., cre, McPherson, J., Luraschi, J., Ushey, K., Atkins, A., Wickham, H., Cheng, J., Chang, W., Iannone, R., Dunning, A., filter), A.Y.(.
+sections L., Schloerke, B., Dervieux, C., Aust, F., Allen, J., Seo, J., Barrett, M., Hyndman, R., Lesur, R., Storey, R., Arslan, R., Oller, S., RStudio, PBC, library), j.F.(., inst/rmd/h/jquery-AUTHORS.txt), j.
+contributors (.
+library; authors listed in, inst/rmd/h/jqueryui-AUTHORS.txt), j.U.
+contributors (.U.
+library; authors listed in, library), M.O.(., library), J.T.(., library), B.
+contributors (., Twitter, library), I.(., library), A.F.(., library), S.J.(.
+js, library), I.S.(.
+js, library), G.F.(., templates), J.M.(., Google, library), I.(., library), D.R.(., library), W.(., Gandy (Font-Awesome), D., Sperry (Ionicons), B., (Ionicons), D., StickyTabs), A.L.(., filter), B.P.J.(.L., filter), A.K.(.L., 2020.
+Rmarkdown: Dynamic Documents for R.
 

### 2021-08-24 analysis/chapters/paper_ana_20181122.Rmd
Completo comentarios de Maru, Fig 13 ahora de 0 a 15 km

@@
   We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
-A multiphysics approach is also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
+A multiphysics approach was also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
 We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
@@
 DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations. 
-However, surface observations, which directly measure the atmospheric state, have limited availability in South America. 
-Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products. 
-Radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
-
 The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts. 
 For example, @ha2014 showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
-
+However, surface observations, which directly measure the atmospheric state, have limited availability in South America.
 Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale.
+Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products. 
 Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
-@miyoshi2012a assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons. 
-@dillon2019 also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation. 
+@miyoshi2012a assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons near Japan. 
+@dillon2019 also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation over Southern South America.
+
+However radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
 Moreover, @bao2015 studied the impact on temperature and humidity forecast over the western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
@@
 
-The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
+The main objective of this work is to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
 The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system. 
@@
 
-The model simulations for the case study were performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2018) model.
-The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. 
+The model simulations for the case study are performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2018) model.
+The resolution is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. 
 The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal resolution and 6-hour temporal resolution, @cisl_rda_ds084.1). 
@@
 Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. 
-For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) was used as an observation operator to calculate model-simulated brightness temperatures.
+For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures.
 
@@
 We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
-All ensemble members used the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
+All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
 
@@
 b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions.
-The topography in meters is also shown (filled contours).
+The topography in meters is also shown (shaded).
 
@@
 The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS).
-We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
-The orange triangles in Figure \@ref(fig:dominio)a, are the surface stations and ship locations included in this experiment.
+We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at National Centers for Environmental Prediction (NCEP) that consist of surface observations from 117 conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
 The orange triangles in Figure \@ref(fig:dominio) a), are the surface station locations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.
+Wind surface observations over oceans (ASCATW) come from scaterometters and are also part of the PREPBUFR files.
 
@@
   # str()
-  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and it source, the available variables, the observation error and the gross check thresholds used.",
+  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.",
       col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
@@
 For this reason, the number of satellite observations varied significantly among cycles.
-This is evident in Figure 3 a) which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
+This is evident in Figure 3 a) which shows the number of  observations assimilated for RAD experiment (see Section \@ref(exp)) at each analysis cycle.
 In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiment.
@@
 
-In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
-The coefficients initialized from the 18 UTC Nov 11, 2018, GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
+In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves the $\beta_i$ that provides the best fit between the simulation and the observations.
+The coefficients were initialized at 18 UTC Nov 11 with the GFS system coefficients, and were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
 Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.
@@
       escape = FALSE,
-      col.names = linebreak(c("Sensor", "Plataform", "Assimilated\nchannels", "Percentage\nover total")),
+      col.names = linebreak(c("Sensor", "Platform", "Assimilated\nchannels", "Percentage\nover total")),
       caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>%
@@
 
-## Experimental design
+## Experimental design {#exp}
 
@@
 The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations.
-In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet.
+In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet [@vera2006].
 In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown.
@@
 
-(ref:cycle) Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
-The zoomed section shows the hourly assimilation hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
-The two Intensive Observation Periods (IOP) missions from the RELAMPAGO Field Campaign are shown.
+(ref:cycle) Diagram of the analysis cycles between Nov 20, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
+The zoomed section shows the hourly assimilation that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
+The two IOPs missions from the RELAMPAGO Field Campaign are shown.
 
@@
 
-where $O$ and $X$ stand for independent observations and the analysis respectively.
-
-For the comparison of the first-guess’ precipitation with the precipitation estimates, we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following @maldonado2021 and adaptations from @roberts2008.
+where $O$ and $X$ stand for independent observations and the simulations respectively.
+
+For the comparison of the first-guess’ precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS) for different sized sampling areas and thresholds, following @maldonado2021 and adaptations from @roberts2008:
 
@@
 
-We performed all analysis in this paper using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
-All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2020].
-All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019].
+We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019].
+All the analysis in this paper were performed using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
+All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2019].
 
@@
 Figure \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
-Both ADPUPA and AIRCFR show a generally good agreement between the ensemble spread and the observation error.
+Both ADPUPA and AIRCFT show a generally good agreement between the ensemble spread and the observation error.
 As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread.
@@
 
-For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behavior with a negative bias and an overestimation of the observation error.
+For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias and an overestimation of the observation error.
 Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero.
@@
 
-The impacts on the wind components are shown in Figure Figure \@ref(fig:UV-diff) along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
+The impacts on the wind components are shown in Figure \@ref(fig:UV-diff) along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
 The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
@@
 This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
-Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
-
-The assimilation of radiances produced a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure  \@ref(fig:UV-diff)c and f).
+Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
+
+The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure  \@ref(fig:UV-diff)c and f).
 For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of ASWS observations.
 On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind.
-When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in  Figure \@ref(fig:UV-diff)f.
+When we look at the meridional wind field at 200 hPa at different times, we observe that the outflow from the MCS is even more intense than in the other experiments, while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in  Figure \@ref(fig:UV-diff)f.
 
@@
 RAD has lower precipitable water content than AUT and SATWND but higher than CONV.
-The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compare to CONV.
+The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
 
@@
 The 0-6 km wind shear is more intense in AUT, SATWND, and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
-Moreover, in this same region, these experiments show larger MCAPE values.
-Wind shear over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowsky2010].
-
-(ref:TQ-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a).
-
-```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=6, fig.align="center",fig.pos="t"}
+Moreover, in this same region, these experiments show larger MCAPE values than CONV.
+Note that wind shears over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowsky2010].
+
+(ref:TQ-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+
+```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=6, fig.align="center",fig.pos="th"}
 files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
@@
 
-(ref:UV-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a).
+(ref:UV-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
 In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f) are shown.
 
-```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=6, fig.align="center", fig.pos="th"}
+```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=6, fig.align="center", fig.pos="ht"}
 perfiles %>% 
@@
 The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period.
-In all the experiments, the accumulated precipitation in the short-range forecasts underestimates the precipitation.
+In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated.
 This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation.
 AUT, SATWND, and RAD better capture the timing and location of convective initiation.
-On RAD the initiation is delayed when we compare with AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
+On RAD the initiation is delayed when we compare with AUT and SATWND, possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
 After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
 
-(ref:pp-hov) Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right), for the ensemble mean of each experiment, averaged over the inner domain (red box in Figure \@ref(fig:dominio)a).
+(ref:pp-hov) Hövmoller of mean hourly accumulated precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67$^{\circ}$W and 54.5$^{\circ}$W.
 Contours drawn every 0.5 $mmh^{-1}$, starting at 0.5 $mmh^{-1}$.
@@
 The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
-The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).
-
-(ref:fss) FSS calculated over a 6-hours moving window for a) 1 mm and b) 25 mm on a 10 km and c) 1 mm and a d) 25 mm on a 100 km scale, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
+The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).
+
+(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
 
@@
 
-To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed maximum reflectivity.
+To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity.
 These experiments were chosen because they represent the analysis with the minimum and the maximum number of assimilated observations.
-In addition, these experiments are the worst and best performing experiments in terms of the short-range precipitation forecast skill.
-Overall none of the short-range forecasts is able to capture the mesoscale details in the precipitation distribution.
-This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately capture the strength of the convective band associated with the MCS.
+In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the short-range precipitation forecast skill (Figure \@ref(fig:pp-hov)).
+Overall none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
+This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately represent the strength of the convective band associated with the MCS.
 Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field.
-RAD better represents the observed features of the MSC showing a more strong and organized MCS over the domain center at 10 and 13 UTC (first and second columns in Figure \@ref(fig:dbz-mean)).
-The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in the RAD experiment.
-CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
-Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between experiment and observations is quite high in the regions where radar data are available.
+RAD better represents the observed features of the system showing a  stronger and more organized MCS than CONV, over the domain center at 10 and 13 UTC (first and second columns in Figure \@ref(fig:dbz-mean)).
+The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in RAD.
+In addition, CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
+Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between experiment and observations is quite high in the regions where radar data are available particularly for RAD.
 
 Finally, Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+
 IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments.
@@
 The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND, and RAD.
-On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C), indicating a more moist atmosphere in the analysis.
-Overall, the RMSE grows with height, this is probably because at upper levels, moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
+On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C) for all the experiments, indicating a more moist atmosphere in the analyses than the observed.
+Overall, the RMSE grows with height, probably because at upper levels moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
 The zonal wind has a significant positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced by assimilating satellite radiances.
@@
 
-(ref:dbz-mean) Observed maximum reflectivity in the column (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third rows) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.
+(ref:dbz-mean) Maximum reflectivity in the column (COLMAX in $dBZ$), observed (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.
 
@@
 
-For the IOP 8 (bottom row in Figure \@ref(fig:soundings)), the densely observed area was behind the MCS, but far enough from it to be directly affected by its mesoscale circulations.
+For the IOP 8 (bottom row in Figure \@ref(fig:soundings)), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulations.
 This area was also behind the cold front thus affected by low-level cold advection.
-Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilations of ASWS.
+Below 5 km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
+Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilation of ASWS.
 The RMSE for AUT, SATWND, and RAD remains below 2$^{\circ}$C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact.
@@
 The zonal wind is overestimated on the analyses with a positive bias at middle levels.
-For this period only RAD shows an improvement in the upper troposphere.
+For this period only RAD shows an improvement with respect to CONV in the upper troposphere.
 At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND, and RAD.
-In fact, low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
-
-(ref:soundings) RMSE (solid line) and Bias (dash line) of a) temperature, b) dew point temperature, c) u wind and d) v wind calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
-The blue line correspond to CONV, the ligh blue line to AUT, SATWND is represented with an orange line and RAD with a red line.
+In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
+
+(ref:soundings) RMSE (solid line) and Bias (dash line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+The blue line correspond to CONV, the light blue line to AUT, SATWND is represented with an orange line and RAD with a red line.
 
@@
   melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est") %>% 
+  .[lev <= 15000] %>% 
   ggplot(aes(lev/1000, value.est)) +
@@
 
-We performed four experiments to assess the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances, into a regional frequent-update ensemble-based data assimilation system.
+We performed four experiments to assess the impact of different observation types, including conventional data (CONV), frequent surface observations (AUT), satellite-derived winds (SATWND), and clear-sky satellite radiances (RAD), into a regional frequent-update ensemble-based data assimilation system.
 We used a case study approach of a massive MCS developed over Southern South America on Nov 22, 2018.
@@
 We evaluated the reliability of the ensemble using the RCRV calculated for each type of observation used in RAD.
-Conventional observations including surface observation and upper air observations from soundings and aircraft show a $sd RCRV$ near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
-For satellite-derived winds and radiance observations, the $sd RCRV$ is lower than 1.
-This could be the result of an underdispersive ensemble or an overestimation of the observation errors.
-Taking into account that the errors of this type of observation are usually overestimated to avoid strong impacts on the analysis for poor quality observations and the results obtained for more reliable observations such as soundings, we conclude that the assembly has a reasonable spread.
-
-In terms of the analysis, we found that ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly on the PBL.
+Conventional observations including surface data and upper air observations from soundings and aircraft, show a $sd RCRV$ near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
+For satellite-derived winds and radiance observations, the $sd RCRV$ is lower than 1, which could be the result of an underdispersive ensemble or an overestimation of the observation errors.
+However, taking into account that the errors of this type of observation are usually overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the assembly has a reasonable spread.
+
+In terms of the analysis, we found that ASWS observations, which have high spatial and temporal resolution, produce major impacts throughout the troposphere and mainly on the PBL.
 In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
-We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature state of the MCS leading to an increase in the accumulated precipitation compared with the other experiments.
+We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
 However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates.
@@
 Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during the IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement on low level when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind.
+For the pre-convective environment (IOP 7), we observed an improvement on low levels when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind.
 In particular, ASWS observations correct a dry bias already observed by @ruiz2010.
-For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels for the temperature but the opposite effect for the dew point temperature.
+For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels including more observations for the temperature but the opposite effect for the dew point temperature.
 In this case, the analyses have a moist bias that increases with the assimilation of ASWS and radiances observations.
 We observed a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
-It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period and they require further analysis.
+It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period, which requires further analysis.
 
@@
 
-We are very thankful to the Atmospheric and Sea Research Center (CIMA), the University of Buenos Aires (UBA) and the National Scientific and Technical Research Council (CONICET) who support this study.
+We are very thankful to the Atmospheric and Sea Research Center (CIMA), the University of Buenos Aires (UBA), and the National Scientific and Technical Research Council (CONICET) who support this study.
 We acknowledge the Argentinian National Meteorological Service (ANMS) for kindly providing the radar observations used for validation.
 We also acknowledge the Cheyenne HPC resources (doi:10.5065/D6RX99HX) from NCAR's Computational and Information Systems Laboratory, National Science Foundation (project code UIUC0012).
-Also, PICT 2017-2033 and PICT 2016-0710 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
+Also, PICT 2017-2033,  PICT 2016-0710 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
 

### 2021-08-24 analysis/chapters/paper_ana_20181122.tex
Completo comentarios de Maru, Fig 13 ahora de 0 a 15 km

@@
   We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
-A multiphysics approach is also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
+A multiphysics approach was also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
 We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
@@
 DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations.
-However, surface observations, which directly measure the atmospheric state, have limited availability in South America.
-Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products.
-Radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
-
 The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts.
 For example, Ha and Snyder (2014) showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
-
+However, surface observations, which directly measure the atmospheric state, have limited availability in South America.
 Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale.
+Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products.
 Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e.~AIRS and IASI) in a regional domain application.
-Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons.
-Dillon et al. (2019) also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation.
+Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons near Japan.
+Dillon et al. (2019) also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation over Southern South America.
+
+However radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
 Moreover, Bao et al.
@@
 
-The main objectives of this work are: to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
+The main objective of this work is to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
 The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system.
@@
 
-The model simulations for the case study were performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al.
+The model simulations for the case study are performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al.
 (2018)) model.
-The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
+The resolution is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
 The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal resolution and 6-hour temporal resolution, National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
@@
 Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
-For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) was used as an observation operator to calculate model-simulated brightness temperatures.
+For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) is used as an observation operator to calculate model-simulated brightness temperatures.
 
@@
 We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain--Fritsch (Kain, 2004), Grell--Freitas (Grell and Freitas, 2013), and Betts--Miller--Janjic (Janjić, 1994)) and 3 planetary boundary layer schemes (Yonsei University Scheme (Hong, Noh, et al., 2006), Mellor--Yamada--Janjic Scheme (Janjić, 1994), and Mellor--Yamada Nakanishi Niino (Nakanishi and Niino, 2009)).
-All ensemble members used the same land-surface model (Noah-MP, Chen and Dudhia (2001)), microphysics (WRF single-moment 6--class scheme (Hong, Kim, et al., 2006)), and radiation processes (RRTMG shortwave and longwave scheme (Iacono et al., 2008)) parameterizations.
+All ensemble members use the same land-surface model (Noah-MP, Chen and Dudhia (2001)), microphysics (WRF single-moment 6--class scheme (Hong, Kim, et al., 2006)), and radiation processes (RRTMG shortwave and longwave scheme (Iacono et al., 2008)) parameterizations.
 
@@
 b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions.
-The topography in meters is also shown (filled contours).}\label{fig:dominio}
+The topography in meters is also shown (shaded).}\label{fig:dominio}
 \end{figure}
@@
 The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS).
-We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
-The orange triangles in Figure \ref{fig:dominio}a, are the surface stations and ship locations included in this experiment.
+We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at National Centers for Environmental Prediction (NCEP) that consist of surface observations from 117 conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
 The orange triangles in Figure \ref{fig:dominio} a), are the surface station locations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.
+Wind surface observations over oceans (ASCATW) come from scaterometters and are also part of the PREPBUFR files.
 
@@
 
-\caption{\label{tab:table-obs}Characteristics of the assimilated observations: The code for each observation type and it source, the available variables, the observation error and the gross check thresholds used.}
+\caption{\label{tab:table-obs}Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.}
 \centering
@@
 For this reason, the number of satellite observations varied significantly among cycles.
-This is evident in Figure 3 a) which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure.
+This is evident in Figure 3 a) which shows the number of observations assimilated for RAD experiment (see Section \ref{exp}) at each analysis cycle.
 In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 \% of the total amount of assimilated radiances in our experiment.
@@
 
-In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves for the \(\beta_i\) that provides the best fit between the simulation and the observations.
-The coefficients initialized from the 18 UTC Nov 11, 2018, GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
+In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves the \(\beta_i\) that provides the best fit between the simulation and the observations.
+The coefficients were initialized at 18 UTC Nov 11 with the GFS system coefficients, and were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
 Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.
@@
 \toprule
-Sensor & Plataform & \makecell[r]{Assimilated\\channels} & \makecell[l]{Percentage\\over total}\\
+Sensor & Platform & \makecell[r]{Assimilated\\channels} & \makecell[l]{Percentage\\over total}\\
 \midrule
@@
 
-\hypertarget{experimental-design}{%
-\subsection{Experimental design}\label{experimental-design}}
+\hypertarget{exp}{%
+\subsection{Experimental design}\label{exp}}
 
@@
 The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations.
-In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet.
+In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet (Vera et al., 2006).
 In Figure \ref{fig:obs-cycle}a the number of assimilated observations over time is shown.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
-The zoomed section shows the hourly assimilation hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
-The two Intensive Observation Periods (IOP) missions from the RELAMPAGO Field Campaign are shown.}\label{fig:cycle}
+\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Diagram of the analysis cycles between Nov 20, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
+The zoomed section shows the hourly assimilation that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
+The two IOPs missions from the RELAMPAGO Field Campaign are shown.}\label{fig:cycle}
 \end{figure}
@@
 
-where \(O\) and \(X\) stand for independent observations and the analysis respectively.
-
-For the comparison of the first-guess' precipitation with the precipitation estimates, we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al.
-(2021) and adaptations from Roberts (2008).
+where \(O\) and \(X\) stand for independent observations and the simulations respectively.
+
+For the comparison of the first-guess' precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al.
+(2021) and adaptations from Roberts (2008):
 
@@
 
-We performed all analysis in this paper using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
-All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2020; Xie, 2015).
-All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019).
+We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019).
+All the analysis in this paper were performed using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
+All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2019; Xie, 2015).
 
@@
 Figure \ref{fig:rcrv-profile}a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
-Both ADPUPA and AIRCFR show a generally good agreement between the ensemble spread and the observation error.
+Both ADPUPA and AIRCFT show a generally good agreement between the ensemble spread and the observation error.
 As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread.
@@
 
-For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) in low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) shows a larger departure from the expected behavior with a negative bias and an overestimation of the observation error.
+For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) in low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) show a larger departure from the expected behavior with a negative bias and an overestimation of the observation error.
 Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero.
@@
 
-The impacts on the wind components are shown in Figure Figure \ref{fig:UV-diff} along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
+The impacts on the wind components are shown in Figure \ref{fig:UV-diff} along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
 The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
@@
 This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
-Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
-
-The assimilation of radiances produced a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c and f).
+Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
+
+The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c and f).
 For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of ASWS observations.
 On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind.
-When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in Figure \ref{fig:UV-diff}f.~
+When we look at the meridional wind field at 200 hPa at different times, we observe that the outflow from the MCS is even more intense than in the other experiments, while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in Figure \ref{fig:UV-diff}f.~
 
@@
 RAD has lower precipitable water content than AUT and SATWND but higher than CONV.
-The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compare to CONV.
+The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
 
@@
 The 0-6 km wind shear is more intense in AUT, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values.
-Moreover, in this same region, these experiments show larger MCAPE values.
-Wind shear over 15 \(ms^{-1}\) have been associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowsky and Richardson, 2010).
-
-\begin{figure*}[t]
+Moreover, in this same region, these experiments show larger MCAPE values than CONV.
+Note that wind shears over 15 \(ms^{-1}\) have been associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowsky and Richardson, 2010).
+
+\begin{figure*}[th]
 
@@
 
-\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a).}\label{fig:TQ-diff}
+\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:TQ-diff}
 \end{figure*}
 
-\begin{figure*}[th]
+\begin{figure*}[ht]
 
@@
 
-\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a).
+\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.
 In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f) are shown.}\label{fig:UV-diff}
@@
 The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period.
-In all the experiments, the accumulated precipitation in the short-range forecasts underestimates the precipitation.
+In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated.
 This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation.
 AUT, SATWND, and RAD better capture the timing and location of convective initiation.
-On RAD the initiation is delayed when we compare with AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
+On RAD the initiation is delayed when we compare with AUT and SATWND, possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
 After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
@@
 \begin{figure*}[h]
-\includegraphics{../figures/pp-hov-1} \caption{Hövmoller of mean hourly accumulated precipitation for each latitude band observed (left) and simulated (right), for the ensemble mean of each experiment, averaged over the inner domain (red box in Figure \ref{fig:dominio}a).
+\includegraphics{../figures/pp-hov-1} \caption{Hövmoller of mean hourly accumulated precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W.
 Contours drawn every 0.5 \(mmh^{-1}\), starting at 0.5 \(mmh^{-1}\).}\label{fig:pp-hov}
@@
 The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
-The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).
+The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).
 
@@
 \includegraphics{../figures/fss-1.png}
-\caption{\label{fig:fss}FSS calculated over a 6-hours moving window for a) 1 mm and b) 25 mm on a 10 km and c) 1 mm and a d) 25 mm on a 100 km scale, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.}
+\caption{\label{fig:fss}FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.}
 \end{figure}
 
-To complement the analysis, Figure \ref{fig:dbz-mean} shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22th along with the observed maximum reflectivity.
+To complement the analysis, Figure \ref{fig:dbz-mean} shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity.
 These experiments were chosen because they represent the analysis with the minimum and the maximum number of assimilated observations.
-In addition, these experiments are the worst and best performing experiments in terms of the short-range precipitation forecast skill.
-Overall none of the short-range forecasts is able to capture the mesoscale details in the precipitation distribution.
-This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately capture the strength of the convective band associated with the MCS.
+In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the short-range precipitation forecast skill (Figure \ref{fig:pp-hov}).
+Overall none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
+This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately represent the strength of the convective band associated with the MCS.
 Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field.
-RAD better represents the observed features of the MSC showing a more strong and organized MCS over the domain center at 10 and 13 UTC (first and second columns in Figure \ref{fig:dbz-mean}).
-The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in the RAD experiment.
-CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
-Before and after the times shown in Figure \ref{fig:dbz-mean}, the agreement between experiment and observations is quite high in the regions where radar data are available.
+RAD better represents the observed features of the system showing a stronger and more organized MCS than CONV, over the domain center at 10 and 13 UTC (first and second columns in Figure \ref{fig:dbz-mean}).
+The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in RAD.
+In addition, CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
+Before and after the times shown in Figure \ref{fig:dbz-mean}, the agreement between experiment and observations is quite high in the regions where radar data are available particularly for RAD.
 
 Finally, Figure \ref{fig:soundings} shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+
 IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments.
@@
 The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND, and RAD.
-On middle levels (below 10 km) the bias is large and positive (8 - 10\(^{\circ}\)C), indicating a more moist atmosphere in the analysis.
-Overall, the RMSE grows with height, this is probably because at upper levels, moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
+On middle levels (below 10 km) the bias is large and positive (8 - 10\(^{\circ}\)C) for all the experiments, indicating a more moist atmosphere in the analyses than the observed.
+Overall, the RMSE grows with height, probably because at upper levels moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
 The zonal wind has a significant positive bias of approximately 10 \(ms^{-1}\) in the upper troposphere which is mainly reduced by assimilating satellite radiances.
@@
 \begin{figure*}[h]
-\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Observed maximum reflectivity in the column (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third rows) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.}\label{fig:dbz-mean}
+\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.}\label{fig:dbz-mean}
 \end{figure*}
 
-For the IOP 8 (bottom row in Figure \ref{fig:soundings}), the densely observed area was behind the MCS, but far enough from it to be directly affected by its mesoscale circulations.
+For the IOP 8 (bottom row in Figure \ref{fig:soundings}), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulations.
 This area was also behind the cold front thus affected by low-level cold advection.
-Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilations of ASWS.
+Below 5 km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
+Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilation of ASWS.
 The RMSE for AUT, SATWND, and RAD remains below 2\(^{\circ}\)C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact.
@@
 The zonal wind is overestimated on the analyses with a positive bias at middle levels.
-For this period only RAD shows an improvement in the upper troposphere.
+For this period only RAD shows an improvement with respect to CONV in the upper troposphere.
 At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND, and RAD.
-In fact, low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
+In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
 
@@
 
-\caption{RMSE (solid line) and Bias (dash line) of a) temperature, b) dew point temperature, c) u wind and d) v wind calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
-The blue line correspond to CONV, the ligh blue line to AUT, SATWND is represented with an orange line and RAD with a red line.}\label{fig:soundings}
+\caption{RMSE (solid line) and Bias (dash line) of a) temperature (\(K\)), b) dew point temperature (\(K\)), c) u wind (\(ms^{-1}\)) and d) v wind (\(ms^{-1}\)) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+The blue line correspond to CONV, the light blue line to AUT, SATWND is represented with an orange line and RAD with a red line.}\label{fig:soundings}
 \end{figure*}
@@
 We evaluated the reliability of the ensemble using the RCRV calculated for each type of observation used in RAD.
-Conventional observations including surface observation and upper air observations from soundings and aircraft show a \(sd RCRV\) near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
-For satellite-derived winds and radiance observations, the \(sd RCRV\) is lower than 1.
-This could be the result of an underdispersive ensemble or an overestimation of the observation errors.
-Taking into account that the errors of this type of observation are usually overestimated to avoid strong impacts on the analysis for poor quality observations and the results obtained for more reliable observations such as soundings, we conclude that the assembly has a reasonable spread.
-
-In terms of the analysis, we found that ASWS observations that have high spatial and temporal resolution produce major impacts throughout the troposphere and mainly on the PBL.
+Conventional observations including surface data and upper air observations from soundings and aircraft, show a \(sd RCRV\) near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
+For satellite-derived winds and radiance observations, the \(sd RCRV\) is lower than 1, which could be the result of an underdispersive ensemble or an overestimation of the observation errors.
+However, taking into account that the errors of this type of observation are usually overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the assembly has a reasonable spread.
+
+In terms of the analysis, we found that ASWS observations, which have high spatial and temporal resolution, produce major impacts throughout the troposphere and mainly on the PBL.
 In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
-We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature state of the MCS leading to an increase in the accumulated precipitation compared with the other experiments.
+We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
 However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates.
@@
 Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during the IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement on low level when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind.
+For the pre-convective environment (IOP 7), we observed an improvement on low levels when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind.
 In particular, ASWS observations correct a dry bias already observed by Ruiz et al.
 (2010).
-For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels for the temperature but the opposite effect for the dew point temperature.
+For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels including more observations for the temperature but the opposite effect for the dew point temperature.
 In this case, the analyses have a moist bias that increases with the assimilation of ASWS and radiances observations.
 We observed a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
-It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period and they require further analysis.
+It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period, which requires further analysis.
 
@@
 
-We are very thankful to the Atmospheric and Sea Research Center (CIMA), the University of Buenos Aires (UBA) and the National Scientific and Technical Research Council (CONICET) who support this study.
+We are very thankful to the Atmospheric and Sea Research Center (CIMA), the University of Buenos Aires (UBA), and the National Scientific and Technical Research Council (CONICET) who support this study.
 We acknowledge the Argentinian National Meteorological Service (ANMS) for kindly providing the radar observations used for validation.
 We also acknowledge the Cheyenne HPC resources (doi:10.5065/D6RX99HX) from NCAR's Computational and Information Systems Laboratory, National Science Foundation (project code UIUC0012).
-Also, PICT 2017-2033 and PICT 2016-0710 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
+Also, PICT 2017-2033, PICT 2016-0710 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-allaire2020}{}%
-Allaire, J.J., Xie {[}aut, Y., cre, McPherson, J., Luraschi, J., Ushey, K., Atkins, A., Wickham, H., Cheng, J., Chang, W., Iannone, R., Dunning, A., filter), A.Y.(.
-sections L., Schloerke, B., Dervieux, C., Aust, F., Allen, J., Seo, J., Barrett, M., Hyndman, R., Lesur, R., Storey, R., Arslan, R., Oller, S., RStudio, PBC, library), j.F.(., inst/rmd/h/jquery-AUTHORS.txt), j.
-contributors (.
-library; authors listed in, inst/rmd/h/jqueryui-AUTHORS.txt), j.U.
-contributors (.U.
-library; authors listed in, library), M.O.(., library), J.T.(., library), B.
-contributors (., Twitter, library), I.(., library), A.F.(., library), S.J.(.
-js, library), I.S.(.
-js, library), G.F.(., templates), J.M.(., Google, library), I.(., library), D.R.(., library), W.(., Gandy (Font-Awesome), D., Sperry (Ionicons), B., (Ionicons), D., StickyTabs), A.L.(., filter), B.P.J.(.L., filter), A.K.(.L., 2020.
-Rmarkdown: Dynamic Documents for R.
+\leavevmode\hypertarget{ref-allaire2019}{}%
+Allaire, J., Horner, J., Xie, Y., Marti, V., Porte, N., 2019. Markdown: Render markdown with the c library 'sundown'.
 
@@
 
+\leavevmode\hypertarget{ref-vera2006}{}%
+Vera, C., Baez, J., Douglas, M., Emmanuel, C.B., Marengo, J., Meitin, J., Nicolini, M., Nogues-Paegle, J., Paegle, J., Penalba, O., Salio, P., Saulo, C., Silva Dias, M.A., Dias, P.S., Zipser, E., 2006.
+The South American Low-Level Jet Experiment.
+Bull.
+Amer.
+Meteor.
+Soc.
+87, 63--78.
+doi:\href{https://doi.org/10.1175/BAMS-87-1-63}{10.1175/BAMS-87-1-63}
+
 \leavevmode\hypertarget{ref-weston2019}{}%

### 2021-09-07 poster2.Rmd
poster 2 - figura analisis

---
title: "Poster 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function(fill = NA) {
  list(geom_sf(data = map, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       # geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)),
       scale_x_longitude(ticks = 5),
       scale_y_latitude(ticks = 5))
}

arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10, 
                                file.dir = "~/mesoda/analysis/data/derived_data/")
arg_topo[, lon := metR::ConvertLongitude(lon)]

cordillera <- list(
  ggnewscale::new_scale_fill(),
  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat,
                                         z = h),
                    breaks = seq(1500, 8000, by = 700)),
  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
                      oob = scales::squish))

experiments <- c("E4", "E5", "E6", "E7")
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311")

satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
```

## Analysis

```{r summary-fields}

files <- Sys.glob(paste0(derived_data, "pw/pw_2*"))

dates <- c("20181122000000", "20181122060000")

pw <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "pw_{date}_{exp}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "pw", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      .[]
  }
}) %>% 
  rbindlist() %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

files <- c(Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122000000.nc")),
           Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122060000.nc")))

mucape <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
      .[]
  }
}) %>% 
  rbindlist()

fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
                     exp = c("E2", "E5", "E6", "E8"))

ana <- lapply(seq_len(nrow(fcsts)), function(f) {

  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>%     .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
    .[, exp := fcsts[f, 2]] %>%  
    .[]
}) %>% 
  rbindlist()

cor <- ana[bottom_top %in% c(1, 16)] %>% 
  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
  .[, cortante := Mag(u_16 - u_1, v_16 - v_1)] %>% 
  .[, ":="(u_1 = NULL, 
           u_16 = NULL,
           v_1 = NULL,
           v_16 = NULL)]

ana[bottom_top %between% c(1, 7), .(v_mean = mean(v)), 
    by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)),
                    proj = norargentina_lambert) +
  colorspace::scale_fill_continuous_divergingx(super = ScaleDiscretised,
                                               palette = "PRGn", mid = 25,
                                               guide = guide_colorsteps(barwidth = 25,
                                                                        barheight = 0.5, 
                                                                        title.position = "left", 
                                                                        title.vjust = 1,
                                                                        frame.colour = "black")) +
  geom_contour2(aes(z = v_mean, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(-10, -5)) +
  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
                        labels = c("-10", "-5")) +
  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

  # ana[bottom_top %between% c(1, 10), .(theta_mean = mean(theta)), 
  #     by = .(south_north, west_east, date, exp)] %>% 
  # .[pw, on = .NATURAL] %>% 
  # ggplot(aes(x, y)) +
  # geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
  #                   proj = norargentina_lambert,
  #                   breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
  # scale_fill_distiller(super = ScaleDiscretised,
  #                      palette = "RdYlBu", direction = -1,
  #                      guide = guide_colorsteps(barwidth = 25,
  #                                               barheight = 0.5, 
  #                                               title.position = "left", 
  #                                               title.vjust = 1,
  #                                               frame.colour = "black")) +
  # labs(fill = "Potential\ntemperature (K)") +
  # cordillera +
  # geom_mapa() +
  # facet_grid(. ~ exp, 
  #            labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
  #                                        E6 = "SATWND", E8 = "RAD"))) +
  # theme_minimal() +
  # theme(legend.position = "bottom",
  #       panel.ontop = TRUE,
  #       panel.grid = element_line(linetype = 3),
  #       strip.text = element_blank(), 
  #       strip.background = element_blank()) +

mucape %>% 
  .[variable == "mcape"] %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
  .[cor, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)), 
                    proj = norargentina_lambert, limits = c(500, 6000)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       palette = "YlOrRd", direction = 1,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.5, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_contour2(aes(z = cortante, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(15, 30)) +
  scale_size_continuous(breaks = c(15, 30), range = c(0.8, 0.4),
                        labels = c("15", "30")) +
  labs(fill = latex2exp::TeX("MCAPE ($JKg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($ms^{-1}$)")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) 
```

### 2021-09-14 analysis/chapters/paper_ana_20181122.Rmd
Changes after LS review

@@
 ---
-title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis"
+title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis and First-guess"
 author:
@@
 A multiphysics approach was also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, ASWS combining CONV and dense automatic surface weather station networks, SATWND, combining ASWS with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
@@
 
-(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AUT c) SATWND and d) RAD experiments calculate over 2.5$^{\circ}$ boxes.
+(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) ASWS c) SATWND and d) RAD experiments calculate over 2.5$^{\circ}$ boxes.
 
@@
   geom_mapa() +
-  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
                                                E6 = "SATWND", E8 = "RAD"))) +
@@
 
-(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
+(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), ASWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
 
@@
   scale_color_manual(values = colores_exp, 
-                     labels = c(E2 = "CONV", E5 = "AUT", 
+                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                 E6 = "SATWND", E8 = "RAD")) +
-  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "AUT", 
+  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
                                                             E6 = "SATWND", E8 = "RAD")) +
@@
   scale_date() +
-  labs(y = "Number of obs per cycle", x = NULL,
+  labs(y = "Number of obs per cycle", x = "Hour (UTC)",
        color = NULL, shape = NULL) +
@@
 The CONV experiment uses only convectional observations from PREPBUFR.
-In a second experiment, referred to as AUT, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from ASWS.
-In the third experiment, referred to as SATWND, we assimilate all the observations of the AUT experiment and the satellite-derived motion winds.
+In a second experiment, referred to as ASWS, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from ASWS.
+In the third experiment, referred to as SATWND, we assimilate all the observations of the ASWS experiment and the satellite-derived motion winds.
 Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbit satellites as described in Section \@ref(sat).
@@
 ) %>% 
-  kbl(booktabs = TRUE, col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
+  kbl(booktabs = TRUE, col.names = c("Obs type", "CONV", "ASWS", "SATWND", "RAD"),
       align = "lcccc",
@@
 
-knitr::include_graphics(here("analysis", "figures", "analysis_cycle.pdf"))
+knitr::include_graphics(here("analysis", "figures", "analysis_cycle.png"))
 
@@
   tag_facets() +
-  labs(x = NULL, y = NULL,
+  labs(x = "Pressure (hPa)", y = NULL,
        linetype = NULL, color = "Upper air\nobservations") +
@@
         legend.spacing.y = unit(0.1, 'mm'),
+        legend.title = element_text(size = 6),
         tagger.panel.tag.text = element_text(size = 8)) +
@@
                         labels = c("mean RCRV", "sd RCRV")) +
-  scale_color_brewer(palette = "Set1") +
+  # scale_color_brewer(palette = "Set1") +
   scale_x_level(breaks = c(1000, 500, 250, 125, 50), limits = c(1000, 50)) +
@@
   tag_facets(tag_pool = c("c")) +
-  labs(x = NULL, y = NULL,
+  labs(x = NULL, y = "mean RCRV / sd RCRV",
        linetype = NULL, color = "Satellite-derived\nwind") +
@@
         legend.spacing.y = unit(0.1, 'mm'),
+        legend.title = element_text(size = 6),
         tagger.panel.tag.text = element_text(size = 8))
@@
 During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
-The warm difference shown in AUT with respect to CONV  in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV.
+The warm difference shown in ASWS with respect to CONV  in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV.
 This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection.
@@
 No clear systematic impact is found above the PBL during this period.
-During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AUT experiment but with the opposite sign.
+During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the ASWS experiment but with the opposite sign.
 
 Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \@ref(fig:TQ-diff)d).
-The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the nighttime.
+The PBL in the ASWS experiment is consistently moister than in the CONV experiment and this is enhanced during the nighttime.
 The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@ruiz2010].
@@
 The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
-AUT and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
-RAD has lower precipitable water content than AUT and SATWND but higher than CONV.
+ASWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
+RAD has lower precipitable water content than ASWS and SATWND but higher than CONV.
 The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
@@
 The potential temperature on the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed on the temperature profiles in Figure \@ref(fig:TQ-diff)a-c.
-On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL than AUT due to the assimilation of radiance observations.
+On average the PBL in ASWS and SATWND is colder than in CONV while RAD shows a warmer PBL than ASWS due to the assimilation of radiance observations.
 Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear.
 The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$.
-MCAPE in the RAD experiment is lower compared to AUT or SATWND.
+MCAPE in the RAD experiment is lower compared to ASWS or SATWND.
 This is consistent with less humidity in the PBL with respect to these experiments and may be partially compensated by a slightly warmer PBL in the RAD experiment.
-The 0-6 km wind shear is more intense in AUT, SATWND, and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
+The 0-6 km wind shear is more intense in ASWS, SATWND, and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
 Moreover, in this same region, these experiments show larger MCAPE values than CONV.
@@
 
-(ref:TQ-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+(ref:TQ-diff) Difference between experiments ASWS-CONV (a and d), SATWND-ASWS (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
 
@@
   facet_wrap(~variable, ncol = 3,
-             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
-                                              "E6_E5" = "SATWND - AUT",
+             labeller = labeller(variable = c("E5_E2" = "ASWS - CONV",
+                                              "E6_E5" = "SATWND - ASWS",
                                               "E8_E6" = "RAD - SATWND"))) +
@@
 
-(ref:UV-diff) Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f) are shown.
+(ref:UV-diff) Difference between experiments ASWS-CONV (a and d), SATWND-ASWS (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+In contours the u wind for ASWS (a), SATWND (b) and RAD (c) and v wind for ASWS (d), SATWND (e) and RAD (f) are shown.
 
@@
   facet_wrap(~variable, ncol = 3,
-             labeller = labeller(variable = c("E5_E2" = "**AUT** - CONV",
-                                              "E6_E5" = "**SATWND** - AUT",
+             labeller = labeller(variable = c("E5_E2" = "**ASWS** - CONV",
+                                              "E6_E5" = "**SATWND** - ASWS",
                                               "E8_E6" = "**RAD** - SATWND"))) +
@@
 All fields correspond to 00 UTC Nov, 22.
+Grey filled contours corresponde to  topography over 1500 meter over sea level.
 
@@
 
-files <- Sys.glob(paste0(derived_data, "pw/*"))[c(1, 4, 5, 7)]
+files <- Sys.glob(paste0(derived_data, "pw/pw_ana_E[2,5,6,8]*"))
 
@@
   facet_grid(. ~ exp, 
-             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+             labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
                                          E6 = "SATWND", E8 = "RAD"))) +
@@
 This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation.
-AUT, SATWND, and RAD better capture the timing and location of convective initiation.
-On RAD the initiation is delayed when we compare with AUT and SATWND, possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
+ASWS, SATWND, and RAD better capture the timing and location of convective initiation.
+On RAD the initiation is delayed when we compare with ASWS and SATWND, possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
 After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
@@
   scale_y_latitude(breaks = seq(-40, -20, 5), labels = NULL) +
-  facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+  facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
                                                           E6 = "SATWND", E8 = "RAD"))) +
@@
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
-AUT and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
+ASWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
 The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
@@
 
-(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
+(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the CONV (blue line), ASWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
 
@@
   # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "ASWS", 
                                                       E6 = "SATWND", E8 = "RAD")) +
@@
 For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that is slightly reduced by the assimilation of ASWS and radiances.
-The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND, and RAD is small above the PBL.
-The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND, and RAD.
+The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the ASWS, SATWND, and RAD is small above the PBL.
+The dew point temperature shows a small dry bias in low levels that is corrected on ASWS, SATWND, and RAD.
 On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C) for all the experiments, indicating a more moist atmosphere in the analyses than the observed.
@@
 Below 5 km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilation of ASWS.
-The RMSE for AUT, SATWND, and RAD remains below 2$^{\circ}$C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact.
+Between 5 and 12 km, the bias in each experiment is cold and is reduced in ASWS, SATWND, and RAD to the assimilation of ASWS.
+The RMSE for ASWS, SATWND, and RAD remains below 2$^{\circ}$C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact.
 In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
@@
 The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
-SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C.
+SATWND presents the lowest RMSE overall while ASWS and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C.
 The zonal wind is overestimated on the analyses with a positive bias at middle levels.
 For this period only RAD shows an improvement with respect to CONV in the upper troposphere.
-At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND, and RAD.
+At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in ASWS, SATWND, and RAD.
 In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
@@
 (ref:soundings) RMSE (solid line) and Bias (dash line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
-The blue line correspond to CONV, the light blue line to AUT, SATWND is represented with an orange line and RAD with a red line.
+The blue line correspond to CONV, the light blue line to ASWS, SATWND is represented with an orange line and RAD with a red line.
 
@@
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = c(colores_exp, "black"), labels = c(E2 = "CONV", E5 = "AUT", 
+  scale_color_manual(values = c(colores_exp, "black"), labels = c(E2 = "CONV", E5 = "ASWS", 
                                                                   E6 = "SATWND", E8 = "RAD")) +
@@
 
-We performed four experiments to assess the impact of different observation types, including conventional data (CONV), frequent surface observations (AUT), satellite-derived winds (SATWND), and clear-sky satellite radiances (RAD), into a regional frequent-update ensemble-based data assimilation system.
+We performed four experiments to assess the impact of different observation types, including conventional data (CONV), frequent surface observations (ASWS), satellite-derived winds (SATWND), and clear-sky satellite radiances (RAD), into a regional frequent-update ensemble-based data assimilation system.
 We used a case study approach of a massive MCS developed over Southern South America on Nov 22, 2018.

### 2021-11-22 R/read_diag_rad.R
big update

@@ -9,31 +9,33 @@
 #' @importFrom magrittr %>%
 #'
 #' @export
-read_diag_rad <- function(file_list, exp) {
+read_diag_rad <- function(file_list, this_exp) {
 
   diag <- purrr::map(seq_along(file_list), function(f){
     if (file.size(file_list[f]) == 0) {
       return(NULL)
     }
+
     meta <- unglue::unglue(basename(file_list[f]), "asim_{sensor}_{plat}_{date}.{mem}")
     # print(f)
     out <- data.table::fread(file_list[f], na.strings = c("0.100000E+12"), fill = TRUE) %>%
-      dplyr::mutate(across(V3:V35, ~as.numeric(.x)))
+      dplyr::mutate(across(V3:V33, ~as.numeric(.x)))
     # .[V10 == 1] %>%
 
     if (file.size(file_list[f]) != 0) {
       out[, ":="(date = lubridate::ymd_hms(meta[[1]][["date"]]),
                  mem = meta[[1]][["mem"]],
-                 exp = exp)] %>%
+                 exp = this_exp)] %>%
         .[, id := 1:.N, by = mem]
     }
     out
   }) %>%
     data.table::rbindlist()
 
-  colnames(diag) <- c("sensor", "channel", "freq", "lat", "lon", "peakwt", "press", "dhr", "tb_obs", "tbc", "tbcnob",
-                      "varch", "errinv", "qc", "emis", "tlapchn", "rzen", "razi", "rlnd", "rice", "rsnw", "rcld",
-                      "rcldp", paste0("pred", seq(12)), "date", "mem", "exp", "id")
+
+  setnames(diag, colnames(diag), c("sensor", "channel", "freq", "lat", "lon", "press", "dhr", "tb_obs", "tbc", "tbcnob",
+                      "errinv", "qc", "emis", "tlapchn", "rzen", "razi", "rlnd", "rice", "rsnw", "rcld",
+                      "rcldp", paste0("pred", seq(12)), "date", "mem", "exp", "id"))
   return(diag)
 }
 

### 2021-11-22 analysis/scripts/acumulados_wrf.R

@@ -137,8 +137,8 @@ saveRDS(pp_wrf_mean, paste0(exp, "_fcst_", format(ini_date, "%Y%m%d%H"), "_mean_
 # Acumulados cada 6 horas -----------------------------------------------------
 
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
-wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E7"
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
+exp <- "E8"
 run <- "fcst"
 
 ini_date <- ymd_hms("20181122060000")

### 2021-11-22 analysis/scripts/altura_estaciones.R

@@ -9,7 +9,7 @@ dominio <- fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>
   .[, c("x", "y") := wrf_project(lon, lat)]
 
 
-oficiales <- fread("/home/paola.corrales/E2_asim_conv_20181121120000.ensmean",
+oficiales <- fread("/home/paola.corrales/datosmunin3/EXP/E2/ANA/20181121120000/diagfiles/asim_conv_20181121120000.ensmean",
                    na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00", "0.100E+12")) %>%
   .[, c("V2", "V4") := NULL] %>%
   setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "alt", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>%
@@ -17,7 +17,7 @@ oficiales <- fread("/home/paola.corrales/E2_asim_conv_20181121120000.ensmean",
   .[!str_detect(stationID, pattern = "[A-Z]")] %>%
   .[, source := "Sfc - Official"]
 
-no_oficiales <- fread("/home/paola.corrales/E5_asim_conv_20181121120000.ensmean",
+no_oficiales <- fread("/home/paola.corrales/datosmunin3/EXP/E5/ANA/20181121120000/diagfiles/asim_conv_20181121120000.ensmean",
                       na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00", "0.100E+12")) %>%
   .[, c("V2", "V4") := NULL] %>%
   setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "alt", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>%

### 2021-11-22 analysis/scripts/count_obs.R

@@ -1,4 +1,4 @@
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E6/ANA/*/diagfiles/asim*ensmean")
+files <- Sys.glob("/home/paola.corrales3/datosmunin/EXP/E6/ANA/*/diagfiles/asim*ensmean")
 
 diag <- read_diag_conv(files, "E6", variable = "uv")
 
@@ -40,7 +40,7 @@ multiespectrales <- c("airs_aqua",
 
 asimilable <- satinfo[iuse == 1, .N, by = sensor]
 
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E8/ANA/*/diagfiles/asim*ensmean")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")
 
 files <- files[str_detect(files, "conv", negate = TRUE)]
 
@@ -60,6 +60,21 @@ diag[, tipo := fifelse(sensor %in% multiespectrales, "multiespectral", "no-multi
 
 diag[tipo == "no-multiespectral", unique(sensor)]
 
+
+a <- diag[, tipo := fifelse(sensor %in% multiespectrales, "multiespectral", "no-multiespectral")] %>%
+  satinfo[., on = c("sensor", "channel")] %>%
+  .[errinv %between% c(0.0000316227, 1000) &
+      qc == 0 &
+      iuse == 1 &
+      peakwt %between% c(50, 1200)] %>%
+  asimilable[., on = c("sensor")] %>%
+  separate(sensor, into = c("sensor", "plataforma"), sep = "_") %>%
+  .[, .N, by = .(sensor, plataforma, channel)]
+
+
+
+a %>%
+  .[sensor == "mhs"]
 
 
 # Porcentaje de obs por experimento ---------------------------------------

### 2021-11-22 analysis/scripts/download_goes_aws.R

@@ -1,0 +1,36 @@
+library(mesoda)
+library(metR)
+library(tidyverse)
+library(lubridate)
+library(data.table)
+
+year <- 2018
+day <- 324
+hour <- 00
+keystart <- paste0("ABI-L1b-RadF/", year, "/", day, "/", hour, "/*")
+
+goes16 <- aws.s3::get_bucket_df(bucket = 'noaa-goes16',
+                                marker = keystart,
+                                max = 7000)
+
+goes16 <- goes16 %>%
+  setDT() %>%
+  .[, start := lubridate::parse_date_time(str_extract(str_extract(Key, "s\\d{14}"), "\\d{13}"), "%Y%j%H%%M%S") ] %>%
+  .[, channel := as.numeric(str_extract(str_extract(Key, "C\\d{2}"), "\\d{2}"))] %>%
+  .[minute(start) == 0 & channel %in% c(7:16) & day(start) < 24] %>%
+  .[]
+
+
+for (i in 1:nrow(goes16)) {
+
+  key <- goes16$Key[i]
+  file <- paste0("/home/paola.corrales/datosmunin3/DATA/GOES16/", key)
+  dir <- dirname(file)
+
+  dir.create(dir, recursive = TRUE)
+
+  print(paste("Downloading file", i, "from", nrow(goes16)))
+
+  aws.s3::save_object(key, bucket = 'noaa-goes16', file = file)
+
+}

### 2021-11-22 analysis/scripts/ejemplo_focus.R

@@ -1,0 +1,73 @@
+# Función para convertir una matriz tidy en matriz, aplicar una función y
+# devolver un vector ordenado
+with_matrix <- function(formula, data = NULL, fun, fill = NULL) {
+  f <- as.character(formula)
+
+  if (length(f) == 1) {  # formula.tool did its thing
+    f <- stringr::str_split(f, "~", n = 2)[[1]]
+  } else {
+    f <- f[-1]
+  }
+
+  value.var <- stringr::str_squish(f[!stringr::str_detect(f, "\\|")])
+
+  matrix.vars <- f[stringr::str_detect(f, "\\|")]
+  matrix.vars <- stringr::str_split(matrix.vars,"\\|", n = 2)[[1]]
+
+  row.vars <- stringr::str_squish(stringr::str_split(matrix.vars[1], "\\+")[[1]])
+  col.vars <- stringr::str_squish(stringr::str_split(matrix.vars[2], "\\+")[[1]])
+
+  if (is.null(data)) {
+    formula <- Formula::as.Formula(formula)
+    data <- data.table::as.data.table(eval(quote(model.frame(formula, data  = data))))
+  } else {
+    # Check if columns are indata
+    all.cols <- c(value.var, row.vars, col.vars)
+    missing.cols <- all.cols[!(all.cols %in% colnames(data))]
+    if (length(missing.cols) != 0) {
+      stopf("Columns not found in data: %s.", paste0(missing.cols, collapse = ", "))
+    }
+    data <- data.table::setDT(data)[, (all.cols), with = FALSE]
+  }
+
+  data.table::setDT(data)
+  dcast.formula <- stringr::str_squish(f[stringr::str_detect(f, "\\|")])
+  dcast.formula <- stats::as.formula(stringr::str_replace(dcast.formula, "\\|", "~"))
+  value.var <- stringr::str_squish(f[!stringr::str_detect(f, "\\|")])
+
+  g <- metR:::.tidy2matrix(data, dcast.formula, value.var, fill = fill)
+  id.name <- "ff19bdd67ff5f59cdce2824074707d20"
+  data[, (id.name) := 1:.N]
+  id <- c(metR:::.tidy2matrix(data, dcast.formula, id.name)$matrix)
+
+  g$matrix <- c(fun(g$matrix))
+
+  c(g$matrix)[order(id)]
+}
+
+
+library(ggplot2)
+library(terra)
[…]

### 2021-11-22 analysis/scripts/interp_obs_parallel.R

@@ -14,12 +14,12 @@ library(foreach)
 library(doParallel)
 
 
-obs_path <- "/home/paola.corrales/datosmunin/EXP/derived_data/interp_obs/"
-fcst_path <- "/home/paola.corrales/datosmunin/EXP"
+obs_path <- "/home/paola.corrales/datosmunin3/EXP/derived_data/interp_obs/"
+fcst_path <- "/home/paola.corrales/datosmunin3/EXP"
 
-ini_date <- ymd_hms("20181122000000")
+ini_date <- ymd_hms("20181122060000")
 
-lead_time <- 36
+lead_time <- 30
 
 # out <- foreach(l = 0:lead_time),
 # .packages = c("data.table", "metR", "lubridate", "interp", "dplyr"),
@@ -46,7 +46,7 @@ purrr::map_dbl(c(0:lead_time), function(l) {
     .[, id := .N, by = .(mem, lon, lat, var)]
 
   #Leo pronostico para ese tiempo
-  files_fcst <- Sys.glob(paste0(fcst_path, "/E[4-7]/FCST/", format(ini_date, "%Y%m%d%H"),
+  files_fcst <- Sys.glob(paste0(fcst_path, "/E[2,5,6,8]/FCST/", format(ini_date, "%Y%m%d%H"),
                                 "/NPP/NPP_",
                                 format(ini_date, "%Y"), "-",
                                 format(ini_date, "%m"), "-",
@@ -57,7 +57,7 @@ purrr::map_dbl(c(0:lead_time), function(l) {
 
   fcst <- purrr::map(files_fcst, function(f) {
 
-    fcst_exp <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{fcst_exp}/FCST/{info}")
+    fcst_exp <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{fcst_exp}/FCST/{info}")
     fcst <- ReadNetCDF(f, vars = c("XLONG", "XLAT", p = "PSFC",
                                    t = "T2",q = "Q2", u = "U10", v = "V10")) %>%
       .[, fcst_exp := fcst_exp[[1]][["fcst_exp"]]] %>%
@@ -102,7 +102,7 @@ purrr::map_dbl(c(0:lead_time), function(l) {
     temp <- fcst_interp[obs, on = c("lon", "lat", "var", "ens" = "mem")]
 
     path_out <- paste0(obs_path, format(ini_date, "%Y%m%d%H"),
[…]

### 2021-11-22 analysis/scripts/satwnd-aut.R

@@ -1,0 +1,85 @@
+files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
+                    pattern = "perfiles_ana")
+
+perfiles <- purrr::map(files, function(f) {
+  exp <- unglue(basename(f), "perfiles_{run}_{exp}.csv")
+  fread(f) %>%
+    .[, exp := exp[[1]][["exp"]]]
+
+}) %>%
+  rbindlist() %>%
+  .[, date := as_datetime(date)] %>%
+  .[date %between% c(as_datetime("2018-11-20 18:00:00"),
+                     as_datetime("2018-11-23 12:00:00"))]
+
+
+perfiles %>%
+  dcast(lev + date ~ exp, value.var = "T") %>%
+  .[, ":="(E5_E2 = E5 - E2,
+           E6_E5 = E6 - E5,
+           E8_E6 = E8 - E6)] %>%
+  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>%
+  .[variable == "E6_E5"] %>%
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = value, fill = stat(level)),
+                    breaks = c(seq(-0.2, 0.2, 0.02), Inf)) +
+  geom_contour2(aes(z = value), color = "white", size = 0.1,
+                breaks = c(seq(-0.2, 0.2, 0.02), Inf)) +
+  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
+                                                            barheight = 0.3,
+                                                            title.position = "left",
+                                                            title.vjust = 1,
+                                                            frame.colour = "black")) +
+  labs(fill = "Temperature (K)",
+       x = NULL,
+       y = NULL) +
+  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_wrap(~variable, ncol = 3,
+             labeller = labeller(variable = c("E5_E2" = "AUT - CONV",
+                                              "E6_E5" = "SATWND - AUT",
+                                              "E8_E6" = "RAD - SATWND"))) +
+  tag_facets(position = list(x = 0.1, y = 0.96)) +
+  theme_minimal(base_size = 8) +
+  theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 8),
+        panel.ontop = TRUE,
[…]

### 2021-11-22 analysis/scripts/sondeos_est.R

@@ -13,7 +13,7 @@ IOP <- tribble(
 ) %>% setDT()
 
 
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E[2,5,6,8]/FCST/*/sondeos/*")
 
 
 estadisticos <- list()
@@ -37,7 +37,8 @@ for (m in seq_len(miembros)) {
     rbindlist() %>%
     .[p >= 50] %>%
     .[, iop := fcase(launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")), "IOP07",
-                     launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08")] %>%
+                     launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")), "IOP08",
+                     default = "No IOP")] %>%
     .[variable %in% c("t", "td", "u", "v") & !str_detect(site, "/") & !is.na(fcst_value) & !is.na(iop)] %>%
     .[site != "Sao Borja, Brazil"] %>%
     .[, lev := cut_round(alt, c(seq(0, 3000, 500), seq(4000, 21000, 1000)))]

### 2021-11-22 analysis/scripts/superficie_est.R

@@ -10,16 +10,16 @@ fisica <- data.table(mem = as.character(formatC(1:60, flag = "0", width = 3)),
                                     "BMJ-MYNN2",
                                     "GF-MYNN2"), length.out = 60)) %>% setDT()
 
-files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112200/interp*"))
+files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112206/interp*new.rds"))
 
 obs <- purrr::map(files, function(f) {
   print(f)
   read_rds(f) %>%
     .[, id := 1:.N, by = ens] %>%
-    .[, fcst_exp := "2018112200"] %>%
+    .[, fcst_exp := "2018112206"] %>%
     .[, mem := formatC(ens, flag = "0", width = 3)] %>%
     .[, exp := NULL] %>%
-    melt(measure.vars = c("E4", "E5", "E6", "E7"),
+    melt(measure.vars = c("E2", "E5", "E6", "E8"),
          variable.name = "exp",
          value.name = "fcst") %>%
     .[, ":="(obs = if_else(usage.flag == 1, obs, NA_real_),
@@ -35,13 +35,13 @@ obs <- purrr::map(files, function(f) {
 }) %>%
   rbindlist()
 
-write_rds(obs, paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds"))
+write_rds(obs, paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem_new.rds"))
 
 
 
 # mean --------------------------------------------------------------------
 
-files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112206/interp*"))
+files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112200/interp*"))
 
 obs <- purrr::map(files, function(f) {
   print(f)
@@ -50,7 +50,7 @@ obs <- purrr::map(files, function(f) {
     .[, fcst_exp := "2018112206"] %>%
     .[, mem := formatC(ens, flag = "0", width = 3)] %>%
     .[, exp := NULL] %>%
-    melt(measure.vars = c("E4", "E5", "E6", "E7"),
+    melt(measure.vars = c("E2", "E5", "E6", "E8"),
          variable.name = "exp",
          value.name = "fcst") %>%
     .[, ":="(obs = if_else(usage.flag == 1, obs, NA_real_),
[…]

### 2021-11-22 analysis/scripts/wrap_fss.R

@@ -9,15 +9,15 @@ library(mesoda)
 # Constante
 #future::plan("cluster", workers = 3)
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
-wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E5"
-run <- "ana_ens"
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
+exp <- "E8"
+run <- "fcst_ens"
 
-ini_date <- ymd_hms("20181120180000")
-ciclos <- 67
+ini_date <- ymd_hms("20181122060000")
+ciclos <- 31
 
 acumulado <- 6
-q <- c(1, 5, 10, 25) #10mm, para arrancar pensando en pp acumulada
+q <- c(1, 5, 10, 25, 50) #10mm, para arrancar pensando en pp acumulada
 w <- c(1, 5, 11, 51, 101) #ancho de cada box en puntos de grilla
 
 # Inicio
@@ -76,8 +76,8 @@ fss_out <- purrr::map_dfr(dates, function(d) {
                         acumulado, "h.rds")
 
     pp_wrf <- readRDS(files_wrf) %>%
-      .[, date := lead_time] %>%
-      .[, exp := exp] %>%
+      # .[, date := lead_time] %>%
+      # .[, exp := exp] %>%
       .[date == d]
 
   } else {

### 2021-11-22 temp_R/goes_functions.R

@@ -4,14 +4,14 @@
 goes_projection <- function(x, y, ncfile) {
 
     proj_info <- ncdf4::ncatt_get(ncfile, "goes_imager_projection")
-  h <- proj_info$perspective_point_height 
-  
+  h <- proj_info$perspective_point_height
+
   map_proj <- paste0("+proj=geos",
                      " +h=",  proj_info$perspective_point_height,
                      " +lon_0=", proj_info$longitude_of_projection_origin,
                      " +sweep=", proj_info$sweep_angle_axis,
                      " +ellps=GRS80")
-  
+
   x_atr <- ncdf4::ncatt_get(ncfile, "x")
   x <- (x*x_atr$scale_factor + x_atr$add_offset)*h
   y_atr <- ncdf4::ncatt_get(ncfile, "y")
@@ -21,7 +21,7 @@ goes_projection <- function(x, y, ncfile) {
 }
 
 # ggplot2::ggplot() +
-#   
+#
 #   scattermore::geom_scattermost(data_sub[!is.na(z_norm), .(lon, lat)], interpolate = TRUE, pixels = c(512, 512),
 #                                 viridisLite::viridis(100)[1 + 99*data_sub[!is.na(z_norm), z_norm]]) +
 #   geom_path(data = map, aes(long, lat, group = group)) +
@@ -30,19 +30,19 @@ goes_projection <- function(x, y, ncfile) {
 
 # scale_color_topes -------------------------------------------------------
 
-scale_color_topes <- function(colours = hex, 
-                              limits = c(-90, 50), 
+scale_color_topes <- function(colours = hex,
+                              limits = c(-90, 50),
                               breaks = seq(-90, 50, 20),
                               guide = guide_colorbar(barheight = 15),
                               ...) {
   # https://github.com/garciafido/cima-goes/tree/master/src/cima/goes/img
-  topes <- data.table::fread("paleta_topes")
+  topes <- data.table::fread("/home/paola.corrales/Analisis_2018112022/paleta_topes")
   colours <- rgb(topes$V2, topes$V3, topes$V4, maxColorValue = 255)
-  
-  scale_color_gradientn(colours = colours, 
-                        limits = limits, 
-                        breaks = breaks, 
-                        guide = guide, ...) 
[…]

### 2021-11-22 analysis/chapters/DA_goes16.Rmd
QC abi

---
title: "Assimilating GOES-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)
source(here::here("temp_R", "goes_functions.R"))

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function() {
  list(geom_sf(data = map, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE),
       coord_sf(xlim = c(-76, -52), ylim = c(-41, -20)))
}

#dominio de interés
xlim <- c(2600, 3950)
ylim <- c(3700, 4750)
```

## Cloud mask

```{r}
nc <- ncdf4::nc_open('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183250000341_e20183250011108_c20183250011312.nc')

cld_msk <- ReadNetCDF('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183250000341_e20183250011108_c20183250011312.nc', 
                      vars = 'BCM', 
                      subset = list(x = xlim,
                                    y = ylim)) %>% 
  .[, c("lon", "lat") := goes_projection(x, y, nc)]

cld_msk %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = BCM == 0)) +
  scale_color_manual(values = c("darkgrey", "white"), guide = NULL) +
  geom_mapa() +
  labs(x = NULL, y = NULL,
       title = "Cloud Mask", subtitle = "Time: 2018-11-21 00 UTC")

```

## Channel 8

```{r}
f <- '~/datosmunin3/DATA/GOES16/ABI-L1b-RadF/2018/325/00/OR_ABI-L1b-RadF-M3C08_G16_s20183250000341_e20183250011108_c20183250011152.nc'

nc <- ncdf4::nc_open(f)

ch_8 <- ReadNetCDF(f, vars = c(rad = "Rad", 
                               qf = "DQF", 
                               sd = "std_dev_radiance_value_of_valid_pixels", 
                               pixel = "valid_pixel_count"), 
                   subset = list(x = xlim, y = ylim)) %>% 
  .[, rad := calculate_rad(rad, nc)] %>% 
  .[, tb := rad_to_tb(rad, nc)] %>% 
  .[, c("lon", "lat") := goes_projection(x, y, nc)]

ch_8 %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = tb)) +
  scale_color_topes() +
  geom_mapa() +
  labs(x = NULL, y = NULL)

ch_8 %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = tb)) +
  scale_color_topes() +
  geom_point(data = cld_msk[BCM == 1], color = "darkgrey", size = 0.9) +
  geom_mapa() +
  labs(x = NULL, y = NULL)

```

## Channel 10

```{r}

files <- Sys.glob("~/datosmunin3/DATA/GOES16/ABI-L1b-RadF/2018/324/*/OR_ABI-L1b-RadF-M3C10_G16*")

ch10 <- map(files, function(f){

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/DATA/GOES16/ABI-L1b-RadF/2018/{day}/{hour}/OR_ABI-L1b-RadF-M3C10_G16_{file}.nc")

  d <- chron::month.day.year(324, c(month = 1, day = 1, year = 2018))

  nc <- ncdf4::nc_open(f)

  ReadNetCDF(f, vars = c(rad = "Rad", 
                         qf = "DQF", 
                         sd = "std_dev_radiance_value_of_valid_pixels"), 
             subset = list(x = xlim, y = ylim)) %>% 
    .[, time := make_datetime(d$year, d$month, d$day, meta[[1]][["hour"]])] %>% 
    .[, rad := calculate_rad(rad, nc)] %>% 
    .[, tb := rad_to_tb(rad, nc)] %>% 
    .[, c("lon", "lat") := goes_projection(x, y, nc)]

}) %>% 
  rbindlist()

```
## Standar deviation

```{r}

ch10 %>%  
  # data.table::copy() %>% 
  .[, tb_std := with_matrix(tb ~ x | y, fun = rolling_2d_sd), 
    by = .(time)]

a %>%   
  .[tb_std >= 0.5] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = tb_std)) +
  scale_color_viridis_c() +
  geom_mapa()

fechas <- unique(ch10$time)

for(d in fechas) {

  ch10 %>%   
  .[tb_std >= 0.5 & time == d] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = tb_std), size = 0.5) +
  scale_color_viridis_c() +
  geom_mapa() +
  labs(title = "SDTB >= 0.5",
       subtitle = paste0("Time: ", as_datetime(d)))

  ggsave(paste0("std_",d,".png"))

}

# xlim <- c(2600, 3950)
# ylim <- c(3700, 4750)
# 
# tb_array <- matrix(a$tb, 
#                    nrow = 4750-3700+1, 
#                    ncol = 3950-2600+1,
#                    byrow = TRUE)
# 
# filled.contour(tb_array)
# 
# sd_array <- raster::focal(as(tb_array, "RasterLayer"),
#               w = matrix(1/9, ncol = 3, nrow = 3),
#               fun = sd,
#               na.rm = TRUE)
# 
# filled.contour(as(sd_array, "matrix"))

```

## Configuration

* 11/21 - 00 UTC
* Only 10 members (small test)
* Only assimilates ch 8 (upper level water vapor)
* Thinning 60 km

```{r}
diag <- fread("~/datosmunin3/EXP/prueba_abi/diag/diag_abi_g16_ges.ensmean.csv") %>% 
  .[, ":="(V5 = ConvertLongitude(V5),
           V2 = V2+6)]

colnames(diag) <- c("sensor", "channel", "freq", "lat", "lon", "peak", "press", "dhr", "tb_obs", "tbc", "tbcnob", "errinv", "qc", "usage", "emis", "tlapchn", "rzen", "razi", "rlnd", "rice", "rsnw", "rcld", "rcldp", paste0("pred", seq(12)))
```

### Observations minus background

```{r}
diag %>% 
  .[channel %in% c(8:10)] %>% 
  ggplot(aes(lon, lat)) +
  # geom_contour_fill(aes(z = tbc, fill = ..level..)) +
  geom_point(aes(color = tbc), size = 1) +
  scale_colour_divergent() +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "O-B [TB]") +
  theme_minimal()

```
### Quality Control flags

* 2 = crtm issues
* 3 = gross check failed
* 5 = surface issues
* 7 = cloud contamination
* 50 = standar deviation to big
* 53 = issues with emissivity from surface

```{r}
diag %>% 
    .[channel %in% c(8:10)] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "QC") +
  theme_minimal()
```

## Assimilable observations

qc == 0, error !=0 and I shoud check the peakwt variable for outliers.

```{r}
diag %>% 
  .[channel %in% c(8:10) &
      errinv != 0 & qc == 0] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = errinv)) +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "Error") +
  theme_minimal()
```

```{r}
diag %>% 
   .[channel %in% c(8:10) &
      errinv != 0 & qc == 0] %>% 
  ggplot(aes(lon, lat)) +
  # geom_contour_fill(aes(z = tbc, fill = ..level..)) +
  geom_point(aes(color = tbc), size = 1) +
  scale_colour_divergent() +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "O-B [TB]") +
  theme_minimal()

```

usage flag = 4 on satinfo for the 3 channels. This means: assimilate without bias correction.

```{r}
diag %>% 
  .[channel %in% c(8:10)] %>% 
  melt(measure.vars = c("tbc", "tbcnob")) %>% 
  # .[variable == "tbcnob"] %>% 
  ggplot(aes(value)) + 
  geom_density(aes(color = variable)) +
  scale_color_discrete(labels = c("BC", "no BC")) +
  facet_wrap(~channel) +
  labs(x = "diff O-B", color = NULL) +
  theme_minimal()
```

## Quality control analysis

```{r}
files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")

diag <- map(files, function(f) {

  # meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000/diagfiles/asim_abi_g16_20181121000000.ensmean")

  read_diag_rad(f, "EG1") %>% 
    # .[, run := meta[[1]][["exp"]]] %>% 
    .[, channel := channel + 6]

}) %>% 
  rbindlist()

```

* **Gross check:** if tbc > errf reject obs. errf = min(3*errf, ermax_rad)

* **Standar Deviation:** toss data from channel 10 and surface channels if standar deviations for channel 10 is >= 0.5. The std dev for each pixel is calculated from the bt 3x3 pixel area. For detection of cloud border? what about areas where the surface have a impact?

* **Surface checks:** toss data depending on the channel and surface type.

```{r}

diag %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(factor(channel))) +
  geom_bar(aes(fill = factor(qc))) +
  scale_fill_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                    labels = c("0" = "pass qc", 
                               "3" = "gross check",
                               "5" = "surface check",
                               "50" = "std dev")) +
  facet_wrap(~date, ncol = 7) +
  labs(fill = NULL, x = "Channel") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

Looks like there are less observations rejected due to the standar deviation threshold on the last cycles. 00 UTC is 21 PM in Argentina, so it may be asociated to a diurnal cycle?

```{r}
diag %>% 
    .[channel %in% c(10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("grey50", "grey70", "grey80", "cyan4"),
                    labels = c("0" = "pass qc", 
                               "3" = "gross check",
                               "5" = "surface check",
                               "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_wrap(~date, ncol = 4) +
  labs(subtitle = "Channel 10", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
Only a few observations are rejected at each cycle. These observations are located over The Andes (surface type = snow)

```{r}
diag[channel %in% c(8:10) & qc == 5] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(rsnw)), size = 0.9) +
  geom_mapa() +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_grid(channel~date) +
  labs(subtitle = "QC = 5 (surface type)", 
       x =  NULL, y = NULL,
       color = "snow?") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
diag[channel %in% c(8:10) & qc == 3] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  .[, gross := min(3*errinv, 2)] %>% 
  ggplot(aes(tbc)) +
  geom_histogram() +
  # geom_point(aes(color = tbc), size = 0.9) +
  # scale_color_divergent() +
  # geom_mapa() +
  facet_grid(channel~date) +
  labs(x = "O-B", subtitle = "QC = 3 (Gross check)") +
  theme_minimal()
```

## Analysis minus guess

```{r}

files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/20181121000000_*/analysis.mem001")

ana <- map(files_ana, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000_{exp}/analysis.mem001")

  ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
                  vars = c(T_ana = "T", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[, run := meta[[1]][["exp"]]]

}) %>% 
  rbindlist()

gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181121000000/wrfarw.mem001", 
                  vars = c(T_guess = "T", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
    .[, a_g := T_ana - T_guess] %>%
  .[level %in% c("low", "middle", "upper") & run %in% c("all", "ch08")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_grid(run~level) +
  theme_minimal()

ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
    .[, a_g := T_ana - T_guess] %>%
  .[bottom_top == 1] %>% 
  dcast(x + y ~ run, value.var = "a_g") %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = all - noabi, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa()

```

```{r}
ana <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
                  vars = c(T_ana = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] 
gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/analysis.mem001",
# gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181120180000/wrfarw.mem001", 
                  vars = c(T_guess = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(2:7)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()

```
```{r}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(10:19)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()
```
```{r}
ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
    .[, a_g := T_ana - T_guess] %>%
  .[level %in% c("low", "middle", "upper")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_wrap(~level) +
  theme_minimal()

ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
    .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(13, 18, 25)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_wrap(~bottom_top) +
  theme_minimal()

```

```{r}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
   .[bottom_top > 10 & lon %between% c(-55, -50)] %>% 
  .[, .(mean_lat = mean(a_g)), by = .(lat, bottom_top)] %>% 
  ggplot(aes(lat, bottom_top)) +
  geom_contour_fill(aes(z = mean_lat, fill = stat(level))) +
  scale_fill_divergent_discretised()
```

### 2021-11-23 analysis/chapters/DA_goes16.Rmd
analisis preliminar ABI

@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
+
+source("~/mesoda/analysis/scripts/ejemplo_focus.R")
 
@@
 
+```{r}
+diag %>% 
+  .[channel %in% c(8:10)] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = factor(qc)), size = 0.9) +
+  geom_mapa() +
+  scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
+                     labels = c("0" = "pass qc", 
+                                "3" = "gross check",
+                                "5" = "surface check",
+                                "50" = "std dev")) +
+  guides(color = guide_legend(override.aes = list(size = 2))) +
+  facet_grid(channel~date) +
+  labs(subtitle = "Assimilated obs", x =  NULL, y = NULL, color = "QC") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
 Looks like there are less observations rejected due to the standar deviation threshold on the last cycles. 00 UTC is 21 PM in Argentina, so it may be asociated to a diurnal cycle?
@@
 ```{r}
+files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/analysis.mem001")
 
-files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/20181121000000_*/analysis.mem001")
+coord <- ReadNetCDF(files_ana[1], vars = c(lon = "XLONG", lat = "XLAT"), subset = list(bottom_top = 1))
+
+levs <- c(200, 300, 400, 500, 600, 700, 800, 900, 1000)
 
@@
 
-  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000_{exp}/analysis.mem001")
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/{date}/analysis.mem001")
 
-  ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
-                  vars = c(T_ana = "T", lon = "XLONG", lat = "XLAT")) %>% 
+  ReadNetCDF(f, vars = c(t_ana = "T", p = "P", "PB", q_ana = "QVAPOR",
+                         lon = "XLONG", lat = "XLAT")) %>% 
+    .[, ":="(t_ana = tk(t_ana, p + PB, T_BASE = 290) - 273.15,
+             p = p + PB,
+             PB = NULL)] %>% 
+    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
+    .[, .(lev = levs,
+          t_ana = approx(p, t_ana, xout = levs*100)$y,
+          q_ana = approx(p, q_ana, xout = levs*100)$y),
+      by = .(date, south_north, west_east)] %>%
+    coord[., on = .NATURAL] %>% 
     .[, c("x", "y") := wrf_project(lon, lat)] %>% 
-    .[, run := meta[[1]][["exp"]]]
+    .[]
 
@@
 
-gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181121000000/wrfarw.mem001", 
-                  vars = c(T_guess = "T", lon = "XLONG", lat = "XLAT")) %>% 
-  .[, c("x", "y") := wrf_project(lon, lat)]
+files_gue <- Sys.glob("~/datosmunin3/EXP/prueba_abi/GUESS/*00/wrfarw.mem001")
+
+gue <- map(files_gue, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/GUESS/{date}/wrfarw.mem001")
+
+  ReadNetCDF(f, vars = c(t_gue = "T", p = "P", "PB", q_gue = "QVAPOR",
+                         lon = "XLONG", lat = "XLAT")) %>% 
+    .[, ":="(t_gue = tk(t_gue, p + PB, T_BASE = 290) - 273.15,
+             p = p + PB,
+             PB = NULL)] %>% 
+    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
+    .[, .(lev = levs,
+          t_gue = approx(p, t_gue, xout = levs*100)$y,
+          q_gue = approx(p, q_gue, xout = levs*100)$y),
+      by = .(date, south_north, west_east)] %>%
+    coord[., on = .NATURAL] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+    .[]
+
+}) %>% 
+  rbindlist()
+
+```
+
+```{r}
+dates <- unique(ana$date)
+
+for (i in 1:length(dates)) {
 
   ana[gue, on = .NATURAL] %>% 
-  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
-                         bottom_top %between% c(11, 15), "low",
-                         bottom_top %between% c(16, 20), "middle",
-                         bottom_top %between% c(21, 26), "upper",
-                         bottom_top %between% c(27, 36), "the rest"))] %>% 
-    .[, a_g := T_ana - T_guess] %>%
-  .[level %in% c("low", "middle", "upper") & run %in% c("all", "ch08")] %>% 
+  .[date == dates[i]] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  .[, a_g := t_ana - t_gue] %>%
+  .[lev %in% c(300, 500, 700)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(trans = mesoda::symlog_trans()) +
+  geom_mapa() +
+  labs(x = NULL, y = NULL, fill = "A-B",
+       title = "Temperature",
+       subtitle = as.character(dates[i])) +
+  facet_grid(date~lev) +
+  theme_minimal()
+
+  ggsave(paste0("t_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)
+
+  ana[gue, on = .NATURAL] %>% 
+  .[date == dates[i]] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  .[, a_g := q_ana - q_gue] %>%
+  .[lev %in% c(300, 500, 700)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(trans = mesoda::symlog_trans()) +
+  geom_mapa() +
+  labs(x = NULL, y = NULL, fill = "A-B",
+       title = "Humidity",
+       subtitle = as.character(dates[i])) +
+  facet_grid(date~lev) +
+  theme_minimal()
+
+  ggsave(paste0("q_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)
+
+}
+```
+
+```{r}
+
+for (i in 1:length(levs)) {
+
+  ana[gue, on = .NATURAL] %>% 
+  .[lev == levs[i]] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  .[, a_g := t_ana - t_gue] %>%
   ggplot(aes(x, y)) +
@@
   geom_mapa() +
-  labs(x = NULL, y = NULL, fill = "A-B") +
-  facet_grid(run~level) +
+  labs(x = NULL, y = NULL, fill = "A-B",
+       title = "Temperature",
+       subtitle = paste(as.character(levs[i]), "hPa")) +
+  facet_wrap(~date, ncol = 4) +
   theme_minimal()
 
+  ggsave(paste0("t_ab_", levs[i], ".png"), height = 4)
+
   ana[gue, on = .NATURAL] %>% 
-  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
-                         bottom_top %between% c(11, 15), "low",
-                         bottom_top %between% c(16, 20), "middle",
-                         bottom_top %between% c(21, 26), "upper",
-                         bottom_top %between% c(27, 36), "the rest"))] %>% 
-    .[, a_g := T_ana - T_guess] %>%
-  .[bottom_top == 1] %>% 
-  dcast(x + y ~ run, value.var = "a_g") %>% 
+  .[lev == levs[i]] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  .[, a_g := q_ana - q_gue] %>%
   ggplot(aes(x, y)) +
-  geom_contour_fill(aes(z = all - noabi, fill = ..level..), proj = norargentina_lambert) +
+  geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
   scale_fill_divergent_discretised() +
-  geom_mapa()
+  geom_mapa() +
+  labs(x = NULL, y = NULL, fill = "A-B",
+       title = "Humidity",
+       subtitle = paste(as.character(levs[i]), "hPa")) +
+   facet_wrap(~date, ncol = 4) +
+  theme_minimal()
+
+  ggsave(paste0("q_ab_", levs[i], ".png"), height = 4)
+
+}
+
+```
+```{r}
+files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_conv*.ensmean")
+
+read_diag_conv(files, "EG1") %>% 
+  .[type %in% c(120)] %>% 
+  unique(., by = c("stationID", "lat", "lon", "date")) %>% 
+  .[] %>% 
+  ggplot(aes(ConvertLongitude(lon), lat)) + 
+  geom_point() +
+  geom_mapa() +
+  labs(x = NULL, y = NULL, 
+       title = "Soundings",
+       subtitle = "2018-11-21 00 UTC")
 
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 ana <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
@@
   .[, c("x", "y") := wrf_project(lon, lat)]
-gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/analysis.mem001",
-# gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181120180000/wrfarw.mem001", 
+
+gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181120180000/wrfarw.mem001",
                   vars = c(T_guess = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
@@
   geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
-  scale_fill_divergent_discretised() +
+  # scale_fill_divergent_discretised() +
   geom_mapa() +
@@
 ```
-```{r}
+
+```{r eval=FALSE, include=FALSE}
 ana[gue, on = .NATURAL] %>% 
@@
 ```
-```{r}
+```{r eval=FALSE, include=FALSE}
 ana[gue, on = .NATURAL] %>% 
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 ana[gue, on = .NATURAL] %>% 

### 2021-11-23 R/read_diag_rad.R

@@ -18,8 +18,9 @@ read_diag_rad <- function(file_list, this_exp) {
 
     meta <- unglue::unglue(basename(file_list[f]), "asim_{sensor}_{plat}_{date}.{mem}")
     # print(f)
-    out <- data.table::fread(file_list[f], na.strings = c("0.100000E+12"), fill = TRUE) %>%
-      dplyr::mutate(across(V3:V33, ~as.numeric(.x)))
+    out <- data.table::fread(file_list[f], na.strings = c("0.100000E+12"), fill = TRUE)
+    # %>%
+    #   dplyr::mutate(across(V3:V33, ~as.numeric(.x)))
     # .[V10 == 1] %>%
 
     if (file.size(file_list[f]) != 0) {

### 2021-11-23 analysis/scripts/ASWS.R

@@ -1,0 +1,42 @@
+librayr(tidyverse)
+library(ggplot2)
+
+df <- map_df( Sys.glob("~/datosmunin3/DATA/OBS/datain/ADPSFC_n3/201811*"), ~fread(.x, na.strings = c("1e+11",
+                                                                                                     "1.000000e+11",
+                                                                                                     "1.0000e+11"))) %>%
+  setnames(paste0("V", c(1:10)), c("lat", "lon", "alt", "stationID", "date", "t", "q", "u", "v", "p"))
+
+
+n_stations <-
+  df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54), .N, by = stationID] %>%
+  .[str_detect(stationID, "SMN", negate = TRUE)]
+
+df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
+  .[str_detect(stationID, "SMN", negate = TRUE)] %>%
+  melt(measure.vars = c("t", "q", "u", "v", "p")) %>%
+  .[, is_na := fifelse(is.na(value), "NA", "OBS")] %>%
+  .[, .N, by = .(stationID, variable, is_na)] %>%
+  dcast(stationID + variable ~ is_na) %>%
+  .[!is.na(OBS)] %>%
+  dcast(stationID ~ variable, value.var = "OBS")
+
+df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
+  .[str_detect(stationID, "SMN", negate = TRUE)] %>%
+  melt(measure.vars = c("t", "q", "u", "v", "p")) %>%
+  .[, is_na := fifelse(is.na(value), "NA", "OBS")] %>%
+  .[, .N, by = .(stationID, variable, is_na)] %>%
+  dcast(stationID + variable ~ is_na) %>%
+  .[!is.na(OBS), .N, by = variable]
+
+SMN <- %>%
+  df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
+  # unique(by = c("stationID")) %>%
+  .[str_detect(stationID, "SMN")] %>%
+  unique(by = "date")
+
+df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
+  unique(by = c("lat", "lon")) %>%
+  ggplot(aes(lon, lat)) +
+  geom_point() +
+  geom_point(data = square, size = 0.1) +
+  geom_mapa()

### 2021-11-23 analysis/scripts/ejemplo_focus.R

@@ -45,29 +45,29 @@ with_matrix <- function(formula, data = NULL, fun, fill = NULL) {
   c(g$matrix)[order(id)]
 }
 
-
+#
 library(ggplot2)
-library(terra)
-#> terra version 1.4.11
+# library(terra)
+# #> terra version 1.4.11
 library(raster)
-#> Loading required package: sp
-# Hace desvio estándard en una ventana movil de 3x3
+# #> Loading required package: sp
+# # Hace desvio estándard en una ventana movil de 3x3
 rolling_2d_sd <- function(m) {
   as(m, "RasterLayer") %>%
     raster::focal(w = matrix(1/9, ncol = 3, nrow = 3), fun = sd) %>%
     as("matrix")
 
 }
-
-library(metR)
-
-
-
-
-
-temperature[lev == 200] |>
-  data.table::copy() |>
-  D(, air_std := with_matrix(air ~ lon | lat, fun = rolling_2d_sd)) |>
-  ggplot(aes(lon, lat)) +
-  geom_contour_fill(aes(z = air_std)) +
-  geom_contour2(aes(z = air))
+#
+# library(metR)
+#
+#
+#
+#
+#
+# temperature[lev == 200] |>
+#   data.table::copy() |>
+#   D(, air_std := with_matrix(air ~ lon | lat, fun = rolling_2d_sd)) |>
+#   ggplot(aes(lon, lat)) +
+#   geom_contour_fill(aes(z = air_std)) +
+#   geom_contour2(aes(z = air))

### 2022-03-10 R/areal_coverage.R
new functions

@@ -1,0 +1,138 @@
+#' Calculate areal coverage
+#'
+#' @param exp name of the experiment
+#' @param run if "ana" or "fcst
+#' @param ini_date init date and time
+#' @param ciclos number of hourly cycles/fcst
+#' @param acumulado time of accumulated
+#' @param q pp threshold
+#' @param lon_band longitud limits
+#' @param lat_band latitude limits
+#' @param wrf_path where the simulations are located
+#' @param out_path where the output file will be saved
+#'
+#' @export
+
+
+calculate_areal_coverage <- function(exp, run, ini_date, ciclos, acumulado = 1,
+                                     q = c(1, 5, 10, 25, 50),
+                                     lon_band = c(-66.5, -61.5),
+                                     lat_band = c(-35.5, -29),
+                                     wrf_path = "/home/paola.corrales/datosmunin3/EXP/",
+                                     out_path = "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/") {
+
+  ini_date <- lubridate::ymd_hms(ini_date)
+
+
+  if (run == "ana") {
+
+    dates <- seq.POSIXt(ini_date + lubridate::hours(acumulado), by = "hour",
+                        length.out = ciclos - acumulado)
+
+    pp_wrf_out <- purrr::map_dbl(dates, function(d) {
+
+      date <- d
+
+      print(d)
+
+      files_wrf <-  purrr::map(seq(0, acumulado - 1), function(l) {
+        list.files(path = paste0(wrf_path, exp, "/", toupper(run), "/", format(date - lubridate::hours(l), "%Y%m%d%H%M%S")),
+                   full.names = TRUE,
+                   recursive = TRUE,
+                   pattern = "analysis.mem*")
+      }) %>% unlist()
+
+
+      pp_wrf <- purrr::map(files_wrf, function(f) {
+
+        metadatos <- unglue::unglue(basename(f), "analysis.mem0{mem}")
+
+        metR::ReadNetCDF(f, vars = c("RAINNC", "RAINC", "RAINSH",
+                               lon = "XLONG", lat = "XLAT")) %>%
+          .[, ":="(pp_acum = RAINNC + RAINC + RAINSH,
+                   exp = exp,
+                   date = date,
+                   mem = metadatos[[1]][["mem"]])] %>%
[…]

### 2022-03-10 R/probability_matching.R

@@ -1,0 +1,141 @@
+#' Calculate probability matching
+#'
+#' @param exp name of the experiment
+#' @param run if "ana" or "fcst
+#' @param ini_date init date and time
+#' @param ciclos number of hourly cycles/fcst
+#' @param acumulado time of accumulated
+#' @param wrf_path where the simulations are located
+#' @param out_path where the output file will be saved
+#'
+#' @export
+
+
+calculate_probability_marching <- function(exp, run, ini_date, ciclos, acumulado = 1,
+                                           wrf_path = "/home/paola.corrales/datosmunin3/EXP/",
+                                           out_path = "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/") {
+
+  ini_date <- lubridate::ymd_hms(ini_date)
+
+  members <- 60
+  dates <- seq.POSIXt(ini_date, by = "hour",
+                      length.out = ciclos)
+
+
+  if (run == "ana") {
+    pp_wrf_out <- purrr::map_dbl(dates, function(d) {
+
+      date <- d
+
+      print(d)
+
+      files_wrf <- list.files(path = paste0(wrf_path, exp, "/", toupper(run), "/", format(date, "%Y%m%d%H%M%S")),
+                              full.names = TRUE,
+                              recursive = TRUE,
+                              pattern = "analysis.mem*")
+
+      pp_wrf <- purrr::map(files_wrf, function(f) {
+
+        metadatos <- unglue::unglue(basename(f), "analysis.mem0{mem}")
+
+        metR::ReadNetCDF(f, vars = c("RAINNC", "RAINC", "RAINSH",
+                                     lon = "XLONG", lat = "XLAT")) %>%
+          .[, ":="(pp_acum = RAINNC + RAINC + RAINSH,
+                   mem = metadatos[[1]][["mem"]])] %>%
+          .[, ":="(RAINNC = NULL,
+                   RAINC = NULL,
+                   RAINSH = NULL,
+                   Time = NULL)]
+      }) %>%
+        data.table::rbindlist()
+
+
+      pp_wrf_mean <- pp_wrf[, .(ens_mean = mean(pp_acum)), by = .(lon, lat)]
+
+      pp_wrf_mean[, rank := data.table::frank(-ens_mean, ties.method = "first")]
+
[…]

### 2022-03-10 R/td.R

@@ -3,7 +3,7 @@
 #' @param QVAPOR vector con la humedad especifica en kg/kg
 #' @param P vector con la presión en Pa
 #'
-#'  @export
+#' @export
 td <- function(QVAPOR, P) {
 
   # qv en kg/kg

### 2022-03-10 analysis/chapters/paper_ana_20181122.Rmd
Fix figures and captions

@@
                    labels = function(x) {
-                     fifelse(hour(x) == 0, format(x, "%H\n%b %d"), format(x, "%H"))
+                     fifelse(hour(x) == 0, format(x, "%H UTC\n%b %d"), format(x, "%H"))
                    }) 
@@
 #Color blind friendly but E4 y E7 parecidos en Desaturated
-colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311") 
+colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E9 = "#CC3311") 
 # colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
@@
 A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
-Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. 
+Observations are assimilated in a 4D approach by comparing them with the corresponding first-guess state at 10-minute intervals. 
 For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures.
@@
 
-(ref:dominio) a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
-b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions.
+(ref:dominio) a) The horizontal domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+b) Locations of radiosonde launches during the experiment.
+Green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions.
 The topography in meters is also shown (shaded).
@@
   footnote(symbol = c("Observation error varied with height.",
-                      "Observations over 600 hPa are rejected.",
+                      "Observations above 600 hPa are rejected.",
                       "Observation error depends on the report type."),
@@
       col.names = linebreak(c("Sensor", "Platform", "Assimilated\nchannels", "Percentage\nover total")),
-      caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>%
+      caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation, and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>%
   kable_styling(font_size = 7) %>% 
@@
 
-(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) ASWS c) SATWND and d) RAD experiments calculate over 2.5$^{\circ}$ boxes.
+(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) ASWS c) SATWND, and d) RAD experiments calculated over 2.5$^{\circ}$ boxes.
 
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E9/ANA/*/diagfiles/asim*ensmean")
 
@@
 
-rad <- read_diag_rad(files, "E8") %>% 
+rad <- read_diag_rad(files, "E9") %>% 
   satinfo[., on = c("sensor", "channel")] %>% 
@@
 
-rad_conv <- read_diag_conv(files, exp = "E8", member = "000")
+rad_conv <- read_diag_conv(files, exp = "E9", member = "000")
 
@@
   facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
-                                               E6 = "SATWND", E8 = "RAD"))) +
+                                               E6 = "SATWND", E9 = "RAD"))) +
   tagger::tag_facets() +
@@
                      labels = c(E2 = "CONV", E5 = "ASWS", 
-                                E6 = "SATWND", E8 = "RAD")) +
+                                E6 = "SATWND", E9 = "RAD")) +
   scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
-                                                            E6 = "SATWND", E8 = "RAD")) +
+                                                            E6 = "SATWND", E9 = "RAD")) +
   scale_y_log10() +
@@
                      labels = c(E2 = "CONV", E5 = "ASWS", 
-                                E6 = "SATWND", E8 = "RAD")) +
+                                E6 = "SATWND", E9 = "RAD")) +
   scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
-                                                            E6 = "SATWND", E8 = "RAD")) +
+                                                            E6 = "SATWND", E9 = "RAD")) +
   # geom_tile(aes(fill = obs_cycle)) +
@@
   # scale_color_discrete() +
-  scale_x_level(breaks = c(1000, 500, 250, 125, 50)) +
+  scale_x_level(name =  "Pressure (hPa)", breaks = c(1000, 500, 250, 125, 50)) +
   coord_flip() +
@@
   theme(legend.box = "vertical",
-        legend.spacing.y = unit(0.1, 'mm'),
+        # legend.spacing.y = unit(0.1, 'mm'),
         legend.title = element_text(size = 6),
@@
   rbindlist() %>% 
+  .[sensor %in% c("airs", "amsua", "iasi", "mhs")] %>% 
   separate(sensor, into = c("sensor", "plat"), sep = "_") %>% 
@@
 
-(ref:TQ-diff) Difference between experiments ASWS-CONV (a and d), SATWND-ASWS (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and b) ASWS-CONV, b) and e) SATWND-ASWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
 
@@
            E6_E5 = E6 - E5,
-           E8_E6 = E8 - E6)] %>% 
-  melt(measure.vars = c("E5_E2", "E6_E5", "E8_E6")) %>% 
+           E9_E6 = E9 - E6)] %>% 
+  melt(measure.vars = c("E5_E2", "E6_E5", "E9_E6")) %>% 
   ggplot(aes(date, lev)) +
@@
        x = NULL,
-       y = NULL) +
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+       y = "Pressure (hPa)") +
+  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
@@
                                               "E6_E5" = "SATWND - ASWS",
-                                              "E8_E6" = "RAD - SATWND"))) +
+                                              "E9_E6" = "RAD - SATWND"))) +
   tag_facets(position = list(x = 0.1, y = 0.96)) +
@@
                 breaks = c(seq(-1.6, 1.6, 0.2), Inf)) +
-  scale_fill_divergent_discretised(limits = c(-1.6, 1.6),
-                                   guide = guide_colorsteps(barwidth = 25,
+  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                             barheight = 0.3, 
@@
 
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
   labs(fill = latex2exp::TeX("Specific \nhumidity ($gKg^{-1}$)"),
-       y = NULL,
+       y = "Pressure (hPa)",
        x = NULL) +
@@
                                               "E6_E5" = "SATWND - ASWS",
-                                              "E8_E6" = "RAD - SATWND"))) +
+                                              "E9_E6" = "RAD - SATWND"))) +
   tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
@@
 
-(ref:UV-diff) Difference between experiments ASWS-CONV (a and d), SATWND-ASWS (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-In contours the u wind for ASWS (a), SATWND (b) and RAD (c) and v wind for ASWS (d), SATWND (e) and RAD (f) are shown.
+(ref:UV-diff)  Difference between analysis ensemble mean experiments a) and b) ASWS-CONV, b) and e) SATWND-ASWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+Black contours corresponde to u wind for ASWS (a), SATWND (b) and RAD (c) and v wind for ASWS (d), SATWND (e) and RAD (f).
 
@@
                 breaks = c(seq(-2, 2, 0.2), Inf)) +
-  geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>%
+  geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E9_E6 = E9)] %>%
                   melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
                 breaks = seq(0, 30, 3), size = 0.2, linetype = 1, color = "grey30") +
-  geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>%
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E9_E6 = E9)] %>%
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value),
@@
        y = NULL) +
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
@@
                                               "E6_E5" = "**SATWND** - ASWS",
-                                              "E8_E6" = "**RAD** - SATWND"))) +
+                                              "E9_E6" = "**RAD** - SATWND"))) +
   tag_facets(position = list(x = 0.1, y = 0.96)) +
@@
                 breaks = c(-Inf, seq(-1.6, 1.6, 0.2), Inf)) +
-  scale_fill_divergent_discretised(limits = c(-1.6, 1.6),
-                                   guide = guide_colorsteps(barwidth = 25,
+  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
                                                             barheight = 0.3, 
@@
                                                             frame.colour = "black")) +
-  geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
+  geom_contour2(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E9_E6 = E9)] %>% 
                   melt(id.vars = c("lev", "date")) %>% unique(), 
@@
                 breaks = seq(-10, 4, 2), size = 0.2, color = "grey30") +
-  geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E8_E6 = E8)] %>% 
+  geom_text_contour(data = ~.x[, .(lev, date, E5_E2 = E5, E6_E5 = E6, E9_E6 = E9)] %>% 
                       melt(id.vars = c("lev", "date")) %>% unique(), aes(z = value), 
@@
   scale_linetype(guide = NULL) +
-  scale_y_level(breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
@@
                                               "E6_E5" = "**SATWND** - ASWS",
-                                              "E8_E6" = "**RAD** - SATWND"))) +
+                                              "E9_E6" = "**RAD** - SATWND"))) +
   tag_facets(tag_pool = c("d", "e", "f"), position = list(x = 0.1, y = 0.96)) +
@@
 (ref:summary-fields) a) Precipitable water (shaded, $kgm^{-2}$) and average northerly wind on the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, $ms^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $ms^{-1}$ for each experiment.
-All fields correspond to 00 UTC Nov, 22.
+All fields correspond to the analysis ensemble mean for 00 UTC Nov, 22.
 Grey filled contours corresponde to  topography over 1500 meter over sea level.
@@
 
-files <- Sys.glob(paste0(derived_data, "pw/pw_ana_E[2,5,6,8]*"))
+files <- Sys.glob(paste0(derived_data, "pw/pw_ana_E[2,5,6,9]*"))
 
@@
 
-files <- Sys.glob(paste0(derived_data, "mucape/mcape_ana_E*_20181122000000.nc"))[c(1, 4, 5, 7)]
+files <- Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,9]_20181122000000.nc"))
 
@@
 fcsts <- expand_grid(fcsts = c("20181122000000"),
-                     exp = c("E2", "E5", "E6", "E8"))
+                     exp = c("E2", "E5", "E6", "E9"))
 
@@
              labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
-                                         E6 = "SATWND", E8 = "RAD"))) +
+                                         E6 = "SATWND", E9 = "RAD"))) +
   theme_minimal() +
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2568]_ana*mean_acum_1h.rds")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2569]_ana*_pp_pmm.rds")
 
@@
   readRDS(f) %>% 
-    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)]
+    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+    .[, ":="(rank = NULL, ens_mean = NULL)] %>% 
+  setnames(c("pp_pmm"), c("pp_acum"))
 
@@
   .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
-  .[end_date %between% c(as_datetime("20181120190000"), as_datetime("20181123120000"))] %>% 
+  .[end_date %between% c(as_datetime("20181121180000"), as_datetime("20181123120000"))] %>% 
   .[, .(pp_acum = mean(pp_acum),
@@
   facet_wrap(~ exp) +
-  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
+  tag_facets(position = list(x = 0.1, y = 0.96)) +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Accumulated\nprecipitation ($mmh^{-1}$)")) +
   theme_minimal(base_size = 9) +
@@
   ppacum_1h %>% 
-  .[exp %in% c("E2", "E5", "E6", "E8")] %>%
+  .[exp %in% c("E2", "E5", "E6", "E9")] %>%
   .[, lat := lat[1], by = .(y)] %>% 
   .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y, lat)] %>% 
+   .[date %between% c(as_datetime("20181121180000"), as_datetime("20181123120000"))] %>% 
   ggplot(aes(date, lat)) +
@@
   facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
-                                                          E6 = "SATWND", E8 = "RAD"))) +
-  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
+                                                          E6 = "SATWND", E9 = "RAD"))) +
+  tag_facets(tag_pool = c("b", "c", "d", "e"), position = list(x = 0.1, y = 0.96)) +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Accumulated\nprecipitation ($mmh^{-1}$)")) +
   theme_minimal(base_size = 9) +
@@
 
-(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the CONV (blue line), ASWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
+(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), ASWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
 
 ```{r fss, fig.cap="(ref:fss)", fig.width=3.5, fig.height=3.5, fig.pos="t"}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2-8].csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2569].csv")
 
@@
 fss %>% 
-  .[w %in% c(1, 11) & q %in% c(1, 25) & exp %in% c("E2", "E5", "E6", "E8")] %>% 
+  .[w %in% c(1, 11) & q %in% c(1, 25) & exp %in% c("E2", "E5", "E6", "E9")] %>% 
   # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
@@
   scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "ASWS", 
-                                                      E6 = "SATWND", E8 = "RAD")) +
+                                                      E6 = "SATWND", E9 = "RAD")) +
   scale_date(20181122000000, 20181123000000, "12 hours") +
@@
 
-(ref:dbz-mean) Maximum reflectivity in the column (COLMAX in $dBZ$), observed (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.
+(ref:dbz-mean) Maximum reflectivity in the column (COLMAX in $dBZ$), observed (upper row) and 1-hr forecast ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22.
 
@@
 
-rad_loc <- read_csv(here("analysis", "data", "derived_data", "radar_loc.csv"))
+rad_loc <- fread(here("analysis", "data", "derived_data", "radar_loc.csv")) %>% 
+  .[, run := "OBSERVATIONS"]
 
@@
            "20181122160000", "20181122190000")
-files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[28]_", dates, ".nc"))
+files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[29]_", dates, ".nc"))
 
@@
   rbindlist()  %>% 
-  .[, run := fifelse(exp == "E8", "RAD", "CONV")] %>% 
+  .[, run := fifelse(exp == "E9", "RAD", "CONV")] %>% 
   .[, ":="(south_north = NULL,
@@
   # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
-  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
+  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.2, color = "grey20") +
   geom_mapa() +
   coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)) +
-  scale_x_longitude(ticks = 5, labels = NULL) +
-  facet_grid(run ~ date, labeller = labeller(date = c("2018-11-22 10:00:00" = "10 UTC",
+  scale_x_longitude(ticks = 5) +
+  facet_grid(factor(run, levels = c("OBSERVATIONS", "CONV", "RAD")) ~ date, 
+             labeller = labeller(date = c("2018-11-22 10:00:00" = "10 UTC",
                                                       "2018-11-22 13:00:00" = "13 UTC",
@@
   theme(legend.position = "bottom",
-        axis.text.x = element_blank(),
+        # axis.text.x = element_blank(),
         plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
@@
 
-(ref:soundings) RMSE (solid line) and Bias (dash line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
-The blue line correspond to CONV, the light blue line to ASWS, SATWND is represented with an orange line and RAD with a red line.
+(ref:soundings) RMSE (solid line) and Bias (dashed line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+The blue line corresponds to CONV, the light blue line to ASWS, SATWND is represented with an orange line, and RAD with a red line.
 
 ```{r soundings, fig.cap="(ref:soundings)", fig.env="figure*", fig.pos="t", fig.align="center"}
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos/ANA/sondeo_E[2568]_ana*")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos/ANA/sondeo_E[2569]_ana*")
 
@@
   scale_color_manual(values = c(colores_exp, "black"), labels = c(E2 = "CONV", E5 = "ASWS", 
-                                                                  E6 = "SATWND", E8 = "RAD")) +
+                                                                  E6 = "SATWND", E9 = "RAD")) +
   coord_flip() +
   facet_grid(iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "a) Temperature",
-                                              "td" = "b) Dew point temperature",
-                                              "u" = "c) U wind",
-                                              "v" = "d) V wind"),
+             labeller = labeller(variable = c("t" = "Temperature (K)",
+                                              "td" = "Dew point temperature (K)",
+                                              "u" = "U wind (ms-1)",
+                                              "v" = "V wind (ms-1)"),
                                  iop = c("IOP07" = "IOP 7",
                                          "IOP08" = "IOP 8"))) +
-  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
+   tagger::tag_facets(position = list(x = 0.05, y = 0.95)) +
+  labs(y = "Bias/RMSE", x = "Height (Km)", color = NULL, linetype = NULL) +
   theme_minimal(base_size = 9) +

### 2022-04-11 analysis/chapters/paper_ana_20181122.Rmd
new version of paper

@@
 abstract: |
-  In this paper we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF.
-We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November 22, 2018.
+  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north-eastern Argentina where the surface and upper air observing networks are relatively coarse.
+We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed during November 22, 2018.
 The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
-This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
-
-  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
-A multiphysics approach was also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, ASWS combining CONV and dense automatic surface weather station networks, SATWND, combining ASWS with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
+The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
+
+  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members multiphisics  ensemble.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
+We also run ensemble forecasts initialized from each experiment show and evaluated their skill to predict precipitation.
+We found that the assimilated observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
 journal: Atmospheric Research
@@
 A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
-Southern South America has been shown to have the highest frequency of favorable conditions for high-impact meteorological events [@brooks2003] and the highest occurrence of large hail events [@cecil2012].
+Southern South America has been shown to have one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer.
 This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014].
-
-Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to several factors. 
-On the one hand, mesoscale phenomena have limited predictability, while on the other hand, there is great difficulty in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and in the order of minutes).
-To solve this problem, data assimilation (DA) at the mesoscale is a topic that has gained increasing interest since it can provide appropriate initial conditions for the initialization of high-resolution forecasts [@sun2014].
-
-DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations. 
-The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts. 
-For example, @ha2014 showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
-However, surface observations, which directly measure the atmospheric state, have limited availability in South America.
-Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale.
-Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products. 
-Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
-@miyoshi2012a assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons near Japan. 
-@dillon2019 also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation over Southern South America.
-
-However radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
-Moreover, @bao2015 studied the impact on temperature and humidity forecast over the western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
-Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
-Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts [@gao2015].
-
-Having analyses with higher temporal resolution than those available from global forecasting systems (e.g.
-at hourly or sub-hourly) and using an increased number of regional observations that are not assimilated in global DA systems are some of the advantages of regional DA.
-Previous work has shown promising results on mesoscale DA in the region using different approaches such as LETKF; [@dillon2016; @dillonInreview] and 3D-Var [@goncalvesdegoncalves2015]. 
-In particular, the use of ensemble-based methods has shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances.
-Moreover, the use of ensembles also provides valuable information about the uncertainties in the forecast.
-Recently a rapid-refresh ensemble-based DA system has been experimentally implemented during the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) Field Campaign (central Argentina, November-December 2018, @nesbitt2021) with promising results [@dillonInreview].
-This system assimilated data from high-resolution surface networks, AMDAR data from local aircraft flights, soundings, AIRS retrievals, high-resolution GOES-16 wind estimates, and local radar data every hour with a grid spacing of 10 km.
-The forecast initialized from the analyses shows an overall similar performance to cold-start initializations from the Global Forecast System analysis, and even a positive impact in some cases.
-
-One important challenge of regional data assimilation systems is the assessment of their representation of the true atmospheric state and the skill of the forecasts initialized using their analysis as initial conditions.
-In this sense, the observations and measurements taken during the intense observing period of RELAMPAGO are an invaluable resource to validate a regional data assimilation system.
-
-The main objective of this work is to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
-The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system. 
-The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
-On Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
-The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours. 
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MSC. 
+Recently, the RELAMPAGO-CACTI field campaign @nesbitt2021 has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
+
+Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability, and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes).
+Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for the initialization of high-resolution forecasts [@sun2014] and thus has received increasing attention in the last decades.
+
+For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used.
+
+In that regard, @wheatley2010 investigated the impact of assimilating surface pressure data in a mesoscale ensemble-based data assimilation system but found limited impact in two case studies related to mesoscale convective systems.
+@ha2014 showed that assimilating temperature and dew point temperature from high resolution surface weather station networks systematically improved the structure of the simulated planetary boundary layer and improved the short range precipitation forecast over the U.S.
+@chang2017, @bae2022 and @chen2016  reported beneficial effects of assimilating surface weather stations in a high resolution DA system using an ensemble Kalman filter, a 3-dimensional variational system and a 4-dimensional variational system respectively.
+Positive impacts have been found on the temperature and moisture forecast in the planetary boundary layer and in the location of precipitation systems.
+@sobash2015 showed in a mesoscale ensemble based DA system that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
+@maejima2019 investigated the impact of 1-minute frequency assimilation of surface observations using observation simulation experiments on a heavy rainfall case founding that assimilating high frequency and spatially dense observations lead to a better representation of mesoscale circulation even though the number of observations provided by surface stations is much less than the one provided by weather radars.
+@gasperoni2018 conducted a case study to evaluate the impact of assimilating observations produced by private weather stations which are not incorporated in the operational analysis.
+They found a positive effect of these observations upon the initiation of deep moist convection along a dryline.
+This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.
+
+The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs) has been investigated in the context of limited-area mesoscale data assimilation.
+Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.
+@wu2014, @cherubini2006, and @sawada2019, among many others).
+Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm.
+However, some works indicated mixed impacts (e.g.
+@sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecasted intensity).
+As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
+@zhao2021 and @zhao2021a assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
+They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts.
+@otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.
+
+While the assimilation of radiance observations into global models is well established (), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low top models used for this application.
+@bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
+@wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis.
+They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere.
+They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on the short-range forecasts (12 and 24 h).
+Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land.
+Moreover, to the best of the author's knowledge, there are few to none studies related to the direct assimilation of radiance observations over South America.
+
+In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed.
+Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources [e.g.
+@dillon2016; @dillon2021; @goncalvesdegoncalves2015].
+In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 AMVs, and satellite temperature and moisture retrievals with promising results.
+However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above.
+The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated, and ensemble-based DA system.
+In particular, in this paper we will focus on the impact in the context of mid-latitude mesoscale convective system events.
+Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.
+
+To reach this goal, we conduct several DA experiments for a case study of a gigantic Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
+
 The paper is organized as follows.
@@
 
-The model simulations for the case study are performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2018) model.
-The resolution is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa. 
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal resolution and 6-hour temporal resolution, @cisl_rda_ds084.1). 
+## Case Study
+
+On Nov 22, a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
+The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours. 
+To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
+
+## Data assimilation system configuration
+
+The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2008) model.
+The horizontal grid spacing is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the model at 50 hPa. 
+The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). 
 The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.
 
-The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @hu2018).
+The analyses are generated using the LETKF implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @shao2016). 
 A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
-Observations are assimilated in a 4D approach by comparing them with the corresponding first-guess state at 10-minute intervals. 
+Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals. 
 For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures.
-
-To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillonInreview for all types of observations.
-A relaxation-to-prior inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
-
-We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis.
+In addition to being already coupled to the GSI system.
+
+We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis.
 Since the publicly available GFS ensemble members change at each cycle we used random perturbations to generate the initial ensemble perturbations.
-This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021.
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
-These random perturbations are also applied at the boundaries to keep the ensemble spread within the domain.
+This method helps to prevent an underestimation of the ensemble spread [@ouaraini2015], and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.
 
@@
 
-(ref:dominio) a) The horizontal domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
-b) Locations of radiosonde launches during the experiment.
-Green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions.
+To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations.
+A relaxation-to-prior inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
+
+(ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+b) Locations of radiosonde launches during the RELAMPAGO.
+Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions.
 The topography in meters is also shown (shaded).
@@
                      breaks = c("Sfc - Non-official", "Sfc - Official"),
-                     labels = c("ASWS","CSWS"), values = c(15, 17),
+                     labels = c("AWS","CSWS"), values = c(15, 17),
                      guide = guide_legend(override.aes = list(size = 2))) +
@@
                      breaks = c("Sfc - Non-official", "Sfc - Official"),
-                     labels = c("ASWS","CSWS"),
+                     labels = c("AWS","CSWS"),
                      guide = guide_legend(override.aes = list(size = 2))) +
@@
 
-### Conventional and Satellite-derived wind
-
-The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS).
-We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at National Centers for Environmental Prediction (NCEP) that consist of surface observations from 117 conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
-The orange triangles in Figure \@ref(fig:dominio) a), are the surface station locations included in this experiment.
-The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.
-Wind surface observations over oceans (ASCATW) come from scaterometters and are also part of the PREPBUFR files.
-
-Satellite-derived wind observations also come from GDAS and include information from GOES-16 (estimated from the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (visible and water vapor channels).
-Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations).
-
-We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of various public and private automatic station networks over Southern South America and are included in the RELAMPAGO Data Set repository [@garcia2019].
+### Conventional
+
+The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream.
+We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP).
+These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft.
+The orange triangles in Figure \@ref(fig:dominio) indicate the location of the surface station included in this experiment.
+The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes.
+Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.
+
+Table \@ref(tab:table-obs) lists all the observation types (i.e., surface pressure, temperature, specific humidity, and wind) available for each source, together with their associated errors.
+The observation errors were specified following the GSI default configuration.
+In some cases, the error varies with height and depends on the specific platform (aircraft and satellite-derived wind).
+In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and the model-simulated observation based on the first-guess) with a predefined threshold that depends on the observation error (also included in Table \@ref(tab:table-obs)).
+
+### AWS networks
+
+We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America.
+The dataset used in this study has been obtained at the RELAMPAGO Data Set repository [@garcia2019].
 These stations are indicated as green squares in Figure \@ref(fig:dominio)a.
 They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases.
-The variables measured by the stations varied between networks; all stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.
-
-Table \@ref(tab:table-obs) lists all the observation types (i.e., pressure, temperature, specific humidity, and wind) available for each source, together with their associated errors.
-The observation errors were specified following the GSI default configuration.
-In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind).
-In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and the model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \@ref(tab:table-obs)).
+All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.
+Observation errors used to assimilate these observations are the same as for the CSWS (see Table \@ref(tab:table-obs))
+
+### Satellite derived winds
+
+Satellite-derived wind observations are also included in the PREPBUFR files available every 6 hs,  and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels).
+Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations).
+Observation errors used to assimilate these observations are indicated in Table \@ref(tab:table-obs).
 
@@
 Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study.
-This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
 Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
 For this reason, the number of satellite observations varied significantly among cycles.
-This is evident in Figure 3 a) which shows the number of  observations assimilated for RAD experiment (see Section \@ref(exp)) at each analysis cycle.
-In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiment.
+In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 % of the total amount of assimilated radiances in our experiment.
 The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM.
 The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
-The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top (50 hPa) used in the experiments.
-
-The channels accepted for assimilation and their associated error were defined using the default GSI configuration.
+The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, on the observation quality (based on available data quality information) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
-Also, observations over the sea are preferred over those over land or snow [@hu2018].
+Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
+Also, observations over the sea are preferred to those over land or snow [@hu2018].
 
@@
 The bias correction (BC) has an air-mass dependent and an angle-dependent component [@zhu2014] and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$.
-Then, the bias corrected brightness temperature can be obtained as:
+Then, the bias corrected brightness temperature (BT_{bc}) can be obtained as:
 
@@
 
-GSI has a constant offset bias correction term ($p_0 = 1$) and the following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
+GSI has a constant offset bias correction term ($p_0 = 1$) and the remaining predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
 Scan angle-dependent bias is modeled as a 4th-order polynomial [@zhu2014].
@@
 In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves the $\beta_i$ that provides the best fit between the simulation and the observations.
-The coefficients were initialized at 18 UTC Nov 11 with the GFS system coefficients, and were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
-Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.
+The coefficients were initialized at 18 UTC Nov 18 with the GFS system coefficients.
+The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.
 
@@
 For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes [@weston2019; @zhu2016].
-For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
-Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
+For the infrared channels, cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
+Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenith angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
 It also performs an emissivity check for observations over land for both infrared and microwave radiances.
@@
       col.names = linebreak(c("Sensor", "Platform", "Assimilated\nchannels", "Percentage\nover total")),
-      caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation, and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>%
+      caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>%
   kable_styling(font_size = 7) %>% 
@@
 
-(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) ASWS c) SATWND, and d) RAD experiments calculated over 2.5$^{\circ}$ boxes.
+(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AWS c) SATWND, and d) RAD experiments calculated over 2.5$^{\circ}$ boxes.
 
@@
   geom_mapa() +
-  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
+  facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                                E6 = "SATWND", E9 = "RAD"))) +
@@
 
-(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), ASWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
+(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
 
@@
   scale_color_manual(values = colores_exp, 
-                     labels = c(E2 = "CONV", E5 = "ASWS", 
+                     labels = c(E2 = "CONV", E5 = "AWS", 
                                 E6 = "SATWND", E9 = "RAD")) +
-  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "ASWS", 
+  scale_shape_manual(values = c(15, 16, 17, 18), labels = c(E2 = "CONV", E5 = "AWS", 
                                                             E6 = "SATWND", E9 = "RAD")) +
@@
 
-To evaluate the performance of the ensemble-based data assimilation system presented in this article, we use the following observations datasets:
-
-* The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
+To evaluate the performance of the ensemble-based DA system presented in this article, we use the following observational datasets:
+
+* The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the IOPs along with the operational radiosondes.
 Only the soundings that did not enter the assimilation system were used for validation.
@@
 The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes.
-For this work, only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
+For this work, only the maximum reflectivity in the column (COLMAX) closest to the analysis time was used.
 
@@
 
-To investigate the impact of different observations upon the analysis, four data assimilation experiments were performed using different observation sets (Table \@ref(tab:table-exp)).
-The CONV experiment uses only convectional observations from PREPBUFR.
-In a second experiment, referred to as ASWS, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from ASWS.
-In the third experiment, referred to as SATWND, we assimilate all the observations of the ASWS experiment and the satellite-derived motion winds.
-Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbit satellites as described in Section \@ref(sat).
+To investigate the impact of different observations upon the analysis, four DA experiments were performed using different observation sets (Table \@ref(tab:table-exp)).
+The CONV experiment uses only conventional observations from PREPBUFR.
+In a second experiment, referred to as AWS, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from AWS.
+In the third experiment, referred to as SATWND, we assimilate all the observations of the AWS experiment and the satellite-derived winds.
+Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbiting satellites as described in Section \@ref(sat).
 
 The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \@ref(fig:obs-horizontal).
-The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations.
-In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet [@vera2006].
+The larger number of assimilated observations over the center and east of the domain corresponds to the AWS observations.
 In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown.
-Local maxima at 12 and 00 UTC  found mainly in CONV is produced by the operational soundings.
+Local maxima at 12 and 00 UTC found mainly in CONV are attributed to operational soundings.
 The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
 The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors.
-The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b) shows a maximum in low levels due to the ASWS observations.
-Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa).
+The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b) shows a maximum in low levels due to the AWS observations.
+Satellite-derived winds available in this particular case are maximized at the upper troposphere (between 500-250 hPa).
 Above 850 hPa, most of the observations correspond to radiance observations.
@@
 All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
-The initial ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \@ref(fig:cycle)).
+The initial 60-member ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \@ref(fig:cycle)).
+
+Ensemble forecast initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
+Both forecasts are integrated until 12 UTC Nov 23.
+All forecasts use the same domain and ensemble configuration as the analysis.
+The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1).
 
@@
   "Conventional (PREPBUFR)", "x", "x", "x", "x",
-  "Conventional (ASWS)", " ", "x", "x", "x",
+  "Conventional (AWS)", " ", "x", "x", "x",
   "Satellite-derived winds", " ", " ", "x", "x",
@@
 ) %>% 
-  kbl(booktabs = TRUE, col.names = c("Obs type", "CONV", "ASWS", "SATWND", "RAD"),
+  kbl(booktabs = TRUE, col.names = c("Obs type", "CONV", "AWS", "SATWND", "RAD"),
       align = "lcccc",
@@
 A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper.
-These aspects include a validation of how the uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations that are not assimilated.
+These aspects include a validation of how the uncertainty is quantified in the first-guess and in the analysis, and how different experiments fit an independent set of observations that are not assimilated.
 
@@
 
-where $x_o$ is the observation and its error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
+where $x_o$ is the assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble. 
 The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
@@
 where $M$ is the sample size.
-A reliable system will have no bias ($mean RCRV = 0$), and a standard deviation equal to 1 ($sd RCRV = 1$).
+A consistent system will have no bias ($mean RCRV = 0$) and a standard deviation equal to 1 ($sd RCRV = 1$).
 If the ensemble has a positive (negative) bias, $mean RCRV$ will be positive (negative).
-A $sd RCRV > 1$ ($sd RCRV < 1$), indicates that the ensemble is underdispersive (overdispersive).
-
-The fit of the first-guess and analysis to a set of independent observations is computed based on the Root Mean Square Error (RMSE) and the BIAS:
+A $sd RCRV > 1$ ($sd RCRV < 1$) indicates that the ensemble is underdispersive (overdispersive).
+
+The fit of the first-guess and analysis to a set of independent observations, the high-resolution radiosondes from RELAMPAGO,  is computed based on the Root Mean Square Error (RMSE) and the BIAS:
 
@@
 
-For the comparison of the first-guess’ precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS) for different sized sampling areas and thresholds, following @maldonado2021 and adaptations from @roberts2008:
+For the comparison of the first-guess precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds.
+Following @maldonado2021 the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members :
 
@@
 
-where ${P_x}$ is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and ${P_o}$ is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise.
-This probabilistic formulation uses the information from the full ensemble thus avoiding the smoothing effect of the ensemble mean which can particularly affect localized extreme precipitation values.
-The FSS was applied to the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
+where $P_{xi}$ is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and $P_{oi}$ is the proportion of grid points in each sampling area in which the observed accumulated precipitation is over the same threshold.
+Since the precipitation forecast of all the ensemble members is considered in the computation of $P_{xi}$, the smoothing effect of the ensemble mean is avoided.
+The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
 
@@
 We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019].
-All the analysis in this paper were performed using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
+All the analyses in this paper were performed using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
 All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2019].
@@
 
-## Ensemble reliability
+## Ensemble consistency
 
 To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment.
-As this experiment assimilates all types of observations used in this work, it is possible to analyze the reliability of the ensemble by comparing it with each type of observation.
+As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation.
 Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid.
 The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error.
-For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas to the west of the domain where the $sd RCRV$ can be as high as 4.5.
+For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas in the west of the domain the $sd RCRV$ can be as high as 4.5.
 In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.
@@
 AIRCAR presents an irregular profile with $sd RCRV$ values that suggest that the error for this type of observation is overestimated.
-ADPUPA and AIRCAR present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dillonInreview.
-
-For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias and an overestimation of the observation error.
-Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero.
+ADPUPA and AIRCAR present a $mean RCRV$ profile near zero at middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dillon2021.
+
+For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) at low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error.
+Wind estimations derived from water vapor channels are abundant above 500 hPa where their bias is close to zero.
 The only exception are the EUMETSAT observations which contribute very little in the region.
 
-The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) show almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor.
+The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) show almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor (not shown).
 This indicates that the bias correction algorithm works as expected.
@@
 
-Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
-Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate for the system.
-
-(ref:rcrv-sfc) $sd RCRV$ calculated for surface observations (from PREPBUFR and ASWS) of a) wind, and b) temperature, averaged over 2.5º boxes.
+Overall, these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
+Moreover, these results suggest the inflation parameter $\alpha = 0.9$ is adequate for the system.
+
+(ref:rcrv-sfc) $sd RCRV$ calculated for surface observations (from PREPBUFR and AWS) of a) wind, and b) temperature, averaged over 2.5º boxes.
 
@@
 The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS.
-Figure \@ref(fig:TQ-diff)a-c shows the difference between experiments in the spatially averaged vertical profile of temperature.
+Figure \@ref(fig:TQ-diff)a-c shows the analysis difference between experiments in the spatially averaged vertical profile of temperature.
 By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
-During the first day, the assimilation of ASWS observations results in a colder PBL.
-This cooling effect has a clear diurnal cycle being stronger during nighttime (Figure \@ref(fig:TQ-diff)a).
-During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
-The warm difference shown in ASWS with respect to CONV  in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV.
+During the first day, the assimilation of AWS observations results in a colder PBL.
+This cooling effect has a clear diurnal cycle, being stronger during nighttime (Figure \@ref(fig:TQ-diff)a).
+During the second day of the experiment, the impact of AWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
+The warm difference shown in AWS-CONV between 500 and 200 hPa is produced by the development of stronger convection in AWS compared to CONV.
 This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection.
-Figure \@ref(fig:TQ-diff)b, indicates that SATWND observations have an almost negligible impact on the mean temperature through the experiment.
+Figure \@ref(fig:TQ-diff)b indicates that SATWND observations have an almost negligible impact on the mean temperature throughout the experiment.
 This result is unexpected since mid-to-upper level winds can have a profound impact on the organization and evolution of MCSs over the region.
-During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of ASWS observations.
+During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations.
 No clear systematic impact is found above the PBL during this period.
-During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the ASWS experiment but with the opposite sign.
-
-Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \@ref(fig:TQ-diff)d).
-The PBL in the ASWS experiment is consistently moister than in the CONV experiment and this is enhanced during the nighttime.
+During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AWS experiment but with the opposite sign.
+
+Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f), the impact of assimilating AWS with finer spatial and temporal resolution is most substantial at low levels (Figure \@ref(fig:TQ-diff)d).
+The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime.
 The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@ruiz2010].
@@
 In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
-As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels.
+As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels (Figure \@ref(fig:TQ-diff)e).
 Figure \@ref(fig:TQ-diff)f shows that radiances reduce low-middle level moisture during the first day of the experiment.
@@
 
-The impacts on the wind components are shown in Figure \@ref(fig:UV-diff) along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
-The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
-There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with a stronger reduction of the northerly wind during night hours.
-This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment.
-After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain.
-After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels.
-During Nov 22 and 23 the impact of assimilating ASWS produces an increase of northerly wind in upper levels.
-This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
-Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
-
-The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure  \@ref(fig:UV-diff)c and f).
-For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of ASWS observations.
-On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind.
-When we look at the meridional wind field at 200 hPa at different times, we observe that the outflow from the MCS is even more intense than in the other experiments, while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in  Figure \@ref(fig:UV-diff)f.
-
-To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles).
-At that moment the first convective cells were developing on the southern region of the domain along the cold front.
-Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
-We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
-The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
-ASWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
-RAD has lower precipitable water content than ASWS and SATWND but higher than CONV.
-The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
-
-The potential temperature on the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed on the temperature profiles in Figure \@ref(fig:TQ-diff)a-c.
-On average the PBL in ASWS and SATWND is colder than in CONV while RAD shows a warmer PBL than ASWS due to the assimilation of radiance observations.
-Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear.
-The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$.
-MCAPE in the RAD experiment is lower compared to ASWS or SATWND.
-This is consistent with less humidity in the PBL with respect to these experiments and may be partially compensated by a slightly warmer PBL in the RAD experiment.
-The 0-6 km wind shear is more intense in ASWS, SATWND, and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
-Moreover, in this same region, these experiments show larger MCAPE values than CONV.
-Note that wind shears over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowsky2010].
-
-(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and b) ASWS-CONV, b) and e) SATWND-ASWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-
-```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=6, fig.align="center",fig.pos="th"}
+(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+
+```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=5.5, fig.align="center",fig.pos="th"}
 files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
@@
   facet_wrap(~variable, ncol = 3,
-             labeller = labeller(variable = c("E5_E2" = "ASWS - CONV",
-                                              "E6_E5" = "SATWND - ASWS",
+             labeller = labeller(variable = c("E5_E2" = "AWS - CONV",
+                                              "E6_E5" = "SATWND - AWS",
                                               "E9_E6" = "RAD - SATWND"))) +
@@
 ```
-
-(ref:UV-diff)  Difference between analysis ensemble mean experiments a) and b) ASWS-CONV, b) and e) SATWND-ASWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-Black contours corresponde to u wind for ASWS (a), SATWND (b) and RAD (c) and v wind for ASWS (d), SATWND (e) and RAD (f).
-
-```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=6, fig.align="center", fig.pos="ht"}
+The impacts on the wind components are shown in Figure \@ref(fig:UV-diff), along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations (for example, Figure \@ref(fig:UV-diff)a shows the zonal wind difference between AWS and CONV and the zonal wind for AWS).
+The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis.
+There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with a stronger reduction of the northerly wind during night hours.
+This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment.
+After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain.
+After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels.
+During Nov 22 and 23 the impact of assimilating AWS produces an increase of northerly wind in upper levels.
+This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
+Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger; the systematic impact is overall smaller than the one produced by assimilating data from AWS  (Figure  \@ref(fig:UV-diff)b, e).
+
+The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \@ref(fig:UV-diff)c).
+For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of AWS observations during the nights previous to the development of the MCS.
+At upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind speed.
+The meridional wind field at 200 hPa at different times shows that the outflow from the MCS is even more intense than in the other experiments, while the southerly wind ahead of the MCS also increases producing an average reduction of the northerly wind (Figure \@ref(fig:UV-diff)f).
+
+(ref:UV-diff)  Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+Black contours correspond to u wind for AWS (a), SATWND (b) and RAD (c) and v wind for AWS (d), SATWND (e) and RAD (f).
+
+```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=5.5, fig.align="center", fig.pos="ht"}
 perfiles %>% 
@@
   facet_wrap(~variable, ncol = 3,
-             labeller = labeller(variable = c("E5_E2" = "**ASWS** - CONV",
-                                              "E6_E5" = "**SATWND** - ASWS",
+             labeller = labeller(variable = c("E5_E2" = "**AWS** - CONV",
+                                              "E6_E5" = "**SATWND** - AWS",
                                               "E9_E6" = "**RAD** - SATWND"))) +
@@
 ```
+To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles).
+At that moment the first convective cells were developing over the southern region of the domain along the cold front.
+Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
+We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
+The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
+AWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
+RAD has lower precipitable water content than AWS and SATWND, but higher than CONV.
+The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating AWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
+
+The analyzed distribution of temperature and moisture in the PBL (Figure \@ref(fig:summary-fields)b) resembles the observations (Figure \@ref(fig:TQ-diff)a-c).
+On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
+Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear.
+The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$.
+MCAPE in the RAD experiment is lower compared to AWS or SATWND.
+This is consistent with less humidity in the PBL with respect to these experiments but may be partially compensated by a slightly warmer PBL in the RAD experiment.
+The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
+Moreover, in this same region, these experiments show larger MCAPE values than CONV.
+Note that wind shears over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowski2010].
 
@@
 All fields correspond to the analysis ensemble mean for 00 UTC Nov, 22.
-Grey filled contours corresponde to  topography over 1500 meter over sea level.
+Grey filled contours correspond to  topography over 1500 meter over sea level.
 
@@
   facet_grid(. ~ exp, 
-             labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
+             labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                          E6 = "SATWND", E9 = "RAD"))) +
@@
 
-Firstly we analyze the impact of different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \@ref(fig:pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitude range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and the estimated precipitation by IMERG.
-The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period.
+Firstly we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
+Figure \@ref(fig:pp-hov)(a) shows the mean hourly accumulated precipitation as estimated by IMERG, and (b-e) the probability matching mean (PM) [@clark2017] for the first-guess hourly accumulated precipitation as averaged between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments.
+The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \@ref(fig:pp-hov)a).
 In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated.
-This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation.
-ASWS, SATWND, and RAD better capture the timing and location of convective initiation.
-On RAD the initiation is delayed when we compare with ASWS and SATWND, possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
-After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
+This is particularly evident in CONV (Figure \@ref(fig:pp-hov)b) where the convection initiation is delayed and occurs further north with respect to the observed initiation.
+AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figure \@ref(fig:pp-hov)c-e).
+AWS and RAD show a more fragmented distribution compared with SATWND, possibly due to the development of less organized convection during Nov 22.
+After 18 UTC Nov 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
 
@@
   scale_y_latitude(breaks = seq(-40, -20, 5), labels = NULL) +
-  facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
+  facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
                                                           E6 = "SATWND", E9 = "RAD"))) +
@@
 
-To quantify the spatial match between the observed precipitation and the short-range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
-All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature stage of the MCS.
+To quantify the spatial match between the observed precipitation and the short-range forecast precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
+All experiments show values of FSS below 0.6 around 03 UTC Nov 22 indicating that the experiments were not able to accurately simulate the convective environment and the initialization of the convection.
+However, we notice that the FSS values for RAD around 00 UTC Nov 22 are higher than for the other experiment, indicating that radiance observations have a positive impact on the analysis.
+Moreover, the FSS values improve during the mature stage of the MCS for all spatial scales and thresholds.
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
-ASWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
-The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
+AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
+The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22.
 The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).
 
-(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), ASWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
+(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), AWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
 
@@
   # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "ASWS", 
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AWS", 
                                                       E6 = "SATWND", E9 = "RAD")) +
@@
   coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
-  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
-                                            "10" = "10 mm", "25" = "25 mm"),
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mmh-1", "5" = "5 mm", 
+                                            "10" = "10 mm", "25" = "25 mmh-1"),
                                       w = c("1" = "10 km", "5" = "50 km", "11" = "100 km", 
@@
 
-To complement the analysis, Figure \@ref(fig:dbz-mean) shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity.
-These experiments were chosen because they represent the analysis with the minimum and the maximum number of assimilated observations.
-In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the short-range precipitation forecast skill (Figure \@ref(fig:pp-hov)).
-Overall none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
-This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately represent the strength of the convective band associated with the MCS.
-Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field.
+To complement the analysis, Figure \@ref(fig:dbz-mean) shows the observed maximum reflectivity in the vertical column (COLMAX) and the ensemble mean COLMAX for the CONV and RAD experiments at different times between 10 and 19 UTC Nov 22.
+These experiments were chosen because they represent the analysis with the minimum (CONV) and maximum (RAD) number of assimilated observations.
+In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the 1-hour precipitation forecast skill (Figure \@ref(fig:fss)).
+Overall, none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
+This is partially expected considering the coarse horizontal grid spacing (10 km), which is not enough to appropriately represent the strength of the convective band associated with the MCS.
+Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecast reflectivity field.
 RAD better represents the observed features of the system showing a  stronger and more organized MCS than CONV, over the domain center at 10 and 13 UTC (first and second columns in Figure \@ref(fig:dbz-mean)).
-The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in RAD.
+The convective cells that initiate after 16 UTC along the warm front in the northeast part of the domain are well captured by both experiments but are better represented in terms of strength in RAD.
 In addition, CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
-Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between experiment and observations is quite high in the regions where radar data are available particularly for RAD.
-
-Finally, Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between experiment and observations is quite good in the regions where radar data are available, especially for RAD.
+
+Finally, Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes), and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
 
@@
 The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection.
-For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that is slightly reduced by the assimilation of ASWS and radiances.
-The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the ASWS, SATWND, and RAD is small above the PBL.
-The dew point temperature shows a small dry bias in low levels that is corrected on ASWS, SATWND, and RAD.
-On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C) for all the experiments, indicating a more moist atmosphere in the analyses than the observed.
+For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that is slightly reduced by the assimilation of AWS and radiances (Figure \@ref(fig:soundings)a).
+The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AWS, SATWND, and RAD is small above the PBL.
+The dew point temperature shows a small dry bias in low levels that is corrected in AWS, SATWND, and RAD (Figure \@ref(fig:soundings)b).
+At middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C) for all the experiments, indicating a more moist atmosphere in the analyses than observed.
 Overall, the RMSE grows with height, probably because at upper levels moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
-The zonal wind has a significant positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced by assimilating satellite radiances.
-It is somewhat surprising that no significant effect is observed by assimilating satellite-derived winds.
-The assimilation of ASWS and radiance observations increases the northerly wind bias and RMSE near the surface.
+The zonal wind for AWS and SATWND has a substantial positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced to CONV values by assimilating satellite radiances (Figure \@ref(fig:soundings)c).
+It is somewhat surprising that no substantial effect is observed by assimilating satellite-derived winds.
+The assimilation of AWS and radiance observations increases the northerly wind bias and RMSE near the surface (Figure \@ref(fig:soundings)d).
 
@@
 
-For the IOP 8 (bottom row in Figure \@ref(fig:soundings)), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulations.
+For IOP 8 (Figure \@ref(fig:soundings)e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation.
 This area was also behind the cold front thus affected by low-level cold advection.
-Below 5 km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in ASWS, SATWND, and RAD to the assimilation of ASWS.
-The RMSE for ASWS, SATWND, and RAD remains below 2$^{\circ}$C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact.
+Below 5 km all experiments show a warm bias which decreases with the assimilation of AWS and radiances.
+Between 5 and 12 km, the bias in each experiment is cold and is reduced in AWS, SATWND, and RAD.
+The RMSE for AWS, SATWND, and RAD remains below 2$^{\circ}$C for almost all levels with substantial improvement in upper levels where radiance observations have the largest impact.
 In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
 This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region.
-The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
-SATWND presents the lowest RMSE overall while ASWS and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C.
-The zonal wind is overestimated on the analyses with a positive bias at middle levels.
+The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation, but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
+SATWND presents the lowest RMSE overall while AWS and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C.
+The zonal wind is overestimated in the analyses with a positive bias at middle levels.
 For this period only RAD shows an improvement with respect to CONV in the upper troposphere.
-At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in ASWS, SATWND, and RAD.
-In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
+At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD.
+In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).
 
 (ref:soundings) RMSE (solid line) and Bias (dashed line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
-The blue line corresponds to CONV, the light blue line to ASWS, SATWND is represented with an orange line, and RAD with a red line.
+The blue line corresponds to CONV, the light blue line to AWS, SATWND is represented with an orange line, and RAD with a red line.
 
@@
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = c(colores_exp, "black"), labels = c(E2 = "CONV", E5 = "ASWS", 
+  scale_color_manual(values = c(colores_exp, "black"), labels = c(E2 = "CONV", E5 = "AWS", 
                                                                   E6 = "SATWND", E9 = "RAD")) +
@@
 ```
+## Ensemble forescast validation
+
+In this section, we analyse the 60-members ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
+We calculated the FSS for the ensemble forecast in 6-hours moving windows for the same thresholds and spatial scales as for the analysis to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
+CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
+AWS, SATWND, and RAD show an important improvement in the FSS values, particularly for the higher threshold (Figure \@ref(fig:fssfcst)b, d).
+Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
+
+The satellite-derived wind observations show a clearly positive impact on the forecast, in contrast to what was seen when comparing the analysis with independent observations in terms of precipitation.
+Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the analysis to IMERG estimations.
+
+(ref:fssfcst) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the forecasts initialized from CONV (blue line), AWS (light blue line), SATWND (orange line) and RAD (red line) experiments at 00 UTC (solid line) and 06 UTC (dashed line).
+
+```{r fssfcst, fig.cap="(ref:fssfcst)", fig.width=3.5, fig.height=4}
+
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,9]_ens.csv")
+
+fss <- purrr::map(files, function(f) {
+
+  meta <- unglue(basename(f), "fss_6h_fcst_ens_{ini_date}_{exp}_ens.csv")
+
+  fread(f) %>% 
+    .[, date := as_datetime(date)] %>% 
+    .[, ini_date := ymd_h(meta[[1]][["ini_date"]])]
+}) %>% 
+  rbindlist()
+
+fss %>% 
+  .[q %in% c(1, 25) & w %in% c(1, 11)] %>% 
+  ggplot(aes(date, fss)) +
+  geom_hline(yintercept = 1, color = "darkgray") +
+  geom_line(aes(color = exp, linetype = factor(ini_date))) +
+  # geom_point(aes(color = exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT",
+                                                      E6 = "SATWND", E9 = "RAD")) +
+  scale_linetype_manual(values = c(1, 2), labels = c("2018-11-22 00:00:00" = "OO UTC", 
+                                   "2018-11-22 06:00:00" = "06 UTC")) +
+  scale_date(20181122000000, 20181123000000, "12 hours") +
+  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mmh-1", "5" = "5 mm", 
+                                            "10" = "10 mm", "25" = "25 mmh-1", "50" = "50 mm"),
+                                      w = c("1" = "10 km", "5" = "50 km", 
+                                            "11" = "100 km", 
+                                            "51" = "500 km", "101" = "1000 km"))) +
+  tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
+  labs(color = NULL, linetype = NULL, x = NULL, y = "FSS") +
+  theme_minimal(base_size = 8) +
+  theme(legend.position = "bottom",
+        legend.box = "vertical", 
+        legend.margin = margin())
+
+```
 
@@
 
-We performed four experiments to assess the impact of different observation types, including conventional data (CONV), frequent surface observations (ASWS), satellite-derived winds (SATWND), and clear-sky satellite radiances (RAD), into a regional frequent-update ensemble-based data assimilation system.
-We used a case study approach of a massive MCS developed over Southern South America on Nov 22, 2018.
-
-We evaluated the reliability of the ensemble using the RCRV calculated for each type of observation used in RAD.
+In this paper, we investigate for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional data assimilation system.
+In particular, we explore the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors.
+Southern South America is a particularly interesting region due to the heterogeneity and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
+This combined with a climatology characterized by frequent organized convective events makes mesoscale data assimilation particularly challenging.
+To consider these issues, we used a case study approach of a massive mesoscale convective system that developed over Southern South America, particularly in a region where the surface and upper-air observing networks are too coarse to capture variability at the mesoscale and that took place on Nov 22, 2018.
+
+We evaluated the consistency of the ensemble using the RCRV calculated for each type of observation used in RAD.
 Conventional observations including surface data and upper air observations from soundings and aircraft, show a $sd RCRV$ near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
 For satellite-derived winds and radiance observations, the $sd RCRV$ is lower than 1, which could be the result of an underdispersive ensemble or an overestimation of the observation errors.
-However, taking into account that the errors of this type of observation are usually overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the assembly has a reasonable spread.
-
-In terms of the analysis, we found that ASWS observations, which have high spatial and temporal resolution, produce major impacts throughout the troposphere and mainly on the PBL.
-In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
-We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
+However, taking into account that the observation error for satellite-derived wind is overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the ensemble has a reasonable spread.
+
+In terms of the analysis, we found that AWS observations, which have high spatial and temporal resolution, produce major impacts throughout the troposphere and mainly on the PBL.
+In particular, the precipitable water content and the low level meridional circulation in the AWS experiment lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
+We also found positive results when assimilating radiance observations that produce an adequate development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
 However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates.
-While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels available for this case study.
+While the assimilation of satellite-derived wind does not produce a noticeable impact on the analysis with respect to the AWS experiment, this is possibly due to the small number of observations in low levels available for this case study and its large observation error.
+However, there are substantial improvements for the accumulated precipitation distribution.
 A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels.
 
-Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during the IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement on low levels when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind.
-In particular, ASWS observations correct a dry bias already observed by @ruiz2010.
+Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during IOPs 7 and 8.
+For the pre-convective environment (IOP 7), we observed an improvement at low level when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
+In particular, AWS observations correct a dry bias already observed by @ruiz2010.
 For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels including more observations for the temperature but the opposite effect for the dew point temperature.
-In this case, the analyses have a moist bias that increases with the assimilation of ASWS and radiances observations.
-We observed a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
+In this case, the analyses have a moist bias that increases with the assimilation of AWS and radiance observations.
+We noted a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
 It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period, which requires further analysis.
 
-To summarize this study, we conclude that the assimilation of surface observations with higher spatial and temporal resolution and clear-sky radiances from polar orbit satellites had an overly positive impact on the development of the studied MCS and its associated precipitation.
-In the future, we will analyze whether these impacts carry over in time to the independent forecasts initialized from the analyses.
+We also analyzed the performance of independent ensemble forecasts intialized from the analyses to estimate the precipitation during Nov 22.
+The forecast initialized from AWS, SATWND, and RAD are able to estimate the precipitation substantially better than CONV.
+In particular, the continuous assimilation of atellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
+
+To summarize, in this study we conclude that the assimilation of surface observations with higher spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
+Moreover, the ensemble forecasts initialized from the analysis show promising results to estimate the precipitation of severe events.
+In the future, we will further analyze the impacts in the  independent forecasts  and evaluate the implementation of other sources of observations like  GPS radio occultation data and radiances from geostationary orbiting satellites.
 
@@
 We also acknowledge the Cheyenne HPC resources (doi:10.5065/D6RX99HX) from NCAR's Computational and Information Systems Laboratory, National Science Foundation (project code UIUC0012).
-Also, PICT 2017-2033,  PICT 2016-0710 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
+Also, PICT 2017-2033 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
 

### 2022-04-11 analysis/chapters/paper_ana_20181122.tex
new version of paper

@@
             pdfauthor={},
-            pdftitle={Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis},
+            pdftitle={Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO.
+Part I: Impact on the Analysis and First-guess},
             colorlinks=false,
@@
 
-  \title{Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis}
+  \title{Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis and First-guess}
     \author[UBA,CIMA,CNRS]{Paola Belén Corrales\corref{1}}
@@
   \begin{abstract}
-  In this paper we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF.
-We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November 22, 2018.
+  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north-eastern Argentina where the surface and upper air observing networks are relatively coarse.
+We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed during November 22, 2018.
 The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
-This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
-
-  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance.
-A multiphysics approach was also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations).
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
+The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
+
+  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members multiphisics ensemble.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files, AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
+We also run ensemble forecasts initialized from each experiment show and evaluated their skill to predict precipitation.
+We found that the assimilated observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
   \end{abstract}
@@
 A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
-Southern South America has been shown to have the highest frequency of favorable conditions for high-impact meteorological events (Brooks et al., 2003) and the highest occurrence of large hail events (Cecil and Blankenship, 2012).
+Southern South America has been shown to have one of the highest frequencies in the world of favorable conditions for high-impact meteorological events (Brooks et al., 2003) and large hail events (Cecil and Blankenship, 2012), particularly during austral spring and summer.
 This is also confirmed by observational evidence and high impact weather reports (Matsudo et al., 2015; Rasmussen et al., 2014).
-
-Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to several factors.
-On the one hand, mesoscale phenomena have limited predictability, while on the other hand, there is great difficulty in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and in the order of minutes).
-To solve this problem, data assimilation (DA) at the mesoscale is a topic that has gained increasing interest since it can provide appropriate initial conditions for the initialization of high-resolution forecasts (Sun et al., 2014).
-
-DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations.
-The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts.
-For example, Ha and Snyder (2014) showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
-However, surface observations, which directly measure the atmospheric state, have limited availability in South America.
-Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale.
-Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products.
-Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e.~AIRS and IASI) in a regional domain application.
-Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons near Japan.
-Dillon et al. (2019) also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation over Southern South America.
-
-However radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.
-Moreover, Bao et al.
-(2015) studied the impact on temperature and humidity forecast over the western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
-Recently Zhu et al.
+Recently, the RELAMPAGO-CACTI field campaign Nesbitt et al.
+(2021) has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
+
+Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability, and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes).
+Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for the initialization of high-resolution forecasts (Sun et al., 2014) and thus has received increasing attention in the last decades.
+
+For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used.
+
+In that regard, Wheatley and Stensrud (2010) investigated the impact of assimilating surface pressure data in a mesoscale ensemble-based data assimilation system but found limited impact in two case studies related to mesoscale convective systems.
+Ha and Snyder (2014) showed that assimilating temperature and dew point temperature from high resolution surface weather station networks systematically improved the structure of the simulated planetary boundary layer and improved the short range precipitation forecast over the U.S.
+Chang et al.
+(2017), Bae and Min (2022) and Chen et al.
+(2016) reported beneficial effects of assimilating surface weather stations in a high resolution DA system using an ensemble Kalman filter, a 3-dimensional variational system and a 4-dimensional variational system respectively.
+Positive impacts have been found on the temperature and moisture forecast in the planetary boundary layer and in the location of precipitation systems.
+Sobash and Stensrud (2015) showed in a mesoscale ensemble based DA system that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
+Maejima et al.
+(2019) investigated the impact of 1-minute frequency assimilation of surface observations using observation simulation experiments on a heavy rainfall case founding that assimilating high frequency and spatially dense observations lead to a better representation of mesoscale circulation even though the number of observations provided by surface stations is much less than the one provided by weather radars.
+Gasperoni et al.
+(2018) conducted a case study to evaluate the impact of assimilating observations produced by private weather stations which are not incorporated in the operational analysis.
+They found a positive effect of these observations upon the initiation of deep moist convection along a dryline.
+This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.
+
+The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs) has been investigated in the context of limited-area mesoscale data assimilation.
+Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.~Wu et al.
+(2014), Cherubini et al.
+(2006), and Sawada et al.
+(2019), among many others).
+Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm.
+However, some works indicated mixed impacts (e.g.~Sawada et al.
+(2019) reported an improvement in the forecast of the track of the storm but a degradation in the forecasted intensity).
+As stated in J.
+Zhao et al.
+(2021a) and J.
+Zhao et al.
+(2021b), the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
+J.
+Zhao et al.
+(2021a) and J.
+Zhao et al.
+(2021b) assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
+They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts.
+Otsuka et al.
+(2015) and Mallick and Jones (2020) found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.
+
+While the assimilation of radiance observations into global models is well established (), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low top models used for this application.
+Bao et al.
+(2015) studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+More recently, Zhu et al.
 (2019) studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
-Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al., 2015).
-
-Having analyses with higher temporal resolution than those available from global forecasting systems (e.g.~at hourly or sub-hourly) and using an increased number of regional observations that are not assimilated in global DA systems are some of the advantages of regional DA.
-Previous work has shown promising results on mesoscale DA in the region using different approaches such as LETKF; (Dillon et al., n.d., 2016) and 3D-Var (gustavo Goncalves de Goncalves et al., 2015).
-In particular, the use of ensemble-based methods has shown great potential for data assimilation in these scales, as they are able to characterize the flow-dependent evolution of the background error covariances.
-Moreover, the use of ensembles also provides valuable information about the uncertainties in the forecast.
-Recently a rapid-refresh ensemble-based DA system has been experimentally implemented during the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) Field Campaign (central Argentina, November-December 2018, Nesbitt et al.
-(2021)) with promising results (Dillon et al., n.d.).
-This system assimilated data from high-resolution surface networks, AMDAR data from local aircraft flights, soundings, AIRS retrievals, high-resolution GOES-16 wind estimates, and local radar data every hour with a grid spacing of 10 km.
-The forecast initialized from the analyses shows an overall similar performance to cold-start initializations from the Global Forecast System analysis, and even a positive impact in some cases.
-
-One important challenge of regional data assimilation systems is the assessment of their representation of the true atmospheric state and the skill of the forecasts initialized using their analysis as initial conditions.
-In this sense, the observations and measurements taken during the intense observing period of RELAMPAGO are an invaluable resource to validate a regional data assimilation system.
-
-The main objective of this work is to evaluate the impact of different observation types, including frequent surface observations, satellite-derived winds, and clear-sky satellite radiances from different platforms, into a regional frequent-update ensemble-based data assimilation system.
-The experiments are conducted using the Local Ensemble Transform Kalman filter implemented within the GSI-WRF data assimilation system.
-The evaluation is performed using a case study approach involving a gigantic Mesoscale Convective System (MCS) developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
-On Nov, 22 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
-The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MSC.
+Wang and Randriamampianina (2021) studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis.
+They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere.
+They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on the short-range forecasts (12 and 24 h).
+Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land.
+Moreover, to the best of the author's knowledge, there are few to none studies related to the direct assimilation of radiance observations over South America.
+
+In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed.
+Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources (e.g.~Dillon et al., 2016, 2021; gustavo Goncalves de Goncalves et al., 2015).
+In particular Dillon et al.
+(2021) assimilated high resolution surface weather station networks, GOES-16 AMVs, and satellite temperature and moisture retrievals with promising results.
+However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above.
+The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated, and ensemble-based DA system.
+In particular, in this paper we will focus on the impact in the context of mid-latitude mesoscale convective system events.
+Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.
+
+To reach this goal, we conduct several DA experiments for a case study of a gigantic Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
+
 The paper is organized as follows.
@@
 
-The model simulations for the case study are performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al.
-(2018)) model.
-The resolution is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal resolution and 6-hour temporal resolution, National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
+\hypertarget{case-study}{%
+\subsection{Case Study}\label{case-study}}
+
+On Nov 22, a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
+The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
+To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
+
+\hypertarget{data-assimilation-system-configuration}{%
+\subsection{Data assimilation system configuration}\label{data-assimilation-system-configuration}}
+
+The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al.
+(2008)) model.
+The horizontal grid spacing is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the model at 50 hPa.
+The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
 Department of Commerce (2015)).
@@
 
-The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, Hunt et al.
-(2007)) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Hu et al.
-(2018)).
+The analyses are generated using the LETKF implementation (V1.3, Hunt et al. (2007)) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Shao et al. (2016)).
 A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within \(\pm\) 30 minutes of the analysis time are assimilated.
 Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
-For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) is used as an observation operator to calculate model-simulated brightness temperatures.
-
-To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in Dillon et al.
-(n.d.) for all types of observations.
-A relaxation-to-prior inflation (Whitaker and Hamill, 2012) is applied with an inflation parameter \(\alpha=0.9\) to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
-
-We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis.
+For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al.
+(2006)) is used as an observation operator to calculate model-simulated brightness temperatures.
+In addition to being already coupled to the GSI system.
+
+We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis.
 Since the publicly available GFS ensemble members change at each cycle we used random perturbations to generate the initial ensemble perturbations.
-This method also helps to prevent an underestimation of the ensemble spread (Ouaraini et al., 2015).
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in Necker et al.
@@
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
-These random perturbations are also applied at the boundaries to keep the ensemble spread within the domain.
+This method helps to prevent an underestimation of the ensemble spread (Ouaraini et al., 2015), and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.
 
@@
 
+To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in Dillon et al.
+(2021) for all types of observations.
+A relaxation-to-prior inflation (Whitaker and Hamill, 2012) is applied with an inflation parameter \(\alpha=0.9\) to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
-b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions.
+\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+b) Locations of radiosonde launches during the RELAMPAGO.
+Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions.
 The topography in meters is also shown (shaded).}\label{fig:dominio}
@@
 
-\hypertarget{conventional-and-satellite-derived-wind}{%
-\subsubsection{Conventional and Satellite-derived wind}\label{conventional-and-satellite-derived-wind}}
-
-The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS).
-We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at National Centers for Environmental Prediction (NCEP) that consist of surface observations from 117 conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft.
-The orange triangles in Figure \ref{fig:dominio} a), are the surface station locations included in this experiment.
-The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.
-Wind surface observations over oceans (ASCATW) come from scaterometters and are also part of the PREPBUFR files.
-
-Satellite-derived wind observations also come from GDAS and include information from GOES-16 (estimated from the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (visible and water vapor channels).
-Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 \% of the observations).
-
-We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of various public and private automatic station networks over Southern South America and are included in the RELAMPAGO Data Set repository (Garcia et al., 2019).
+\hypertarget{conventional}{%
+\subsubsection{Conventional}\label{conventional}}
+
+The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream.
+We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP).
+These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft.
+The orange triangles in Figure \ref{fig:dominio} indicate the location of the surface station included in this experiment.
+The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes.
+Wind surface observations over oceans (ASCATW) come from scatterometers and are also included in the PREPBUFR files.
+
+Table \ref{tab:table-obs} lists all the observation types (i.e., surface pressure, temperature, specific humidity, and wind) available for each source, together with their associated errors.
+The observation errors were specified following the GSI default configuration.
+In some cases, the error varies with height and depends on the specific platform (aircraft and satellite-derived wind).
+In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and the model-simulated observation based on the first-guess) with a predefined threshold that depends on the observation error (also included in Table \ref{tab:table-obs}).
+
+\hypertarget{aws-networks}{%
+\subsubsection{AWS networks}\label{aws-networks}}
+
+We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America.
+The dataset used in this study has been obtained at the RELAMPAGO Data Set repository (Garcia et al., 2019).
 These stations are indicated as green squares in Figure \ref{fig:dominio}a.
 They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases.
-The variables measured by the stations varied between networks; all stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.
-
-Table \ref{tab:table-obs} lists all the observation types (i.e., pressure, temperature, specific humidity, and wind) available for each source, together with their associated errors.
-The observation errors were specified following the GSI default configuration.
-In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind).
-In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and the model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \ref{tab:table-obs}).
+All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.
+Observation errors used to assimilate these observations are the same as for the CSWS (see Table \ref{tab:table-obs})
+
+\hypertarget{satellite-derived-winds}{%
+\subsubsection{Satellite derived winds}\label{satellite-derived-winds}}
+
+Satellite-derived wind observations are also included in the PREPBUFR files available every 6 hs, and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels).
+Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 \% of the observations).
+Observation errors used to assimilate these observations are indicated in Table \ref{tab:table-obs}.
 
@@
 \multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{*} Observation error varied with height.}\\
-\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{**} Observations over 600 hPa are rejected.}\\
+\multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{**} Observations above 600 hPa are rejected.}\\
 \multicolumn{5}{l}{\rule{0pt}{1em}\textsuperscript{+} Observation error depends on the report type.}\\
@@
 Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study.
-This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
 Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
 For this reason, the number of satellite observations varied significantly among cycles.
-This is evident in Figure 3 a) which shows the number of observations assimilated for RAD experiment (see Section \ref{exp}) at each analysis cycle.
-In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 \% of the total amount of assimilated radiances in our experiment.
+In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 \% of the total amount of assimilated radiances in our experiment.
 The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM.
 The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
-The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top (50 hPa) used in the experiments.
-
-The channels accepted for assimilation and their associated error were defined using the default GSI configuration.
+The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-Firstly a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, on the observation quality (based on available data quality information) and on the number of available channels (from the same pixel and sensor) that passed the quality control.
-Also, observations over the sea are preferred over those over land or snow (Hu et al., 2018).
+Firstly a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
+Also, observations over the sea are preferred to those over land or snow (Hu et al., 2018).
 
@@
 The bias correction (BC) has an air-mass dependent and an angle-dependent component (Zhu et al., 2014) and it is calculated as a multi-linear function of N predictors \(p_i(x)\), with associated coefficients \(\beta_i\).
-Then, the bias corrected brightness temperature can be obtained as:
+Then, the bias corrected brightness temperature (BT\_\{bc\}) can be obtained as:
 
@@
 
-GSI has a constant offset bias correction term (\(p_0 = 1\)) and the following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
+GSI has a constant offset bias correction term (\(p_0 = 1\)) and the remaining predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
 Scan angle-dependent bias is modeled as a 4th-order polynomial (Zhu et al., 2014).
@@
 In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves the \(\beta_i\) that provides the best fit between the simulation and the observations.
-The coefficients were initialized at 18 UTC Nov 11 with the GFS system coefficients, and were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section.
-Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.
+The coefficients were initialized at 18 UTC Nov 18 with the GFS system coefficients.
+The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.
 
@@
 For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes (Weston et al., 2019; Zhu et al., 2016).
-For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
-Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
+For the infrared channels, cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
+Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenith angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
 It also performs an emissivity check for observations over land for both infrared and microwave radiances.
@@
 \midrule
-AIRS & AQUA & 118 & 39.37 \%\\
+AIRS & AQUA & 52 & 31.63 \%\\
 \cmidrule{1-4}
- & AQUA & 7 & 1.13 \%\\
+ & N15 & 2 & 3.31 \%\\
 \cmidrule{2-4}
- & N19 & 13 & 1.55 \%\\
+ & N18 & 2 & 4.45 \%\\
 \cmidrule{2-4}
- & N15 & 13 & 3.93 \%\\
+\multirow[t]{-3}{*}{\raggedright\arraybackslash AMSUA} & METOP-A & 2 & 2.08 \%\\
+\cmidrule{1-4}
+ & METOP-A & 66 & 52.72 \%\\
 \cmidrule{2-4}
- & N18 & 13 & 3.17 \%\\
+\multirow[t]{-2}{*}{\raggedright\arraybackslash IASI} & METOP-B & 68 & 3.47 \%\\
+\cmidrule{1-4}
+ & N19 & 2 & 0.68 \%\\
 \cmidrule{2-4}
-\multirow[t]{-5}{*}{\raggedright\arraybackslash AMSUA} & METOP-A & 13 & 1.3 \%\\
-\cmidrule{1-4}
- & N20 & 21 & 1.16 \%\\
+ & METOP-A & 3 & 0.8 \%\\
 \cmidrule{2-4}
-\multirow[t]{-2}{*}{\raggedright\arraybackslash ATMS} & NPP & 21 & 0.86 \%\\
-\cmidrule{1-4}
- & N19 & 14 & 1.79 \%\\
-\cmidrule{2-4}
-\multirow[t]{-2}{*}{\raggedright\arraybackslash HIRS4} & METOP-A & 14 & 1.37 \%\\
-\cmidrule{1-4}
- & METOP-A & 174 & 40.07 \%\\
-\cmidrule{2-4}
-\multirow[t]{-2}{*}{\raggedright\arraybackslash IASI} & METOP-B & 174 & 2.94 \%\\
-\cmidrule{1-4}
- & N19 & 5 & 0.44 \%\\
-\cmidrule{2-4}
- & METOP-A & 5 & 0.46 \%\\
-\cmidrule{2-4}
-\multirow[t]{-3}{*}{\raggedright\arraybackslash MHS} & METOP-B & 5 & 0.44 \%\\
+\multirow[t]{-3}{*}{\raggedright\arraybackslash MHS} & METOP-B & 3 & 0.85 \%\\
 \bottomrule
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/obs-horizontal-1} \caption{Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AUT c) SATWND and d) RAD experiments calculate over 2.5\(^{\circ}\) boxes.}\label{fig:obs-horizontal}
+\includegraphics[width=1\linewidth]{../figures/obs-horizontal-1} \caption{Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AWS c) SATWND, and d) RAD experiments calculated over 2.5\(^{\circ}\) boxes.}\label{fig:obs-horizontal}
 \end{figure}
@@
 \begin{figure*}
-\includegraphics{../figures/obs-cycle-1} \caption{a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.}\label{fig:obs-cycle}
+\includegraphics{../figures/obs-cycle-1} \caption{a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.}\label{fig:obs-cycle}
 \end{figure*}
@@
 
-To evaluate the performance of the ensemble-based data assimilation system presented in this article, we use the following observations datasets:
+To evaluate the performance of the ensemble-based DA system presented in this article, we use the following observational datasets:
 
@@
 \item
-  The Multi-Network Composite Highest Resolution Radiosonde Data (Earth Observing Laboratory, 2020) from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements.
+  The Multi-Network Composite Highest Resolution Radiosonde Data (Earth Observing Laboratory, 2020) from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the IOPs along with the operational radiosondes.
 Only the soundings that did not enter the assimilation system were used for validation.
@@
 The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network (de Elía et al., 2017) with a temporal frequency of 10 minutes.
-For this work, only the maximum reflectivity in the column (COLMAX) field closest to the analysis time was used.
+For this work, only the maximum reflectivity in the column (COLMAX) closest to the analysis time was used.
 \end{itemize}
@@
 
-To investigate the impact of different observations upon the analysis, four data assimilation experiments were performed using different observation sets (Table \ref{tab:table-exp}).
-The CONV experiment uses only convectional observations from PREPBUFR.
-In a second experiment, referred to as AUT, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from ASWS.
-In the third experiment, referred to as SATWND, we assimilate all the observations of the AUT experiment and the satellite-derived motion winds.
-Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbit satellites as described in Section \ref{sat}.
+To investigate the impact of different observations upon the analysis, four DA experiments were performed using different observation sets (Table \ref{tab:table-exp}).
+The CONV experiment uses only conventional observations from PREPBUFR.
+In a second experiment, referred to as AWS, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from AWS.
+In the third experiment, referred to as SATWND, we assimilate all the observations of the AWS experiment and the satellite-derived winds.
+Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbiting satellites as described in Section \ref{sat}.
 
 The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \ref{fig:obs-horizontal}.
-The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations.
-In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet (Vera et al., 2006).
+The larger number of assimilated observations over the center and east of the domain corresponds to the AWS observations.
 In Figure \ref{fig:obs-cycle}a the number of assimilated observations over time is shown.
-Local maxima at 12 and 00 UTC found mainly in CONV is produced by the operational soundings.
+Local maxima at 12 and 00 UTC found mainly in CONV are attributed to operational soundings.
 The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
 The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors.
-The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the ASWS observations.
-Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa).
+The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the AWS observations.
+Satellite-derived winds available in this particular case are maximized at the upper troposphere (between 500-250 hPa).
 Above 850 hPa, most of the observations correspond to radiance observations.
@@
 All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
-The initial ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \ref{fig:cycle}).
+The initial 60-member ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \ref{fig:cycle}).
+
+Ensemble forecast initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
+Both forecasts are integrated until 12 UTC Nov 23.
+All forecasts use the same domain and ensemble configuration as the analysis.
+The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
+Department of Commerce (2015)).
 
@@
 \toprule
-Obs type & CONV & AUT & SATWND & RAD\\
+Obs type & CONV & AWS & SATWND & RAD\\
 \midrule
 Conventional (PREPBUFR) & x & x & x & x\\
-Conventional (ASWS) &  & x & x & x\\
+Conventional (AWS) &  & x & x & x\\
 Satellite-derived winds &  &  & x & x\\
@@
 
-where \(x_o\) is the observation and its error \(\sigma_o\), the mean \(m\), and the standard deviation \(\sigma\) of the ensemble.
+where \(x_o\) is the assimilated observation and its error \(\sigma_o\), the ensemble mean of the analysis in observational space \(m\), and the standard deviation \(\sigma\) of the ensemble.
 The average of \(RCRV\) computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
@@
 where \(M\) is the sample size.
-A reliable system will have no bias (\(mean RCRV = 0\)), and a standard deviation equal to 1 (\(sd RCRV = 1\)).
+A consistent system will have no bias (\(mean RCRV = 0\)) and a standard deviation equal to 1 (\(sd RCRV = 1\)).
 If the ensemble has a positive (negative) bias, \(mean RCRV\) will be positive (negative).
-A \(sd RCRV > 1\) (\(sd RCRV < 1\)), indicates that the ensemble is underdispersive (overdispersive).
-
-The fit of the first-guess and analysis to a set of independent observations is computed based on the Root Mean Square Error (RMSE) and the BIAS:
+A \(sd RCRV > 1\) (\(sd RCRV < 1\)) indicates that the ensemble is underdispersive (overdispersive).
+
+The fit of the first-guess and analysis to a set of independent observations, the high-resolution radiosondes from RELAMPAGO, is computed based on the Root Mean Square Error (RMSE) and the BIAS:
 
@@
 
-For the comparison of the first-guess' precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al.
-(2021) and adaptations from Roberts (2008):
+For the comparison of the first-guess precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds.
+Following Maldonado et al.
+(2021) the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members :
 
@@
 
-where \({P_x}\) is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and \({P_o}\) is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise.
-This probabilistic formulation uses the information from the full ensemble thus avoiding the smoothing effect of the ensemble mean which can particularly affect localized extreme precipitation values.
-The FSS was applied to the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
+where \(P_{xi}\) is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and \(P_{oi}\) is the proportion of grid points in each sampling area in which the observed accumulated precipitation is over the same threshold.
+Since the precipitation forecast of all the ensemble members is considered in the computation of \(P_{xi}\), the smoothing effect of the ensemble mean is avoided.
+The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
 
@@
 We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019).
-All the analysis in this paper were performed using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
+All the analyses in this paper were performed using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
 All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2019; Xie, 2015).
@@
 
-\hypertarget{ensemble-reliability}{%
-\subsection{Ensemble reliability}\label{ensemble-reliability}}
+\hypertarget{ensemble-consistency}{%
+\subsection{Ensemble consistency}\label{ensemble-consistency}}
 
 To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the \(mean RCRV\) and the \(sd RCRV\) for the RAD experiment.
-As this experiment assimilates all types of observations used in this work, it is possible to analyze the reliability of the ensemble by comparing it with each type of observation.
+As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation.
 Figure 5 shows the \(sd RCRV\) for surface observations box-averaged to a 2.5° grid.
 The \(sd RCRV\) for wind observations (Figure \ref{fig:rcrv-sfc}a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error.
-For the temperature (Figure \ref{fig:rcrv-sfc}b), the results are similar except that for some areas to the west of the domain where the \(sd RCRV\) can be as high as 4.5.
+For the temperature (Figure \ref{fig:rcrv-sfc}b), the results are similar except that for some areas in the west of the domain the \(sd RCRV\) can be as high as 4.5.
 In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.
@@
 AIRCAR presents an irregular profile with \(sd RCRV\) values that suggest that the error for this type of observation is overestimated.
-ADPUPA and AIRCAR present a \(mean RCRV\) profile near zero on middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in Ruiz et al.
+ADPUPA and AIRCAR present a \(mean RCRV\) profile near zero at middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in Ruiz et al.
 (2010) and Dillon et al.
-(n.d.).
-
-For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) in low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) show a larger departure from the expected behavior with a negative bias and an overestimation of the observation error.
-Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero.
+(2021).
+
+For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) at low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error.
+Wind estimations derived from water vapor channels are abundant above 500 hPa where their bias is close to zero.
 The only exception are the EUMETSAT observations which contribute very little in the region.
 
-The mean RCRV profiles calculated from the radiance observations (Figure \ref{fig:rcrv-profile}d) show almost no bias and the same happens if the \(mean RCRV\) is calculated over each channel of each sensor.
+The mean RCRV profiles calculated from the radiance observations (Figure \ref{fig:rcrv-profile}d) show almost no bias and the same happens if the \(mean RCRV\) is calculated over each channel of each sensor (not shown).
 This indicates that the bias correction algorithm works as expected.
@@
 
-Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
-Moreover, these results suggest the inflation parameter \(\beta = 0.9\) is adequate for the system.
+Overall, these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
+Moreover, these results suggest the inflation parameter \(\alpha = 0.9\) is adequate for the system.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/rcrv-sfc-1} \caption{\(sd RCRV\) calculated for surface observations (from PREPBUFR and ASWS) of a) wind, and b) temperature, averaged over 2.5º boxes.}\label{fig:rcrv-sfc}
+\includegraphics[width=1\linewidth]{../figures/rcrv-sfc-1} \caption{\(sd RCRV\) calculated for surface observations (from PREPBUFR and AWS) of a) wind, and b) temperature, averaged over 2.5º boxes.}\label{fig:rcrv-sfc}
 \end{figure}
@@
 The analysis is performed over a smaller domain (red box in Figure \ref{fig:dominio}a) to focus on the region most directly affected by the MCS.
-Figure \ref{fig:TQ-diff}a-c shows the difference between experiments in the spatially averaged vertical profile of temperature.
+Figure \ref{fig:TQ-diff}a-c shows the analysis difference between experiments in the spatially averaged vertical profile of temperature.
 By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
-During the first day, the assimilation of ASWS observations results in a colder PBL.
-This cooling effect has a clear diurnal cycle being stronger during nighttime (Figure \ref{fig:TQ-diff}a).
-During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
-The warm difference shown in AUT with respect to CONV in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV.
+During the first day, the assimilation of AWS observations results in a colder PBL.
+This cooling effect has a clear diurnal cycle, being stronger during nighttime (Figure \ref{fig:TQ-diff}a).
+During the second day of the experiment, the impact of AWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS.
+The warm difference shown in AWS-CONV between 500 and 200 hPa is produced by the development of stronger convection in AWS compared to CONV.
 This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection.
-Figure \ref{fig:TQ-diff}b, indicates that SATWND observations have an almost negligible impact on the mean temperature through the experiment.
+Figure \ref{fig:TQ-diff}b indicates that SATWND observations have an almost negligible impact on the mean temperature throughout the experiment.
 This result is unexpected since mid-to-upper level winds can have a profound impact on the organization and evolution of MCSs over the region.
-During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of ASWS observations.
+During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations.
 No clear systematic impact is found above the PBL during this period.
-During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AUT experiment but with the opposite sign.
-
-Comparing the specific humidity in the experiments (Figure \ref{fig:TQ-diff}d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \ref{fig:TQ-diff}d).
-The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the nighttime.
+During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AWS experiment but with the opposite sign.
+
+Comparing the specific humidity in the experiments (Figure \ref{fig:TQ-diff}d-f), the impact of assimilating AWS with finer spatial and temporal resolution is most substantial at low levels (Figure \ref{fig:TQ-diff}d).
+The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime.
 The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region (Ruiz et al., 2010).
@@
 In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
-As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels.
+As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels (Figure \ref{fig:TQ-diff}e).
 Figure \ref{fig:TQ-diff}f shows that radiances reduce low-middle level moisture during the first day of the experiment.
@@
 
-The impacts on the wind components are shown in Figure \ref{fig:UV-diff} along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations.
-The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis.
+\begin{figure*}[th]
+
+{\centering \includegraphics{../figures/TQ-diff-1}
+
+}
+
+\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:TQ-diff}
+\end{figure*}
+
+The impacts on the wind components are shown in Figure \ref{fig:UV-diff}, along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations (for example, Figure \ref{fig:UV-diff}a shows the zonal wind difference between AWS and CONV and the zonal wind for AWS).
+The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis.
 There is a diurnal cycle in the impact of surface weather stations on the meridional velocity (Figure \ref{fig:UV-diff}d) with a stronger reduction of the northerly wind during night hours.
 This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment.
-After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain.
+After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain.
 After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels.
-During Nov 22 and 23 the impact of assimilating ASWS produces an increase of northerly wind in upper levels.
+During Nov 22 and 23 the impact of assimilating AWS produces an increase of northerly wind in upper levels.
 This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
-Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.
-
-The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c and f).
-For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of ASWS observations.
-On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind.
-When we look at the meridional wind field at 200 hPa at different times, we observe that the outflow from the MCS is even more intense than in the other experiments, while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in Figure \ref{fig:UV-diff}f.~
-
-To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles).
-At that moment the first convective cells were developing on the southern region of the domain along the cold front.
+Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger; the systematic impact is overall smaller than the one produced by assimilating data from AWS (Figure \ref{fig:UV-diff}b, e).
+
+The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c).
+For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of AWS observations during the nights previous to the development of the MCS.
+At upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind speed.
+The meridional wind field at 200 hPa at different times shows that the outflow from the MCS is even more intense than in the other experiments, while the southerly wind ahead of the MCS also increases producing an average reduction of the northerly wind (Figure \ref{fig:UV-diff}f).
+
+\begin{figure*}[ht]
+
+{\centering \includegraphics{../figures/UV-diff-1}
+
+}
+
+\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.
+Black contours correspond to u wind for AWS (a), SATWND (b) and RAD (c) and v wind for AWS (d), SATWND (e) and RAD (f).}\label{fig:UV-diff}
+\end{figure*}
+
+To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles).
+At that moment the first convective cells were developing over the southern region of the domain along the cold front.
 Figure \ref{fig:summary-fields}a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
@@
 The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
-AUT and SATWND experiments are very similar, with values of precipitable water over 55 \(kgm^{-2}\) north of 30\(^{\circ}\)S and a similar vertical distribution of specific humidity (not shown).
-RAD has lower precipitable water content than AUT and SATWND but higher than CONV.
-The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
-
-The potential temperature on the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed on the temperature profiles in Figure \ref{fig:TQ-diff}a-c.~On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL than AUT due to the assimilation of radiance observations.
+AWS and SATWND experiments are very similar, with values of precipitable water over 55 \(kgm^{-2}\) north of 30\(^{\circ}\)S and a similar vertical distribution of specific humidity (not shown).
+RAD has lower precipitable water content than AWS and SATWND, but higher than CONV.
+The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating AWS -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
+
+The analyzed distribution of temperature and moisture in the PBL (Figure \ref{fig:summary-fields}b) resembles the observations (Figure \ref{fig:TQ-diff}a-c).
+On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
 Figure \ref{fig:summary-fields}c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 \(km\) wind shear.
 The values of MCAPE in CONV do not exceed 2000 \(JKg^{-1}\) while the rest of the experiments show maximum MCAPE over 4000 \(JKg^{-1}\).
-MCAPE in the RAD experiment is lower compared to AUT or SATWND.
-This is consistent with less humidity in the PBL with respect to these experiments and may be partially compensated by a slightly warmer PBL in the RAD experiment.
-The 0-6 km wind shear is more intense in AUT, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values.
+MCAPE in the RAD experiment is lower compared to AWS or SATWND.
+This is consistent with less humidity in the PBL with respect to these experiments but may be partially compensated by a slightly warmer PBL in the RAD experiment.
+The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values.
 Moreover, in this same region, these experiments show larger MCAPE values than CONV.
-Note that wind shears over 15 \(ms^{-1}\) have been associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowsky and Richardson, 2010).
-
-\begin{figure*}[th]
-
-{\centering \includegraphics{../figures/TQ-diff-1}
-
-}
-
-\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:TQ-diff}
-\end{figure*}
-
-\begin{figure*}[ht]
-
-{\centering \includegraphics{../figures/UV-diff-1}
-
-}
-
-\caption{Difference between experiments AUT-CONV (a and d), SATWND-AUT (b and e) and RAD-SATWND (c and f) for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.
-In contours the u wind for AUT (a), SATWND (b) and RAD (c) and v wind for AUT (d), SATWND (e) and RAD (f) are shown.}\label{fig:UV-diff}
-\end{figure*}
+Note that wind shears over 15 \(ms^{-1}\) have been associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowski and Richardson, 2010).
 
@@
 \includegraphics[width=1\linewidth]{../figures/summary-fields-1} \caption{a) Precipitable water (shaded, \(kgm^{-2}\)) and average northerly wind on the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, \(ms^{-1}\)), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and \textasciitilde0-6 km wind shear over 15 and 30 \(ms^{-1}\) for each experiment.
-All fields correspond to 00 UTC Nov, 22.}\label{fig:summary-fields}
+All fields correspond to the analysis ensemble mean for 00 UTC Nov, 22.
+Grey filled contours correspond to topography over 1500 meter over sea level.}\label{fig:summary-fields}
 \end{figure*}
@@
 
-Firstly we analyze the impact of different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \ref{fig:pp-hov} shows the first-guess hourly accumulated precipitation averaged over a longitude range between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments, and the estimated precipitation by IMERG.
-The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period.
+Firstly we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
+Figure \ref{fig:pp-hov}(a) shows the mean hourly accumulated precipitation as estimated by IMERG, and (b-e) the probability matching mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments.
+The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \ref{fig:pp-hov}a).
 In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated.
-This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation.
-AUT, SATWND, and RAD better capture the timing and location of convective initiation.
-On RAD the initiation is delayed when we compare with AUT and SATWND, possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue.
-After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
+This is particularly evident in CONV (Figure \ref{fig:pp-hov}b) where the convection initiation is delayed and occurs further north with respect to the observed initiation.
+AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figure \ref{fig:pp-hov}c-e).
+AWS and RAD show a more fragmented distribution compared with SATWND, possibly due to the development of less organized convection during Nov 22.
+After 18 UTC Nov 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
 
@@
 
-To quantify the spatial match between the observed precipitation and the short-range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}).
-All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature stage of the MCS.
+To quantify the spatial match between the observed precipitation and the short-range forecast precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}).
+All experiments show values of FSS below 0.6 around 03 UTC Nov 22 indicating that the experiments were not able to accurately simulate the convective environment and the initialization of the convection.
+However, we notice that the FSS values for RAD around 00 UTC Nov 22 are higher than for the other experiment, indicating that radiance observations have a positive impact on the analysis.
+Moreover, the FSS values improve during the mature stage of the MCS for all spatial scales and thresholds.
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
-AUT and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
-The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22.
+AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
+The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22.
 The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).
@@
 \includegraphics{../figures/fss-1.png}
-\caption{\label{fig:fss}FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the CONV (blue line), AUT (ligh blue line), SATWND (orange line) and RAD (red line) experiments.}
+\caption{\label{fig:fss}FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), AWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.}
 \end{figure}
 
-To complement the analysis, Figure \ref{fig:dbz-mean} shows the maximum reflectivity in the column (COLMAX) for the CONV and RAD experiments ensemble mean at different times between 10 and 19 UTC Nov, 22 along with the observed maximum reflectivity.
-These experiments were chosen because they represent the analysis with the minimum and the maximum number of assimilated observations.
-In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the short-range precipitation forecast skill (Figure \ref{fig:pp-hov}).
-Overall none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
-This is partially expected considering the relatively low horizontal resolution (10 km) which is not enough to appropriately represent the strength of the convective band associated with the MCS.
-Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecasted reflectivity field.
+To complement the analysis, Figure \ref{fig:dbz-mean} shows the observed maximum reflectivity in the vertical column (COLMAX) and the ensemble mean COLMAX for the CONV and RAD experiments at different times between 10 and 19 UTC Nov 22.
+These experiments were chosen because they represent the analysis with the minimum (CONV) and maximum (RAD) number of assimilated observations.
+In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the 1-hour precipitation forecast skill (Figure \ref{fig:fss}).
+Overall, none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
+This is partially expected considering the coarse horizontal grid spacing (10 km), which is not enough to appropriately represent the strength of the convective band associated with the MCS.
+Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecast reflectivity field.
 RAD better represents the observed features of the system showing a stronger and more organized MCS than CONV, over the domain center at 10 and 13 UTC (first and second columns in Figure \ref{fig:dbz-mean}).
-The convective cells that initiate after 16 UTC over the warm front in the northeast part of the domain are well captured by both experiments, but better represented in terms of strength in RAD.
+The convective cells that initiate after 16 UTC along the warm front in the northeast part of the domain are well captured by both experiments but are better represented in terms of strength in RAD.
 In addition, CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
-Before and after the times shown in Figure \ref{fig:dbz-mean}, the agreement between experiment and observations is quite high in the regions where radar data are available particularly for RAD.
-
-Finally, Figure \ref{fig:soundings} shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+Before and after the times shown in Figure \ref{fig:dbz-mean}, the agreement between experiment and observations is quite good in the regions where radar data are available, especially for RAD.
+
+Finally, Figure \ref{fig:soundings} shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes), and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
 
@@
 The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection.
-For the temperature, CONV presents a warm bias on the PBL of 1\(^{\circ}\)C and an RMSE between 1 and 2.5\(^{\circ}\)C that is slightly reduced by the assimilation of ASWS and radiances.
-The RMSE for temperature remains between 0.5 and 2\(^{\circ}\)C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND, and RAD is small above the PBL.
-The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND, and RAD.
-On middle levels (below 10 km) the bias is large and positive (8 - 10\(^{\circ}\)C) for all the experiments, indicating a more moist atmosphere in the analyses than the observed.
+For the temperature, CONV presents a warm bias on the PBL of 1\(^{\circ}\)C and an RMSE between 1 and 2.5\(^{\circ}\)C that is slightly reduced by the assimilation of AWS and radiances (Figure \ref{fig:soundings}a).
+The RMSE for temperature remains between 0.5 and 2\(^{\circ}\)C throughout the troposphere and lower stratosphere and the impact of additional observations in the AWS, SATWND, and RAD is small above the PBL.
+The dew point temperature shows a small dry bias in low levels that is corrected in AWS, SATWND, and RAD (Figure \ref{fig:soundings}b).
+At middle levels (below 10 km) the bias is large and positive (8 - 10\(^{\circ}\)C) for all the experiments, indicating a more moist atmosphere in the analyses than observed.
 Overall, the RMSE grows with height, probably because at upper levels moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
-The zonal wind has a significant positive bias of approximately 10 \(ms^{-1}\) in the upper troposphere which is mainly reduced by assimilating satellite radiances.
-It is somewhat surprising that no significant effect is observed by assimilating satellite-derived winds.
-The assimilation of ASWS and radiance observations increases the northerly wind bias and RMSE near the surface.
+The zonal wind for AWS and SATWND has a substantial positive bias of approximately 10 \(ms^{-1}\) in the upper troposphere which is mainly reduced to CONV values by assimilating satellite radiances (Figure \ref{fig:soundings}c).
+It is somewhat surprising that no substantial effect is observed by assimilating satellite-derived winds.
+The assimilation of AWS and radiance observations increases the northerly wind bias and RMSE near the surface (Figure \ref{fig:soundings}d).
 
 \begin{figure*}[h]
-\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecasted ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov, 22.}\label{fig:dbz-mean}
+\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22.}\label{fig:dbz-mean}
 \end{figure*}
 
-For the IOP 8 (bottom row in Figure \ref{fig:soundings}), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulations.
+For IOP 8 (Figure \ref{fig:soundings}e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation.
 This area was also behind the cold front thus affected by low-level cold advection.
-Below 5 km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND, and RAD to the assimilation of ASWS.
-The RMSE for AUT, SATWND, and RAD remains below 2\(^{\circ}\)C for almost all levels with significant improvement in upper levels where radiance observations have the largest impact.
+Below 5 km all experiments show a warm bias which decreases with the assimilation of AWS and radiances.
+Between 5 and 12 km, the bias in each experiment is cold and is reduced in AWS, SATWND, and RAD.
+The RMSE for AWS, SATWND, and RAD remains below 2\(^{\circ}\)C for almost all levels with substantial improvement in upper levels where radiance observations have the largest impact.
 In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
 This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region.
-The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
-SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7\(^{\circ}\)C.
-The zonal wind is overestimated on the analyses with a positive bias at middle levels.
+The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation, but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
+SATWND presents the lowest RMSE overall while AWS and RAD are comparable between 4 and 12 km with values that do not exceed 7\(^{\circ}\)C.
+The zonal wind is overestimated in the analyses with a positive bias at middle levels.
 For this period only RAD shows an improvement with respect to CONV in the upper troposphere.
-At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND, and RAD.
-In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
+At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD.
+In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).
 
@@
 
-\caption{RMSE (solid line) and Bias (dash line) of a) temperature (\(K\)), b) dew point temperature (\(K\)), c) u wind (\(ms^{-1}\)) and d) v wind (\(ms^{-1}\)) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
-The blue line correspond to CONV, the light blue line to AUT, SATWND is represented with an orange line and RAD with a red line.}\label{fig:soundings}
+\caption{RMSE (solid line) and Bias (dashed line) of a) temperature (\(K\)), b) dew point temperature (\(K\)), c) u wind (\(ms^{-1}\)) and d) v wind (\(ms^{-1}\)) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+The blue line corresponds to CONV, the light blue line to AWS, SATWND is represented with an orange line, and RAD with a red line.}\label{fig:soundings}
 \end{figure*}
+
+\hypertarget{ensemble-forescast-validation}{%
+\subsection{Ensemble forescast validation}\label{ensemble-forescast-validation}}
+
+In this section, we analyse the 60-members ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
+We calculated the FSS for the ensemble forecast in 6-hours moving windows for the same thresholds and spatial scales as for the analysis to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}).
+CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
+AWS, SATWND, and RAD show an important improvement in the FSS values, particularly for the higher threshold (Figure \ref{fig:fssfcst}b, d).
+Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
+
+The satellite-derived wind observations show a clearly positive impact on the forecast, in contrast to what was seen when comparing the analysis with independent observations in terms of precipitation.
+Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the analysis to IMERG estimations.
+
+\begin{figure}
+\centering
+\includegraphics{../figures/fssfcst-1.png}
+\caption{\label{fig:fssfcst}FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the forecasts initialized from CONV (blue line), AWS (light blue line), SATWND (orange line) and RAD (red line) experiments at 00 UTC (solid line) and 06 UTC (dashed line).}
+\end{figure}
 
@@
 
-We performed four experiments to assess the impact of different observation types, including conventional data (CONV), frequent surface observations (AUT), satellite-derived winds (SATWND), and clear-sky satellite radiances (RAD), into a regional frequent-update ensemble-based data assimilation system.
-We used a case study approach of a massive MCS developed over Southern South America on Nov 22, 2018.
-
-We evaluated the reliability of the ensemble using the RCRV calculated for each type of observation used in RAD.
+In this paper, we investigate for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional data assimilation system.
+In particular, we explore the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors.
+Southern South America is a particularly interesting region due to the heterogeneity and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
+This combined with a climatology characterized by frequent organized convective events makes mesoscale data assimilation particularly challenging.
+To consider these issues, we used a case study approach of a massive mesoscale convective system that developed over Southern South America, particularly in a region where the surface and upper-air observing networks are too coarse to capture variability at the mesoscale and that took place on Nov 22, 2018.
+
+We evaluated the consistency of the ensemble using the RCRV calculated for each type of observation used in RAD.
 Conventional observations including surface data and upper air observations from soundings and aircraft, show a \(sd RCRV\) near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
 For satellite-derived winds and radiance observations, the \(sd RCRV\) is lower than 1, which could be the result of an underdispersive ensemble or an overestimation of the observation errors.
-However, taking into account that the errors of this type of observation are usually overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the assembly has a reasonable spread.
-
-In terms of the analysis, we found that ASWS observations, which have high spatial and temporal resolution, produce major impacts throughout the troposphere and mainly on the PBL.
-In particular, the precipitable water content and the low level meridional circulation lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
-We also found positive results when assimilating radiance observations that produce a better development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
+However, taking into account that the observation error for satellite-derived wind is overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the ensemble has a reasonable spread.
+
+In terms of the analysis, we found that AWS observations, which have high spatial and temporal resolution, produce major impacts throughout the troposphere and mainly on the PBL.
+In particular, the precipitable water content and the low level meridional circulation in the AWS experiment lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
+We also found positive results when assimilating radiance observations that produce an adequate development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
 However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates.
-While the assimilation of satellite-derived wind does not impact substantially on the analysis, this is possible due to the small number of observations in low levels available for this case study.
+While the assimilation of satellite-derived wind does not produce a noticeable impact on the analysis with respect to the AWS experiment, this is possibly due to the small number of observations in low levels available for this case study and its large observation error.
+However, there are substantial improvements for the accumulated precipitation distribution.
 A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels.
 
-Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during the IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement on low levels when assimilating ASWS observations on the temperature, dew point temperature, and meridional wind.
-In particular, ASWS observations correct a dry bias already observed by Ruiz et al.
+Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during IOPs 7 and 8.
+For the pre-convective environment (IOP 7), we observed an improvement at low level when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
+In particular, AWS observations correct a dry bias already observed by Ruiz et al.
 (2010).
 For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels including more observations for the temperature but the opposite effect for the dew point temperature.
-In this case, the analyses have a moist bias that increases with the assimilation of ASWS and radiances observations.
-We observed a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
+In this case, the analyses have a moist bias that increases with the assimilation of AWS and radiance observations.
+We noted a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
 It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period, which requires further analysis.
 
-To summarize this study, we conclude that the assimilation of surface observations with higher spatial and temporal resolution and clear-sky radiances from polar orbit satellites had an overly positive impact on the development of the studied MCS and its associated precipitation.
-In the future, we will analyze whether these impacts carry over in time to the independent forecasts initialized from the analyses.
+We also analyzed the performance of independent ensemble forecasts intialized from the analyses to estimate the precipitation during Nov 22.
+The forecast initialized from AWS, SATWND, and RAD are able to estimate the precipitation substantially better than CONV.
+In particular, the continuous assimilation of atellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
+
+To summarize, in this study we conclude that the assimilation of surface observations with higher spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
+Moreover, the ensemble forecasts initialized from the analysis show promising results to estimate the precipitation of severe events.
+In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like GPS radio occultation data and radiances from geostationary orbiting satellites.
 
@@
 We also acknowledge the Cheyenne HPC resources (doi:10.5065/D6RX99HX) from NCAR's Computational and Information Systems Laboratory, National Science Foundation (project code UIUC0012).
-Also, PICT 2017-2033, PICT 2016-0710 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
+Also, PICT 2017-2033 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
 
@@
 Allaire, J., Horner, J., Xie, Y., Marti, V., Porte, N., 2019. Markdown: Render markdown with the c library 'sundown'.
+
+\leavevmode\hypertarget{ref-bae2022}{}%
+Bae, J.-H., Min, K.-H., 2022.
+Forecast Characteristics of Radar Data Assimilation Based on the Scales of Precipitation Systems.
+Remote Sensing 14, 605.
+doi:\href{https://doi.org/10.3390/rs14030605}{10.3390/rs14030605}
 
@@
 The spatial distribution of severe thunderstorm and tornado environments from global reanalysis data.
-Atmospheric Research 67-68, 73--94.
+Atmospheric Research 67--68, 73--94.
 doi:\href{https://doi.org/10.1016/S0169-8095(03)00045-0}{10.1016/S0169-8095(03)00045-0}
@@
 
+\leavevmode\hypertarget{ref-chang2017}{}%
+Chang, W., Jacques, D., Fillion, L., Baek, S.-J., 2017.
+Assimilation of Hourly Surface Observations with the Canadian High-Resolution Ensemble Kalman Filter.
+Atmosphere-Ocean 55, 247--263.
+doi:\href{https://doi.org/10.1080/07055900.2017.1384361}{10.1080/07055900.2017.1384361}
+
 \leavevmode\hypertarget{ref-chen2001}{}%
@@
 
+\leavevmode\hypertarget{ref-chen2016}{}%
+Chen, X., Zhao, K., Sun, J., Zhou, B., Lee, W.-C., 2016.
+Assimilating surface observations in a four-dimensional variational Doppler radar data assimilation system to improve the analysis and forecast of a squall line case.
+Adv.
+Atmos.
+Sci.
+33, 1106--1119.
+doi:\href{https://doi.org/10.1007/s00376-016-5290-0}{10.1007/s00376-016-5290-0}
+
+\leavevmode\hypertarget{ref-cherubini2006}{}%
+Cherubini, T., Businger, S., Velden, C., Ogasawara, R., 2006.
+The Impact of Satellite-Derived Atmospheric Motion Vectors on Mesoscale Forecasts over Hawaii.
+Monthly Weather Review 134, 2009--2020.
+doi:\href{https://doi.org/10.1175/MWR3163.1}{10.1175/MWR3163.1}
+
+\leavevmode\hypertarget{ref-clark2017}{}%
+Clark, A.J., 2017.
+Generation of Ensemble Mean Precipitation Forecasts from Convection-Allowing Ensembles.
+Weather and Forecasting 32, 1569--1583.
+doi:\href{https://doi.org/10.1175/WAF-D-16-0199.1}{10.1175/WAF-D-16-0199.1}
+
 \leavevmode\hypertarget{ref-Cheyenne2019}{}%
@@
 
-\leavevmode\hypertarget{ref-dillon2019}{}%
-Dillon, M.E., García Skabar, Y., Kalnay, E., Ruiz, J.J., Collini, E.A., 2019.
-Sensibilidad de un sistema de asimilación de datos por ensambles a diferentes configuraciones, implementado en el sur de Sudamérica.
-Meteorológica 44, 15--34.
-
-\leavevmode\hypertarget{ref-dillonInreview}{}%
-Dillon, M.E., Maldonado, P., Corrales, P., García Skabar, Y., Ruiz, J.J., Sacco, M., Cutraro, F., Mingari, L., Matsudo, C., Vidal, L., Rugna, M., Hobouchian, M.P., Salio, P., Nesbitt, S., Saulo, C., Kalnay, E., Miyoshi, T., n.d.
-A Rapid Refresh ensemble based Data Assimilation and Forecast system for the RELAMPAGO field campaign.
-Atmospheric Research.
+\leavevmode\hypertarget{ref-dillon2021}{}%
+Dillon, M.E., Maldonado, P., Corrales, P., Skabar, Y.G., Ruiz, J., Sacco, M., Cutraro, F., Mingari, L., Matsudo, C., Vidal, L., Rugna, M., Hobouchian, M.P., Salio, P., Nesbitt, S., Saulo, C., Kalnay, E., Miyoshi, T., 2021.
+A rapid refresh ensemble based data assimilation and forecast system for the RELAMPAGO field campaign.
+Atmospheric Research 105858.
+doi:\href{https://doi.org/10.1016/j.atmosres.2021.105858}{10.1016/j.atmosres.2021.105858}
 
@@
 
-\leavevmode\hypertarget{ref-gao2015}{}%
-Gao, F., Huang, X.-Y., Jacobs, N.A., Wang, H., 2015.
-Assimilation of wind speed and direction observations: Results from real observation experiments.
-Tellus A: Dynamic Meteorology and Oceanography 67, 27132.
-doi:\href{https://doi.org/10.3402/tellusa.v67.27132}{10.3402/tellusa.v67.27132}
-
 \leavevmode\hypertarget{ref-garcia2019}{}%
 Garcia, F., Ruiz, J., Salio, P., Bechis, H., Nesbitt, S., 2019. Argentina mesonet data. Version 1.1. UCAR/NCAR - earth observing laboratory.
+
+\leavevmode\hypertarget{ref-gasperoni2018}{}%
+Gasperoni, N.A., Wang, X., Brewster, K.A., Carr, F.H., 2018.
+Assessing Impacts of the High-Frequency Assimilation of Surface Observations for the Forecast of Convection Initiation on 3 April 2014 within the Dallas--Fort Worth Test Bed.
+Monthly Weather Review 146, 3845--3872.
+doi:\href{https://doi.org/10.1175/MWR-D-18-0177.1}{10.1175/MWR-D-18-0177.1}
 
@@
 
+\leavevmode\hypertarget{ref-maejima2019}{}%
+Maejima, Y., Miyoshi, T., Kunii, M., Seko, H., Sato, K., 2019.
+Impact of Dense and Frequent Surface Observations on 1-Minute-Update Severe Rainstorm Prediction: A Simulation Study.
+Journal of the Meteorological Society of Japan.
+Ser.
+II 97, 253--273.
+doi:\href{https://doi.org/10.2151/jmsj.2019-014}{10.2151/jmsj.2019-014}
+
 \leavevmode\hypertarget{ref-maldonado2021}{}%
@@
 
-\leavevmode\hypertarget{ref-markowsky2010}{}%
-Markowsky, P., Richardson, Y., 2010.
+\leavevmode\hypertarget{ref-mallick2020}{}%
+Mallick, S., Jones, T.A., 2020.
+Assimilation of GOES-16 satellite derived winds into the warn-on-forecast system.
+Atmospheric Research 245, 105131.
+doi:\href{https://doi.org/10.1016/j.atmosres.2020.105131}{10.1016/j.atmosres.2020.105131}
+
+\leavevmode\hypertarget{ref-markowski2010}{}%
+Markowski, P., Richardson, Y., 2010.
 Organization of Isolated Convection, in: Mesoscale Meteorology in Midlatitudes.
@@
 Matsudo, C., García Skabar, Y., Ruiz, J., Vidal, L., Salio, P., 2015. Verification of WRF-ARW convective-resolving forecasts over Southeastern South America. Mausam 66, 445--456.
-
-\leavevmode\hypertarget{ref-miyoshi2012a}{}%
-Miyoshi, T., Kunii, M., 2012.
-Using AIRS retrievals in the WRF-LETKF system to improve regional numerical weather prediction.
-Tellus A: Dynamic Meteorology and Oceanography 64, 18408.
-doi:\href{https://doi.org/10.3402/tellusa.v64i0.18408}{10.3402/tellusa.v64i0.18408}
 
@@
 doi:\href{https://doi.org/10.1175/BAMS-D-20-0029.1}{10.1175/BAMS-D-20-0029.1}
+
+\leavevmode\hypertarget{ref-otsuka2015}{}%
+Otsuka, M., Kunii, M., Seko, H., Shimoji, K., Hayashi, M., Yamashita, K., 2015.
+Assimilation Experiments of MTSAT Rapid Scan Atmospheric Motion Vectors on a Heavy Rainfall Event.
+Journal of the Meteorological Society of Japan.
+Ser.
+II 93, 459--475.
+doi:\href{https://doi.org/10.2151/jmsj.2015-030}{10.2151/jmsj.2015-030}
 
@@
 
+\leavevmode\hypertarget{ref-sawada2019}{}%
+Sawada, M., Ma, Z., Mehra, A., Tallapragada, V., Oyama, R., Shimoji, K., 2019.
+Impacts of Assimilating High-Resolution Atmospheric Motion Vectors Derived from Himawari-8 on Tropical Cyclone Forecast in HWRF.
+Monthly Weather Review 147, 3721--3740.
+doi:\href{https://doi.org/10.1175/MWR-D-18-0261.1}{10.1175/MWR-D-18-0261.1}
+
+\leavevmode\hypertarget{ref-shao2016}{}%
+Shao, H., Derber, J., Huang, X.-Y., Hu, M., Newman, K., Stark, D., Lueken, M., Zhou, C., Nance, L., Kuo, Y.-H., Brown, B., 2016.
+Bridging Research to Operations Transitions: Status and Plans of Community GSI.
+Bulletin of the American Meteorological Society 97, 1427--1440.
+doi:\href{https://doi.org/10.1175/BAMS-D-13-00245.1}{10.1175/BAMS-D-13-00245.1}
+
 \leavevmode\hypertarget{ref-singh2016}{}%
@@
 
-\leavevmode\hypertarget{ref-skamarock2018}{}%
-Skamarock, W.C., Klemp, J.B., Dudhia, J., Gill, D.O., Barker, D.M., Duda, M.G., Huang, X.-Y., Wang, W., Powers, J.G., 2018. A Description of the Advanced Research WRF Version 3.
+\leavevmode\hypertarget{ref-skamarock2008}{}%
+Skamarock, W.C., Klemp, J.B., Dudhia, J., Gill, D.O., Barker, D.M., Duda, M.G., Huang, X.-Y., Wang, W., Powers, J.G., 2008. A Description of the Advanced Research WRF Version 3.
+
+\leavevmode\hypertarget{ref-sobash2015}{}%
+Sobash, R.A., Stensrud, D.J., 2015.
+Assimilating Surface Mesonet Observations with the EnKF to Improve Ensemble Forecasts of Convection Initiation on 29 May 2012.
+Monthly Weather Review 143, 3700--3725.
+doi:\href{https://doi.org/10.1175/MWR-D-14-00126.1}{10.1175/MWR-D-14-00126.1}
 
@@
 
-\leavevmode\hypertarget{ref-vera2006}{}%
-Vera, C., Baez, J., Douglas, M., Emmanuel, C.B., Marengo, J., Meitin, J., Nicolini, M., Nogues-Paegle, J., Paegle, J., Penalba, O., Salio, P., Saulo, C., Silva Dias, M.A., Dias, P.S., Zipser, E., 2006.
-The South American Low-Level Jet Experiment.
-Bull.
-Amer.
-Meteor.
-Soc.
-87, 63--78.
-doi:\href{https://doi.org/10.1175/BAMS-87-1-63}{10.1175/BAMS-87-1-63}
+\leavevmode\hypertarget{ref-wang2021}{}%
+Wang, Z.Q., Randriamampianina, R., 2021.
+The Impact of Assimilating Satellite Radiance Observations in the Copernicus European Regional Reanalysis (CERRA).
+Remote Sensing 13, 426.
+doi:\href{https://doi.org/10.3390/rs13030426}{10.3390/rs13030426}
 
@@
 doi:\href{https://doi.org/10.21957/ewahn9ce}{10.21957/ewahn9ce}
+
+\leavevmode\hypertarget{ref-wheatley2010}{}%
+Wheatley, D.M., Stensrud, D.J., 2010.
+The Impact of Assimilating Surface Pressure Observations on Severe Weather Events in a WRF Mesoscale Ensemble System.
+Monthly Weather Review 138, 1673--1694.
+doi:\href{https://doi.org/10.1175/2009MWR3042.1}{10.1175/2009MWR3042.1}
 
@@
 
+\leavevmode\hypertarget{ref-wu2014}{}%
+Wu, T.-C., Liu, H., Majumdar, S.J., Velden, C.S., Anderson, J.L., 2014.
+Influence of Assimilating Satellite-Derived Atmospheric Motion Vector Observations on Numerical Analyses and Forecasts of Tropical Cyclone Track and Intensity.
+Monthly Weather Review 142, 49--71.
+doi:\href{https://doi.org/10.1175/MWR-D-13-00023.1}{10.1175/MWR-D-13-00023.1}
+
 \leavevmode\hypertarget{ref-xie2015}{}%
 Xie, Y., 2015. Dynamic documents with R and knitr, Second. ed. Chapman and Hall/CRC, Boca Raton, Florida.
+
+\leavevmode\hypertarget{ref-zhao2021}{}%
+Zhao, J., Gao, J., Jones, T.A., Hu, J., 2021a.
+Impact of Assimilating High-Resolution Atmospheric Motion Vectors on Convective Scale Short-Term Forecasts: 1.
+Observing System Simulation Experiment (OSSE).
+Journal of Advances in Modeling Earth Systems 13, e2021MS002484.
+doi:\href{https://doi.org/10.1029/2021MS002484}{10.1029/2021MS002484}
+
+\leavevmode\hypertarget{ref-zhao2021a}{}%
+Zhao, J., Gao, J., Jones, T.A., Hu, J., 2021b.
+Impact of Assimilating High-Resolution Atmospheric Motion Vectors on Convective Scale Short-Term Forecasts: 2.
+Assimilation Experiments of GOES-16 Satellite Derived Winds.
+Journal of Advances in Modeling Earth Systems 13, e2021MS002486.
+doi:\href{https://doi.org/10.1029/2021MS002486}{10.1029/2021MS002486}
 

### 2022-04-11 analysis/scripts/RCRV_rad.R

@@ -6,7 +6,7 @@ library(metR)
 
 sensor <- "iasi"
 
-path <- paste0("/home/paola.corrales/datosmunin/EXP/E8/ANA/*/diagfiles/asim_", sensor, "*ensmean*")
+path <- paste0("/home/paola.corrales/datosmunin3/EXP/E9/ANA/*/diagfiles/asim_", sensor, "*ensmean*")
 
 files <- Sys.glob(path)
 
@@ -28,12 +28,12 @@ satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>%
 
 perfiles <- purrr::map(dates, function(d) {
 
-  files_date <- Sys.glob(paste0("/home/paola.corrales/datosmunin/EXP/E8/ANA/", d, "/diagfiles/asim_iasi*_", d, ".mem", formatC(1: 60, width = 3, flag = "0")))
+  files_date <- Sys.glob(paste0("/home/paola.corrales/datosmunin3/EXP/E9/ANA/", d, "/diagfiles/asim_iasi*_", d, ".mem", formatC(1: 60, width = 3, flag = "0")))
 
   files_date <- files_date[str_detect(files_date, "metop-a")]
   print(files_date[1])
 
-  out <- read_diag_rad(files_date, exp = "E8") %>%
+  out <- read_diag_rad(files_date, "E9") %>%
     satinfo[., on = c("sensor", "channel")] %>%
     .[iuse == 1] %>%
     .[, ":="(level = cut_round(peakwt, breaks = c(seq(0, 500, 100), seq(550, 1050, 50))))] %>%

### 2022-04-11 analysis/scripts/acumulados_imerg.R

@@ -18,27 +18,24 @@ wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
 ini_date <- ymd_hms("20181120000000")
 ciclos <- 85
 
-acumulado <- 6
+acumulado <- 1
 
 # Inicio
 dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
                     length.out = ciclos - acumulado)
 
 # Modelo
-pp_wrf <- purrr::map("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181122060000/analysis.ensmean", function(f) {
-  ReadNetCDF(f, vars = c("RAINNC", "RAINC", "RAINSH",
+pp_wrf <- ReadNetCDF("/home/paola.corrales/datosmunin3/EXP/E9/FCST/2018112200_det/wrfout_d02_2018-11-22_00:00:00",
+             vars = c("RAINNC", "RAINC", "RAINSH",
                          lon = "XLONG", lat = "XLAT")) %>%
-    .[, ":="(pp_acum = RAINNC + RAINC + RAINSH,
-             exp = exp,
-             date = date)] %>%
+    .[, ":="(pp_acum = RAINNC + RAINC + RAINSH)] %>%
     .[, ":="(RAINNC = NULL,
              RAINC = NULL,
              RAINSH = NULL,
-             Time = NULL)]
-}) %>%
-  rbindlist() %>%
-  .[, .(pp_acum = sum(pp_acum, na.rm = TRUE)), by = .(lon, lat, date, exp)] %>%
-  .[, c("x", "y") := wrf_project(lon, lat)]
+             Time = NULL)] %>%
+  .[, .(pp_acum = sum(pp_acum, na.rm = TRUE)), by = .(lon, lat)] %>%
+  .[, c("x", "y") := wrf_project(lon, lat, round = c(-3, -3))]
+
 
 
 interpolate <- function(x, y, pp) {
@@ -73,7 +70,7 @@ pp_acum_imerg <- furrr::future_map_dfr(dates, function(d) {
     rbindlist() %>%
     setnames(c("Grid/lon", "Grid/lat"), c("lon", "lat")) %>%
     .[, .(pp_acum = sum(pp)), by = .(lon, lat)] %>%
-    .[, c("x", "y") := wrf_project(lon, lat)]
+    .[, c("x", "y") := wrf_project(lon, lat, round = c(-3, -3))]
 
 
[…]

### 2022-04-11 analysis/scripts/acumulados_wrf.R

@@ -78,7 +78,7 @@ saveRDS(pp_wrf_mean, paste0(exp, "_fcst_", format(ini_date, "%Y%m%d%H"), "_mean_
 # Acumulados horarios -----------------------------------------------------
 
 
-acumulado <- 1
+acumulado <- 6
 q <- c(1, 5, 10, 25, 50)
 
 # Inicio
@@ -218,14 +218,14 @@ saveRDS(pp_wrf$pp_wrf_mean, paste0(exp, "_fcst_", format(ini_date, "%Y%m%d%H"), 
 # Acumulados horarios ANA ------------------------------------------------
 
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
-wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-exp <- "E8"
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
+exp <- "E9"
 run <- "ana"
 
 ini_date <- ymd_hms("20181120180000")
 ciclos <- 67
 
-acumulado <- 6
+acumulado <- 1
 q <- c(1, 5, 10, 25, 50)
 
 
@@ -280,8 +280,8 @@ pp_wrf_out <- purrr::map_dbl(dates, function(d) {
     rbindlist()
 
 
-  saveRDS(pp_prop, paste0(exp, "_ana_", format(date, "%Y%m%d%H"), "_prop_acum_", acumulado, "h.rds"))
-  saveRDS(pp_wrf_mean, paste0(exp, "_ana_", format(date, "%Y%m%d%H"), "_mean_acum_", acumulado, "h.rds"))
+  saveRDS(pp_prop, paste0(wrf_path, "/derived_data/ppacum/", exp, "_ana_", format(date, "%Y%m%d%H"), "_prop_acum_", acumulado, "h.rds"))
+  saveRDS(pp_wrf_mean, paste0(wrf_path, "/derived_data/ppacum/", exp, "_ana_", format(date, "%Y%m%d%H"), "_mean_acum_", acumulado, "h.rds"))
 
 d
 })

### 2022-04-11 analysis/scripts/campos_ana.R

@@ -19,7 +19,7 @@ td_colors <- c("#341A02", "#56300B", "#7D4813", "#A46522", "#C6893E", "#DBB773",
                "#DCF2D9", "#BAE4B8", "#8BD290", "#59BC71", "#259F57",
                "#A9CFE4", "#5FA9D2", "#227CB9", "#00488F", "#0000AC")
 
-wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
 ini_date <- ymd_hms("20181120180000")
 ciclos <- 67
 
@@ -33,9 +33,9 @@ for (d in seq_along(dates)) {
   print(dates[d])
 
 
-  files_ana <- Sys.glob(paste0(wrf_path, "E[2,5,6,8]/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.ensmean"))
+  files_ana <- Sys.glob(paste0(wrf_path, "E[9]/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.ensmean"))
 
-  files_gue <- Sys.glob(paste0(wrf_path, "E[2,5,6,8]/GUESS/", format(dates[d], "%Y%m%d%H%M%S"), "/wrfarw.ensmean"))
+  files_gue <- Sys.glob(paste0(wrf_path, "E[9]/GUESS/", format(dates[d], "%Y%m%d%H%M%S"), "/wrfarw.ensmean"))
 
   ana <- purrr::map(files_ana, function(f) {
 

### 2022-04-11 analysis/scripts/count_obs.R

@@ -40,19 +40,18 @@ multiespectrales <- c("airs_aqua",
 
 asimilable <- satinfo[iuse == 1, .N, by = sensor]
 
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/diagfiles/asim*ensmean")
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E9/ANA/*/diagfiles/asim*ensmean")
 
 files <- files[str_detect(files, "conv", negate = TRUE)]
 
-diag <- read_diag_rad(files, "E8")
+diag <- read_diag_rad(files, "E9")
 
 diag[, tipo := fifelse(sensor %in% multiespectrales, "multiespectral", "no-multiespectral")] %>%
   satinfo[., on = c("sensor", "channel")] %>%
   .[errinv %between% c(0.0000316227, 1000) &
   qc == 0 &
-  iuse == 1 &
-  peakwt %between% c(0.001, 1200)] %>%
-  .[, .(prop = round(.N*100/456774, 2)), by = sensor] %>%
+  iuse > 0] %>%
+  .[, .(prop = round(.N*100/146511, 2)), by = sensor] %>%
   asimilable[., on = c("sensor")] %>%
  separate(sensor, into = c("sensor", "plataforma"), sep = "_") %>%
   write_csv("analysis/data/derived_data/tabla_radianzas.csv")

### 2022-04-11 analysis/scripts/download_goes_aws.R

@@ -5,19 +5,19 @@ library(lubridate)
 library(data.table)
 
 year <- 2018
-day <- 324
+day <- 312
 hour <- 00
 keystart <- paste0("ABI-L1b-RadF/", year, "/", day, "/", hour, "/*")
 
 goes16 <- aws.s3::get_bucket_df(bucket = 'noaa-goes16',
                                 marker = keystart,
-                                max = 7000)
+                                max = 20000)
 
 goes16 <- goes16 %>%
   setDT() %>%
   .[, start := lubridate::parse_date_time(str_extract(str_extract(Key, "s\\d{14}"), "\\d{13}"), "%Y%j%H%%M%S") ] %>%
   .[, channel := as.numeric(str_extract(str_extract(Key, "C\\d{2}"), "\\d{2}"))] %>%
-  .[minute(start) == 0 & channel %in% c(7:16) & day(start) < 24] %>%
+  .[minute(start) == 0 & channel %in% c(7:16) & day(start) < 20] %>%
   .[]
 
 

### 2022-04-11 analysis/scripts/wrap_fss.R

@@ -10,8 +10,9 @@ library(mesoda)
 #future::plan("cluster", workers = 3)
 imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
 wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
-exp <- "E8"
+exp <- "E9"
 run <- "fcst_ens"
+domain <- "d02"  #only for fcst_det
 
 ini_date <- ymd_hms("20181122060000")
 ciclos <- 31
@@ -25,8 +26,8 @@ dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
                     length.out = ciclos - acumulado)
 
 
+# pp_imerg_all <- readRDS(paste0(imerg_path, "IMERG_", acumulado, "h_d02.rds"))
 pp_imerg_all <- readRDS(paste0(imerg_path, "IMERG_", acumulado, "h.rds"))
-
 
 fss_out <- purrr::map_dfr(dates, function(d) {
 
@@ -80,6 +81,30 @@ fss_out <- purrr::map_dfr(dates, function(d) {
       # .[, exp := exp] %>%
       .[date == d]
 
+  } else if (run == "fcst_det") {
+
+    files_wrf <- purrr::map(seq(acumulado, 0), function(l) {
+
+      paste0(wrf_path, exp, "/FCST/", format(ini_date, "%Y%m%d%H"), "_det/wrfout_", domain, "_",
+                        format(date - hours(l), "%Y-%m-%d_%H:%M:%S"))
+
+    }) %>% unlist()
+
+    pp_wrf <- purrr::map(files_wrf, function(f) {
+      ReadNetCDF(f, vars = c("PREC_ACC_C", "PREC_ACC_NC",
+                             lon = "XLONG", lat = "XLAT")) %>%
+        .[, ":="(pp_acum = PREC_ACC_C + PREC_ACC_NC,
+                 exp = exp,
+                 date = date)] %>%
+        .[, ":="(PREC_ACC_C = NULL,
+                 PREC_ACC_NC = NULL,
+                 Time = NULL)]
+    }) %>%
+      rbindlist() %>%
+      .[, .(pp_acum = sum(pp_acum, na.rm = TRUE)), by = .(lon, lat, date, exp)] %>%
+      .[, c("x", "y") := wrf_project(lon, lat, round = c(-3, -3))]
+
+
   } else {
 
[…]

### 2022-04-11 analysis/chapters/paper_ana_20181122.Rmd
fix title

@@
 ---
-title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis and First-guess"
+title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO"
 author:

### 2022-05-03 R/probability_matching.R
PM for dbz

@@ -11,7 +11,7 @@
 #' @export
 
 
-calculate_probability_marching <- function(exp, run, ini_date, ciclos, acumulado = 1,
+calculate_probability_marching_pp <- function(exp, run, ini_date, ciclos, acumulado = 1,
                                            wrf_path = "/home/paola.corrales/datosmunin3/EXP/",
                                            out_path = "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/") {
 
@@ -128,7 +128,146 @@ calculate_probability_marching <- function(exp, run, ini_date, ciclos, acumulado
   }
 }
 
+#' @rdname calculate_probability_marching_pp
+#'
+#'
+#' @export
+#'
+exp <- "E2"
+run <- "ana"
+dates <- lubridate::ymd_h("2018112210", "2018112213", "2018112216", "2018112219")
+calculate_probability_marching_dbz <- function(exp, run, dates,
+                                              wrf_path = "/home/paola.corrales/datosmunin3/EXP/",
+                                              out_path = "/home/paola.corrales/datosmunin3/EXP/derived_data/dbz/") {
 
+  library(reticulate)
+  use_condaenv("/home/jruiz/share/anaconda3/envs/wrfpy")
+  py_run_string("import os")
+  py_run_string("os.environ['PROJ_LIB'] = '/home/jruiz/share/anaconda3/share/proj'")
+  wrf <- import("wrf")
+  ncdf <- import("netCDF4")
+
+
+  # ini_date <- lubridate::ymd_hms(ini_date)
+
+  members <- 60
+  # dates <- seq.POSIXt(ini_date, by = "hour",
+  #                     length.out = ciclos)
+
+
+  if (run == "ana") {
+    pp_wrf_out <- purrr::map_dbl(dates, function(d) {
+
+      date <- d
+
+      print(d)
+
+      files_wrf <- list.files(path = paste0(wrf_path, exp, "/", toupper(run), "/", format(date, "%Y%m%d%H%M%S")),
+                              full.names = TRUE,
+                              recursive = TRUE,
+                              pattern = "analysis.mem*")
+
+      dbz_wrf <- purrr::map(files_wrf, function(f) {
+
+        metadatos <- unglue::unglue(basename(f), "analysis.mem0{mem}")
+
[…]

### 2022-05-03 temp.R

@@ -1,114 +1,88 @@
-library(tidyverse)
-library(lubridate)
-library(data.table)
-library(unglue)
+# Calcula y guarda acumulados del pronóstico para todo el periodo
+
 library(metR)
+library(tidyverse)
+library(data.table)
+library(lubridate)
+library(unglue)
 library(mesoda)
-library(patchwork)
-source(here::here("temp_R/postprocesamiento.R"))
-
-map_arg <- rnaturalearth::ne_states(country = c("argentina"),
-                                    returnclass = "sf")
-map_limitrofes <- rnaturalearth::ne_countries(country = c("Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")
-
-geom_mapa <- function(fill = NA) {
-  list(geom_sf(data = map_arg, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
-       geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
-       coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)),
-       scale_x_longitude(ticks = 5),
-       scale_y_latitude(ticks = 5))
-}
-
-wrf_path <- "/home/paola.corrales/datosmunin/EXP/"
-ini_date <- ymd_hms("20181120180000")
-ciclos <- 67
-
-dates <- seq(ini_date, by = "hour",
-             length.out = ciclos)
 
 
+imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
+exp <- "E2"
+run <- "fcst"
 
-for (d in seq_along(dates)) {
+ini_date <- ymd_hms("20181122000000")
+ciclos <- 37
+acumulado <- 1
 
-  print(dates[d])
-
-
-  files_ana <- Sys.glob(paste0(wrf_path, "E[2,5,6,8]/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.ensmean"))
-
-  files_gue <- Sys.glob(paste0(wrf_path, "E[2,5,6,8]/GUESS/", format(dates[d], "%Y%m%d%H%M%S"), "/wrfarw.ensmean"))
-
-  ana <- purrr::map(files_ana, function(f) {
-
[…]

### 2022-05-03 analysis/chapters/paper_ana_20181122.Rmd
Revised version

@@
 ---
-title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO"
+title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign"
 author:
@@
   affiliation: NCAR
+- name: Stephen W. Nesbitt
+  affiliation: Illinois
 date: "`r Sys.Date()`"
@@
 abstract: |
-  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north-eastern Argentina where the surface and upper air observing networks are relatively coarse.
-We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed during November 22, 2018.
+  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse.
+We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018.
 The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
-The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
-
-  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members multiphisics  ensemble.
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files, AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
+The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign.
+
+  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND, and satellite radiances from different microwave and infrared sensors.
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-We also run ensemble forecasts initialized from each experiment show and evaluated their skill to predict precipitation.
-We found that the assimilated observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
+We also run ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
+We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
 journal: Atmospheric Research
@@
   address: National Center for Atmospheric Research, Boulder, Colorado.
+- code: Illinois
+  address: Department of Atmospheric Sciences, University of Illinois at Urbana–Champaign, Urbana, Illinois.
 bibliography: 
@@
 A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
-Southern South America has been shown to have one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer.
+Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer.
 This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014].
-Recently, the RELAMPAGO-CACTI field campaign @nesbitt2021 has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
-
-Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability, and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes).
-Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for the initialization of high-resolution forecasts [@sun2014] and thus has received increasing attention in the last decades.
+Recently, the RELAMPAGO-CACTI field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
+
+Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes).
+Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for high-resolution numerical forecasts [@sun2014] and thus has received increasing attention in the last decades.
 
@@
 
-In that regard, @wheatley2010 investigated the impact of assimilating surface pressure data in a mesoscale ensemble-based data assimilation system but found limited impact in two case studies related to mesoscale convective systems.
-@ha2014 showed that assimilating temperature and dew point temperature from high resolution surface weather station networks systematically improved the structure of the simulated planetary boundary layer and improved the short range precipitation forecast over the U.S.
-@chang2017, @bae2022 and @chen2016  reported beneficial effects of assimilating surface weather stations in a high resolution DA system using an ensemble Kalman filter, a 3-dimensional variational system and a 4-dimensional variational system respectively.
-Positive impacts have been found on the temperature and moisture forecast in the planetary boundary layer and in the location of precipitation systems.
-@sobash2015 showed in a mesoscale ensemble based DA system that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
-@maejima2019 investigated the impact of 1-minute frequency assimilation of surface observations using observation simulation experiments on a heavy rainfall case founding that assimilating high frequency and spatially dense observations lead to a better representation of mesoscale circulation even though the number of observations provided by surface stations is much less than the one provided by weather radars.
-@gasperoni2018 conducted a case study to evaluate the impact of assimilating observations produced by private weather stations which are not incorporated in the operational analysis.
-They found a positive effect of these observations upon the initiation of deep moist convection along a dryline.
+In that regard, several authors investigated the impact of assimilating surface weather data (e.g.
+@wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016 using different assimilation methodologies.
+Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.
+@sobash2015 showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
+More recently, @gasperoni2018 evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis.
+They found a positive effect of these observations upon the initiation of deep moist convection.
 This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.
 
-The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs) has been investigated in the context of limited-area mesoscale data assimilation.
-Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.
-@wu2014, @cherubini2006, and @sawada2019, among many others).
+The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs), has also been investigated in the context of limited-area mesoscale DA.
+Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.,  @wu2014, @cherubini2006, and @sawada2019, among many others).
 Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm.
 However, some works indicated mixed impacts (e.g.
-@sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecasted intensity).
+@sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity).
 As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
@@
 
-While the assimilation of radiance observations into global models is well established (), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low top models used for this application.
+While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low model tops used for this application.
 @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
@@
 They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere.
-They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on the short-range forecasts (12 and 24 h).
+They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts.
 Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land.
-Moreover, to the best of the author's knowledge, there are few to none studies related to the direct assimilation of radiance observations over South America.
+Moreover, to the best of our knowledge, there are none studies related to the direct assimilation of radiance observations over South America.
 
@@
 However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above.
-The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated, and ensemble-based DA system.
-In particular, in this paper we will focus on the impact in the context of mid-latitude mesoscale convective system events.
+The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based DA system.
+In particular, we will focus on the impact in the context of mid-latitude mesoscale convective system events.
 Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.
@@
 The paper is organized as follows.
-The data assimilation system, the experimental design, and the observations used are presented in section 2.
+The DA system, the experimental design, and the observations used are presented in section 2.
 Results are discussed in section 3 and finally, conclusions are summarized in section 4.
@@
 
-On Nov 22, a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS. 
+On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
+To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
-
-## Data assimilation system configuration
-
-The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), @skamarock2008) model.
-The horizontal grid spacing is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the model at 50 hPa. 
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). 
+
+## Data assimilation system configuration {#config}
+
+The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW, @skamarock2008) model. 
+The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa. 
+The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1). 
 The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.
@@
 For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; @han2006) is used as an observation operator to calculate model-simulated brightness temperatures.
-In addition to being already coupled to the GSI system.
 
 We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis.
-Since the publicly available GFS ensemble members change at each cycle we used random perturbations to generate the initial ensemble perturbations.
+with random perturbations to generate the initial ensemble perturbations.
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021.
@@
 
-(ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+(ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
 b) Locations of radiosonde launches during the RELAMPAGO.
@@
   scale_fill_gradient(low = "#f2f2f2", high = "#333333",
-                      name = NULL,
+                      name = "Altitude (m)",
                       breaks = seq(0, 6000, 1000),
@@
 These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft.
-The orange triangles in Figure \@ref(fig:dominio) indicate the location of the surface station included in this experiment.
+The orange triangles in Figure \@ref(fig:dominio) indicate the location of the surface stations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes.
@@
 We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America.
-The dataset used in this study has been obtained at the RELAMPAGO Data Set repository [@garcia2019].
+The dataset used in this study has been obtained from the RELAMPAGO Data Set repository [@garcia2019].
 These stations are indicated as green squares in Figure \@ref(fig:dominio)a.
 They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases.
-All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.
-Observation errors used to assimilate these observations are the same as for the CSWS (see Table \@ref(tab:table-obs))
+All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure, and 605 provide wind information.
+Observation errors used to assimilate these observations are the same as for the CSWS (see Table \@ref(tab:table-obs)).
 
@@
 
-Satellite-derived wind observations are also included in the PREPBUFR files available every 6 hs,  and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels).
+Satellite-derived wind observations are also included in the PREPBUFR files available every 6 h,  and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels).
 Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations).
@@
   # str()
-  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.",
+  kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error, and the gross check thresholds used.",
       col.names = c("Code", "Platform", "Variable", "Error", "Gross check"),
@@
 
-Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study.
+Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study.
 This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
@@
 The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM.
-The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
+The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere.
 The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
+First, a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
 Also, observations over the sea are preferred to those over land or snow [@hu2018].
@@
 The bias correction (BC) has an air-mass dependent and an angle-dependent component [@zhu2014] and it is calculated as a multi-linear function of N predictors $p_i(x)$, with associated coefficients $\beta_i$.
-Then, the bias corrected brightness temperature (BT_{bc}) can be obtained as:
+Then, the bias corrected brightness temperature ($BT_{bc}$) can be obtained as:
 
@@
 
-In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves the $\beta_i$ that provides the best fit between the simulation and the observations.
-The coefficients were initialized at 18 UTC Nov 18 with the GFS system coefficients.
+In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
+The coefficients were initialized at 18 UTC Nov 18, 2018 with the GFS system coefficients.
 The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.
@@
 
-(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AWS c) SATWND, and d) RAD experiments calculated over 2.5$^{\circ}$ boxes.
+(ref:obs-horizontal) Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AWS, c) SATWND, and d) RAD experiments calculated over 2.5$^{\circ}$ boxes.
 
@@
 
-(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.
+(ref:obs-cycle) a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamonds and line) experiments.
 
@@
 Only the soundings that did not enter the assimilation system were used for validation.
-The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio)b).
+The experiment period covers IOP missions 7 and 8, during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio)b).
 
@@
 * Radar observations are used to perform a qualitative and visual assessment of the convection features.
-The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes.
+The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes.
 For this work, only the maximum reflectivity in the column (COLMAX) closest to the analysis time was used.
@@
 In the third experiment, referred to as SATWND, we assimilate all the observations of the AWS experiment and the satellite-derived winds.
-Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbiting satellites as described in Section \@ref(sat).
+Finally, a fourth experiment referred to as RAD assimilates all available clear-sky radiances from sensors onboard polar orbiting satellites as described in section \@ref(sat).
 
@@
 The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
-The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors.
+The maxima at 13-14 and 01-02 UTC in RAD correspond to the contribution of the multispectral sensors.
 The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b) shows a maximum in low levels due to the AWS observations.
-Satellite-derived winds available in this particular case are maximized at the upper troposphere (between 500-250 hPa).
+Satellite-derived winds are maximized at the upper troposphere (between 500-250 hPa).
 Above 850 hPa, most of the observations correspond to radiance observations.
@@
 All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
-The initial 60-member ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \@ref(fig:cycle)).
-
-Ensemble forecast initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
+The initial 60-member ensemble is generated as explained in section \@ref(config) from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \@ref(fig:cycle)).
+
+Ensemble forecasts initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
 Both forecasts are integrated until 12 UTC Nov 23.
@@
 
-(ref:cycle) Diagram of the analysis cycles between Nov 20, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
+(ref:cycle) Figure 4.
+Diagram of the analysis cycles between 18 UTC Nov 20, and 12 UTC Nov 23 plus spin up period of 6 hours.
 The zoomed section shows the hourly assimilation that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
-The two IOPs missions from the RELAMPAGO Field Campaign are shown.
+The two IOP missions from the RELAMPAGO field campaign and the ensemble forecast initialized at 00 and 06 UTC Nov 22 are shown.
 
@@
 
-where $M$ is the sample size.
+where $M$ is the ensemble size.
 A consistent system will have no bias ($mean RCRV = 0$) and a standard deviation equal to 1 ($sd RCRV = 1$).
-If the ensemble has a positive (negative) bias, $mean RCRV$ will be positive (negative).
-A $sd RCRV > 1$ ($sd RCRV < 1$) indicates that the ensemble is underdispersive (overdispersive).
+If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative.
+A $sd RCRV > 1$ () indicates that the ensemble is underdispersive and an $sd RCRV < 1$ indicates that the ensemble is overdispersive.
 
@@
 
-where $O$ and $X$ stand for independent observations and the simulations respectively.
-
-For the comparison of the first-guess precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds.
+where $O$ and $X$ stand for independent observations, and the simulations respectively and N is the sample size.
+
+For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds.
 Following @maldonado2021 the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members :
@@
 Figure \@ref(fig:rcrv-profile) shows the mean and standard deviation of the RCRV for the upper-air observations.
-Figure \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
+Figures \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
 Both ADPUPA and AIRCFT show a generally good agreement between the ensemble spread and the observation error.
@@
 
-(ref:rcrv-sfc) $sd RCRV$ calculated for surface observations (from PREPBUFR and AWS) of a) wind, and b) temperature, averaged over 2.5º boxes.
+(ref:rcrv-sfc) First guess $sd RCRV$ calculated for surface observations (from PREPBUFR and AWS) of a) wind, and b) temperature averaged over 2.5º boxes for the RAD experiment.
+Observations were aggregated every hourly cycle for the entire experiment period.
 
@@
 
-(ref:rcrv-profile) Vertical profiles of $mean RCRV$ (dash line) and $sd RCRV$ (solid line) for  a) temperature and b) wind of sounding and aircraft observations, c) satellite-derived wind observations, and d) brightness temperature observations.
+(ref:rcrv-profile) Vertical profiles of first guess $mean RCRV$ (dashed line) and $sd RCRV$ (solid line) for  a) temperature and b) wind of sounding and aircraft observations, c) satellite-derived wind observations, and d) brightness temperature observations for the RAD experiment.
+Observations were aggregated every hourly cycle for the entire experiment period.
 
@@
 The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS.
-Figure \@ref(fig:TQ-diff)a-c shows the analysis difference between experiments in the spatially averaged vertical profile of temperature.
+Figures \@ref(fig:TQ-diff)a-c show the analysis difference between experiments in the spatially averaged vertical profile of temperature.
 By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
@@
 This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection.
-Figure \@ref(fig:TQ-diff)b indicates that SATWND observations have an almost negligible impact on the mean temperature throughout the experiment.
-This result is unexpected since mid-to-upper level winds can have a profound impact on the organization and evolution of MCSs over the region.
-During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations.
+Although the mid-to-upper circulation can have an important impact on the organization and evolution of the MCS over the region, the satellite-derived winds did not have an appreciable impact on the mean temperature and humidity (Figure \@ref(fig:TQ-diff)b-e), possibly due to the large observation errors used for the assimilation.
+During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations (Figure \@ref(fig:TQ-diff)c).
 No clear systematic impact is found above the PBL during this period.
@@
 
-Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f), the impact of assimilating AWS with finer spatial and temporal resolution is most substantial at low levels (Figure \@ref(fig:TQ-diff)d).
+Comparing the specific humidity in the experiments (Figures \@ref(fig:TQ-diff)d-f), the impact of assimilating AWS with fine spatial and temporal resolution is most substantial at low levels (Figure \@ref(fig:TQ-diff)d).
 The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime.
@@
 In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
-As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels (Figure \@ref(fig:TQ-diff)e).
+As for the temperature, the systematic impact of satellite-derived winds on moisture is small (Figure \@ref(fig:TQ-diff)e).
 Figure \@ref(fig:TQ-diff)f shows that radiances reduce low-middle level moisture during the first day of the experiment.
-The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov, 22.
-
-(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov 22.
+
+(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
 
@@
 ```
+
 The impacts on the wind components are shown in Figure \@ref(fig:UV-diff), along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations (for example, Figure \@ref(fig:UV-diff)a shows the zonal wind difference between AWS and CONV and the zonal wind for AWS).
-The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis.
+The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis (Figures \@ref(fig:UV-diff)a,b).
 There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with a stronger reduction of the northerly wind during night hours.
@@
 After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain.
-After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels.
-During Nov 22 and 23 the impact of assimilating AWS produces an increase of northerly wind in upper levels.
+After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels (Figures  \@ref(fig:UV-diff)d, f).
+During Nov 22 and 23 the impact of assimilating AWS observations produces an increase of northerly wind in upper levels.
 This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
-Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger; the systematic impact is overall smaller than the one produced by assimilating data from AWS  (Figure  \@ref(fig:UV-diff)b, e).
+Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is largest; the systematic impact is overall smaller than the one produced by assimilating data from AWS  (Figures  \@ref(fig:UV-diff)b, e).
 
 The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \@ref(fig:UV-diff)c).
-For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of AWS observations during the nights previous to the development of the MCS.
+For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of AWS observations during the nights, between 03 and 12 UTC, previous to the development of the MCS  (Figure \@ref(fig:UV-diff)f).
 At upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind speed.
@@
 
-(ref:UV-diff)  Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-Black contours correspond to u wind for AWS (a), SATWND (b) and RAD (c) and v wind for AWS (d), SATWND (e) and RAD (f).
+(ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+Black contours correspond to negative u wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD.
+Dashed contours correspond to negative v wind.
 
@@
 ```
+
 To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles).
@@
 We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
-The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
+The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
 AWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
@@
 
-The analyzed distribution of temperature and moisture in the PBL (Figure \@ref(fig:summary-fields)b) resembles the observations (Figure \@ref(fig:TQ-diff)a-c).
+The analyzed distribution of temperature and moisture in the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed in the temperature profiles (Figure \@ref(fig:TQ-diff)a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer that SATWND.
 On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
@@
 Moreover, in this same region, these experiments show larger MCAPE values than CONV.
-Note that wind shears over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowski2010].
-
-(ref:summary-fields) a) Precipitable water (shaded, $kgm^{-2}$) and average northerly wind on the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, $ms^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $ms^{-1}$ for each experiment.
-All fields correspond to the analysis ensemble mean for 00 UTC Nov, 22.
-Grey filled contours correspond to  topography over 1500 meter over sea level.
+Note that wind shear over 15 $ms^{-1}$ is associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowski2010].
+
+(ref:summary-fields) a) Precipitable water (shaded, $kgm^{-2}$) and average northerly wind over the first 7 sigma levels (from the surface up to approximately 800 hPa, contours, $ms^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels), and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $ms^{-1}$ for each experiment.
+All fields correspond to the analysis ensemble mean for 00 UTC Nov 22.
+Grey filled contours correspond to  topography over 1500 meters above sea level.
 
@@
 
-Firstly we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \@ref(fig:pp-hov)(a) shows the mean hourly accumulated precipitation as estimated by IMERG, and (b-e) the probability matching mean (PM) [@clark2017] for the first-guess hourly accumulated precipitation as averaged between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments.
+First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
+Figure \@ref(fig:pp-hov)a shows the mean hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) [@clark2017] for the first-guess hourly accumulated precipitation as averaged between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments.
 The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \@ref(fig:pp-hov)a).
 In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated.
-This is particularly evident in CONV (Figure \@ref(fig:pp-hov)b) where the convection initiation is delayed and occurs further north with respect to the observed initiation.
-AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figure \@ref(fig:pp-hov)c-e).
+This is particularly evident in CONV (Figure \@ref(fig:pp-hov)b), where the convection initiation is delayed and occurs further north with respect to the observed initiation.
+AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figures \@ref(fig:pp-hov)c-e).
 AWS and RAD show a more fragmented distribution compared with SATWND, possibly due to the development of less organized convection during Nov 22.
-After 18 UTC Nov 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
-
-(ref:pp-hov) Hövmoller of mean hourly accumulated precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67$^{\circ}$W and 54.5$^{\circ}$W.
+After 18 UTC Nov 22, RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of enhanced development of the convection.
+
+(ref:pphov) Hövmoller diagram of probability matched mean hourly accumulated 1-h forecast precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67$^{\circ}$W and 54.5$^{\circ}$W.
 Contours drawn every 0.5 $mmh^{-1}$, starting at 0.5 $mmh^{-1}$.
 
-```{r pp-hov, fig.cap="(ref:pp-hov)", fig.height=4, fig.env="figure*", fig.pos="h"}
+```{r pp-hov, fig.cap="(ref:pphov)", fig.height=4, fig.env="figure*", fig.pos="h"}
 
@@
   theme(legend.position = "bottom",
+        panel.spacing = unit(1, "lines"),
         tagger.panel.tag.text = element_text(size = 9)) +
@@
 To quantify the spatial match between the observed precipitation and the short-range forecast precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
-All experiments show values of FSS below 0.6 around 03 UTC Nov 22 indicating that the experiments were not able to accurately simulate the convective environment and the initialization of the convection.
-However, we notice that the FSS values for RAD around 00 UTC Nov 22 are higher than for the other experiment, indicating that radiance observations have a positive impact on the analysis.
-Moreover, the FSS values improve during the mature stage of the MCS for all spatial scales and thresholds.
+All experiments show similar values of FSS during the initialization of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
+This indicates that radiance observations have a positive impact on the analysis.
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
-AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
-The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22.
+AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for this case study.
+The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22 (Figure \@ref(fig:fss)b,d).
 The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).
 
-(ref:fss) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), AWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.
+(ref:fss) FSS calculated over 1-h forecast precipitation accumulated in a 6-hour moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), AWS (light blue line), SATWND (orange line) and RAD (red line) experiments.
 
@@
   .[w %in% c(1, 11) & q %in% c(1, 25) & exp %in% c("E2", "E5", "E6", "E9")] %>% 
-  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
+  .[, ":="(q_label = factor(q, labels = c("1~mmh^{-1}", "25~mmh^{-1}")),
+           w_label = factor(w, labels = c("10~km", "100~km")))] %>% 
   ggplot(aes(date, fss)) +
@@
   geom_line(aes(color = exp)) +
-  # geom_point(aes(color = exp)) +
   scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AWS", 
@@
   coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
-  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mmh-1", "5" = "5 mm", 
-                                            "10" = "10 mm", "25" = "25 mmh-1"),
-                                      w = c("1" = "10 km", "5" = "50 km", "11" = "100 km", 
-                                            "51" = "500 km", "101" = "1000 km"))) +
+  facet_grid(w_label ~ q_label, labeller = label_parsed) +
   tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
@@
 In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the 1-hour precipitation forecast skill (Figure \@ref(fig:fss)).
-Overall, none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
+Overall, none of the short-range forecasts capture the mesoscale details in the reflectivity distribution.
 This is partially expected considering the coarse horizontal grid spacing (10 km), which is not enough to appropriately represent the strength of the convective band associated with the MCS.
@@
 
-Finally, Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes), and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
-
-IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments.
+Finally, Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov 21 (including 30 radiosondes), and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+
+IOP 7 (Figures \@ref(fig:soundings)a-d) provides a good characterization of the pre-convective environment during the first day of our experiments.
 The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection.
-For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that is slightly reduced by the assimilation of AWS and radiances (Figure \@ref(fig:soundings)a).
-The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AWS, SATWND, and RAD is small above the PBL.
-The dew point temperature shows a small dry bias in low levels that is corrected in AWS, SATWND, and RAD (Figure \@ref(fig:soundings)b).
-At middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C) for all the experiments, indicating a more moist atmosphere in the analyses than observed.
-Overall, the RMSE grows with height, probably because at upper levels moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
-The zonal wind for AWS and SATWND has a substantial positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced to CONV values by assimilating satellite radiances (Figure \@ref(fig:soundings)c).
-It is somewhat surprising that no substantial effect is observed by assimilating satellite-derived winds.
-The assimilation of AWS and radiance observations increases the northerly wind bias and RMSE near the surface (Figure \@ref(fig:soundings)d).
-
-(ref:dbz-mean) Maximum reflectivity in the column (COLMAX in $dBZ$), observed (upper row) and 1-hr forecast ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22.
+In general, the experiments show a similar RMSE and bias for all the variables.
+AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias.
+However, AWS degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
+
+(ref:dbz-mean) Maximum reflectivity in the column (COLMAX in $dBZ$), observed (upper row) and 1-hr forecast ensemble mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018.
+Black circles in first row show the observation range of each radar.
 
@@
            "20181122160000", "20181122190000")
-files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[29]_", dates, ".nc"))
+files <- Sys.glob(paste0(derived_data, "dbz/E[29]_ana_", dates, "_dbz_pmm.rds"))
 
@@
 
-  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
-  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
-    .[, ":="(exp = meta[[1]][["exp"]],
-             date = ymd_hms(meta[[1]][["date"]]))] %>% 
-    .[, c("x", "y") := wrf_project(lon, lat)] 
+  read_rds(f)
+
+  # meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
+  # ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
+  #   .[, ":="(exp = meta[[1]][["exp"]],
+  #            date = ymd_hms(meta[[1]][["date"]]))] %>% 
+  #   .[, c("x", "y") := wrf_project(lon, lat)] 
 }) %>% 
@@
   .[, run := fifelse(exp == "E9", "RAD", "CONV")] %>%
-  .[, ":="(south_north = NULL,
-           west_east = NULL,
+  .[, ":="(ens_mean = NULL,
+           rank = NULL,
            exp = NULL)]
 
-values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
+# values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21,
+#             22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5,
+#             42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60,
+#             61.5, 63, 64.5, 66, 69, 76.5)
+
+values <- c(9, 12, 15, 16.5, 18, 19.5, 21,
             22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5,
@@
 
-colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
-             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
+colours <- c("#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
              "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
@@
                     drop = FALSE, 
+                    # limits = c(10, 76.5),
                     guide = guide_colorsteps(barwidth = 25,
@@
 
-For IOP 8 (Figure \@ref(fig:soundings)e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation.
-This area was also behind the cold front thus affected by low-level cold advection.
-Below 5 km all experiments show a warm bias which decreases with the assimilation of AWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in AWS, SATWND, and RAD.
-The RMSE for AWS, SATWND, and RAD remains below 2$^{\circ}$C for almost all levels with substantial improvement in upper levels where radiance observations have the largest impact.
-In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
-This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region.
-The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation, but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
-SATWND presents the lowest RMSE overall while AWS and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C.
-The zonal wind is overestimated in the analyses with a positive bias at middle levels.
-For this period only RAD shows an improvement with respect to CONV in the upper troposphere.
-At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD.
-In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).
-
-(ref:soundings) RMSE (solid line) and Bias (dashed line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+For IOP 8 (Figures \@ref(fig:soundings)e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation.
+This area was also behind the cold front and affected by low-level cold advection.
+The assimilation of AWS, SATWND, and RAD reduces the cold bias and RMSE for temperature between 5 and 12 km and the RMSE in the PBL compared with CONV (Figure \@ref(fig:soundings)e).
+The reduction of bias and RMSE is also important for dew point temperature (Figure \@ref(fig:soundings)f) with SATWND showing the bigger impact followed by AWS and RAD.
+The zonal wind is overestimated in the analyses and only RAD shows an improvement with respect to CONV in the upper troposphere (Figure \@ref(fig:soundings)g).
+At low levels the meridional wind (Figure \@ref(fig:soundings)g) presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD.
+In fact, low level biases in these experiments are higher than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).
+
+(ref:soundings) RMSE (solid line) and Bias (dashed line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing the analysis of each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
 The blue line corresponds to CONV, the light blue line to AWS, SATWND is represented with an orange line, and RAD with a red line.
@@
   .[lev <= 15000] %>% 
+  .[, ":="(iop_labels = factor(iop, labels = c("IOP~7", "IOP~8")),
+           variable_labels = factor(variable, labels = c("Temperature~(K)", "Dew~point~temperature~(K)", 
+                                                         "U~wind~(ms^{-1})", "V~wind~(ms^{-1})")))] %>% 
   ggplot(aes(lev/1000, value.est)) +
@@
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = c(colores_exp, "black"), labels = c(E2 = "CONV", E5 = "AWS", 
+  scale_color_manual(values = c(colores_exp), labels = c(E2 = "CONV", E5 = "AWS", 
                                                                   E6 = "SATWND", E9 = "RAD")) +
   coord_flip() +
-  facet_grid(iop~variable, scales = "free_x", 
-             labeller = labeller(variable = c("t" = "Temperature (K)",
-                                              "td" = "Dew point temperature (K)",
-                                              "u" = "U wind (ms-1)",
-                                              "v" = "V wind (ms-1)"),
-                                 iop = c("IOP07" = "IOP 7",
-                                         "IOP08" = "IOP 8"))) +
+  facet_grid(iop_labels ~ variable_labels, scales = "free_x", 
+             labeller = label_parsed) +
    tagger::tag_facets(position = list(x = 0.05, y = 0.95)) +
@@
 
-In this section, we analyse the 60-members ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
-We calculated the FSS for the ensemble forecast in 6-hours moving windows for the same thresholds and spatial scales as for the analysis to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
+In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
+We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the 1-h forecast to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
 CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
-AWS, SATWND, and RAD show an important improvement in the FSS values, particularly for the higher threshold (Figure \@ref(fig:fssfcst)b, d).
-Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
-
-The satellite-derived wind observations show a clearly positive impact on the forecast, in contrast to what was seen when comparing the analysis with independent observations in terms of precipitation.
-Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the analysis to IMERG estimations.
-
-(ref:fssfcst) FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the forecasts initialized from CONV (blue line), AWS (light blue line), SATWND (orange line) and RAD (red line) experiments at 00 UTC (solid line) and 06 UTC (dashed line).
+AWS, SATWND, and RAD show improvements in the FSS values, particularly for the higher threshold (Figure \@ref(fig:fssfcst)b, d).
+Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD tham the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
+
+The satellite-derived wind observations show a clearly positive impact on the forecast, in contrast to what was seen when comparing the 1-h forecast with independent observations in terms of precipitation.
+Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the 1-h forecast to IMERG estimations.
+
+(ref:fssfcst) FSS calculated over a 6-hour moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on 10 km (a and b) and 100 km (c and d) scales, for the forecasts initialized from CONV (blue line), AWS (light blue line), SATWND (orange line), and RAD (red line) experiments at 00 UTC (solid line) and 06 UTC (dashed line), Nov 22.
 
@@
   .[q %in% c(1, 25) & w %in% c(1, 11)] %>% 
+  .[, ":="(q_label = factor(q, labels = c("1~mmh^{-1}", "25~mmh^{-1}")),
+           w_label = factor(w, labels = c("10~km", "100~km")))] %>% 
   ggplot(aes(date, fss)) +
@@
   # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT",
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AWS",
                                                       E6 = "SATWND", E9 = "RAD")) +
@@
   scale_date(20181122000000, 20181123000000, "12 hours") +
-  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mmh-1", "5" = "5 mm", 
-                                            "10" = "10 mm", "25" = "25 mmh-1", "50" = "50 mm"),
-                                      w = c("1" = "10 km", "5" = "50 km", 
-                                            "11" = "100 km", 
-                                            "51" = "500 km", "101" = "1000 km"))) +
+  facet_grid(w_label ~ q_label, labeller = label_parsed) +
   tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
@@
 
-In this paper, we investigate for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional data assimilation system.
+In this paper, we investigate for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system.
 In particular, we explore the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors.
 Southern South America is a particularly interesting region due to the heterogeneity and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
-This combined with a climatology characterized by frequent organized convective events makes mesoscale data assimilation particularly challenging.
+This combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.
 To consider these issues, we used a case study approach of a massive mesoscale convective system that developed over Southern South America, particularly in a region where the surface and upper-air observing networks are too coarse to capture variability at the mesoscale and that took place on Nov 22, 2018.
@@
 
-In terms of the analysis, we found that AWS observations, which have high spatial and temporal resolution, produce major impacts throughout the troposphere and mainly on the PBL.
+In terms of the analysis, we found that AWS observations, which have high spatial and temporal resolution, produce impacts throughout the troposphere, especially in the PBL.
 In particular, the precipitable water content and the low level meridional circulation in the AWS experiment lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
@@
 
-Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement at low level when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
+Taking advantage of the observations taken during the RELAMPAGO field campaign, we compared the experiments with the soundings launched during IOPs 7 and 8.
+For the pre-convective environment (IOP 7), we observed an improvement at low levels when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
 In particular, AWS observations correct a dry bias already observed by @ruiz2010.
-For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels including more observations for the temperature but the opposite effect for the dew point temperature.
+For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels when assimilating more observations for the temperature but the opposite effect for the dew point temperature.
 In this case, the analyses have a moist bias that increases with the assimilation of AWS and radiance observations.
@@
 
-We also analyzed the performance of independent ensemble forecasts intialized from the analyses to estimate the precipitation during Nov 22.
-The forecast initialized from AWS, SATWND, and RAD are able to estimate the precipitation substantially better than CONV.
-In particular, the continuous assimilation of atellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
+We also analyzed the performance of independent ensemble forecasts intialized from the analyses to forecast the precipitation during Nov 22.
+The forecast initialized from AWS, SATWND, and RAD are able to forecast the precipitation substantially better than CONV.
+In particular, the continuous assimilation of satellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
 
 To summarize, in this study we conclude that the assimilation of surface observations with higher spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
-Moreover, the ensemble forecasts initialized from the analysis show promising results to estimate the precipitation of severe events.
+Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events.
 In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like  GPS radio occultation data and radiances from geostationary orbiting satellites.
@@
 We are very thankful to the Atmospheric and Sea Research Center (CIMA), the University of Buenos Aires (UBA), and the National Scientific and Technical Research Council (CONICET) who support this study.
-We acknowledge the Argentinian National Meteorological Service (ANMS) for kindly providing the radar observations used for validation.
+We acknowledge the Sistema Nacional de Radares Meteorológicos supported by the Secretaría de Infraestructura y Política Hídrica for kindly providing the radar observations used for validation and the National Meteorological Service for facilitating the access to the data.
 We also acknowledge the Cheyenne HPC resources (doi:10.5065/D6RX99HX) from NCAR's Computational and Information Systems Laboratory, National Science Foundation (project code UIUC0012).
-Also, PICT 2017-2033 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
+Also, PICT 2017-2233 and PICT 2018-3202 projects of the National Agency for the Promotion of Research, Technological Development and Innovation from Argentina partially funded this project.
 

### 2022-05-03 analysis/chapters/paper_ana_20181122.tex
Revised version

@@
             pdfauthor={},
-            pdftitle={Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO.
-Part I: Impact on the Analysis and First-guess},
+            pdftitle={Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO campaign},
             colorlinks=false,
@@
 
-  \title{Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis and First-guess}
+  \title{Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO campaign}
     \author[UBA,CIMA,CNRS]{Paola Belén Corrales\corref{1}}
@@
     \author[NCAR]{Craig S. Schwartz}
+
+    \author[Illinois]{Stephen W. Nesbitt}
 
@@
     \address[NCAR]{National Center for Atmospheric Research, Boulder, Colorado.}
+    \address[Illinois]{Department of Atmospheric Sciences, University of Illinois at Urbana--Champaign, Urbana, Illinois.}
       \cortext[1]{Corresponding Author}
@@
   \begin{abstract}
-  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north-eastern Argentina where the surface and upper air observing networks are relatively coarse.
-We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed during November 22, 2018.
+  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse.
+We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018.
 The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
-The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during November 2018.
-
-  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members multiphisics ensemble.
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files, AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI).
+The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign.
+
+  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND, and satellite radiances from different microwave and infrared sensors.
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-We also run ensemble forecasts initialized from each experiment show and evaluated their skill to predict precipitation.
-We found that the assimilated observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
+We also run ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
+We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
   \end{abstract}
@@
 A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
-Southern South America has been shown to have one of the highest frequencies in the world of favorable conditions for high-impact meteorological events (Brooks et al., 2003) and large hail events (Cecil and Blankenship, 2012), particularly during austral spring and summer.
+Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events (Brooks et al., 2003) and large hail events (Cecil and Blankenship, 2012), particularly during austral spring and summer.
 This is also confirmed by observational evidence and high impact weather reports (Matsudo et al., 2015; Rasmussen et al., 2014).
-Recently, the RELAMPAGO-CACTI field campaign Nesbitt et al.
-(2021) has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
-
-Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability, and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes).
-Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for the initialization of high-resolution forecasts (Sun et al., 2014) and thus has received increasing attention in the last decades.
+Recently, the RELAMPAGO-CACTI field campaign (Nesbitt et al., 2021) has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
+
+Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes).
+Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for high-resolution numerical forecasts (Sun et al., 2014) and thus has received increasing attention in the last decades.
 
@@
 
-In that regard, Wheatley and Stensrud (2010) investigated the impact of assimilating surface pressure data in a mesoscale ensemble-based data assimilation system but found limited impact in two case studies related to mesoscale convective systems.
-Ha and Snyder (2014) showed that assimilating temperature and dew point temperature from high resolution surface weather station networks systematically improved the structure of the simulated planetary boundary layer and improved the short range precipitation forecast over the U.S.
-Chang et al.
-(2017), Bae and Min (2022) and Chen et al.
-(2016) reported beneficial effects of assimilating surface weather stations in a high resolution DA system using an ensemble Kalman filter, a 3-dimensional variational system and a 4-dimensional variational system respectively.
-Positive impacts have been found on the temperature and moisture forecast in the planetary boundary layer and in the location of precipitation systems.
-Sobash and Stensrud (2015) showed in a mesoscale ensemble based DA system that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
-Maejima et al.
-(2019) investigated the impact of 1-minute frequency assimilation of surface observations using observation simulation experiments on a heavy rainfall case founding that assimilating high frequency and spatially dense observations lead to a better representation of mesoscale circulation even though the number of observations provided by surface stations is much less than the one provided by weather radars.
-Gasperoni et al.
-(2018) conducted a case study to evaluate the impact of assimilating observations produced by private weather stations which are not incorporated in the operational analysis.
-They found a positive effect of these observations upon the initiation of deep moist convection along a dryline.
+In that regard, several authors investigated the impact of assimilating surface weather data (e.g.~Wheatley and Stensrud (2010), Ha and Snyder (2014), Chang et al.
+(2017), Bae and Min (2022), Banos et al.
+(2021), Maejima et al.
+(2019), and Chen et al.
+(2016) using different assimilation methodologies.
+Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.
+Sobash and Stensrud (2015) showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
+More recently, Gasperoni et al.
+(2018) evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis.
+They found a positive effect of these observations upon the initiation of deep moist convection.
 This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.
 
-The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs) has been investigated in the context of limited-area mesoscale data assimilation.
-Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.~Wu et al.
+The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs), has also been investigated in the context of limited-area mesoscale DA.
+Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., Wu et al.
 (2014), Cherubini et al.
@@
 However, some works indicated mixed impacts (e.g.~Sawada et al.
-(2019) reported an improvement in the forecast of the track of the storm but a degradation in the forecasted intensity).
+(2019) reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity).
 As stated in J.
@@
 
-While the assimilation of radiance observations into global models is well established (), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low top models used for this application.
+While the assimilation of radiance observations into global models is well established (Eyre et al., 2020), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low model tops used for this application.
 Bao et al.
@@
 
-On Nov 22, a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
+On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
+To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
-
-\hypertarget{data-assimilation-system-configuration}{%
-\subsection{Data assimilation system configuration}\label{data-assimilation-system-configuration}}
-
-The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al.
+
+\hypertarget{config}{%
+\subsection{Data assimilation system configuration}\label{config}}
+
+The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW, Skamarock et al.
 (2008)) model.
-The horizontal grid spacing is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the model at 50 hPa.
-The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
+The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa.
+The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
 Department of Commerce (2015)).
@@
 Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
-For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al.
-(2006)) is used as an observation operator to calculate model-simulated brightness temperatures.
-In addition to being already coupled to the GSI system.
+For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) is used as an observation operator to calculate model-simulated brightness temperatures.
 
 We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis.
-Since the publicly available GFS ensemble members change at each cycle we used random perturbations to generate the initial ensemble perturbations.
+with random perturbations to generate the initial ensemble perturbations.
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in Necker et al.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
 b) Locations of radiosonde launches during the RELAMPAGO.
@@
 These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft.
-The orange triangles in Figure \ref{fig:dominio} indicate the location of the surface station included in this experiment.
+The orange triangles in Figure \ref{fig:dominio} indicate the location of the surface stations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes.
@@
 We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America.
-The dataset used in this study has been obtained at the RELAMPAGO Data Set repository (Garcia et al., 2019).
+The dataset used in this study has been obtained from the RELAMPAGO Data Set repository (Garcia et al., 2019).
 These stations are indicated as green squares in Figure \ref{fig:dominio}a.
 They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases.
-All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.
-Observation errors used to assimilate these observations are the same as for the CSWS (see Table \ref{tab:table-obs})
+All stations measure temperature, but only 395 stations provide humidity, 422 provide pressure, and 605 provide wind information.
+Observation errors used to assimilate these observations are the same as for the CSWS (see Table \ref{tab:table-obs}).
 
@@
 
-Satellite-derived wind observations are also included in the PREPBUFR files available every 6 hs, and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels).
+Satellite-derived wind observations are also included in the PREPBUFR files available every 6 h, and consist of estimations from GOES-16 (using the visible, infrared, and water vapor channels) and METEOSAT 8 and 11 (using the visible and water vapor channels).
 Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 \% of the observations).
@@
 
-\caption{\label{tab:table-obs}Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.}
+\caption{\label{tab:table-obs}Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error, and the gross check thresholds used.}
 \centering
@@
 
-\multirow{-4}{3.5em}{\raggedright\arraybackslash CSWS   ASWS} & \multirow{-4}{4.5em}{\raggedright\arraybackslash Surface weather stations} & Wind & 2.2 $ms^{-1}$ & 6 $ms^{-1}$\\
+\multirow{-4}{3.5em}{\raggedright\arraybackslash CSWS   AWS} & \multirow{-4}{4.5em}{\raggedright\arraybackslash Surface weather stations} & Wind & 2.2 $ms^{-1}$ & 6 $ms^{-1}$\\
 \cmidrule{1-5}
@@
 \cmidrule{1-5}
-ASCATW & Advance Scatterometers & Wind & 1.5 $ms^{-1}$ & 5 $ms^{-1}$\\
+ASCATW & Advanced Scatterometers & Wind & 1.5 $ms^{-1}$ & 5 $ms^{-1}$\\
 \cmidrule{1-5}
@@
 The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM.
-The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere.
+The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere.
 The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-Firstly a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
+First, a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
 Also, observations over the sea are preferred to those over land or snow (Hu et al., 2018).
@@
 The bias correction (BC) has an air-mass dependent and an angle-dependent component (Zhu et al., 2014) and it is calculated as a multi-linear function of N predictors \(p_i(x)\), with associated coefficients \(\beta_i\).
-Then, the bias corrected brightness temperature (BT\_\{bc\}) can be obtained as:
+Then, the bias corrected brightness temperature (\(BT_{bc}\)) can be obtained as:
 
@@
 
-In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves the \(\beta_i\) that provides the best fit between the simulation and the observations.
-The coefficients were initialized at 18 UTC Nov 18 with the GFS system coefficients.
+In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves for the \(\beta_i\) that provides the best fit between the simulation and the observations.
+The coefficients were initialized at 18 UTC Nov 18, 2018 with the GFS system coefficients.
 The assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each time.
@@
 
-\caption{\label{tab:table-rad}List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.}
+\caption{\label{tab:table-rad}List of the available sensors over several platforms, the number of accepted channels for the assimilation, and the percentage of assimilated observations calculated over all radiance observations and all cycles.}
 \centering
@@
 \cmidrule{1-4}
- & N15 & 2 & 3.31 \%\\
+ & NOAA15 & 2 & 3.31 \%\\
 \cmidrule{2-4}
- & N18 & 2 & 4.45 \%\\
+ & NOAA18 & 2 & 4.45 \%\\
 \cmidrule{2-4}
@@
 \cmidrule{1-4}
- & N19 & 2 & 0.68 \%\\
+ & NOAA19 & 2 & 0.68 \%\\
 \cmidrule{2-4}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/obs-horizontal-1} \caption{Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AWS c) SATWND, and d) RAD experiments calculated over 2.5\(^{\circ}\) boxes.}\label{fig:obs-horizontal}
+\includegraphics[width=1\linewidth]{../figures/obs-horizontal-1} \caption{Horizontal spatial distribution of the mean available observations per analysis cycle for the a) CONV, b) AWS, c) SATWND, and d) RAD experiments calculated over 2.5\(^{\circ}\) boxes.}\label{fig:obs-horizontal}
 \end{figure}
@@
 \begin{figure*}
-\includegraphics{../figures/obs-cycle-1} \caption{a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.}\label{fig:obs-cycle}
+\includegraphics{../figures/obs-cycle-1} \caption{a) Number of assimilated observations per cycle and b) time averaged number of assimilated observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AWS (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamonds and line) experiments.}\label{fig:obs-cycle}
 \end{figure*}
@@
 Only the soundings that did not enter the assimilation system were used for validation.
-The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \ref{fig:dominio}b).
+The experiment period covers IOP missions 7 and 8, during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \ref{fig:dominio}b).
 \item
@@
   Radar observations are used to perform a qualitative and visual assessment of the convection features.
-The data comes from 9 radars located on the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network (de Elía et al., 2017) with a temporal frequency of 10 minutes.
+The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network (de Elía et al., 2017) with a temporal frequency of 10 minutes.
 For this work, only the maximum reflectivity in the column (COLMAX) closest to the analysis time was used.
@@
 In the third experiment, referred to as SATWND, we assimilate all the observations of the AWS experiment and the satellite-derived winds.
-Finally, a fourth experiment referred to as RAD assimilates all available clear-sky observations from sensors onboard polar orbiting satellites as described in Section \ref{sat}.
+Finally, a fourth experiment referred to as RAD assimilates all available clear-sky radiances from sensors onboard polar orbiting satellites as described in section \ref{sat}.
 
@@
 The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
-The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors.
+The maxima at 13-14 and 01-02 UTC in RAD correspond to the contribution of the multispectral sensors.
 The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the AWS observations.
-Satellite-derived winds available in this particular case are maximized at the upper troposphere (between 500-250 hPa).
+Satellite-derived winds are maximized at the upper troposphere (between 500-250 hPa).
 Above 850 hPa, most of the observations correspond to radiance observations.
@@
 All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
-The initial 60-member ensemble is generated from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \ref{fig:cycle}).
-
-Ensemble forecast initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
+The initial 60-member ensemble is generated as explained in section \ref{config} from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \ref{fig:cycle}).
+
+Ensemble forecasts initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
 Both forecasts are integrated until 12 UTC Nov 23.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Diagram of the analysis cycles between Nov 20, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours.
+\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Figure 4.
+Diagram of the analysis cycles between 18 UTC Nov 20, and 12 UTC Nov 23 plus spin up period of 6 hours.
 The zoomed section shows the hourly assimilation that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
-The two IOPs missions from the RELAMPAGO Field Campaign are shown.}\label{fig:cycle}
+The two IOP missions from the RELAMPAGO field campaign and the ensemble forecast initialized at 00 and 06 UTC Nov 22 are shown.}\label{fig:cycle}
 \end{figure}
@@
 
-where \(M\) is the sample size.
+where \(M\) is the ensemble size.
 A consistent system will have no bias (\(mean RCRV = 0\)) and a standard deviation equal to 1 (\(sd RCRV = 1\)).
-If the ensemble has a positive (negative) bias, \(mean RCRV\) will be positive (negative).
-A \(sd RCRV > 1\) (\(sd RCRV < 1\)) indicates that the ensemble is underdispersive (overdispersive).
+If the ensemble has a positive bias, \(mean RCRV\) will be positive, on the opposite, if the ensemble has a negative bias, \(mean RCRV\) will be negative.
+A \(sd RCRV > 1\) () indicates that the ensemble is underdispersive and an \(sd RCRV < 1\) indicates that the ensemble is overdispersive.
 
@@
 
-where \(O\) and \(X\) stand for independent observations and the simulations respectively.
-
-For the comparison of the first-guess precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds.
+where \(O\) and \(X\) stand for independent observations, and the simulations respectively and N is the sample size.
+
+For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds.
 Following Maldonado et al.
@@
 Figure \ref{fig:rcrv-profile} shows the mean and standard deviation of the RCRV for the upper-air observations.
-Figure \ref{fig:rcrv-profile}a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
+Figures \ref{fig:rcrv-profile}a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT).
 Both ADPUPA and AIRCFT show a generally good agreement between the ensemble spread and the observation error.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/rcrv-sfc-1} \caption{\(sd RCRV\) calculated for surface observations (from PREPBUFR and AWS) of a) wind, and b) temperature, averaged over 2.5º boxes.}\label{fig:rcrv-sfc}
+\includegraphics[width=1\linewidth]{../figures/rcrv-sfc-1} \caption{First guess \(sd RCRV\) calculated for surface observations (from PREPBUFR and AWS) of a) wind, and b) temperature averaged over 2.5º boxes for the RAD experiment.
+Observations were aggregated every hourly cycle for the entire experiment period.}\label{fig:rcrv-sfc}
 \end{figure}
@@
 \begin{figure*}
-\includegraphics{../figures/rcrv-profile-1} \caption{Vertical profiles of \(mean RCRV\) (dash line) and \(sd RCRV\) (solid line) for a) temperature and b) wind of sounding and aircraft observations, c) satellite-derived wind observations, and d) brightness temperature observations.}\label{fig:rcrv-profile}
+\includegraphics{../figures/rcrv-profile-1} \caption{Vertical profiles of first guess \(mean RCRV\) (dashed line) and \(sd RCRV\) (solid line) for a) temperature and b) wind of sounding and aircraft observations, c) satellite-derived wind observations, and d) brightness temperature observations for the RAD experiment.
+Observations were aggregated every hourly cycle for the entire experiment period.}\label{fig:rcrv-profile}
 \end{figure*}
@@
 The analysis is performed over a smaller domain (red box in Figure \ref{fig:dominio}a) to focus on the region most directly affected by the MCS.
-Figure \ref{fig:TQ-diff}a-c shows the analysis difference between experiments in the spatially averaged vertical profile of temperature.
+Figures \ref{fig:TQ-diff}a-c show the analysis difference between experiments in the spatially averaged vertical profile of temperature.
 By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
@@
 This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection.
-Figure \ref{fig:TQ-diff}b indicates that SATWND observations have an almost negligible impact on the mean temperature throughout the experiment.
-This result is unexpected since mid-to-upper level winds can have a profound impact on the organization and evolution of MCSs over the region.
-During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations.
+Although the mid-to-upper circulation can have an important impact on the organization and evolution of the MCS over the region, the satellite-derived winds did not have an appreciable impact on the mean temperature and humidity (Figure \ref{fig:TQ-diff}b-e), possibly due to the large observation errors used for the assimilation.
+During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensates for the cooling effect of AWS observations (Figure \ref{fig:TQ-diff}c).
 No clear systematic impact is found above the PBL during this period.
@@
 
-Comparing the specific humidity in the experiments (Figure \ref{fig:TQ-diff}d-f), the impact of assimilating AWS with finer spatial and temporal resolution is most substantial at low levels (Figure \ref{fig:TQ-diff}d).
+Comparing the specific humidity in the experiments (Figures \ref{fig:TQ-diff}d-f), the impact of assimilating AWS with fine spatial and temporal resolution is most substantial at low levels (Figure \ref{fig:TQ-diff}d).
 The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime.
@@
 In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
-As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels (Figure \ref{fig:TQ-diff}e).
+As for the temperature, the systematic impact of satellite-derived winds on moisture is small (Figure \ref{fig:TQ-diff}e).
 Figure \ref{fig:TQ-diff}f shows that radiances reduce low-middle level moisture during the first day of the experiment.
-The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov, 22.
+The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov 22.
 
@@
 
-\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:TQ-diff}
+\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:TQ-diff}
 \end{figure*}
@@
 The impacts on the wind components are shown in Figure \ref{fig:UV-diff}, along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations (for example, Figure \ref{fig:UV-diff}a shows the zonal wind difference between AWS and CONV and the zonal wind for AWS).
-The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis.
+The assimilation of AWS produces a more easterly wind and a less northerly wind at low levels during the first two days of analysis (Figures \ref{fig:UV-diff}a,b).
 There is a diurnal cycle in the impact of surface weather stations on the meridional velocity (Figure \ref{fig:UV-diff}d) with a stronger reduction of the northerly wind during night hours.
@@
 After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain.
-After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels.
-During Nov 22 and 23 the impact of assimilating AWS produces an increase of northerly wind in upper levels.
+After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels (Figures \ref{fig:UV-diff}d, f).
+During Nov 22 and 23 the impact of assimilating AWS observations produces an increase of northerly wind in upper levels.
 This could be a consequence of a stronger MCS with an increased polar side upper level outflow.
-Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger; the systematic impact is overall smaller than the one produced by assimilating data from AWS (Figure \ref{fig:UV-diff}b, e).
+Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is largest; the systematic impact is overall smaller than the one produced by assimilating data from AWS (Figures \ref{fig:UV-diff}b, e).
 
 The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c).
-For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of AWS observations during the nights previous to the development of the MCS.
+For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of AWS observations during the nights, between 03 and 12 UTC, previous to the development of the MCS (Figure \ref{fig:UV-diff}f).
 At upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind speed.
@@
 
-\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.
-Black contours correspond to u wind for AWS (a), SATWND (b) and RAD (c) and v wind for AWS (d), SATWND (e) and RAD (f).}\label{fig:UV-diff}
+\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.
+Black contours correspond to negative u wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD.
+Dashed contours correspond to negative v wind.}\label{fig:UV-diff}
 \end{figure*}
@@
 We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
-The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
+The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
 AWS and SATWND experiments are very similar, with values of precipitable water over 55 \(kgm^{-2}\) north of 30\(^{\circ}\)S and a similar vertical distribution of specific humidity (not shown).
@@
 
-The analyzed distribution of temperature and moisture in the PBL (Figure \ref{fig:summary-fields}b) resembles the observations (Figure \ref{fig:TQ-diff}a-c).
+The analyzed distribution of temperature and moisture in the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed in the temperature profiles (Figure \ref{fig:TQ-diff}a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer that SATWND.
 On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
@@
 Moreover, in this same region, these experiments show larger MCAPE values than CONV.
-Note that wind shears over 15 \(ms^{-1}\) have been associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowski and Richardson, 2010).
+Note that wind shear over 15 \(ms^{-1}\) is associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowski and Richardson, 2010).
 
 \begin{figure*}
-\includegraphics[width=1\linewidth]{../figures/summary-fields-1} \caption{a) Precipitable water (shaded, \(kgm^{-2}\)) and average northerly wind on the first 7 sigma levels (from the surface and up to approximately 800 hPa, contours, \(ms^{-1}\)), b) Average potential temperature for the PBL (first 10 sigma levels) and c) Maximum CAPE and \textasciitilde0-6 km wind shear over 15 and 30 \(ms^{-1}\) for each experiment.
-All fields correspond to the analysis ensemble mean for 00 UTC Nov, 22.
-Grey filled contours correspond to topography over 1500 meter over sea level.}\label{fig:summary-fields}
+\includegraphics[width=1\linewidth]{../figures/summary-fields-1} \caption{a) Precipitable water (shaded, \(kgm^{-2}\)) and average northerly wind over the first 7 sigma levels (from the surface up to approximately 800 hPa, contours, \(ms^{-1}\)), b) Average potential temperature for the PBL (first 10 sigma levels), and c) Maximum CAPE and \textasciitilde0-6 km wind shear over 15 and 30 \(ms^{-1}\) for each experiment.
+All fields correspond to the analysis ensemble mean for 00 UTC Nov 22.
+Grey filled contours correspond to topography over 1500 meters above sea level.}\label{fig:summary-fields}
 \end{figure*}
@@
 
-Firstly we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \ref{fig:pp-hov}(a) shows the mean hourly accumulated precipitation as estimated by IMERG, and (b-e) the probability matching mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments.
+First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
+Figure \ref{fig:pp-hov}a shows the mean hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments.
 The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \ref{fig:pp-hov}a).
 In all the experiments, the accumulated precipitation in the short-range forecasts is underestimated.
-This is particularly evident in CONV (Figure \ref{fig:pp-hov}b) where the convection initiation is delayed and occurs further north with respect to the observed initiation.
-AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figure \ref{fig:pp-hov}c-e).
+This is particularly evident in CONV (Figure \ref{fig:pp-hov}b), where the convection initiation is delayed and occurs further north with respect to the observed initiation.
+AWS, SATWND, and RAD better capture the timing and location of convective initiation (Figures \ref{fig:pp-hov}c-e).
 AWS and RAD show a more fragmented distribution compared with SATWND, possibly due to the development of less organized convection during Nov 22.
-After 18 UTC Nov 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the convection.
+After 18 UTC Nov 22, RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of enhanced development of the convection.
 
 \begin{figure*}[h]
-\includegraphics{../figures/pp-hov-1} \caption{Hövmoller of mean hourly accumulated precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W.
+\includegraphics{../figures/pp-hov-1} \caption{Hövmoller diagram of probability matched mean hourly accumulated 1-h forecast precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W.
 Contours drawn every 0.5 \(mmh^{-1}\), starting at 0.5 \(mmh^{-1}\).}\label{fig:pp-hov}
@@
 To quantify the spatial match between the observed precipitation and the short-range forecast precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}).
-All experiments show values of FSS below 0.6 around 03 UTC Nov 22 indicating that the experiments were not able to accurately simulate the convective environment and the initialization of the convection.
-However, we notice that the FSS values for RAD around 00 UTC Nov 22 are higher than for the other experiment, indicating that radiance observations have a positive impact on the analysis.
-Moreover, the FSS values improve during the mature stage of the MCS for all spatial scales and thresholds.
+All experiments show similar values of FSS during the initialization of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
+This indicates that radiance observations have a positive impact on the analysis.
 The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS.
-AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study.
-The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22.
+AWS and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for this case study.
+The assimilation of radiances led to an overall improvement of the 1-hour forecast precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov 22 (Figure \ref{fig:fss}b,d).
 The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and also for spatial scales above 500 km, not shown).
@@
 \includegraphics{../figures/fss-1.png}
-\caption{\label{fig:fss}FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), AWS (ligh blue line), SATWND (orange line) and RAD (red line) experiments.}
+\caption{\label{fig:fss}FSS calculated over 1-h forecast precipitation accumulated in a 6-hour moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on 10 km (a and b) and 100 km (c and d) scales, for the first-guess of CONV (blue line), AWS (light blue line), SATWND (orange line) and RAD (red line) experiments.}
 \end{figure}
@@
 In addition, they are the worst (CONV) and best (RAD) performing experiments in terms of the 1-hour precipitation forecast skill (Figure \ref{fig:fss}).
-Overall, none of the short-range forecasts is able to capture the mesoscale details in the reflectivity distribution.
+Overall, none of the short-range forecasts capture the mesoscale details in the reflectivity distribution.
 This is partially expected considering the coarse horizontal grid spacing (10 km), which is not enough to appropriately represent the strength of the convective band associated with the MCS.
@@
 
-Finally, Figure \ref{fig:soundings} shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes), and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
-
-IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments.
+Finally, Figure \ref{fig:soundings} shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov 21 (including 30 radiosondes), and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes).
+
+IOP 7 (Figures \ref{fig:soundings}a-d) provides a good characterization of the pre-convective environment during the first day of our experiments.
 The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection.
-For the temperature, CONV presents a warm bias on the PBL of 1\(^{\circ}\)C and an RMSE between 1 and 2.5\(^{\circ}\)C that is slightly reduced by the assimilation of AWS and radiances (Figure \ref{fig:soundings}a).
-The RMSE for temperature remains between 0.5 and 2\(^{\circ}\)C throughout the troposphere and lower stratosphere and the impact of additional observations in the AWS, SATWND, and RAD is small above the PBL.
-The dew point temperature shows a small dry bias in low levels that is corrected in AWS, SATWND, and RAD (Figure \ref{fig:soundings}b).
-At middle levels (below 10 km) the bias is large and positive (8 - 10\(^{\circ}\)C) for all the experiments, indicating a more moist atmosphere in the analyses than observed.
-Overall, the RMSE grows with height, probably because at upper levels moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures.
-The zonal wind for AWS and SATWND has a substantial positive bias of approximately 10 \(ms^{-1}\) in the upper troposphere which is mainly reduced to CONV values by assimilating satellite radiances (Figure \ref{fig:soundings}c).
-It is somewhat surprising that no substantial effect is observed by assimilating satellite-derived winds.
-The assimilation of AWS and radiance observations increases the northerly wind bias and RMSE near the surface (Figure \ref{fig:soundings}d).
+In general, the experiments show a similar RMSE and bias for all the variables.
+AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias.
+However, AWS degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \ref{fig:soundings}c).
 
 \begin{figure*}[h]
-\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast ensemble mean maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22.}\label{fig:dbz-mean}
+\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast ensemble mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018.
+Black circles in first row show the observation range of each radar.}\label{fig:dbz-mean}
 \end{figure*}
 
-For IOP 8 (Figure \ref{fig:soundings}e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation.
-This area was also behind the cold front thus affected by low-level cold advection.
-Below 5 km all experiments show a warm bias which decreases with the assimilation of AWS and radiances.
-Between 5 and 12 km, the bias in each experiment is cold and is reduced in AWS, SATWND, and RAD.
-The RMSE for AWS, SATWND, and RAD remains below 2\(^{\circ}\)C for almost all levels with substantial improvement in upper levels where radiance observations have the largest impact.
-In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations.
-This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region.
-The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation, but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.
-SATWND presents the lowest RMSE overall while AWS and RAD are comparable between 4 and 12 km with values that do not exceed 7\(^{\circ}\)C.
-The zonal wind is overestimated in the analyses with a positive bias at middle levels.
-For this period only RAD shows an improvement with respect to CONV in the upper troposphere.
-At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD.
-In fact, low level biases in these experiments are higher (in module) than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).
+For IOP 8 (Figures \ref{fig:soundings}e-h), the densely observed area was behind the MCS, but far enough from it to not be directly affected by its mesoscale circulation.
+This area was also behind the cold front and affected by low-level cold advection.
+The assimilation of AWS, SATWND, and RAD reduces the cold bias and RMSE for temperature between 5 and 12 km and the RMSE in the PBL compared with CONV (Figure \ref{fig:soundings}e).
+The reduction of bias and RMSE is also important for dew point temperature (Figure \ref{fig:soundings}f) with SATWND showing the bigger impact followed by AWS and RAD.
+The zonal wind is overestimated in the analyses and only RAD shows an improvement with respect to CONV in the upper troposphere (Figure \ref{fig:soundings}g).
+At low levels the meridional wind (Figure \ref{fig:soundings}g) presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AWS, SATWND, and RAD.
+In fact, low level biases in these experiments are higher than in the CONV experiment, indicating a detrimental effect of the additional observations (possibly associated with the effect of AWS).
 
@@
 
-\caption{RMSE (solid line) and Bias (dashed line) of a) temperature (\(K\)), b) dew point temperature (\(K\)), c) u wind (\(ms^{-1}\)) and d) v wind (\(ms^{-1}\)) calculated by comparing each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+\caption{RMSE (solid line) and Bias (dashed line) of a) temperature (\(K\)), b) dew point temperature (\(K\)), c) u wind (\(ms^{-1}\)) and d) v wind (\(ms^{-1}\)) calculated by comparing the analysis of each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
 The blue line corresponds to CONV, the light blue line to AWS, SATWND is represented with an orange line, and RAD with a red line.}\label{fig:soundings}
@@
 
-In this section, we analyse the 60-members ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
-We calculated the FSS for the ensemble forecast in 6-hours moving windows for the same thresholds and spatial scales as for the analysis to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}).
+In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
+We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the 1-h forecast to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}).
 CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
-AWS, SATWND, and RAD show an important improvement in the FSS values, particularly for the higher threshold (Figure \ref{fig:fssfcst}b, d).
-Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
-
-The satellite-derived wind observations show a clearly positive impact on the forecast, in contrast to what was seen when comparing the analysis with independent observations in terms of precipitation.
-Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the analysis to IMERG estimations.
+AWS, SATWND, and RAD show improvements in the FSS values, particularly for the higher threshold (Figure \ref{fig:fssfcst}b, d).
+Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD tham the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
+
+The satellite-derived wind observations show a clearly positive impact on the forecast, in contrast to what was seen when comparing the 1-h forecast with independent observations in terms of precipitation.
+Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the 1-h forecast to IMERG estimations.
 
@@
 \includegraphics{../figures/fssfcst-1.png}
-\caption{\label{fig:fssfcst}FSS calculated over a 6-hours moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on a 10 km (a and b) and 100 km (c and d) scales, for the forecasts initialized from CONV (blue line), AWS (light blue line), SATWND (orange line) and RAD (red line) experiments at 00 UTC (solid line) and 06 UTC (dashed line).}
+\caption{\label{fig:fssfcst}FSS calculated over a 6-hour moving window for 1 mm (a and c) and 25 mm (b and d) thresholds, on 10 km (a and b) and 100 km (c and d) scales, for the forecasts initialized from CONV (blue line), AWS (light blue line), SATWND (orange line), and RAD (red line) experiments at 00 UTC (solid line) and 06 UTC (dashed line), Nov 22.}
 \end{figure}
@@
 
-Taking advantage of the observations taken during the RELAMPAGO field campaign we compared the experiments with the soundings launched during IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement at low level when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
+Taking advantage of the observations taken during the RELAMPAGO field campaign, we compared the experiments with the soundings launched during IOPs 7 and 8.
+For the pre-convective environment (IOP 7), we observed an improvement at low levels when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
 In particular, AWS observations correct a dry bias already observed by Ruiz et al.
 (2010).
-For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels including more observations for the temperature but the opposite effect for the dew point temperature.
+For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels when assimilating more observations for the temperature but the opposite effect for the dew point temperature.
 In this case, the analyses have a moist bias that increases with the assimilation of AWS and radiance observations.
@@
 
-We also analyzed the performance of independent ensemble forecasts intialized from the analyses to estimate the precipitation during Nov 22.
-The forecast initialized from AWS, SATWND, and RAD are able to estimate the precipitation substantially better than CONV.
-In particular, the continuous assimilation of atellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
+We also analyzed the performance of independent ensemble forecasts intialized from the analyses to forecast the precipitation during Nov 22.
+The forecast initialized from AWS, SATWND, and RAD are able to forecast the precipitation substantially better than CONV.
+In particular, the continuous assimilation of satellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
 
 To summarize, in this study we conclude that the assimilation of surface observations with higher spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
-Moreover, the ensemble forecasts initialized from the analysis show promising results to estimate the precipitation of severe events.
+Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events.
 In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like GPS radio occultation data and radiances from geostationary orbiting satellites.
@@
 doi:\href{https://doi.org/10.3390/rs14030605}{10.3390/rs14030605}
+
+\leavevmode\hypertarget{ref-banos2021}{}%
+Banos, I.H., Mayfield, W.D., Ge, G., Sapucci, L.F., Carley, J.R., Nance, L., 2021.
+Assessment of the data assimilation framework for the Rapid Refresh Forecast System v0.1 and impacts on forecasts of convective storms.
+Geoscientific Model Development Discussions 1--36.
+doi:\href{https://doi.org/10.5194/gmd-2021-289}{10.5194/gmd-2021-289}
 
@@
 Earth Observing Laboratory, U. -, 2020. Multi-network composite highest resolution radiosonde data. Version 1.3. UCAR/NCAR - earth observing laboratory.
+
+\leavevmode\hypertarget{ref-eyre2020}{}%
+Eyre, J.R., English, S.J., Forsythe, M., 2020.
+Assimilation of satellite data in numerical weather prediction.
+Part I: The early years.
+Quarterly Journal of the Royal Meteorological Society 146, 49--68.
+doi:\href{https://doi.org/10.1002/qj.3654}{10.1002/qj.3654}
 

### 2022-05-03 analysis/chapters/experimento_20181122.Rmd
tidying folders.

@@
 ```{r dbz-ens}
-files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[4-7]_0*_20181122100000.nc"))
+files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[5,6,8]_0*_20181122130000.nc"))
 
@@
 
+values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
+            22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
+            42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
+            61.5, 63, 64.5, 66, 69, 76.5)
+
+colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
+             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
+             "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
+             "#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
+             "#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
+             "#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
+             "#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
+             "#95E3C5", "#85E0BD")
+
 dbz %>% 
@@
   scale_fill_manual(values = colours, 
-                    drop = FALSE, guide = guide_colorsteps(barwidth = 0.8, 
-                                                           barheight = 30)) +
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 25,
+                                             barheight = 0.4, 
+                                             title.position = "left", 
+                                             title.vjust = 1,
+                                             frame.colour = "black")) +
   geom_mapa() +
@@
                       labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                  E6 = "SATWND", E7 = "RAD"))) +
+                                                  E6 = "SATWND", E8 = "RAD"))) +
   labs(fill = NULL) +
   theme_minimal() +
-  theme(#legend.position = "bottom",
+  theme(legend.position = "bottom",
     legend.box.spacing = margin(0, 0, 0, 0),
@@
 # rm(dbz)
+
+dbz %>% 
+  .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(exp, date, lon, lat, x, y)] %>% 
+  # .[date %in% unique(date)[1:2]] %>% 
+  .[, date := factor(date)] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    labels = function(x) JumpBy(x, 3, fill = ""),
+                    drop = FALSE, 
+                    guide = guide_colorsteps(barwidth = 25,
+                                             barheight = 0.4, 
+                                             title.position = "left", 
+                                             title.vjust = 1,
+                                             frame.colour = "black")) +
+  geom_mapa() +
+  # ggh4x::facet_nested(date + cumulus ~ exp, 
+  #                     labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+  #                                                 E6 = "SATWND", E8 = "RAD"))) +
+  facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
+                                                  E6 = "SATWND", E8 = "RAD"))) +
+  labs(fill = NULL) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+    legend.box.spacing = margin(0, 0, 0, 0),
+    panel.ontop = TRUE,
+    panel.grid = element_line(linetype = 3))
 ```

### 2022-05-03 analysis/chapters/forecast_20181122.Rmd
tidying folders.

--- 2021-03-03 analysis/chapters/forecast_20181122.Rmd
@@
                     breaks = seq(1500, 8000, by = 700)),
-  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
+  scale_fill_gradient(low = "#E2E6E6", high = "#7E8E8E", guide = "none",
                       oob = scales::squish))
@@
 
-experiments <- c("E4", "E5", "E6", "E7")
-#Color blind friendly but E4 y E7 parecidos en Desaturated
-colores_exp <- c(E4 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E7 = "#CC3311") 
-# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
+experiments <- c("E2", "E5", "E6", "E8")
+#Color blind friendly but E4 y E8 parecidos en Desaturated
+colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE8733", E9 = "#CC3311") 
+# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E8 = "#882255")
 satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
   setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
-derived_data <- "/home/paola.corrales/datosmunin/EXP/derived_data/"
+derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
 
@@
 ```{r}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[4-7]_ens.csv")
+files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,8,9]_ens.csv")
 
@@
 fss %>% 
+  .[q %in% c(1, 25) & w %in% c(1, 11)] %>% 
   ggplot(aes(date, fss)) +
@@
   # geom_point(aes(color = exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT",
+                                                      E6 = "SATWND", E8 = "viejo", E9 = "RAD")) +
+  scale_linetype_manual(values = c(1, 2), labels = c("2018-11-22 00:00:00" = "OO UTC", 
+                                   "2018-11-22 06:00:00" = "06 UTC")) +
   scale_date(20181122000000, 20181123000000, "12 hours") +
   facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
-                                            "10" = "10 mm", "25" = "25 mm"),
-                                      w = c("1" = "1 km", "5" = "50 km", 
+                                            "10" = "10 mm", "25" = "25 mm", "50" = "50 mm"),
+                                      w = c("1" = "10 km", "5" = "50 km", 
                                             "11" = "100 km", 
@@
 
+```{r}
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2569]_fcst_201811220*pp_pmm.rds")
+
+ppacum_1h <- purrr::map(files, function(f) {
+
+  meta <- unglue(basename(f), "{exp}_fcst_{inidate}_pp_pmm.rds")
+
+  readRDS(f) %>% 
+    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+    .[, ":="(rank = NULL, 
+             ens_mean = NULL,
+             ini_date = ymd_h(meta[[1]][["inidate"]]))] %>% 
+  setnames(c("pp_pmm"), c("pp_acum"))
+
+}) %>% rbindlist()
+
+imerg <- readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
+  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+  .[x %between% c(-990000, 990000) & y %between% c(-1190000, 1190000)] %>% 
+  .[end_date %between% c(as_datetime("20181122060000"), as_datetime("20181123120000"))] %>% 
+  .[, exp := "IMERG"]
+# %>% 
+#   .[, .(pp_acum = mean(pp_acum),
+#         lat = mean(lat),
+#         exp = "IMERG"), by = .(end_date, y)]
+
+ppacum_1h[, .(mm = sum(pp_acum)/47561), by = .(exp, date, ini_date)] %>% 
+  ggplot(aes(date, mm)) +
+  geom_line(aes(color =  exp, linetype = factor(ini_date))) +
+    scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT",
+                                                      E6 = "SATWND", E9 = "RAD")) +
+  geom_line(data = imerg[, .(mm = sum(pp_acum)/47561), by = .(exp, end_date)], 
+            aes(x = end_date, linetype = "IMERG"), color = "black") +
+  labs(x = "Time", y = "Total pp [mm/grid points]", color = NULL, linetype =  NULL) +
+  theme_minimal()
+```
+
+```{r}
+readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
+  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
+  .[end_date %between% c(as_datetime("20181122060000"), as_datetime("20181123120000"))] %>% 
+  .[, .(pp_acum = mean(pp_acum),
+        lat = mean(lat),
+        exp = "IMERG"), by = .(end_date, y)] %>% 
+  ggplot(aes(end_date, lat)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(0.5, 13, 0.5), Inf)) +
+  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(0.5, 13, 0.5)) +
+  # scale_fill_viridis_c(option = "A",
+  scale_fill_distiller(palette = "YlGnBu",
+                       direction = 1,
+                       labels = function(x) JumpBy(x, 2, fill = ""),
+                       super = ScaleDiscretised,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.3, 
+                                                title.position = "left", 
+                                                title.vjust = 1,
+                                                frame.colour = "black")) +
+  scale_date(ini = 20181122060000, break_bin = "12 hours") +
+  scale_y_latitude(breaks = seq(-40, -20, 5)) +
+  facet_wrap(~ exp) +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
+  theme_minimal(base_size = 9) +
+  theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 9)) +
+
+ppacum_1h %>% 
+  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000)] %>% 
+  .[, lat := lat[1], by = .(y)] %>% 
+  .[, .(pp_acum = mean(pp_acum)), by = .(exp, date, y, lat)] %>% 
+   # .[date %between% c(as_datetime("20181121180000"), as_datetime("20181123120000"))] %>% 
+  ggplot(aes(date, lat)) +
+  geom_contour_fill(aes(z = pp_acum, fill = stat(level)), breaks = c(seq(0.5, 13, 0.5), Inf)) +
+  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(0.5, 13, 0.5)) +
+  # scale_fill_viridis_c(option = "A",
+  scale_fill_distiller(palette = "YlGnBu",
+                       direction = 1,
+                       labels = function(x) JumpBy(x, 2, fill = ""),
+                       super = ScaleDiscretised,
+                       guide = guide_colorsteps(barwidth = 25,
+                                                barheight = 0.3, 
+                                                title.position = "left", 
+                                                title.vjust = 1,
+                                                frame.colour = "black")) +
+  scale_date(ini = 20181122060000, break_bin = "12 hours") +
+  scale_y_latitude(breaks = seq(-40, -20, 5), labels = NULL) +
+  facet_wrap(~ exp, ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "ASWS", 
+                                                          E6 = "SATWND", E9 = "RAD"))) +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
+  theme_minimal(base_size = 9) +
+  theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 9)) +
+
+  plot_layout(ncol = 2, widths = c(0.2, 0.8), guides = "collect") & theme(legend.position = 'bottom')
+```
+```{r}
+
+imerg <- read_rds("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds")
+
+q <- c(1, 5, 10, 25, 50)
+imerg <- purrr::map(q, function(f){
+
+  imerg %>%
+    .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
+    .[lon %between% lon_band & lat %between% lat_band] %>%
+    .[, .(area = sum(pp_acum > f)/.N), by = .(end_date)] %>%
+    .[, umbral := f] %>%
+    .[]
+}) %>%
+  rbindlist()
+
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2569]_fcst_*_area_acum_1h.rds")
+
+pp_area <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/{exp}_fcst_{ini_date}_area_acum_1h.rds")
+
+  readRDS(f) %>% 
+    .[, ":="(ini_date = ymd_h(meta[[1]][["ini_date"]]))]
+  # %>% 
+  #   .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)]
+
+}) %>% rbindlist()
+
+pp_area %>% 
+  .[umbral %in% c(1, 5, 10) ] %>%
+  .[, .(mean_area = mean(area),
+        max_area = max(area),
+        min_area = min(area)), by = .(umbral, date, exp, ini_date)] %>% 
+  ggplot(aes(date, mean_area)) +
+  geom_line(data = imerg[umbral %in% c(1, 5, 10)], 
+            aes(end_date, area, color = "OBS"),
+            linetype = 1) +
+  geom_line(aes(color = exp, linetype = factor(ini_date))) +
+  geom_ribbon(aes(ymin = min_area, ymax = max_area, fill = exp, group = interaction(exp, ini_date)), alpha = .1) +
+  scale_color_manual(values = c(OBS = "grey20", colores_exp),
+                     labels =  c(E2 = "CONV", E5 = "ASWS", E6 = "SATWND",
+                                E9 = "RAD", OBS = "OBS")) +
+    scale_fill_manual(values = c(OBS = "grey20", colores_exp),
+                 guide = NULL,
+                     labels = c(E2 = "CONV", E5 = "ASWS", E6 = "SATWND",
+                                E9 = "RAD", OBS = "OBS")) +
+  scale_y_continuous(label = scales::percent_format()) +
+  scale_x_datetime(limits = c(ymd_h(2018112200), ymd_h(2018112312))) +
+  facet_wrap(~umbral, scales = "free_y", 
+             labeller = labeller(umbral = c("1" = "1 mm", "5" = "5 mm", 
+                                       "10" = "10 mm", "25" = "25 mm"))) +
+  labs(y = "Areal coverage", x = "Date", 
+       fill = NULL, color = NULL, linetype = "FCST init") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
 ### Soundings
@@
 ```{r message=FALSE, warning=FALSE}
-files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
-files <- files[str_detect(files, "_01_")]
+# files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E[2,5,6,8]/FCST/*/sondeos/*")
+# files <- files[str_detect(files, "_01_")]
+# 
+# sondeos <- purrr::map(files, function(f){
+# 
+#   fread(f) %>%
+#     .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>%
+#     .[, launch_time := as_datetime(launch_time)] %>%
+#     .[, fcst_exp := basename(dirname(dirname(f)))]
+# 
+# }) %>%
+#   rbindlist()
 
-sondeos <- purrr::map(files, function(f){
-
-  fread(f) %>% 
-    .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
-    .[, launch_time := as_datetime(launch_time)] %>% 
-    .[, fcst_exp := basename(dirname(dirname(f)))]
-
-}) %>% 
-  rbindlist()
-
-miembros <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_agregados.rds") %>% 
+miembros <- read_rds("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos_agregados.rds") %>% 
   .[, mem := formatC(mem, width = 3, flag = "0")] %>% 
   fisica[., on = "mem"] %>% 
-  separate(fisica, into = c("cumulus", "pbl"))
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  setDT() %>% 
+  .[, BIAS := -BIAS]
 
-media <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_media.rds") %>% 
-  .[, .(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
-        BIAS = mean(value - fcst_value, na.rm = TRUE)), 
+media <- read_rds("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos_media.rds") %>% 
+  .[, .(RMSE = sqrt(mean((fcst_value - value)^2, na.rm = TRUE)), 
+        BIAS = mean(fcst_value - value, na.rm = TRUE)), 
     by = .(iop, exp, variable, lev, fcst_exp)] %>% 
@@
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  coord_flip(xlim = c(0, 15)) +
   facet_grid(fcst_exp + iop~variable, scales = "free_x", 
@@
                   ymax = max_rmse, fill = exp), alpha = 0.2) +
-  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                                   E6 = "SATWND", E7 = "RAD")) +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E8 = "RAD")) +
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
-  coord_flip() +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  coord_flip(xlim = c(0,  15)) +
   facet_grid(fcst_exp + iop~variable, scales = "free_x", 
@@
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   coord_flip() +
@@
 ```{r rmse-sondeos-param}
-pbl <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, lev, variable)]
+pbl <- miembros[, .(RMSE = mean(RMSE),
+                    tipo = "pbl"), 
+                by = .(exp, fcst_exp, pbl, lev, variable, iop)] %>% 
+  setnames("pbl", "param")
 
@@
   .[estadistico == "RMSE"] %>% 
-  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
-  .[, ":="(pbl = "Mean",
-           RMSE = value.est)] %>% 
+  .[, c("iop", "exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
+  .[, ":="(param = "Mean",
+           RMSE = value.est, 
+           tipo = "Mean")] %>% 
   .[, value.est := NULL] %>% 
@@
   # geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(y = RMSE, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
+  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
   scale_linetype_manual(values = c(1, 2, 3, 4)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   coord_flip() +
-  facet_grid(fcst_exp ~ variable, scales = "free_x", 
+  facet_grid(factor(fcst_exp) +iop~variable, scales = "free_x", 
              labeller = labeller(variable = c("t" = "Temperature",
@@
 
-cum <- miembros[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, lev, variable)]
+cum <- miembros[, .(RMSE = mean(RMSE),
+                    tipo = "cumulus"), 
+                by = .(exp, fcst_exp, cumulus, lev, variable, iop)] %>% 
+  setnames("cumulus", "param")
 
@@
   .[estadistico == "RMSE"] %>% 
-  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
-  .[, ":="(cumulus = "Mean",
-           RMSE = value.est)] %>% 
+  .[, c("iop", "exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
+  .[, ":="(param = "Mean",
+           RMSE = value.est,
+           tipo = "Mean")] %>% 
   .[, value.est := NULL] %>% 
@@
   # geom_hline(yintercept = 0, color = "grey30") +
-  geom_line(aes(y = RMSE, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
+  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
   scale_linetype_manual(values = c(4, 2, 3, 1)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   coord_flip() +
-  facet_grid(fcst_exp ~ variable, scales = "free_x", 
+  facet_grid(factor(fcst_exp) + iop~variable, scales = "free_x", 
              labeller = labeller(variable = c("t" = "Temperature",
@@
 
+rbind(pbl, cum) %>% 
+  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
+    by = .(variable, exp, param, fcst_exp)] %>% 
+  ggplot(aes(RMSE_mean)) +
+  geom_histogram(aes(fill = exp), 
+                 position = "identity", 
+                 alpha = 0.5) +
+   scale_fill_manual(values = colores_exp, 
+                     labels = c(E2 = "CONV", E5 = "AUT",
+                                E6 = "SATWND", E8 = "RAD")) +
+  facet_grid(fcst_exp ~ variable, scales = "free_x")
+
+rbind(pbl, cum) %>%
+  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
+    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(RMSE_mean, exp)) +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  facet_grid(fcst_exp ~variable, scales = "free_x") +
+  labs(y = NULL, x = "mean RMSE", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
+
 ```
 
-```{r bias-sondeos-media}
+```{r bias-sondeos-ensamble}
 extremos <- miembros %>% 
@@
   geom_hline(yintercept = 0, color = "grey30") +
-  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                                   E6 = "SATWND", E7 = "RAD")) +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E8 = "RAD")) +
   geom_line(aes(color = exp, linetype = estadistico)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   coord_flip() +
@@
   geom_hline(yintercept = 0, color = "grey30") +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   coord_flip() +
@@
   ggplot(aes(lev/1000, BIAS)) +
-  geom_line(aes(y = BIAS, color = exp, linetype = pbl, group = interaction(exp, pbl))) +
+  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
   geom_hline(yintercept = 0, color = "grey30") +
   scale_linetype_manual(values = c(1, 2, 3, 4)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   coord_flip() +
@@
   ggplot(aes(lev/1000, BIAS)) +
-  geom_line(aes(y = BIAS, color = exp, linetype = cumulus, group = interaction(exp, cumulus))) +
+  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
   geom_hline(yintercept = 0, color = "grey30") +
   scale_linetype_manual(values = c(4, 2, 3, 1)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   coord_flip() +
@@
   theme_minimal()
+
+rbind(pbl, cum) %>%
+  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
+    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(BIAS_mean, exp)) +
+  geom_vline(xintercept = 0, color = "grey70") +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  facet_grid(fcst_exp ~ variable, scales = "free_x") +
+  labs(y = NULL, x = "mean BIAS", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
 
@@
 #     .[, exp := NULL] %>%
-#     melt(measure.vars = c("E4", "E5", "E6", "E7"),
+#     melt(measure.vars = c("E2", "E5", "E6", "E8"),
 #          variable.name = "exp",
@@
 
-obs_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean.rds"))
-obs_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean.rds"))
+media <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean_new.rds")),
+               read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean_new.rds"))) %>% 
+  .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
+           BIAS = fifelse(var == "q", BIAS*1000, BIAS))] %>% 
+  .[, BIAS := -BIAS]
 
-obs_ens_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds"))
-obs_ens_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds"))
+miembros <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem_new.rds")),
+                  read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem_new.rds"))) %>% 
+    .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
+           BIAS = fifelse(var == "q", BIAS*1000, BIAS))] %>% 
+  .[, BIAS := -BIAS]
 
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
+media %>% 
+  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico") %>% 
+  .[estadistico == "RMSE"] %>% 
   ggplot(aes(date, value)) +
   geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   scale_date() +
@@
   geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   scale_date() +
@@
 
-miembros <- rbind(obs_ens_00, obs_ens_06)
-
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
+media %>% 
   extremos[., on = .NATURAL] %>% 
-  ggplot(aes(date, value)) +
+  .[exp == "E8"] %>% 
+  ggplot(aes(date, RMSE)) +
   geom_ribbon(aes(ymin = min_rmse, 
-                  ymax = max_rmse, fill = exp), alpha = 0.2) +
-  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                                   E6 = "SATWND", E7 = "RAD")) +
+                  ymax = max_rmse, fill = exp, group = interaction(exp, fcst_exp)), alpha = 0.2) +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E8 = "RAD")) +
   geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   scale_date() +
@@
 
-rbind(obs_00, obs_06) %>% 
-  .[usage.flag == 1] %>% 
-  .[, .(rmse = sqrt(mean((obs - fcst)^2, na.rm = TRUE)),
-        bias = mean(obs -  fcst, na.rm = TRUE)), by = .(var, date, exp, fcst_exp)] %>% 
-  melt(measure.vars = c("rmse", "bias"), variable.name = "estadistico") %>% 
-  .[estadistico == "rmse"] %>% 
-  miembros[., on = .NATURAL] %>% 
-  ggplot(aes(date, value)) +
-  geom_line(alpha = 0.1, 
+media %>% 
+  .[, mem := "000"] %>% 
+  rbind(., miembros) %>% 
+  .[exp == "E8"] %>% 
+  ggplot(aes(date, RMSE)) +
+  geom_line(data = ~.x[mem != "000"], alpha = 0.1, 
             aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
-  geom_line(aes(color = exp, linetype = fcst_exp)) +
-  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  geom_line(data = ~.x[mem == "000"], aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   scale_date() +
@@
   geom_line(aes(y = mean_RMSE, color = exp, linetype = fcst_exp)) +
-   scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
-                                                      E6 = "SATWND", E7 = "RAD")) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
   scale_date() +
@@
 ```
+```{r}
+
+extremos <- miembros %>% 
+  .[, .(max_bias = max(BIAS),
+        min_bias = min(BIAS)), by = .(var, exp, fcst_exp, date)]
+
+media %>% 
+  extremos[., on = .NATURAL] %>% 
+  .[exp == "E8"] %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_ribbon(aes(ymin = min_bias, 
+                  ymax = max_bias, fill = exp, group = interaction(exp, fcst_exp)), alpha = 0.2) +
+  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                                   E6 = "SATWND", E8 = "RAD")) +
+  geom_line(aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+
+media %>% 
+  .[, mem := "000"] %>% 
+  rbind(., miembros) %>% 
+  .[exp == "E8"] %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_line(data = ~.x[mem != "000"], alpha = 0.1, 
+            aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
+  geom_line(data = ~.x[mem == "000"], aes(color = exp, linetype = fcst_exp)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_wrap(~ var, scales = "free_y") +
+  theme_minimal()
+
+```
+
+```{r}
+cum <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
+   setnames("cumulus", "param") %>% 
+  .[, tipo := "cumulus"]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
+  rbind(., cum) %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_line(aes(color = exp, linetype = param)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(4, 2, 3, 1)) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+
+pbl <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
+   setnames("pbl", "param") %>% 
+  .[, tipo := "pbl"]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
+  rbind(., pbl) %>% 
+  ggplot(aes(date, BIAS)) +
+  geom_line(aes(color = exp, linetype = param)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(1, 2, 3, 4)) +
+  scale_date() +
+  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+
+rbind(pbl, cum) %>%
+  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
+    by = .(var, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(BIAS_mean, exp)) +
+  geom_vline(xintercept = 0, color = "grey70") +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  facet_grid(fcst_exp ~ var, scales = "free_x") +
+  labs(y = NULL, x = "mean BIAS", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
+```
+
+```{r}
+cum <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
+   setnames("cumulus", "param") %>% 
+  .[, tipo := "cumulus"]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
+  rbind(., cum) %>% 
+  ggplot(aes(date, RMSE)) +
+  geom_line(aes(color = exp, linetype = param)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(4, 2, 3, 1)) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+
+pbl <- fisica[miembros, on = .NATURAL] %>% 
+  separate(fisica, into = c("cumulus", "pbl")) %>% 
+  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
+   setnames("pbl", "param") %>% 
+  .[, tipo := "pbl"]
+
+media %>% 
+  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
+  .[, ":="(param = "Mean",
+           tipo = "Mean")] %>% 
+  rbind(., pbl) %>% 
+  ggplot(aes(date, RMSE)) +
+  geom_line(aes(color = exp, linetype = param)) +
+  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
+                                                      E6 = "SATWND", E8 = "RAD")) +
+  # geom_hline(yintercept = 0, color = "grey30") +
+  scale_linetype_manual(values = c(1, 2, 3, 4)) +
+  scale_date() +
+  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
+  facet_grid(var~fcst_exp, scales = "free_y") +
+  theme_minimal()
+
+rbind(pbl, cum) %>%
+  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
+    by = .(var, exp, param, fcst_exp, tipo)] %>% 
+  ggplot(aes(RMSE_mean, exp)) +
+  geom_point(aes(color = exp, shape = param),
+             size = 3) +
+  scale_color_manual(guide = NULL, values = colores_exp, 
+                      labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
+  scale_y_discrete(labels = c(E2 = "CONV", E5 = "AUT", 
+                                 E6 = "SATWND", E8 = "RAD")) +
+  facet_grid(fcst_exp ~ var, scales = "free_x") +
+  labs(y = NULL, x = "mean RMSE", shape = NULL) +
+  theme_minimal() +
+  theme(aspect.ratio = 1/2,
+        panel.background = element_rect(fill = "grey98",
+                                        color = "grey98"))
+```
 
@@
 ```{r}
-files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_mean_acum_30h.rds"))
+files <- Sys.glob(paste0(derived_data, "/ppacum/E[2,5,6,8]_fcst_201811220*_mean_acum_30h.rds"))
 
@@
   geom_mapa() +
-  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                          E6 = "SATWND", E7 = "RAD"))) +
+  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
   labs(fill = NULL) +
@@
 ```{r}
-files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_prop_acum_30h.rds"))
+files <- Sys.glob(paste0(derived_data, "/ppacum/E[2,5,6,8]_fcst_201811220*_prop_acum_30h.rds"))
 
@@
   geom_mapa() +
-  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                          E6 = "SATWND", E7 = "RAD"))) +
+  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
   labs(fill = "100 mm") +
@@
     geom_mapa() +
-    facet_wrap(~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
-                                                 E6 = "SATWND", E7 = "RAD"))) +
+    facet_wrap(~exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                 E6 = "SATWND", E8 = "RAD"))) +
     labs(caption = crtm[,date[1]]) +
@@
 ```
+
+### Campos de variables
+
+```{r}
+df <- data.frame(
+  lead_time = rep(c(6, 12, 18), 4),
+  exp = rep(c("E2", "E5", "E6", "E8"), each = 3)
+)
+
+files <- purrr::map2_chr(df$lead_time, df$exp, function(l, e) {
+  paste0("/home/paola.corrales/datosmunin3/EXP/", e, "/FCST/2018112200/NPP/NPP_2018-11-22_00_FC", 
+         formatC(l, width = 2, flag = "0"), ".nc")
+})
+
+coord <- ReadNetCDF(files[1], vars = c("XLAT", "XLONG"))
+
+var_00 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("Q", "T", "V"), subset = list(lev = c(800, 1000))) %>% 
+    .[, .(Q = mean(Q),
+          T = mean(T),
+          V = mean(V)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+
+mcape_00 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("MCAPE", "MDBZ")) %>% 
+    .[, .(MCAPE = mean(MCAPE, na.rm = TRUE),
+          MDBZ = mean(MDBZ, na.rm = TRUE)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+```
+
+```{r}
+df <- data.frame(
+  lead_time = rep(c(0, 6, 12), 4),
+  exp = rep(c("E2", "E5", "E6", "E8"), each = 3)
+)
+
+files <- purrr::map2_chr(df$lead_time, df$exp, function(l, e) {
+  paste0("/home/paola.corrales/datosmunin3/EXP/", e, "/FCST/2018112206/NPP/NPP_2018-11-22_06_FC", 
+         formatC(l, width = 2, flag = "0"), ".nc")
+})
+
+coord <- ReadNetCDF(files[1], vars = c("XLAT", "XLONG"))
+
+var_06 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("Q", "T", "V"), subset = list(lev = c(800, 1000))) %>% 
+    .[, .(Q = mean(Q),
+          T = mean(T),
+          V = mean(V)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+
+mcape_06 <- purrr::map(files, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/FCST/{fcst_exp}/NPP/NPP_{x}_FC{lead}.nc")
+
+  ReadNetCDF(f, vars = c("MCAPE", "MDBZ")) %>% 
+    .[, .(MCAPE = mean(MCAPE, na.rm = TRUE),
+          MDBZ = mean(MDBZ, na.rm = TRUE)), by = .(lat, lon)] %>% 
+    .[, ":="(exp = meta[[1]][["exp"]],
+             fcst_exp = meta[[1]][["fcst_exp"]],
+             lead_date = ymd_h(meta[[1]][["fcst_exp"]]) + hours(as.numeric(meta[[1]][["lead"]])))]
+
+}) %>% rbindlist()
+```
+
+```{r}
+var_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = Q, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 20, 2)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       limits = c(1, 20),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+var_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = Q, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 20, 2)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "BrBG", direction = 1,
+                       limits = c(1, 20),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+```{r}
+var_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = V, fill = stat(level_d)), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(name = NULL,
+                                   guide = guide_colorsteps(#inside = FALSE,
+                                     barwidth = 25,
+                                     barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+var_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = V, fill = stat(level_d)), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised(name = NULL,
+                                   guide = guide_colorsteps(#inside = FALSE,
+                                     barwidth = 25,
+                                     barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+```
+
+```{r}
+var_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = T - 273.15, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 40, 5)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "OrRd", direction = 1,
+                       limits = c(1, 40),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+var_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = T - 273.15, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 40, 5)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "OrRd", direction = 1,
+                       limits = c(1, 40),
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+```{r}
+mcape_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MCAPE, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 4500, 500)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+mcape_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MCAPE, fill = stat(level_d)), proj = norargentina_lambert,
+                    breaks = seq(0, 4500, 500)) +
+  scale_fill_distiller(name = NULL, super = ScaleDiscretised,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(#inside = FALSE,
+                         barwidth = 25,
+                         barheight = 0.8)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```
+```{r}
+values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
+22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
+42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
+61.5, 63, 64.5, 66, 69, 76.5)
+
+colours <- c("#3D4170", "#3D4E8C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
+"#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
+"#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
+"#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
+"#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
+"#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
+"#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
+"#95E3C5", "#85E0BD")
+
+mcape_00 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MDBZ, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    drop = FALSE, guide = guide_colorsteps(barwidth = 0.8, 
+                                                                         barheight = 30)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+
+mcape_06 %>% 
+  coord[., on = .NATURAL] %>% 
+  .[, c("x", "y") := wrf_project(XLONG, XLAT)] %>% 
+  .[, lead_date := factor(lead_date)] %>%
+
+  ggplot(aes(x, y)) +
+  # geom_point(aes(color = MCAPE)) +
+  geom_contour_fill(aes(z = MDBZ, fill = stat(level)), 
+                    proj = norargentina_lambert,
+                    breaks = values) +
+  scale_fill_manual(values = colours, 
+                    drop = FALSE, guide = guide_colorsteps(barwidth = 0.8, 
+                                                                         barheight = 30)) +
+  geom_mapa() +
+  cordillera +
+  facet_grid(lead_date ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
+                                                          E6 = "SATWND", E8 = "RAD"))) +
+  theme_minimal() +
+  theme(panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3))
+```

### 2022-05-03 analysis/chapters/DA_goes16.Rmd
more organization

@@
 library(kableExtra)
-source(here::here("temp_R", "goes_functions.R"))
+# source(here::here("temp_R", "goes_functions.R"))
 
@@
 ```{r}
-nc <- ncdf4::nc_open('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183250000341_e20183250011108_c20183250011312.nc')
+  nc <- ncdf4::nc_open('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183240000342_e20183240011109_c20183240011270.nc')
 
-cld_msk <- ReadNetCDF('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183250000341_e20183250011108_c20183250011312.nc', 
+cld_msk <- ReadNetCDF('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183240000342_e20183240011109_c20183240011270.nc', 
                       vars = 'BCM', 
@@
   labs(x = NULL, y = NULL,
-       title = "Cloud Mask", subtitle = "Time: 2018-11-21 00 UTC")
+       title = "Cloud Mask", subtitle = "Time: 2018-11-20 00 UTC")
 
@@
 ```{r}
-files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")
+files <- Sys.glob("~/datosmunin3/EXP/E10/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")
 
@@
 
-  # meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000/diagfiles/asim_abi_g16_20181121000000.ensmean")
-
-  read_diag_rad(f, "EG1") %>% 
-    # .[, run := meta[[1]][["exp"]]] %>% 
+  read_diag_rad(f, "E10") %>% 
     .[, channel := channel + 6]
@@
 
+files <- Sys.glob("~/datosmunin3/EXP/EG2/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")
+
+diag_eg2 <- map(files, function(f) {
+
+  read_diag_rad(f, "EG2") %>% 
+    .[, channel := channel + 6]
+
+}) %>% 
+  rbindlist()
+
+```
+
+```{r}
+rbind(diag, diag_eg2) %>% 
+  .[channel %in% c(8:10), .N, by = .(qc, date, channel, exp)] %>% 
+  ggplot(aes(date, N)) +
+  geom_point(aes(color = factor(qc))) +
+  geom_line(aes(color = factor(qc))) +
+  # scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
+  #                    labels = c("0" = "pass qc", 
+  #                               "3" = "gross check",
+  #                               "5" = "surface check",
+  #                               "50" = "std dev")) +
+  labs(x = NULL, y = NULL, color = NULL) +
+  facet_grid(exp~channel,  scales = "free_y",
+             labeller = labeller(channel = channels)) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
 ```
@@
   facet_wrap(~date, ncol = 7) +
-  labs(fill = NULL, x = "Channel") +
+  labs(fill = NULL, x = "Channel", subtitle = "EG1") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
+```{r}
+
+diag_eg2 %>% 
+  .[channel %in% c(8:10)] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  ggplot(aes(factor(channel))) +
+  geom_bar(aes(fill = factor(qc))) +
+  scale_fill_manual(values = c("darkorange","grey40", "grey60", "grey80"),
+                    labels = c("0" = "pass qc", 
+                               "3" = "gross check",
+                               "5" = "surface check",
+                               "50" = "std dev")) +
+  facet_wrap(~date, ncol = 7) +
+  labs(fill = NULL, x = "Channel", subtitle = "EG2") +
   theme_minimal() +
@@
 
+```{r}
+diag_eg2 %>% 
+  .[channel %in% c(8:10)] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = factor(qc)), size = 0.9) +
+  geom_mapa() +
+  scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
+                     labels = c("0" = "pass qc", 
+                                "3" = "gross check",
+                                "5" = "surface check",
+                                "50" = "std dev")) +
+  guides(color = guide_legend(override.aes = list(size = 2))) +
+  facet_grid(channel~date) +
+  labs(subtitle = "Assimilated obs (EG2)", x =  NULL, y = NULL, color = "QC") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
+
 Looks like there are less observations rejected due to the standar deviation threshold on the last cycles. 00 UTC is 21 PM in Argentina, so it may be asociated to a diurnal cycle?
@@
   labs(subtitle = "Channel 10", x =  NULL, y = NULL, color = "QC") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+diag_eg2 %>% 
+  .[channel %in% c(10)] %>% 
+  .[, date := format(date, "%m/%d %H UTC")] %>%
+  ggplot(aes(ConvertLongitude(lon), lat)) +
+  geom_point(aes(color = factor(qc)), size = 0.9) +
+  geom_mapa() +
+  scale_color_manual(values = c("grey50", "grey70", "grey80", "cyan4"),
+                     labels = c("0" = "pass qc", 
+                                "3" = "gross check",
+                                "5" = "surface check",
+                                "50" = "std dev")) +
+  guides(color = guide_legend(override.aes = list(size = 2))) +
+  facet_wrap(~date, ncol = 4) +
+  labs(subtitle = "Channel 10 (EG2)", x =  NULL, y = NULL, color = "QC") +
   theme_minimal() +
@@
 ```{r}
-files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/analysis.mem001")
+files_ana <- Sys.glob("~/datosmunin3/EXP/E9/ANA/*00/analysis.mem001")[c(1:7)]
 
@@
 
-  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/{date}/analysis.mem001")
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/E6/ANA/{date}/analysis.mem001")
 
@@
 
-files_gue <- Sys.glob("~/datosmunin3/EXP/prueba_abi/GUESS/*00/wrfarw.mem001")
+files_gue <- Sys.glob("~/datosmunin3/EXP/EG2/GUESS/*00/wrfarw.mem001")
 
@@
 
-  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/GUESS/{date}/wrfarw.mem001")
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/EG2/GUESS/{date}/wrfarw.mem001")
 
@@
 
+files_conv <- Sys.glob("~/datosmunin3/EXP/E2/ANA/*00/analysis.mem001")[1:7]
+
+ana_conv <- map(files_conv, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/E2/ANA/{date}/analysis.mem001")
+
+  ReadNetCDF(f, vars = c(t_conv = "T", p = "P", "PB", q_conv = "QVAPOR",
+                         lon = "XLONG", lat = "XLAT")) %>% 
+    .[, ":="(t_conv = tk(t_conv, p + PB, T_BASE = 290) - 273.15,
+             p = p + PB,
+             PB = NULL)] %>% 
+    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
+    .[, .(lev = levs,
+          t_conv = approx(p, t_conv, xout = levs*100)$y,
+          q_conv = approx(p, q_conv, xout = levs*100)$y),
+      by = .(date, south_north, west_east)] %>%
+    coord[., on = .NATURAL] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+    .[]
+
+}) %>% 
+  rbindlist()
+
+files_conv <- Sys.glob("~/datosmunin3/EXP/*/ANA/20181121000000/analysis.ensmean")[c(1:5,7:8)]
+
+ana_conv <- map(files_conv, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{run}/ANA/20181121000000/analysis.ensmean")
+
+  ReadNetCDF(f, vars = c(t_conv = "T", p = "P", "PB", q_conv = "QVAPOR",
+                         lon = "XLONG", lat = "XLAT")) %>% 
+    .[, ":="(t_conv = tk(t_conv, p + PB, T_BASE = 290) - 273.15,
+             p = p + PB,
+             PB = NULL)] %>% 
+    .[, date := meta[[1]][["run"]]] %>% 
+    .[, .(lev = levs,
+          t_conv = approx(p, t_conv, xout = levs*100)$y,
+          q_conv = approx(p, q_conv, xout = levs*100)$y),
+      by = .(date, south_north, west_east)] %>%
+    coord[., on = .NATURAL] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+    .[]
+
+}) %>% 
+  rbindlist()
+```
+
+```{r}
+temp <- ana_conv %>% 
+  dcast(x + y + lev ~ date, value.var = "q_conv") %>% 
+  .[, ":="(aut = E5 - E2,
+           satwnd = E6 - E5,
+           rad_bad = E7 - E6,
+           rad = E8 - E6,
+           # abi_old = prueba_abi - E2,
+           abi = EG2 - E6,
+           abi_sat = EG2 - E5,
+           abi_sat_aut = EG2 - E2)] %>% 
+  melt(measure.vars = c("aut", "satwnd", "rad_bad", "rad", "abi", "abi_sat", "abi_sat_aut"))
+
+temp2 <- ana_conv %>% 
+  dcast(x + y + lev ~ date, value.var = "t_conv") %>% 
+  .[, ":="(aut = E5 - E2,
+           satwnd = E6 - E5,
+           rad_bad = E7 - E6,
+           rad = E8 - E6,
+           # abi_old = prueba_abi - E2,
+           abi = EG2 - E6,
+           abi_sat = EG2 - E5,
+           abi_sat_aut = EG2 - E2)] %>% 
+  melt(measure.vars = c("aut", "satwnd", "rad_bad", "rad", "abi", "abi_sat", "abi_sat_aut"))
+
+for (i in 1:length(levs)) {
+  temp %>% 
+    .[lev == levs[i]] %>% 
+    ggplot(aes(x, y)) +
+    geom_contour_fill(aes(z = value*1000, fill = ..level..), proj = norargentina_lambert) +
+    scale_fill_divergent_discretised(trans = mesoda::symlog_trans(),
+                                     guide = guide_colorsteps(barhight = 10)) +
+    geom_mapa() +
+    labs(x = NULL, y = NULL, fill = NULL,
+         title = "Humidity",
+         subtitle = as.character(levs[i])) +
+    facet_wrap(~variable, ncol = 4) +
+    theme_minimal()
+
+   ggsave(paste0("impact_2018112100_q_", levs[i], ".png"), height = 4)
+
+  temp2 %>% 
+    .[lev == levs[i]] %>% 
+    ggplot(aes(x, y)) +
+    geom_contour_fill(aes(z = value, fill = ..level..), proj = norargentina_lambert) +
+    scale_fill_divergent_discretised(trans = mesoda::symlog_trans(),
+                                     guide = guide_colorsteps(barhight = 10)) +
+    geom_mapa() +
+    labs(x = NULL, y = NULL, fill = NULL,
+         title = "Temperature",
+         subtitle = as.character(levs[i])) +
+    facet_wrap(~variable, ncol = 4) +
+    theme_minimal()
+
+   ggsave(paste0("impact_2018112100t_", levs[i], ".png"), height = 4)
+
+}
 ```
@@
 
-  ggsave(paste0("t_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)
+  ggsave(paste0("EG2_t_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)
 
@@
 
-  ggsave(paste0("q_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)
+  ggsave(paste0("EG2_q_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)
 
@@
 
-  ggsave(paste0("t_ab_", levs[i], ".png"), height = 4)
+  ggsave(paste0("EG2_t_ab_", levs[i], ".png"), height = 4)
 
@@
 
-  ggsave(paste0("q_ab_", levs[i], ".png"), height = 4)
+  ggsave(paste0("EG2_q_ab_", levs[i], ".png"), height = 4)
 
@@
 ```
+
+```{r}
+for (i in 1:length(levs)) {
+
+  ana[ana_conv, on = .NATURAL] %>% 
+    .[lev == levs[i]] %>% 
+    .[, date := format(date, "%m/%d %H UTC")] %>%
+    .[, a_g := t_ana - t_conv] %>%
+    ggplot(aes(x, y)) +
+    geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
+    scale_fill_divergent_discretised() +
+    geom_mapa() +
+    labs(x = NULL, y = NULL, fill = "SATWND-CONV",
+         title = "Temperature",
+         subtitle = paste(as.character(levs[i]), "hPa")) +
+    facet_wrap(~date, ncol = 4) +
+    theme_minimal()
+
+  ggsave(paste0("E62_t_ac_", levs[i], ".png"), height = 4)
+
+  ana[ana_conv, on = .NATURAL] %>% 
+    .[lev == levs[i]] %>% 
+    .[, date := format(date, "%m/%d %H UTC")] %>%
+    .[, a_g := q_ana - q_conv] %>%
+    ggplot(aes(x, y)) +
+    geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
+    scale_fill_divergent_discretised() +
+    geom_mapa() +
+    labs(x = NULL, y = NULL, fill = "SATWND-CONV",
+         title = "Humidity",
+         subtitle = paste(as.character(levs[i]), "hPa")) +
+    facet_wrap(~date, ncol = 4) +
+    theme_minimal()
+
+  ggsave(paste0("E62_q_ac_", levs[i], ".png"), height = 4)
+
+}
+```
+```{r}
+files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*_*/analysis.mem001")[c(1,3:5)]
+
+ana <- map(files_ana, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000_{run}/analysis.mem001")
+
+  ReadNetCDF(f, vars = c(t_ana = "T", p = "P", "PB", q_ana = "QVAPOR",
+                         lon = "XLONG", lat = "XLAT")) %>% 
+    .[, ":="(t_ana = tk(t_ana, p + PB, T_BASE = 290) - 273.15,
+             p = p + PB,
+             PB = NULL)] %>% 
+    .[, run := meta[[1]][["run"]]] %>% 
+    .[, .(lev = levs,
+          t_ana = approx(p, t_ana, xout = levs*100)$y,
+          q_ana = approx(p, q_ana, xout = levs*100)$y),
+      by = .(run, south_north, west_east)] %>%
+    coord[., on = .NATURAL] %>% 
+    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
+    .[]
+
+}) %>% 
+  rbindlist()
+
+ana %>% 
+  .[lev == levs[i]] %>% 
+  # .[, date := format(date, "%m/%d %H UTC")] %>%
+  # .[, a_g := q_ana - q_conv] %>%
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = q_ana*1000, fill = ..level..), proj = norargentina_lambert) +
+  # scale_fill_divergent_discretised() +
+  geom_mapa() +
+  labs(x = NULL, y = NULL, fill = NULL,
+       title = "Humidity",
+       subtitle = paste(as.character(levs[i]), "hPa")) +
+  facet_wrap(~run, ncol = 4) +
+  theme_minimal()
+
+ana[run != "all"] %>% 
+  dcast(x + y + lev ~ run, value.var = "q_ana") %>% 
+  .[ana[run == "all"], on = .NATURAL] %>% 
+  .[, ":="(all_8 = t_ana - ch08,
+           all_9 = t_ana - ch09,
+           all_0 = t_ana - ch10)] %>% 
+  melt(measure.vars = c("all_8", "all_9", "all_0")) %>% 
+  .[lev == levs[4]] %>% 
+  ggplot(aes(x, y)) +
+  geom_contour_fill(aes(z = value, fill = ..level..), proj = norargentina_lambert) +
+  scale_fill_divergent_discretised() +
+  geom_mapa() +
+  labs(x = NULL, y = NULL, fill = NULL,
+       title = "Humidity",
+       subtitle = paste(as.character(levs[i]), "hPa")) +
+  facet_wrap(~variable, ncol = 4) +
+  theme_minimal()
+
+```
+
 ```{r}

### 2022-05-03 analysis/chapters/goes.Rmd
more organization

---
title: "Deterministic hires forecasts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(tagger)
library(knitr)
library(kableExtra)
#source(here::here("temp_R", "goes_functions.R"))

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function() {
  list(geom_sf(data = map, fill = NA, color = "black", size = 0.1, inherit.aes = FALSE),
       coord_sf(xlim = c(-76, -52), ylim = c(-41, -20)))
}

#dominio de interés
xlim <- c(2600, 3950)
ylim <- c(3700, 4750)

lon_band <- c(-66.5, -61.5)
lat_band <- c(-35.5, -29)

scale_date <- function(ini = 20181120180000, end = 20181123120000,
                       break_bin = "6 hour") {
  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
                   labels = function(x) {
                     fifelse(hour(x) == 0, format(x, "%H\n%b %d"), format(x, "%H"))
                   }) 
}

derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"

colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", 
                 E9 = "#CC3311", E10 = "chartreuse4")

channels <- c("8" = "Ch 8 - upper level", 
              "9" = "Ch 9 - middle level", 
              "10" = "Ch 10 - lower level")

satinfo <- fread("~/mesoda/analysis/data/derived_data/satinfo.txt") %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
```

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_1h_fcst_det_201811220*_E*_d02.csv")

fss_fcst <- purrr::map(files, function(f) {
  meta <- unglue(f, "/home/paola.corrales/mesoda/analysis/data/derived_data/fss_1h_fcst_det_{init}_{exp}_d02.csv")

  fread(f) %>% 
    .[, date := as_datetime(date)] %>% 
    .[, run := meta[[1]][["init"]]]
}) %>% 
  rbindlist()

# 
# files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E*[1-9]*.csv")
# 
# fss_ana <- purrr::map(files, function(f) {
#   
#   fread(f) %>% 
#     .[, date := as_datetime(date)] %>% 
#     .[, run := "ana"]
# }) %>% 
#   rbindlist()

fss_fcst %>% 
  .[w %in% c(1, 5, 11) & q %in% c(1, 5, 25)] %>% 
  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = run)) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = c(colores_exp, EG3 = "blue"), 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", EG3 = "only ABI",
                                E9 = "RAD", E10 = "ABI")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "2 km", "5" = "10 km", "11" = "20 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
  labs(color = NULL, x = NULL, y = "FSS", linetype = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")

fss_fcst %>% 
  .[w %in% c(1, 5, 11) & q %in% c(1, 5, 25)] %>% 
  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = run)) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = c(colores_exp, EG3 = "blue"), 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", EG3 = "only ABI",
                                E9 = "RAD", E10 = "ABI")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123100000))) +
  facet_grid(w + q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "2 km", "5" = "10 km", "11" = "20 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
  labs(color = NULL, x = NULL, y = "FSS", linetype = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

### 2022-05-03 analysis/chapters/poster2.Rmd
more organization

---
title: "Poster 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function(fill = NA) {
  list(geom_sf(data = map, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       # geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       coord_sf(ylim = c(-42, -19), xlim = c(-76, -50)),
       scale_x_longitude(ticks = 5),
       scale_y_latitude(ticks = 5))
}

arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10, 
                                file.dir = "~/mesoda/analysis/data/derived_data/")
arg_topo[, lon := metR::ConvertLongitude(lon)]

cordillera <- list(
  ggnewscale::new_scale_fill(),
  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat,
                                         z = h),
                    breaks = seq(1500, 8000, by = 700)),
  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
                      oob = scales::squish))

scale_date <- function(ini = 20181120180000, end = 20181123120000,
                       break_bin = "6 hour") {
  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
                   labels = function(x) {
                     fifelse(hour(x) == 0, format(x, "%H UTC\n%b %d"), format(x, "%H"))
                   }) 
}

experiments <- c("E4", "E5", "E6", "E7")
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311")

satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
```

## Analysis

```{r summary-fields}

files <- Sys.glob(paste0(derived_data, "pw/pw_2*"))

dates <- c("20181122000000", "20181122060000")

pw <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "pw_{date}_{exp}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "pw", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      .[]
  }
}) %>% 
  rbindlist() %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

files <- c(Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122000000.nc")),
           Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122060000.nc")))

mucape <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
      .[]
  }
}) %>% 
  rbindlist()

fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
                     exp = c("E2", "E5", "E6", "E8"))

ana <- lapply(seq_len(nrow(fcsts)), function(f) {

  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>%     .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
    .[, exp := fcsts[f, 2]] %>%  
    .[]
}) %>% 
  rbindlist()

cor <- ana[bottom_top %in% c(1, 16)] %>% 
  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
  .[, cortante := Mag(u_16 - u_1, v_16 - v_1)] %>% 
  .[, ":="(u_1 = NULL, 
           u_16 = NULL,
           v_1 = NULL,
           v_16 = NULL)]

ana[bottom_top %between% c(1, 7), .(v_mean = mean(v)), 
    by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)),
                    proj = norargentina_lambert) +
  colorspace::scale_fill_continuous_divergingx(super = ScaleDiscretised,
                                               palette = "PRGn", mid = 25,
                                               guide = guide_colorsteps(barwidth = 25,
                                                                        barheight = 0.5, 
                                                                        title.position = "left", 
                                                                        title.vjust = 1,
                                                                        frame.colour = "black")) +
  geom_contour2(aes(z = v_mean, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(-10, -5)) +
  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
                        labels = c("-10", "-5")) +
  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 10),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

ggsave("pw_ana.png", width = 30, height = 20, units = "cm")

mucape %>% 
  .[variable == "mcape"] %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
  .[cor, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)), 
                    proj = norargentina_lambert, limits = c(500, 6000)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       palette = "YlOrRd", direction = 1,
                       labels = function(x) JumpBy(x, 2, fill = ""),
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.5, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_contour2(aes(z = cortante, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(15, 30)) +
  scale_size_continuous(breaks = c(15, 30), range = c(0.8, 0.4),
                        labels = c("15", "30")) +
  labs(fill = latex2exp::TeX("MCAPE ($JKg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($ms^{-1}$)")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 10),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

ggsave("mcape_ana.png", width = 30, height = 20, units = "cm")
```
## Pronóstico

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,8]_ens.csv")

fss <- purrr::map(files, function(f) {

  meta <- unglue(basename(f), "fss_6h_fcst_ens_{ini_date}_{exp}_ens.csv")

  fread(f) %>% 
    .[, date := as_datetime(date)] %>% 
    .[, ini_date := ymd_h(meta[[1]][["ini_date"]])]
}) %>% 
  rbindlist()

fss %>% 
  .[w %in% c(1, 11) & q %in% c(5, 25) & exp %in% c("E2", "E5", "E6", "E8")] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = factor(ini_date))) +
  geom_vline(xintercept = c(as_datetime("2018-11-22 14:00:00"), as_datetime("2018-11-23 06:00:00")), 
             size = 0.5,
             color = "darkgrey") +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  scale_linetype_discrete(labels = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC")) +
  scale_date(20181122000000, 20181123120000, "12 hours") +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "10 km", "5" = "50 km", 
                                            "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL, y = "FSS", linetype = NULL) +
  theme_minimal()

ggsave("fss2.png", width = 15, height = 10, unit = "cm")
```

```{r}

values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
            22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
            42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
            61.5, 63, 64.5, 66, 69, 76.5)

colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
             "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
             "#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
             "#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
             "#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
             "#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
             "#95E3C5", "#85E0BD")

coord <- ReadNetCDF("~/datosmunin3/EXP/E2/FCST/2018112200/NPP/NPP_2018-11-22_00_FC14.nc", vars = c("XLAT", "XLONG"))

# files <- c(Sys.glob("~/datosmunin3/EXP/E[2,5,6,8]/FCST/201811220*/NPP/NPP_2018-11-22_00_FC14.nc"),
#            Sys.glob("~/datosmunin3/EXP/E[2,5,6,8]/FCST/201811220*/NPP/NPP_2018-11-22_06_FC08.nc"))
# 
# pp <- purrr::map_df(files, function(f) {
#   
#   d <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/FCST/{ini_date}/NPP/NPP_2018-11-22_{other}_FC{val_date}.nc")
#   
#   ReadNetCDF(f, vars = "PP") %>% 
#     .[, .(PP = mean(PP)), by = .(lon, lat)] %>%
#     .[, ":="(fecha_ini = d[[1]][["ini_date"]],
#              hora_val = d[[1]][["val_date"]],
#              exp = d[[1]][["exp"]])]
#   
# }) %>% 
#   coord[., on = .NATURAL] %>% 
#   .[, ":="(lon = XLONG, lat = XLAT)] %>% 
#   .[, ":="(XLONG = NULL, XLAT = NULL)] %>% 
#   .[, c("x", "y") := wrf_project(lon, lat)]

files <- Sys.glob("~/datosmunin3/EXP/derived_data/ppacum/E[2,5,6,8]_fcst_201811220[0,6]_mean_acum_1h.rds")

pp <- map_df(files, function(f) {

  d <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/{exp}_fcst_{ini_date}_mean_acum_1h.rds")

  if (d[[1]][["exp"]] %in% c("E2", "E8")) {

    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               lead_time = date)]

  } else { 
    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               date = lead_time)]

  }

}) %>% 
  .[lead_time %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

pp %>%  
  # .[lead_time == as_datetime("20181123060000")] %>%
  .[, lead_time := factor(lead_time)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level)),  breaks = c(seq(1, 45, 5), Inf),
                    proj = norargentina_lambert) +
  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 45, 5)) +
  scale_fill_distiller(palette = "YlGnBu",
                       direction = 1,, 
                       labels = function(x) JumpBy(x, 3, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.4, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  # coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)) +
  # scale_x_longitude(ticks = 5, labels = NULL) +
  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
  facet_grid(lead_time + fecha_ini ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                                                      E6 = "SATWND", E8 = "RAD"),
                                                              fecha_ini = c("2018112200" = "00 UTC", "2018112206" = "06 UTC"))) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

obs_pp <- readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000) & 
      end_date %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

obs_pp %>% 
  .[, end_date := factor(end_date)] %>% 
  # .[end_date == as_datetime("20181122140000")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level)),  breaks = c(seq(1, 45, 5), Inf),
                    proj = norargentina_lambert) +
  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 45, 5)) +
  scale_fill_distiller(palette = "YlGnBu",
                       direction = 1,, 
                       labels = function(x) JumpBy(x, 3, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.4, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  facet_wrap(~end_date) +
  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) 
```

```{r}
files <- Sys.glob("~/datosmunin3/EXP/derived_data/ppacum/E[2,5,6,8]_fcst_201811220[0,6]_prop_acum_1h.rds")

pp <- map_df(files, function(f) {

  d <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/{exp}_fcst_{ini_date}_prop_acum_1h.rds")

  if (d[[1]][["exp"]] %in% c("E2", "E8")) {

    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               lead_time = date)]

  } else { 
    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               exp = d[[1]][["exp"]],
               date = lead_time)]

  }

}) %>% 
  .[lead_time %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

obs_pp <- readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000) & 
      end_date %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

pp %>%  
  # .[lead_time == as_datetime("20181123060000")] %>%
  .[umbral == 5] %>% 
  .[, lead_time := factor(lead_time)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = prop*100, fill = stat(level)),  breaks = seq(10, 100, 10),
                    proj = norargentina_lambert) +
  scale_fill_distiller(palette = "Spectral",
                       direction = -1,
                       labels = function(x) JumpBy(x, 1, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.4,
                                                title.position = "left",
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_contour2(data = obs_pp %>% .[, lead_time := end_date], aes(z = pp_acum), color = "red", size = 0.2, 
                breaks = c(5),
                proj = norargentina_lambert) +
  geom_mapa() +
  # coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)) +
  # scale_x_longitude(ticks = 5, labels = NULL) +
  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Probability of 5 $mmh^{-1}$\n precipitation")) +
  facet_grid(lead_time + fecha_ini ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", E6 = "SATWND", E8 = "RAD"),
                                 fecha_ini = c("2018112200" = "FCST: 00 UTC", "2018112206" = " FCST: 06 UTC"),
                                 lead_time = c("2018-11-22 14:00:00" = "11/22 14:00", "2018-11-23 06:00:00" = "11/23 06:00"))) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

ggsave("pp.png", width = 20, height = 20, units = "cm")
```

### 2022-05-03 analysis/chapters/presentacion_RAD.Rmd
more organization

---
title: "Poster"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function() {
  list(geom_sf(data = map, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE),
       coord_sf(xlim = c(-76, -52), ylim = c(-41, -20)))
}

sensores <- c("amsua_n15",
              "amsua_n18",
              "amsua_n19",
              "amsua_metop-b",
              "amsua_aqua",
              "airs_aqua",
              "iasi_metop-a",
              "iasi_metop-b", 
              "hirs4_n19",
              "hirs4_metop-a",
              "hirs4_metop-b",
              "mhs_metop-a",
              "mhs_metop-b",
              "mhs_n18",
              "mhs_n19",
              "mhs_metop-a",
              "atms_npp",
              "atms_n20",
              "cris-fsr_npp",
              "cris-fsr_n20"
)

multiespectrales <- c("airs_aqua",
                      "iasi_metop-a",
                      "iasi_metop-b")
```

## Coeficientes

```{r}
files <- c(Sys.glob("/home/paola.corrales/datosmunin3/EXP/prueba_BC/ANA/*/satbias_enkf/satbias_2*"),
           Sys.glob("/home/paola.corrales/datosmunin3/EXP/E8/ANA/*/satbias_enkf/satbias_2*"))

satbias <- map(files, function(f){

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{periodo}/ANA/{date_ana}/satbias_enkf/satbias_{date}")

  read_satbias(f) %>% 
    as.data.table() %>% 
    .[, ":="(date = ymd_hms(meta[[1]][["date"]]),
             periodo = meta[[1]][["periodo"]])]

}) %>% 
  rbindlist()

satbias <- satbias[sensor %in% sensores]

satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))

satbias <- satinfo[satbias, on = c("sensor", "channel")]

```

```{r coef-amsua-n15, fig.cap="Coeficientes de corrección de bias en cada ciclo de asimilación para los canales 5 a 9 del sensor AMSU-A de NOAA-15."}

periodos <- tibble(periodo = c("prueba_Bc", "E8"),
                   start = ymd_hms(c(20181111180000, 20181120180000)),
                   end = ymd_hms(c(20181118050000, 20181123120000)))

this_sensor <- sensores[1]

satbias[sensor == this_sensor & iuse == 1] %>% 
  .[channel %in% c(4:9)] %>%
  melt(measure.vars = paste0("coeff", c(1, 3:5, 8))) %>% 
  ggplot(aes(date, value)) +
  geom_path(aes(color = variable, group = interaction(variable, periodo)), size= 0.5) +
  geom_rect(data = periodos, aes(xmin = start, xmax = end, fill = periodo),
                ymin = -Inf, ymax = Inf, alpha = 0.1, inherit.aes = FALSE) +
  scale_color_brewer(type = "qual", palette = "Dark2",
                     labels = c("coeff1" = "Offset",
                                "coeff3" = "CLW",
                                "coeff4" = "TLR^2",
                                "coeff5" = "TLR",
                                "coeff8" = "Emissivity")) +
  scale_fill_manual(values = c("grey70", "black"), labels = c("prueba_Bc" = "training", "E8" = "RAD")) +
  # scale_linetype_discrete(name = "", labels = c("pre-RAD1", "pre-RAD2")) +
  scale_x_datetime(date_labels = "%d") +
  labs(color = NULL, x = "Date", y = NULL, fill = NULL) +
  facet_wrap(~channel, scales = "free_y", ncol = 3, labeller = labeller(channel = c("4" = "Channel 4",
                                                                                    "5" = "Channel 5",
                                                                                    "6" = "Channel 6",
                                                                                    "7" = "Channel 7",
                                                                                    "8" = "Channel 8",
                                                                                    "9" = "Channel 9"))) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
files <- Sys.glob("~/datosmunin3/EXP/E8/ANA/2018112*/diagfiles/asim*.ensmean")

files <- files[str_detect(files, "conv", negate = TRUE)]

rad <- read_diag_rad(files, "E8") %>% 
  satinfo[., on = c("sensor", "channel")] %>% 
  .[,  type :=  fifelse(sensor %in% multiespectrales, "multiespectral", "otro")] %>% 
  separate(sensor, into = c("sensor", "plat"), sep = "_")

rad[errinv %between% c(0.0000316227, 1000) & 
      qc == 0 & 
      iuse == 1 &
      peakwt %between% c(0.001, 1200), 
    .(tbc_mean = mean(tbc),
      tbcnob_mean = mean(tbcnob)), by = .(sensor, channel, date)] %>% 
  # .[, diff := abs(tbcnob_mean) - abs(tbc_mean)] %>% 
  # .[, range(diff, na.rm = TRUE)]
  melt(id.vars = c("sensor", "channel", "date")) %>% 
  ggplot(aes(value)) +
  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
                 alpha= 0.3, position = "identity") +
  geom_density(aes(color = variable, y = stat(density))) +
  geom_rug(data = ~.x[variable == "diff"]) +
  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
  scale_fill_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
  scale_color_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
  geom_vline(xintercept = 0, color = "grey40") +
  coord_cartesian(xlim = c(-2.5, 2.5)) +
  facet_wrap(~sensor, labeller = labeller(sensor = toupper)) +
  labs(fill = NULL, x = NULL, color = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")

rad[errinv %between% c(0.0000316227, 1000) & 
      qc == 0 & 
      iuse == 1 &
      peakwt %between% c(0.001, 1200) &
      sensor == "amsua" &
      plat == "n15", 
    .(tbc_mean = mean(tbc),
      tbcnob_mean = mean(tbcnob),
      .N), by = .(sensor, channel, date)] %>% 
   melt(id.vars = c("sensor", "channel", "date")) %>% 
  .[channel %in% c(4:9)] %>% 
  ggplot(aes(value)) +
  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
                 alpha= 0.3, position = "identity") +
  geom_density(aes(color = variable, y = stat(density))) +
  geom_rug(data = ~.x[variable == "diff"]) +
  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
  scale_fill_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
  scale_color_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
  geom_vline(xintercept = 0, color = "grey40") +
  coord_cartesian(xlim = c(-2.5, 2.5)) +
  facet_wrap(~channel, labeller = labeller(sensor = toupper)) +
  labs(fill = NULL, x = NULL, color = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
rad[errinv %between% c(0.0000316227, 1000) & 
      qc == 0 & 
      iuse == 1 &
      peakwt %between% c(0.001, 1200)] %>% 
  .[, lev := factor(cut_round(peakwt, breaks = seq(0, 1050, 50)))] %>% 
  .[, .(omb = mean(tbc),
        n = .N), by = .(date, lev, exp)] %>% 
  .[lev != 1025] %>% 
  ggplot(aes(date, fct_rev(lev))) +
  geom_tile(aes(fill = n)) +
  scale_fill_viridis_c(trans = scales::log10_trans(), breaks = c(1, 10, 100, 1000, 10000),
                       guide = guide_colourbar(barheight = 8,
                                               barwidth = 0.5)) +

  scale_date() +
  labs(x = NULL, fill = NULL, y = "pressure (hPa)") +
  theme_minimal() +
  theme(aspect.ratio = 1/1.5) 
```
```{r fss, fig.cap="FSS calculated over a 6-hours moving window for different thresholds and scales."}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2-8].csv")

fss <- purrr::map(files, function(f) {

  fread(f) %>% 
    .[, date := as_datetime(date)]
}) %>% 
  rbindlist()

colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311")

fss %>% 
  .[w %in% c(1, 11) & q %in% c(1, 25) & exp %in% c("E2", "E5", "E6", "E8")] %>% 
  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp)) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "10 km", "5" = "50 km", "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL, y = "FSS") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r dbz-mean, fig.cap="Observed maximum reflectivity in the column (upper row) and ensemble mean maximum reflectivity for CONV and RAD at different times (bottom two rows).", include=FALSE}
proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"

rad_loc <- read_csv(here("analysis", "data", "derived_data", "radar_loc.csv"))

files <- Sys.glob("/home/jruiz/datosmunin3/datos/DATOS_RADAR/MOSAICO/*.nc")[c(1, 7)]
# Solo un par de horarios 10, 13, 16 y 19 UTC

dbz_obs <- purrr::map(files, function(f) {

  ReadNetCDF(f, vars = "DBZH_cor") %>%
    .[!is.na(DBZH_cor)] %>% 
    .[, .(DBZH_cor = max(DBZH_cor, na.rm = TRUE)), by = .(x, y)] %>% 
    .[, max_dbz := 10 * log10(DBZH_cor)] %>% 
    .[, ":="(date = ymd_hms(str_remove(basename(f), ".nc")),
             DBZH_cor = NULL)] %>% 
    .[, c("lon", "lat") := proj4::project(list(x, y), proj_radar, inverse = TRUE)]

}) %>% rbindlist() %>% 
  .[, run := "OBSERVATIONS"]

dates <- c("20181122100000", 
           "20181122160000")
files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[28]_", dates, ".nc"))

dbz_ana <- purrr::map(files, function(f){

  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
    .[, ":="(exp = meta[[1]][["exp"]],
             date = ymd_hms(meta[[1]][["date"]]))] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] 
}) %>% 
  rbindlist()  %>% 
  .[, run := fifelse(exp == "E8", "RAD", "CONV")]

values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
            22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
            42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
            61.5, 63, 64.5, 66, 69, 76.5)

colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
             "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
             "#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
             "#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
             "#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
             "#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
             "#95E3C5", "#85E0BD")

dbz_obs %>% 
  # .[date == unique(date)[1]] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
                    proj = proj_radar,
                    breaks = values) +
  scale_fill_manual(values = colours, 
                    labels = function(x) JumpBy(x, 3, fill = ""),
                    drop = FALSE, 
                    guide = guide_colorsteps(barwidth = 20, 
                                             barheight = 0.5)) +
  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
  geom_mapa() +
  facet_grid(run ~ date) +
  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
  theme_minimal() +
  theme(legend.position = "bottom",
        # legend.box.spacing = margin(0, 0, 0, 0),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

  dbz_ana %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = max_dbz, fill = stat(level)), 
                    proj = norargentina_lambert,
                    breaks = values) +
  scale_fill_manual(values = colours, 
                    labels = function(x) JumpBy(x, 3, fill = ""),
                    drop = FALSE, 
                    guide = guide_colorsteps(barwidth = 20, 
                                             barheight = 0.5)) +
  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
  geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.4) +
  geom_mapa() +
  facet_grid(run ~ date) +
  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
  theme_minimal() +
  theme(legend.position = "bottom",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

  plot_layout(ncol = 1, heights = c(1, 2), guides = "collect") & 
  theme(legend.position = 'bottom')

```
```{r}
oficiales <- fread(here("analysis", "data", "raw_data", "E2_asim_conv_20181121120000.ensmean"), 
                   na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
  .[, c("V2", "V4") := NULL] %>% 
  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>% 
  .[type %in% c(181, 281)] %>% unique(by = c("stationID")) %>%
  .[!str_detect(stationID, pattern = "[A-Z]")] %>% 
  .[, source := "Sfc - Official"]

no_oficiales <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
                      na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
  .[, c("V2", "V4") := NULL] %>% 
  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr")) %>% 
  .[type %in% c(181, 187)] %>% 
  .[!str_detect(stationID, "SMN")] %>%
  .[!(stationID %in% oficiales$stationID)] %>% 
  unique(by = c("stationID")) %>% 
  .[, source := "Sfc - Non-official"]

obs <- rbind(no_oficiales, oficiales)

lista_sondeos <- fread(here("analysis/data/derived_data/lista_sondeos.csv")) %>% 
  .[, periodo := fcase(nominal_launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")),  "IOP07", 
                       nominal_launch_time %between% c(ymd_hms("20181122140000"), ymd_hms("20181122200000")),  "IOP08", 
                       default = "Others")
  ]

dominio <- fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

dominio %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = NULL,
                      breaks = seq(0, 6000, 1000),
                      guide = guide_colorstrip(barwidth = 15,
                                               barheight = 0.5)) +
  geom_mapa() +
  geom_point(data = obs, aes(ConvertLongitude(lon), lat, 
                             color = source, shape = source), 
             size = 1, alpha = 0.8) + 
  scale_shape_manual(name = NULL,  
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("ASWS","CSWS"), values = c(15, 17),
                     guide = guide_legend(override.aes = list(size = 2))) +
  scale_color_manual(name =  NULL, 
                     values = c("Sfc - Non-official" = "#00695c", 
                                "Sfc - Official" = "#FD8002"),
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("ASWS","CSWS"),
                     guide = guide_legend(override.aes = list(size = 2))) +
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  # geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom", 
        legend.box = "vertical") 
  # ,
```

### 2022-05-03 analysis/paper/paper_ana_20181122.Rmd
more organization

--- 2022-05-03 analysis/chapters/paper_ana_20181122.Rmd
@@
 arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10, 
-                                file.dir = "~/mesoda/analysis/data/derived_data/")
+                                file.dir = here("analysis/data/derived_data/"))
 arg_topo[, lon := metR::ConvertLongitude(lon)]
@@
 
-experiments <- c("E4", "E5", "E6", "E7")
-#Color blind friendly but E4 y E7 parecidos en Desaturated
 colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E9 = "#CC3311")
-# colores_exp <- c(E4 = "#225555", E5 = "#CCBB44", E6 = "#CC6677", E7 = "#882255")
+
 satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
   setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
+
 derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
@@
 ```{r obs-horizontal, fig.cap="(ref:obs-horizontal)", out.width="100%", fig.height=7}
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean")[1:67]
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean")
 
@@
 
-files <- files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
 
@@
 
-files <- files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]
+files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]
 
@@
                                                             E6 = "SATWND", E9 = "RAD")) +
-  # geom_tile(aes(fill = obs_cycle)) +
-  # scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
-  #                      guide = guide_colorbar(barheight = NULL)) +
   scale_y_log10() +
@@
 
-  purrr::map(Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/rcrv_*_perfil.csv")[1:6], function(f) {
+  purrr::map(Sys.glob(here("analysis/data/derived_data/rcrv_*_perfil.csv"))[1:6], function(f) {
     meta <- unglue::unglue(basename(f), "rcrv_{sensor}_perfil.csv")
@@
 ```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=5.5, fig.align="center",fig.pos="th"}
-files <- list.files(derived_data, recursive = TRUE, full.names = TRUE,
-                    pattern = "perfiles_ana")
+files <- Sys.glob(paste0(derived_data, "perfiles_ana_E[2,5,6,9].csv"))
 
@@
 
-files <- Sys.glob(paste0(derived_data, "pw/pw_ana_E[2,5,6,9]*"))
+files <- Sys.glob(paste0(derived_data, "pw/pw_ana_E[2,5,6,9]_20181122000000.nc"))
 
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/E[2569]_ana*_pp_pmm.rds")
+files <- Sys.glob(paste0(derived_data, "ppacum/E[2569]_ana*_pp_pmm.rds"))
 
@@
 
-readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
+readRDS(here("analysis/data/derived_data/IMERG_1h.rds")) %>% 
   .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
@@
 ```{r fss, fig.cap="(ref:fss)", fig.width=3.5, fig.height=3.5, fig.pos="t"}
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[2569].csv")
+files <- Sys.glob(here("analysis/data/derived_data/fss_6h_ana_ens_E[2569].csv"))
 
@@
 
-  # meta <- unglue(basename(f), "maxdbz_ana_{exp}_{date}.nc")
-  # ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
-  #   .[, ":="(exp = meta[[1]][["exp"]],
-  #            date = ymd_hms(meta[[1]][["date"]]))] %>% 
-  #   .[, c("x", "y") := wrf_project(lon, lat)] 
 }) %>% 
@@
            exp = NULL)]
-
-# values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21,
-#             22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5,
-#             42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60,
-#             61.5, 63, 64.5, 66, 69, 76.5)
 
@@
                                              frame.colour = "black")) +
-  # geom_point(data = rad_loc, aes(lon, lat, color = factor(nradar)), size = 2) +
   geom_path(data = rad_loc, aes(lon, lat, group = ID), size = 0.2, color = "grey20") +
@@
   theme(legend.position = "bottom",
-        # axis.text.x = element_blank(),
         plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
@@
 ```{r soundings, fig.cap="(ref:soundings)", fig.env="figure*", fig.pos="t", fig.align="center"}
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/derived_data/sondeos/ANA/sondeo_E[2569]_ana*")
+files <- Sys.glob(paste0(derived_data, "sondeos/ANA/sondeo_E[2569]_ana*")
 
@@
 
-files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,9]_ens.csv")
+files <- Sys.glob(here("analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,9]_ens.csv"))
 

### 2022-05-03 ASWS.R

@@ -1,42 +1,0 @@
-librayr(tidyverse)
-library(ggplot2)
-
-df <- map_df( Sys.glob("~/datosmunin3/DATA/OBS/datain/ADPSFC_n3/201811*"), ~fread(.x, na.strings = c("1e+11",
-                                                                                                     "1.000000e+11",
-                                                                                                     "1.0000e+11"))) %>%
-  setnames(paste0("V", c(1:10)), c("lat", "lon", "alt", "stationID", "date", "t", "q", "u", "v", "p"))
-
-
-n_stations <-
-  df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54), .N, by = stationID] %>%
-  .[str_detect(stationID, "SMN", negate = TRUE)]
-
-df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
-  .[str_detect(stationID, "SMN", negate = TRUE)] %>%
-  melt(measure.vars = c("t", "q", "u", "v", "p")) %>%
-  .[, is_na := fifelse(is.na(value), "NA", "OBS")] %>%
-  .[, .N, by = .(stationID, variable, is_na)] %>%
-  dcast(stationID + variable ~ is_na) %>%
-  .[!is.na(OBS)] %>%
-  dcast(stationID ~ variable, value.var = "OBS")
-
-df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
-  .[str_detect(stationID, "SMN", negate = TRUE)] %>%
-  melt(measure.vars = c("t", "q", "u", "v", "p")) %>%
-  .[, is_na := fifelse(is.na(value), "NA", "OBS")] %>%
-  .[, .N, by = .(stationID, variable, is_na)] %>%
-  dcast(stationID + variable ~ is_na) %>%
-  .[!is.na(OBS), .N, by = variable]
-
-SMN <- %>%
-  df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
-  # unique(by = c("stationID")) %>%
-  .[str_detect(stationID, "SMN")] %>%
-  unique(by = "date")
-
-df[lat %between% c(-41.5, -20) & lon %between% c(-75, -54)] %>%
-  unique(by = c("lat", "lon")) %>%
-  ggplot(aes(lon, lat)) +
-  geom_point() +
-  geom_point(data = square, size = 0.1) +
-  geom_mapa()

### 2022-05-03 analysis/scripts/AmG.R

@@ -1,0 +1,224 @@
+library(tidyverse)
+library(lubridate)
+library(data.table)
+library(unglue)
+library(metR)
+library(mesoda)
+source(here::here("temp_R/postprocesamiento.R"))
+
+map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")
+
+geom_mapa <- function() {
+  list(geom_sf(data = map, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE),
+       coord_sf(xlim = c(-76, -52), ylim = c(-41, -20)))
+}
+
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
+ini_date <- ymd_hms("20181120180000")
+ciclos <- 67/3
+
+dates <- seq(ini_date, by = "3 hour",
+             length.out = ciclos)
+
+coord <- ReadNetCDF(paste0(wrf_path, "E9/ANA/20181120180000/analysis.ensmean"),
+                    vars = c(lon = "XLONG", lat = "XLAT"), subset = list(bottom_top = 1))
+
+
+levs <- c(300, 500, 700)
+
+for (d in seq_along(dates)) {
+
+  print(dates[d])
+
+files_ana <- Sys.glob(paste0(wrf_path, "E[1,6,9]*/ANA/", format(dates[d], "%Y%m%d%H%M%S"), "/analysis.ensmean"))
+
+ana <- map(files_ana, function(f) {
+
+  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/ANA/{date}/analysis.ensmean")
+
+  ReadNetCDF(f, vars = c(t_ana = "T", p = "P", "PB", "QVAPOR",
+                         lon = "XLONG", lat = "XLAT")) %>%
+    .[, ":="(t_ana = tk(t_ana, p + PB, T_BASE = 290) - 273.15,
+             rh_ana = rh(QVAPOR, p + PB, tk(t_ana, p + PB, T_BASE = 290)),
+             p = p + PB)] %>%
+    .[, PB := NULL] %>%
+    .[, ":="(date = ymd_hms(meta[[1]][["date"]]),
+             exp = meta[[1]][["exp"]])] %>%
+    .[, .(lev = levs,
+          t_ana = approx(p, t_ana, xout = levs*100)$y,
+          rh_ana = approx(p, rh_ana, xout = levs*100)$y),
[…]

### 2022-05-03 analysis/scripts/R.R

@@ -1,0 +1,114 @@
+library(tidyverse)
+library(data.table)
+library(metR)
+library(mesoda)
+
+
+Rcpp::cppFunction('double outer(NumericVector x, NumericVector y) {
+  double n = x.size();
+  double total = 0;
+  for(int i = 0; i < n; ++i) {
+    for (int j = 0; j < n; ++j) {
+      total += x[i]*y[j];
+    }
+  }
+  total /= n*n;
+  return total;
+}')
+
+# outer <- cov
+
+path_data <- "/home/paola.corrales/datosmunin3/EXP/"
+
+files_gue <- Sys.glob(paste0(path_data, "E10/ANA/*/diagfiles/asim_abi_g16_*.ensmean"))[1:50]
+files_ana <- Sys.glob(paste0(path_data, "E10/ANA/*/diagfiles_ana/asim_abi_g16_*.ensmean"))[1:50]
+
+d <- map2(files_gue, files_ana, function(fg, fa) {
+
+  meta <- unglue::unglue(fg, "/home/paola.corrales/datosmunin3/EXP/E10/ANA/{date}/diagfiles/asim_abi_g16_{date2}.ensmean")
+  gue <- read_diag_rad(fg, "E10") %>%
+    .[channel %in% c(8:10) & qc == 0, .(channel, lat, lon, tbc)] %>%
+    setnames("tbc", "gue")
+
+  ana <- read_diag_rad(fa, "E10") %>%
+    .[channel %in% c(8:10) & qc == 0, .(channel, lat, lon, tbc)] %>%
+    setnames("tbc", "ana")
+
+  ana[gue, on = c("channel", "lat", "lon") ] %>%
+    na.omit(.) %>%
+    melt(measure.vars = c("ana", "gue")) %>%
+    dcast(lat + lon ~ channel + variable) %>%
+    na.omit(.) %>%
+    # .[, .(cor_88 = mean(`8_ana` %*% t(`8_gue`)),
+    #       cor_89 = mean(`8_ana` %*% t(`9_gue`)),
+    #       cor_810 = mean(`8_ana` %*% t(`10_gue`)),
+    #       cor_98 = mean(`9_ana` %*% t(`8_gue`)),
+    #       cor_99 = mean(`9_ana` %*% t(`9_gue`)),
+    #       cor_910 = mean(`9_ana` %*% t(`10_gue`)),
[…]

### 2022-05-03 analysis/scripts/temp.R

@@ -1,0 +1,88 @@
+# Calcula y guarda acumulados del pronóstico para todo el periodo
+
+library(metR)
+library(tidyverse)
+library(data.table)
+library(lubridate)
+library(unglue)
+library(mesoda)
+
+
+imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
+wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
+exp <- "E2"
+run <- "fcst"
+
+ini_date <- ymd_hms("20181122000000")
+ciclos <- 37
+acumulado <- 1
+
+dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
+                    length.out = ciclos - acumulado)
+
+
+q <- c(1, 5, 10, 25, 50)
+
+# Inicio
+
+pp_wrf <- purrr::map(dates, function(d) {
+
+  print(d)
+  date <- d
+
+  lead_time <- as.numeric(difftime(date, ini_date, units = "hours"))
+
+  files_wrf <-  purrr::map(c(acumulado, 0), function(l) {
+    paste0(wrf_path, exp, "/", toupper(run), "/",
+           format(ini_date, "%Y%m%d%H"), "/NPP/NPP_",
+           format(ini_date, "%Y"), "-",
+           format(ini_date, "%m"), "-",
+           format(ini_date, "%d"), "_",
+           format(ini_date, "%H"), "_FC",
+           formatC(lead_time - l, width = 2, flag = "0"),
+           ".nc")
+  }) %>% unlist()
+
+
+  pp_wrf <- purrr::map(files_wrf, function(f) {
+
+    if (!file.exists(f)) {
+      return(NULL)
+    }
+
+    ReadNetCDF(f, vars = c("PP", "XLAT", "XLONG")) %>%
+      .[, ":="(lon = XLONG, lat = XLAT)] %>%
+      .[, ":="(XLONG = NULL, XLAT = NULL)] %>%
+      .[, ":="(exp = exp,
+               date = date,
+               t = basename(f))]
+
+  }) %>%
+    rbindlist() %>%
+    .[, .(pp_acum = diff(PP)), by = .(lon, lat, date, exp, ens)] %>%
+    .[, c("x", "y") := wrf_project(lon, lat)]
+
+
[…]

### 2022-05-03 temp.R

@@ -1,88 +1,0 @@
-# Calcula y guarda acumulados del pronóstico para todo el periodo
-
-library(metR)
-library(tidyverse)
-library(data.table)
-library(lubridate)
-library(unglue)
-library(mesoda)
-
-
-imerg_path <- "/home/paola.corrales/mesoda/analysis/data/derived_data/"
-wrf_path <- "/home/paola.corrales/datosmunin3/EXP/"
-exp <- "E2"
-run <- "fcst"
-
-ini_date <- ymd_hms("20181122000000")
-ciclos <- 37
-acumulado <- 1
-
-dates <- seq.POSIXt(ini_date + hours(acumulado), by = "hour",
-                    length.out = ciclos - acumulado)
-
-
-q <- c(1, 5, 10, 25, 50)
-
-# Inicio
-
-pp_wrf <- purrr::map(dates, function(d) {
-
-  print(d)
-  date <- d
-
-  lead_time <- as.numeric(difftime(date, ini_date, units = "hours"))
-
-  files_wrf <-  purrr::map(c(acumulado, 0), function(l) {
-    paste0(wrf_path, exp, "/", toupper(run), "/",
-           format(ini_date, "%Y%m%d%H"), "/NPP/NPP_",
-           format(ini_date, "%Y"), "-",
-           format(ini_date, "%m"), "-",
-           format(ini_date, "%d"), "_",
-           format(ini_date, "%H"), "_FC",
-           formatC(lead_time - l, width = 2, flag = "0"),
-           ".nc")
-  }) %>% unlist()
-
-
-  pp_wrf <- purrr::map(files_wrf, function(f) {
-
-    if (!file.exists(f)) {
-      return(NULL)
-    }
-
-    ReadNetCDF(f, vars = c("PP", "XLAT", "XLONG")) %>%
-      .[, ":="(lon = XLONG, lat = XLAT)] %>%
-      .[, ":="(XLONG = NULL, XLAT = NULL)] %>%
-      .[, ":="(exp = exp,
-               date = date,
-               t = basename(f))]
-
-  }) %>%
-    rbindlist() %>%
-    .[, .(pp_acum = diff(PP)), by = .(lon, lat, date, exp, ens)] %>%
-    .[, c("x", "y") := wrf_project(lon, lat)]
-
-
[…]

### 2022-05-04 analysis/scripts/file_list.R
files list for zenodo

@@ -1,0 +1,91 @@
+library(here)
+
+derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
+
+dates <- c("20181122100000", "20181122130000",
+           "20181122160000", "20181122190000")
+
+derived_data_paper <- "/home/paola.corrales/datosmunin3/EXP/paper_ana_20181122-derived_data/"
+
+dir.create(derived_data_paper)
+
+# sample_obs
+dir.create(paste0(derived_data_paper, "sample_obs"))
+file.copy(
+  c(here("analysis", "data", "raw_data", "E2_asim_conv_20181121120000.ensmean")
+    ,here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean")
+    ,here("analysis/data/derived_data/lista_sondeos.csv")
+    ,here("analysis", "data", "derived_data", "dominio_hgt.csv")),
+  paste0(derived_data_paper, "sample_obs/"))
+
+# tables
+dir.create(paste0(derived_data_paper, "tables"))
+file.copy(
+  c(here("analysis/data/derived_data/table1.csv")
+    ,here("analysis", "data", "derived_data", "tabla_radianzas.csv")),
+  paste0(derived_data_paper, "tables/"))
+
+# omb_diagfiles (subfolders for each exp)
+dir.create(paste0(derived_data_paper, "omb_diafiles"))
+dir.create(paste0(derived_data_paper, "omb_diafiles/E2"))
+dir.create(paste0(derived_data_paper, "omb_diafiles/E5"))
+dir.create(paste0(derived_data_paper, "omb_diafiles/E6"))
+dir.create(paste0(derived_data_paper, "omb_diafiles/E9"))
+
+purrr::map(Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean"), function(f) {
+  file.copy(f, paste0(derived_data_paper, "omb_diagfiles/E2/"))
+})
+purrr::map(Sys.glob("/home/paola.corrales/datosmunin3/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67], function(f) {
+  file.copy(f, paste0(derived_data_paper, "omb_diagfiles/E5/"))
+})
[…]

### 2022-05-04 analysis/paper/paper_ana_20181122.Rmd
change path to make the thing reproducible

@@
 
-square <- fread(here("analysis", "data", "derived_data", "dominio_square.csv"))
-square2 <- fread(here("analysis", "data", "derived_data", "dominio_square2.csv"))
-coord <- fread(here("analysis", "data", "derived_data", "coordenadas.csv"))
+square <- fread(here("analysis/data/derived_data/sample_obs/dominio_square.csv"))
+square2 <- fread(here("analysis/data/derived_data/sample_obs/dominio_square2.csv"))
+coord <- fread(here("analysis/data/derived_data/sample_obs/coordenadas.csv"))
 
@@
 
-satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
-  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
-
 derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
@@
 
-oficiales <- fread(here("analysis", "data", "raw_data", "E2_asim_conv_20181121120000.ensmean"), 
+oficiales <- fread(here("analysis/data/derived_data/sample_obs/E2_asim_conv_20181121120000.ensmean"), 
                    na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
@@
 
-no_oficiales <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
+no_oficiales <- fread(here("analysis/data/derived_data/sample_obs/E5_asim_conv_20181121120000.ensmean"), 
                       na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
@@
 
-lista_sondeos <- fread(here("analysis/data/derived_data/lista_sondeos.csv")) %>% 
+lista_sondeos <- fread(here("analysis/data/derived_data/sample_obs/lista_sondeos.csv")) %>% 
   .[, periodo := fcase(nominal_launch_time %between% c(ymd_hms("20181121150000"), ymd_hms("20181121210000")),  "IOP 7", 
@@
 
-dominio <- fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
+dominio <- fread(here("analysis/data/derived_data/sample_obs/dominio_hgt.csv")) %>% 
   .[, c("x", "y") := wrf_project(lon, lat)]
@@
 ```{r table-obs}
-fread(here("analysis/data/derived_data/table1.csv")) %>% 
-  # str()
+fread(here("analysis/data/derived_data/tables/table1.csv")) %>% 
   kbl(caption = "Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error, and the gross check thresholds used.",
@@
 ```{r table-rad}
-fread(here("analysis", "data", "derived_data", "tabla_radianzas.csv")) %>% 
+fread(here("analysis/data/derived_data/tables/tabla_radianzas.csv")) %>% 
   .[, ":="(sensor = toupper(sensor),
@@
 ```{r obs-horizontal, fig.cap="(ref:obs-horizontal)", out.width="100%", fig.height=7}
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E2/ANA/*/diagfiles/asim*ensmean")
+satinfo <- fread(here("analysis/data/derived_data/sample_obs/satinfo.txt")) %>% 
+  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
+
+files <- Sys.glob(here("analysis/data/derived_data/omb_diagfiles/E2/asim*ensmean"))
 
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E5/ANA/*/diagfiles/asim*ensmean")[1:67]
+files <- Sys.glob(here("analysis/data/derived_data/omb_diagfiles/E5/asim*ensmean"))
 
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E6/ANA/*/diagfiles/asim*ensmean")[1:67]
+files <- Sys.glob(here("analysis/data/derived_data/omb_diagfiles/E6/asim*ensmean"))
 
@@
 
-files <- Sys.glob("/home/paola.corrales/datosmunin3/EXP/E9/ANA/*/diagfiles/asim*ensmean")
+files <- Sys.glob(here("analysis/data/derived_data/omb_diagfiles/E9/asim*ensmean"))
 
@@
 
-For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds.
-Following @maldonado2021 the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members :
-
-$$\mathit{FSS} = 1-\frac{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i)^{2}+\frac{1}{N}\sum_{i=1}^{N} ({P_o}_i)^{2}} $$
+For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute a Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds.
+Following @maldonado2021 the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members:
+
+$$\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} $$
 
@@
 
-rbind(fread(here("analysis", "data", "derived_data", "rcrv_V_box.csv")),
-      fread(here("analysis", "data", "derived_data", "rcrv_t_box.csv"))) %>%   
+rbind(fread(here("analysis/data/derived_data/RCRV/rcrv_V_box.csv")),
+      fread(here("analysis/data/derived_data/RCRV/rcrv_t_box.csv"))) %>%   
   .[, type.obs := type] %>% 
@@
 
-rbind(fread(here("analysis", "data", "derived_data", "rcrv_V_perfil.csv")), 
-      fread(here("analysis", "data", "derived_data", "rcrv_t_perfil.csv"))) %>% 
+rbind(fread(here("analysis/data/derived_data/RCRV/rcrv_V_perfil.csv")), 
+      fread(here("analysis/data/derived_data/RCRV/rcrv_t_perfil.csv"))) %>% 
   obs.type[., on = "type"] %>%   
@@
 
-  read_csv(here::here("analysis", "data", "derived_data", "rcrv_satwind.csv")) %>% 
+  read_csv(here("analysis/data/derived_data/RCRV/rcrv_satwind.csv")) %>% 
   setDT() %>% 
@@
 
-  purrr::map(Sys.glob(here("analysis/data/derived_data/rcrv_*_perfil.csv"))[1:6], function(f) {
+  purrr::map(Sys.glob(here("analysis/data/derived_data/RCRV/rcrv_*_perfil.csv"))[1:6], function(f) {
     meta <- unglue::unglue(basename(f), "rcrv_{sensor}_perfil.csv")
@@
 ```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=5.5, fig.align="center",fig.pos="th"}
-files <- Sys.glob(paste0(derived_data, "perfiles_ana_E[2,5,6,9].csv"))
+files <- Sys.glob(here("analysis/data/derived_data/analysis_variables/perfiles_ana_E[2569].csv"))
 
@@
 
-files <- Sys.glob(paste0(derived_data, "pw/pw_ana_E[2,5,6,9]_20181122000000.nc"))
+files <- Sys.glob(here("analysis/data/derived_data/analysis_variables/pw/pw_ana_E[2569]_20181122000000.nc"))
 
@@
 
-files <- Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,9]_20181122000000.nc"))
+files <- Sys.glob(here("analysis/data/derived_data/analysis_variables/mcape/mcape_ana_E[2569]_20181122000000.nc"))
 
@@
 
+dir <- here("analysis/data/derived_data/analysis_variables")
+
 ana <- lapply(seq_len(nrow(fcsts)), function(f) {
 
-  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
-    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
-    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
-    .[, theta := ReadNetCDF(paste0(derived_data, "/theta_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(th = "theta"), out = "vector")] %>%
+  tmp <- ReadNetCDF(paste0(dir, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
+    .[, v := ReadNetCDF(paste0(dir, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
+    .[, q := ReadNetCDF(paste0(dir, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>% 
+    .[, theta := ReadNetCDF(paste0(dir, "/theta_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(th = "theta"), out = "vector")] %>%
     .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
@@
 
-files <- Sys.glob(paste0(derived_data, "ppacum/E[2569]_ana*_pp_pmm.rds"))
+files <- Sys.glob(here("analysis/data/derived_data/analysis_variables/pp_pmm/E[2569]_ana*_pp_pmm.rds"))
 
@@
 
-readRDS(here("analysis/data/derived_data/IMERG_1h.rds")) %>% 
+readRDS(here("analysis/data/derived_data/observations/IMERG_1h.rds")) %>% 
   .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
@@
 ```{r fss, fig.cap="(ref:fss)", fig.width=3.5, fig.height=3.5, fig.pos="t"}
-files <- Sys.glob(here("analysis/data/derived_data/fss_6h_ana_ens_E[2569].csv"))
+files <- Sys.glob(here("analysis/data/derived_data/FSS/fss_6h_ana_ens_E[2569].csv"))
 
@@
 
-(ref:dbz-mean) Maximum reflectivity in the column (COLMAX in $dBZ$), observed (upper row) and 1-hr forecast ensemble mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018.
+(ref:dbz-mean) Maximum reflectivity in the column (COLMAX in $dBZ$), observed (upper row) and 1-hr forecast probability matched mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018.
 Black circles in first row show the observation range of each radar.
@@
 
-files <- Sys.glob("/home/jruiz/datosmunin3/datos/DATOS_RADAR/MOSAICO/*.nc")[seq(1, 12, 3)]
+files <- Sys.glob(here("analysis/data/derived_data/observations/*.nc"))
 # Solo un par de horarios 10, 13, 16 y 19 UTC
@@
 
-dates <- c("20181122100000", "20181122130000", 
-           "20181122160000", "20181122190000")
-files <- Sys.glob(paste0(derived_data, "dbz/E[29]_ana_", dates, "_dbz_pmm.rds"))
+files <- Sys.glob(here("analysis/data/derived_data/analysis_variables/dbz_pmm/E[29]_ana_*_dbz_pmm.rds"))
 
@@
 ```{r soundings, fig.cap="(ref:soundings)", fig.env="figure*", fig.pos="t", fig.align="center"}
-files <- Sys.glob(paste0(derived_data, "sondeos/ANA/sondeo_E[2569]_ana*")
+files <- Sys.glob(here("analysis/data/derived_data/soundings/sondeo_E[2569]_ana*"))
 
@@
 
-files <- Sys.glob(here("analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,9]_ens.csv"))
+files <- Sys.glob(here("analysis/data/derived_data/FSS/fss_6h_fcst_ens_*_E[2569]_ens.csv"))
 

### 2022-05-10 analysis/paper/paper_ana_20181122.Rmd
comentarios de Maru

@@
   We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND, and satellite radiances from different microwave and infrared sensors.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-We also run ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
+We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
 We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
@@
 In that regard, several authors investigated the impact of assimilating surface weather data (e.g.
-@wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016 using different assimilation methodologies.
+@wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016) using different assimilation methodologies.
 Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.
@@
 
-We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis.
-with random perturbations to generate the initial ensemble perturbations.
+We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis with random perturbations to generate the initial ensemble perturbations.
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021.
@@
 These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft.
-The orange triangles in Figure \@ref(fig:dominio) indicate the location of the surface stations included in this experiment.
+The orange triangles in Figure \@ref(fig:dominio)a indicate the location of the surface stations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes.
@@
 Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study.
-This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \@ref(tab:table-rad)).
 Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
@@
 
-In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations.
+In the GSI system, the $\beta_i$ coefficients are trained using a variational estimation method which solves the $\beta_i$ that provides the best fit between the simulation and the observations.
 The coefficients were initialized at 18 UTC Nov 18, 2018 with the GFS system coefficients.
@@
 
-The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \@ref(fig:obs-horizontal).
-The larger number of assimilated observations over the center and east of the domain corresponds to the AWS observations.
-In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown.
-Local maxima at 12 and 00 UTC found mainly in CONV are attributed to operational soundings.
-The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
-The maxima at 13-14 and 01-02 UTC in RAD correspond to the contribution of the multispectral sensors.
-The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b) shows a maximum in low levels due to the AWS observations.
-Satellite-derived winds are maximized at the upper troposphere (between 500-250 hPa).
-Above 850 hPa, most of the observations correspond to radiance observations.
-
-All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
-The initial 60-member ensemble is generated as explained in section \@ref(config) from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \@ref(fig:cycle)).
-
-Ensemble forecasts initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
-Both forecasts are integrated until 12 UTC Nov 23.
-All forecasts use the same domain and ensemble configuration as the analysis.
-The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1).
-
 ```{r table-exp}
@@
 
+The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \@ref(fig:obs-horizontal).
+The larger number of assimilated observations over the center and east of the domain corresponds to the AWS observations.
+In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown.
+Local maxima at 12 and 00 UTC found mainly in CONV are attributed to operational soundings.
+The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
+The maxima at 13-14 and 01-02 UTC in RAD correspond to the contribution of the multispectral sensors.
+The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b) shows a maximum in low levels due to the AWS observations.
+Satellite-derived winds are maximized at the upper troposphere (between 500-250 hPa).
+Above 850 hPa, most of the observations correspond to radiance observations.
+
+All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
+The initial 60-member ensemble is generated as explained in section \@ref(config) from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \@ref(fig:cycle)).
+
+Ensemble forecasts initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
+Both forecasts are integrated until 12 UTC Nov 23.
+All forecasts use the same domain and ensemble configuration as the analysis.
+The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1).
+
 (ref:cycle) Figure 4.
@@
 
-where $O$ and $X$ stand for independent observations, and the simulations respectively and N is the sample size.
-
-For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute a Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds.
-Following @maldonado2021 the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members:
+where $O$ and $X$ stand for independent observations and the simulations respectively, and N is the sample size.
+
+For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute the Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds:
 
@@
 
-where $P_{xi}$ is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and $P_{oi}$ is the proportion of grid points in each sampling area in which the observed accumulated precipitation is over the same threshold.
-Since the precipitation forecast of all the ensemble members is considered in the computation of $P_{xi}$, the smoothing effect of the ensemble mean is avoided.
+Where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold.
+Following @maldonado2021, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area.
 The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
@@
 As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation.
-Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid.
+Figure \@ref(fig:rcrv-sfc) shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid.
 The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error.
@@
 
-The analyzed distribution of temperature and moisture in the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed in the temperature profiles (Figure \@ref(fig:TQ-diff)a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer that SATWND.
+The analyzed distribution of temperature and moisture in the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed in the temperature profiles (Figure \@ref(fig:TQ-diff)a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer than in SATWND.
 On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
@@
 First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \@ref(fig:pp-hov)a shows the mean hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) [@clark2017] for the first-guess hourly accumulated precipitation as averaged between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments.
+Figure \@ref(fig:pp-hov)a shows the hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) [@clark2017] for the first-guess hourly accumulated precipitation as averaged between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments.
 The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \@ref(fig:pp-hov)a).
@@
 
-To quantify the spatial match between the observed precipitation and the short-range forecast precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
-All experiments show similar values of FSS during the initialization of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
+To quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
+All experiments show similar values of FSS during the initiation of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
 This indicates that radiance observations have a positive impact on the analysis.
@@
 
-rad_loc <- fread(here("analysis", "data", "derived_data", "radar_loc.csv")) %>% 
+rad_loc <- fread(here("analysis/data/derived_data/sample_obs/radar_loc.csv")) %>% 
   .[, run := "OBSERVATIONS"]
@@
 The assimilation of AWS, SATWND, and RAD reduces the cold bias and RMSE for temperature between 5 and 12 km and the RMSE in the PBL compared with CONV (Figure \@ref(fig:soundings)e).
-The reduction of bias and RMSE is also important for dew point temperature (Figure \@ref(fig:soundings)f) with SATWND showing the bigger impact followed by AWS and RAD.
+The reduction of bias and RMSE is also important for dew point temperature (Figure \@ref(fig:soundings)f) with SATWND showing the biggest impact followed by AWS and RAD.
 The zonal wind is overestimated in the analyses and only RAD shows an improvement with respect to CONV in the upper troposphere (Figure \@ref(fig:soundings)g).
@@
 
-In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
-We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the 1-h forecast to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
+In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
+We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
 CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
 AWS, SATWND, and RAD show improvements in the FSS values, particularly for the higher threshold (Figure \@ref(fig:fssfcst)b, d).
-Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD tham the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
+Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD than the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
 
@@
 
-```{r fssfcst, fig.cap="(ref:fssfcst)", fig.width=3.5, fig.height=4}
+```{r fssfcst, fig.cap="(ref:fssfcst)", fig.width=3.5, fig.height=3.5}
 
@@
 
-To summarize, in this study we conclude that the assimilation of surface observations with higher spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
+To summarize, in this study we conclude that the assimilation of surface observations with high spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
 Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events.
@@
 
-# Acknowledgements
+# Code and data availability
+
+A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at <https://github.com/paocorrales/mesoda>.
+The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/[doi], reference number [reference number].
+
+# Acknowledgements {.unnumbered}
 

### 2022-05-10 analysis/paper/paper_ana_20181122.tex
comentarios de Maru

--- 2022-05-03 analysis/chapters/paper_ana_20181122.tex
@@
   We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND, and satellite radiances from different microwave and infrared sensors.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-We also run ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
+We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
 We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
@@
 (2019), and Chen et al.
-(2016) using different assimilation methodologies.
+(2016)) using different assimilation methodologies.
 Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.
@@
 
-We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis.
-with random perturbations to generate the initial ensemble perturbations.
+We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis with random perturbations to generate the initial ensemble perturbations.
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in Necker et al.
@@
 These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft.
-The orange triangles in Figure \ref{fig:dominio} indicate the location of the surface stations included in this experiment.
+The orange triangles in Figure \ref{fig:dominio}a indicate the location of the surface stations included in this experiment.
 The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for radiosondes.
@@
 Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study.
-This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2).
+This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \ref{tab:table-rad}).
 Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath.
@@
 
-In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves for the \(\beta_i\) that provides the best fit between the simulation and the observations.
+In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves the \(\beta_i\) that provides the best fit between the simulation and the observations.
 The coefficients were initialized at 18 UTC Nov 18, 2018 with the GFS system coefficients.
@@
 
-The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \ref{fig:obs-horizontal}.
-The larger number of assimilated observations over the center and east of the domain corresponds to the AWS observations.
-In Figure \ref{fig:obs-cycle}a the number of assimilated observations over time is shown.
-Local maxima at 12 and 00 UTC found mainly in CONV are attributed to operational soundings.
-The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
-The maxima at 13-14 and 01-02 UTC in RAD correspond to the contribution of the multispectral sensors.
-The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the AWS observations.
-Satellite-derived winds are maximized at the upper troposphere (between 500-250 hPa).
-Above 850 hPa, most of the observations correspond to radiance observations.
-
-All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
-The initial 60-member ensemble is generated as explained in section \ref{config} from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \ref{fig:cycle}).
-
-Ensemble forecasts initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
-Both forecasts are integrated until 12 UTC Nov 23.
-All forecasts use the same domain and ensemble configuration as the analysis.
-The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
-Department of Commerce (2015)).
-
 \begin{table}
@@
 
+The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \ref{fig:obs-horizontal}.
+The larger number of assimilated observations over the center and east of the domain corresponds to the AWS observations.
+In Figure \ref{fig:obs-cycle}a the number of assimilated observations over time is shown.
+Local maxima at 12 and 00 UTC found mainly in CONV are attributed to operational soundings.
+The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage.
+The maxima at 13-14 and 01-02 UTC in RAD correspond to the contribution of the multispectral sensors.
+The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the AWS observations.
+Satellite-derived winds are maximized at the upper troposphere (between 500-250 hPa).
+Above 850 hPa, most of the observations correspond to radiance observations.
+
+All the assimilation experiments start at 18 UTC Nov 20, 2018 and continue until 12 UTC Nov, 23 (totaling 67 hours/assimilation cycles).
+The initial 60-member ensemble is generated as explained in section \ref{config} from a spin-up run without assimilating observations performed between 12 UTC and 18 UTC Nov, 20 (Figure \ref{fig:cycle}).
+
+Ensemble forecasts initialized from the different analysis experiments at 00 and 06 UTC Nov 22 were performed to evaluate the impact of the different observing networks on short range precipitation forecasts.
+Both forecasts are integrated until 12 UTC Nov 23.
+All forecasts use the same domain and ensemble configuration as the analysis.
+The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
+Department of Commerce (2015)).
+
 \begin{figure}
@@
 
-where \(O\) and \(X\) stand for independent observations, and the simulations respectively and N is the sample size.
-
-For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds.
+where \(O\) and \(X\) stand for independent observations and the simulations respectively, and N is the sample size.
+
+For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute the Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds:
+
+\[\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} \]
+
+Where \(P_{oi}\) is the fraction of grid points in the \(i-th\) sampling area in which the observed accumulated precipitation is greater than a specified threshold.
 Following Maldonado et al.
-(2021) the definition of the FSS is slightly modified in that the probability of having rain exceeding a particular threshold and over a particular region is computed taking into account all the ensemble members :
-
-\[\mathit{FSS} = 1-\frac{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\frac{1}{N}\sum_{i=1}^{N} ({P_x}_i)^{2}+\frac{1}{N}\sum_{i=1}^{N} ({P_o}_i)^{2}} \]
-
-where \(P_{xi}\) is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and \(P_{oi}\) is the proportion of grid points in each sampling area in which the observed accumulated precipitation is over the same threshold.
-Since the precipitation forecast of all the ensemble members is considered in the computation of \(P_{xi}\), the smoothing effect of the ensemble mean is avoided.
+(2021), \(P_{xi}\) is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the \(i-th\) sampling area.
 The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
@@
 As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation.
-Figure 5 shows the \(sd RCRV\) for surface observations box-averaged to a 2.5° grid.
+Figure \ref{fig:rcrv-sfc} shows the \(sd RCRV\) for surface observations box-averaged to a 2.5° grid.
 The \(sd RCRV\) for wind observations (Figure \ref{fig:rcrv-sfc}a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error, and the observation error.
@@
 
-The analyzed distribution of temperature and moisture in the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed in the temperature profiles (Figure \ref{fig:TQ-diff}a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer that SATWND.
+The analyzed distribution of temperature and moisture in the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed in the temperature profiles (Figure \ref{fig:TQ-diff}a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer than in SATWND.
 On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
@@
 First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \ref{fig:pp-hov}a shows the mean hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments.
+Figure \ref{fig:pp-hov}a shows the hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments.
 The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \ref{fig:pp-hov}a).
@@
 
-To quantify the spatial match between the observed precipitation and the short-range forecast precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}).
-All experiments show similar values of FSS during the initialization of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
+To quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}).
+All experiments show similar values of FSS during the initiation of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
 This indicates that radiance observations have a positive impact on the analysis.
@@
 \begin{figure*}[h]
-\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast ensemble mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018.
+\includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast probability matched mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018.
 Black circles in first row show the observation range of each radar.}\label{fig:dbz-mean}
@@
 The assimilation of AWS, SATWND, and RAD reduces the cold bias and RMSE for temperature between 5 and 12 km and the RMSE in the PBL compared with CONV (Figure \ref{fig:soundings}e).
-The reduction of bias and RMSE is also important for dew point temperature (Figure \ref{fig:soundings}f) with SATWND showing the bigger impact followed by AWS and RAD.
+The reduction of bias and RMSE is also important for dew point temperature (Figure \ref{fig:soundings}f) with SATWND showing the biggest impact followed by AWS and RAD.
 The zonal wind is overestimated in the analyses and only RAD shows an improvement with respect to CONV in the upper troposphere (Figure \ref{fig:soundings}g).
@@
 
-In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment.
-We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the 1-h forecast to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}).
+In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
+We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}).
 CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
 AWS, SATWND, and RAD show improvements in the FSS values, particularly for the higher threshold (Figure \ref{fig:fssfcst}b, d).
-Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD tham the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
+Moreover, the late initialization at 06 UTC performs better for AWS, SATWND, and RAD than the forecast initialized at 00 UTC, highlighting the positive impact of the observations assimilated between 00 and 06 UTC.
 
@@
 
+\hypertarget{code-and-data-availability}{%
+\section{Code and data availability}\label{code-and-data-availability}}
+
+A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at \url{https://github.com/paocorrales/mesoda}.
+The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/{[}doi{]}, reference number {[}reference number{]}.
+
 \hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
 

### 2022-05-12 _dependencies.R
renv

@@ -1,0 +1,4 @@
+library(rticles)
+library(rnaturalearthhires)
+library(rgeos)
+library(udunits2)

### 2022-05-12 analysis/paper/paper_ana_20181122.Rmd
For review

@@
 journal: Atmospheric Research
-#layout: authoryear,preprint,review,12pt
-layout: final,5p,times,twocolumn,authoryear
+layout: authoryear,preprint,review,12pt
+#layout: final,5p,times,twocolumn,authoryear
 address:
@@
                            labels = c("a)", "b)"), label_size = 11), legend, 
-                 ncol = 1, rel_heights = c(1, 0.05)))
+                 ncol = 1, rel_heights = c(1, 0.08)))
 
@@
 Where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold.
-Following @maldonado2021, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area.
+Following @roberts2020, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area.
 The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
@@
 A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at <https://github.com/paocorrales/mesoda>.
-The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/[doi], reference number [reference number].
+The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5072/zenodo.1060834, version 0.9.0.
 

### 2022-05-12 analysis/paper/paper_ana_20181122.tex
For review

@@
-\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
+\documentclass[authoryear,preprint,review,12pt]{elsarticle} %review=doublespace preprint=single 5p=2 column
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
@@
 \usepackage{lineno} % add
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
 \usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
 %%%%%%%%%%%%%%%% end my additions to header
@@
 \bibliographystyle{elsarticle-harv}
-\usepackage{longtable}
 \usepackage{graphicx}
@@
 
-% Pandoc citation processing
-
-% Pandoc header
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 \usepackage{booktabs}
@@
 \begin{document}
+
 \begin{frontmatter}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Figure 4.
+\includegraphics[width=1\linewidth]{../figures/analysis_cycle} \caption{Figure 4.
 Diagram of the analysis cycles between 18 UTC Nov 20, and 12 UTC Nov 23 plus spin up period of 6 hours.
@@
 Where \(P_{oi}\) is the fraction of grid points in the \(i-th\) sampling area in which the observed accumulated precipitation is greater than a specified threshold.
-Following Maldonado et al.
-(2021), \(P_{xi}\) is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the \(i-th\) sampling area.
+Following Roberts et al.
+(2020), \(P_{xi}\) is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the \(i-th\) sampling area.
 The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
@@
 A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at \url{https://github.com/paocorrales/mesoda}.
-The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/{[}doi{]}, reference number {[}reference number{]}.
+The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5072/zenodo.1060834, version 0.9.0.
 
@@
 
+\leavevmode\hypertarget{ref-roberts2020}{}%
+Roberts, B., Gallo, B.T., Jirak, I.L., Clark, A.J., Dowell, D.C., Wang, X., Wang, Y., 2020.
+What Does a Convection-Allowing Ensemble of Opportunity Buy Us in Forecasting Thunderstorms?
+Weather and Forecasting 35, 2293--2316.
+doi:\href{https://doi.org/10.1175/WAF-D-20-0069.1}{10.1175/WAF-D-20-0069.1}
+
 \leavevmode\hypertarget{ref-roberts2008}{}%

### 2022-05-16 analysis/paper/paper_ana_20181122.Rmd
Steve's comments

@@
 The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign.
-
 We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
@@
 - code: Illinois
-  address: Department of Atmospheric Sciences, University of Illinois at Urbana–Champaign, Urbana, Illinois.
+  address: Department of Atmospheric Sciences, University of Illinois Urbana–Champaign, Urbana, Illinois.
 bibliography: 
@@
   text: Corresponding Author
-keywords: Regional Data Assimilation, Surface Obsevations, Satellite Observations 
+keywords: Regional Data Assimilation, Surface Observations, Satellite Observations 
 numbersections: true
-# header-includes: 
+header-includes: 
 #  - \usepackage{gensymb}
@@
 #  - \usepackage{natbib}
-#  - \usepackage{lineno}
-#  - \linenumbers 
+  - \usepackage{lineno}
+  - \linenumbers 
 ---
@@
 
-Severe weather events cause significant human and material losses around the world.
-A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
+Severe weather events cause significant human and economic losses around the world.
+A large number of these phenomena are associated with the occurrence of deep moist convection, including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
 Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer.
@@
 
-While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low model tops used for this application.
+While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, bias correction, and the relatively low model tops used for this application.
 @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
@@
 They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts.
-Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land.
-Moreover, to the best of our knowledge, there are none studies related to the direct assimilation of radiance observations over South America.
+Given these mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land.
+Moreover, to the best of our knowledge, there are no studies related to the direct assimilation of radiance observations over South America.
 
@@
 
-To reach this goal, we conduct several DA experiments for a case study of a gigantic Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
+To reach this goal, we conduct several DA experiments for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
 
@@
 (ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
-b) Locations of radiosonde launches during the RELAMPAGO.
+b) Locations of radiosonde launches during RELAMPAGO.
 Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions.
@@
 The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere.
+
 The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-First, a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
+First, a spatial data thinning is applied using a 60 km grid following @singh2016, @jones2013, and @lin2017a, where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
 Also, observations over the sea are preferred to those over land or snow [@hu2018].
@@
 
-* Radar observations are used to perform a qualitative and visual assessment of the convection features.
+* Radar observations are used to perform a qualitative and visual assessment of the convective features.
 The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes.
@@
 $$\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}$$
-
 where $M$ is the ensemble size.
@@
 $$\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)$$
-
 where $O$ and $X$ stand for independent observations and the simulations respectively, and N is the sample size.
@@
 $$\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} $$
-
-Where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold.
+where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold.
 Following @roberts2020, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area.

### 2022-05-16 analysis/paper/paper_ana_20181122.tex
Steve's comments

@@
 
+\usepackage{lineno}
+\linenumbers
 \usepackage{booktabs}
@@
     \address[NCAR]{National Center for Atmospheric Research, Boulder, Colorado.}
-    \address[Illinois]{Department of Atmospheric Sciences, University of Illinois at Urbana--Champaign, Urbana, Illinois.}
+    \address[Illinois]{Department of Atmospheric Sciences, University of Illinois Urbana--Champaign, Urbana, Illinois.}
       \cortext[1]{Corresponding Author}
@@
   \end{abstract}
-   \begin{keyword} Regional Data Assimilation, Surface Obsevations, Satellite Observations\end{keyword}
+   \begin{keyword} Regional Data Assimilation, Surface Observations, Satellite Observations\end{keyword}
  \end{frontmatter}
@@
 
-Severe weather events cause significant human and material losses around the world.
-A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
+Severe weather events cause significant human and economic losses around the world.
+A large number of these phenomena are associated with the occurrence of deep moist convection, including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
 Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events (Brooks et al., 2003) and large hail events (Cecil and Blankenship, 2012), particularly during austral spring and summer.
@@
 
-While the assimilation of radiance observations into global models is well established (Eyre et al., 2020), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low model tops used for this application.
+While the assimilation of radiance observations into global models is well established (Eyre et al., 2020), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, bias correction, and the relatively low model tops used for this application.
 Bao et al.
@@
 They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts.
-Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land.
-Moreover, to the best of our knowledge, there are none studies related to the direct assimilation of radiance observations over South America.
+Given these mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land.
+Moreover, to the best of our knowledge, there are no studies related to the direct assimilation of radiance observations over South America.
 
 In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed.
-Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources (e.g.~Dillon et al., 2016, 2021; gustavo Goncalves de Goncalves et al., 2015).
+Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources (e.g.~Dillon et al., 2016, 2021; Goncalves de Goncalves et al., 2015).
 In particular Dillon et al.
@@
 \includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
-b) Locations of radiosonde launches during the RELAMPAGO.
+b) Locations of radiosonde launches during RELAMPAGO.
 Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions.
@@
 The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere.
+
 The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-First, a spatial data thinning is applied using a 60 km grid following (Jones et al., 2013; Lin et al., 2017; Singh et al., 2016), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
+First, a spatial data thinning is applied using a 60 km grid following Singh et al.
+(2016), Jones et al.
+(2013), and Lin et al.
+(2017), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
 Also, observations over the sea are preferred to those over land or snow (Hu et al., 2018).
@@
 \item
-  Radar observations are used to perform a qualitative and visual assessment of the convection features.
+  Radar observations are used to perform a qualitative and visual assessment of the convective features.
 The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network (de Elía et al., 2017) with a temporal frequency of 10 minutes.
@@
 \[\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}\]
-
 where \(M\) is the ensemble size.
@@
 \[\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)\]
-
 where \(O\) and \(X\) stand for independent observations and the simulations respectively, and N is the sample size.
@@
 \[\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} \]
-
-Where \(P_{oi}\) is the fraction of grid points in the \(i-th\) sampling area in which the observed accumulated precipitation is greater than a specified threshold.
+where \(P_{oi}\) is the fraction of grid points in the \(i-th\) sampling area in which the observed accumulated precipitation is greater than a specified threshold.
 Following Roberts et al.
@@
 
+\leavevmode\hypertarget{ref-goncalvesdegoncalves2015}{}%
+Goncalves de Goncalves, L.G., Sapucci, L., Vendrasco, E., de Mattos, J.G., Ferreira, C., Khamis, E., Cruz, N., 2015.
+A rapid update data assimilation cycle over South America using 3DVar and EnKF, in: The 20th International TOVS Study Conference (ITSC-20).
+The 20th International TOVS Study Conference (ITSC-20), Lake Geneva, Wisconsin, USA.
+
 \leavevmode\hypertarget{ref-grell2013}{}%
@@
 
-\leavevmode\hypertarget{ref-goncalvesdegoncalves2015}{}%
-gustavo Goncalves de Goncalves, L., Sapucci, L., Vendrasco, E., de Mattos, J.G., Ferreira, C., Khamis, E., Cruz, N., 2015.
-A rapid update data assimilation cycle over South America using 3DVar and EnKF, in: The 20th International TOVS Study Conference (ITSC-20).
-The 20th International TOVS Study Conference (ITSC-20), Lake Geneva, Wisconsin, USA.
-
 \leavevmode\hypertarget{ref-ha2014}{}%

### 2022-07-21 analysis/paper/paper_ana_20181122.Rmd
review 1

@@
       base_format: rticles::elsevier_article
+      citation_package: natbib
 abstract: |
@@
 The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
-The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign.
-We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
+The MCS developed during the Intense Observing Period (IOP) of the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign.
+We used the GSI-4DLETKF data assimilation package to produce analyses by assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
+We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
 We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
@@
 bibliography: 
-  - paper_ANA.bib 
+  - paperANA.bib 
   - packages.bib
@@
 #  - \usepackage{chngcntr}
-#  - \usepackage{natbib}
+  - \usepackage{natbib}
   - \usepackage{lineno}
@@
 The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs), has also been investigated in the context of limited-area mesoscale DA.
-Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.,  @wu2014, @cherubini2006, and @sawada2019, among many others).
+Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g.,  @wu2014, @cherubini2006, and @sawada2019, and many others).
 Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm.
@@
 As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
-@zhao2021 and @zhao2021a assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
+\citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
 They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts.
@@
 
-The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW, @skamarock2008) model. 
+The forecast model uses the non-hydrostatic Advanced Research version of Weather Research and Forecasting (WRF-ARW V3.9.1, @skamarock2008). 
 The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa. 
@@
 
-The analyses are generated using the LETKF implementation (V1.3, @hunt2007) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @shao2016). 
+The analyses are generated using the LETKF implementation (V1.3, @hunt2007) of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; @shao2016). 
 A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within $\pm$ 30 minutes of the analysis time are assimilated. 
@@
 
-In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system.
-We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
+We used multi-physics scheme to better represent the uncertainty in the model formulation within the DA system.
+We generated 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
 All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
@@
 
-The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
+The channels adopted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
@@
 
-$$\mathit{BT_{bc}} =\mathit{ BT} + \sum_{i = 0}^{N} \beta_i p_i (x)$$
+\begin{equation}
+  \mathrm{\mathit{BT_{bc}} =\mathit{ BT} + \sum_{i = 0}^{N} \beta_i p_i (x)}
+  (\#eq:eq1)
+\end{equation}
 
@@
 
-To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:
-
-$$RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}$$
-
-where $x_o$ is the assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble. 
-The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
-
-$$\mathit{mean RCRV} = E[RCRV]$$
+To evaluate the statistical consistency of the uncertainty quantification in the ensemble system we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:
+
+\begin{equation}
+  \mathrm{RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}}
+  (\#eq:eq2)
+\end{equation}
+
+where $x_o$ is a assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble. 
+The average of $RCRV$ computed over all the realizations or analysis cycles, represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
+
+\begin{equation}
+  \mathrm{\mathit{mean RCRV} = E[RCRV]}
+  (\#eq:eq3)
+\end{equation}
 
@@
 
-$$\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}$$
+\begin{equation}
+  \mathrm{\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}}
+  (\#eq:eq4)
+\end{equation}
+
 where $M$ is the ensemble size.
@@
 If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative.
-A $sd RCRV > 1$ () indicates that the ensemble is underdispersive and an $sd RCRV < 1$ indicates that the ensemble is overdispersive.
+A $sd RCRV > 1$ indicates that the difference between the observations and the ensemble mean is greater than combined uncertainty of the observations and the model, hence, the ensemble is underdispersive.
+On the contrary, an $sd RCRV < 1$ will indicate that the ensemble is overdispersive [@candille2007].
 
@@
 
-$$\mathit{RMSE} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)^{2}}$$
-
-$$\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)$$
+\begin{equation}
+  \mathrm{\mathit{RMSE} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)^{2}}}
+  (\#eq:eq5)
+\end{equation}
+
+\begin{equation}
+  \mathrm{\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)}
+  (\#eq:eq6)
+\end{equation}
+
 where $O$ and $X$ stand for independent observations and the simulations respectively, and N is the sample size.
@@
 
-$$\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} $$
+\begin{equation}
+  \mathrm{\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}}}
+  (\#eq:eq7)
+\end{equation}
+
 where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold.
@@
   tag_facets(tag_pool = c("c")) +
-  labs(x = NULL, y = "mean RCRV / sd RCRV",
+  labs(x = NULL, y = latex2exp::TeX("mean RCRV (----) / sd RCRV (――)"),
        linetype = NULL, color = "Satellite-derived\nwind") +

### 2022-07-21 analysis/paper/paper_ana_20181122.Rmd
Figures

@@
 
-(ref:cycle) Figure 4.
-Diagram of the analysis cycles between 18 UTC Nov 20, and 12 UTC Nov 23 plus spin up period of 6 hours.
+(ref:cycle) Diagram of the analysis cycles between 18 UTC Nov 20, and 12 UTC Nov 23 plus spin up period of 6 hours.
 The zoomed section shows the hourly assimilation that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
@@
 
-(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $gkg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+(ref:TQ-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in $K$) and specific humidity (d, e, and f in $g\ kg^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
 
@@
   scale_date(ini = 20181121000000, break_bin = "12 hours") +
-  labs(fill = latex2exp::TeX("Specific \nhumidity ($gKg^{-1}$)"),
+  labs(fill = latex2exp::TeX("Specific \nhumidity ($g$ $Kg^{-1}$)"),
        y = "Pressure (hPa)",
@@
 
-(ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $ms^{-1}$) and v wind (d, e, and f in $ms^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-Black contours correspond to negative u wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD.
+(ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $m\ s^{-1}$) and v wind (d, e, and f in $m\ s^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+Black contours correspond to u wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD.
 Dashed contours correspond to negative v wind.
@@
                     size = 2, stroke = 0.1) +
-  labs(fill = latex2exp::TeX("U ($ms^{-1}$)"),
+  labs(fill = latex2exp::TeX("U ($m$ $s^{-1}$)"),
        x = NULL,
@@
                     size = 2, stroke = 0.1) +
-  labs(fill = latex2exp::TeX("V ($ms^{-1}$)"),
+  labs(fill = latex2exp::TeX("V ($m$ $s^{-1}$)"),
        x = NULL,
@@
 Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear.
-The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$.
+The values of MCAPE in CONV do not exceed 2000 $J\ Kg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $J\ Kg^{-1}$.
 MCAPE in the RAD experiment is lower compared to AWS or SATWND.
 This is consistent with less humidity in the PBL with respect to these experiments but may be partially compensated by a slightly warmer PBL in the RAD experiment.
-The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values.
+The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 $m\ s^{-1}$ at the southern tip of the region with positive MCAPE values.
 Moreover, in this same region, these experiments show larger MCAPE values than CONV.
-Note that wind shear over 15 $ms^{-1}$ is associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowski2010].
-
-(ref:summary-fields) a) Precipitable water (shaded, $kgm^{-2}$) and average northerly wind over the first 7 sigma levels (from the surface up to approximately 800 hPa, contours, $ms^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels), and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $ms^{-1}$ for each experiment.
+Note that wind shear over 15 $m\ s^{-1}$ is associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowski2010].
+
+(ref:summary-fields) a) Precipitable water (shaded, $kg\ m^{-2}$) and average northerly wind over the first 7 sigma levels (from the surface up to approximately 800 hPa, contours, $m\ s^{-1}$), b) Average potential temperature for the PBL (first 10 sigma levels), and c) Maximum CAPE and ~0-6 km wind shear over 15 and 30 $m\ s^{-1}$ for each experiment.
 All fields correspond to the analysis ensemble mean for 00 UTC Nov 22.
@@
                         labels = c("-10", "-5")) +
-  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
+  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kg$ $m^{-2}$)"), size = latex2exp::TeX("V ($m$ $s^{-1})")) +
   cordillera +
@@
                         labels = c("15", "30")) +
-  labs(fill = latex2exp::TeX("MCAPE ($JKg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($ms^{-1}$)")) +
+  labs(fill = latex2exp::TeX("MCAPE ($J$ $Kg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($m$ $s^{-1}$)")) +
   cordillera +
@@
 (ref:pphov) Hövmoller diagram of probability matched mean hourly accumulated 1-h forecast precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67$^{\circ}$W and 54.5$^{\circ}$W.
-Contours drawn every 0.5 $mmh^{-1}$, starting at 0.5 $mmh^{-1}$.
+Contours drawn every 0.5 $mm\ h^{-1}$, starting at 0.5 $mm\ h^{-1}$.
 
@@
   tag_facets(position = list(x = 0.1, y = 0.96)) +
-  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Accumulated\nprecipitation ($mmh^{-1}$)")) +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Accumulated\nprecipitation ($mm$ $h^{-1}$)")) +
   theme_minimal(base_size = 9) +
@@
   tag_facets(tag_pool = c("b", "c", "d", "e"), position = list(x = 0.1, y = 0.96)) +
-  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Accumulated\nprecipitation ($mmh^{-1}$)")) +
+  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Accumulated\nprecipitation ($mm$ $h^{-1}$)")) +
   theme_minimal(base_size = 9) +
@@
   .[w %in% c(1, 11) & q %in% c(1, 25) & exp %in% c("E2", "E5", "E6", "E9")] %>% 
-  .[, ":="(q_label = factor(q, labels = c("1~mmh^{-1}", "25~mmh^{-1}")),
+  .[, ":="(q_label = factor(q, labels = c("1~mm~h^{-1}", "25~mm~h^{-1}")),
            w_label = factor(w, labels = c("10~km", "100~km")))] %>% 
@@
 
-(ref:soundings) RMSE (solid line) and Bias (dashed line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($ms^{-1}$) and d) v wind ($ms^{-1}$) calculated by comparing the analysis of each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+(ref:soundings) RMSE (solid line) and Bias (dashed line) of a) temperature ($K$), b) dew point temperature ($K$), c) u wind ($m\ s^{-1}$) and d) v wind ($m\ s^{-1}$) calculated by comparing the analysis of each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
 The blue line corresponds to CONV, the light blue line to AWS, SATWND is represented with an orange line, and RAD with a red line.
@@
            variable_labels = factor(variable, labels = c("Temperature~(K)", "Dew~point~temperature~(K)", 
-                                                         "U~wind~(ms^{-1})", "V~wind~(ms^{-1})")))] %>% 
+                                                         "U~wind~(m~s^{-1})", "V~wind~(m~s^{-1})")))] %>% 
   ggplot(aes(lev/1000, value.est)) +
@@
   .[q %in% c(1, 25) & w %in% c(1, 11)] %>% 
-  .[, ":="(q_label = factor(q, labels = c("1~mmh^{-1}", "25~mmh^{-1}")),
+  .[, ":="(q_label = factor(q, labels = c("1~mm~h^{-1}", "25~mm~h^{-1}")),
            w_label = factor(w, labels = c("10~km", "100~km")))] %>% 
@@
                                                       E6 = "SATWND", E9 = "RAD")) +
-  scale_linetype_manual(values = c(1, 2), labels = c("2018-11-22 00:00:00" = "OO UTC", 
+  scale_linetype_manual(values = c(1, 2), labels = c("2018-11-22 00:00:00" = "00 UTC", 
                                    "2018-11-22 06:00:00" = "06 UTC")) +

### 2022-07-26 analysis/paper/paper_ana_20181122.Rmd
Posible weather field y otros

@@
 For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used.
-
 In that regard, several authors investigated the impact of assimilating surface weather data (e.g.
@@
 
-While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, bias correction, and the relatively low model tops used for this application.
+While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting sateññite observations), bias correction, and the relatively low model tops used for this application.
 @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
@@
 
-## Case Study
-
-On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
+## Case overview
+
+On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)a) that rapidly grew upscale into an exceptionally large MCS.
+During that day several surface stations reported lightning, gustfronts and heavy rain.
+To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS (Figure \@ref(fig:case)c).
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
+
+(ref:case) GOES-16 infrared imagery corresponding to channel 13 (13 mu) for a) 00 UTC, b) 12 UTC Nov 22, and c) 00 UTC Nov 23.
+
+```{r case, fig.cap="(ref:case)"}
+
+campos <- ReadNetCDF(here("analysis/data/derived_data/reanalysis/geopotential.nc")) %>% 
+  dcast(time + longitude + latitude ~ level) %>% 
+  .[order(time, -latitude, longitude)] %>% 
+  .[, ":="(msl = ReadNetCDF(here("analysis/data/derived_data/reanalysis/pressure-pw.nc"), vars = "msl", out = "vector")[[1]],
+           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]],
+           tcw = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw", out = "vector")[[1]],
+           tcwv = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcwv", out = "vector")[[1]],
+           td = ReadNetCDF(here("analysis/data/derived_data/reanalysis/td2m.nc"), vars = "d2m", out = "vector")[[1]])]
+
+campos <-  ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw")
+
+campos %>%
+
+  .[time != ymd_h(2018112312)] %>% 
+  .[, ":="(espesor = (`500`-`1000`)/100, time = factor(time))] %>%
+  ggplot(aes(longitude, latitude)) +
+  geom_contour_fill(aes(z = cape, fill = stat(level_d)), breaks = seq(500, 4000, 500)) +
+  scale_fill_distiller(super = ScaleDiscretised,
+                       name = "CAPE",
+                       # palette = "PRGn", direction = 1,
+                       palette = "YlOrRd", direction = 1,
+                       guide = guide_colorsteps(barwidth = 25, mid = 25,
+                                                barheight = 0.5, 
+                                                title.position = "left", 
+                                                title.vjust = 1,
+                                                frame.colour = "black")) +
+  geom_mapa() +
+  geom_contour2(aes(z = espesor, label = ..level..),
+                size = 0.2, color = "darkred", linetype = 2,
+                label_size = 2, label.placer = label_placer_n(n = 3)) +
+  # label_size = 2, label.placer = label_placer_fraction(frac = 0.75)) +
+  geom_contour2(aes(z = msl/100, label = ..level..),
+                size = 0.4, color = "black", linetype = 1,
+                label_size = 2) +
+  cordillera +
+  coord_sf(ylim = c(-45, -22), xlim = c(-75, -52)) +
+  facet_wrap(vars(time), labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
+                                                      "2018-11-22 12:00:00" = "12 UTC Nov 22",
+                                                      "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
+  labs(x = NULL, y = NULL, fill = "CAPE" ) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+```
 
@@
 To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations.
-A relaxation-to-prior inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
+A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
 
@@
 Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is largest; the systematic impact is overall smaller than the one produced by assimilating data from AWS  (Figures  \@ref(fig:UV-diff)b, e).
+The reason of the small impact observed in SATWND could be asociated to the large observation error used for satellite-derived wind observatios.
 
@@
 (ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $m\ s^{-1}$) and v wind (d, e, and f in $m\ s^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-Black contours correspond to u wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD.
-Dashed contours correspond to negative v wind.
+Black contours correspond to u wind and aashed contours to negative v wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD since those experiments are the ones with more assimilated observations in each panel.
 

### 2022-07-29 analysis/paper/paper_ana_20181122.Rmd
Boundary conditions

@@
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
-This method helps to prevent an underestimation of the ensemble spread [@ouaraini2015], and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.
+This method helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
+All experiments presented here uses the same boundary conditions from GFS, generated using the same method as for the inicial conditions to maintain proper levels of ensemble spread within the domain.
 
@@
 The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime.
-The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@ruiz2010].
+The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@casaretto2022, @matsudo2021, @ruiz2010].
 We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL.
@@
 
-To quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)).
+We computed the FSS to quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments (Figure \@ref(fig:fss)).
+For each threshold and spatial scale, we applied Equation \@red(eq:eq7) in 6-hours rolling windows throughout the experiment period.
 All experiments show similar values of FSS during the initiation of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
@@
 In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
-We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
+We calculated the FSS for the ensemble forecasts in 6-hour rollig windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
 CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.

### 2022-08-08 analysis/paper/paper_ana_20181122.Rmd
case overview, smothed conours and others

@@
 scale_date <- function(ini = 20181120180000, end = 20181123120000,
-                       break_bin = "6 hour") {
-  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
+                       break_bin = "6 hour", ...) {
+  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin), ..., expand = c(0, 0),
                    labels = function(x) {
@@
 
-For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used.
+For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used (@gustafsson2018).
 In that regard, several authors investigated the impact of assimilating surface weather data (e.g.
@@
 
-On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)a) that rapidly grew upscale into an exceptionally large MCS.
+Previoursly to the development of this case study, the center and north of Argentina was inmersed in a warm and humid air mass with high values of convective available potential energy (CAPE) (Figure \@ref(fig:case)a).
+On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)b) that rapidly grew upscale into an exceptionally large MCS.
 During that day several surface stations reported lightning, gustfronts and heavy rain.
@@
 
-(ref:case) GOES-16 infrared imagery corresponding to channel 13 (13 mu) for a) 00 UTC, b) 12 UTC Nov 22, and c) 00 UTC Nov 23.
+(ref:case) ERA5 Reanalysis of pressure (hPa) at sea level (black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) for 00 and 12 UTC Nov 22 and 00 UTC Nov 23.
 
@@
 
-campos <-  ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw")
+# campos <-  ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw")
 
@@
                                                       "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
+  tag_facets() +
   labs(x = NULL, y = NULL, fill = "CAPE" ) +
@@
 AIRCAR presents an irregular profile with $sd RCRV$ values that suggest that the error for this type of observation is overestimated.
-ADPUPA and AIRCAR present a $mean RCRV$ profile near zero at middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dillon2021.
+ADPUPA and AIRCAR present a $mean RCRV$ profile near zero at middle and upper levels.
+At low levels, the $mean RCRV$ profile is positive, showing a cold bias present in the model, a characteristic already studied in @ruiz2010 and @dillon2021.
 
@@
   .[pw, on = .NATURAL] %>% 
+  .[, v_mean_suave := metR:::smooth2d(x, y, v_mean, kx = 0.2, ky = 0.2), by = .(exp)] %>% 
   ggplot(aes(x, y)) +
@@
                                                                         frame.colour = "black")) +
-  geom_contour2(aes(z = v_mean, size = stat(level)),
+  geom_contour2(data = ~.[x %between% c(-950000, 950000)], aes(z = v_mean_suave, size = stat(level)),
                 proj = norargentina_lambert, color = "black", 
                 breaks = c(-10, -5)) +
-  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
+  scale_size_continuous(breaks = c(-10, -5), range = c(0.6, 0.4),
                         labels = c("-10", "-5")) +
@@
         strip.text = element_blank(), 
-        strip.background = element_blank()) +
+        strip.background = element_blank())
 
@@
   .[cor, on = .NATURAL] %>% 
+  .[, cortante_suave := metR:::smooth2d(x, y, cortante, kx = 0.2, ky = 0.2), by = .(exp)] %>% 
   ggplot(aes(x, y)) +
@@
                                                 frame.colour = "black")) +
-  geom_contour2(aes(z = cortante, size = stat(level)),
+  geom_contour2(data = ~.[x %between% c(-950000, 950000) &
+                           y %between% c(-1170000, 1170000)], aes(z = cortante_suave, size = stat(level)),
                 proj = norargentina_lambert, color = "black", 
                 breaks = c(15, 30)) +
-  scale_size_continuous(breaks = c(15, 30), range = c(0.8, 0.4),
+  scale_size_continuous(breaks = c(15, 30), range = c(0.6, 0.4),
                         labels = c("15", "30")) +
@@
 In addition, CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
-Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between experiment and observations is quite good in the regions where radar data are available, especially for RAD.
+Before and after the times shown in Figure \@ref(fig:dbz-mean), the agreement between location of the observed convective cells and the simulated in the experiment is quite good in the regions where radar data are available, especially for RAD.
 
@@
 AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias.
-However, AWS degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
+However, AWS increments (Figure \@red(fig:UV-diff)d) degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
 
@@
 In particular, we explore the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors.
-Southern South America is a particularly interesting region due to the heterogeneity and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
+Southern South America is a particularly interesting region due to the heterogeneity in topography and weather and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
 This combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.

### 2022-08-08 analysis/paper/paper_ana_20181122.Rmd
RSRV para satelite

@@
 
-For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) at low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error.
+Satellite-derived winds observations varies in number depending on the satellite and the level.
+In Figure \@ref(fig:rcrv-profile)c we included the $RCRV$ calculated at least with 100 observations for each satellite and level.
+At low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error.
 Wind estimations derived from water vapor channels are abundant above 500 hPa where their bias is close to zero.

### 2022-08-09 analysis/paper/paper_ana_20181122.Rmd
Agrega tabla de miembros

@@
 We generated 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
+The distribution of these schemes among the 60 ensemble members is outlined in Table \@ref(tab:miembros-desc).
+The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevamnt physical processes that are not resolved by the model.
 All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
+
+```{r miembros-desc}
+fisica %>% 
+  copy() %>% 
+  .[, c("Cumulus", "PBL") := tstrsplit(fisica, split = "-")] %>% 
+  .[, .(mem = paste(mem, collapse = ", ")), by = .(PBL, Cumulus)] %>% 
+  dcast(Cumulus ~ PBL, value.var  = "mem") %>% 
+  kable(booktabs = TRUE, caption = "Generation of the 60-member RRR multi-physics ensemble as a combination of Cumulus and PBL parameterizations, and GEFS members for boundary conditions.", 
+        align = "cccc") %>% 
+  add_header_above(c(" " = 1, PBL = 3)) %>% 
+  kable_classic_2(full_width = FALSE) %>% 
+  kable_styling(font_size = 6,
+                position = "center") %>% 
+  column_spec(1, )
+```
 

### 2022-08-09 analysis/paper/paper_ana_20181122.Rmd
Agrega algunas figuras

--- 2022-08-08 analysis/paper/paper_ana_20181122.Rmd
@@
 journal: Atmospheric Research
-layout: authoryear,preprint,review,12pt
-#layout: final,5p,times,twocolumn,authoryear
+# layout: authoryear,preprint,review,12pt
+layout: final,5p,times,twocolumn,authoryear
 address:
@@
 
-```{r case, fig.cap="(ref:case)"}
+```{r case, fig.cap="(ref:case)", fig.width=7, fig.height=3.5, fig.align="center", fig.env = "figure*"}
 
@@
   labs(x = NULL, y = NULL, fill = "CAPE" ) +
-  theme_minimal() +
+  theme_minimal(base_size = 8) +
   theme(legend.position = "bottom")
@@
 We generated 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
+The distribution of these schemes among the 60 ensemble members is outlined in Table \@ref(tab:miembros-desc).
+The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevamnt physical processes that are not resolved by the model.
 All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
+
+```{r miembros-desc}
+fisica %>% 
+  copy() %>% 
+  .[, c("Cumulus", "PBL") := tstrsplit(fisica, split = "-")] %>% 
+  .[, .(mem = paste(as.numeric(mem), collapse = ", ")), by = .(PBL, Cumulus)] %>% 
+  dcast(Cumulus ~ PBL, value.var  = "mem") %>% 
+  kable(booktabs = TRUE, caption = "Generation of the 60-member RRR multi-physics ensemble as a combination of Cumulus and PBL parameterizations, and GEFS members for boundary conditions.", 
+        align = "cccc") %>% 
+  add_header_above(c(" " = 1, PBL = 3)) %>% 
+  kable_classic_2(full_width = FALSE) %>% 
+  kable_styling(font_size = 6,
+                position = "center") %>% 
+  # column_spec(1, width = "1em") %>% 
+  column_spec(2:4, width = "7em")
+```
 
@@
 (ref:UV-diff) Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in $m\ s^{-1}$) and v wind (d, e, and f in $m\ s^{-1}$) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
-Black contours correspond to u wind and aashed contours to negative v wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD since those experiments are the ones with more assimilated observations in each panel.
+Black contours correspond to u wind and dashed contours to negative v wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD since those experiments are the ones with more assimilated observations in each panel.
 
@@
 
+(ref:era5-cap) Difference between analysis ensemble mean experiments and ERA5 for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity ($g\ Kg^{-1}$, e--h) and meridional wind ($m\ s^{-1}$, i--l) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+
+```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.height=7, fig.align="center"}
+era <- fread("~/datosmunin3/DATA_EXP/ERA5/perfiles_ana_era5.csv")
+
+files <- Sys.glob(here::here("/home/paola.corrales/datosmunin3/EXP//derived_data/perfiles_ana_E[2569]_new.csv"))
+
+exp <- purrr::map(files, function(f) {
+  exp <- unglue(basename(f), "perfiles_{run}_{exp}_new.csv")
+  fread(f) %>% 
+    .[, exp := exp[[1]][["exp"]]]
+
+}) %>% 
+  rbindlist() %>% 
+  .[, date := as_datetime(date)] %>% 
+  .[date %between% c(as_datetime("2018-11-20 18:00:00"),
+                     as_datetime("2018-11-23 12:00:00"))]
+
+era[exp, on = c("lev", "date")] %>% 
+  .[lev != 1000] %>%
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = T - i.T, fill = stat(level)), 
+                    breaks = seq(-3, 3, 0.5)) +
+  geom_contour2(aes(z = T - i.T), color = "white", size = 0.1,
+                breaks = seq(-3, 3, 0.5)) +
+  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
+                                                            barheight = 0.3,
+                                                            title.position = "left", 
+                                                            title.vjust = 1,
+                                                            frame.colour = "black")) +
+  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
+                                                              E6 = "SATWND", E9 = "RAD"))) +
+  labs(fill = "Temperature (K)",
+       x = NULL,
+       y = "Pressure (hPa)") +
+  tag_facets(position = list(x = 0.1, y = 0.96)) +
+  theme_minimal(base_size = 8) +
+  theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 8),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+era[exp, on = c("lev", "date")] %>% 
+  .[lev != 1000] %>%
+  ggplot(aes(date, lev)) +
+  geom_contour_fill(aes(z = QVAPOR - i.QVAPOR, fill = stat(level)), 
+                    breaks = seq(-0.0015, 0.0025, 0.0004)) +
+  geom_contour2(aes(z = QVAPOR - i.QVAPOR), color = "white", size = 0.1,
+                breaks = seq(-0.0015, 0.0025, 0.0004)) +
+  scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
+                                                            barheight = 0.3,
+                                                            title.position = "left", 
+                                                            title.vjust = 1,
+                                                            frame.colour = "black")) +
+  scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+  scale_date(ini = 20181121000000, break_bin = "12 hours") +
+  facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
+                                                              E6 = "SATWND", E9 = "RAD"))) +
+  labs(fill = latex2exp::TeX("Specific \nhumidity ($g$ $Kg^{-1}$)"),
+       x = NULL,
+       y = "Pressure (hPa)") +
+  tag_facets(position = list(x = 0.1, y = 0.96), tag_pool = c("e", "f", "g", "h")) +
+  theme_minimal(base_size = 8) +
+  theme(legend.position = "bottom",
+        tagger.panel.tag.text = element_text(size = 8),
+        panel.ontop = TRUE,
+        panel.grid = element_line(linetype = 3)) +
+
+  era[exp, on = c("lev", "date")] %>% 
+    .[lev != 1000] %>%
+    ggplot(aes(date, lev)) +
+    geom_contour_fill(aes(z = V - i.V, fill = stat(level)), 
+                      breaks = seq(-7, 4, 1)) +
+    geom_contour2(aes(z = V - i.V), color = "white", size = 0.1,
+                  breaks = seq(-7, 4, 1)) +
+    scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
+                                                              barheight = 0.3,
+                                                              title.position = "left", 
+                                                              title.vjust = 1,
+                                                              frame.colour = "black")) +
+    scale_y_level(name = "Pressure (hPa)", breaks = c(1000, 850, 750, 500, 300, 200, 100)) +
+    scale_date(ini = 20181121000000, break_bin = "12 hours") +
+    facet_wrap(vars(exp), ncol = 4, labeller = labeller(exp = c(E2 = "CONV", E5 = "AWS", 
+                                                                E6 = "SATWND", E9 = "RAD"))) +
+    labs(fill = latex2exp::TeX("V ($m$ $s^{-1}$)"),
+         x = NULL,
+         y = "Pressure (hPa)") +
+    tag_facets(position = list(x = 0.1, y = 0.96), tag_pool = c("i", "j", "k", "l")) +
+    theme_minimal(base_size = 8) +
+    theme(legend.position = "bottom",
+          tagger.panel.tag.text = element_text(size = 8),
+          panel.ontop = TRUE,
+          panel.grid = element_line(linetype = 3)) +
+
+  plot_layout(ncol = 1, heights = c(1, 1, 1)) 
+```
+
 The analyzed distribution of temperature and moisture in the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed in the temperature profiles (Figure \@ref(fig:TQ-diff)a-c) where AWS produces a colder PBL than CONV while the PBL in RAD is warmer than in SATWND.
@@
         strip.text = element_blank(), 
-        strip.background = element_blank())
+        strip.background = element_blank()) +
 

### 2022-08-22 analysis/paper/paper_ana_20181122.Rmd
ERA5 y otros comentarios menores

@@
 output: 
+  # bookdown::word_document2: default
   bookdown::pdf_book: 
@@
       citation_package: natbib
+# always_allow_html: true      
 abstract: |
@@
   - packages.bib
+  - era.bib
 csl: elsevier-harvard.csl
@@
 #  - \usepackage{chngcntr}
-  - \usepackage{natbib}
+#  - \usepackage{natbib}
   - \usepackage{lineno}
-  - \linenumbers 
+#  - \linenumbers 
 ---
@@
 This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014].
-Recently, the RELAMPAGO-CACTI field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
+Recently, the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
 
@@
 
-For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used (@gustafsson2018).
-In that regard, several authors investigated the impact of assimilating surface weather data (e.g.
-@wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016) using different assimilation methodologies.
-Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.
-@sobash2015 showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
-More recently, @gasperoni2018 evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis.
-They found a positive effect of these observations upon the initiation of deep moist convection.
-This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.
-
-The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs), has also been investigated in the context of limited-area mesoscale DA.
+For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used [@gustafsson2018].
+Assimilating information on temperature, moisture and wind in the planetary boundary layer improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts upon the planetary boundary layer structure and the location and timing of precipitating systems (e.g.
+@wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016).
+
+Particularly relevant for regional mesoscale DA systems in the region of interest is that South America is characterized by a limited number of conventional observations (i.e., radiosondes, surface weather stations) and operational networks are not dense enough to capture mesoscale details.
+In this context, analyzing the potential impact of non-conventional sources of observations is essential to improve mesoscale numerical weather prediction (NWP) over South America using DA.
+There have been only a few published efforts on regional mesoscale DA, but they have all shown promising results [e.g.
+@dillon2016; @dillon2021; @goncalvesdegoncalves2015].
+In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts.
+Similarly to @gasperoni2018, @dillon2021 included private weather station networks which are not incorporated in the operational analysis.
+However, the impact of different observation types on the analysis quality has not been addressed.
+
+The impact of non-conventional high spatial and temporal resolution observations, such as AMVs, has been investigated in the context of regional mesoscale DA.
 Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., @wu2014, @cherubini2006, and @sawada2019, and many others).
@@
 
-While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting sateññite observations), bias correction, and the relatively low model tops used for this application.
+While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting satellite observations), bias correction, and the relatively low model tops used for this application.
 @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
@@
 
-In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed.
-Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources [e.g.
-@dillon2016; @dillon2021; @goncalvesdegoncalves2015].
-In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 AMVs, and satellite temperature and moisture retrievals with promising results.
-However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above.
-The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based DA system.
-In particular, we will focus on the impact in the context of mid-latitude mesoscale convective system events.
-Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.
+The main objective of this work is thus to contribute to the quantification and comparison of the impact of high resolution automatic weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based regional DA system.
+This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps.
+In particular, we will focus on the impact in the context of a mid-latitude mesoscale convective system.
 
@@
 
-Previoursly to the development of this case study, the center and north of Argentina was inmersed in a warm and humid air mass with high values of convective available potential energy (CAPE) (Figure \@ref(fig:case)a).
+Previously to the development of this case study, the center and north of Argentina was immersed in a warm and humid air mass with high values of convective available potential energy (CAPE) (Figure \@ref(fig:case)a).
 On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)b) that rapidly grew upscale into an exceptionally large MCS.
@@
   .[, ":="(msl = ReadNetCDF(here("analysis/data/derived_data/reanalysis/pressure-pw.nc"), vars = "msl", out = "vector")[[1]],
-           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]],
-           tcw = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw", out = "vector")[[1]],
-           tcwv = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcwv", out = "vector")[[1]],
-           td = ReadNetCDF(here("analysis/data/derived_data/reanalysis/td2m.nc"), vars = "d2m", out = "vector")[[1]])]
-
-# campos <-  ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "tcw")
+           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]])]
 
@@
                 label_size = 2, label.placer = label_placer_n(n = 3)) +
-  # label_size = 2, label.placer = label_placer_fraction(frac = 0.75)) +
   geom_contour2(aes(z = msl/100, label = ..level..),
@@
 The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1).
+“In this case a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km aproximately to 10 km).
+This approach is also based in recent studies which suggest that using multiple nested domains do not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g.
+@liang2019, @beck2004).
 The domain covers the area indicated in Figure \@ref(fig:dominio) to capture the development of the MCS during the simulated period.
@@
 
-We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis with random perturbations to generate the initial ensemble perturbations.
+We use a 60-member ensemble where the initial ensemble mean and the mean boundary conditions are taken from the GFS deterministic analysis.
+A set of 60 perturbations are  randomly generated to perturb the initial state as well as the  boundary conditions during the length of the experiment.
+Perturbing the boundary conditions helps to reduce the impact of errors in the driving global model and helps to keep a larger ensemble spread throughout the domain and during the length of the experiment [@ouaraini2015].
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021.
 In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
-This method helps to prevent an underestimation of the ensemble spread [@ouaraini2015].
-All experiments presented here uses the same boundary conditions from GFS, generated using the same method as for the inicial conditions to maintain proper levels of ensemble spread within the domain.
+The random perturbations are the same across different experiments to ensure that the differences between experiments are only related to changes in the number and type of assimilated observations.
 
@@
 To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations.
-A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
+A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ following @maldonado2020 to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
 
@@
 For the infrared channels, cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
+Moreover, GSI checks the difference between the observations and simulated brightness temperature with height to detect cloudy pixels.
 Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenith angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
@@
 
+* ERA5 hourly data on pressure levels from 1959 to present [@era5pressure].
+The variables of interest (air temperature, humidity and wind) were interpolated to the model grid to compare them with the analysis of each experiment.
+
 * The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the IOPs along with the operational radiosondes.
@@
 
-where $x_o$ is a assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble. 
-The average of $RCRV$ computed over all the realizations or analysis cycles, represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
+where $x_o$ is the assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble.
+The $RCRV$ is the ratio between the distance between the observations and the forecast and its expected standard deviation assuming the statistical independence between the forecast error (estimated from the ensemble spread) and the observation error.
+The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
 
@@
 
-The standard deviation of the $RCRV$ or $sd RCRV$ measures the agreement of the ensemble spread and the observational error with respect to the distance between the ensemble mean and the observations, and then the systematic over- or under- dispersion of the ensemble:
+If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative.
+The standard deviation of the $RCRV$ or $sd RCRV$ is defined as:
 
 \begin{equation}
-  \mathrm{\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}}
+  \mathrm{\mathit{sd RCRV} = \sqrt{\frac{1}{M -1}\sum_{i=1}^{M}(\mathit{RCRV_i} - \mathit{mean RCRV})^2}}
   (\#eq:eq4)
@@
 where $M$ is the ensemble size.
+The $sd RCRV$ measures how large is the distance between the forecast and the observations with respect to the expected distance (given by the combination of the ensemble spread and the observation error).
+Assuming that the observation error is accurately estimated, an $sd RCRV > 1$ indicates that the ensemble is underdispersive (i.e.
+the distance between the observations and the forecasts is larger than expected) and an $sd RCRV < 1$ indicates that the ensemble is overdispersive (i.e.
+the distance between the observations and the forecasts is lower than expected).
 A consistent system will have no bias ($mean RCRV = 0$) and a standard deviation equal to 1 ($sd RCRV = 1$).
-If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative.
-A $sd RCRV > 1$ indicates that the difference between the observations and the ensemble mean is greater than combined uncertainty of the observations and the model, hence, the ensemble is underdispersive.
-On the contrary, an $sd RCRV < 1$ will indicate that the ensemble is overdispersive [@candille2007].
 
@@
 For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas in the west of the domain the $sd RCRV$ can be as high as 4.5.
-In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.
+These higher values of $sdRCRV$ can be associated with systematic errors arising from heigh differences between the model surface and the observations.
+Small scale circulations associated with the complex terrain and not well resolved by the model can also contribute to increase the distance between the forecast and the observations.
+These aspects are usually not captured by the ensemble spread unless a well tuned space dependent inflation scheme is used thus leading to greater sdRCRV values.
 
@@
 Overall, these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
-Moreover, these results suggest the inflation parameter $\alpha = 0.9$ is adequate for the system.
+Moreover, these results suggest the relaxation-to-prior spread inflation inflation parameter $\alpha = 0.9$ is adequate for the system.
 
@@
 ```
-
-To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles).
-At that moment the first convective cells were developing over the southern region of the domain along the cold front.
-Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
-We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
-The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
-AWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
-RAD has lower precipitable water content than AWS and SATWND, but higher than CONV.
-The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating AWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
+We also compared ensemble mean analyses to ERA5 reanalyses [@era5pressure], which suports Figures
+\@ref(fig:TQ-diff) and \@ref(fig:UV-diff).
+Specifically, Figure \@ref(fig:era5)a shows a warm bias in low levels (i.e.
+CONV is warmer than ERA5) that decreases in Figure \@ref(fig:era5)b when we assimilate the AWS observations.
+In the same direction, Figure \@ref(fig:TQ-diff)a shows a negative difference between AWS and CONV meaning that the AWS observations are cooling the low levels.
+Comparing ERA5 with RAD (Figure \@ref(fig:era5)d), there is a see a small increase in the warm bias, associated with the warming produced by the radiance observations (Figure \@ref(fig:TQ-diff)c).
+A similar effect can be observed for specific humidity, AWS observations parcialy corrects the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces this effect.
+The impact on the wind components is minor and only the meridional wind is included in Figure \@ref(fig:era5).
+In this case Figures \@ref(fig:era5)i-l show that the radiance observations are mainly responsible for the positive impact observed in the analysis.
+Moreover, these adjustments due to assimilating radiance and AWS observations lead to ensemble mean analyses closer to ERA5 reanalyses.
 
@@
 
-```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.height=7, fig.align="center"}
-era <- fread("~/datosmunin3/DATA_EXP/ERA5/perfiles_ana_era5.csv")
-
-files <- Sys.glob(here::here("/home/paola.corrales/datosmunin3/EXP//derived_data/perfiles_ana_E[2569]_new.csv"))
-
-exp <- purrr::map(files, function(f) {
-  exp <- unglue(basename(f), "perfiles_{run}_{exp}_new.csv")
-  fread(f) %>% 
-    .[, exp := exp[[1]][["exp"]]]
-
-}) %>% 
-  rbindlist() %>% 
-  .[, date := as_datetime(date)] %>% 
-  .[date %between% c(as_datetime("2018-11-20 18:00:00"),
-                     as_datetime("2018-11-23 12:00:00"))]
-
-era[exp, on = c("lev", "date")] %>% 
+```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.width=7, fig.height=6, fig.align="center"}
+era <- fread(here("analysis/data/derived_data/reanalysis/perfiles_ana_era5.csv"))
+
+era[perfiles, on = c("lev", "date")] %>% 
   .[lev != 1000] %>%
@@
 ```
+To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles).
+At that moment the first convective cells were developing over the southern region of the domain along the cold front.
+Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
+We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
+The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
+AWS and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown).
+RAD has lower precipitable water content than AWS and SATWND, but higher than CONV.
+The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating AWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.
 
@@
 On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
+A warmer PBL increases the potencial inestability and helps to generate a suitable enviroment for the development of deep convection.
 Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear.
@@
 AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias.
-However, AWS increments (Figure \@red(fig:UV-diff)d) degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
+However, in this region (Figure \@ref(fig:domain)b) and for this period, AWS increments (Figure \@ref(fig:UV-diff)d) degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
 
@@
 Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the 1-h forecast to IMERG estimations.
+The reason for why the forecasts inicialized from RAD degrades over time needs to be further study.
 
@@
 Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events.
-In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like  GPS radio occultation data and radiances from geostationary orbiting satellites.
+In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like GPS radio occultation data and radiances from geostationary orbiting satellites like GOES-16.
 

### 2022-08-23 _dependencies.R
renv

@@ -2,3 +2,4 @@ library(rticles)
 library(rnaturalearthhires)
 library(rgeos)
 library(udunits2)
+library(rgdal)

### 2022-08-23 R/probability_matching.R
scripts

@@ -141,7 +141,7 @@ calculate_probability_marching_dbz <- function(exp, run, dates,
                                               out_path = "/home/paola.corrales/datosmunin3/EXP/derived_data/dbz/") {
 
   library(reticulate)
-  use_condaenv("/home/jruiz/share/anaconda3/envs/wrfpy")
+  use_condaenv("/home/paola.corrales/.local/share/r-miniconda/envs/r-reticulate")
   py_run_string("import os")
   py_run_string("os.environ['PROJ_LIB'] = '/home/jruiz/share/anaconda3/share/proj'")
   wrf <- import("wrf")

### 2022-08-23 analysis/paper/paper_ana_20181122.Rmd
paper and biblio

@@
   address: Department of Atmospheric Sciences, University of Illinois Urbana–Champaign, Urbana, Illinois.
-bibliography: 
-  - paperANA.bib 
-  - packages.bib
-  - era.bib
+bibliography: [paperANA.bib, packages.bib, era.bib]
 csl: elsevier-harvard.csl
@@
 #  - \usepackage{natbib}
-  - \usepackage{lineno}
+#  - \usepackage{lineno}
 #  - \linenumbers 
@@
   theme(legend.position = "bottom", 
+        panel.background = element_rect(fill = "#fbfbfb", color = NA),
         plot.margin = unit(c(0, 0, 0, 0), "cm")) +
@@
         legend.margin = unit(c(0, 0, 0, 0), "cm"),
+        panel.background = element_rect(fill = "#fbfbfb", color = NA),
         plot.margin = unit(c(0, 0, 0, 0), "cm"))
@@
   theme(legend.box = "vertical",
-        # legend.spacing.y = unit(0.1, 'mm'),
         legend.title = element_text(size = 6),
+        panel.background = element_rect(fill = "#fbfbfb", color = NA),
         tagger.panel.tag.text = element_text(size = 8)) +
@@
         legend.title = element_text(size = 6),
+        panel.background = element_rect(fill = "#fbfbfb", color = NA),
         tagger.panel.tag.text = element_text(size = 8)) +
@@
         legend.title = element_text(size = 6),
+        panel.background = element_rect(fill = "#fbfbfb", color = NA),
         tagger.panel.tag.text = element_text(size = 8))
@@
 
-```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.width=7, fig.height=6, fig.align="center"}
+```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.width=7.2, fig.height=5, fig.align="center"}
 era <- fread(here("analysis/data/derived_data/reanalysis/perfiles_ana_era5.csv"))
@@
        y = "Pressure (hPa)") +
-  tag_facets(position = list(x = 0.1, y = 0.96)) +
+  tag_facets(position = list(x = 0.1, y = 0.95)) +
   theme_minimal(base_size = 8) +
@@
         tagger.panel.tag.text = element_text(size = 8),
+        plot.margin = margin(0, 0, 0, 0, "cm"),
         panel.ontop = TRUE,
@@
 
-era[exp, on = c("lev", "date")] %>% 
+era[perfiles, on = c("lev", "date")] %>% 
   .[lev != 1000] %>%
@@
        y = "Pressure (hPa)") +
-  tag_facets(position = list(x = 0.1, y = 0.96), tag_pool = c("e", "f", "g", "h")) +
+  tag_facets(position = list(x = 0.1, y = 0.95), tag_pool = c("e", "f", "g", "h")) +
   theme_minimal(base_size = 8) +
@@
 
-  era[exp, on = c("lev", "date")] %>% 
+  era[perfiles, on = c("lev", "date")] %>% 
     .[lev != 1000] %>%
@@
          y = "Pressure (hPa)") +
-    tag_facets(position = list(x = 0.1, y = 0.96), tag_pool = c("i", "j", "k", "l")) +
+    tag_facets(position = list(x = 0.1, y = 0.95), tag_pool = c("i", "j", "k", "l")) +
     theme_minimal(base_size = 8) +
@@
           tagger.panel.tag.text = element_text(size = 8),
+          plot.margin = margin(0, 0, 0, 0, "cm"),
           panel.ontop = TRUE,
@@
   theme_minimal(base_size = 8) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        panel.background = element_rect(fill = "#fbfbfb", color = NA))
 ```
@@
   theme_minimal(base_size = 9) +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        panel.background = element_rect(fill = "#fbfbfb", color = NA))
 ```
@@
         legend.box = "vertical", 
-        legend.margin = margin())
+        legend.margin = margin(0, 0, 0, 0),
+        panel.background = element_rect(fill = "#fbfbfb", color = NA))
 

### 2022-08-23 _dependencies.R
Actualiza dependencias

@@ -3,3 +3,4 @@ library(rnaturalearthhires)
 library(rgeos)
 library(udunits2)
 library(rgdal)
+library(magick)

### 2022-08-23 analysis/paper/paper_ana_20181122.Rmd
Actualiza al nuevo elsevier

@@
 journal: Atmospheric Research
-# layout: authoryear,preprint,review,12pt
-layout: final,5p,times,twocolumn,authoryear
+#layout: authoryear,preprint,review,12pt
+# layout: final,5p,times,twocolumn,authoryear
+classoption: preprint, 3p, authoryear
 address:
@@
     address: Department of Atmospheric Sciences, University of Illinois Urbana–Champaign, Urbana, Illinois.
-bibliography: [paperANA.bib, packages.bib, era.bib]
-csl: elsevier-harvard.csl
+bibliography: "paperANA.bib,packages.bib,era.bib"
+# csl: elsevier-harvard.csl
 footnote:
@@
 numbersections: true
-header-includes: 
+# header-includes: 
 #  - \usepackage{gensymb}
@@
            plataforma = toupper(plataforma),
-           prop = paste0(prop, " \\", "%"))] %>% 
+           prop = paste0(prop, " ", "%"))] %>%
   .[order(.[,'sensor']), ] %>%
@@
   kbl(booktabs = TRUE,
-      escape = FALSE,
-      col.names = linebreak(c("Sensor", "Platform", "Assimilated\nchannels", "Percentage\nover total")),
+      escape = TRUE,
+      col.names = linebreak(c("Sensor", "Platform", "Assimilated channels", "Percentage over total")),
       caption = "List of the available sensors over several platforms, the number of accepted channels for the assimilation, and the percentage of assimilated observations calculated over all radiance observations and all cycles.") %>%
@@
 AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias.
-However, in this region (Figure \@ref(fig:domain)b) and for this period, AWS increments (Figure \@ref(fig:UV-diff)d) degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
+However, in this region (Figure \@ref(fig:dominio)b) and for this period, AWS increments (Figure \@ref(fig:UV-diff)d) degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
 

### 2022-08-24 analysis/paper/paper_ana_20181122.Rmd
Chau we

@@
 abstract: |
-  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse.
+  This paper evaluates the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse.
 We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018.
@@
 The MCS developed during the Intense Observing Period (IOP) of the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign.
-We used the GSI-4DLETKF data assimilation package to produce analyses by assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
-We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
+The GSI-4DLETKF data assimilation package is used to produce analyses by assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
+Four assimilation experiments are conducted using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
+The assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
-We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
+Ensemble forecasts initialized from each experiment were also similated to evaluated their skill to predict precipitation.
+The hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
 journal: Atmospheric Research
@@
 # layout: final,5p,times,twocolumn,authoryear
-classoption: preprint, 3p, authoryear
+classoption: preprint, 5p, authoryear, twocolumn
 address:
@@
 For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used [@gustafsson2018].
-Assimilating information on temperature, moisture and wind in the planetary boundary layer improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts upon the planetary boundary layer structure and the location and timing of precipitating systems (e.g.
+Assimilating information on temperature, moisture and wind in the planetary boundary layer (PBL) improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts upon the PBL structure and the location and timing of precipitating systems (e.g.
 @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016).
 
-Particularly relevant for regional mesoscale DA systems in the region of interest is that South America is characterized by a limited number of conventional observations (i.e., radiosondes, surface weather stations) and operational networks are not dense enough to capture mesoscale details.
+Particularly relevant for regional mesoscale DA systems in the region of interest is that South America is characterized by a limited number of conventional observations (i.e., radiosondes, surface weather stations) and operational networks that are not dense enough to capture mesoscale details.
 In this context, analyzing the potential impact of non-conventional sources of observations is essential to improve mesoscale numerical weather prediction (NWP) over South America using DA.
@@
 @sawada2019 reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity).
-As stated in @zhao2021 and @zhao2021a, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
+As stated in \citet{zhao2021, zhao2021a}, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
 \citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
@@
 This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps.
-In particular, we will focus on the impact in the context of a mid-latitude mesoscale convective system.
-
-To reach this goal, we conduct several DA experiments for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
+In particular, this paper focus on the impact in the context of a mid-latitude mesoscale convective system.
+To reach this goal, several DA experiments are conducted for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
 
@@
 
-(ref:case) ERA5 Reanalysis of pressure (hPa) at sea level (black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) for 00 and 12 UTC Nov 22 and 00 UTC Nov 23.
+(ref:case) ERA5 Reanalysis of pressure (hPa) at sea level (black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) for a) 00 and b) 12 UTC Nov 22 and c) 00 UTC Nov 23.
 
@@
 The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1).
-“In this case a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km aproximately to 10 km).
+In this case a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km approximately to 10 km).
 This approach is also based in recent studies which suggest that using multiple nested domains do not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g.
@@
 
-We use a 60-member ensemble where the initial ensemble mean and the mean boundary conditions are taken from the GFS deterministic analysis.
+A 60-member ensemble is used where the initial ensemble mean and the mean boundary conditions are taken from the GFS deterministic analysis.
 A set of 60 perturbations are randomly generated to perturb the initial state as well as the boundary conditions during the length of the experiment.
@@
 The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021.
-In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
-The random perturbations are the same across different experiments to ensure that the differences between experiments are only related to changes in the number and type of assimilated observations.
-
-We used multi-physics scheme to better represent the uncertainty in the model formulation within the DA system.
-We generated 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
+In this way, the nearly hydrostatic and geostrophic equilibrium of larger scales is preserved.
+The random perturbations used are the same across experiments to ensure that the differences between experiments are only related to changes in the number and type of assimilated observations.
+
+A multi-physics scheme is used to better represent the uncertainty in the model formulation within the DA system.
+9 different model configurations are generated consisting of the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]).
 The distribution of these schemes among the 60 ensemble members is outlined in Table \@ref(tab:miembros-desc).
-The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevamnt physical processes that are not resolved by the model.
+The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevant physical processes that are not resolved by the model.
 All ensemble members use the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class scheme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.
@@
   dcast(Cumulus ~ PBL, value.var  = "mem") %>% 
-  kable(booktabs = TRUE, caption = "Generation of the 60-member RRR multi-physics ensemble as a combination of Cumulus and PBL parameterizations, and GEFS members for boundary conditions.", 
+  kable(booktabs = TRUE, caption = "Generation of the 60-member multi-physics ensemble as a combination of Cumulus and PBL parameterizations.", 
         align = "cccc") %>% 
@@
 
-To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in @dillon2021 for all types of observations.
+To reduce the effect of spurious correlations in the estimation of error covariances, a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) is used as in @dillon2021 for all types of observations.
 A relaxation-to-prior spread inflation [@whitaker2012] is applied with an inflation parameter $\alpha=0.9$ following @maldonado2020 to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
@@
 The conventional observations used are part of the Global Data Assimilation System (GDAS) data stream.
-We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP).
+Conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at the National Centers for Environmental Prediction (NCEP) are assimilated.
 These consist of surface observations from 117 Conventional Surface Weather Stations (CSWS), ships, and upper-air observations from 13 radiosondes sites and aircraft.
@@
 
-We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America.
+Data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America are also assimilated.
 The dataset used in this study has been obtained from the RELAMPAGO Data Set repository [@garcia2019].
@@
 
-To evaluate the performance of the ensemble-based DA system presented in this article, we use the following observational datasets:
+To evaluate the performance of the ensemble-based DA system presented in this article, the following observational datasets were used:
 
@@
 The CONV experiment uses only conventional observations from PREPBUFR.
-In a second experiment, referred to as AWS, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from AWS.
-In the third experiment, referred to as SATWND, we assimilate all the observations of the AWS experiment and the satellite-derived winds.
+In a second experiment, referred to as AWS, all the observations included in CONV are assimilated plus the 10-minute frequency surface observations from AWS.
+In the third experiment, referred to as SATWND, the observations from the AWS experiment along with the satellite-derived winds are assimilated.
 Finally, a fourth experiment referred to as RAD assimilates all available clear-sky radiances from sensors onboard polar orbiting satellites as described in section \@ref(sat).
@@
 
-To evaluate the statistical consistency of the uncertainty quantification in the ensemble system we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:
+To evaluate the statistical consistency of the uncertainty quantification in the ensemble system the Reduced Centered Random Variable (RCRV, @candille2007) is used which is defined as:
 
@@
 where $x_o$ is the assimilated observation and its error $\sigma_o$, the ensemble mean of the analysis in observational space $m$, and the standard deviation $\sigma$ of the ensemble.
-The $RCRV$ is the ratio between the distance between the observations and the forecast and its expected standard deviation assuming the statistical independence between the forecast error (estimated from the ensemble spread) and the observation error.
-The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
+The $RCRV$ is the ratio of the distance between the observations and the forecast and its expected standard deviation assuming the statistical independence between the forecast error (estimated from the ensemble spread) and the observation error.
+The average of $RCRV$ computed over all the analysis cycles represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
 
@@
 
-For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute the Fractions Skill Score (FSS, @roberts2008) for different neighborhood length scales and thresholds:
+For the comparison of the first-guess precipitation with the IMERG precipitation estimates, the Fractions Skill Score (FSS, @roberts2008) is computed for different neighborhood length scales and thresholds:
 
@@
 Following @roberts2020, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area.
-The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
+The FSS was computed from the accumulated precipitation over 6 hr rolling windows by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
 
@@
 
-We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019].
-All the analyses in this paper were performed using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
+All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne [@Cheyenne2019].
+All the analyses in this paper were conducted using the R programming language [@rcoreteam2020], using data.table [@dowle2020] and metR [@campitelli2020] packages.
 All graphics are made using ggplot2 [@wickham2009] and the paper was rendered using knitr and rmarkdown [@xie2015; @allaire2019].
@@
 
-To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment.
+To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, the $mean RCRV$ and the $sd RCRV$ is calculated for the RAD experiment.
 As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation.
@@
 For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas in the west of the domain the $sd RCRV$ can be as high as 4.5.
-These higher values of $sdRCRV$ can be associated with systematic errors arising from heigh differences between the model surface and the observations.
+These higher values of $sdRCRV$ can be associated with systematic errors arising from high differences between the model surface and the observations.
 Small scale circulations associated with the complex terrain and not well resolved by the model can also contribute to increase the distance between the forecast and the observations.
@@
 ADPUPA and AIRCAR present a $mean RCRV$ profile near zero at middle and upper levels.
-At low levels, the $mean RCRV$ profile is positive, showing a cold bias present in the model, a characteristic already studied in @ruiz2010 and @dillon2021.
 
 Satellite-derived winds observations varies in number depending on the satellite and the level.
-In Figure \@ref(fig:rcrv-profile)c we included the $RCRV$ calculated at least with 100 observations for each satellite and level.
+In Figure \@ref(fig:rcrv-profile)c only the $RCRV$ calculated with at least 100 observations for each satellite and level is included.
 At low levels, where there are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error.
@@
 
-In this section, we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
+This section presents the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
 The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS.
 Figures \@ref(fig:TQ-diff)a-c show the analysis difference between experiments in the spatially averaged vertical profile of temperature.
-By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
+By averaging the differences between two experiments the systematic impact produced by different observing systems on the analyzed state can be isolated.
 During the first day, the assimilation of AWS observations results in a colder PBL.
@@
 The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@casaretto2022, @matsudo2021, @ruiz2010].
-We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL.
-In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
+The moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL.
+In the experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
 As for the temperature, the systematic impact of satellite-derived winds on moisture is small (Figure \@ref(fig:TQ-diff)e).
@@
 This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment.
-After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain.
+After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain to the northeast.
 After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels (Figures \@ref(fig:UV-diff)d, f).
@@
 Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is largest; the systematic impact is overall smaller than the one produced by assimilating data from AWS (Figures \@ref(fig:UV-diff)b, e).
-The reason of the small impact observed in SATWND could be asociated to the large observation error used for satellite-derived wind observatios.
+The reason of the small impact observed in SATWND could be associated to the large observation error used for satellite-derived wind observations.
 
@@
 ```
-We also compared ensemble mean analyses to ERA5 reanalyses [@era5pressure], which suports Figures
+The ensemble mean analyses are also compared to ERA5 reanalyses [@era5pressure], which supports Figures
 \@ref(fig:TQ-diff) and \@ref(fig:UV-diff).
 Specifically, Figure \@ref(fig:era5)a shows a warm bias in low levels (i.e.
-CONV is warmer than ERA5) that decreases in Figure \@ref(fig:era5)b when we assimilate the AWS observations.
+CONV is warmer than ERA5) that decreases in Figure \@ref(fig:era5)b when the AWS observations are assimilated.
 In the same direction, Figure \@ref(fig:TQ-diff)a shows a negative difference between AWS and CONV meaning that the AWS observations are cooling the low levels.
-Comparing ERA5 with RAD (Figure \@ref(fig:era5)d), there is a see a small increase in the warm bias, associated with the warming produced by the radiance observations (Figure \@ref(fig:TQ-diff)c).
-A similar effect can be observed for specific humidity, AWS observations parcialy corrects the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces this effect.
+Comparing ERA5 with RAD (Figure \@ref(fig:era5)d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations (Figure \@ref(fig:TQ-diff)c).
+A similar effect can be observed for specific humidity, AWS observations partially corrects the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces this effect.
 The impact on the wind components is minor and only the meridional wind is included in Figure \@ref(fig:era5).
@@
 ```
-To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles).
-At that moment the first convective cells were developing over the southern region of the domain along the cold front.
+
+To investigate how changes in the PBL can modify the pre-convective environment, the analysis mean horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE are compared.
+At 00 UTC Nov 22 (after 30 assimilation cycles) the first convective cells were developing over the southern region of the domain along the cold front.
 Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
-We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
+It shows that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
 The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
@@
 On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
-A warmer PBL increases the potencial inestability and helps to generate a suitable enviroment for the development of deep convection.
+A warmer PBL increases the potential instability and helps to generate a suitable environment for the development of deep convection.
 Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear.
@@
 
-First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
+First, the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation is analyzed.
 Figure \@ref(fig:pp-hov)a shows the hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) [@clark2017] for the first-guess hourly accumulated precipitation as averaged between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments.
@@
 
-We computed the FSS to quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments (Figure \@ref(fig:fss)).
-For each threshold and spatial scale, we applied Equation \@red(eq:eq7) in 6-hours rolling windows throughout the experiment period.
+The FSS is computed to quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments (Figure \@ref(fig:fss)).
+For each threshold and spatial scale, Equation \@red(eq:eq7) is applied in 6-hours rolling windows throughout the experiment period.
 All experiments show similar values of FSS during the initiation of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
@@
 This is partially expected considering the coarse horizontal grid spacing (10 km), which is not enough to appropriately represent the strength of the convective band associated with the MCS.
-Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecast reflectivity field.
 RAD better represents the observed features of the system showing a stronger and more organized MCS than CONV, over the domain center at 10 and 13 UTC (first and second columns in Figure \@ref(fig:dbz-mean)).
@@
 AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias.
-However, in this region (Figure \@ref(fig:dominio)b) and for this period, AWS increments (Figure \@ref(fig:UV-diff)d) degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
+However, in this region (Figure \@ref(fig:dominio)b) and for this period, AWS increments (Figure \@ref(fig:UV-diff)d) degrades the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \@ref(fig:soundings)c).
 
@@
 
-In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
-We calculated the FSS for the ensemble forecasts in 6-hour rollig windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
+This section analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
+The FSS is again calculated for the ensemble forecasts in 6-hour rolling windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
 CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
@@
 Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the 1-h forecast to IMERG estimations.
-The reason for why the forecasts inicialized from RAD degrades over time needs to be further study.
+The reason for why the forecasts initialized from RAD degrades over time needs to be further study.
+However, it is possible that the assimilation of observations associated with channels affected by the surface is contributing to the degradation of the PBL in the analysis and subsequently in the forecasts.
+@lin2014 observed limited impact when assimilating AIRS observations and attributes this result to the use of surface channels where the uncertainties associated with emissivity are large.
 
@@
 
-We also analyzed the performance of independent ensemble forecasts intialized from the analyses to forecast the precipitation during Nov 22.
+We also analyzed the performance of independent ensemble forecasts initialized from the analyses to forecast the precipitation during Nov 22.
 The forecast initialized from AWS, SATWND, and RAD are able to forecast the precipitation substantially better than CONV.

### 2022-08-24 analysis/paper/paper_ana_20181122.Rmd
Nuevas conclusiones

@@
 The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25$^{\circ}$ horizontal grid spacing and 6-hour temporal resolution; @cisl_rda_ds084.1).
-In this case a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km approximately to 10 km).
-This approach is also based in recent studies which suggest that using multiple nested domains do not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g.
+In this case, a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25$^{\circ}$ or 25 km approximately to 10 km).
+This approach is also based on recent studies which suggest that using multiple nested domains does not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g.
 @liang2019, @beck2004).
@@
 
-```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.width=7.2, fig.height=5, fig.align="center"}
+```{r era5, fig.cap="(ref:era5-cap)", fig.env="figure*", fig.width=7.2, fig.height=6, fig.align="center", fig.pos="ht"}
 era <- fread(here("analysis/data/derived_data/reanalysis/perfiles_ana_era5.csv"))
@@
 
-```{r dbz-mean, fig.cap="(ref:dbz-mean)", out.width="100%", fig.height=6.5, fig.width=8, fig.env="figure*", fig.pos="h"}
+```{r dbz-mean, fig.cap="(ref:dbz-mean)", out.width="100%", fig.height=6.5, fig.width=8, fig.env="figure*", fig.pos="ht"}
 proj_radar <- "+proj=tmerc +lon_0=-61.5 +lat_0=-32 +k_0=1 +ellps=WGS84"
@@
 However, it is possible that the assimilation of observations associated with channels affected by the surface is contributing to the degradation of the PBL in the analysis and subsequently in the forecasts.
-@lin2014 observed limited impact when assimilating AIRS observations and attributes this result to the use of surface channels where the uncertainties associated with emissivity are large.
+@lim2014 observed limited impact when assimilating AIRS observations and attributes this result to the use of surface channels where the uncertainties associated with emissivity are large.
 
@@
 
-In this paper, we investigate for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system.
-In particular, we explore the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors.
+This paper investigates for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system.
+In particular, the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors is explored.
 Southern South America is a particularly interesting region due to the heterogeneity in topography and weather and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
 This combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.
-To consider these issues, we used a case study approach of a massive mesoscale convective system that developed over Southern South America, particularly in a region where the surface and upper-air observing networks are too coarse to capture variability at the mesoscale and that took place on Nov 22, 2018.
-
-We evaluated the consistency of the ensemble using the RCRV calculated for each type of observation used in RAD.
-Conventional observations including surface data and upper air observations from soundings and aircraft, show a $sd RCRV$ near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
-For satellite-derived winds and radiance observations, the $sd RCRV$ is lower than 1, which could be the result of an underdispersive ensemble or an overestimation of the observation errors.
+To consider these issues, a case study approach of a massive mesoscale convective system that developed over Southern South America was used, particularly in a region where the surface and upper-air observing networks are too coarse to capture variability at the mesoscale and that took place on Nov 22, 2018.
+
+Firstly, the consistency of the ensemble was evaluated using the RCRV to ensure a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
+While Conventional observations including surface data and upper air observations from soundings and aircraft, show a $sd RCRV$ near 1, satellite-derived winds and radiance observations, show $sd RCRV$ is lower than 1.
+This could be the result of an underdispersive ensemble or an overestimation of the observation errors.
 However, taking into account that the observation error for satellite-derived wind is overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the ensemble has a reasonable spread.
@@
 In particular, the precipitable water content and the low level meridional circulation in the AWS experiment lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
-We also found positive results when assimilating radiance observations that produce an adequate development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
-However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates.
-While the assimilation of satellite-derived wind does not produce a noticeable impact on the analysis with respect to the AWS experiment, this is possibly due to the small number of observations in low levels available for this case study and its large observation error.
+AWS observations also help to reduce a warm and dry bias present in the model producing an analysis closer to the ERA5 reanalysis.
+When comparing the analysis with soundings during the pre-convective environment (IOP 7), we observed an improvement at low levels when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
+
+We also found positive results when assimilating radiance observations that produce an adequate development of the convection and its circulation, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
+However, this observations weakened the impact of AWS in the PBL slightly increasing the warm and dry bias.
+While this needs to be further study, it could be related to the assimilation of channels affected by the surface or a sub optimal bias correction.
+The assimilation of radiances also improves the mid and upper level wind during IOP 8 after the MCS crossed the center of the domain.
+
+The assimilation of satellite-derived wind does not produce a noticeable impact on the analysis with respect to the AWS experiment, this is possibly due to the small number of observations in low levels available for this case study and its large observation error.
 However, there are substantial improvements for the accumulated precipitation distribution.
 A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels.
-
-Taking advantage of the observations taken during the RELAMPAGO field campaign, we compared the experiments with the soundings launched during IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement at low levels when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
-In particular, AWS observations correct a dry bias already observed by @ruiz2010.
-For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels when assimilating more observations for the temperature but the opposite effect for the dew point temperature.
-In this case, the analyses have a moist bias that increases with the assimilation of AWS and radiance observations.
-We noted a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
-It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period, which requires further analysis.
 
@@
 In particular, the continuous assimilation of satellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
+The reason why the forecast initialized from RAD does not perform better than SATWND needs to be further study.
 
@@
 A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at <https://github.com/paocorrales/mesoda>.
-The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5072/zenodo.1060834, version 0.9.0.
+The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5281/zenodo.7015913, version 0.9.1.
 

### 2022-08-24 analysis/paper/paper_ana_20181122.Rmd
Conclusión en líneas
Typos

@@
 Four assimilation experiments are conducted using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
-The assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
+The assimilation of observations with high temporal and spatial frequency generates an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-Ensemble forecasts initialized from each experiment were also similated to evaluated their skill to predict precipitation.
+Ensemble forecasts initialized from each experiment were also simulated to evaluate their skill to predict precipitation.
 The hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
@@
 For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used [@gustafsson2018].
-Assimilating information on temperature, moisture and wind in the planetary boundary layer (PBL) improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts upon the PBL structure and the location and timing of precipitating systems (e.g.
+Assimilating information on temperature, moisture, and wind in the planetary boundary layer (PBL) improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts on the PBL structure and the location and timing of precipitating systems (e.g.
 @wheatley2010, @ha2014, @chang2017, @bae2022, @banos2021, @maejima2019, and @chen2016).
@@
 @dillon2016; @dillon2021; @goncalvesdegoncalves2015].
-In particular @dillon2021 assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts.
-Similarly to @gasperoni2018, @dillon2021 included private weather station networks which are not incorporated in the operational analysis.
+In particular, @dillon2021 assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts.
+Similar to @gasperoni2018, @dillon2021 included private weather station networks which are not incorporated in the operational analysis.
 However, the impact of different observation types on the analysis quality has not been addressed.
@@
 This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps.
-In particular, this paper focus on the impact in the context of a mid-latitude mesoscale convective system.
+In particular, this paper focuses on the impact in the context of a mid-latitude mesoscale convective system.
 To reach this goal, several DA experiments are conducted for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
@@
 On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)b) that rapidly grew upscale into an exceptionally large MCS.
-During that day several surface stations reported lightning, gustfronts and heavy rain.
+During that day several surface stations reported lightning, gustfronts, and heavy rain.
 To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS (Figure \@ref(fig:case)c).
@@
 
-### Satellite derived winds
+### Satellite-derived winds
 
@@
 Assuming that the observation error is accurately estimated, an $sd RCRV > 1$ indicates that the ensemble is underdispersive (i.e.
-the distance between the observations and the forecasts is larger than expected) and an $sd RCRV < 1$ indicates that the ensemble is overdispersive (i.e.
+the distance between the observations and the forecasts is larger than expected), and an $sd RCRV < 1$ indicates that the ensemble is overdispersive (i.e.
 the distance between the observations and the forecasts is lower than expected).
@@
 
-Satellite-derived winds observations varies in number depending on the satellite and the level.
+Satellite-derived winds observations vary in number depending on the satellite and the level.
 In Figure \@ref(fig:rcrv-profile)c only the $RCRV$ calculated with at least 100 observations for each satellite and level is included.
@@
 Overall, these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
-Moreover, these results suggest the relaxation-to-prior spread inflation inflation parameter $\alpha = 0.9$ is adequate for the system.
+Moreover, these results suggest the relaxation-to-prior spread inflation parameter $\alpha = 0.9$ is adequate for the system.
 
@@
 Comparing the specific humidity in the experiments (Figures \@ref(fig:TQ-diff)d-f), the impact of assimilating AWS with fine spatial and temporal resolution is most substantial at low levels (Figure \@ref(fig:TQ-diff)d).
-The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime.
+The PBL in the AWS experiment is consistently moister than in the CONV experiment, particularly at nighttime.
 The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@casaretto2022, @matsudo2021, @ruiz2010].
@@
 ```
+
 The ensemble mean analyses are also compared to ERA5 reanalyses [@era5pressure], which supports Figures
@@
 Comparing ERA5 with RAD (Figure \@ref(fig:era5)d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations (Figure \@ref(fig:TQ-diff)c).
-A similar effect can be observed for specific humidity, AWS observations partially corrects the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces this effect.
+A similar effect can be observed for specific humidity, AWS observations partially correct the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces this effect.
 The impact on the wind components is minor and only the meridional wind is included in Figure \@ref(fig:era5).
 In this case Figures \@ref(fig:era5)i-l show that the radiance observations are mainly responsible for the positive impact observed in the analysis.
-Moreover, these adjustments due to assimilating radiance and AWS observations lead to ensemble mean analyses closer to ERA5 reanalyses.
+Moreover, these adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5 reanalyses.
 
@@
 
-This section analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
+This section analyzes the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
 The FSS is again calculated for the ensemble forecasts in 6-hour rolling windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \@ref(fig:fssfcst)).
@@
 Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the 1-h forecast to IMERG estimations.
-The reason for why the forecasts initialized from RAD degrades over time needs to be further study.
+The reason why the forecasts initialized from RAD degrade over time needs to be further study.
 However, it is possible that the assimilation of observations associated with channels affected by the surface is contributing to the degradation of the PBL in the analysis and subsequently in the forecasts.
-@lim2014 observed limited impact when assimilating AIRS observations and attributes this result to the use of surface channels where the uncertainties associated with emissivity are large.
+@lim2014 observed limited impact when assimilating AIRS observations and attribute this result to the use of surface channels where the uncertainties associated with emissivity are large.
 
@@
 
-This paper investigates for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system.
-In particular, the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors is explored.
-Southern South America is a particularly interesting region due to the heterogeneity in topography and weather and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
-This combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.
-To consider these issues, a case study approach of a massive mesoscale convective system that developed over Southern South America was used, particularly in a region where the surface and upper-air observing networks are too coarse to capture variability at the mesoscale and that took place on Nov 22, 2018.
+Southern South America is a particularly interesting region due to the heterogeneity in topography and coarse resolution of the operational observing network (considering both surface based and upper air observations).
+This, combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.
+This paper investigates, for the first time in South America, using a case-study approach, the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system on the mesoscale.
+This case study corresponds to a massive MCS that developed over Southern South America on Nov 22, 2018 during the RELAMPAGO field campaign.
+In particular, the impact on the analysis quality of assimilating frequent and relatively dense surface observations, satellite-derived winds, and satellite clear-sky radiances from multiple sensors is explored.
 
 Firstly, the consistency of the ensemble was evaluated using the RCRV to ensure a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
-While Conventional observations including surface data and upper air observations from soundings and aircraft, show a $sd RCRV$ near 1, satellite-derived winds and radiance observations, show $sd RCRV$ is lower than 1.
-This could be the result of an underdispersive ensemble or an overestimation of the observation errors.
-However, taking into account that the observation error for satellite-derived wind is overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the ensemble has a reasonable spread.
-
-In terms of the analysis, we found that AWS observations, which have high spatial and temporal resolution, produce impacts throughout the troposphere, especially in the PBL.
-In particular, the precipitable water content and the low level meridional circulation in the AWS experiment lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
-AWS observations also help to reduce a warm and dry bias present in the model producing an analysis closer to the ERA5 reanalysis.
-When comparing the analysis with soundings during the pre-convective environment (IOP 7), we observed an improvement at low levels when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
-
-We also found positive results when assimilating radiance observations that produce an adequate development of the convection and its circulation, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
-However, this observations weakened the impact of AWS in the PBL slightly increasing the warm and dry bias.
-While this needs to be further study, it could be related to the assimilation of channels affected by the surface or a sub optimal bias correction.
-The assimilation of radiances also improves the mid and upper level wind during IOP 8 after the MCS crossed the center of the domain.
-
-The assimilation of satellite-derived wind does not produce a noticeable impact on the analysis with respect to the AWS experiment, this is possibly due to the small number of observations in low levels available for this case study and its large observation error.
-However, there are substantial improvements for the accumulated precipitation distribution.
+While conventional observations, including surface data and upper air observations from soundings and aircraft, show a near 1, satellite-derived winds and radiance observations show a lower than 1.
+The latter could be the result of either an underdispersive ensemble or an overestimation of the observation errors.
+However, considering that the observation error of satellite-derived wind is overestimated in the system configuration (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the ensemble has a reasonable spread.
+
+In terms of the analysis, AWS observations, which have high spatial and temporal resolution, produced impacts throughout the troposphere, especially in the PBL.
+In particular, precipitable water content and low level meridional circulation in the AWS experiment led to the development of deep convection and heavy precipitation closer to observations.
+AWS observations also helped to reduce the warm and dry bias present in the model, producing an analysis closer to the ERA5 reanalysis.
+During the pre-convective environment (IOP 7), assimilating temperature, dew point temperature, and meridional wind from AWS improved the analysis at low levels when compared with observed soundings.
+
+We also found positive results when assimilating radiance observations, which produced a better development of the convection and its circulation, mainly during the mature stage of the MCS, leading to increased accumulated precipitation compared with the other experiments.
+However, these observations weakened the impact of AWS in the PBL, slightly increasing the warm and dry bias.
+While this needs to be further studied, it could be related to the assimilation of channels affected by the surface or sub optimal bias correction.
+The assimilation of radiances also improved mid and upper level wind during IOP 8 after the MCS crossed the center of the domain.
+The assimilation of satellite-derived wind did not produce a noticeable impact on the analysis with respect to the AWS experiment.
+This is possibly due to the small number of observations in low levels available for this case study and their large observation error.
+However, there are substantial improvements in the accumulated precipitation distribution.
 A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels.
 
-We also analyzed the performance of independent ensemble forecasts initialized from the analyses to forecast the precipitation during Nov 22.
-The forecast initialized from AWS, SATWND, and RAD are able to forecast the precipitation substantially better than CONV.
-In particular, the continuous assimilation of satellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
-The reason why the forecast initialized from RAD does not perform better than SATWND needs to be further study.
+We also analyzed the performance of independent ensemble precipitation forecasts initialized from the analyses during Nov 22.
+The forecasts initialized from AWS, SATWND, and RAD were able to forecast the precipitation substantially better than CONV.
+In particular, continuous assimilation of satellite-derived wind and radiance observations improved the latest initialization but only satellite-derived wind observations produced a positive impact that persisted throughout the forecast.
+Why the forecast initialized from RAD did not perform better than SATWND needs to be further study.
 
 To summarize, in this study we conclude that the assimilation of surface observations with high spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
-Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events.
-In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like GPS radio occultation data and radiances from geostationary orbiting satellites like GOES-16.
+Moreover, ensemble forecasts initialized from the analysis showed promising results for predicting extreme severe precipitation events.
+In the future, we will further analyze the impact of these observations in independent forecasts and evaluate the assimilation of other sources of observations such as GPS radio occultation data and radiances from geostationary orbiting satellites like GOES-16.
 
@@
 
-# Acknowledgements {.unnumbered}
+# Acknowledgments {.unnumbered}
 

### 2022-08-27 R/goes_functions.R
goes functions

@@ -1,0 +1,100 @@
+#' Goes Proyection
+#'
+#' @param x regular grid
+#' @param y regular grid
+#' @param ncfile NetCDF file
+#'
+#' @export
+
+
+# projection --------------------------------------------------------------
+
+goes_projection <- function(x, y, ncfile) {
+
+  proj_info <- ncdf4::ncatt_get(ncfile, "goes_imager_projection")
+  h <- proj_info$perspective_point_height
+
+  map_proj <- paste0("+proj=geos",
+                     " +h=",  proj_info$perspective_point_height,
+                     " +lon_0=", proj_info$longitude_of_projection_origin,
+                     " +sweep=", proj_info$sweep_angle_axis,
+                     " +ellps=GRS80")
+
+  x_atr <- ncdf4::ncatt_get(ncfile, "x")
+  x <- (x*x_atr$scale_factor + x_atr$add_offset)*h
+  y_atr <- ncdf4::ncatt_get(ncfile, "y")
+  y <- (y*y_atr$scale_factor + y_atr$add_offset)*h
+
+  proj4::project(list(x, y), map_proj, inverse = TRUE)
+}
+
+
+#' Color scale for cloud tops
+#'
+#' @param colours hex format
+#' @param limits vector with limits
+#' @param breaks vector with breaks
+#' @param guide
+#' @param ... others
+#'
+#' @export
+#'
+# scale_color_topes -------------------------------------------------------
+
+scale_color_topes <- function(colours = hex,
+                              limits = c(-90, 50),
+                              breaks = seq(-90, 50, 20),
+                              guide = guide_colorbar(barheight = 15),
+                              ...) {
+  # https://github.com/garciafido/cima-goes/tree/master/src/cima/goes/img
+  topes <- data.table::fread(here::here("R/paleta_topes"))
+  colours <- rgb(topes$V2, topes$V3, topes$V4, maxColorValue = 255)
+
+  scale_color_gradientn(colours = colours,
+                        limits = limits,
+                        breaks = breaks,
+                        guide = guide, ...)
+}
+
+#' Radiance to BT
+#'
+#' @param rad radiance
+#' @param ncfile NetCDF file
+#' @param h constant
+#' @param c constant
[…]

### 2022-08-27 analysis/paper/paper_ana_20181122.Rmd
Revisión

@@
     footnote: 1
-  - name: Victoria Galligani
+  - name: V. Galligani
     affiliation: UBA,CIMA,CNRS
@@
 output: 
-  # bookdown::word_document2: default
   bookdown::pdf_book:
@@
   This paper evaluates the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse.
-We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018.
+A case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018 was used.
 The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
@@
 The GSI-4DLETKF data assimilation package is used to produce analyses by assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
-Four assimilation experiments are conducted using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
+Four assimilation experiments are conducted using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks (AWS), SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
 The assimilation of observations with high temporal and spatial frequency generates an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
@@
 # layout: final,5p,times,twocolumn,authoryear
-classoption: preprint, 5p, authoryear, twocolumn
+classoption: preprint, 3p, authoryear,review, 12pt
+linenumbers: true
+numbersections: true
 address:
   - code: UBA
-    address: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
+    address: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmosfera y los Oceanos. Buenos Aires, Argentina.
   - code: CIMA
-    address: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
+    address: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmosfera (CIMA). Buenos Aires, Argentina.
   - code: CNRS
@@
   - code: SMN
-    address: Servicio Meteorológico Nacional de Argentina.
+    address: Servicio Meteorologico Nacional de Argentina.
   - code: CONICET
-    address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).
+    address: CONICET (Consejo Nacional de Investigaciones Cientificas y Tecnicas).
   - code: INPE
@@
 keywords: Regional Data Assimilation, Surface Observations, Satellite Observations 
-numbersections: true
-# header-includes: 
+header-includes:
 #  - \usepackage{gensymb}
@@
 #  - \usepackage{lineno}
-#  - \linenumbers 
+  - \usepackage[utf8]{inputenc}
 ---
@@
 While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting satellite observations), bias correction, and the relatively low model tops used for this application.
-@bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
-More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
-@wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis.
+@bao2015 studied the impact of assimilating cloud-cleared microwave and infrared radiance data polar orbiting instruments on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+More recently, @zhu2019 studied the impact of assimilating clear sky polar orbiting satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
+@wang2021 studied the impact of assimilating clear sky radiances in the high-resolution Copernicus European Regional Reanalysis.
 They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere.
@@
 
-Previously to the development of this case study, the center and north of Argentina was immersed in a warm and humid air mass with high values of convective available potential energy (CAPE) (Figure \@ref(fig:case)a).
-On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells (Figure \@ref(fig:case)b) that rapidly grew upscale into an exceptionally large MCS.
-During that day several surface stations reported lightning, gustfronts, and heavy rain.
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS (Figure \@ref(fig:case)c).
+Previously to the development of this case study, the center and north of Argentina was immersed in a warm and humid air mass with high values of convective available potential energy (CAPE), as shown by  ERA 5 Reanalysis [@era5pressure] in Figure \@ref(fig:case)a.
+On Nov 22, 2018 a cold front crossed the center of Argentina (Figure \@ref(fig:case)b).
+This cold front triggered isolated convective cells that rapidly grew upscale into an exceptionally large MCS (Figure \@ref(fig:case)d,e).
+During that day several surface stations reported lightning, strong wind gusts, and heavy rain.
+To the north of the region, a warm and humid environment contributed to the development of isolated convection that ultimately grew and merged with the MCS (Figure \@ref(fig:case)f).
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
 
-(ref:case) ERA5 Reanalysis of pressure (hPa) at sea level (black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) for a) 00 and b) 12 UTC Nov 22 and c) 00 UTC Nov 23.
-
-```{r case, fig.cap="(ref:case)", fig.width=7, fig.height=3.5, fig.align="center", fig.env = "figure*"}
-
+(ref:case) ERA5 Reanalysis of sea level pressure (hPa, black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) and GOES-16 channel 13 brightness temperature for a,d) 00 and b,e) 12 UTC Nov 22 and c,f) 00 UTC Nov 23.
+
+```{r case, fig.cap="(ref:case)", fig.width=6, fig.height=4, fig.align="center", fig.env = "figure*"}
+#fig.width=7, fig.height=3.5,
 campos <- ReadNetCDF(here("analysis/data/derived_data/reanalysis/geopotential.nc")) %>% 
@@
   .[, ":="(msl = ReadNetCDF(here("analysis/data/derived_data/reanalysis/pressure-pw.nc"), vars = "msl", out = "vector")[[1]],
-           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]])]
+           cape = ReadNetCDF(here("analysis/data/derived_data/reanalysis/cape-pw.nc"), vars = "cape", out = "vector")[[1]])] %>% 
+  setnames(c("longitude", "latitude"), c("lon", "lat")) %>% 
+    .[time != ymd_h(2018112312)]
+
+xlim <- c(2700, 3950)
+ylim <- c(3800, 4950)
+
+files <- Sys.glob(here("analysis/data/derived_data/goes16/*"))
+goes <- map(files, function(f) {
+  nc <- ncdf4::nc_open(f)
+  meta <- unglue::unglue(basename(f), "OR_ABI-L1b-RadF-M3{channel}_G16_s{sdate}_e{edate}_c{cdate}.nc")
+  ReadNetCDF(f, vars = c(rad = "Rad", time = "t"), 
+             subset = list(x = xlim, y = ylim)) %>% 
+  .[, rad := calculate_rad(rad, nc)] %>% 
+  .[, tb := rad_to_tb(rad, nc)] %>% 
+  .[, time := as_datetime(time, origin = "2000-01-01 12:00:00")] %>%
+  .[, c("lon", "lat") := goes_projection(x, y, nc)] %>% 
+  .[]
+}) %>% 
+  rbindlist() %>% 
+  .[, time := floor_date(time, "hour")]
 
 campos %>% 
-
-  .[time != ymd_h(2018112312)] %>% 
+  .[, source := "era5"] %>%
+  .[, time := factor(time)] %>% 
   .[, ":="(espesor = (`500`-`1000`)/100, time = factor(time))] %>%
-  ggplot(aes(longitude, latitude)) +
+  ggplot(aes(lon, lat)) +
   geom_contour_fill(aes(z = cape, fill = stat(level_d)), breaks = seq(500, 4000, 500)) +
@@
                        palette = "YlOrRd", direction = 1,
-                       guide = guide_colorsteps(barwidth = 25, mid = 25,
-                                                barheight = 0.5,
-                                                title.position = "left",
-                                                title.vjust = 1,
+                       guide = guide_colorsteps(barwidth = 0.5, #mid = 25,
+                                                barheight = 8,
+                                                order = 2, 
+                                                # title.position = "left",
+                                                # title.vjust = 1,
                                                 frame.colour = "black")) +
-  geom_mapa() +
   geom_contour2(aes(z = espesor, label = ..level..),
@@
   cordillera +
+  scattermore::geom_scattermore(data = goes[, source := "goes"], aes(color = tb)) +
+  scale_color_topes(guide = guide_colorbar(barwidth = 0.5,
+                                           barheight = 8,
+                                           order = 99,
+                                           frame.colour = "black")) +
+  # guides(fill = guide_legend(" CAPE", order = 0),
+  #        color = guide_legend("BT (ºC)", order = 2)) +
+  geom_mapa() +
   coord_sf(ylim = c(-45, -22), xlim = c(-75, -52)) +
-  facet_wrap(vars(time), labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
+  facet_grid(source ~ time, labeller = labeller(time = c("2018-11-22 00:00:00" = "00 UTC Nov 22",
                                                          "2018-11-22 12:00:00" = "12 UTC Nov 22",
-                                                      "2018-11-23 00:00:00" = "00 UTC Nov 23"))) +
+                                                         "2018-11-23 00:00:00" = "00 UTC Nov 23"),
+                                                source = c("era5" = "ERA5",
+                                                           "goes" = "GOES-16 Channel 13"))) +
   tag_facets() +
-  labs(x = NULL, y = NULL, fill = "CAPE" ) +
-  theme_minimal(base_size = 8) +
-  theme(legend.position = "bottom")
+  labs(x = NULL, y = NULL, fill = "CAPE", color = "BT (ºC)") +
+  theme_minimal(base_size = 9) +
+  theme(tagger.panel.tag.background = element_rect(color = "white"))
 
@@
 
-```{r dominio, fig.cap="(ref:dominio)", out.width="100%"}
+```{r dominio, fig.cap="(ref:dominio)", out.width="98%", fig.width=5}
 
@@
   theme(legend.box = "horizontal",
-        legend.position = c(0.1, 0.95),
+        legend.position = c(0.15, 0.95),
         legend.background = element_rect(fill = "white", color = "white"))
@@
                       breaks = seq(0, 6000, 1000),
-                      guide = guide_colorstrip(barwidth = 20,
+                      guide = guide_colorstrip(barwidth = 18,
                                                barheight = 0.5)) +
@@
   theme(legend.box = "horizontal",
-        legend.position = c(0.16, 0.92),
+        legend.position = c(0, 1),
         legend.background = element_rect(fill = "white", color = "white"))
@@
 Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations).
-Observation errors used to assimilate these observations are indicated in Table \@ref(tab:table-obs).
+Observation errors used to assimilate these observations follow the GSI default configuration and are indicated in Table \@ref(tab:table-obs).
 
@@
 \begin{equation}
-  \mathrm{RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}}
+  \mathrm{RCRV = \frac{m - x_o}{\sqrt{\sigma_o^2 + \sigma^2}}}
   (\#eq:eq2)
@@
 ADPUPA and AIRCAR present a $mean RCRV$ profile near zero at middle and upper levels.
+At low levels, the meanRCRV profile is positive, showing a cold bias present in the model, a characteristic already studied in @ruiz2010  and @dillon2021.
 
@@
 
-```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=5.5, fig.align="center",fig.pos="th"}
+```{r TQ-diff, fig.cap="(ref:TQ-diff)", fig.env="figure*", fig.height=5.5, fig.width=6, fig.align="center",fig.pos="ht"}
 files <- Sys.glob(here("analysis/data/derived_data/analysis_variables/perfiles_ana_E[2569].csv"))
@@
 
-```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=5.5, fig.align="center", fig.pos="ht"}
+```{r UV-diff, fig.cap="(ref:UV-diff)", fig.env="figure*", fig.height=5.5, fig.width=6, fig.align="center", fig.pos="ht"}
 perfiles %>% 
@@
 
-The ensemble mean analyses are also compared to ERA5 reanalyses [@era5pressure], which supports Figures
-\@ref(fig:TQ-diff) and \@ref(fig:UV-diff).
+The difference between ERA5 reanalyses [@era5pressure`] and the ensemble mean analyses are also compared in Figure \@ref(fig:era5), which supports Figures \@ref(fig:TQ-diff) and \@ref(fig:UV-diff).
 Specifically, Figure \@ref(fig:era5)a shows a warm bias in low levels (i.e.
@@
 In the same direction, Figure \@ref(fig:TQ-diff)a shows a negative difference between AWS and CONV meaning that the AWS observations are cooling the low levels.
-Comparing ERA5 with RAD (Figure \@ref(fig:era5)d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations (Figure \@ref(fig:TQ-diff)c).
-A similar effect can be observed for specific humidity, AWS observations partially correct the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces this effect.
-The impact on the wind components is minor and only the meridional wind is included in Figure \@ref(fig:era5).
-In this case Figures \@ref(fig:era5)i-l show that the radiance observations are mainly responsible for the positive impact observed in the analysis.
-Moreover, these adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5 reanalyses.
-
-(ref:era5-cap) Difference between analysis ensemble mean experiments and ERA5 for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity ($g\ Kg^{-1}$, e--h) and meridional wind ($m\ s^{-1}$, i--l) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+Comparing ERA5-RAD (Figure \@ref(fig:era5)d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations as shown in Figure \@ref(fig:TQ-diff)c.
+A similar effect can be observed for specific humidity, AWS observations partially correct the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces the positive impact of AWS.
+The impact on the wind components is minor so only the meridional wind is included in Figures \@ref(fig:era5)i-l, which show that the radiance observations are mainly responsible for the positive impact observed in the analysis by reducing the distance ERA5-RAD, particularly during the mature stage of the MCS.
+Overall, the adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5 reanalyses.
+
+(ref:era5-cap) Difference between ERA5 and the analysis ensemble mean experiments for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity ($g\ Kg^{-1}$, e--h) and meridional wind ($m\ s^{-1}$, i--l) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
 
@@
 
-```{r pp-hov, fig.cap="(ref:pphov)", fig.height=4, fig.env="figure*", fig.pos="h"}
+```{r pp-hov, fig.cap="(ref:pphov)", fig.height=4, fig.width=6, fig.env="figure*", fig.pos="h", fig.align="center"}
 
@@
 
-```{r soundings, fig.cap="(ref:soundings)", fig.env="figure*", fig.pos="t", fig.align="center"}
+```{r soundings, fig.cap="(ref:soundings)", fig.width=6, fig.height=4, fig.env="figure*", fig.pos="ht", fig.align="center"}
 files <- Sys.glob(here("analysis/data/derived_data/soundings/sondeo_E[2569]_ana*"))
@@
 This, combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.
-This paper investigates, for the first time in South America, using a case-study approach, the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system on the mesoscale.
+This paper investigates, for the first time in South America, using a case-study approach, the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system.
 This case study corresponds to a massive MCS that developed over Southern South America on Nov 22, 2018 during the RELAMPAGO field campaign.
@@
 
-Firstly, the consistency of the ensemble was evaluated using the RCRV to ensure a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
-While conventional observations, including surface data and upper air observations from soundings and aircraft, show a near 1, satellite-derived winds and radiance observations show a lower than 1.
-The latter could be the result of either an underdispersive ensemble or an overestimation of the observation errors.
-However, considering that the observation error of satellite-derived wind is overestimated in the system configuration (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the ensemble has a reasonable spread.
-
-In terms of the analysis, AWS observations, which have high spatial and temporal resolution, produced impacts throughout the troposphere, especially in the PBL.
-In particular, precipitable water content and low level meridional circulation in the AWS experiment led to the development of deep convection and heavy precipitation closer to observations.
-AWS observations also helped to reduce the warm and dry bias present in the model, producing an analysis closer to the ERA5 reanalysis.
-During the pre-convective environment (IOP 7), assimilating temperature, dew point temperature, and meridional wind from AWS improved the analysis at low levels when compared with observed soundings.
-
-We also found positive results when assimilating radiance observations, which produced a better development of the convection and its circulation, mainly during the mature stage of the MCS, leading to increased accumulated precipitation compared with the other experiments.
-However, these observations weakened the impact of AWS in the PBL, slightly increasing the warm and dry bias.
-While this needs to be further studied, it could be related to the assimilation of channels affected by the surface or sub optimal bias correction.
-The assimilation of radiances also improved mid and upper level wind during IOP 8 after the MCS crossed the center of the domain.
-The assimilation of satellite-derived wind did not produce a noticeable impact on the analysis with respect to the AWS experiment.
-This is possibly due to the small number of observations in low levels available for this case study and their large observation error.
-However, there are substantial improvements in the accumulated precipitation distribution.
-A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels.
-
-We also analyzed the performance of independent ensemble precipitation forecasts initialized from the analyses during Nov 22.
-The forecasts initialized from AWS, SATWND, and RAD were able to forecast the precipitation substantially better than CONV.
+Firstly, the consistency of the ensemble was evaluated to ensure a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
+While conventional observations departures are consistent with the ensemble spread and assumed observation errors, satellite-derived winds and radiance observations departures are lower than expected.
+The latter could be the result of an overestimation of the observation errors which is usually introduced to avoid the detrimental impact on the analysis of poor quality observations.
+In this case study, all the observation types considered (i.e.
+automatic weather stations, satellite derived winds and clear-sky radiances from polar orbiting satellites) improves the quality of the analysis and of the short range forecast with respect to the conventional observation network.
+In terms of the analysis, automatic weather station observations, which have high spatial and temporal resolution, produced impacts mainly within the PBL but which occasionally extends throughout the troposphere during the periods where moist convection is stronger within the domain.
+These observations also helped to reduce the warm and dry bias present in the model, producing an analysis closer to the ERA5 reanalysis.
+During the pre-convective environment, assimilating surface temperature, dew point temperature, and meridional wind improved the analysis at low levels when compared with observed soundings.
+In particular, when these observations are assimilated, precipitable water content and low level meridional circulation led to the enhancement of deep convection and heavy precipitation which is closer to observations.
+
+Positive results were also found when assimilating radiance observations, which produced a better development of the convection and its associated outflow circulation, mainly during the mature stage of the MCS, leading to increased accumulated precipitation compared to the case in which these observations are not assimilated.
+However, these observations weakened the impact of automatic weather station observations within the PBL, slightly increasing the warm and dry bias with respect to ERA5.
+While this needs to be further studied, it could be related to the assimilation of channels affected by the surface or sub-optimal bias correction.
+Comparing the experiment with independet soundings, the assimilation of radiances improved mid and upper level wind.
+
+The assimilation of satellite-derived wind did not produce a noticeable impact on the analysis.
+This is possibly due to the relatively small number of observations in low levels available for this case study and their large observation error.
+However, there are improvements in the 1-h forecast accumulated precipitation distribution.
+A more comprehensive analysis is necessary to understand the mechanisms behind the impact of these observations on longer range forecasts.
+
+The evaluation of the performance of independent ensemble precipitation forecasts initialized from the analyses during Nov 22 showed that the forecasts initialized from AWS, SATWND, and RAD were able to forecast the precipitation substantially better than CONV.
 In particular, continuous assimilation of satellite-derived wind and radiance observations improved the latest initialization but only satellite-derived wind observations produced a positive impact that persisted throughout the forecast.
@@
 
-To summarize, in this study we conclude that the assimilation of surface observations with high spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
+To summarize, in this case study we found that the assimilation of surface observations with high spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
 Moreover, ensemble forecasts initialized from the analysis showed promising results for predicting extreme severe precipitation events.
-In the future, we will further analyze the impact of these observations in independent forecasts and evaluate the assimilation of other sources of observations such as GPS radio occultation data and radiances from geostationary orbiting satellites like GOES-16.
+In the future, we will further analyze the impact of these observations upon short-range forecasts over longer periods and evaluate the assimilation of other sources of observations such as GPS radio occultation data and radiances from geostationary orbiting satellites like GOES-16.
 
@@
 A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at <https://github.com/paocorrales/mesoda>.
-The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5281/zenodo.7015913, version 0.9.1.
+The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5281/zenodo.7015913, version 0.9.2.
 

### 2022-08-27 analysis/paper/paper_ana_20181122.tex

@@
-\documentclass[authoryear,preprint,review,12pt]{elsarticle} %review=doublespace preprint=single 5p=2 column
+\documentclass[preprint, 3p, authoryear,review, 12pt]{elsarticle} %review=doublespace preprint=single 5p=2 column
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+
 \usepackage[hyphens]{url}
@@
 \usepackage{lineno} % add
+  \linenumbers % turns line numbering on
 
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\bibliographystyle{elsarticle-harv}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+
 \usepackage{graphicx}
@@
             pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
 
@@
 
-\usepackage{lineno}
-\linenumbers
+\usepackage[utf8]{inputenc}
 \usepackage{booktabs}
@@
   \title{Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO campaign}
-    \author[UBA,CIMA,CNRS]{Paola Belén Corrales\corref{1}}
+    \author[UBA,CIMA,CNRS]{Paola Belén Corrales%
+  %
+  \fnref{1}}
    \ead{paola.corrales@cima.fcen.uba.ar} 
-    \author[UBA,CIMA,CNRS]{Victoria Galligani}
+    \author[UBA,CIMA,CNRS]{V. Galligani}
 
@@
 
-      \address[UBA]{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.}
-    \address[CIMA]{CONICET -- Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.}
-    \address[CNRS]{CNRS -- IRD -- CONICET -- UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.}
-    \address[SMN]{Servicio Meteorológico Nacional de Argentina.}
-    \address[CONICET]{CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).}
-    \address[INPE]{National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.}
-    \address[NCAR]{National Center for Atmospheric Research, Boulder, Colorado.}
-    \address[Illinois]{Department of Atmospheric Sciences, University of Illinois Urbana--Champaign, Urbana, Illinois.}
-      \cortext[1]{Corresponding Author}
+      \affiliation[UBA]{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmosfera y los Oceanos. Buenos Aires, Argentina.}
+    \affiliation[CIMA]{CONICET -- Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmosfera (CIMA). Buenos Aires, Argentina.}
+    \affiliation[CNRS]{CNRS -- IRD -- CONICET -- UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.}
+    \affiliation[SMN]{Servicio Meteorologico Nacional de Argentina.}
+    \affiliation[CONICET]{CONICET (Consejo Nacional de Investigaciones Cientificas y Tecnicas).}
+    \affiliation[INPE]{National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.}
+    \affiliation[NCAR]{National Center for Atmospheric Research, Boulder, Colorado.}
+    \affiliation[Illinois]{Department of Atmospheric Sciences, University of Illinois Urbana--Champaign, Urbana, Illinois.}
+    \cortext[cor1]{Corresponding author}
+    \fntext[1]{Corresponding Author}
 
   \begin{abstract}
-  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse.
-We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018.
+  This paper evaluates the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse.
+A case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018 was used.
 The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay.
-The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign.
-We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
-We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
-We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
+The MCS developed during the Intense Observing Period (IOP) of the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign.
+The GSI-4DLETKF data assimilation package is used to produce analyses by assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble.
+Four assimilation experiments are conducted using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files; AWS, combining CONV and dense automatic surface weather station networks (AWS), SATWND, combining AWS with satellite-derived winds, and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors.
+The assimilation of observations with high temporal and spatial frequency generates an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study.
 The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
-We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation.
-We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
+Ensemble forecasts initialized from each experiment were also simulated to evaluate their skill to predict precipitation.
+The hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
   \end{abstract}
-   \begin{keyword} Regional Data Assimilation, Surface Observations, Satellite Observations\end{keyword}
+    \begin{keyword}
+    Regional Data Assimilation, Surface Observations, Satellite Observations \sep 
+    Regional Data Assimilation, Surface Observations, Satellite Observations
+  \end{keyword}
+
  \end{frontmatter}
@@
 A large number of these phenomena are associated with the occurrence of deep moist convection, including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning.
-Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events (Brooks et al., 2003) and large hail events (Cecil and Blankenship, 2012), particularly during austral spring and summer.
-This is also confirmed by observational evidence and high impact weather reports (Matsudo et al., 2015; Rasmussen et al., 2014).
-Recently, the RELAMPAGO-CACTI field campaign (Nesbitt et al., 2021) has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
+Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events \citep{brooks2003} and large hail events \citep{cecil2012}, particularly during austral spring and summer.
+This is also confirmed by observational evidence and high impact weather reports \citep{matsudo2015, rasmussen2014}.
+Recently, the Remote sensing of Electrification, Lightning, And Mesoscale/microscale Processes with Adaptive Ground Observations (RELAMPAGO) field campaign \citep{nesbitt2021} has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.
 
 Forecasting mesoscale meteorological phenomena and particularly deep moist convection is a scientific and technological challenge due to its limited predictability and the difficulties in diagnosing the state of the atmosphere at small spatial and short temporal scales (for example from 1 to 10 kilometers and on the order of minutes).
-Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for high-resolution numerical forecasts (Sun et al., 2014) and thus has received increasing attention in the last decades.
-
-For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used.
-
-In that regard, several authors investigated the impact of assimilating surface weather data (e.g.~Wheatley and Stensrud (2010), Ha and Snyder (2014), Chang et al.
-(2017), Bae and Min (2022), Banos et al.
-(2021), Maejima et al.
-(2019), and Chen et al.
-(2016)) using different assimilation methodologies.
-Most of these studies reported the beneficial impacts of assimilating temperature and dew point observations upon the planetary boundary layer structure and the location and timing of precipitating systems.
-Sobash and Stensrud (2015) showed using a mesoscale DA system, that the positive impact upon convection initiation and the short range precipitation forecast is achieved if data is assimilated frequently (in the order of minutes, rather than in the order of hours).
-More recently, Gasperoni et al.
-(2018) evaluated the impact of assimilating observations produced by private weather station networks which are not incorporated in the operational analysis.
-They found a positive effect of these observations upon the initiation of deep moist convection.
-This result is particularly important for data sparse regions such as Southern South America, where operational networks are not dense enough to capture mesoscale details.
-
-The impact of other types of high spatial and temporal resolution observations, such as atmospheric motion vectors (AMVs), has also been investigated in the context of limited-area mesoscale DA.
-Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., Wu et al.
-(2014), Cherubini et al.
-(2006), and Sawada et al.
-(2019), among many others).
+Mesoscale data assimilation (DA) is an approach that can provide appropriate initial conditions for high-resolution numerical forecasts \citep{sun2014} and thus has received increasing attention in the last decades.
+
+For DA methods to be successful, observing networks with sufficient temporal and spatial resolution capable of capturing mesoscale variability should be used \citep{gustafsson2018}.
+Assimilating information on temperature, moisture, and wind in the planetary boundary layer (PBL) improves mesoscale model initialization, and several authors have reported the resultant beneficial impacts on the PBL structure and the location and timing of precipitating systems (e.g.
+\citet{wheatley2010}, \citet{ha2014}, \citet{chang2017}, \citet{bae2022}, \citet{banos2021}, \citet{maejima2019}, and \citet{chen2016}).
+
+Particularly relevant for regional mesoscale DA systems in the region of interest is that South America is characterized by a limited number of conventional observations (i.e., radiosondes, surface weather stations) and operational networks that are not dense enough to capture mesoscale details.
+In this context, analyzing the potential impact of non-conventional sources of observations is essential to improve mesoscale numerical weather prediction (NWP) over South America using DA.
+There have been only a few published efforts on regional mesoscale DA, but they have all shown promising results \citep[e.g.][]{dillon2016, dillon2021, goncalvesdegoncalves2015}.
+In particular, \citet{dillon2021} assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts.
+Similar to \citet{gasperoni2018}, \citet{dillon2021} included private weather station networks which are not incorporated in the operational analysis.
+However, the impact of different observation types on the analysis quality has not been addressed.
+
+The impact of non-conventional high spatial and temporal resolution observations, such as AMVs, has been investigated in the context of regional mesoscale DA.
+Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., \citet{wu2014}, \citet{cherubini2006}, and \citet{sawada2019}, and many others).
 Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm.
-However, some works indicated mixed impacts (e.g.~Sawada et al.
-(2019) reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity).
-As stated in J.
-Zhao et al.
-(2021a) and J.
-Zhao et al.
-(2021b), the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
-J.
-Zhao et al.
-(2021a) and J.
-Zhao et al.
-(2021b) assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
+However, some works indicated mixed impacts (e.g.
+\citet{sawada2019} reported an improvement in the forecast of the track of the storm but a degradation in the forecast intensity).
+As stated in \citet{zhao2021, zhao2021a}, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
+\citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
 They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts.
-Otsuka et al.
-(2015) and Mallick and Jones (2020) found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.
-
-While the assimilation of radiance observations into global models is well established (Eyre et al., 2020), the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, bias correction, and the relatively low model tops used for this application.
-Bao et al.
-(2015) studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
-More recently, Zhu et al.
-(2019) studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
-Wang and Randriamampianina (2021) studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis.
+\citet{otsuka2015} and \citet{mallick2020} found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.
+
+While the assimilation of radiance observations into global models is well established \citep{eyre2020}, the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage (in the case of polar-orbiting satellite observations), bias correction, and the relatively low model tops used for this application.
+\citet{bao2015} studied the impact of assimilating cloud-cleared microwave and infrared radiance data polar orbiting instruments on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data.
+More recently, \citet{zhu2019} studied the impact of assimilating clear sky polar orbiting satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
+\citet{wang2021} studied the impact of assimilating clear sky radiances in the high-resolution Copernicus European Regional Reanalysis.
 They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere.
@@
 
-In particular, in South America, the relatively scarcity of conventional observations such as radiosondes increases the potential impact of the observation sources previously discussed.
-Previous work has shown promising results of mesoscale DA in South America using some of the above mentioned data sources (e.g.~Dillon et al., 2016, 2021; Goncalves de Goncalves et al., 2015).
-In particular Dillon et al.
-(2021) assimilated high resolution surface weather station networks, GOES-16 AMVs, and satellite temperature and moisture retrievals with promising results.
-However, no work has been conducted yet over this region to investigate the potential contribution of the datasets discussed above.
-The main objective of this work is to contribute to the quantification and comparison of the impact of high resolution surface weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based DA system.
-In particular, we will focus on the impact in the context of mid-latitude mesoscale convective system events.
-Another particular goal of this paper is to investigate the impact of these data sources in a region where the conventional observation network is rather sparse and where the potential contributions of these observing systems is larger.
-
-To reach this goal, we conduct several DA experiments for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
+The main objective of this work is thus to contribute to the quantification and comparison of the impact of high resolution automatic weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based regional DA system.
+This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps.
+In particular, this paper focuses on the impact in the context of a mid-latitude mesoscale convective system.
+To reach this goal, several DA experiments are conducted for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.
 
@@
 
-\hypertarget{case-study}{%
-\subsection{Case Study}\label{case-study}}
-
-On Nov 22, 2018 a cold front crossed the center of Argentina triggering isolated convective cells that rapidly grew upscale into an exceptionally large MCS.
-To the north of the region, a warm front contributed to the development of isolated multicells that ultimately grew and merged with the MCS.
+\hypertarget{case-overview}{%
+\subsection{Case overview}\label{case-overview}}
+
+Previously to the development of this case study, the center and north of Argentina was immersed in a warm and humid air mass with high values of convective available potential energy (CAPE), as shown by ERA 5 Reanalysis \citep{era5pressure} in Figure \ref{fig:case}a.
+On Nov 22, 2018 a cold front crossed the center of Argentina (Figure \ref{fig:case}b).
+This cold front triggered isolated convective cells that rapidly grew upscale into an exceptionally large MCS (Figure \ref{fig:case}d,e).
+During that day several surface stations reported lightning, strong wind gusts, and heavy rain.
+To the north of the region, a warm and humid environment contributed to the development of isolated convection that ultimately grew and merged with the MCS (Figure \ref{fig:case}f).
 The MCS traveled approximately 2500 km from south to north, dissipating over Paraguay and Southern Brazil after 42 hours.
+
+\begin{figure*}
+
+{\centering \includegraphics{../figures/case-1}
+
+}
+
+\caption{ERA5 Reanalysis of sea level pressure (hPa, black contours), 1000-500 hPa thickness (red dashed contours) and convective available potential energy (shaded) and GOES-16 channel 13 brightness temperature for a,d) 00 and b,e) 12 UTC Nov 22 and c,f) 00 UTC Nov 23.}\label{fig:case}
+\end{figure*}
 
@@
 
-The model simulations for the case study are performed using version 3.9.1 of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW, Skamarock et al.
-(2008)) model.
+The forecast model uses the non-hydrostatic Advanced Research version of Weather Research and Forecasting (WRF-ARW V3.9.1, \citet{skamarock2008}).
 The horizontal grid spacing is 10 km (150 x 200 grid points) in the horizontal and 37 levels in the vertical with the top of the model at 50 hPa.
-The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
-Department of Commerce (2015)).
+The initial and boundary conditions are provided by the Global Forecast System (GFS) analysis (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; \citet{cisl_rda_ds084.1}).
+In this case, a single nesting approach is used since the resolution gap between the driving model and the regional model is not too large (0.25\(^{\circ}\) or 25 km approximately to 10 km).
+This approach is also based on recent studies which suggest that using multiple nested domains does not necessarily lead to improved precipitation forecasts in regional domains, particularly in areas of complex terrain (e.g.
+\citet{liang2019}, \citet{beck2004}).
 The domain covers the area indicated in Figure \ref{fig:dominio} to capture the development of the MCS during the simulated period.
 
-The analyses are generated using the LETKF implementation (V1.3, Hunt et al. (2007)) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Shao et al. (2016)).
+The analyses are generated using the LETKF implementation (V1.3, \citet{hunt2007}) of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; \citet{shao2016}).
 A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within \(\pm\) 30 minutes of the analysis time are assimilated.
 Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
-For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) is used as an observation operator to calculate model-simulated brightness temperatures.
-
-We use a 60-member ensemble where the ensemble mean at the beginning of the DA cycle is initialized using the GFS deterministic analysis with random perturbations to generate the initial ensemble perturbations.
-The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in Necker et al.
-(2020) and Maldonado et al.
-(2021).
-In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales.
-This method helps to prevent an underestimation of the ensemble spread (Ouaraini et al., 2015), and the random perturbations are also applied at the boundaries to maintain proper levels of ensemble spread within the domain.
-
-In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system.
-We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain--Fritsch (Kain, 2004), Grell--Freitas (Grell and Freitas, 2013), and Betts--Miller--Janjic (Janjić, 1994)) and 3 planetary boundary layer schemes (Yonsei University Scheme (Hong, Noh, et al., 2006), Mellor--Yamada--Janjic Scheme (Janjić, 1994), and Mellor--Yamada Nakanishi Niino (Nakanishi and Niino, 2009)).
-All ensemble members use the same land-surface model (Noah-MP, Chen and Dudhia (2001)), microphysics (WRF single-moment 6--class scheme (Hong, Kim, et al., 2006)), and radiation processes (RRTMG shortwave and longwave scheme (Iacono et al., 2008)) parameterizations.
-
-To reduce the effect of spurious correlations in the estimation of error covariances, we use a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) as in Dillon et al.
-(2021) for all types of observations.
-A relaxation-to-prior inflation (Whitaker and Hamill, 2012) is applied with an inflation parameter \(\alpha=0.9\) to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
+For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; \citet{han2006}) is used as an observation operator to calculate model-simulated brightness temperatures.
+
+A 60-member ensemble is used where the initial ensemble mean and the mean boundary conditions are taken from the GFS deterministic analysis.
+A set of 60 perturbations are randomly generated to perturb the initial state as well as the boundary conditions during the length of the experiment.
+Perturbing the boundary conditions helps to reduce the impact of errors in the driving global model and helps to keep a larger ensemble spread throughout the domain and during the length of the experiment \citep{ouaraini2015}.
+The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5\(^{\circ}\) horizontal grid spacing with a smooth time evolution as in \citet{necker2020} and \citet{maldonado2021}.
+In this way, the nearly hydrostatic and geostrophic equilibrium of larger scales is preserved.
+The random perturbations used are the same across experiments to ensure that the differences between experiments are only related to changes in the number and type of assimilated observations.
+
+A multi-physics scheme is used to better represent the uncertainty in the model formulation within the DA system.
+9 different model configurations are generated consisting of the combination of 3 moist convection schemes (Kain--Fritsch \citep{kain2004}, Grell--Freitas \citep{grell2013}, and Betts--Miller--Janjic \citep{janjic1994}) and 3 planetary boundary layer schemes (Yonsei University Scheme \citep{hong2006}, Mellor--Yamada--Janjic Scheme \citep{janjic1994}, and Mellor--Yamada Nakanishi Niino \citep{nakanishi2009}).
+The distribution of these schemes among the 60 ensemble members is outlined in Table \ref{tab:miembros-desc}.
+The multi-physics approach is also introduced in order to represent the uncertainty associated with the more relevant physical processes that are not resolved by the model.
+All ensemble members use the same land-surface model (Noah-MP, \citet{chen2001}), microphysics (WRF single-moment 6--class scheme \citep{hong2006a}), and radiation processes (RRTMG shortwave and longwave scheme \citep{iacono2008}) parameterizations.
+
+\begin{table}
+
+\caption{\label{tab:miembros-desc}Generation of the 60-member multi-physics ensemble as a combination of Cumulus and PBL parameterizations.}
+\centering
+\fontsize{6}{8}\selectfont
+\begin{tabular}[t]{c>{\centering\arraybackslash}p{7em}>{\centering\arraybackslash}p{7em}>{\centering\arraybackslash}p{7em}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{PBL} \\
+\cmidrule(l{3pt}r{3pt}){2-4}
+Cumulus & MYJ & MYNN2 & YSU\\
+\midrule
+BMJ & 5, 14, 23, 32, 41, 50, 59 & 8, 17, 26, 35, 44, 53 & 2, 11, 20, 29, 38, 47, 56\\
+GF & 6, 15, 24, 33, 42, 51, 60 & 9, 18, 27, 36, 45, 54 & 3, 12, 21, 30, 39, 48, 57\\
+KF & 4, 13, 22, 31, 40, 49, 58 & 7, 16, 25, 34, 43, 52 & 1, 10, 19, 28, 37, 46, 55\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+To reduce the effect of spurious correlations in the estimation of error covariances, a horizontal localization radius of 180 km and a vertical localization radius of 0.4 (in log pressure coordinates) is used as in \citet{dillon2021} for all types of observations.
+A relaxation-to-prior spread inflation \citep{whitaker2012} is applied with an inflation parameter \(\alpha=0.9\) following \citet{maldonado2020} to mitigate the impact of sampling errors and to consider model errors not accounted for by the multi-model ensemble approach.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
+\includegraphics[width=0.98\linewidth]{../figures/dominio-1} \caption{a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles).
 b) Locations of radiosonde launches during RELAMPAGO.
@@
 
-We also assimilate data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America.
-The dataset used in this study has been obtained from the RELAMPAGO Data Set repository (Garcia et al., 2019).
+Data from 866 Automatic Weather Stations (AWS) that are part of 17 public and private surface networks over Southern South America are also assimilated.
+The dataset used in this study has been obtained from the RELAMPAGO Data Set repository \citep{garcia2019}.
 These stations are indicated as green squares in Figure \ref{fig:dominio}a.
@@
 \hypertarget{satellite-derived-winds}{%
-\subsubsection{Satellite derived winds}\label{satellite-derived-winds}}
+\subsubsection{Satellite-derived winds}\label{satellite-derived-winds}}
 
@@
 Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 \% of the observations).
-Observation errors used to assimilate these observations are indicated in Table \ref{tab:table-obs}.
+Observation errors used to assimilate these observations follow the GSI default configuration and are indicated in Table \ref{tab:table-obs}.
 
@@
 
-The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
+The channels adopted for assimilation and their associated errors were defined taking into account the low model top (50 hPa).
 The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically.
-First, a spatial data thinning is applied using a 60 km grid following Singh et al.
-(2016), Jones et al.
-(2013), and Lin et al.
-(2017), where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
-Also, observations over the sea are preferred to those over land or snow (Hu et al., 2018).
+First, a spatial data thinning is applied using a 60 km grid following \citet{singh2016}, \citet{jones2013}, and \citet{lin2017a}, where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control.
+Also, observations over the sea are preferred to those over land or snow \citep{hu2018}.
 
 The thinned observations were then bias corrected.
-The bias correction (BC) has an air-mass dependent and an angle-dependent component (Zhu et al., 2014) and it is calculated as a multi-linear function of N predictors \(p_i(x)\), with associated coefficients \(\beta_i\).
+The bias correction (BC) has an air-mass dependent and an angle-dependent component \citep{zhu2014} and it is calculated as a multi-linear function of N predictors \(p_i(x)\), with associated coefficients \(\beta_i\).
 Then, the bias corrected brightness temperature (\(BT_{bc}\)) can be obtained as:
 
-\[\mathit{BT_{bc}} =\mathit{ BT} + \sum_{i = 0}^{N} \beta_i p_i (x)\]
+\begin{equation}
+  \mathrm{\mathit{BT_{bc}} =\mathit{ BT} + \sum_{i = 0}^{N} \beta_i p_i (x)}
+  \label{eq:eq1}
+\end{equation}
 
 GSI has a constant offset bias correction term (\(p_0 = 1\)) and the remaining predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity.
-Scan angle-dependent bias is modeled as a 4th-order polynomial (Zhu et al., 2014).
+Scan angle-dependent bias is modeled as a 4th-order polynomial \citep{zhu2014}.
 
@@
 In our experiments, only clear-sky observations are used.
-For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes (Weston et al., 2019; Zhu et al., 2016).
+For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes \citep{weston2019, zhu2016}.
 For the infrared channels, cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms.
+Moreover, GSI checks the difference between the observations and simulated brightness temperature with height to detect cloudy pixels.
 Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenith angle (over 60°) to reject channels near the visible range that can be contaminated with reflection.
@@
 \toprule
-Sensor & Platform & \makecell[r]{Assimilated\\channels} & \makecell[l]{Percentage\\over total}\\
+Sensor & Platform & Assimilated channels & Percentage over total\\
 \midrule
@@
 \item
-  The Multi-Network Composite Highest Resolution Radiosonde Data (Earth Observing Laboratory, 2020) from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the IOPs along with the operational radiosondes.
+  ERA5 hourly data on pressure levels from 1959 to present \citep{era5pressure}.
+The variables of interest (air temperature, humidity and wind) were interpolated to the model grid to compare them with the analysis of each experiment.
+\item
+  The Multi-Network Composite Highest Resolution Radiosonde Data \citep{sondeos} from the RELAMPAGO field campaign database consisting of high-resolution radiosondes launched from several locations during the IOPs along with the operational radiosondes.
 Only the soundings that did not enter the assimilation system were used for validation.
@@
 \item
-  The Satellite precipitation estimation IMERG Final Run with 0.01\(^{\circ}\) spatial resolution and 30 minutes temporal resolution (Huffman et al., 2018) was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
+  The Satellite precipitation estimation IMERG Final Run with 0.01\(^{\circ}\) spatial resolution and 30 minutes temporal resolution \citep{huffman2018} was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
 \item
   Radar observations are used to perform a qualitative and visual assessment of the convective features.
-The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network (de Elía et al., 2017) with a temporal frequency of 10 minutes.
+The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network \citep{deelia2017} with a temporal frequency of 10 minutes.
 For this work, only the maximum reflectivity in the column (COLMAX) closest to the analysis time was used.
@@
 The CONV experiment uses only conventional observations from PREPBUFR.
-In a second experiment, referred to as AWS, we assimilate all the observations included in CONV plus the 10-minute frequency surface observations from AWS.
-In the third experiment, referred to as SATWND, we assimilate all the observations of the AWS experiment and the satellite-derived winds.
+In a second experiment, referred to as AWS, all the observations included in CONV are assimilated plus the 10-minute frequency surface observations from AWS.
+In the third experiment, referred to as SATWND, the observations from the AWS experiment along with the satellite-derived winds are assimilated.
 Finally, a fourth experiment referred to as RAD assimilates all available clear-sky radiances from sensors onboard polar orbiting satellites as described in section \ref{sat}.
@@
 All forecasts use the same domain and ensemble configuration as the analysis.
-The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; National Centers for Environmental Prediction, National Weather Service, NOAA, U.S.
-Department of Commerce (2015)).
+The boundary conditions for the ensemble members are generated by adding random perturbations to the GFS deterministic forecast (0.25\(^{\circ}\) horizontal grid spacing and 6-hour temporal resolution; \citet{cisl_rda_ds084.1}).
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/analysis_cycle} \caption{Figure 4.
-Diagram of the analysis cycles between 18 UTC Nov 20, and 12 UTC Nov 23 plus spin up period of 6 hours.
+\includegraphics[width=1\linewidth]{../figures/analysis_cycle} \caption{Diagram of the analysis cycles between 18 UTC Nov 20, and 12 UTC Nov 23 plus spin up period of 6 hours.
 The zoomed section shows the hourly assimilation that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours.
@@
 
-To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use the Reduced Centered Random Variable (RCRV, Candille et al. (2007)) which is defined as:
-
-\[RCRV = \frac{x_o - m}{\sqrt{\sigma_o^2 + \sigma^2}}\]
+To evaluate the statistical consistency of the uncertainty quantification in the ensemble system the Reduced Centered Random Variable (RCRV, \citet{candille2007}) is used which is defined as:
+
+\begin{equation}
+  \mathrm{RCRV = \frac{m - x_o}{\sqrt{\sigma_o^2 + \sigma^2}}}
+  \label{eq:eq2}
+\end{equation}
 
 where \(x_o\) is the assimilated observation and its error \(\sigma_o\), the ensemble mean of the analysis in observational space \(m\), and the standard deviation \(\sigma\) of the ensemble.
-The average of \(RCRV\) computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
-
-\[\mathit{mean RCRV} = E[RCRV]\]
-
-The standard deviation of the \(RCRV\) or \(sd RCRV\) measures the agreement of the ensemble spread and the observational error with respect to the distance between the ensemble mean and the observations, and then the systematic over- or under- dispersion of the ensemble:
-
-\[\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}\]
+The \(RCRV\) is the ratio of the distance between the observations and the forecast and its expected standard deviation assuming the statistical independence between the forecast error (estimated from the ensemble spread) and the observation error.
+The average of \(RCRV\) computed over all the analysis cycles represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:
+
+\begin{equation}
+  \mathrm{\mathit{mean RCRV} = E[RCRV]}
+  \label{eq:eq3}
+\end{equation}
+
+If the ensemble has a positive bias, \(mean RCRV\) will be positive, on the opposite, if the ensemble has a negative bias, \(mean RCRV\) will be negative.
+The standard deviation of the \(RCRV\) or \(sd RCRV\) is defined as:
+
+\begin{equation}
+  \mathrm{\mathit{sd RCRV} = \sqrt{\frac{1}{M -1}\sum_{i=1}^{M}(\mathit{RCRV_i} - \mathit{mean RCRV})^2}}
+  \label{eq:eq4}
+\end{equation}
+
 where \(M\) is the ensemble size.
+The \(sd RCRV\) measures how large is the distance between the forecast and the observations with respect to the expected distance (given by the combination of the ensemble spread and the observation error).
+Assuming that the observation error is accurately estimated, an \(sd RCRV > 1\) indicates that the ensemble is underdispersive (i.e.~the distance between the observations and the forecasts is larger than expected), and an \(sd RCRV < 1\) indicates that the ensemble is overdispersive (i.e.~the distance between the observations and the forecasts is lower than expected).
 A consistent system will have no bias (\(mean RCRV = 0\)) and a standard deviation equal to 1 (\(sd RCRV = 1\)).
-If the ensemble has a positive bias, \(mean RCRV\) will be positive, on the opposite, if the ensemble has a negative bias, \(mean RCRV\) will be negative.
-A \(sd RCRV > 1\) () indicates that the ensemble is underdispersive and an \(sd RCRV < 1\) indicates that the ensemble is overdispersive.
 
@@
 
-\[\mathit{RMSE} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)^{2}}\]
-
-\[\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)\]
+\begin{equation}
+  \mathrm{\mathit{RMSE} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)^{2}}}
+  \label{eq:eq5}
+\end{equation}
+
+\begin{equation}
+  \mathrm{\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)}
+  \label{eq:eq6}
+\end{equation}
+
 where \(O\) and \(X\) stand for independent observations and the simulations respectively, and N is the sample size.
 
-For the comparison of the first-guess precipitation with the IMERG precipitation estimates, we compute the Fractions Skill Score (FSS, Roberts (2008)) for different neighborhood length scales and thresholds:
-
-\[\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} \]
+For the comparison of the first-guess precipitation with the IMERG precipitation estimates, the Fractions Skill Score (FSS, \citet{roberts2008}) is computed for different neighborhood length scales and thresholds:
+
+\begin{equation}
+  \mathrm{\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}}}
+  \label{eq:eq7}
+\end{equation}
+
 where \(P_{oi}\) is the fraction of grid points in the \(i-th\) sampling area in which the observed accumulated precipitation is greater than a specified threshold.
-Following Roberts et al.
-(2020), \(P_{xi}\) is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the \(i-th\) sampling area.
-The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
+Following \citet{roberts2020}, \(P_{xi}\) is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the \(i-th\) sampling area.
+The FSS was computed from the accumulated precipitation over 6 hr rolling windows by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
 
@@
 
-We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019).
-All the analyses in this paper were performed using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
-All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2019; Xie, 2015).
+All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne \citep{Cheyenne2019}.
+All the analyses in this paper were conducted using the R programming language \citep{rcoreteam2020}, using data.table \citep{dowle2020} and metR \citep{campitelli2020} packages.
+All graphics are made using ggplot2 \citep{wickham2009} and the paper was rendered using knitr and rmarkdown \citep{xie2015, allaire2019}.
 
@@
 
-To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the \(mean RCRV\) and the \(sd RCRV\) for the RAD experiment.
+To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, the \(mean RCRV\) and the \(sd RCRV\) is calculated for the RAD experiment.
 As this experiment assimilates all types of observations used in this work, it is possible to analyze the consistency of the ensemble by comparing it with each type of observation.
@@
 For the temperature (Figure \ref{fig:rcrv-sfc}b), the results are similar except that for some areas in the west of the domain the \(sd RCRV\) can be as high as 4.5.
-In this region, the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.
+These higher values of \(sdRCRV\) can be associated with systematic errors arising from high differences between the model surface and the observations.
+Small scale circulations associated with the complex terrain and not well resolved by the model can also contribute to increase the distance between the forecast and the observations.
+These aspects are usually not captured by the ensemble spread unless a well tuned space dependent inflation scheme is used thus leading to greater sdRCRV values.
 
@@
 AIRCAR presents an irregular profile with \(sd RCRV\) values that suggest that the error for this type of observation is overestimated.
-ADPUPA and AIRCAR present a \(mean RCRV\) profile near zero at middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in Ruiz et al.
-(2010) and Dillon et al.
-(2021).
-
-For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) at low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error.
+ADPUPA and AIRCAR present a \(mean RCRV\) profile near zero at middle and upper levels.
+At low levels, the meanRCRV profile is positive, showing a cold bias present in the model, a characteristic already studied in \citet{ruiz2010} and \citet{dillon2021}.
+
+Satellite-derived winds observations vary in number depending on the satellite and the level.
+In Figure \ref{fig:rcrv-profile}c only the \(RCRV\) calculated with at least 100 observations for each satellite and level is included.
+At low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) show a larger departure from the expected behavior with a negative bias, and a possible overestimation of the observation error.
 Wind estimations derived from water vapor channels are abundant above 500 hPa where their bias is close to zero.
@@
 Overall, these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work.
-Moreover, these results suggest the inflation parameter \(\alpha = 0.9\) is adequate for the system.
+Moreover, these results suggest the relaxation-to-prior spread inflation parameter \(\alpha = 0.9\) is adequate for the system.
 
@@
 
-In this section, we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
+This section presents the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection.
 The analysis is performed over a smaller domain (red box in Figure \ref{fig:dominio}a) to focus on the region most directly affected by the MCS.
 Figures \ref{fig:TQ-diff}a-c show the analysis difference between experiments in the spatially averaged vertical profile of temperature.
-By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state.
+By averaging the differences between two experiments the systematic impact produced by different observing systems on the analyzed state can be isolated.
 During the first day, the assimilation of AWS observations results in a colder PBL.
@@
 Comparing the specific humidity in the experiments (Figures \ref{fig:TQ-diff}d-f), the impact of assimilating AWS with fine spatial and temporal resolution is most substantial at low levels (Figure \ref{fig:TQ-diff}d).
-The PBL in the AWS experiment is consistently moister than in the CONV experiment particularly at nighttime.
-The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region (Ruiz et al., 2010).
-We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL.
-In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
+The PBL in the AWS experiment is consistently moister than in the CONV experiment, particularly at nighttime.
+The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region \citep[\citet{matsudo2021}, \citet{ruiz2010}]{casaretto2022}.
+The moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL.
+In the experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections.
 As for the temperature, the systematic impact of satellite-derived winds on moisture is small (Figure \ref{fig:TQ-diff}e).
@@
 
-\begin{figure*}[th]
+\begin{figure*}[ht]
 
@@
 
-\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(gkg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:TQ-diff}
+\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of temperature (a, b, and c, in \(K\)) and specific humidity (d, e, and f in \(g\ kg^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:TQ-diff}
 \end{figure*}
@@
 This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment.
-After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain.
+After 18 UTC Nov 22, the opposite effect is observed when the MCS is moving through the domain to the northeast.
 After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels (Figures \ref{fig:UV-diff}d, f).
@@
 Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is largest; the systematic impact is overall smaller than the one produced by assimilating data from AWS (Figures \ref{fig:UV-diff}b, e).
+The reason of the small impact observed in SATWND could be associated to the large observation error used for satellite-derived wind observations.
 
@@
 
-\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in \(ms^{-1}\)) and v wind (d, e, and f in \(ms^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.
-Black contours correspond to negative u wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD.
-Dashed contours correspond to negative v wind.}\label{fig:UV-diff}
-\end{figure*}
-
-To investigate how changes in the PBL can modify the pre-convective environment, we compare the horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov 22 (after 30 assimilation cycles).
-At that moment the first convective cells were developing over the southern region of the domain along the cold front.
+\caption{Difference between analysis ensemble mean experiments a) and d) AWS-CONV, b) and e) SATWND-AWS, and c) and f) RAD-SATWND for the spatially averaged vertical profiles of u wind (a, b, and c, in \(m\ s^{-1}\)) and v wind (d, e, and f in \(m\ s^{-1}\)) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.
+Black contours correspond to u wind and dashed contours to negative v wind for (a) AWS, (b) SATWND, and (c) RAD and v wind for (d) AWS, (e) SATWND, and (f) RAD since those experiments are the ones with more assimilated observations in each panel.}\label{fig:UV-diff}
+\end{figure*}
+
+The difference between ERA5 reanalyses \citep[`]{era5pressure} and the ensemble mean analyses are also compared in Figure \ref{fig:era5}, which supports Figures \ref{fig:TQ-diff} and \ref{fig:UV-diff}.
+Specifically, Figure \ref{fig:era5}a shows a warm bias in low levels (i.e.~CONV is warmer than ERA5) that decreases in Figure \ref{fig:era5}b when the AWS observations are assimilated.
+In the same direction, Figure \ref{fig:TQ-diff}a shows a negative difference between AWS and CONV meaning that the AWS observations are cooling the low levels.
+Comparing ERA5-RAD (Figure \ref{fig:era5}d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations as shown in Figure \ref{fig:TQ-diff}c.~A similar effect can be observed for specific humidity, AWS observations partially correct the dry bias present in Figure \ref{fig:era5}e and the assimilation of radiance observations reduces the positive impact of AWS.
+The impact on the wind components is minor so only the meridional wind is included in Figures \ref{fig:era5}i-l, which show that the radiance observations are mainly responsible for the positive impact observed in the analysis by reducing the distance ERA5-RAD, particularly during the mature stage of the MCS.
+Overall, the adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5 reanalyses.
+
+\begin{figure*}[ht]
+
+{\centering \includegraphics{../figures/era5-1}
+
+}
+
+\caption{Difference between ERA5 and the analysis ensemble mean experiments for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity (\(g\ Kg^{-1}\), e--h) and meridional wind (\(m\ s^{-1}\), i--l) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:era5}
+\end{figure*}
+
+To investigate how changes in the PBL can modify the pre-convective environment, the analysis mean horizontal distribution of the low level northerly flow (for the first 7 sigma levels), precipitable water, low level temperature, and CAPE are compared.
+At 00 UTC Nov 22 (after 30 assimilation cycles) the first convective cells were developing over the southern region of the domain along the cold front.
 Figure \ref{fig:summary-fields}a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours).
-We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
+It shows that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations.
 The moisture increase is particularly strong at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place.
@@
 On average the PBL in AWS and SATWND is colder than in CONV, while RAD shows a warmer PBL than AWS due to the assimilation of radiance observations.
+A warmer PBL increases the potential instability and helps to generate a suitable environment for the development of deep convection.
 Figure \ref{fig:summary-fields}c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 \(km\) wind shear.
-The values of MCAPE in CONV do not exceed 2000 \(JKg^{-1}\) while the rest of the experiments show maximum MCAPE over 4000 \(JKg^{-1}\).
+The values of MCAPE in CONV do not exceed 2000 \(J\ Kg^{-1}\) while the rest of the experiments show maximum MCAPE over 4000 \(J\ Kg^{-1}\).
 MCAPE in the RAD experiment is lower compared to AWS or SATWND.
 This is consistent with less humidity in the PBL with respect to these experiments but may be partially compensated by a slightly warmer PBL in the RAD experiment.
-The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values.
+The 0-6 km wind shear is more intense in AWS, SATWND, and RAD reaching values over 15 \(m\ s^{-1}\) at the southern tip of the region with positive MCAPE values.
 Moreover, in this same region, these experiments show larger MCAPE values than CONV.
-Note that wind shear over 15 \(ms^{-1}\) is associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowski and Richardson, 2010).
+Note that wind shear over 15 \(m\ s^{-1}\) is associated with the development of more intense and organized MCSs \citep{chen2015} and also with conditions favorable for supercells \citep{markowski2010}.
 
 \begin{figure*}
-\includegraphics[width=1\linewidth]{../figures/summary-fields-1} \caption{a) Precipitable water (shaded, \(kgm^{-2}\)) and average northerly wind over the first 7 sigma levels (from the surface up to approximately 800 hPa, contours, \(ms^{-1}\)), b) Average potential temperature for the PBL (first 10 sigma levels), and c) Maximum CAPE and \textasciitilde0-6 km wind shear over 15 and 30 \(ms^{-1}\) for each experiment.
+\includegraphics[width=1\linewidth]{../figures/summary-fields-1} \caption{a) Precipitable water (shaded, \(kg\ m^{-2}\)) and average northerly wind over the first 7 sigma levels (from the surface up to approximately 800 hPa, contours, \(m\ s^{-1}\)), b) Average potential temperature for the PBL (first 10 sigma levels), and c) Maximum CAPE and \textasciitilde0-6 km wind shear over 15 and 30 \(m\ s^{-1}\) for each experiment.
 All fields correspond to the analysis ensemble mean for 00 UTC Nov 22.
@@
 
-First, we analyze the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation.
-Figure \ref{fig:pp-hov}a shows the hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) (Clark, 2017) for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments.
+First, the impact of assimilating different observation types in terms of the representation of the MCS and its associated precipitation is analyzed.
+Figure \ref{fig:pp-hov}a shows the hourly accumulated precipitation as estimated by IMERG, and the probability matched mean (PM) \citep{clark2017} for the first-guess hourly accumulated precipitation as averaged between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W as a function of time and latitude in the different experiments.
 The heaviest precipitation (over 12 \(mmh^{-1}\)) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period (Figure \ref{fig:pp-hov}a).
@@
 \begin{figure*}[h]
-\includegraphics{../figures/pp-hov-1} \caption{Hövmoller diagram of probability matched mean hourly accumulated 1-h forecast precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W.
-Contours drawn every 0.5 \(mmh^{-1}\), starting at 0.5 \(mmh^{-1}\).}\label{fig:pp-hov}
-\end{figure*}
-
-To quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments, we compute the FSS in 6-hour moving windows for different thresholds and spatial scales (Figure \ref{fig:fss}).
+
+{\centering \includegraphics{../figures/pp-hov-1}
+
+}
+
+\caption{Hövmoller diagram of probability matched mean hourly accumulated 1-h forecast precipitation for each latitude band estimated by IMERG (left) and simulated (right), for the ensemble mean of each experiment, averaged over a longitude range between 67\(^{\circ}\)W and 54.5\(^{\circ}\)W.
+Contours drawn every 0.5 \(mm\ h^{-1}\), starting at 0.5 \(mm\ h^{-1}\).}\label{fig:pp-hov}
+\end{figure*}
+
+The FSS is computed to quantify the spatial match between the observed precipitation and the first-guess hourly accumulated precipitation for the different experiments (Figure \ref{fig:fss}).
+For each threshold and spatial scale, Equation @red(eq:eq7) is applied in 6-hours rolling windows throughout the experiment period.
 All experiments show similar values of FSS during the initiation of the convection before 06 UTC Nov 22 except for RAD which performs better than the rest of the experiments during this period.
@@
 This is partially expected considering the coarse horizontal grid spacing (10 km), which is not enough to appropriately represent the strength of the convective band associated with the MCS.
-Also, the ensemble mean produces a smoothing effect that reduces the amplitude of local maxima in the forecast reflectivity field.
 RAD better represents the observed features of the system showing a stronger and more organized MCS than CONV, over the domain center at 10 and 13 UTC (first and second columns in Figure \ref{fig:dbz-mean}).
@@
 In addition, CONV captures the location of the MCS, but the convection seems to be less organized and much weaker than in RAD.
-Before and after the times shown in Figure \ref{fig:dbz-mean}, the agreement between experiment and observations is quite good in the regions where radar data are available, especially for RAD.
+Before and after the times shown in Figure \ref{fig:dbz-mean}, the agreement between location of the observed convective cells and the simulated in the experiment is quite good in the regions where radar data are available, especially for RAD.
 
@@
 AWS observations were able to reduce the RMSE for temperature and dew point temperature in the PBL and reduce a small dry bias.
-However, AWS degradated the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \ref{fig:soundings}c).
-
-\begin{figure*}[h]
+However, in this region (Figure \ref{fig:dominio}b) and for this period, AWS increments (Figure \ref{fig:UV-diff}d) degrades the zonal wind between 7 and 12 km increasing the bias and RMSE (Figure \ref{fig:soundings}c).
+
+\begin{figure*}[ht]
 \includegraphics[width=1\linewidth]{../figures/dbz-mean-1} \caption{Maximum reflectivity in the column (COLMAX in \(dBZ\)), observed (upper row) and 1-hr forecast probability matched mean column maximum reflectivity for CONV (second row) and RAD (third row) at 10 UTC (first column), 13 UTC (second column), 16 UTC (third column), and 19 UTC (fourth column) Nov 22, 2018.
@@
 
-\begin{figure*}[t]
+\begin{figure*}[ht]
 
@@
 
-\caption{RMSE (solid line) and Bias (dashed line) of a) temperature (\(K\)), b) dew point temperature (\(K\)), c) u wind (\(ms^{-1}\)) and d) v wind (\(ms^{-1}\)) calculated by comparing the analysis of each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
+\caption{RMSE (solid line) and Bias (dashed line) of a) temperature (\(K\)), b) dew point temperature (\(K\)), c) u wind (\(m\ s^{-1}\)) and d) v wind (\(m\ s^{-1}\)) calculated by comparing the analysis of each experiment with the RELAMPAGO soundings during IOP 7 and IOP 8.
 The blue line corresponds to CONV, the light blue line to AWS, SATWND is represented with an orange line, and RAD with a red line.}\label{fig:soundings}
@@
 
-In this section, we analyze the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
-We calculated the FSS for the ensemble forecasts in 6-hour moving windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}).
+This section analyzes the 60-member ensemble forecast initialized at 00 and 06 UTC Nov 22 from each experiment that runs for 36 and 30 h respectively, until 12 UTC Nov 23.
+The FSS is again calculated for the ensemble forecasts in 6-hour rolling windows for the same thresholds and spatial scales as for the first-guess hourly accumulated precipitation to quantify the skill of the forecasts to predict precipitation (Figure \ref{fig:fssfcst}).
 CONV forecasts perform very poorly in terms of the FSS compared with the experiments that include other sources of observations.
@@
 Conversely, the radiance observations resulted in a neutral to a slightly negative impact on the forecast as opposed to what was seen when comparing the 1-h forecast to IMERG estimations.
+The reason why the forecasts initialized from RAD degrade over time needs to be further study.
+However, it is possible that the assimilation of observations associated with channels affected by the surface is contributing to the degradation of the PBL in the analysis and subsequently in the forecasts.
+\citet{lim2014} observed limited impact when assimilating AIRS observations and attribute this result to the use of surface channels where the uncertainties associated with emissivity are large.
 
@@
 
-In this paper, we investigate for the first time in South America the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system.
-In particular, we explore the impact upon the analysis quality of assimilating frequent and relatively dense surface observations, satellite derived winds, and satellite clear-sky radiances from multiple sensors.
-Southern South America is a particularly interesting region due to the heterogeneity and sometimes coarse resolution of the operational observing network (considering both surface based and upper air observations).
-This combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.
-To consider these issues, we used a case study approach of a massive mesoscale convective system that developed over Southern South America, particularly in a region where the surface and upper-air observing networks are too coarse to capture variability at the mesoscale and that took place on Nov 22, 2018.
-
-We evaluated the consistency of the ensemble using the RCRV calculated for each type of observation used in RAD.
-Conventional observations including surface data and upper air observations from soundings and aircraft, show a \(sd RCRV\) near 1 indicating a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
-For satellite-derived winds and radiance observations, the \(sd RCRV\) is lower than 1, which could be the result of an underdispersive ensemble or an overestimation of the observation errors.
-However, taking into account that the observation error for satellite-derived wind is overestimated (to avoid strong impacts on the analysis for poor quality estimations) and that the results obtained for more reliable observations (such as soundings) were good, we conclude that the ensemble has a reasonable spread.
-
-In terms of the analysis, we found that AWS observations, which have high spatial and temporal resolution, produce impacts throughout the troposphere, especially in the PBL.
-In particular, the precipitable water content and the low level meridional circulation in the AWS experiment lead to the development of deep convection and heavy precipitation closer to the observed in this case study.
-We also found positive results when assimilating radiance observations that produce an adequate development of the convection, mainly during the mature stage of the MCS, leading to an increase in the accumulated precipitation compared with the other experiments.
-However, all experiments underestimated the 1-hour accumulated precipitation when compared to IMERG estimates.
-While the assimilation of satellite-derived wind does not produce a noticeable impact on the analysis with respect to the AWS experiment, this is possibly due to the small number of observations in low levels available for this case study and its large observation error.
-However, there are substantial improvements for the accumulated precipitation distribution.
-A more comprehensive analysis is necessary to assess the possible impacts on middle and upper levels.
-
-Taking advantage of the observations taken during the RELAMPAGO field campaign, we compared the experiments with the soundings launched during IOPs 7 and 8.
-For the pre-convective environment (IOP 7), we observed an improvement at low levels when assimilating AWS observations for the temperature, dew point temperature, and meridional wind.
-In particular, AWS observations correct a dry bias already observed by Ruiz et al.
-(2010).
-For IOP 8, after the MCS crossed the center of the domain, we found a positive impact in low levels when assimilating more observations for the temperature but the opposite effect for the dew point temperature.
-In this case, the analyses have a moist bias that increases with the assimilation of AWS and radiance observations.
-We noted a negative covariance between temperature and moisture content over this region that could lead to a moistening of the PBL.
-It is not clear what produces these negative covariances but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content during this period, which requires further analysis.
-
-We also analyzed the performance of independent ensemble forecasts intialized from the analyses to forecast the precipitation during Nov 22.
-The forecast initialized from AWS, SATWND, and RAD are able to forecast the precipitation substantially better than CONV.
-In particular, the continuous assimilation of satellite-derived wind and radiance observations helps to improve the latest initialization but only the satellite-derived wind observations produces a positive impact that persists throughout the forecast.
-
-To summarize, in this study we conclude that the assimilation of surface observations with high spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
-Moreover, the ensemble forecasts initialized from the analysis show promising results to predict the precipitation of severe events.
-In the future, we will further analyze the impacts in the independent forecasts and evaluate the implementation of other sources of observations like GPS radio occultation data and radiances from geostationary orbiting satellites.
+Southern South America is a particularly interesting region due to the heterogeneity in topography and coarse resolution of the operational observing network (considering both surface based and upper air observations).
+This, combined with a climatology characterized by frequent organized convective events makes mesoscale DA particularly challenging.
+This paper investigates, for the first time in South America, using a case-study approach, the impact of different observation systems on the performance of an ensemble-based mesoscale regional DA system.
+This case study corresponds to a massive MCS that developed over Southern South America on Nov 22, 2018 during the RELAMPAGO field campaign.
+In particular, the impact on the analysis quality of assimilating frequent and relatively dense surface observations, satellite-derived winds, and satellite clear-sky radiances from multiple sensors is explored.
+
+Firstly, the consistency of the ensemble was evaluated to ensure a good agreement between the ensemble spread and the observational errors with respect to the distance between the ensemble mean and the observations.
+While conventional observations departures are consistent with the ensemble spread and assumed observation errors, satellite-derived winds and radiance observations departures are lower than expected.
+The latter could be the result of an overestimation of the observation errors which is usually introduced to avoid the detrimental impact on the analysis of poor quality observations.
+In this case study, all the observation types considered (i.e.~automatic weather stations, satellite derived winds and clear-sky radiances from polar orbiting satellites) improves the quality of the analysis and of the short range forecast with respect to the conventional observation network.
+In terms of the analysis, automatic weather station observations, which have high spatial and temporal resolution, produced impacts mainly within the PBL but which occasionally extends throughout the troposphere during the periods where moist convection is stronger within the domain.
+These observations also helped to reduce the warm and dry bias present in the model, producing an analysis closer to the ERA5 reanalysis.
+During the pre-convective environment, assimilating surface temperature, dew point temperature, and meridional wind improved the analysis at low levels when compared with observed soundings.
+In particular, when these observations are assimilated, precipitable water content and low level meridional circulation led to the enhancement of deep convection and heavy precipitation which is closer to observations.
+
+Positive results were also found when assimilating radiance observations, which produced a better development of the convection and its associated outflow circulation, mainly during the mature stage of the MCS, leading to increased accumulated precipitation compared to the case in which these observations are not assimilated.
+However, these observations weakened the impact of automatic weather station observations within the PBL, slightly increasing the warm and dry bias with respect to ERA5.
+While this needs to be further studied, it could be related to the assimilation of channels affected by the surface or sub-optimal bias correction.
+Comparing the experiment with independet soundings, the assimilation of radiances improved mid and upper level wind.
+
+The assimilation of satellite-derived wind did not produce a noticeable impact on the analysis.
+This is possibly due to the relatively small number of observations in low levels available for this case study and their large observation error.
+However, there are improvements in the 1-h forecast accumulated precipitation distribution.
+A more comprehensive analysis is necessary to understand the mechanisms behind the impact of these observations on longer range forecasts.
+
+The evaluation of the performance of independent ensemble precipitation forecasts initialized from the analyses during Nov 22 showed that the forecasts initialized from AWS, SATWND, and RAD were able to forecast the precipitation substantially better than CONV.
+In particular, continuous assimilation of satellite-derived wind and radiance observations improved the latest initialization but only satellite-derived wind observations produced a positive impact that persisted throughout the forecast.
+Why the forecast initialized from RAD did not perform better than SATWND needs to be further study.
+
+To summarize, in this case study we found that the assimilation of surface observations with high spatial and temporal resolution, satellite-derived winds, and clear-sky radiances from polar orbiting satellites had an overall positive impact on the development of the studied MCS and its associated precipitation.
+Moreover, ensemble forecasts initialized from the analysis showed promising results for predicting extreme severe precipitation events.
+In the future, we will further analyze the impact of these observations upon short-range forecasts over longer periods and evaluate the assimilation of other sources of observations such as GPS radio occultation data and radiances from geostationary orbiting satellites like GOES-16.
 
@@
 A version-controlled repository of the code used to create this analysis, including the code used to download the data can be found at \url{https://github.com/paocorrales/mesoda}.
-The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5072/zenodo.1060834, version 0.9.0.
-
-\hypertarget{acknowledgements}{%
-\section*{Acknowledgements}\label{acknowledgements}}
-\addcontentsline{toc}{section}{Acknowledgements}
+The derived data that support the findings of this study are also openly available in Zenodo at http://doi.org/10.5281/zenodo.7015913, version 0.9.2.
+
+\hypertarget{acknowledgments}{%
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-allaire2019}{}%
-Allaire, J., Horner, J., Xie, Y., Marti, V., Porte, N., 2019. Markdown: Render markdown with the c library 'sundown'.
-
-\leavevmode\hypertarget{ref-bae2022}{}%
-Bae, J.-H., Min, K.-H., 2022.
-Forecast Characteristics of Radar Data Assimilation Based on the Scales of Precipitation Systems.
-Remote Sensing 14, 605.
-doi:\href{https://doi.org/10.3390/rs14030605}{10.3390/rs14030605}
-
-\leavevmode\hypertarget{ref-banos2021}{}%
-Banos, I.H., Mayfield, W.D., Ge, G., Sapucci, L.F., Carley, J.R., Nance, L., 2021.
-Assessment of the data assimilation framework for the Rapid Refresh Forecast System v0.1 and impacts on forecasts of convective storms.
-Geoscientific Model Development Discussions 1--36.
-doi:\href{https://doi.org/10.5194/gmd-2021-289}{10.5194/gmd-2021-289}
-
-\leavevmode\hypertarget{ref-bao2015}{}%
-Bao, Y., Xu, J., Powell Jr., A.M., Shao, M., Min, J., Pan, Y., 2015.
-Impacts of AMSU-A, MHS and IASI data assimilation on temperature and humidity forecasts with GSI--WRF over the western United States.
-Atmos.
-Meas.
-Tech.
-8, 4231--4242.
-doi:\href{https://doi.org/10.5194/amt-8-4231-2015}{10.5194/amt-8-4231-2015}
-
-\leavevmode\hypertarget{ref-brooks2003}{}%
-Brooks, H.E., Lee, J.W., Craven, J.P., 2003.
-The spatial distribution of severe thunderstorm and tornado environments from global reanalysis data.
-Atmospheric Research 67--68, 73--94.
-doi:\href{https://doi.org/10.1016/S0169-8095(03)00045-0}{10.1016/S0169-8095(03)00045-0}
-
-\leavevmode\hypertarget{ref-campitelli2020}{}%
-Campitelli, E., 2020. metR: Tools for Easier Analysis of Meteorological Fields.
-
-\leavevmode\hypertarget{ref-candille2007}{}%
-Candille, G., Côté, C., Houtekamer, P.L., Pellerin, G., 2007.
-Verification of an Ensemble Prediction System against Observations.
-Mon.
-Wea.
-Rev.
-135, 2688--2699.
-doi:\href{https://doi.org/10.1175/MWR3414.1}{10.1175/MWR3414.1}
-
-\leavevmode\hypertarget{ref-cecil2012}{}%
-Cecil, D.J., Blankenship, C.B., 2012.
-Toward a Global Climatology of Severe Hailstorms as Estimated by Satellite Passive Microwave Imagers.
-Journal of Climate 25, 687--703.
-doi:\href{https://doi.org/10.1175/JCLI-D-11-00130.1}{10.1175/JCLI-D-11-00130.1}
-
-\leavevmode\hypertarget{ref-chang2017}{}%
-Chang, W., Jacques, D., Fillion, L., Baek, S.-J., 2017.
-Assimilation of Hourly Surface Observations with the Canadian High-Resolution Ensemble Kalman Filter.
-Atmosphere-Ocean 55, 247--263.
-doi:\href{https://doi.org/10.1080/07055900.2017.1384361}{10.1080/07055900.2017.1384361}
-
-\leavevmode\hypertarget{ref-chen2001}{}%
-Chen, F., Dudhia, J., 2001.
-Coupling an Advanced Land Surface--Hydrology Model with the Penn State--NCAR MM5 Modeling System.
-Part I: Model Implementation and Sensitivity.
-Monthly Weather Review 129, 569--585.
-doi:\href{https://doi.org/10.1175/1520-0493(2001)129\%3C0569:CAALSH\%3E2.0.CO;2}{10.1175/1520-0493(2001)129\textless0569:CAALSH\textgreater2.0.CO;2}
-
-\leavevmode\hypertarget{ref-chen2015}{}%
-Chen, Q., Fan, J., Hagos, S., Gustafson, W.I., Berg, L.K., 2015.
-Roles of wind shear at different vertical levels: Cloud system organization and properties.
-Journal of Geophysical Research: Atmospheres 120, 6551--6574.
-doi:\href{https://doi.org/10.1002/2015JD023253}{10.1002/2015JD023253}
-
-\leavevmode\hypertarget{ref-chen2016}{}%
-Chen, X., Zhao, K., Sun, J., Zhou, B., Lee, W.-C., 2016.
-Assimilating surface observations in a four-dimensional variational Doppler radar data assimilation system to improve the analysis and forecast of a squall line case.
-Adv.
-Atmos.
-Sci.
-33, 1106--1119.
-doi:\href{https://doi.org/10.1007/s00376-016-5290-0}{10.1007/s00376-016-5290-0}
-
-\leavevmode\hypertarget{ref-cherubini2006}{}%
-Cherubini, T., Businger, S., Velden, C., Ogasawara, R., 2006.
-The Impact of Satellite-Derived Atmospheric Motion Vectors on Mesoscale Forecasts over Hawaii.
-Monthly Weather Review 134, 2009--2020.
-doi:\href{https://doi.org/10.1175/MWR3163.1}{10.1175/MWR3163.1}
-
-\leavevmode\hypertarget{ref-clark2017}{}%
-Clark, A.J., 2017.
-Generation of Ensemble Mean Precipitation Forecasts from Convection-Allowing Ensembles.
-Weather and Forecasting 32, 1569--1583.
-doi:\href{https://doi.org/10.1175/WAF-D-16-0199.1}{10.1175/WAF-D-16-0199.1}
-
-\leavevmode\hypertarget{ref-Cheyenne2019}{}%
-Computational and Information Systems Laboratory, 2019.
-Cheyenne: HPE/SGI ICE XA System (University Community Computing).
-National Center for Atmospheric Research Boulder, CO.
-doi:\href{https://doi.org/doi:10.5065/D6RX99HX}{doi:10.5065/D6RX99HX}
-
-\leavevmode\hypertarget{ref-deelia2017}{}%
-de Elía, R., Vidal, L., Lohigorry, P., 2017. El SMN y la red argentina de radares meteorológicos {[}WWW Document{]}. URL \url{http://hdl.handle.net/20.500.12160/625}
-
-\leavevmode\hypertarget{ref-dillon2021}{}%
-Dillon, M.E., Maldonado, P., Corrales, P., Skabar, Y.G., Ruiz, J., Sacco, M., Cutraro, F., Mingari, L., Matsudo, C., Vidal, L., Rugna, M., Hobouchian, M.P., Salio, P., Nesbitt, S., Saulo, C., Kalnay, E., Miyoshi, T., 2021.
-A rapid refresh ensemble based data assimilation and forecast system for the RELAMPAGO field campaign.
-Atmospheric Research 105858.
-doi:\href{https://doi.org/10.1016/j.atmosres.2021.105858}{10.1016/j.atmosres.2021.105858}
-
-\leavevmode\hypertarget{ref-dillon2016}{}%
-Dillon, M.E., Skabar, Y.G., Ruiz, J., Kalnay, E., Collini, E.A., Echevarría, P., Saucedo, M., Miyoshi, T., Kunii, M., 2016.
-Application of the WRF-LETKF Data Assimilation System over Southern South America: Sensitivity to Model Physics.
-Wea.
-Forecasting 31, 217--236.
-doi:\href{https://doi.org/10.1175/WAF-D-14-00157.1}{10.1175/WAF-D-14-00157.1}
-
-\leavevmode\hypertarget{ref-dowle2020}{}%
-Dowle, M., Srinivasan, A., 2020. Data.Table: Extension of 'data.frame'.
-
-\leavevmode\hypertarget{ref-sondeos}{}%
-Earth Observing Laboratory, U. -, 2020. Multi-network composite highest resolution radiosonde data. Version 1.3. UCAR/NCAR - earth observing laboratory.
-
-\leavevmode\hypertarget{ref-eyre2020}{}%
-Eyre, J.R., English, S.J., Forsythe, M., 2020.
-Assimilation of satellite data in numerical weather prediction.
-Part I: The early years.
-Quarterly Journal of the Royal Meteorological Society 146, 49--68.
-doi:\href{https://doi.org/10.1002/qj.3654}{10.1002/qj.3654}
-
-\leavevmode\hypertarget{ref-garcia2019}{}%
-Garcia, F., Ruiz, J., Salio, P., Bechis, H., Nesbitt, S., 2019. Argentina mesonet data. Version 1.1. UCAR/NCAR - earth observing laboratory.
-
-\leavevmode\hypertarget{ref-gasperoni2018}{}%
-Gasperoni, N.A., Wang, X., Brewster, K.A., Carr, F.H., 2018.
-Assessing Impacts of the High-Frequency Assimilation of Surface Observations for the Forecast of Convection Initiation on 3 April 2014 within the Dallas--Fort Worth Test Bed.
-Monthly Weather Review 146, 3845--3872.
-doi:\href{https://doi.org/10.1175/MWR-D-18-0177.1}{10.1175/MWR-D-18-0177.1}
-
-\leavevmode\hypertarget{ref-goncalvesdegoncalves2015}{}%
-Goncalves de Goncalves, L.G., Sapucci, L., Vendrasco, E., de Mattos, J.G., Ferreira, C., Khamis, E., Cruz, N., 2015.
-A rapid update data assimilation cycle over South America using 3DVar and EnKF, in: The 20th International TOVS Study Conference (ITSC-20).
-The 20th International TOVS Study Conference (ITSC-20), Lake Geneva, Wisconsin, USA.
-
-\leavevmode\hypertarget{ref-grell2013}{}%
-Grell, G.A., Freitas, S.R., 2013.
-A scale and aerosol aware stochastic convective parameterization for weather and air quality modeling.
-Atmos.
-Chem.
-Phys.
-Discuss.
-13, 23845--23893.
-doi:\href{https://doi.org/10.5194/acpd-13-23845-2013}{10.5194/acpd-13-23845-2013}
-
-\leavevmode\hypertarget{ref-ha2014}{}%
-Ha, S.-Y., Snyder, C., 2014.
-Influence of Surface Observations in Mesoscale Data Assimilation Using an Ensemble Kalman Filter.
-Monthly Weather Review 142, 1489--1508.
-doi:\href{https://doi.org/10.1175/MWR-D-13-00108.1}{10.1175/MWR-D-13-00108.1}
-
-\leavevmode\hypertarget{ref-han2006}{}%
-Han, Y., Van Delst, P., Liu, Q., Weng, F., Yan, B., Treadon, R., Derber, J., 2006. JCSDA Community Radiative Transfer Model (CRTM)---version 1, NOAA Technical Report NESDIS 122. Washington, D.C.
-
-\leavevmode\hypertarget{ref-hong2006a}{}%
-Hong, S.-Y., Kim, J.-H., Lim, J.-o., Dudhia, J., 2006. The WRF Single Moment 6-Class Microphysics Scheme (WSM6). Journal of the Korean Meteorological Society 42, 129--151.
-
-\leavevmode\hypertarget{ref-hong2006}{}%
-Hong, S.-Y., Noh, Y., Dudhia, J., 2006.
-A New Vertical Diffusion Package with an Explicit Treatment of Entrainment Processes.
-Mon.
-Wea.
-Rev.
-134, 2318--2341.
-doi:\href{https://doi.org/10.1175/MWR3199.1}{10.1175/MWR3199.1}
-
-\leavevmode\hypertarget{ref-hu2018}{}%
-Hu, M., Ge, G., Zhou, C., Stark, D., Shao, H., Newman, K., Beck, J., Zhang, X., 2018. Grid-point Statistical Interpolation (GSI) User's Guide Version 3.7. Developmental Testbed Center.
-
-\leavevmode\hypertarget{ref-huffman2018}{}%
-Huffman, G., Bolvin, D., Braithwaite, D., Hsu, K., Joyce, R., Kidd, C., Nelkin, E., Sorooshian, S., Tan, J., Xie, P., 2018.
-NASA Global Precipitation Measurement (GPM) Integrated Multi-satellitE Retrievals for GPM (IMERG).
-National Aeronautics and Space Administration (NASA).
-
-\leavevmode\hypertarget{ref-hunt2007}{}%
-Hunt, B.R., Kostelich, E.J., Szunyogh, I., 2007.
-Efficient data assimilation for spatiotemporal chaos: A local ensemble transform Kalman filter.
-Physica D: Nonlinear Phenomena 230, 112--126.
-doi:\href{https://doi.org/10.1016/j.physd.2006.11.008}{10.1016/j.physd.2006.11.008}
-
-\leavevmode\hypertarget{ref-iacono2008}{}%
-Iacono, M.J., Delamere, J.S., Mlawer, E.J., Shephard, M.W., Clough, S.A., Collins, W.D., 2008.
-Radiative forcing by long-lived greenhouse gases: Calculations with the AER radiative transfer models.
-J.
-Geophys.
-Res.
-113, D13103.
-doi:\href{https://doi.org/10.1029/2008JD009944}{10.1029/2008JD009944}
-
-\leavevmode\hypertarget{ref-janjic1994}{}%
-Janjić, Z.I., 1994.
-The Step-Mountain Eta Coordinate Model: Further Developments of the Convection, Viscous Sublayer, and Turbulence Closure Schemes.
-Mon.
-Wea.
-Rev.
-122, 927--945.
-doi:\href{https://doi.org/10.1175/1520-0493(1994)122\%3C0927:TSMECM\%3E2.0.CO;2}{10.1175/1520-0493(1994)122\textless0927:TSMECM\textgreater2.0.CO;2}
-
-\leavevmode\hypertarget{ref-jones2013}{}%
-Jones, T.A., Otkin, J.A., Stensrud, D.J., Knopfmeier, K., 2013.
-Assimilation of Satellite Infrared Radiances and Doppler Radar Observations during a Cool Season Observing System Simulation Experiment.
-Mon.
-Wea.
-Rev.
-141, 3273--3299.
-doi:\href{https://doi.org/10.1175/MWR-D-12-00267.1}{10.1175/MWR-D-12-00267.1}
-
-\leavevmode\hypertarget{ref-kain2004}{}%
-Kain, J.S., 2004. The Kain--Fritsch Convective Parameterization: An Update. JOURNAL OF APPLIED METEOROLOGY 43, 12.
-
-\leavevmode\hypertarget{ref-lin2017a}{}%
-Lin, H., Weygandt, S.S., Benjamin, S.G., Hu, M., 2017.
-Satellite Radiance Data Assimilation within the Hourly Updated Rapid Refresh.
-Wea.
-Forecasting 32, 1273--1287.
-doi:\href{https://doi.org/10.1175/WAF-D-16-0215.1}{10.1175/WAF-D-16-0215.1}
-
-\leavevmode\hypertarget{ref-maejima2019}{}%
-Maejima, Y., Miyoshi, T., Kunii, M., Seko, H., Sato, K., 2019.
-Impact of Dense and Frequent Surface Observations on 1-Minute-Update Severe Rainstorm Prediction: A Simulation Study.
-Journal of the Meteorological Society of Japan.
-Ser.
-II 97, 253--273.
-doi:\href{https://doi.org/10.2151/jmsj.2019-014}{10.2151/jmsj.2019-014}
-
-\leavevmode\hypertarget{ref-maldonado2021}{}%
-Maldonado, P., Ruiz, J., Saulo, C., 2021.
-Sensitivity to Initial and Boundary Perturbations in Convective-Scale Ensemble-Based Data Assimilation: Imperfect-Model OSSEs.
-SOLA 17, 96--102.
-doi:\href{https://doi.org/10.2151/sola.2021-015}{10.2151/sola.2021-015}
-
-\leavevmode\hypertarget{ref-mallick2020}{}%
-Mallick, S., Jones, T.A., 2020.
-Assimilation of GOES-16 satellite derived winds into the warn-on-forecast system.
-Atmospheric Research 245, 105131.
-doi:\href{https://doi.org/10.1016/j.atmosres.2020.105131}{10.1016/j.atmosres.2020.105131}
-
-\leavevmode\hypertarget{ref-markowski2010}{}%
-Markowski, P., Richardson, Y., 2010.
-Organization of Isolated Convection, in: Mesoscale Meteorology in Midlatitudes.
-John Wiley \& Sons, Ltd, pp.
-201--244.
-doi:\href{https://doi.org/10.1002/9780470682104.ch8}{10.1002/9780470682104.ch8}
-
-\leavevmode\hypertarget{ref-matsudo2015}{}%
-Matsudo, C., García Skabar, Y., Ruiz, J., Vidal, L., Salio, P., 2015. Verification of WRF-ARW convective-resolving forecasts over Southeastern South America. Mausam 66, 445--456.
-
-\leavevmode\hypertarget{ref-nakanishi2009}{}%
-Nakanishi, M., Niino, H., 2009.
-Development of an Improved Turbulence Closure Model for the Atmospheric Boundary Layer.
-JMSJ 87, 895--912.
-doi:\href{https://doi.org/10.2151/jmsj.87.895}{10.2151/jmsj.87.895}
-
-\leavevmode\hypertarget{ref-cisl_rda_ds084.1}{}%
-National Centers for Environmental Prediction, National Weather Service, NOAA, U.S. Department of Commerce, 2015. NCEP GFS 0.25 degree global forecast grids historical archive.
-
-\leavevmode\hypertarget{ref-necker2020}{}%
-Necker, T., Geiss, S., Weissmann, M., Ruiz, J., Miyoshi, T., Lien, G., 2020.
-A convective‐scale 1,000‐member ensemble simulation and potential applications.
-Q.J.R.
-Meteorol.
-Soc 146, 1423--1442.
-doi:\href{https://doi.org/10.1002/qj.3744}{10.1002/qj.3744}
-
-\leavevmode\hypertarget{ref-nesbitt2021}{}%
-Nesbitt, S.W., Salio, P.V., Ávila, E., Bitzer, P., Carey, L., Chandrasekar, V., Deierling, W., Dominguez, F., Dillon, M.E., Garcia, C.M., Gochis, D., Goodman, S., Hence, D.A., Kosiba, K.A., Kumjian, M.R., Lang, T., Luna, L.M., Marquis, J., Marshall, R., McMurdie, L.A., Nascimento, E.L., Rasmussen, K.L., Roberts, R., Rowe, A.K., Ruiz, J.J., Sabbas, E.F.M.T.S., Saulo, A.C., Schumacher, R.S., Skabar, Y.G., Machado, L.A.T., Trapp, R.J., Varble, A., Wilson, J., Wurman, J., Zipser, E.J., Arias, I., Bechis, H., Grover, M.A., 2021.
-A storm safari in Subtropical South America: Proyecto RELAMPAGO.
-Bulletin of the American Meteorological Society -1, 1--64.
-doi:\href{https://doi.org/10.1175/BAMS-D-20-0029.1}{10.1175/BAMS-D-20-0029.1}
-
-\leavevmode\hypertarget{ref-otsuka2015}{}%
-Otsuka, M., Kunii, M., Seko, H., Shimoji, K., Hayashi, M., Yamashita, K., 2015.
-Assimilation Experiments of MTSAT Rapid Scan Atmospheric Motion Vectors on a Heavy Rainfall Event.
-Journal of the Meteorological Society of Japan.
-Ser.
-II 93, 459--475.
-doi:\href{https://doi.org/10.2151/jmsj.2015-030}{10.2151/jmsj.2015-030}
-
-\leavevmode\hypertarget{ref-ouaraini2015}{}%
-Ouaraini, R.E., Berre, L., Fischer, C., Sayouty, E.H., 2015.
-Sensitivity of regional ensemble data assimilation spread to perturbations of lateral boundary conditions.
-Tellus A: Dynamic Meteorology and Oceanography 67, 28502.
-doi:\href{https://doi.org/10.3402/tellusa.v67.28502}{10.3402/tellusa.v67.28502}
-
-\leavevmode\hypertarget{ref-rasmussen2014}{}%
-Rasmussen, K.L., Zuluaga, M.D., Houze, R.A., 2014.
-Severe convection and lightning in subtropical South America.
-Geophysical Research Letters 41, 7359--7366.
-doi:\href{https://doi.org/10.1002/2014GL061767}{10.1002/2014GL061767}
-
-\leavevmode\hypertarget{ref-rcoreteam2020}{}%
-R Core Team, 2020. R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria.
-
-\leavevmode\hypertarget{ref-roberts2020}{}%
-Roberts, B., Gallo, B.T., Jirak, I.L., Clark, A.J., Dowell, D.C., Wang, X., Wang, Y., 2020.
-What Does a Convection-Allowing Ensemble of Opportunity Buy Us in Forecasting Thunderstorms?
-Weather and Forecasting 35, 2293--2316.
-doi:\href{https://doi.org/10.1175/WAF-D-20-0069.1}{10.1175/WAF-D-20-0069.1}
-
-\leavevmode\hypertarget{ref-roberts2008}{}%
-Roberts, N., 2008.
-Assessing the spatial and temporal variation in the skill of precipitation forecasts from an NWP model.
-Meteorological Applications 15, 163--169.
-doi:\href{https://doi.org/10.1002/met.57}{10.1002/met.57}
-
-\leavevmode\hypertarget{ref-ruiz2010}{}%
-Ruiz, J.J., Saulo, C., Nogués-Paegle, J., 2010.
-WRF Model Sensitivity to Choice of Parameterization over South America: Validation against Surface Variables.
-Mon.
-Wea.
-Rev.
-138, 3342--3355.
-doi:\href{https://doi.org/10.1175/2010MWR3358.1}{10.1175/2010MWR3358.1}
-
-\leavevmode\hypertarget{ref-sawada2019}{}%
-Sawada, M., Ma, Z., Mehra, A., Tallapragada, V., Oyama, R., Shimoji, K., 2019.
-Impacts of Assimilating High-Resolution Atmospheric Motion Vectors Derived from Himawari-8 on Tropical Cyclone Forecast in HWRF.
-Monthly Weather Review 147, 3721--3740.
-doi:\href{https://doi.org/10.1175/MWR-D-18-0261.1}{10.1175/MWR-D-18-0261.1}
-
-\leavevmode\hypertarget{ref-shao2016}{}%
-Shao, H., Derber, J., Huang, X.-Y., Hu, M., Newman, K., Stark, D., Lueken, M., Zhou, C., Nance, L., Kuo, Y.-H., Brown, B., 2016.
-Bridging Research to Operations Transitions: Status and Plans of Community GSI.
-Bulletin of the American Meteorological Society 97, 1427--1440.
-doi:\href{https://doi.org/10.1175/BAMS-D-13-00245.1}{10.1175/BAMS-D-13-00245.1}
-
-\leavevmode\hypertarget{ref-singh2016}{}%
-Singh, R., Ojha, S.P., Kishtawal, C.M., Pal, P.K., Kiran Kumar, A.S., 2016.
-Impact of the assimilation of INSAT-3D radiances on short-range weather forecasts: Assimilation of INSAT-3D Radiances.
-Q.J.R.
-Meteorol.
-Soc.
-142, 120--131.
-doi:\href{https://doi.org/10.1002/qj.2636}{10.1002/qj.2636}
-
-\leavevmode\hypertarget{ref-skamarock2008}{}%
-Skamarock, W.C., Klemp, J.B., Dudhia, J., Gill, D.O., Barker, D.M., Duda, M.G., Huang, X.-Y., Wang, W., Powers, J.G., 2008. A Description of the Advanced Research WRF Version 3.
-
-\leavevmode\hypertarget{ref-sobash2015}{}%
-Sobash, R.A., Stensrud, D.J., 2015.
-Assimilating Surface Mesonet Observations with the EnKF to Improve Ensemble Forecasts of Convection Initiation on 29 May 2012.
-Monthly Weather Review 143, 3700--3725.
-doi:\href{https://doi.org/10.1175/MWR-D-14-00126.1}{10.1175/MWR-D-14-00126.1}
-
-\leavevmode\hypertarget{ref-sun2014}{}%
-Sun, J., Xue, M., Wilson, J.W., Zawadzki, I., Ballard, S.P., Onvlee-Hooimeyer, J., Joe, P., Barker, D.M., Li, P.-W., Golding, B., Xu, M., Pinto, J., 2014.
-Use of NWP for Nowcasting Convective Precipitation: Recent Progress and Challenges.
-Bulletin of the American Meteorological Society 95, 409--426.
-doi:\href{https://doi.org/10.1175/BAMS-D-11-00263.1}{10.1175/BAMS-D-11-00263.1}
-
-\leavevmode\hypertarget{ref-wang2021}{}%
-Wang, Z.Q., Randriamampianina, R., 2021.
-The Impact of Assimilating Satellite Radiance Observations in the Copernicus European Regional Reanalysis (CERRA).
-Remote Sensing 13, 426.
-doi:\href{https://doi.org/10.3390/rs13030426}{10.3390/rs13030426}
-
-\leavevmode\hypertarget{ref-weston2019}{}%
-Weston, P., Geer, A., Bormann, N., Bormann, N., 2019.
-Investigations into the assimilation of AMSU-A in the presence of cloud and precipitation {[}WWW Document{]}.
-EUMETSAT/ECMWF Fellowship Programme Research Report.
-doi:\href{https://doi.org/10.21957/ewahn9ce}{10.21957/ewahn9ce}
-
-\leavevmode\hypertarget{ref-wheatley2010}{}%
-Wheatley, D.M., Stensrud, D.J., 2010.
-The Impact of Assimilating Surface Pressure Observations on Severe Weather Events in a WRF Mesoscale Ensemble System.
-Monthly Weather Review 138, 1673--1694.
-doi:\href{https://doi.org/10.1175/2009MWR3042.1}{10.1175/2009MWR3042.1}
-
-\leavevmode\hypertarget{ref-whitaker2012}{}%
-Whitaker, J.S., Hamill, T.M., 2012.
-Evaluating Methods to Account for System Errors in Ensemble Data Assimilation.
-Mon.
-Wea.
-Rev.
-140, 3078--3089.
-doi:\href{https://doi.org/10.1175/MWR-D-11-00276.1}{10.1175/MWR-D-11-00276.1}
-
-\leavevmode\hypertarget{ref-wickham2009}{}%
-Wickham, H., 2009. Ggplot2: Elegant Graphics for Data Analysis, Use R! Springer-Verlag, New York. doi:\href{https://doi.org/10.1007/978-0-387-98141-3}{10.1007/978-0-387-98141-3}
-
-\leavevmode\hypertarget{ref-wu2014}{}%
-Wu, T.-C., Liu, H., Majumdar, S.J., Velden, C.S., Anderson, J.L., 2014.
-Influence of Assimilating Satellite-Derived Atmospheric Motion Vector Observations on Numerical Analyses and Forecasts of Tropical Cyclone Track and Intensity.
-Monthly Weather Review 142, 49--71.
-doi:\href{https://doi.org/10.1175/MWR-D-13-00023.1}{10.1175/MWR-D-13-00023.1}
-
-\leavevmode\hypertarget{ref-xie2015}{}%
-Xie, Y., 2015. Dynamic documents with R and knitr, Second. ed. Chapman and Hall/CRC, Boca Raton, Florida.
-
-\leavevmode\hypertarget{ref-zhao2021}{}%
-Zhao, J., Gao, J., Jones, T.A., Hu, J., 2021a.
-Impact of Assimilating High-Resolution Atmospheric Motion Vectors on Convective Scale Short-Term Forecasts: 1.
-Observing System Simulation Experiment (OSSE).
-Journal of Advances in Modeling Earth Systems 13, e2021MS002484.
-doi:\href{https://doi.org/10.1029/2021MS002484}{10.1029/2021MS002484}
-
-\leavevmode\hypertarget{ref-zhao2021a}{}%
-Zhao, J., Gao, J., Jones, T.A., Hu, J., 2021b.
-Impact of Assimilating High-Resolution Atmospheric Motion Vectors on Convective Scale Short-Term Forecasts: 2.
-Assimilation Experiments of GOES-16 Satellite Derived Winds.
-Journal of Advances in Modeling Earth Systems 13, e2021MS002486.
-doi:\href{https://doi.org/10.1029/2021MS002486}{10.1029/2021MS002486}
-
-\leavevmode\hypertarget{ref-zhu2019}{}%
-Zhu, K., Xue, M., Pan, Y., Hu, M., Benjamin, S.G., Weygandt, S.S., Lin, H., 2019.
-The Impact of Satellite Radiance Data Assimilation within a Frequently Updated Regional Forecast System Using a GSI-based Ensemble Kalman Filter.
-Adv.
-Atmos.
-Sci.
-36, 1308--1326.
-doi:\href{https://doi.org/10.1007/s00376-019-9011-3}{10.1007/s00376-019-9011-3}
-
-\leavevmode\hypertarget{ref-zhu2014}{}%
-Zhu, Y., Derber, J., Collard, A., Dee, D., Treadon, R., Gayno, G., Jung, J.A., 2014.
-Enhanced radiance bias correction in the National Centers for Environmental Prediction's Gridpoint Statistical Interpolation data assimilation system.
-Quarterly Journal of the Royal Meteorological Society 140, 1479--1492.
-doi:\href{https://doi.org/10.1002/qj.2233}{10.1002/qj.2233}
-
-\leavevmode\hypertarget{ref-zhu2016}{}%
-Zhu, Y., Liu, E., Mahajan, R., Thomas, C., Groff, D., Van Delst, P., Collard, A., Kleist, D., Treadon, R., Derber, J.C., 2016.
-All-Sky Microwave Radiance Assimilation in NCEP's GSI Analysis System.
-Mon.
-Wea.
-Rev.
-144, 4709--4735.
-doi:\href{https://doi.org/10.1175/MWR-D-15-0445.1}{10.1175/MWR-D-15-0445.1}
+\renewcommand\refname{References}
+\bibliography{paperANA.bib,packages.bib,era.bib}
 

### 2022-09-27 analysis/paper/paper_ana_20181122.tex
Paper aproved!!!

@@
-\documentclass[preprint, 3p, authoryear,review, 12pt]{elsarticle} %review=doublespace preprint=single 5p=2 column
+\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
@@
 \usepackage{lineno} % add
-  \linenumbers % turns line numbering on
 
@@
 There have been only a few published efforts on regional mesoscale DA, but they have all shown promising results \citep[e.g.][]{dillon2016, dillon2021, goncalvesdegoncalves2015}.
-In particular, \citet{dillon2021} assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts.
+In particular, \citet{dillon2021} assimilated high resolution surface weather station networks, GOES-16 satellite-derived winds, and satellite temperature and moisture retrievals over central Argentina with positive impacts.
 Similar to \citet{gasperoni2018}, \citet{dillon2021} included private weather station networks which are not incorporated in the operational analysis.
@@
 
-The impact of non-conventional high spatial and temporal resolution observations, such as AMVs, has been investigated in the context of regional mesoscale DA.
+The impact of non-conventional high spatial and temporal resolution observations, such as satellite-derived winds, has been investigated in the context of regional mesoscale DA.
 Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., \citet{wu2014}, \citet{cherubini2006}, and \citet{sawada2019}, and many others).
-Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm.
+Most of these studies reported an overall positive impact of the assimilation of satellite-derived winds for this type of storm.
 However, some works indicated mixed impacts (e.g.
@@
 As stated in \citet{zhao2021, zhao2021a}, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
-\citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
-They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts.
-\citet{otsuka2015} and \citet{mallick2020} found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.
+\citet{zhao2021, zhao2021a} assimilated GOES-16 satellite-derived winds into a storm-scale three-dimensional variational DA system during three high impact weather events.
+They reported positive impacts of satellite-derived winds on the characterization of the storm environment and improved short range precipitation forecasts.
+\citet{otsuka2015} and \citet{mallick2020} found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency satellite-derived winds.
 
@@
 
-The main objective of this work is thus to contribute to the quantification and comparison of the impact of high resolution automatic weather stations, AMVs, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based regional DA system.
+The main objective of this work is thus to contribute to the quantification and comparison of the impact of high resolution automatic weather stations, satellite-derived winds, and clear-sky satellite radiances, into a mesoscale, frequently-updated ensemble-based regional DA system.
 This is particularly important in the efforts to improve mesoscale numerical weather prediction (NWP) over South America where the conventional observation network is rather sparse and other sources of information could potentially fill certain gaps.
@@
 \item
-  The Satellite precipitation estimation IMERG Final Run with 0.01\(^{\circ}\) spatial resolution and 30 minutes temporal resolution \citep{huffman2018} was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
+  The Satellite precipitation estimation IMERG Final Run with 0.1\(^{\circ}\) spatial resolution and 30 minutes temporal resolution \citep{huffman2018} was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
 \item
@@
 
-The difference between ERA5 reanalyses \citep[`]{era5pressure} and the ensemble mean analyses are also compared in Figure \ref{fig:era5}, which supports Figures \ref{fig:TQ-diff} and \ref{fig:UV-diff}.
+The difference between the ensemble mean analyses and ERA5 \citep[`]{era5pressure} are also compared in Figure \ref{fig:era5}, which supports Figures \ref{fig:TQ-diff} and \ref{fig:UV-diff}.
 Specifically, Figure \ref{fig:era5}a shows a warm bias in low levels (i.e.~CONV is warmer than ERA5) that decreases in Figure \ref{fig:era5}b when the AWS observations are assimilated.
 In the same direction, Figure \ref{fig:TQ-diff}a shows a negative difference between AWS and CONV meaning that the AWS observations are cooling the low levels.
-Comparing ERA5-RAD (Figure \ref{fig:era5}d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations as shown in Figure \ref{fig:TQ-diff}c.~A similar effect can be observed for specific humidity, AWS observations partially correct the dry bias present in Figure \ref{fig:era5}e and the assimilation of radiance observations reduces the positive impact of AWS.
-The impact on the wind components is minor so only the meridional wind is included in Figures \ref{fig:era5}i-l, which show that the radiance observations are mainly responsible for the positive impact observed in the analysis by reducing the distance ERA5-RAD, particularly during the mature stage of the MCS.
-Overall, the adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5 reanalyses.
+Comparing with RAD-ERA5 (Figure \ref{fig:era5}d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations as shown in Figure \ref{fig:TQ-diff}c.~A similar effect can be observed for specific humidity, AWS observations partially correct the dry bias present in Figure \ref{fig:era5}e and the assimilation of radiance observations reduces the positive impact of AWS.
+The impact on the wind components is minor and only the meridional wind is included in Figures \ref{fig:era5}i-l, which show that the radiance observations are mainly responsible for the positive impact observed in the analysis by reducing the distance RAD-ERA5, particularly during the mature stage of the MCS.
+Overall, the adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5.
 
@@
 
-\caption{Difference between ERA5 and the analysis ensemble mean experiments for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity (\(g\ Kg^{-1}\), e--h) and meridional wind (\(m\ s^{-1}\), i--l) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:era5}
+\caption{Difference between the analysis ensemble mean experiments and ERA5 for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity (\(g\ Kg^{-1}\), e--h) and meridional wind (\(m\ s^{-1}\), i--l) calculated over the inner domain (red box in Figure \ref{fig:dominio}a) for each analysis cycle.}\label{fig:era5}
 \end{figure*}

### 2022-09-27 analysis/paper/paper_ana_20181122.Rmd

@@
 journal: Atmospheric Research
-#layout: authoryear,preprint,review,12pt
+#layout: preprint, 3p, authoryear,review, 12pt
 # layout: final,5p,times,twocolumn,authoryear
-classoption: preprint, 3p, authoryear,review, 12pt
-linenumbers: true
+classoption: final,5p,times,twocolumn,authoryear
+linenumbers: false
 numbersections: true
@@
 @dillon2016; @dillon2021; @goncalvesdegoncalves2015].
-In particular, @dillon2021 assimilated high resolution surface weather station networks, GOES-16 atmospheric motion vectors (AMVs), and satellite temperature and moisture retrievals over central Argentina with positive impacts.
+In particular, @dillon2021 assimilated high resolution surface weather station networks, GOES-16 satellite-derived winds, and satellite temperature and moisture retrievals over central Argentina with positive impacts.
 Similar to @gasperoni2018, @dillon2021 included private weather station networks which are not incorporated in the operational analysis.
@@
 
-The impact of non-conventional high spatial and temporal resolution observations, such as AMVs, has been investigated in the context of regional mesoscale DA.
+The impact of non-conventional high spatial and temporal resolution observations, such as satellite-derived winds, has been investigated in the context of regional mesoscale DA.
 Many studies have focused on the impact of these observations on the prediction of tropical storms (e.g., @wu2014, @cherubini2006, and @sawada2019, and many others).
-Most of these studies reported an overall positive impact of the assimilation of AMVs for this type of storm.
+Most of these studies reported an overall positive impact of the assimilation of satellite-derived winds for this type of storm.
 However, some works indicated mixed impacts (e.g.
@@
 As stated in \citet{zhao2021, zhao2021a}, the impact of assimilating these data on high impact weather events associated with mid-latitude deep convection over land has received relatively less attention.
-\citet{zhao2021, zhao2021a} assimilated GOES-16 AMVs into a storm-scale three-dimensional variational DA system during three high impact weather events.
-They reported positive impacts of AMVs on the characterization of the storm environment and improved short range precipitation forecasts.
-@otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency AMVs.
+\citet{zhao2021, zhao2021a} assimilated GOES-16 satellite-derived winds into a storm-scale three-dimensional variational DA system during three high impact weather events.
+They reported positive impacts of satellite-derived winds on the characterization of the storm environment and improved short range precipitation forecasts.
+@otsuka2015 and @mallick2020 found a slight improvement in the short-range precipitation forecast due to the storm-scale assimilation of high frequency satellite-derived winds.
 
@@
   scale_fill_viridis_c(option = "D", trans = scales::log10_trans(),
-                       guide = guide_colorbar(barwidth = 20,
+                       guide = guide_colorbar(barwidth = 18,
                                               barheight = 0.5, 
@@
 
-* The Satellite precipitation estimation IMERG Final Run with 0.01$^{\circ}$ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
+* The Satellite precipitation estimation IMERG Final Run with 0.1$^{\circ}$ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.
 
@@
 
-```{r rcrv-sfc, fig.cap="(ref:rcrv-sfc)", out.width="100%", fig.height=5}
+```{r rcrv-sfc, fig.cap="(ref:rcrv-sfc)", out.width="100%", fig.height=4}
 
@@
                        guide = guide_colorsteps(inside = TRUE,
-                                                barwidth = 20,
+                                                barwidth = 18,
                                                 barheight = 0.5, 
@@
 
-The difference between ERA5 reanalyses [@era5pressure`] and the ensemble mean analyses are also compared in Figure \@ref(fig:era5), which supports Figures \@ref(fig:TQ-diff) and \@ref(fig:UV-diff).
+The difference between the ensemble mean analyses and ERA5 [@era5pressure`] are also compared in Figure \@ref(fig:era5), which supports Figures \@ref(fig:TQ-diff) and \@ref(fig:UV-diff).
 Specifically, Figure \@ref(fig:era5)a shows a warm bias in low levels (i.e.
@@
 In the same direction, Figure \@ref(fig:TQ-diff)a shows a negative difference between AWS and CONV meaning that the AWS observations are cooling the low levels.
-Comparing ERA5-RAD (Figure \@ref(fig:era5)d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations as shown in Figure \@ref(fig:TQ-diff)c.
+Comparing with RAD-ERA5 (Figure \@ref(fig:era5)d), there is a small increase in the warm bias, associated with the warming produced by the radiance observations as shown in Figure \@ref(fig:TQ-diff)c.
 A similar effect can be observed for specific humidity, AWS observations partially correct the dry bias present in Figure \@ref(fig:era5)e and the assimilation of radiance observations reduces the positive impact of AWS.
-The impact on the wind components is minor so only the meridional wind is included in Figures \@ref(fig:era5)i-l, which show that the radiance observations are mainly responsible for the positive impact observed in the analysis by reducing the distance ERA5-RAD, particularly during the mature stage of the MCS.
-Overall, the adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5 reanalyses.
-
-(ref:era5-cap) Difference between ERA5 and the analysis ensemble mean experiments for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity ($g\ Kg^{-1}$, e--h) and meridional wind ($m\ s^{-1}$, i--l) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
+The impact on the wind components is minor and only the meridional wind is included in Figures \@ref(fig:era5)i-l, which show that the radiance observations are mainly responsible for the positive impact observed in the analysis by reducing the distance RAD-ERA5, particularly during the mature stage of the MCS.
+Overall, the adjustments due to assimilating radiance and AWS observations lead to an ensemble mean analyses closer to ERA5.
+
+(ref:era5-cap) Difference between the analysis ensemble mean experiments and ERA5 for the spatially averaged vertical profiles of air temperature (K, a--d), specific humidity ($g\ Kg^{-1}$, e--h) and meridional wind ($m\ s^{-1}$, i--l) calculated over the inner domain (red box in Figure \@ref(fig:dominio)a) for each analysis cycle.
 
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = T - i.T, fill = stat(level)), 
+  geom_contour_fill(aes(z = i.T - T, fill = stat(level)), 
                     breaks = seq(-3, 3, 0.5)) +
-  geom_contour2(aes(z = T - i.T), color = "white", size = 0.1,
+  geom_contour2(aes(z = i.T - T), color = "white", size = 0.1,
                 breaks = seq(-3, 3, 0.5)) +
@@
   ggplot(aes(date, lev)) +
-  geom_contour_fill(aes(z = QVAPOR - i.QVAPOR, fill = stat(level)), 
-                    breaks = seq(-0.0015, 0.0025, 0.0004)) +
-  geom_contour2(aes(z = QVAPOR - i.QVAPOR), color = "white", size = 0.1,
-                breaks = seq(-0.0015, 0.0025, 0.0004)) +
+  geom_contour_fill(aes(z = i.QVAPOR - QVAPOR, fill = stat(level)), 
+                    breaks = seq(-0.0025, 0.0015, 0.0004)) +
+  geom_contour2(aes(z = i.QVAPOR - QVAPOR), color = "white", size = 0.1,
+                breaks = seq(-0.0025, 0.0015, 0.0004)) +
   scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
@@
     ggplot(aes(date, lev)) +
-    geom_contour_fill(aes(z = V - i.V, fill = stat(level)), 
-                      breaks = seq(-7, 4, 1)) +
-    geom_contour2(aes(z = V - i.V), color = "white", size = 0.1,
-                  breaks = seq(-7, 4, 1)) +
+    geom_contour_fill(aes(z = i.V - V, fill = stat(level)), 
+                      breaks = seq(-4, 7, 1)) +
+    geom_contour2(aes(z = i.V - V), color = "white", size = 0.1,
+                  breaks = seq(-4, 7, 1)) +
     scale_fill_divergent_discretised(guide = guide_colorsteps(barwidth = 25,
